### 2020-08-04 SLT_paper.Rmd
from Chiara

---
title: "SLT_paper"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Open Data and first look

```{r data}
library(tidyverse)
data_all <- read_delim("/Users/chiarasemenzin/Desktop/dc/final_classifications_PU.csv", delim=',',col_types = cols(.default = "c"))
nb_answers_new <- data_all %>%
  group_by(segmentId_DB) %>%
  count()
print(paste("Number of clips with at least one annotation:", nrow(nb_answers_new)))
not_finished_new <- nb_answers_new %>%
  filter(n<=2)
print(paste("Number of clips with at least one annotation missing:",nrow(not_finished_new)))
three_answers_new <- nb_answers_new %>%
  filter(n>=3)
print(paste("Number of clips with three annotations:",nrow(three_answers_new)))
```

# Canonical Rate analysis by Age & ChildID

```{r CR by Age}
df.sub <- subset(data_all, Zoon_classif != "Junk" | Zoon_classif != "Crying" | Zoon_classif != "Laughing")

### By Age
library(dplyr)
library(ggplot2)

n.can <- as.data.frame(df.sub %>% group_by(Age) %>% filter(Zoon_classif == "Canonical") %>% summarise(n.can = n()))
n.non.can <- as.data.frame(df.sub %>% group_by(Age) %>% filter(Zoon_classif == "Non-Canonical") %>% summarise(n.non.can = n()))
n.tot <- merge(n.can, n.non.can)
n.tot$ratio.can <- n.tot$n.can/(n.tot$n.can + n.tot$n.non.can)
n.tot$ratio.non.can <- n.tot$n.non.can/(n.tot$n.can + n.tot$n.non.can)

### CANONICAL
df.plot <- rbind(data.frame(Age = as.factor(n.tot$Age), Ratio = n.tot$ratio.can, Cat = "Canonical"))
ggplot(df.plot, aes(fill=Cat, y=Ratio, x=Age)) + 
  geom_bar(position="dodge", stat="identity")

### NONCANONICAL
df.plot <- rbind(data.frame(Age = as.factor(n.tot$Age), Ratio = n.tot$ratio.non.can, Cat = "Non-Canonical"))
ggplot(df.plot, aes(fill=Cat, y=Ratio, x=Age)) + 
  geom_bar(position="dodge", stat="identity")

###BOTH

df.plot <- rbind(data.frame(Age = as.factor(n.tot$Age), Ratio = n.tot$ratio.can, Cat = "Canonical"),data.frame(Age = as.factor(n.tot$Age), Ratio = n.tot$ratio.non.can, Cat = "Non-canonical"))

ggplot(df.plot, aes(fill=Cat, y=Ratio, x=Age)) + 
  geom_bar(position="dodge", stat="identity")
```

```{r ChildID}
### By Recording (=Child ID)
n.can <- as.data.frame(df.sub %>% group_by(ChildID) %>% filter(Zoon_classif == "Canonical") %>% summarise(n.can = n()))
n.non.can <- as.data.frame(df.sub %>% group_by(ChildID) %>% filter(Zoon_classif == "Non-Canonical") %>% summarise(n.non.can = n()))
n.tot <- merge(n.can, n.non.can)
n.tot$ratio.can <- n.tot$n.can/(n.tot$n.can + n.tot$n.non.can)
n.tot$ratio.non.can <- n.tot$n.non.can/(n.tot$n.can + n.tot$n.non.can)
df.plot <- rbind(data.frame(ChildID = as.factor(n.tot$ChildID), Ratio = n.tot$ratio.can, Cat = "Canonical"),
                 data.frame(ChildID = as.factor(n.tot$ChildID), Ratio = n.tot$ratio.non.can, Cat = "Non-Canonical"))

ggplot(df.plot, aes(fill=Cat, y=Ratio, x=ChildID)) + 
  geom_bar(position="dodge", stat="identity")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-08-04 chunk_to_segment.py

@@ -1,0 +1,70 @@
+import numpy as np
+from collections import defaultdict
+import json
+
+
+result_final="/Users/chiarasemenzin/Desktop/SLT/result_final_lisa.csv"
+result_zoon="/Users/chiarasemenzin/Desktop/SLT/Metadata_batch2_lisa.csv"
+
+# ----------------------------------------------------------------------------------------
+
+def seg_to_ts(result_final):
+    seg2ts={}
+    with open(result_final) as final:
+        for e,line in enumerate(final):
+            line=line.split(",")
+            if e > 0:
+                seg2ts[line[0]]=(float(line[4]),float(line[5]),line[1])
+    return seg2ts
+
+
+def chunk_to_ts(result_zoon):
+    chunk2ts={}
+    with open(result_zoon) as zoon:
+        for e,line in enumerate(zoon):
+            line=line.split("\t")
+            if e > 0:
+                chunk2ts[line[0]]=[float(i) for i in line[4].split("_")]
+                chunk2ts[line[0]].append(line[2])
+    return chunk2ts
+
+
+def check(segments,chunks):
+    """" Check which chunks belong to which segment"""
+
+    segments_final = defaultdict(list)
+
+    for name,chunk in chunks.items(): 
+        diffs_onset=[]
+        diffs_offeset=[]
+        candidate_seg=dict([i for i in segments.items() if i[1][2]==chunk[2]])
+        for k,seg in candidate_seg.items():
+            if seg[2] == chunk[2]:
+                print("Looking at ",k,seg[2])
+                diffs_onset.append([abs(chunk[0]-seg[0]),k])
+                diffs_offeset.append([abs(chunk[1]-seg[1]),k])
+        min_on=min([i[0] for i in diffs_onset])
+        don=np.array(diffs_onset)
+        try:
+            idx=int(np.where(don ==str(min_on))[0])
+            segment=don[idx,1]
+            segments_final[segment].append(name)
+        except:
+            print("Ambigous segment. Check manually.")
+            continue
+    print(list(segments_final.items())[0:100])
+    return segments_final
+
+
+# ----------------------------------------------------------------------------------------
+
+if __name__ == "__main__":
+    s=seg_to_ts(result_final)
+    c=chunk_to_ts(result_zoon)
+    seg_=check(s,c)
+    json = json.dumps(seg_)
+    f = open("dict.json","w")
+    f.write(json)
+    f.close()
+
+

### 2020-08-04 write_seg_data.py

@@ -1,0 +1,75 @@
+import pandas as pd
+import numpy as np
+from collections import defaultdict, Counter
+import json
+
+# ----------------------------------------------------------------------------------------
+# Load data
+
+result_zoon="/Users/chiarasemenzin/Desktop/SLT/first_class_lisa.csv"
+zoon=pd.read_csv(result_zoon)
+lisa=pd.read_csv("result_final_lisa.csv") #file with metadata and PU classifications
+
+
+with open("/Users/chiarasemenzin/Desktop/SLT/dict.json", "r") as read_file:
+    print("Converting JSON encoded data into Python dictionary")
+    d = json.load(read_file)
+
+
+# ----------------------------------------------------------------------------------------
+
+def agreement_by_chunk(d,classifs):
+    """Determine majority agreement within chunks"""
+    answers = defaultdict(list)
+    for k,v in d.items():
+        for chunk in v:
+            c=Counter(list(classifs.loc[classifs["AudioData"] == chunk+".mp3" ]["Answer"]))
+            if len(c)!= 0:
+                value, count = c.most_common()[0]
+                if count >= 3 and sum([i for i in c.values()])>=3: #at least 3 classifs agree
+                    answers[k].append(value)
+                    print("Okay")
+                else:
+                    print("No majority agreement")
+    return answers
+
+
+def agreement_by_segment(answers):
+    """Determine classification within segment """
+#    TODO
+    seg_dict=[[],[]]
+    for k,v in answers.items():
+        if set(v) != {"Junk"}: # Check Junk is not the only answer
+            v=[value for value in v if value != "Junk"] #remove junk
+            if len(set(v)) == 1: # 100% agreement 
+                classif = v[0]
+            elif len(v)%2 == 0 and len(set(v)) == 2:  # 50/50 
+                if "Canonical" in v and "Non-Canonical" in v: # if Can & NonCan = Can
+                    classif="Canonical"
+                else:
+                    classif = "Mixed_" + list(set(v))[0] + "_" + list(set(v))[1] # Otherwise mixed
+            elif "Canonical" in v and "Non-Canonical" in v: #not sure about this
+                classif="Canonical"
+            else:
+                classif= "Mixed_"+"_".join([i for i in set(v)])
+                print(classif)
+        else:
+            classif="Junk" # Junk only
+        seg_dict[0].append(k)
+        seg_dict[1].append(classif)
[…]

### 2020-08-04 SLT_paper.Rmd
confusion matrices done

@@
 knitr::opts_chunk$set(echo = TRUE)
+library(tidyverse)
+library(dplyr)
+library(ggplot2)
+library(caret)
+
+dodiv=function(x) x/sum(x, na.rm=T)
+
 ```
 
-## Open Data and first look
+```{r}
+read.csv("first_class_lisa.csv")->zoon
+summary(zoon)
 
-```{r data}
-library(tidyverse)
-data_all <- read_delim("/Users/chiarasemenzin/Desktop/dc/final_classifications_PU.csv", delim=',',col_types = cols(.default = "c"))
-nb_answers_new <- data_all %>%
-  group_by(segmentId_DB) %>%
-  count()
-print(paste("Number of clips with at least one annotation:", nrow(nb_answers_new)))
-not_finished_new <- nb_answers_new %>%
-  filter(n<=2)
-print(paste("Number of clips with at least one annotation missing:",nrow(not_finished_new)))
-three_answers_new <- nb_answers_new %>%
-  filter(n>=3)
-print(paste("Number of clips with three annotations:",nrow(three_answers_new)))
+read.csv("Metadata_batch2_lisa.csv",sep="\t")->md
+summary(md)
+#? correct errors in md? onset           offset       chunk_pos
+
+read.csv("result_final_lisa.csv")->lab
+summary(lab)
+
+data_all <- read.csv("final_classifications_PU_zoon.csv")
+
+#remove the word mixed that takes up space and is unnecessary
+data_all$Zoon_classif=factor(gsub("Mixed_","",as.character(data_all$Zoon_classif),fixed=T))
+
+data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=c("Canonical","Non-Canonical",
+                                                             "Crying","Laughing","Junk",levels(data_all$Zoon_classif)[grep("_",levels(data_all$Zoon_classif))]))
+
+# create lab column with easier to read correspondance
+data_all$lab<-as.character(data_all$Major_Choice)
+data_all$lab[data_all$lab=="Non-canonical syllables"]<-"Non-Canonical"
+data_all$lab[data_all$lab=="Canonical syllables"]<-"Canonical"
+data_all$lab[data_all$lab %in% c("Don't mark","None")]<-"Junk"
+data_all$lab=factor(data_all$lab,levels=levels(data_all$Zoon_classif))
+
+```
+
+## correspondence
+i) quantifying the correspondence between expert human annotations and Zooniverse annotations %(confusion matrix) 
+and
+
+```{r}
+table(data_all$lab)
+table(data_all$Zoon_classif)
+
+mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
+
+conf_tab=mycf$table
+
+# this package uses sensitivity & specificity
+#Sensitivity=recall
+#Specificity=precision
+
+```
+
+```{r prec}
+pdf("precision_with_mixed.pdf",height=10,width=10)
+prop_cat=data.frame(conf_tab/colSums(conf_tab)*100) #generates precision because columns
+prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
+colnames(prop_cat)[3]<-"pr"
+
+data.frame(conf_tab)->stall
+stall$id=paste(stall$Lab,stall$Zooniverse)
+stall=merge(stall,prop_cat[c("id","pr")])
+
+ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
+ geom_tile(aes(fill= rescale(pr)), colour = "white") +
+  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=2) +
+  geom_text(aes(label = Freq), vjust = 1,size=2) +
+  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
+     theme(legend.position = "none") +
+  xlab("Zooniverse") + ylab("Lab") +
+  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+dev.off()
+```
+
+```{r recall}
+pdf("recall_with_mixed.pdf",height=10,width=10)
+prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
+prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
+colnames(prop_cat)[3]<-"rec"
+
+data.frame(conf_tab)->stall
+stall$id=paste(stall$Lab,stall$Zooniverse)
+stall=merge(stall,prop_cat[c("id","rec")])
+
+ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
+ geom_tile(aes(fill= rescale(rec)), colour = "white") +
+  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=2) +
+  geom_text(aes(label = Freq), vjust = 1,size=2) +
+  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
+     theme(legend.position = "none") +
+  xlab("Zooniverse") + ylab("Lab") +
+  ggtitle("Recall")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+dev.off()
+```
+
+## What is said about infants
+
+ii) examining the correspondences between descriptors at the child level across the two annotation settings (e.g. CR-AGE correlations).
+
+```{r by child}
+
+#given results above, we map the mixed
+data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical"]<-"Canonical"
+data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical"]<-"Non-Canonical"
+data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical_Crying"]<-"Non-Canonical"
+data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Crying"]<-"Crying"
+data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Crying"]<-"Non-Canonical"
+data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing_Crying"]<-"Non-Canonical"
+
+#and reset the factors for cleanliness
+data_all$Zoon_classif=factor(data_all$Zoon_classif)
+data_all$lab=factor(data_all$lab)
+
+zoon_sums = data_all %>% group_by(ChildID,Zoon_classif) %>% tally()
+colnames(zoon_sums)<-c("ChildID",      "Zoon_classif", "n_zoo")
+lab_sums = data_all %>% group_by(ChildID,lab) %>% tally()
+zoon_sums$id=paste(zoon_sums$ChildID,zoon_sums$Zoon_classif)
+lab_sums$id=paste(lab_sums$ChildID,lab_sums$lab)
+sums=merge(zoon_sums,lab_sums[,c("lab" ,  "n", "id")],by="id",all=T)
+
 ```
@@
 ```{r CR by Age}
-df.sub <- subset(data_all, Zoon_classif != "Junk" | Zoon_classif != "Crying" | Zoon_classif != "Laughing")
+df.sub <- subset(data_all, Zoon_classif %in% c("Canonical","Non-Canonical"))
 
 ### By Age
-library(dplyr)
-library(ggplot2)
-
 n.can <- as.data.frame(df.sub %>% group_by(Age) %>% filter(Zoon_classif == "Canonical") %>% summarise(n.can = n()))

### 2020-08-04 SLT_paper.Rmd
broken, don't use

@@
 library(caret)
+library(scales)
 
@@
 
-```{r by child}
+```{r prepare data by child}
 
@@
 
-zoon_sums = data_all %>% group_by(ChildID,Zoon_classif) %>% tally()
-colnames(zoon_sums)<-c("ChildID",      "Zoon_classif", "n_zoo")
-lab_sums = data_all %>% group_by(ChildID,lab) %>% tally()
+## broken here!!
+zoon_sums = data_all %>% group_by(ChildID,Zoon_classif, Age) %>% count(.drop = FALSE) #it's still dropping empty levels
+colnames(zoon_sums)<-c("ChildID",      "Zoon_classif","Age", "n_zoo")
 zoon_sums$id=paste(zoon_sums$ChildID,zoon_sums$Zoon_classif)
+#add zeros
+table(zoon_sums$ChildID,zoon_sums$Zoon_classif)==0
+mytab=table(zoon_sums$ChildID,zoon_sums$Zoon_classif)
+for(i in 1:dim(mytab)[1]) for(j in 1:dim(mytab)[2]) if(mytab[i,j]==0){
+  zero_line=cbind(rownames(mytab)[i],colnames(mytab)[j],data_all[data_all$ChildID==rownames(mytab)[i],"Age"][1],0)
+  colnames(zero_line)<-colnames(zoon_sums)
+  zoon_sums=rbind(zoon_sums,
+                zero_line)
+}
+
+lab_sums = data_all %>% group_by(ChildID, lab, Age) %>% count(.drop = FALSE)
 lab_sums$id=paste(lab_sums$ChildID,lab_sums$lab)
-sums=merge(zoon_sums,lab_sums[,c("lab" ,  "n", "id")],by="id",all=T)
+sums=merge(zoon_sums,lab_sums[,c("lab", "Age" ,  "n", "id")],by="id",all=T)
+sums$age=ifelse(!is.na(sums$Age.x),sums$Age.x,sums$Age.y)
+summary(sums)
+
+l_nl_z=zoon_sums$Zoon_classif
+
+```
+
+```{r corage}
+
+plot()
 

### 2020-08-05 SLT_paper.Rmd
first full version of analyses, knits

--- 2020-08-04 SLT_paper.Rmd
@@
 title: "SLT_paper"
-output: html_document
+author: "AC"
+date: "2020-03-06 (substantive version), latest minor edits 2020-07-17"
+output:
+  pdf_document:
+    toc: yes
+    toc_depth: 3
+  html_document:
+    df_print: paged
+    toc: yes
+    toc_depth: '3'
 ---
@@
 
+## History:
+
+- 2020-08-05 final first version
+
+## Read data in
+
 ```{r}
-read.csv("first_class_lisa.csv")->zoon
-summary(zoon)
-
-read.csv("Metadata_batch2_lisa.csv",sep="\t")->md
-summary(md)
-#? correct errors in md? onset           offset       chunk_pos
-
-read.csv("result_final_lisa.csv")->lab
-summary(lab)
-
+# read dataset composed with python
 data_all <- read.csv("final_classifications_PU_zoon.csv")
@@
 
+#relevel the factor so that it's easier to read
 data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=c("Canonical","Non-Canonical",
@@
 data_all$lab=factor(data_all$lab,levels=levels(data_all$Zoon_classif))
-
+#apply same factor levels as zooniverse so that we can do symmetrical confusion matrices
 ```
 
-## correspondence
-i) quantifying the correspondence between expert human annotations and Zooniverse annotations %(confusion matrix) 
-and
+## Correspondence between lab & zooniverse annotation at the level of segments
+
+Here we look at to what extent zooniverse and lab annotations match at the level of individual segments. Each data point is one segment (one "vocalization").
 
@@
 
+mycf
+
 ```
+
+### Precision
+
+Precision means: If a segment was called X by zooniverse coders, what proportion of the time was it called X by lab coders?
 
@@
 ```
+
+### Recall
+
+Recall means: If a segment was called X by lab coders, what proportion of the time was it called X by zooniverse coders?
 
@@
 
-## What is said about infants
+## Child level descriptors
 
-ii) examining the correspondences between descriptors at the child level across the two annotation settings (e.g. CR-AGE correlations).
+Although there may be errors at the level of the segment, what we really care about is whether Zooniverse annotations give a reliable image of the child's individual development.
+This is what we look at in this section.
 
-```{r by child}
+```{r prepare data by child}
 
@@
 
-zoon_sums = data_all %>% group_by(ChildID,Zoon_classif) %>% tally()
-colnames(zoon_sums)<-c("ChildID",      "Zoon_classif", "n_zoo")
-lab_sums = data_all %>% group_by(ChildID,lab) %>% tally()
-zoon_sums$id=paste(zoon_sums$ChildID,zoon_sums$Zoon_classif)
-lab_sums$id=paste(lab_sums$ChildID,lab_sums$lab)
-sums=merge(zoon_sums,lab_sums[,c("lab" ,  "n", "id")],by="id",all=T)
+#get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
+ztab=table(data_all$ChildID,data_all$Zoon_classif)
+z_lr=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
+z_cr=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
+
+ltab=table(data_all$ChildID,data_all$lab)
+l_lr=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
+l_cr=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
+
+#put all the ratios together
+if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) ratios=cbind(rownames(ztab),z_lr,z_cr,l_lr,l_cr) else print("oops this code needs to be more complex because we don't have the same kids for the two ratios")
+colnames(ratios)[1]<-"ChildID"
+
+#add age
+ages=aggregate(data_all$Age,by=list(data_all$ChildID),mean) #this is a weird way of adding ages, since all of the ages for a given child should be the same if there is only one recording, and if there are multiple recordings, then we should not get the mean
+
+merge(ratios,ages,by.x="ChildID",by.y="Group.1")->ratios
+colnames(ratios)[dim(ratios)[2]]<-"Age"
+
+#cbinding results in text, so we numerize the ratios
+for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) ratios[,thisvar]=as.numeric(as.character(ratios[,thisvar]))
+summary(ratios)
+```
+
+We first look generally at two measures that have been found to relate to age:
+
+- linguistic ratio = ("Canonical"+"Non-Canonical")/"All vocalizations" (i.e. we remove junk)
+- canonical ratio = "Canonical"/("Canonical"+"Non-Canonical") (i.e. we remove junk + non-linguistic vocalizations)
+
+As expected, linguistic ratio goes up with age.
+
+Surprisingly, canonical ratio goes DOWN with age.
+
+```{r corage}
+prettynames=c("Linguistic Ratio (Zooniverse)","Canonical Ratio (Zooniverse)",
+             "Linguistic Ratio (Lab)","Canonical Ratio (Lab)" )
+names(prettynames)<-c("z_lr","z_cr","l_lr","l_cr")
+for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
+  myr=round(cor.test(ratios[,thisvar],ratios$Age)$estimate,3)
+  plot(ratios[,thisvar]~ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr))
+  abline(lm(ratios[,thisvar]~ratios$Age))
+  }
 
@@
 
-# Canonical Rate analysis by Age & ChildID
+But the key thing for us: Are Zooniverse annotations describing children similar to lab annotations? The answer is clearly yes.
 
-```{r CR by Age}
-df.sub <- subset(data_all, Zoon_classif %in% c("Canonical","Non-Canonical"))
+```{r corlab-zoo}
 
-### By Age
-n.can <- as.data.frame(df.sub %>% group_by(Age) %>% filter(Zoon_classif == "Canonical") %>% summarise(n.can = n()))
-n.non.can <- as.data.frame(df.sub %>% group_by(Age) %>% filter(Zoon_classif == "Non-Canonical") %>% summarise(n.non.can = n()))
-n.tot <- merge(n.can, n.non.can)
-n.tot$ratio.can <- n.tot$n.can/(n.tot$n.can + n.tot$n.non.can)
-n.tot$ratio.non.can <- n.tot$n.non.can/(n.tot$n.can + n.tot$n.non.can)
+#Ling ratio
+pdf("ling_rat_z_vs_l.pdf",height=5,width=5)
+lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]))
+  myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
+  plot(ratios[,"z_lr"]~ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
+       xlim=lims,ylim=lims)
+  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]))
+  lines(c(0,1),c(0,1),lty=2,col="darkgray")
+dev.off()
 
-### CANONICAL
-df.plot <- rbind(data.frame(Age = as.factor(n.tot$Age), Ratio = n.tot$ratio.can, Cat = "Canonical"))
-ggplot(df.plot, aes(fill=Cat, y=Ratio, x=Age)) + 
-  geom_bar(position="dodge", stat="identity")
-
-### NONCANONICAL
-df.plot <- rbind(data.frame(Age = as.factor(n.tot$Age), Ratio = n.tot$ratio.non.can, Cat = "Non-Canonical"))
-ggplot(df.plot, aes(fill=Cat, y=Ratio, x=Age)) + 
-  geom_bar(position="dodge", stat="identity")
-
-###BOTH
-
-df.plot <- rbind(data.frame(Age = as.factor(n.tot$Age), Ratio = n.tot$ratio.can, Cat = "Canonical"),data.frame(Age = as.factor(n.tot$Age), Ratio = n.tot$ratio.non.can, Cat = "Non-canonical"))
-
-ggplot(df.plot, aes(fill=Cat, y=Ratio, x=Age)) + 
-  geom_bar(position="dodge", stat="identity")
+  #CR
+pdf("can_rat_z_vs_l.pdf",height=5,width=5)
+lims=range(c(ratios[,"z_cr"],ratios[,"l_cr"]))
+    myr=round(cor.test(ratios[,"z_cr"],ratios[,"l_cr"])$estimate,3)
+  plot(ratios[,"z_cr"]~ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
+       xlim=lims,ylim=lims)
+  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]),col="darkgray")
+    lines(c(0,1),c(0,1),lty=2,col="darkgray")
+dev.off()
 ```
-
-```{r ChildID}
-### By Recording (=Child ID)
-n.can <- as.data.frame(df.sub %>% group_by(ChildID) %>% filter(Zoon_classif == "Canonical") %>% summarise(n.can = n()))
-n.non.can <- as.data.frame(df.sub %>% group_by(ChildID) %>% filter(Zoon_classif == "Non-Canonical") %>% summarise(n.non.can = n()))
-n.tot <- merge(n.can, n.non.can)
-n.tot$ratio.can <- n.tot$n.can/(n.tot$n.can + n.tot$n.non.can)
-n.tot$ratio.non.can <- n.tot$n.non.can/(n.tot$n.can + n.tot$n.non.can)
-df.plot <- rbind(data.frame(ChildID = as.factor(n.tot$ChildID), Ratio = n.tot$ratio.can, Cat = "Canonical"),
-                 data.frame(ChildID = as.factor(n.tot$ChildID), Ratio = n.tot$ratio.non.can, Cat = "Non-Canonical"))
-
-ggplot(df.plot, aes(fill=Cat, y=Ratio, x=ChildID)) + 
-  geom_bar(position="dodge", stat="identity")
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-08-05 SLT_paper.Rmd
improvements

@@
 ```{r}
+#read demo data created by AC from info in paper
+demo_data=read.csv("demo-data.tsv",sep="\t")
+
 # read dataset composed with python
@@
 This is what we look at in this section.
+In all of these graphs, red points correspond to children diagnosed with Angelman Syndrome, black for low-risk control.
 
@@
 #add age
-ages=aggregate(data_all$Age,by=list(data_all$ChildID),mean) #this is a weird way of adding ages, since all of the ages for a given child should be the same if there is only one recording, and if there are multiple recordings, then we should not get the mean
+#ages=aggregate(data_all$Age,by=list(data_all$ChildID),mean) #this is a weird way of adding ages, since all of the ages for a given child should be the same if there is only one recording, and if there are multiple recordings, then we should not get the mean
+#improvement: now we merge with a demo data tab, but note this is merged with child id, so the problem of multiple recs per child is still there
 
-merge(ratios,ages,by.x="ChildID",by.y="Group.1")->ratios
+merge(ratios,demo_data,by="ChildID")->ratios
 colnames(ratios)[dim(ratios)[2]]<-"Age"
@@
 names(prettynames)<-c("z_lr","z_cr","l_lr","l_cr")
+
+#!!! this is working the opposite than it should! but note that to get angsynd kids to come out in red, I have to do it like this -- it doesn't make sense...
+mycols=c("red","black")
+names(mycols)<-c("Low-RiskControl","AngelmanSyndrome")
+
 for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
   myr=round(cor.test(ratios[,thisvar],ratios$Age)$estimate,3)
-  plot(ratios[,thisvar]~ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr))
+  plot(ratios[,thisvar]~ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr),
+       col=mycols[ratios$Diagnosis])
   abline(lm(ratios[,thisvar]~ratios$Age))
@@
   plot(ratios[,"z_lr"]~ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims)
+       xlim=lims,ylim=lims,
+       col=mycols[ratios$Diagnosis])
   abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]))
@@
   plot(ratios[,"z_cr"]~ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims)
+       xlim=lims,ylim=lims,
+       col=mycols[ratios$Diagnosis])
   abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]),col="darkgray")

### 2020-08-05 SLT_paper.Rmd
improvements

@@
 ```{r prec}
-pdf("precision_with_mixed.pdf",height=10,width=10)
+
+prop_cat=data.frame(conf_tab/colSums(conf_tab)*100) #generates precision because columns
+prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
+colnames(prop_cat)[3]<-"pr"
+
+data.frame(conf_tab)->stall
+stall$id=paste(stall$Lab,stall$Zooniverse)
+stall=merge(stall,prop_cat[c("id","pr")])
+
+ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
+ geom_tile(aes(fill= rescale(pr)), colour = "white") +
+  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=2) +
+  geom_text(aes(label = Freq), vjust = 1,size=2) +
+  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
+     theme(legend.position = "none") +
+  xlab("Zooniverse") + ylab("Lab") +
+  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+```
+
+### Recall
+
+Recall means: If a segment was called X by lab coders, what proportion of the time was it called X by zooniverse coders?
+
+```{r rec}
+prop_cat=data.frame(conf_tab/colSums(conf_tab)*100) #generates precision because columns
+prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
+colnames(prop_cat)[3]<-"pr"
+
+data.frame(conf_tab)->stall
+stall$id=paste(stall$Lab,stall$Zooniverse)
+stall=merge(stall,prop_cat[c("id","pr")])
+
+ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
+ geom_tile(aes(fill= rescale(pr)), colour = "white") +
+  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=2) +
+  geom_text(aes(label = Freq), vjust = 1,size=2) +
+  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
+     theme(legend.position = "none") +
+  xlab("Zooniverse") + ylab("Lab") +
+  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+```
+
+## repeat collapsing
+
+```{r}
+
+#given results above, we map the mixed
+data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical"]<-"Canonical"
+data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical"]<-"Non-Canonical"
+data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical_Crying"]<-"Non-Canonical"
+data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Crying"]<-"Crying"
+data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Crying"]<-"Non-Canonical"
+data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing_Crying"]<-"Non-Canonical"
+
+#and reset the factors for cleanliness
+data_all$Zoon_classif=factor(data_all$Zoon_classif)
+data_all$lab=factor(data_all$lab)
+
+mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
+conf_tab=mycf$table
+
+# this package uses sensitivity & specificity
+#Sensitivity=recall
+#Specificity=precision
+
+mycf
+
+```
+
+```{r prec2}
+pdf("precision.pdf",height=10,width=10)
 prop_cat=data.frame(conf_tab/colSums(conf_tab)*100) #generates precision because columns
@@
 
-### Recall
-
-Recall means: If a segment was called X by lab coders, what proportion of the time was it called X by zooniverse coders?
-
-```{r recall}
-pdf("recall_with_mixed.pdf",height=10,width=10)
+```{r recall2}
+pdf("recall.pdf",height=10,width=10)
 prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
@@
 ```
-
 ## Child level descriptors
@@
 ```{r prepare data by child}
-
-#given results above, we map the mixed
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical"]<-"Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical_Crying"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Crying"]<-"Crying"
-data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Crying"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing_Crying"]<-"Non-Canonical"
-
-#and reset the factors for cleanliness
-data_all$Zoon_classif=factor(data_all$Zoon_classif)
-data_all$lab=factor(data_all$lab)
 

### 2020-08-06 SLT_paper.Rmd
improved figs, added one

@@
  geom_tile(aes(fill= rescale(pr)), colour = "white") +
-  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=2) +
-  geom_text(aes(label = Freq), vjust = 1,size=2) +
-  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
+  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=8) +
+  geom_text(aes(label = Freq), vjust = 1,size=8) +
+  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
      theme(legend.position = "none") +
   xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  ggtitle("Precision")+theme(text = element_text(size=20),
+        axis.text.x = element_text(angle=90, hjust=1))
 dev.off()
@@
  geom_tile(aes(fill= rescale(rec)), colour = "white") +
-  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=2) +
-  geom_text(aes(label = Freq), vjust = 1,size=2) +
-  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
+  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=8) +
+  geom_text(aes(label = Freq), vjust = 1,size=8) +
+  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
      theme(legend.position = "none") +
   xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Recall")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  ggtitle("Recall")+theme(text = element_text(size=20),
+        axis.text.x = element_text(angle=90, hjust=1))
 
@@
 dev.off()
+
+#COMBINED to save space
+pdf("combined.pdf",height=5,width=5)
+lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]),c(ratios[,"z_cr"],ratios[,"l_cr"]))
+  #myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
+
+  plot(ratios[,"z_lr"]~ratios[,"l_lr"],xlab="Laboratory annotations",ylab="Zooniverse annotations",
+       xlim=lims,ylim=lims,
+       pch=20,col=mycols[ratios$Diagnosis])
+    points(ratios[,"z_cr"]~ratios[,"l_cr"], pch=2, col=mycols[ratios$Diagnosis])
+  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]))
+  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]),lty=3)
+ # lines(c(0,1),c(0,1),lty=2,col="darkgray")
+dev.off()
+
 ```

### 2020-08-07 SLT_paper.Rmd
Latest data
+ repo reorganization

@@
 #read demo data created by AC from info in paper
-demo_data=read.csv("demo-data.tsv",sep="\t")
+demo_data=read.csv("./Data/demo-data.tsv",sep="\t")
 
 # read dataset composed with python
-data_all <- read.csv("final_classifications_PU_zoon.csv")
+data_all <- read.csv("./Data/new_classifications_PU_zoon.csv")
 
@@
 ```{r prec2}
-pdf("precision.pdf",height=10,width=10)
+pdf("./Results/precision.pdf",height=10,width=10)
 prop_cat=data.frame(conf_tab/colSums(conf_tab)*100) #generates precision because columns
@@
 ```{r recall2}
-pdf("recall.pdf",height=10,width=10)
+pdf("./Results/recall.pdf",height=10,width=10)
 prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
@@
 #Ling ratio
-pdf("ling_rat_z_vs_l.pdf",height=5,width=5)
+pdf("./Results/ling_rat_z_vs_l.pdf",height=5,width=5)
 lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]))
@@
   #CR
-pdf("can_rat_z_vs_l.pdf",height=5,width=5)
+pdf("./Results/can_rat_z_vs_l.pdf",height=5,width=5)
 lims=range(c(ratios[,"z_cr"],ratios[,"l_cr"]))
@@
 #COMBINED to save space
-pdf("combined.pdf",height=5,width=5)
+pdf("./Results/combined.pdf",height=5,width=5)
 lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]),c(ratios[,"z_cr"],ratios[,"l_cr"]))

### 2020-09-15 SLT_paper.Rmd
Fixed prec/recall

@@
 # read dataset composed with python
+#data_ang <- read.csv("./Data/final_classifications_PU_zoon.csv")
+#data_td <- read.csv("./Data/final_class_september7.csv")
+#data_all<-rbind(data_ang, data_td)
+
+#read demo data created by AC from info in paper
+demo_data=read.csv("./Data/demo-data.tsv",sep="\t")
+# read dataset composed with python
 data_all <- read.csv("./Data/new_classifications_PU_zoon.csv")
-
+#remove the word mixed that takes up space and is unnecessary
+data_all$Zoon_classif=factor(gsub("Mixed_","",as.character(data_all$Zoon_classif),fixed=T))
+#relevel the factor so that it's easier to read
+data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=c("Canonical","Non-Canonical",
+                                  "Crying","Laughing","Junk",levels(data_all$Zoon_classif)[grep("_",levels(data_all$Zoon_classif))]))
 #remove the word mixed that takes up space and is unnecessary
@@
 pdf("./Results/precision.pdf",height=10,width=10)
-prop_cat=data.frame(conf_tab/colSums(conf_tab)*100) #generates precision because columns
+colsums=colSums(conf_tab)
+my_conf_tab=conf_tab
+for(i in 1:5) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
+colSums(my_conf_tab)
+prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
 prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
@@
   xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Precision")+theme(text = element_text(size=20),
+  ggtitle("Precision")+theme(text = element_text(size=10),
         axis.text.x = element_text(angle=90, hjust=1))
+dev.off()
+```
+```{r recall2}
+colsums=colSums(conf_tab)
+pdf("./Results/precision_all.pdf",height=10,width=10)
+prop_cat=data.frame(conf_tab/colSums(conf_tab)*100)  #generates precision because columns
+prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
+colnames(prop_cat)[3]<-"prec"
+
+data.frame(conf_tab)->stall
+stall$id=paste(stall$Lab,stall$Zooniverse)
+stall=merge(stall,prop_cat[c("id","prec")])
+
+ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
+ geom_tile(aes(fill= rescale(prec)), colour = "white") +
+  geom_text(aes(label = paste(round(prec),"%")), vjust = -1,size=8) +
+  geom_text(aes(label = Freq), vjust = 1,size=8) +
+  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
+     theme(legend.position = "none") +
+  xlab("Zooniverse") + ylab("Lab") +
+  ggtitle("Precision")+theme(text = element_text(size=10),
+        axis.text.x = element_text(angle=90, hjust=1))
+
 dev.off()
@@
 ```{r recall2}
-pdf("./Results/recall.pdf",height=10,width=10)
+pdf("./Results/recall_all.pdf",height=10,width=10)
 prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
@@
   xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Recall")+theme(text = element_text(size=20),
+  ggtitle("Recall")+theme(text = element_text(size=10),
         axis.text.x = element_text(angle=90, hjust=1))

### 2020-09-15 SLT_paper.Rmd
found, probably has extra metric

--- 2020-08-07 SLT_paper.Rmd
@@
 library(scales)
+library(irrCAC) #https://cran.r-project.org/web/packages/irrCAC/vignettes/overview.html
 
@@
 - 2020-08-05 final first version
+- 2020-08-08 fixed error in precision (cols didn't add to 1)
+- 2020-08-11 added AC1 as alt to kappa
 
@@
 
-prop_cat=data.frame(conf_tab/colSums(conf_tab)*100) #generates precision because columns
+colsums=colSums(conf_tab)
+my_conf_tab=conf_tab
+for(i in 1:5) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
+colSums(my_conf_tab)
+prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
 prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
@@
   geom_text(aes(label = Freq), vjust = 1,size=2) +
-  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
+  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
      theme(legend.position = "none") +
   xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  ggtitle("Precision")+theme(text = element_text(size=20),
+        axis.text.x = element_text(angle=90, hjust=1))
 
@@
 
-### Recall
-
+### recalll
 Recall means: If a segment was called X by lab coders, what proportion of the time was it called X by zooniverse coders?
@@
 ```{r rec}
-prop_cat=data.frame(conf_tab/colSums(conf_tab)*100) #generates precision because columns
+
+prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
 prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
-colnames(prop_cat)[3]<-"pr"
+colnames(prop_cat)[3]<-"rec"
 
@@
 stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","pr")])
+stall=merge(stall,prop_cat[c("id","rec")])
 
 ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(pr)), colour = "white") +
-  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=2) +
-  geom_text(aes(label = Freq), vjust = 1,size=2) +
-  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
+ geom_tile(aes(fill= rescale(rec)), colour = "white") +
+  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=8) +
+  geom_text(aes(label = Freq), vjust = 1,size=8) +
+  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
      theme(legend.position = "none") +
   xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-
+  ggtitle("Recall")+theme(text = element_text(size=20),
+        axis.text.x = element_text(angle=90, hjust=1))
 ```
 
-## repeat collapsing
+## Repeat collapsing mixed onto the relevant category
+
+Given results above, we map the mixed onto the corresponding key classes.
 
@@
 
-#given results above, we map the mixed
+#
 data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical"]<-"Canonical"
@@
 
+gwet.ac1.table(conf_tab)
+
 ```

### 2020-10-17 Scripts/slt_again.Rmd
Chiara's version, rmd does not compile

---
title: "SLT_paper"
author: "AC"
date: "2020-03-06 (substantive version), latest minor edits 2020-07-17"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
library(tidyverse)
library(dplyr)
library(ggplot2)
library(caret)
library(scales)
dodiv=function(x) x/sum(x, na.rm=T)
```

## History:

- 2020-08-05 final first version

## Read data in

```{r}
#read demo data created by AC from info in paper
# read datasets

demo_data=read.csv("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Data/demo-data.tsv",sep="\t")
data_ang <- read.csv("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Data/classifications_PU_zoon_final17.csv",header=T,sep=",")
data_td <- read.csv("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Data/classifications_PU_zoon_final.csv")
data_all<-rbind(data_ang, data_td)

#add filenames to demo data
demo_data_fn <- demo_data %>% 
     left_join(select(data_all, filename, ChildID), by = c("ChildID"))
demo_data_fn<-unique(demo_data_fn)

#remove the word mixed that takes up space and is unnecessary
data_all$Zoon_classif=factor(gsub("Mixed_","",as.character(data_all$Zoon_classif),fixed=T))
#relevel the factor so that it's easier to read
data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=c("Canonical","Non-Canonical",                   "Crying","Laughing","Junk",levels(data_all$Zoon_classif)[grep("_",levels(data_all$Zoon_classif))]))
# create lab column with easier to read correspondance
data_all$lab<-as.character(data_all$Major_Choice)
data_all$lab[data_all$lab=="Non-canonical syllables"]<-"Non-Canonical"
data_all$lab[data_all$lab=="Canonical syllables"]<-"Canonical"
data_all$lab[data_all$lab %in% c("Don't mark","None")]<-"Junk"
data_all$lab=factor(data_all$lab,levels=levels(data_all$Zoon_classif))
#apply same factor levels as zooniverse so that we can do symmetrical confusion matrices
```

## Correspondence between lab & zooniverse annotation at the level of segments

Here we look at to what extent zooniverse and lab annotations match at the level of individual segments. Each data point is one segment (one "vocalization").

```{r}
table(data_all$lab)
table(data_all$Zoon_classif)
mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
# this package uses sensitivity & specificity
#Sensitivity=recall
#Specificity=precision
mycf
```

### Precision

Precision means: If a segment was called X by zooniverse coders, what proportion of the time was it called X by lab coders?

```{r prec}
colsums=colSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:18) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
colSums(my_conf_tab)
prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"pr"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","pr")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(pr)), colour = "white") +
  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=2) +
  geom_text(aes(label = Freq), vjust = 1,size=1) +
  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Recall

Recall means: If a segment was called X by lab coders, what proportion of the time was it called X by zooniverse coders?

```{r rec}
rowsums=rowSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:18) my_conf_tab[,i]=my_conf_tab[,i]/rowsums[i]
rowSums(my_conf_tab)
prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"rec"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","rec")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(rec)), colour = "white") +
  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=2) +
  geom_text(aes(label = Freq), vjust = 1,size=1) +
  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Recall")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

## repeat collapsing

```{r}
#given results above, we map the mixed
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical"]<-"Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical_Crying"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Crying"]<-"Crying"
data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Crying"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing_Crying"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Canonical"]<-"Canonical"
# +
data_all$Zoon_classif[data_all$Zoon_classif=="Canonical_Crying"]<-"Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Canonical_Laughing"]<-"Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical_Crying"]<-"Non-Canonical"

data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Laughing"]<-"Crying"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Canonical_Laughing"]<-"Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Laughing_Non-Canonical"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Non-Canonical"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Non-Canonical_Laughing"]<-"Non-Canonical" 
data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing"]<-"Non-Canonical"

```

```{r}
#and reset the factors for cleanliness
data_all$Zoon_classif=factor(data_all$Zoon_classif)
data_all$lab=factor(data_all$lab)
sample_data<-cbind(data_all$lab,data_all$Zoon_classif)
gac(data = sample_data, kat = 5, weight = c("unweighted"),
    conf.level = 0.95)
mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
# this package uses sensitivity & specificity
#Sensitivity=recall
#Specificity=precision
mycf
```

```{r prec2}
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/precision_final.pdf",height=10,width=10)
colsums=colSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:5) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
colSums(my_conf_tab)
prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"pr"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","pr")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(pr)), colour = "white") +
  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=8) +
  geom_text(aes(label = Freq), vjust = 1,size=8) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Precision")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))
dev.off()
```

```{r recall2}
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/recall_final.pdf",height=10,width=10)
prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"rec"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","rec")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(rec)), colour = "white") +
  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=8) +
  geom_text(aes(label = Freq), vjust = 1,size=8) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Recall")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))
dev.off()
```

```{r}
data_as_td<-left_join(data_all,demo_data,on="ChildID")
# CM with just AS kids
data_AS<-subset(data_as_td, Diagnosis=="AngelmanSyndrome")
mycf=confusionMatrix(data_AS$lab, data_AS$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
mycf
```
```{r prec2AS}
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/precision_AS.pdf",height=10,width=10)
colsums=colSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:5) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
colSums(my_conf_tab)
prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"pr"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","pr")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(pr)), colour = "white") +
  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=8) +
  geom_text(aes(label = Freq), vjust = 1,size=8) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Precision")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))
dev.off()
```

```{r recall2}
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/recall_AS.pdf",height=10,width=10)
prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"rec"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","rec")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(rec)), colour = "white") +
  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=8) +
  geom_text(aes(label = Freq), vjust = 1,size=8) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Recall")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))
dev.off()
```

```{r}
# CM with just TD kids
data_TD<-subset(data_as_td, Diagnosis=="Low-RiskControl")
mycf=confusionMatrix(data_TD$lab, data_TD$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
mycf
```

```{r prec2AS}
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/precision_TD.pdf",height=10,width=10)
colsums=colSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:5) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
colSums(my_conf_tab)
prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"pr"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","pr")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(pr)), colour = "white") +
  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=8) +
  geom_text(aes(label = Freq), vjust = 1,size=8) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Precision")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))
dev.off()
```

```{r recall2}
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/recall_TD.pdf",height=10,width=10)
prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"rec"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","rec")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(rec)), colour = "white") +
  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=8) +
  geom_text(aes(label = Freq), vjust = 1,size=8) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Recall")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))
dev.off()
```

## Child level descriptors

Although there may be errors at the level of the segment, what we really care about is whether Zooniverse annotations give a reliable image of the child's individual development. This is what we look at in this section. In all of these graphs, red points correspond to children diagnosed with Angelman Syndrome, black for low-risk control.

```{r prepare data by child}
#get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
ztab=table(data_all$filename,data_all$Zoon_classif)
z_lr=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
z_cr=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
ltab=table(data_all$filename,data_all$lab)
l_lr=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
l_cr=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
#put all the ratios together
if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) ratios=cbind(rownames(ztab),z_lr,z_cr,l_lr,l_cr) else print("oops this code needs to be more complex because we don't have the same kids for the two ratios")
colnames(ratios)[1]<-"filename"

#ages=aggregate(data_all$Age,by=list(data_all$ChildID),mean) #this is a weird way of adding ages, since all of the ages for a given child should be the same if there is only one recording, and if there are multiple recordings, then we should not get the mean
#improvement: now we merge with a demo data tab, but note this is merged with child id, so the problem of multiple recs per child is still there

# Created demo_data with filenames. Use filenames instead of childIDs to merge ratios and demo data.

merge(ratios,demo_data_fn,by="filename")->ratios
colnames(ratios)[dim(ratios)[2]]<-"Age"
#cbinding results in text, so we numerize the ratios
for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) ratios[,thisvar]=as.numeric(as.character(ratios[,thisvar]))
summary(ratios)
```

We first look generally at two measures that have been found to relate to age:

- linguistic ratio = ("Canonical"+"Non-Canonical")/"All vocalizations" (i.e. we remove junk)
- canonical ratio = "Canonical"/("Canonical"+"Non-Canonical") (i.e. we remove junk + non-linguistic vocalizations)

As expected, linguistic ratio goes up with age.

Surprisingly, canonical ratio goes DOWN with age.

```{r corage}
prettynames=c("Linguistic Ratio (Zooniverse)","Canonical Ratio (Zooniverse)",
             "Linguistic Ratio (Lab)","Canonical Ratio (Lab)" )
names(prettynames)<-c("z_lr","z_cr","l_lr","l_cr")
#!!! this is working the opposite than it should! but note that to get angsynd kids to come out in red, I have to do it like this -- it doesn't make sense...
# Now it works alright
mycols=c("black","red")
names(mycols)<-c("Low-RiskControl","AngelmanSyndrome")
for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
  myr=round(cor.test(ratios[,thisvar],ratios$Age)$estimate,3)
  plot(ratios[,thisvar]~ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr),
       col=mycols[ratios$Diagnosis])
  abline(lm(ratios[,thisvar]~ratios$Age))
  }
```

But the key thing for us: Are Zooniverse annotations describing children similar to lab annotations? The answer is clearly yes.

```{r corlab-zoo}
#Ling ratio
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/ling_rat_z_vs_l_final.pdf",height=5,width=5)
lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]))
  myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
  plot(ratios[,"z_lr"]~ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[ratios$Diagnosis])
  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]))
  lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
  #CR
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/can_rat_z_vs_l_final.pdf",height=5,width=5)
lims=range(c(ratios[,"z_cr"],ratios[,"l_cr"]))
    myr=round(cor.test(ratios[,"z_cr"],ratios[,"l_cr"])$estimate,3)
  plot(ratios[,"z_cr"]~ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[ratios$Diagnosis])
  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]),col="darkgray")
    lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
#COMBINED to save space
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/combined_final.pdf",height=5,width=5)
lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]),c(ratios[,"z_cr"],ratios[,"l_cr"]))
  #myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
  plot(ratios[,"z_lr"]~ratios[,"l_lr"],xlab="Laboratory annotations",ylab="Zooniverse annotations",
       xlim=lims,ylim=lims,
       pch=20,col=mycols[ratios$Diagnosis])
    points(ratios[,"z_cr"]~ratios[,"l_cr"], pch=2, col=mycols[ratios$Diagnosis])
  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]))
  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]),lty=3)
 # lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
```

```{r corage}
prettynames=c("AS Linguistic Ratio (Zooniverse)","AS Canonical Ratio (Zooniverse)",
             "AS Linguistic Ratio (Lab)","AS Canonical Ratio (Lab)" )
names(prettynames)<-c("z_lr","z_cr","l_lr","l_cr")

AS_ratios<-subset(ratios, Diagnosis=="AngelmanSyndrome")
TD_ratios<-subset(ratios, Diagnosis=="Low-RiskControl")

for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
  myr=round(cor.test(AS_ratios[,thisvar],AS_ratios$Age)$estimate,3)
  plot(AS_ratios[,thisvar]~AS_ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr))
  abline(lm(AS_ratios[,thisvar]~AS_ratios$Age))
  }
```

```{r corage}
prettynames=c("TD Linguistic Ratio (Zooniverse)","TD Canonical Ratio (Zooniverse)",
             "TD Linguistic Ratio (Lab)","TD Canonical Ratio (Lab)" )
for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
  myr=round(cor.test(TD_ratios[,thisvar],TD_ratios$Age)$estimate,3)
  plot(TD_ratios[,thisvar]~TD_ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr))
  abline(lm(TD_ratios[,thisvar]~TD_ratios$Age))
  }
```
```{r corlab-zoo}
#Ling ratio
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/td_ling_rat_z_vs_l_final.pdf",height=5,width=5)
lims=range(c(TD_ratios[,"z_lr"],TD_ratios[,"l_lr"]))
  myr=round(cor.test(TD_ratios[,"z_lr"],TD_ratios[,"l_lr"])$estimate,3)
  plot(TD_ratios[,"z_lr"]~TD_ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[TD_ratios$Diagnosis])
  abline(lm(TD_ratios[,"z_lr"]~TD_ratios[,"l_lr"]))
  lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
  #CR
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/td_can_rat_z_vs_l_final.pdf",height=5,width=5)
lims=range(c(TD_ratios[,"z_cr"],TD_ratios[,"l_cr"]))
    myr=round(cor.test(TD_ratios[,"z_cr"],TD_ratios[,"l_cr"])$estimate,3)
  plot(TD_ratios[,"z_cr"]~TD_ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[TD_ratios$Diagnosis])
  abline(lm(TD_ratios[,"z_cr"]~TD_ratios[,"l_cr"]),col="darkgray")
    lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
```

```{r corlab-zoo}
#Ling ratio
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/AS_ling_rat_z_vs_l_final.pdf",height=5,width=5)
lims=range(c(AS_ratios[,"z_lr"],AS_ratios[,"l_lr"]))
  myr=round(cor.test(AS_ratios[,"z_lr"],AS_ratios[,"l_lr"])$estimate,3)
  plot(AS_ratios[,"z_lr"]~AS_ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[AS_ratios$Diagnosis])
  abline(lm(AS_ratios[,"z_lr"]~AS_ratios[,"l_lr"]))
  lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
  #CR
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/AS_can_rat_z_vs_l_final.pdf",height=5,width=5)
lims=range(c(AS_ratios[,"z_cr"],AS_ratios[,"l_cr"]))
    myr=round(cor.test(AS_ratios[,"z_cr"],AS_ratios[,"l_cr"])$estimate,3)
  plot(AS_ratios[,"z_cr"]~AS_ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[AS_ratios$Diagnosis])
  abline(lm(AS_ratios[,"z_cr"]~AS_ratios[,"l_cr"]),col="darkgray")
    lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
```
```{r}
lin_mod=lm(z_lr~l_lr+Diagnosis,data=ratios)
summary(lin_mod)
```

### 2020-10-17 Scripts/chunk_to_segment.py

@@ -1,0 +1,87 @@
+import numpy as np
+from collections import defaultdict
+import json
+
+# ----------------------------------------------------------------------------------------
+
+#This script creates a dictionary file in json format with mapping from 500ms ->chunks<- to Lab annotation ->segments<-
+
+
+result_final="/Users/chiarasemenzin/Desktop/LSCP/SLT/result_final_lisa.csv"  #file with segment info and lab classifications
+result_zoon="/Users/chiarasemenzin/Desktop/LSCP/SLT/merged_metadata.csv"  #metadata for zooniverse clips
+
+
+# Segment to timestamp
+def seg_to_ts(result_final):
+    seg2ts={}
+    with open(result_final) as final:
+        for e,line in enumerate(final):
+            line=line.split(",")
+            if e > 0:
+                seg2ts[line[0]]=(float(line[4]),float(line[5]),line[1])
+    return seg2ts
+
+# Timestamp to segment
+def chunk_to_ts(result_zoon):
+    chunk2ts={}
+    with open(result_zoon) as zoon:
+        for e,line in enumerate(zoon):
+            line=line.split(",")
+            if e > 0:
+                #chunk2ts[line[0]]=[float(i) for i in line[4].split("_")]
+                chunk2ts[line[1]]=[float(line[4]),float(line[5])] #get timestamps
+                chunk2ts[line[1]].append(line[2]) #get name of chunk
+    return chunk2ts
+
+# Calculate average duration of segment
+def avg_duration(result_final):
+    seg_ts=[]
+    with open(result_final) as final:
+        for e,line in enumerate(final):
+            line=line.split(",")
+            if e > 0:
+                seg_ts.append(float(line[5])-float(line[4]))
+    avg_dur=sum(seg_ts)/len(seg_ts)
+    return avg_duration
+
+def check(segments,chunks):
+    """" Check which chunks belong to which segment, by file"""
+
+    segments_final = defaultdict(list)
+
+    for name,chunk in chunks.items(): 
+        diffs_onset=[]
+        diffs_offeset=[]
+        candidate_seg=dict([i for i in segments.items() if i[1][2]==chunk[2]])
+        for k,seg in candidate_seg.items():
+            if seg[2] == chunk[2]:
+                print("Looking at ",k,seg[2])
+                diffs_onset.append([abs(chunk[0]-seg[0]),k])
+                diffs_offeset.append([abs(chunk[1]-seg[1]),k])
+        min_on=min([i[0] for i in diffs_onset])
[…]

### 2020-10-17 Scripts/merge_datasets_zoon.py

@@ -1,0 +1,117 @@
+import pandas as pd
+from collections import OrderedDict
+import csv
+
+print("Started.")
+
+# define paths to your existing zooniverse classifications and metadata locations
+# These are the files you get as output from the Jupyter Notebook convert_zooniverse.ipynb
+
+classifs=pd.read_csv("/Users/chiarasemenzin/Desktop/LSCP/SLT/zooniverse_data_all_final.csv")
+metadata=pd.read_csv("/Users/chiarasemenzin/Desktop/zoon_metadata/Alex_metadata/MetaData_LAAC_20200504_ac.csv",sep='\t')
+filename_links=pd.read_csv("/Users/chiarasemenzin/filename_links.csv")
+#first_class=pd.read_csv("/Users/chiarasemenzin/Desktop/LSCP/experiment/results/lwl_results/first_classifications_last.csv")
+#classifs=pd.read_csv("/Users/chiarasemenzin/Desktop/zoon_metadata/classifs/zooniverse_class_june.csv")
+#metadata_phonses = pd.read_csv("/Users/chiarasemenzin/Desktop/zoon_metadata/metadata-phonses-merged.csv")
+
+# Clean dataset from first pilot files
+patternDel=".wav.mp3"
+filter = classifs['AudioData'].str.contains(patternDel)
+classifs_filter=classifs[~filter]
+
+
+def merge_dataset_simple(classifs,metadata,get_stats=True):
+    """Match classifications with metadata file without info other than clip name (e.g. rttms) """
+    df=OrderedDict({"AudioData":[], "Answer":[]})
+    print("Creating merged dataset...")
+    counter=0
+    clips=0
+    for i,j in metadata.iterrows():
+        matches=classifs[classifs.AudioData.str.contains(str(j.AudioData))]
+        counter+=matches.shape[0]
+        print("found {} matches".format(matches.shape[0]))
+        if matches.shape[0] != 0:
+            clips+=1
+        for answer in matches.Answer.tolist():
+            df["AudioData"].append(j.AudioData)
+            df["Answer"].append(answer)
+#    print("Count "+str(counter))
+    print("writing to CSV...")
+    df_new = pd.DataFrame(df, columns=df.keys())
+    df_new.to_csv("zoon_classifs_sep15.csv")
+    if get_stats:
+        log = open("/Users/chiarasemenzin/Desktop/SLT/logs/stats_log.txt", "w")
+        log.write("Total labels: {} \nTotal clips: {}".format(counter,clips))
[…]

### 2020-10-17 Scripts/write_seg_data.py

@@ -1,0 +1,90 @@
+import pandas as pd
+import numpy as np
+from collections import defaultdict, Counter
+import json
+
+# ----------------------------------------------------------------------------------------
+# Load data
+
+result_zoon="/Users/chiarasemenzin/Desktop/LSCP/SLT/classifications_17sept.csv" #classifs from zooniverse merged with md
+zoon=pd.read_csv(result_zoon)
+lisa=pd.read_csv("/Users/chiarasemenzin/Desktop/LSCP/SLT/result_final_lisa.csv") #file with metadata and PU classifications
+stats_log="stats_log.txt" #give a name to the log 
+
+with open("/Users/chiarasemenzin/Desktop/LSCP/SLT/dict_3.json", "r") as read_file: # Put here path to json file 
+    print("Converting JSON encoded data into Python dictionary")
+    d = json.load(read_file)
+
+
+# ----------------------------------------------------------------------------------------
+
+
+def agreement_by_chunk(d,classifs):
+    """Determine majority agreement within chunks"""
+    answers = defaultdict(list)
+    no_maj_count=0
+    no_3_labels=0
+    valid_count=0
+    total_labels=0
+    for k,v in d.items():
+        for chunk in v:
+            c=Counter(list(classifs.loc[classifs["AudioData"] == int(chunk)]["Answer"]))
+            total_labels+=len(c)
+            if len(c)!= 0:
+                value, count = c.most_common()[0]
+                print(sum([i for i in c.values()]))
+                if count >= 3 and sum([i for i in c.values()])>=3: #at least 3 classifs agree
+                    answers[k].append(value)
+                    valid_count+=1
+                    #print("Okay")
+                elif count < 3 and sum([i for i in c.values()])>=3:
+                    #print("No majority agreement")
+                    no_maj_count+=1
+                else:
+                    #print("Not enough data")
+                    no_3_labels+=1
+    log = open(stats_log, "w")
+    log.write(" Valid clips: {}\n No majority agreement: {}\n Less than 3 labels: {}, total labels: {}".format(valid_count,no_maj_count,no_3_labels,total_labels))
+    log.close()
+    return answers
+
+
+def agreement_by_segment(answers):
+    """Determine classification within segment """
+#    TODO
+    seg_dict=[[],[]]
+    for k,v in answers.items():
+        if set(v) != {"Junk"}: # Check Junk is not the only answer
[…]

### 2020-10-17 Scripts/add-lena-ana.Rmd
Alex's version minus updates to slt_again, rmds do not compile

---
title: "Attempt to calculate LENA accuracy for ling/nonling"
author: "AC"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(scales)
library(caret)
```

## Read data in

```{r read-data,eval=F}
read.csv("../Metadata/result_final_lisa.csv")->metadata
metadata[1,]

csvs=dir("../../its_to_csv/",pattern="csv")

#count_present=NULL
all_segments=NULL
# extract info from csv's
for(thisits in levels(metadata$filename)){
  csv=grep(thisits,csvs)
  #count_present=c(count_present,length(csv))
  if(length(csv)==1){
    #read in full csv
    read.csv(paste0("../../its_to_csv/",csvs[csv]))->thisdat

    #select down to the lines used in zooniverse
    sel_segments_start=metadata[metadata$filename==thisits,"Starttime"]
    sel_segments_dat=thisdat[thisdat$startTime %in% sel_segments_start,]

    #the following is NOT right: we take the data in even if there is a mismatch in start/end time
    if(dim(sel_segments_dat)[1]!=length(sel_segments_start)) print(paste(thisits, csvs[csv], "have line mismatches:",dim(sel_segments_dat)[1],length(sel_segments_start)))
    if(dim(sel_segments_dat)[1]>0){

     #derive LENA classification at the segment level
    sel_segments_dat$LENA_type=NA
      #if ling info contained, then ling
    sel_segments_dat$LENA_type[is.na(sel_segments_dat$LENA_type) & sel_segments_dat$childUttLen>0 & !is.na(sel_segments_dat$childUttLen)]<-"Can/non-can"
      #elsif there is crying
    sel_segments_dat$LENA_type[is.na(sel_segments_dat$LENA_type) &  !is.na(sel_segments_dat$endCry1)]<-"Crying"
     #elsif there is fixed then laughing
    sel_segments_dat$LENA_type[is.na(sel_segments_dat$LENA_type) &  !is.na(sel_segments_dat$endVfx1)]<-"Laughing"
    #else junk
    sel_segments_dat$LENA_type[is.na(sel_segments_dat$LENA_type)]<-"Junk"

    sel_segments_dat=sel_segments_dat[,c("itsId","startTime", "endTime","LENA_type")]

    #combine with lab annotation
    sel_segments_dat=merge(sel_segments_dat,metadata[metadata$filename==thisits,],by.x="startTime",by.y="Starttime")

    #add to full spreadsheet
    all_segments=rbind(all_segments,sel_segments_dat)     
    }

  } else print(paste(thisits,"not found"))
}
write.csv(all_segments,"LENA_type_lab.csv",row.names = F)
```

## Correspondence between LENA & lab annotation at the level of segments

Here we look at to what extent LENA and lab annotations match at the level of individual segments. Each data point is one segment (one "vocalization").

```{r}

read.csv("LENA_type_lab.csv")->all_segments

## NOOOOOOOTE !!!! REMOVING Num_Agreement < 2
all_segments=all_segments[all_segments$Num_Agreement>=2,]

# create lab column with easier to read correspondance
all_segments$lab<-as.character(all_segments$Major_Choice)
all_segments$lab[all_segments$lab=="Non-canonical syllables"]<-"Non-Canonical"
all_segments$lab[all_segments$lab=="Canonical syllables"]<-"Canonical"
all_segments$lab[all_segments$lab=="Words"]<-"Canonical"  #!!! why didn't I need this before??
all_segments$lab[all_segments$lab %in% c("Don't mark","None")]<-"Junk"

#collapse acros can and non-can
all_segments$lab[all_segments$lab %in% c("Non-Canonical","Canonical")]<-"Can/non-can"

table(all_segments$lab)
table(all_segments$LENA_type)

all_segments$lab=factor(all_segments$lab,levels=c("Can/non-can","Laughing","Crying","Junk"))
all_segments$LENA_type=factor(all_segments$LENA_type,levels=c("Can/non-can","Laughing","Crying","Junk"))

mycf=confusionMatrix(all_segments$lab, all_segments$LENA_type, dnn = c("Lab","LENA"))
conf_tab=mycf$table
# this package uses sensitivity & specificity
#Sensitivity=recall
#Specificity=precision
mycf
```

### Precision

Precision means: If a segment was called X by LENA, what proportion of the time was it called X by lab coders?

```{r prec2}

colsums=colSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:dim(my_conf_tab)[2]) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
colSums(my_conf_tab)
prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
prop_cat$id=paste(prop_cat$Lab,prop_cat$LENA)
colnames(prop_cat)[3]<-"pr"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$LENA)
stall=merge(stall,prop_cat[c("id","pr")])
ggplot(data = stall, mapping = aes(y = Lab, x=LENA)) +
 geom_tile(aes(fill= rescale(pr)), colour = "white") +
  geom_text(aes(label = paste(round(pr),"%")), vjust = -1) +
  geom_text(aes(label = Freq), vjust = 1) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("LENA") + ylab("Lab") +
  ggtitle("Precision")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))

```

### Recall

Recall means: If a segment was called X by lab coders, what proportion of the time was it called X by LENA coders?

```{r recall2}

prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
prop_cat$id=paste(prop_cat$Lab,prop_cat$LENA)
colnames(prop_cat)[3]<-"rec"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$LENA)
stall=merge(stall,prop_cat[c("id","rec")])
ggplot(data = stall, mapping = aes(y = Lab, x=LENA)) +
 geom_tile(aes(fill= rescale(rec)), colour = "white") +
  geom_text(aes(label = paste(round(rec),"%")), vjust = -1) +
  geom_text(aes(label = Freq), vjust = 1) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("LENA") + ylab("Lab") +
  ggtitle("Recall")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))

```

### 2020-10-17 202010_jslhr/analyses.Rmd
first clean up, analyses.Rmd in 202010_jslhr knits

---
title: "Analyses for JSLHR version"
date: "2020-10-17"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
library(tidyverse)
library(dplyr)
library(ggplot2)
library(caret)
library(scales)
library(kableExtra)
library(rel)
dodiv=function(x) x/sum(x, na.rm=T)
```

## History:

- 2020-08-05 final first version
- 2020-10-10 (rehaul), 
- latest minor edits 2020-10-17

## TODO:

- pipeline is still not transparent
- there are duplicate files across raw and derived data
- there are a bunch of files with similar names
- README is old
- a note said read demo data created by AC from info in paper - should be replaced with real demo data

## Read data in

```{r}
# read datasets

demo_data=read.csv("../Derived_Data/demo-data.tsv",sep="\t")
data_ang <- read.csv("../Derived_Data/classifications_PU_zoon_final17.csv",header=T,sep=",")
data_td <- read.csv("../Derived_Data/classifications_PU_zoon_final.csv")
data_all<-rbind(data_ang, data_td)

#add filenames to demo data
demo_data_fn <- demo_data %>% 
     left_join(select(data_all, filename, ChildID), by = c("ChildID"))
demo_data_fn<-unique(demo_data_fn)

#remove the word mixed that takes up space and is unnecessary
data_all$Zoon_classif=factor(gsub("Mixed_","",as.character(data_all$Zoon_classif),fixed=T))
#relevel the factor so that it's easier to read
data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=c("Canonical","Non-Canonical",                   "Crying","Laughing","Junk",levels(data_all$Zoon_classif)[grep("_",levels(data_all$Zoon_classif))]))
# create lab column with easier to read correspondance
data_all$lab<-as.character(data_all$Major_Choice)
data_all$lab[data_all$lab=="Non-canonical syllables"]<-"Non-Canonical"
data_all$lab[data_all$lab=="Canonical syllables"]<-"Canonical"
data_all$lab[data_all$lab %in% c("Don't mark","None")]<-"Junk"
data_all$lab=factor(data_all$lab,levels=levels(data_all$Zoon_classif))
#apply same factor levels as zooniverse so that we can do symmetrical confusion matrices
```

## Correspondence between lab & zooniverse annotation at the level of segments

Here we look at to what extent zooniverse and lab annotations match at the level of individual segments. Each data point is one segment (one "vocalization").

```{r}
table(data_all$lab)
table(data_all$Zoon_classif)
mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
# this package uses sensitivity & specificity
#Sensitivity=recall
#Specificity=precision
mycf
```

### Precision

Precision means: If a segment was called X by zooniverse coders, what proportion of the time was it called X by lab coders?

```{r prec}
colsums=colSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:18) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
colSums(my_conf_tab)
prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"pr"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","pr")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(pr)), colour = "white") +
  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=2) +
  geom_text(aes(label = Freq), vjust = 1,size=1) +
  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Recall

Recall means: If a segment was called X by lab coders, what proportion of the time was it called X by zooniverse coders?

```{r rec}
rowsums=rowSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:18) my_conf_tab[,i]=my_conf_tab[,i]/rowsums[i]
rowSums(my_conf_tab)
prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"rec"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","rec")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(rec)), colour = "white") +
  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=2) +
  geom_text(aes(label = Freq), vjust = 1,size=1) +
  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Recall")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

## repeat collapsing

```{r}
#given results above, we map the mixed
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical"]<-"Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical_Crying"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Crying"]<-"Crying"
data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Crying"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing_Crying"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Canonical"]<-"Canonical"
# +
data_all$Zoon_classif[data_all$Zoon_classif=="Canonical_Crying"]<-"Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Canonical_Laughing"]<-"Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical_Crying"]<-"Non-Canonical"

data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Laughing"]<-"Crying"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Canonical_Laughing"]<-"Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Laughing_Non-Canonical"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Non-Canonical"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Non-Canonical_Laughing"]<-"Non-Canonical" 
data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing"]<-"Non-Canonical"

```

```{r}
#and reset the factors for cleanliness
data_all$Zoon_classif=factor(data_all$Zoon_classif)
data_all$lab=factor(data_all$lab)
sample_data<-cbind(data_all$lab,data_all$Zoon_classif)
gac(data = sample_data, kat = 5, weight = c("unweighted"),
    conf.level = 0.95)
mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
# this package uses sensitivity & specificity
#Sensitivity=recall
#Specificity=precision
mycf
```

```{r prec2}
pdf("../Results/precision_final.pdf",height=10,width=10)
colsums=colSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:5) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
colSums(my_conf_tab)
prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"pr"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","pr")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(pr)), colour = "white") +
  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=8) +
  geom_text(aes(label = Freq), vjust = 1,size=8) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Precision")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))
dev.off()
```

```{r recall2}
pdf("../Results/recall_final.pdf",height=10,width=10)
prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"rec"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","rec")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(rec)), colour = "white") +
  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=8) +
  geom_text(aes(label = Freq), vjust = 1,size=8) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Recall")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))
dev.off()
```

```{r}
data_as_td<-left_join(data_all,demo_data,on="ChildID")
# CM with just AS kids
data_AS<-subset(data_as_td, Diagnosis=="AngelmanSyndrome")
mycf=confusionMatrix(data_AS$lab, data_AS$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
mycf
```
```{r prec2AS}
pdf("../Results/precision_AS.pdf",height=10,width=10)
colsums=colSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:5) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
colSums(my_conf_tab)
prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"pr"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","pr")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(pr)), colour = "white") +
  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=8) +
  geom_text(aes(label = Freq), vjust = 1,size=8) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Precision")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))
dev.off()
```

```{r recall2_AS}
pdf("../Results/recall_AS.pdf",height=10,width=10)
prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"rec"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","rec")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(rec)), colour = "white") +
  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=8) +
  geom_text(aes(label = Freq), vjust = 1,size=8) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Recall")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))
dev.off()
```

```{r}
# CM with just TD kids
data_TD<-subset(data_as_td, Diagnosis=="Low-RiskControl")
mycf=confusionMatrix(data_TD$lab, data_TD$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
mycf
```

```{r prec3AS}
pdf("../Results/precision_TD.pdf",height=10,width=10)
colsums=colSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:5) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
colSums(my_conf_tab)
prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"pr"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","pr")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(pr)), colour = "white") +
  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=8) +
  geom_text(aes(label = Freq), vjust = 1,size=8) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Precision")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))
dev.off()
```

```{r recall3}
pdf("../Results/recall_TD.pdf",height=10,width=10)
prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"rec"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","rec")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(rec)), colour = "white") +
  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=8) +
  geom_text(aes(label = Freq), vjust = 1,size=8) +
  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Recall")+theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))
dev.off()
```

## Child level descriptors

Although there may be errors at the level of the segment, what we really care about is whether Zooniverse annotations give a reliable image of the child's individual development. This is what we look at in this section. In all of these graphs, red points correspond to children diagnosed with Angelman Syndrome, black for low-risk control.

```{r prepare data by child}
#get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
ztab=table(data_all$filename,data_all$Zoon_classif)
z_lr=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
z_cr=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
ltab=table(data_all$filename,data_all$lab)
l_lr=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
l_cr=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
#put all the ratios together
if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) ratios=cbind(rownames(ztab),z_lr,z_cr,l_lr,l_cr) else print("oops this code needs to be more complex because we don't have the same kids for the two ratios")
colnames(ratios)[1]<-"filename"

#ages=aggregate(data_all$Age,by=list(data_all$ChildID),mean) #this is a weird way of adding ages, since all of the ages for a given child should be the same if there is only one recording, and if there are multiple recordings, then we should not get the mean
#improvement: now we merge with a demo data tab, but note this is merged with child id, so the problem of multiple recs per child is still there

# Created demo_data with filenames. Use filenames instead of childIDs to merge ratios and demo data.

merge(ratios,demo_data_fn,by="filename")->ratios
colnames(ratios)[dim(ratios)[2]]<-"Age"
#cbinding results in text, so we numerize the ratios
for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) ratios[,thisvar]=as.numeric(as.character(ratios[,thisvar]))
summary(ratios)
```

## Correlations with age

We first look generally at two measures that have been found to relate to age:

- linguistic ratio = ("Canonical"+"Non-Canonical")/"All vocalizations" (i.e. we remove junk)
- canonical ratio = "Canonical"/("Canonical"+"Non-Canonical") (i.e. we remove junk + non-linguistic vocalizations)

TODO

- make margins smaller

- remove title repetition (age only at the bottom)

```{r corage}

prettynames=c("Linguistic Ratio (Zooniverse)","Canonical Ratio (Zooniverse)",
             "Linguistic Ratio (Lab)","Canonical Ratio (Lab)" )
names(prettynames)<-c("z_lr","z_cr","l_lr","l_cr")
mycols=c("black","red")
names(mycols)<-c("Low-RiskControl","AngelmanSyndrome")
mypch=c(4,20)
names(mypch)<-c("Low-RiskControl","AngelmanSyndrome")

jpeg("../Results/corage.jpg",width=20,height=20,units="cm",res=300)
layout(matrix(c(1:4), 2, 2, byrow = F))
for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {

  plot(ratios[,thisvar]~ratios$Age, pch=mypch[ratios$Diagnosis],xlab="Age (months)",ylab=prettynames[thisvar],
       col=mycols[ratios$Diagnosis])
  abline(lm(ratios[,thisvar]~ratios$Age,subset=c(ratios$Diagnosis=="AngelmanSyndrome")),col="black")
  myr=paste0("r=",round(cor.test(ratios[ratios$Diagnosis=="AngelmanSyndrome",thisvar],ratios$Age[ratios$Diagnosis=="AngelmanSyndrome"])$estimate,3))
  text(mean(ratios$Age[ratios$Diagnosis=="AngelmanSyndrome"]),mean(ratios[ratios$Diagnosis=="AngelmanSyndrome",thisvar]),myr,col="black")
    abline(lm(ratios[,thisvar]~ratios$Age,subset=c(ratios$Diagnosis!="AngelmanSyndrome")),col="red")
    myr=paste0("r=",round(cor.test(ratios[ratios$Diagnosis!="AngelmanSyndrome",thisvar],ratios$Age[ratios$Diagnosis!="AngelmanSyndrome"])$estimate,3))
    text(mean(ratios$Age[ratios$Diagnosis!="AngelmanSyndrome"]),mean(ratios[ratios$Diagnosis!="AngelmanSyndrome",thisvar]),myr,col="red")
}
dev.off()
```

## Correlations across zooniverse and lab

But the key thing for us: Are Zooniverse annotations describing children similar to lab annotations? The answer is clearly yes.

TODO:

- add code to print out the results paragraphs
- change figure generation to have lines fitted to each group

```{r corlab-zoo}
#Ling ratio
pdf("../Results/ling_rat_z_vs_l_final.pdf",height=5,width=5)
lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]))
  myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
  plot(ratios[,"z_lr"]~ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[ratios$Diagnosis])
  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]))
  lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
  #CR
pdf("../Results/can_rat_z_vs_l_final.pdf",height=5,width=5)
lims=range(c(ratios[,"z_cr"],ratios[,"l_cr"]))
    myr=round(cor.test(ratios[,"z_cr"],ratios[,"l_cr"])$estimate,3)
  plot(ratios[,"z_cr"]~ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[ratios$Diagnosis])
  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]),col="darkgray")
    lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()

```

```{r regmod}
lin_mod=lm(z_lr~l_lr*Diagnosis*Age,data=ratios)
summary(lin_mod)
plot(lin_mod)
```

## Code junkyard

This is redundant code

Different pdfs for diff groups

```{r corage2, eval=F}
prettynames=c("AS Linguistic Ratio (Zooniverse)","AS Canonical Ratio (Zooniverse)",
             "AS Linguistic Ratio (Lab)","AS Canonical Ratio (Lab)" )
names(prettynames)<-c("z_lr","z_cr","l_lr","l_cr")

AS_ratios<-subset(ratios, Diagnosis=="AngelmanSyndrome")
TD_ratios<-subset(ratios, Diagnosis=="Low-RiskControl")

for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
  myr=round(cor.test(AS_ratios[,thisvar],AS_ratios$Age)$estimate,3)
  plot(AS_ratios[,thisvar]~AS_ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr))
  abline(lm(AS_ratios[,thisvar]~AS_ratios$Age))
  }
```

```{r corage3, eval=F}
prettynames=c("TD Linguistic Ratio (Zooniverse)","TD Canonical Ratio (Zooniverse)",
             "TD Linguistic Ratio (Lab)","TD Canonical Ratio (Lab)" )
for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
  myr=round(cor.test(TD_ratios[,thisvar],TD_ratios$Age)$estimate,3)
  plot(TD_ratios[,thisvar]~TD_ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr))
  abline(lm(TD_ratios[,thisvar]~TD_ratios$Age))
  }
```

```{r corlab-zoo2, eval=F}
#Ling ratio
pdf("../Results/td_ling_rat_z_vs_l_final.pdf",height=5,width=5)
lims=range(c(TD_ratios[,"z_lr"],TD_ratios[,"l_lr"]))
  myr=round(cor.test(TD_ratios[,"z_lr"],TD_ratios[,"l_lr"])$estimate,3)
  plot(TD_ratios[,"z_lr"]~TD_ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[TD_ratios$Diagnosis])
  abline(lm(TD_ratios[,"z_lr"]~TD_ratios[,"l_lr"]))
  lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
  #CR
pdf("../Results/td_can_rat_z_vs_l_final.pdf",height=5,width=5)
lims=range(c(TD_ratios[,"z_cr"],TD_ratios[,"l_cr"]))
    myr=round(cor.test(TD_ratios[,"z_cr"],TD_ratios[,"l_cr"])$estimate,3)
  plot(TD_ratios[,"z_cr"]~TD_ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[TD_ratios$Diagnosis])
  abline(lm(TD_ratios[,"z_cr"]~TD_ratios[,"l_cr"]),col="darkgray")
    lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
```

```{r corlab-zoo3, eval=F}
#Ling ratio
pdf("../Results/AS_ling_rat_z_vs_l_final.pdf",height=5,width=5)
lims=range(c(AS_ratios[,"z_lr"],AS_ratios[,"l_lr"]))
  myr=round(cor.test(AS_ratios[,"z_lr"],AS_ratios[,"l_lr"])$estimate,3)
  plot(AS_ratios[,"z_lr"]~AS_ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[AS_ratios$Diagnosis])
  abline(lm(AS_ratios[,"z_lr"]~AS_ratios[,"l_lr"]))
  lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
  #CR
pdf("../Results/AS_can_rat_z_vs_l_final.pdf",height=5,width=5)
lims=range(c(AS_ratios[,"z_cr"],AS_ratios[,"l_cr"]))
    myr=round(cor.test(AS_ratios[,"z_cr"],AS_ratios[,"l_cr"])$estimate,3)
  plot(AS_ratios[,"z_cr"]~AS_ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[AS_ratios$Diagnosis])
  abline(lm(AS_ratios[,"z_cr"]~AS_ratios[,"l_cr"]),col="darkgray")
    lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
```

combined pdf

```{r , eval=F}
#COMBINED to save space
pdf("../Results/combined_final.pdf",height=5,width=5)
lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]),c(ratios[,"z_cr"],ratios[,"l_cr"]))
  #myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
  plot(ratios[,"z_lr"]~ratios[,"l_lr"],xlab="Laboratory annotations",ylab="Zooniverse annotations",
       xlim=lims,ylim=lims,
       pch=20,col=mycols[ratios$Diagnosis])
    points(ratios[,"z_cr"]~ratios[,"l_cr"], pch=2, col=mycols[ratios$Diagnosis])
  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]))
  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]),lty=3)
 # lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
```

### 2020-10-17 202010_jslhr/add-lena-ana.Rmd
second clean up, add-lena-ana.Rmd in 202010_jslhr knits

--- 2020-10-17 Scripts/add-lena-ana.Rmd
@@
 ```{r read-data,eval=F}
-read.csv("../Metadata/result_final_lisa.csv")->metadata
+read.csv("../Derived_Data/result_final_lisa.csv")->metadata
 metadata[1,]
 
-csvs=dir("../../its_to_csv/",pattern="csv")
+csvs=dir("../Raw_Data/its_to_csv/",pattern="csv")
 
@@
     #read in full csv
-    read.csv(paste0("../../its_to_csv/",csvs[csv]))->thisdat
+    read.csv(paste0("../Raw_Data/its_to_csv/",csvs[csv]))->thisdat
 
@@
 }
-write.csv(all_segments,"LENA_type_lab.csv",row.names = F)
+write.csv(all_segments,"../Derived_Data/LENA_type_lab.csv",row.names = F)
 ```
@@
 
-read.csv("LENA_type_lab.csv")->all_segments
+read.csv("../Derived_Data/LENA_type_lab.csv")->all_segments
 

### 2020-10-17 chunk_to_segment.py

@@ -1,70 +1,0 @@
-import numpy as np
-from collections import defaultdict
-import json
-
-
-result_final="/Users/chiarasemenzin/Desktop/SLT/result_final_lisa.csv"
-result_zoon="/Users/chiarasemenzin/Desktop/SLT/Metadata_batch2_lisa.csv"
-
-# ----------------------------------------------------------------------------------------
-
-def seg_to_ts(result_final):
-    seg2ts={}
-    with open(result_final) as final:
-        for e,line in enumerate(final):
-            line=line.split(",")
-            if e > 0:
-                seg2ts[line[0]]=(float(line[4]),float(line[5]),line[1])
-    return seg2ts
-
-
-def chunk_to_ts(result_zoon):
-    chunk2ts={}
-    with open(result_zoon) as zoon:
-        for e,line in enumerate(zoon):
-            line=line.split("\t")
-            if e > 0:
-                chunk2ts[line[0]]=[float(i) for i in line[4].split("_")]
-                chunk2ts[line[0]].append(line[2])
-    return chunk2ts
-
-
-def check(segments,chunks):
-    """" Check which chunks belong to which segment"""
-
-    segments_final = defaultdict(list)
-
-    for name,chunk in chunks.items(): 
-        diffs_onset=[]
-        diffs_offeset=[]
-        candidate_seg=dict([i for i in segments.items() if i[1][2]==chunk[2]])
-        for k,seg in candidate_seg.items():
-            if seg[2] == chunk[2]:
-                print("Looking at ",k,seg[2])
-                diffs_onset.append([abs(chunk[0]-seg[0]),k])
-                diffs_offeset.append([abs(chunk[1]-seg[1]),k])
-        min_on=min([i[0] for i in diffs_onset])
-        don=np.array(diffs_onset)
-        try:
-            idx=int(np.where(don ==str(min_on))[0])
-            segment=don[idx,1]
-            segments_final[segment].append(name)
-        except:
-            print("Ambigous segment. Check manually.")
-            continue
-    print(list(segments_final.items())[0:100])
-    return segments_final
-
-
-# ----------------------------------------------------------------------------------------
-
-if __name__ == "__main__":
-    s=seg_to_ts(result_final)
-    c=chunk_to_ts(result_zoon)
-    seg_=check(s,c)
-    json = json.dumps(seg_)
-    f = open("dict.json","w")
-    f.write(json)
-    f.close()
-
-

### 2020-10-17 write_seg_data.py

@@ -1,75 +1,0 @@
-import pandas as pd
-import numpy as np
-from collections import defaultdict, Counter
-import json
-
-# ----------------------------------------------------------------------------------------
-# Load data
-
-result_zoon="/Users/chiarasemenzin/Desktop/SLT/first_class_lisa.csv"
-zoon=pd.read_csv(result_zoon)
-lisa=pd.read_csv("result_final_lisa.csv") #file with metadata and PU classifications
-
-
-with open("/Users/chiarasemenzin/Desktop/SLT/dict.json", "r") as read_file:
-    print("Converting JSON encoded data into Python dictionary")
-    d = json.load(read_file)
-
-
-# ----------------------------------------------------------------------------------------
-
-def agreement_by_chunk(d,classifs):
-    """Determine majority agreement within chunks"""
-    answers = defaultdict(list)
-    for k,v in d.items():
-        for chunk in v:
-            c=Counter(list(classifs.loc[classifs["AudioData"] == chunk+".mp3" ]["Answer"]))
-            if len(c)!= 0:
-                value, count = c.most_common()[0]
-                if count >= 3 and sum([i for i in c.values()])>=3: #at least 3 classifs agree
-                    answers[k].append(value)
-                    print("Okay")
-                else:
-                    print("No majority agreement")
-    return answers
-
-
-def agreement_by_segment(answers):
-    """Determine classification within segment """
-#    TODO
-    seg_dict=[[],[]]
-    for k,v in answers.items():
-        if set(v) != {"Junk"}: # Check Junk is not the only answer
-            v=[value for value in v if value != "Junk"] #remove junk
-            if len(set(v)) == 1: # 100% agreement 
-                classif = v[0]
-            elif len(v)%2 == 0 and len(set(v)) == 2:  # 50/50 
-                if "Canonical" in v and "Non-Canonical" in v: # if Can & NonCan = Can
-                    classif="Canonical"
-                else:
-                    classif = "Mixed_" + list(set(v))[0] + "_" + list(set(v))[1] # Otherwise mixed
-            elif "Canonical" in v and "Non-Canonical" in v: #not sure about this
-                classif="Canonical"
-            else:
-                classif= "Mixed_"+"_".join([i for i in set(v)])
-                print(classif)
-        else:
-            classif="Junk" # Junk only
-        seg_dict[0].append(k)
-        seg_dict[1].append(classif)
[…]

### 2020-10-17 202010_jslhr/analyses.Rmd
paper compiles

@@
 - a note said read demo data created by AC from info in paper - should be replaced with real demo data
+- add code to print out the results paragraphs
 
@@
 
-## Correspondence between lab & zooniverse annotation at the level of segments
+## Data post-processing
 
-Here we look at to what extent zooniverse and lab annotations match at the level of individual segments. Each data point is one segment (one "vocalization").
+We collected a total of 169,628 judgments provided for 33,880 500-ms chunks, corresponding to 11,984 LENA segments.
+Nearly a fifth of chunks did not have at least 3 labels in agreement out of the 5 Zooniverse labels (N = 6,585, 19% of all chunks).
+Of the chunks without a majority agreement, 4341  (66%) contained one or two Junk judgements (out of 5), 6523 (99,9%) had at least two matching judgements (the threshold used for lab-annotated segments), and only 61 (0,01%) had 5 different judgements.
+Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the 81% of chunks that did have at least 3 labels in agreement; this represented 136,703 labels for 27,295 chunks, corresponding to 11,593 LENA segments.
+As the segments average 1.12 seconds in length, this means about 3.8 hours of audio data were annotated by 8 different annotators (3 in the laboratory, 5 on Zooniverse).
 
 ```{r}
-table(data_all$lab)
-table(data_all$Zoon_classif)
-mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
-conf_tab=mycf$table
-# this package uses sensitivity & specificity
-#Sensitivity=recall
-#Specificity=precision
-mycf
-```
+# remove non-majority labels
+#TODO
 
-### Precision
-
-Precision means: If a segment was called X by zooniverse coders, what proportion of the time was it called X by lab coders?
-
-```{r prec}
-colsums=colSums(conf_tab)
-my_conf_tab=conf_tab
-for(i in 1:18) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
-colSums(my_conf_tab)
-prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
-prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
-colnames(prop_cat)[3]<-"pr"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","pr")])
-ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(pr)), colour = "white") +
-  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=2) +
-  geom_text(aes(label = Freq), vjust = 1,size=1) +
-  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
-     theme(legend.position = "none") +
-  xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-```
-
-### Recall
-
-Recall means: If a segment was called X by lab coders, what proportion of the time was it called X by zooniverse coders?
-
-```{r rec}
-rowsums=rowSums(conf_tab)
-my_conf_tab=conf_tab
-for(i in 1:18) my_conf_tab[,i]=my_conf_tab[,i]/rowsums[i]
-rowSums(my_conf_tab)
-prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
-prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
-colnames(prop_cat)[3]<-"rec"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","rec")])
-ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(rec)), colour = "white") +
-  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=2) +
-  geom_text(aes(label = Freq), vjust = 1,size=1) +
-  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
-     theme(legend.position = "none") +
-  xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Recall")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-
-```
-
-## repeat collapsing
-
-```{r}
-#given results above, we map the mixed
+# we map the mixed
 data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical"]<-"Canonical"
@@
 
-```
-
-```{r}
 #and reset the factors for cleanliness
@@
 sample_data<-cbind(data_all$lab,data_all$Zoon_classif)
-gac(data = sample_data, kat = 5, weight = c("unweighted"),
-    conf.level = 0.95)
-mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
-conf_tab=mycf$table
-# this package uses sensitivity & specificity
-#Sensitivity=recall
-#Specificity=precision
-mycf
+
 ```
-
-```{r prec2}
-pdf("../Results/precision_final.pdf",height=10,width=10)
-colsums=colSums(conf_tab)
-my_conf_tab=conf_tab
-for(i in 1:5) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
-colSums(my_conf_tab)
-prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
-prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
-colnames(prop_cat)[3]<-"pr"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","pr")])
-ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(pr)), colour = "white") +
-  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=8) +
-  geom_text(aes(label = Freq), vjust = 1,size=8) +
-  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
-     theme(legend.position = "none") +
-  xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Precision")+theme(text = element_text(size=20),
-        axis.text.x = element_text(angle=90, hjust=1))
-dev.off()
-```
-
-```{r recall2}
-pdf("../Results/recall_final.pdf",height=10,width=10)
-prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
-prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
-colnames(prop_cat)[3]<-"rec"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","rec")])
-ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(rec)), colour = "white") +
-  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=8) +
-  geom_text(aes(label = Freq), vjust = 1,size=8) +
-  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
-     theme(legend.position = "none") +
-  xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Recall")+theme(text = element_text(size=20),
-        axis.text.x = element_text(angle=90, hjust=1))
-dev.off()
-```
-
-```{r}
-data_as_td<-left_join(data_all,demo_data,on="ChildID")
-# CM with just AS kids
-data_AS<-subset(data_as_td, Diagnosis=="AngelmanSyndrome")
-mycf=confusionMatrix(data_AS$lab, data_AS$Zoon_classif, dnn = c("Lab","Zooniverse"))
-conf_tab=mycf$table
-mycf
-```
-```{r prec2AS}
-pdf("../Results/precision_AS.pdf",height=10,width=10)
-colsums=colSums(conf_tab)
-my_conf_tab=conf_tab
-for(i in 1:5) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
-colSums(my_conf_tab)
-prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
-prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
-colnames(prop_cat)[3]<-"pr"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","pr")])
-ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(pr)), colour = "white") +
-  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=8) +
-  geom_text(aes(label = Freq), vjust = 1,size=8) +
-  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
-     theme(legend.position = "none") +
-  xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Precision")+theme(text = element_text(size=20),
-        axis.text.x = element_text(angle=90, hjust=1))
-dev.off()
-```
-
-```{r recall2_AS}
-pdf("../Results/recall_AS.pdf",height=10,width=10)
-prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
-prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
-colnames(prop_cat)[3]<-"rec"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","rec")])
-ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(rec)), colour = "white") +
-  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=8) +
-  geom_text(aes(label = Freq), vjust = 1,size=8) +
-  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
-     theme(legend.position = "none") +
-  xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Recall")+theme(text = element_text(size=20),
-        axis.text.x = element_text(angle=90, hjust=1))
-dev.off()
-```
-
-```{r}
-# CM with just TD kids
-data_TD<-subset(data_as_td, Diagnosis=="Low-RiskControl")
-mycf=confusionMatrix(data_TD$lab, data_TD$Zoon_classif, dnn = c("Lab","Zooniverse"))
-conf_tab=mycf$table
-mycf
-```
-
-```{r prec3AS}
-pdf("../Results/precision_TD.pdf",height=10,width=10)
-colsums=colSums(conf_tab)
-my_conf_tab=conf_tab
-for(i in 1:5) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
-colSums(my_conf_tab)
-prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
-prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
-colnames(prop_cat)[3]<-"pr"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","pr")])
-ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(pr)), colour = "white") +
-  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=8) +
-  geom_text(aes(label = Freq), vjust = 1,size=8) +
-  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
-     theme(legend.position = "none") +
-  xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Precision")+theme(text = element_text(size=20),
-        axis.text.x = element_text(angle=90, hjust=1))
-dev.off()
-```
-
-```{r recall3}
-pdf("../Results/recall_TD.pdf",height=10,width=10)
-prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
-prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
-colnames(prop_cat)[3]<-"rec"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","rec")])
-ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(rec)), colour = "white") +
-  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=8) +
-  geom_text(aes(label = Freq), vjust = 1,size=8) +
-  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
-     theme(legend.position = "none") +
-  xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Recall")+theme(text = element_text(size=20),
-        axis.text.x = element_text(angle=90, hjust=1))
-dev.off()
-```
-
-## Child level descriptors
-
-Although there may be errors at the level of the segment, what we really care about is whether Zooniverse annotations give a reliable image of the child's individual development.
-This is what we look at in this section.
-In all of these graphs, red points correspond to children diagnosed with Angelman Syndrome, black for low-risk control.
 
@@
 colnames(ratios)[dim(ratios)[2]]<-"Age"
+
 #cbinding results in text, so we numerize the ratios
@@
 ```
+## Results
 
-## Correlations with age
+### Descriptive analyses
+
+In this section, we provide descriptive analyses of our dataset.
+According to lab annotators, 16% of segments were canonical, 57% non-canonical, 2% laughing, and 5% crying, with the remaining 20% being categorized as "Don't code".
+Zooniverse data revealed a similar distribution: 15% canonical, 60% non-canonical, 4% laughing, 9% crying, 12% junk.
+Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic proportion = (“Canonical”+“Non-Canonical”)/“All vocalizations” (i.e., we remove junk), and ii) Canonical proportion = “Canonical”/(“Canonical”+“Non-Canonical”) (i.e.
+we remove junk + non-linguistic vocalizations).
+See Figure XX for results.
+
+Figure XX. Correlations between child-level descriptors and age as a function of metric (linguistic ratio in the top row, canonical ratio in the bottom row), annotation method
 
@@
 - make margins smaller
-
+- make the y axes use the same range, at least within metrics (i.e., linguistic ratio plots should both range from .4 to 1.0 so you can compare magnitudes of the actual data points)
 - remove title repetition (age only at the bottom)
@@
 
-## Correlations across zooniverse and lab
+Descriptive analyses on the laboratory annotations showed that the Linguistic proportion increased non-significantly with age when all 20 children were considered together r(18) = .36, CI [-.09,.69], p=0.11; with a near-zero relationship among the older children diagnosed with Angelman Syndrome r(8) = .01, CI [-.62,.63], p=0.96]; and a significant association among than the younger low-risk control children r(8) = .74, CI [0.20,0.93], p=0.014.
+The Canonical proportion exhibited non-significant developmental decreases found among older children diagnosed with Angelman Syndrome r(8) = -.55, CI [-0.87,.11], p=0.09; and marginal developmental increases among low-risk control r(8) = .61, CI -.04 0.89, p=.06.
 
-But the key thing for us: Are Zooniverse annotations describing children similar to lab annotations? The answer is clearly yes.
+Using the Zooniverse annotations, we found that the Linguistic proportion increased non-significantly with age when all 20 children were considered together r(18) = .38, CI [-.07,.70], p<0.1.
+As with laboratory annotations, the association with age was very weak for children diagnosed with Angelman Syndrome r(8) = -.10, CI [-.68,.56], p=.78, whereas younger low-risk control children showed a significant increase with age r(8) = .68, CI [0.08,0.91], p=0.03.
+Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman Syndrome r(8) = -.41, CI [-.82,.28], p=.23; and marginal developmental increases among low-risk control children, r(8) = .60, CI [-.04,0.89], p=.07.
+
+Given that these derived metrics are defined at the child level, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
+We therefore repeated our analyses but deriving our proportions for the Zooniverse data not from the segment-level composite, but rather the individual chunk-level annotations.
+We found that the Linguistic proportion increased with age when all 20 children were considered together [r(18) = .41, -.30 .72, P=.07].
+However, when each group was considered separately, Linguistic proportion was not strongly associated with showed a slight decrease with age in the Angelman Syndrome group [r(8) = -.04, CI -.6 .6, p=0.91], but increased with age an opposite pattern in the low-risk control group [r(8) = .71, CI .15 .92, p=.02].
+Similarly, the Canonical proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman Syndrome r(8) = -.35, CI -.80 .35, p=.32; and developmental increases among low-risk control r(8) = .57, CI -.08 0.88, p=.83.
+
+## Main analyses
+
+Next, we discuss the correspondence between citizen science classifications and the laboratory gold standard, at the level of individual clips.
+Results were visualized and assessed with a confusion matrix.
+We report a Precision plot (Fig.
+1) and a Recall plot (Fig.
+2): the diagonal elements show the number of correct segment-level classifications for each class while the off-diagonal elements show non-matching classifications.
+
+```{r}
+table(data_all$lab)
+table(data_all$Zoon_classif)
+gac(data = sample_data, kat = 5, weight = c("unweighted"),
+    conf.level = 0.95)
+
+mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
+conf_tab=mycf$table
+# this package uses sensitivity & specificity
+#Sensitivity=recall
+#Specificity=precision
+mycf
+```
+
+Precision means: If a segment was called X by zooniverse coders, what proportion of the time was it called X by lab coders?
+
+```{r prec}
+colsums=colSums(conf_tab)
+my_conf_tab=conf_tab
+for(i in 1:dim(conf_tab)[2]) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
+colSums(my_conf_tab)
+prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
+prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
+colnames(prop_cat)[3]<-"pr"
+data.frame(conf_tab)->stall
+stall$id=paste(stall$Lab,stall$Zooniverse)
+stall=merge(stall,prop_cat[c("id","pr")])
+ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
+ geom_tile(aes(fill= rescale(pr)), colour = "white") +
+  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=2) +
+  geom_text(aes(label = Freq), vjust = 1,size=1) +
+  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
+     theme(legend.position = "none") +
+  xlab("Zooniverse") + ylab("Lab") +
+  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+```
+
+Recall means: If a segment was called X by lab coders, what proportion of the time was it called X by zooniverse coders?
+
+```{r rec}
+rowsums=rowSums(conf_tab)
+my_conf_tab=conf_tab
+for(i in 1:dim(conf_tab)[2]) my_conf_tab[,i]=my_conf_tab[,i]/rowsums[i]
+rowSums(my_conf_tab)
+prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
+prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
+colnames(prop_cat)[3]<-"rec"
+data.frame(conf_tab)->stall
+stall$id=paste(stall$Lab,stall$Zooniverse)
+stall=merge(stall,prop_cat[c("id","rec")])
+ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
+ geom_tile(aes(fill= rescale(rec)), colour = "white") +
+  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=2) +
+  geom_text(aes(label = Freq), vjust = 1,size=1) +
+  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
+     theme(legend.position = "none") +
+  xlab("Zooniverse") + ylab("Lab") +
+  ggtitle("Recall")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+```
+
+From both visualizations, it appears that performance is moderate to good, with an overall accuracy of 65% (95% CI = .64-.66), a kappa of .4, and a Gwet's AC1 coefficient of .67 (95% CI = .66,.69). 
+A closer look at the confusion matrix reveals that errors in the classification are not distributed randomly.
+The Non-Canonical label shows the highest agreement between Zooniverse and lab experts (Recall 79%), and it is worth noticing that a majority of segments in the dataset were assigned this classificationattributed this judgment.
+Junk and Canonical show a moderately good agreement (Recall 62% and 58% respectively), with other classes having more variable assignments (49% for Crying and 39% for Junk).
+
+### Child level descriptors
+
+Although there may be errors at the level of the segment, what we really care about is whether Zooniverse annotations give a reliable image of the child's individual development.
+This is what we look at in this section.
+In all of these graphs, red points correspond to children diagnosed with Angelman Syndrome, black for low-risk control.
+
+Although the classification at the clip level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of childrens’ individual development.
+Looking at all 20 children together, we found a strong positive correlation (r (18) = 0.811, CI 0.58-0.92, p<0.001) between Linguistic proportion by child from the Zooniverse and the lab annotators' data.
+We notice a trend for Linguistic Proportions from Zooniverse data to be lower than that from laboratory data (see Fig.
+3).
+When we split by participant group, correlations remain high [Angelman Syndrome r(8) = .88, CI .55-.97, p<0.001; low-risk control r(8) =.82, CI .4 - .96, p<0.01].
+
+Similarly, a strong positive correlation (r=0.937, 0.84 0.97, p<0.001) is found in the Canonical Proportion.
+Unlike the linguistic proportion, the error for Canonical Ratio is very small and does not suggest a systematic over- or under-estimation (see Fig.
+XX).
+When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman Syndrome, who are also older than the low-risk control children [Angelman Syndrome r(8) = .86, CI 0.5-.96, p<0.01; low-risk control r(8) = .96, CI 0.83-0.99, p<0.001].
+When we use chunk-level judgments to derive the our child-level metrics, we observe very similar levels of correlation: linguistic proportion overall r (18) = .84, CI .63 .93, p<0.001 [Angelman Syndrome r(8) = .82, CI .41 .95, p<0.01; low-risk control r(8) = .84, CI .44 0.96, p<0.01]; canonical proportion overall r (18) = .96, CI .90 .98, p<0.001 [Angelman Syndrome r(8) = .85, CI .49 .85, p<0.01; low-risk control r(8) = .97, CI .89 .99, p<0.001].
 
@@
 
-- add code to print out the results paragraphs
 - change figure generation to have lines fitted to each group
@@
 ```
-
-## Code junkyard
-
-This is redundant code
-
-Different pdfs for diff groups
-
-```{r corage2, eval=F}
-prettynames=c("AS Linguistic Ratio (Zooniverse)","AS Canonical Ratio (Zooniverse)",
-             "AS Linguistic Ratio (Lab)","AS Canonical Ratio (Lab)" )
-names(prettynames)<-c("z_lr","z_cr","l_lr","l_cr")
-
-AS_ratios<-subset(ratios, Diagnosis=="AngelmanSyndrome")
-TD_ratios<-subset(ratios, Diagnosis=="Low-RiskControl")
-
-for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
-  myr=round(cor.test(AS_ratios[,thisvar],AS_ratios$Age)$estimate,3)
-  plot(AS_ratios[,thisvar]~AS_ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr))
-  abline(lm(AS_ratios[,thisvar]~AS_ratios$Age))
-  }
-```
-
-```{r corage3, eval=F}
-prettynames=c("TD Linguistic Ratio (Zooniverse)","TD Canonical Ratio (Zooniverse)",
-             "TD Linguistic Ratio (Lab)","TD Canonical Ratio (Lab)" )
-for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
-  myr=round(cor.test(TD_ratios[,thisvar],TD_ratios$Age)$estimate,3)
-  plot(TD_ratios[,thisvar]~TD_ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr))
-  abline(lm(TD_ratios[,thisvar]~TD_ratios$Age))
-  }
-```
-
-```{r corlab-zoo2, eval=F}
-#Ling ratio
-pdf("../Results/td_ling_rat_z_vs_l_final.pdf",height=5,width=5)
-lims=range(c(TD_ratios[,"z_lr"],TD_ratios[,"l_lr"]))
-  myr=round(cor.test(TD_ratios[,"z_lr"],TD_ratios[,"l_lr"])$estimate,3)
-  plot(TD_ratios[,"z_lr"]~TD_ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims,
-       col=mycols[TD_ratios$Diagnosis])
-  abline(lm(TD_ratios[,"z_lr"]~TD_ratios[,"l_lr"]))
-  lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
-  #CR
-pdf("../Results/td_can_rat_z_vs_l_final.pdf",height=5,width=5)
-lims=range(c(TD_ratios[,"z_cr"],TD_ratios[,"l_cr"]))
-    myr=round(cor.test(TD_ratios[,"z_cr"],TD_ratios[,"l_cr"])$estimate,3)
-  plot(TD_ratios[,"z_cr"]~TD_ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims,
-       col=mycols[TD_ratios$Diagnosis])
-  abline(lm(TD_ratios[,"z_cr"]~TD_ratios[,"l_cr"]),col="darkgray")
-    lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
-```
-
-```{r corlab-zoo3, eval=F}
-#Ling ratio
-pdf("../Results/AS_ling_rat_z_vs_l_final.pdf",height=5,width=5)
-lims=range(c(AS_ratios[,"z_lr"],AS_ratios[,"l_lr"]))
-  myr=round(cor.test(AS_ratios[,"z_lr"],AS_ratios[,"l_lr"])$estimate,3)
-  plot(AS_ratios[,"z_lr"]~AS_ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims,
-       col=mycols[AS_ratios$Diagnosis])
-  abline(lm(AS_ratios[,"z_lr"]~AS_ratios[,"l_lr"]))
-  lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
-  #CR
-pdf("../Results/AS_can_rat_z_vs_l_final.pdf",height=5,width=5)
-lims=range(c(AS_ratios[,"z_cr"],AS_ratios[,"l_cr"]))
-    myr=round(cor.test(AS_ratios[,"z_cr"],AS_ratios[,"l_cr"])$estimate,3)
-  plot(AS_ratios[,"z_cr"]~AS_ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims,
-       col=mycols[AS_ratios$Diagnosis])
-  abline(lm(AS_ratios[,"z_cr"]~AS_ratios[,"l_cr"]),col="darkgray")
-    lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
-```
-
-combined pdf
-
-```{r , eval=F}
-#COMBINED to save space
-pdf("../Results/combined_final.pdf",height=5,width=5)
-lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]),c(ratios[,"z_cr"],ratios[,"l_cr"]))
-  #myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
-  plot(ratios[,"z_lr"]~ratios[,"l_lr"],xlab="Laboratory annotations",ylab="Zooniverse annotations",
-       xlim=lims,ylim=lims,
-       pch=20,col=mycols[ratios$Diagnosis])
-    points(ratios[,"z_cr"]~ratios[,"l_cr"], pch=2, col=mycols[ratios$Diagnosis])
-  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]))
-  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]),lty=3)
- # lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
-```

### 2020-10-17 202010_jslhr/sup_mat.Rmd
paper compiles

--- 2020-10-17 Scripts/slt_again.Rmd
@@
 ---
-title: "SLT_paper"
-author: "AC"
-date: "2020-03-06 (substantive version), latest minor edits 2020-07-17"
+title: "Sup mat for JSLHR version"
+date: "2020-10-17"
 output:
@@
 library(scales)
+library(kableExtra)
+library(rel)
 dodiv=function(x) x/sum(x, na.rm=T)
@@
 
-- 2020-08-05 final first version
+- 2020-10-17 first version
+
+## TODO:
+
+- so much!
 
@@
 ```{r}
-#read demo data created by AC from info in paper
 # read datasets
 
-demo_data=read.csv("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Data/demo-data.tsv",sep="\t")
-data_ang <- read.csv("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Data/classifications_PU_zoon_final17.csv",header=T,sep=",")
-data_td <- read.csv("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Data/classifications_PU_zoon_final.csv")
+demo_data=read.csv("../Derived_Data/demo-data.tsv",sep="\t")
+data_ang <- read.csv("../Derived_Data/classifications_PU_zoon_final17.csv",header=T,sep=",")
+data_td <- read.csv("../Derived_Data/classifications_PU_zoon_final.csv")
 data_all<-rbind(data_ang, data_td)
@@
 ```{r prec2}
-pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/precision_final.pdf",height=10,width=10)
+pdf("../Results/precision_final.pdf",height=10,width=10)
 colsums=colSums(conf_tab)
@@
 ```{r recall2}
-pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/recall_final.pdf",height=10,width=10)
+pdf("../Results/recall_final.pdf",height=10,width=10)
 prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
@@
 ```{r prec2AS}
-pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/precision_AS.pdf",height=10,width=10)
+pdf("../Results/precision_AS.pdf",height=10,width=10)
 colsums=colSums(conf_tab)
@@
 
-```{r recall2}
-pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/recall_AS.pdf",height=10,width=10)
+```{r recall2_AS}
+pdf("../Results/recall_AS.pdf",height=10,width=10)
 prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
@@
 
-```{r prec2AS}
-pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/precision_TD.pdf",height=10,width=10)
+```{r prec3AS}
+pdf("../Results/precision_TD.pdf",height=10,width=10)
 colsums=colSums(conf_tab)
@@
 
-```{r recall2}
-pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/recall_TD.pdf",height=10,width=10)
+```{r recall3}
+pdf("../Results/recall_TD.pdf",height=10,width=10)
 prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
@@
 
+## Correlations with age
+
 We first look generally at two measures that have been found to relate to age:
@@
 
-As expected, linguistic ratio goes up with age.
+TODO
 
-Surprisingly, canonical ratio goes DOWN with age.
+- make margins smaller
+
+- remove title repetition (age only at the bottom)
 
 ```{r corage}
+
 prettynames=c("Linguistic Ratio (Zooniverse)","Canonical Ratio (Zooniverse)",
@@
 names(prettynames)<-c("z_lr","z_cr","l_lr","l_cr")
-#!!! this is working the opposite than it should! but note that to get angsynd kids to come out in red, I have to do it like this -- it doesn't make sense...
-# Now it works alright
 mycols=c("black","red")
 names(mycols)<-c("Low-RiskControl","AngelmanSyndrome")
+mypch=c(4,20)
+names(mypch)<-c("Low-RiskControl","AngelmanSyndrome")
+
+jpeg("../Results/corage.jpg",width=20,height=20,units="cm",res=300)
+layout(matrix(c(1:4), 2, 2, byrow = F))
 for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
-  myr=round(cor.test(ratios[,thisvar],ratios$Age)$estimate,3)
-  plot(ratios[,thisvar]~ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr),
+
+  plot(ratios[,thisvar]~ratios$Age, pch=mypch[ratios$Diagnosis],xlab="Age (months)",ylab=prettynames[thisvar],
        col=mycols[ratios$Diagnosis])
-  abline(lm(ratios[,thisvar]~ratios$Age))
+  abline(lm(ratios[,thisvar]~ratios$Age,subset=c(ratios$Diagnosis=="AngelmanSyndrome")),col="black")
+  myr=paste0("r=",round(cor.test(ratios[ratios$Diagnosis=="AngelmanSyndrome",thisvar],ratios$Age[ratios$Diagnosis=="AngelmanSyndrome"])$estimate,3))
+  text(mean(ratios$Age[ratios$Diagnosis=="AngelmanSyndrome"]),mean(ratios[ratios$Diagnosis=="AngelmanSyndrome",thisvar]),myr,col="black")
+    abline(lm(ratios[,thisvar]~ratios$Age,subset=c(ratios$Diagnosis!="AngelmanSyndrome")),col="red")
+    myr=paste0("r=",round(cor.test(ratios[ratios$Diagnosis!="AngelmanSyndrome",thisvar],ratios$Age[ratios$Diagnosis!="AngelmanSyndrome"])$estimate,3))
+    text(mean(ratios$Age[ratios$Diagnosis!="AngelmanSyndrome"]),mean(ratios[ratios$Diagnosis!="AngelmanSyndrome",thisvar]),myr,col="red")
 }
+dev.off()
 ```
+
+## Correlations across zooniverse and lab
 
@@
 
+TODO:
+
+- add code to print out the results paragraphs
+- change figure generation to have lines fitted to each group
+
 ```{r corlab-zoo}
 #Ling ratio
-pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/ling_rat_z_vs_l_final.pdf",height=5,width=5)
+pdf("../Results/ling_rat_z_vs_l_final.pdf",height=5,width=5)
 lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]))
@@
   #CR
-pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/can_rat_z_vs_l_final.pdf",height=5,width=5)
+pdf("../Results/can_rat_z_vs_l_final.pdf",height=5,width=5)
 lims=range(c(ratios[,"z_cr"],ratios[,"l_cr"]))
@@
 dev.off()
-#COMBINED to save space
-pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/combined_final.pdf",height=5,width=5)
-lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]),c(ratios[,"z_cr"],ratios[,"l_cr"]))
-  #myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
-  plot(ratios[,"z_lr"]~ratios[,"l_lr"],xlab="Laboratory annotations",ylab="Zooniverse annotations",
-       xlim=lims,ylim=lims,
-       pch=20,col=mycols[ratios$Diagnosis])
-    points(ratios[,"z_cr"]~ratios[,"l_cr"], pch=2, col=mycols[ratios$Diagnosis])
-  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]))
-  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]),lty=3)
- # lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
+
 ```
 
-```{r corage}
+```{r regmod}
+lin_mod=lm(z_lr~l_lr*Diagnosis*Age,data=ratios)
+summary(lin_mod)
+plot(lin_mod)
+```
+
+## Code junkyard
+
+This is redundant code
+
+Different pdfs for diff groups
+
+```{r corage2, eval=F}
 prettynames=c("AS Linguistic Ratio (Zooniverse)","AS Canonical Ratio (Zooniverse)",
@@
 
-```{r corage}
+```{r corage3, eval=F}
 prettynames=c("TD Linguistic Ratio (Zooniverse)","TD Canonical Ratio (Zooniverse)",
@@
 ```
-```{r corlab-zoo}
+
+```{r corlab-zoo2, eval=F}
 #Ling ratio
-pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/td_ling_rat_z_vs_l_final.pdf",height=5,width=5)
+pdf("../Results/td_ling_rat_z_vs_l_final.pdf",height=5,width=5)
 lims=range(c(TD_ratios[,"z_lr"],TD_ratios[,"l_lr"]))
@@
   #CR
-pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/td_can_rat_z_vs_l_final.pdf",height=5,width=5)
+pdf("../Results/td_can_rat_z_vs_l_final.pdf",height=5,width=5)
 lims=range(c(TD_ratios[,"z_cr"],TD_ratios[,"l_cr"]))
@@
 
-```{r corlab-zoo}
+```{r corlab-zoo3, eval=F}
 #Ling ratio
-pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/AS_ling_rat_z_vs_l_final.pdf",height=5,width=5)
+pdf("../Results/AS_ling_rat_z_vs_l_final.pdf",height=5,width=5)
 lims=range(c(AS_ratios[,"z_lr"],AS_ratios[,"l_lr"]))
@@
   #CR
-pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/AS_can_rat_z_vs_l_final.pdf",height=5,width=5)
+pdf("../Results/AS_can_rat_z_vs_l_final.pdf",height=5,width=5)
 lims=range(c(AS_ratios[,"z_cr"],AS_ratios[,"l_cr"]))
@@
 ```
-```{r}
-lin_mod=lm(z_lr~l_lr+Diagnosis,data=ratios)
-summary(lin_mod)
+
+combined pdf
+
+```{r , eval=F}
+#COMBINED to save space
+pdf("../Results/combined_final.pdf",height=5,width=5)
+lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]),c(ratios[,"z_cr"],ratios[,"l_cr"]))
+  #myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
+  plot(ratios[,"z_lr"]~ratios[,"l_lr"],xlab="Laboratory annotations",ylab="Zooniverse annotations",
+       xlim=lims,ylim=lims,
+       pch=20,col=mycols[ratios$Diagnosis])
+    points(ratios[,"z_cr"]~ratios[,"l_cr"], pch=2, col=mycols[ratios$Diagnosis])
+  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]))
+  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]),lty=3)
+ # lines(c(0,1),c(0,1),lty=2,col="darkgray")
+dev.off()
 ```

### 2020-10-17 202007_SLT/SLT_paper_analyses.Rmd

--- 2020-09-15 SLT_paper.Rmd
@@
 mycf
+
+gwet.ac1.table(conf_tab)
 

### 2020-10-17 202010_jslhr/analyses.Rmd
in progress, knits

@@
 library(rel)
+library(ggpubr)
 dodiv=function(x) x/sum(x, na.rm=T)
@@
 - add code to print out the results paragraphs
+- corage: - adjust margins and remove title repetition?
+- chunk-based analyses for age
 
@@
 
+table_zoon=round(table(data_all$Zoon)/sum(table(data_all$Zoon))*100)
+table_lab=round(table(data_all$lab)/sum(table(data_all$lab))*100)
 ```
@@
 ztab=table(data_all$filename,data_all$Zoon_classif)
-z_lr=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
-z_cr=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
+z_lp=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
+z_cp=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
 ltab=table(data_all$filename,data_all$lab)
-l_lr=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
-l_cr=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
-#put all the ratios together
-if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) ratios=cbind(rownames(ztab),z_lr,z_cr,l_lr,l_cr) else print("oops this code needs to be more complex because we don't have the same kids for the two ratios")
-colnames(ratios)[1]<-"filename"
+l_lp=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
+l_cp=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
+#put all the proportions together
+if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) proportions=cbind(rownames(ztab),z_lp,z_cp,l_lp,l_cp) else print("oops this code needs to be more complex because we don't have the same kids for the two proportions")
+colnames(proportions)[1]<-"filename"
 
@@
 
-# Created demo_data with filenames. Use filenames instead of childIDs to merge ratios and demo data.
+# Created demo_data with filenames. Use filenames instead of childIDs to merge proportions and demo data.
 
-merge(ratios,demo_data_fn,by="filename")->ratios
-colnames(ratios)[dim(ratios)[2]]<-"Age"
+merge(proportions,demo_data_fn,by="filename")->proportions
+colnames(proportions)[dim(proportions)[2]]<-"Age"
 
-#cbinding results in text, so we numerize the ratios
-for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) ratios[,thisvar]=as.numeric(as.character(ratios[,thisvar]))
-summary(ratios)
+#cbinding results in text, so we numerize the proportions
+for(thisvar in c("z_lp","z_cp","l_lp","l_cp")) proportions[,thisvar]=as.numeric(as.character(proportions[,thisvar]))
+summary(proportions)
 ```
@@
 In this section, we provide descriptive analyses of our dataset.
-According to lab annotators, 16% of segments were canonical, 57% non-canonical, 2% laughing, and 5% crying, with the remaining 20% being categorized as "Don't code".
-Zooniverse data revealed a similar distribution: 15% canonical, 60% non-canonical, 4% laughing, 9% crying, 12% junk.
+According to lab annotators, `r table_lab["Canonical"]`% of segments were canonical, `r table_lab["Non-Canonical"]`% non-canonical, `r table_lab["Laughing"]`% laughing, and `r table_lab["Crying"]`% crying, with the remaining `r table_lab["Junk"]`% being categorized as "Don't code".
+Zooniverse data revealed a similar distribution:  `r table_zoon["Canonical"]`% canonical, `r table_zoon["Non-Canonical"]`% non-canonical, `r table_zoon["Laughing"]`% laughing, `r table_zoon["Crying"]`% crying, `r table_zoon["Junk"]`% junk.
 Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic proportion = (“Canonical”+“Non-Canonical”)/“All vocalizations” (i.e., we remove junk), and ii) Canonical proportion = “Canonical”/(“Canonical”+“Non-Canonical”) (i.e.
@@
 
-Figure XX. Correlations between child-level descriptors and age as a function of metric (linguistic ratio in the top row, canonical ratio in the bottom row), annotation method
-
-We first look generally at two measures that have been found to relate to age:
-
-- linguistic ratio = ("Canonical"+"Non-Canonical")/"All vocalizations" (i.e. we remove junk)
-- canonical ratio = "Canonical"/("Canonical"+"Non-Canonical") (i.e. we remove junk + non-linguistic vocalizations)
-
-TODO
-
-- make margins smaller
-- make the y axes use the same range, at least within metrics (i.e., linguistic ratio plots should both range from .4 to 1.0 so you can compare magnitudes of the actual data points)
-- remove title repetition (age only at the bottom)
+Figure XX.
+Correlations between child-level descriptors and age as a function of metric (linguistic ratio in the top row, canonical ratio in the bottom row), annotation method, and child group (Red as TD, and black as AS)
 
@@
              "Linguistic Ratio (Lab)","Canonical Ratio (Lab)" )
-names(prettynames)<-c("z_lr","z_cr","l_lr","l_cr")
+names(prettynames)<-c("z_lp","z_cp","l_lp","l_cp")
 mycols=c("black","red")
@@
 
-jpeg("../Results/corage.jpg",width=20,height=20,units="cm",res=300)
+lp_range=range(proportions[,c("z_lp","l_lp")])
+cp_range=range(proportions[,c("z_cp","l_cp")])
+
+#jpeg("../Results/corage.jpg",width=20,height=20,units="cm",res=300)
 layout(matrix(c(1:4), 2, 2, byrow = F))
-for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
+par(mar = c(4, 4, 1, 1))
 
-  plot(ratios[,thisvar]~ratios$Age, pch=mypch[ratios$Diagnosis],xlab="Age (months)",ylab=prettynames[thisvar],
-       col=mycols[ratios$Diagnosis])
-  abline(lm(ratios[,thisvar]~ratios$Age,subset=c(ratios$Diagnosis=="AngelmanSyndrome")),col="black")
-  myr=paste0("r=",round(cor.test(ratios[ratios$Diagnosis=="AngelmanSyndrome",thisvar],ratios$Age[ratios$Diagnosis=="AngelmanSyndrome"])$estimate,3))
-  text(mean(ratios$Age[ratios$Diagnosis=="AngelmanSyndrome"]),mean(ratios[ratios$Diagnosis=="AngelmanSyndrome",thisvar]),myr,col="black")
-    abline(lm(ratios[,thisvar]~ratios$Age,subset=c(ratios$Diagnosis!="AngelmanSyndrome")),col="red")
-    myr=paste0("r=",round(cor.test(ratios[ratios$Diagnosis!="AngelmanSyndrome",thisvar],ratios$Age[ratios$Diagnosis!="AngelmanSyndrome"])$estimate,3))
-    text(mean(ratios$Age[ratios$Diagnosis!="AngelmanSyndrome"]),mean(ratios[ratios$Diagnosis!="AngelmanSyndrome",thisvar]),myr,col="red")
+#create matrix to hold r results
+myr=matrix(NA,nrow=4,ncol=2)
+rownames(myr)<-c("z_lp","z_cp","l_lp","l_cp")
+colnames(myr)<-c("AS","LR")
+mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
+
+for(thisvar in c("z_lp","z_cp","l_lp","l_cp")) {
+
+  if(thisvar %in% c("z_lp","l_lp")) myrange=lp_range else myrange=cp_range
+
+  plot(proportions[,thisvar]~proportions$Age, pch=mypch[proportions$Diagnosis],xlab="Age (months)",ylab=prettynames[thisvar],
+       col=mycols[proportions$Diagnosis],
+         ylim=myrange)
+  abline(lm(proportions[,thisvar]~proportions$Age,subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black")
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"])
+    myr[thisvar,"AS"]=round(thiscor$estimate,3)
+    mylb[thisvar,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[thisvar,"AS"]=round(thiscor$conf.int[2],3)
+    myn[thisvar,"AS"]=round(thiscor$parameter,3)
+    myp[thisvar,"AS"]=round(thiscor$p.value,3)
+
+    #add r to figure
+  text(mean(proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"]),mean(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar]),paste0("r=",myr[thisvar,"AS"]),col="black")
+
+  #repeat for LR  
+    abline(lm(proportions[,thisvar]~proportions$Age,subset=c(proportions$Diagnosis!="AngelmanSyndrome")),col="red")
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"])
+    myr[thisvar,"LR"]=round(thiscor$estimate,3)
+    mylb[thisvar,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[thisvar,"LR"]=round(thiscor$conf.int[2],3)
+    myn[thisvar,"LR"]=round(thiscor$parameter,3)
+    myp[thisvar,"LR"]=round(thiscor$p.value,3)
+
+    #add r to figure
+
+    text(mean(proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"]),mean(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar]),paste0("r=",myr[thisvar,"LR"]),col="red")
 }
-dev.off()
+#dev.off()
 ```
 
-Descriptive analyses on the laboratory annotations showed that the Linguistic proportion increased non-significantly with age when all 20 children were considered together r(18) = .36, CI [-.09,.69], p=0.11; with a near-zero relationship among the older children diagnosed with Angelman Syndrome r(8) = .01, CI [-.62,.63], p=0.96]; and a significant association among than the younger low-risk control children r(8) = .74, CI [0.20,0.93], p=0.014.
-The Canonical proportion exhibited non-significant developmental decreases found among older children diagnosed with Angelman Syndrome r(8) = -.55, CI [-0.87,.11], p=0.09; and marginal developmental increases among low-risk control r(8) = .61, CI -.04 0.89, p=.06.
+Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic proportion and age differed across the groups.
+There was a near-zero relationship among the older children diagnosed with Angelman Syndrome r(`r myn["l_lp","AS"]`) = `r myr["l_lp","AS"]`, CI [`r mylb["l_lp","AS"]`,`r myhb["l_lp","AS"]`], p=`r myp["l_lp","AS"]`]; and a significant association among younger low-risk control children r(`r myn["l_lp","LR"]`) = `r myr["l_lp","LR"]`, CI [`r mylb["l_lp","LR"]`,`r myhb["l_lp","LR"]`], p=`r myp["l_lp","LR"]`].
+The Canonical proportion exhibited non-significant developmental decreases among older children diagnosed with Angelman Syndrome r(`r myn["l_cp","AS"]`) = `r myr["l_cp","AS"]`, CI [`r mylb["l_cp","AS"]`,`r myhb["l_cp","AS"]`], p=`r myp["l_cp","AS"]`]; and marginal developmental increases among low-risk control r(`r myn["l_cp","LR"]`) = `r myr["l_cp","LR"]`, CI [`r mylb["l_cp","LR"]`,`r myhb["l_cp","LR"]`], p=`r myp["l_cp","LR"]`].
 
-Using the Zooniverse annotations, we found that the Linguistic proportion increased non-significantly with age when all 20 children were considered together r(18) = .38, CI [-.07,.70], p<0.1.
-As with laboratory annotations, the association with age was very weak for children diagnosed with Angelman Syndrome r(8) = -.10, CI [-.68,.56], p=.78, whereas younger low-risk control children showed a significant increase with age r(8) = .68, CI [0.08,0.91], p=0.03.
-Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman Syndrome r(8) = -.41, CI [-.82,.28], p=.23; and marginal developmental increases among low-risk control children, r(8) = .60, CI [-.04,0.89], p=.07.
+Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman Syndrome r(`r myn["z_lp","AS"]`) = `r myr["z_lp","AS"]`, CI [`r mylb["z_lp","AS"]`,`r myhb["z_lp","AS"]`], p=`r myp["z_lp","AS"]`]; whereas younger low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r myr["z_lp","LR"]`, CI [`r mylb["z_lp","LR"]`,`r myhb["z_lp","LR"]`], p=`r myp["z_lp","LR"]`].
+Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman Syndrome r(`r myn["z_cp","AS"]`) = `r myr["z_cp","AS"]`, CI [`r mylb["z_cp","AS"]`,`r myhb["z_cp","AS"]`], p=`r myp["z_cp","AS"]`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r myr["z_cp","LR"]`, CI [`r mylb["z_cp","LR"]`,`r myhb["z_cp","LR"]`], p=`r myp["z_cp","LR"]`].
 
@@
 We therefore repeated our analyses but deriving our proportions for the Zooniverse data not from the segment-level composite, but rather the individual chunk-level annotations.
-We found that the Linguistic proportion increased with age when all 20 children were considered together [r(18) = .41, -.30 .72, P=.07].
-However, when each group was considered separately, Linguistic proportion was not strongly associated with showed a slight decrease with age in the Angelman Syndrome group [r(8) = -.04, CI -.6 .6, p=0.91], but increased with age an opposite pattern in the low-risk control group [r(8) = .71, CI .15 .92, p=.02].
+We found that the Linguistic proportion was not strongly associated with age in the Angelman Syndrome group [r(8) = -.04, CI -.6 .6, p=0.91], but increased with age  in the low-risk control group [r(8) = .71, CI .15 .92, p=.02].
 Similarly, the Canonical proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman Syndrome r(8) = -.35, CI -.80 .35, p=.32; and developmental increases among low-risk control r(8) = .57, CI -.08 0.88, p=.83.
@@
 ```{r}
-table(data_all$lab)
-table(data_all$Zoon_classif)
-gac(data = sample_data, kat = 5, weight = c("unweighted"),
+mygac=gac(data = sample_data, kat = 5, weight = c("unweighted"),
     conf.level = 0.95)
@@
 #Specificity=precision
-mycf
+#mycf
 ```
 
-Precision means: If a segment was called X by zooniverse coders, what proportion of the time was it called X by lab coders?
+```{r figure-prec-rec}
 
-```{r prec}
 colsums=colSums(conf_tab)
@@
 stall=merge(stall,prop_cat[c("id","pr")])
-ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
+prec_plot = ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
  geom_tile(aes(fill= rescale(pr)), colour = "white") +
@@
   ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-```
 
-Recall means: If a segment was called X by lab coders, what proportion of the time was it called X by zooniverse coders?
-
-```{r rec}
+# repeat for recall
 rowsums=rowSums(conf_tab)
@@
 stall=merge(stall,prop_cat[c("id","rec")])
-ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
+rec_plot = ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
  geom_tile(aes(fill= rescale(rec)), colour = "white") +
@@
 
+ggarrange(prec_plot, rec_plot, 
+          labels = c("A", "B"),
+          ncol = 2, nrow = 1)
+
 ```
 
-From both visualizations, it appears that performance is moderate to good, with an overall accuracy of 65% (95% CI = .64-.66), a kappa of .4, and a Gwet's AC1 coefficient of .67 (95% CI = .66,.69). 
-A closer look at the confusion matrix reveals that errors in the classification are not distributed randomly.
-The Non-Canonical label shows the highest agreement between Zooniverse and lab experts (Recall 79%), and it is worth noticing that a majority of segments in the dataset were assigned this classificationattributed this judgment.
-Junk and Canonical show a moderately good agreement (Recall 62% and 58% respectively), with other classes having more variable assignments (49% for Crying and 39% for Junk).
+From both visualizations, it appears that performance is moderate to good, with an overall accuracy of `r round(mycf$overall["Accuracy"]*100)`%, CI = [`r round(mycf$overall["AccuracyLower"]*100)`,`r round(mycf$overall["AccuracyUpper"]*100)`], a kappa of `r round(mycf$overall["Kappa"],3)`, and a Gwet's AC1 coefficient of `r round(mygac$est,3)`, CI = [`r round(mygac$lb,3)`,`r round(mygac$ub,3)`].
 
@@
 
-Although there may be errors at the level of the segment, what we really care about is whether Zooniverse annotations give a reliable image of the child's individual development.
-This is what we look at in this section.
-In all of these graphs, red points correspond to children diagnosed with Angelman Syndrome, black for low-risk control.
+TODO
+
+- fix size
+- point should be dependent on child group
+
+```{r corlab-zoo}
+
+layout(matrix(c(1:2), 1, 2, byrow = F))
+par(mar = c(4, 4, 1, 1))
+
+#create matrix to hold r results
+myr=matrix(NA,nrow=2,ncol=3)
+rownames(myr)<-c("lp","cp")
+colnames(myr)<-c("AS","LR","all")
+mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
+
+lims=range(c(proportions[,"z_lp"],proportions[,"l_lp"]))
+
+  plot(proportions[,"z_lp"]~proportions[,"l_lp"], pch=20,xlab=prettynames["l_lp"],ylab=prettynames["z_lp"],
+       xlim=lims,ylim=lims,
+       col=mycols[proportions$Diagnosis])
+  lines(c(0,1),c(0,1),lty=2,col="darkgray")
+
+  abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black")
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_lp"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
+    myr[1,"AS"]=round(thiscor$estimate,3)
+    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
+    myn[1,"AS"]=round(thiscor$parameter,3)
+    myp[1,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR  
+    abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$Diagnosis!="AngelmanSyndrome")),col="red")
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"],proportions[proportions$Diagnosis!="AngelmanSyndrome","z_lp"])
+    myr[1,"LR"]=round(thiscor$estimate,3)
+    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
+    myn[1,"LR"]=round(thiscor$parameter,3)
+    myp[1,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(proportions$z_lp,proportions$l_lp)
+    myr[1,"all"]=round(thiscor$estimate,3)
+    mylb[1,"all"]=round(thiscor$conf.int[1],3)
+    myhb[1,"all"]=round(thiscor$conf.int[2],3)
+    myn[1,"all"]=round(thiscor$parameter,3)
+    myp[1,"all"]=round(thiscor$p.value,3)
+
+    # repeat whole for cp
+
+    lims=range(c(proportions[,"z_cp"],proportions[,"l_cp"]))
+
+  plot(proportions[,"z_cp"]~proportions[,"l_cp"], pch=20,xlab=prettynames["l_cp"],ylab=prettynames["z_cp"],
+       xlim=lims,ylim=lims,
+       col=mycols[proportions$Diagnosis])
+  lines(c(0,1),c(0,1),lty=2,col="darkgray")
+
+  abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black")
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_cp"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
+    myr[2,"AS"]=round(thiscor$estimate,3)
+    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
+    myn[2,"AS"]=round(thiscor$parameter,3)
+    myp[2,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR  
+    abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$Diagnosis!="AngelmanSyndrome")),col="red")
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"],proportions[proportions$Diagnosis!="AngelmanSyndrome","z_cp"])
+    myr[2,"LR"]=round(thiscor$estimate,3)
+    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
+    myn[2,"LR"]=round(thiscor$parameter,3)
+    myp[2,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(proportions$z_cp,proportions$l_cp)
+    myr[2,"all"]=round(thiscor$estimate,3)
+    mylb[2,"all"]=round(thiscor$conf.int[1],3)
+    myhb[2,"all"]=round(thiscor$conf.int[2],3)
+    myn[2,"all"]=round(thiscor$parameter,3)
+    myp[2,"all"]=round(thiscor$p.value,3)
+```
 
@@
 When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman Syndrome, who are also older than the low-risk control children [Angelman Syndrome r(8) = .86, CI 0.5-.96, p<0.01; low-risk control r(8) = .96, CI 0.83-0.99, p<0.001].
+
 When we use chunk-level judgments to derive the our child-level metrics, we observe very similar levels of correlation: linguistic proportion overall r (18) = .84, CI .63 .93, p<0.001 [Angelman Syndrome r(8) = .82, CI .41 .95, p<0.01; low-risk control r(8) = .84, CI .44 0.96, p<0.01]; canonical proportion overall r (18) = .96, CI .90 .98, p<0.001 [Angelman Syndrome r(8) = .85, CI .49 .85, p<0.01; low-risk control r(8) = .97, CI .89 .99, p<0.001].
 
-TODO:
-
-- change figure generation to have lines fitted to each group
-
-```{r corlab-zoo}
-#Ling ratio
-pdf("../Results/ling_rat_z_vs_l_final.pdf",height=5,width=5)
-lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]))
-  myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
-  plot(ratios[,"z_lr"]~ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims,
-       col=mycols[ratios$Diagnosis])
-  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]))
-  lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
-  #CR
-pdf("../Results/can_rat_z_vs_l_final.pdf",height=5,width=5)
-lims=range(c(ratios[,"z_cr"],ratios[,"l_cr"]))
-    myr=round(cor.test(ratios[,"z_cr"],ratios[,"l_cr"])$estimate,3)
-  plot(ratios[,"z_cr"]~ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims,
-       col=mycols[ratios$Diagnosis])
-  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]),col="darkgray")
-    lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
-
-```
-
 ```{r regmod}
-lin_mod=lm(z_lr~l_lr*Diagnosis*Age,data=ratios)
+lin_mod=lm(z_lp~l_lp*Diagnosis*Age,data=proportions)
 summary(lin_mod)
@@
 ```
+
+## Additional analyses
+
+We additionally explored under what conditions Zooniverse judgments more closely aligned with laboratory judgments:
+
+- take all 5 Zoo judgments or 3 or 1 (may need to run the latter two a couple of times to check for potential variance -- if variance is small, we'll just say so, if it's large then we'll have to write a loop and report CIs)
+- look at whether kids who have a high "junk" proportion according to Zooniverse 
+     - lead to lower Zoo-PU correlations
+     - have lower levels of concordance across Zoo coders
+- look at what happens to TD Zoo-PU correlation when you undersample 100 vocs to 80 and 60 and to the same number of non-junk vocs as in AS

### 2020-10-18 202010_jslhr/analyses.Rmd
all but chunk ana done

@@
 library(ggpubr)
-dodiv=function(x) x/sum(x, na.rm=T)
+library(lme4)
+
 ```
@@
 
-demo_data=read.csv("../Derived_Data/demo-data.tsv",sep="\t")
 data_ang <- read.csv("../Derived_Data/classifications_PU_zoon_final17.csv",header=T,sep=",")
@@
 data_all<-rbind(data_ang, data_td)
-
-#add filenames to demo data
-demo_data_fn <- demo_data %>% 
-     left_join(select(data_all, filename, ChildID), by = c("ChildID"))
-demo_data_fn<-unique(demo_data_fn)
 
@@
 #apply same factor levels as zooniverse so that we can do symmetrical confusion matrices
+
+#add binomials for linguistic proportion
+data_all$lab_ling=ifelse(data_all$lab %in% c("Canonical","Non-Canonical"),1,0)
+data_all$zoo_ling=ifelse(data_all$Zoon_classif %in% c("Canonical","Non-Canonical"),1,0)
+data_all$lab_ling[data_all$lab=="Junk"]<-NA
+data_all$zoo_ling[data_all$lab=="Junk"]<-NA
+
+#add binomials for canonical proportion
+data_all$lab_can=data_all$zoo_can=NA
+data_all$lab_can[data_all$lab=="Canonical"]<-1
+data_all$lab_can[data_all$lab=="Non-Canonical"]<-0
+data_all$zoo_can[data_all$Zoon_classif=="Canonical"]<-1
+data_all$zoo_can[data_all$Zoon_classif=="Non-Canonical"]<-0
+
+demo_data=read.csv("../Derived_Data/demo-data.tsv",sep="\t")
+#add filenames to demo data, to be used later
+demo_data_fn <- demo_data %>% 
+     left_join(select(data_all, filename, ChildID), by = c("ChildID"))
+demo_data_fn<-unique(demo_data_fn)
+
 ```
@@
 ```{r prepare data by child}
+
 #get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
@@
 z_cp=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
+
 ltab=table(data_all$filename,data_all$lab)
@@
 l_cp=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
+
+#and also get junk from zooniverse
+z_junk=ztab[,c("Junk")]/rowSums(ztab)
+
 #put all the proportions together
-if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) proportions=cbind(rownames(ztab),z_lp,z_cp,l_lp,l_cp) else print("oops this code needs to be more complex because we don't have the same kids for the two proportions")
+if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) proportions=cbind(rownames(ztab),z_lp,z_cp,l_lp,l_cp,z_junk) else print("oops this code needs to be more complex because we don't have the same kids for the two proportions")
 colnames(proportions)[1]<-"filename"
@@
 #cbinding results in text, so we numerize the proportions
-for(thisvar in c("z_lp","z_cp","l_lp","l_cp")) proportions[,thisvar]=as.numeric(as.character(proportions[,thisvar]))
+for(thisvar in c("z_lp","z_cp","l_lp","l_cp","z_junk")) proportions[,thisvar]=as.numeric(as.character(proportions[,thisvar]))
+
+#add median junk
+proportions$median_junk<-ifelse(proportions$z_junk<median(proportions$z_junk),"lower_junk","higher_junk")
+
 summary(proportions)
+
+merge(data_all,proportions,all=T)->data_all
 ```
@@
 
-prettynames=c("Linguistic Ratio (Zooniverse)","Canonical Ratio (Zooniverse)",
-             "Linguistic Ratio (Lab)","Canonical Ratio (Lab)" )
+prettynames=c("Linguistic Proportion (Zooniverse)","Canonical Proportion (Zooniverse)",
+             "Linguistic Proportion (Lab)","Canonical Proportion (Lab)" )
 names(prettynames)<-c("z_lp","z_cp","l_lp","l_cp")
@@
 
-TODO
-
-- fix size
-- point should be dependent on child group
-
 ```{r corlab-zoo}
@@
 
-  plot(proportions[,"z_lp"]~proportions[,"l_lp"], pch=20,xlab=prettynames["l_lp"],ylab=prettynames["z_lp"],
+  plot(proportions[,"z_lp"]~proportions[,"l_lp"], xlab=prettynames["l_lp"],ylab=prettynames["z_lp"],
        xlim=lims,ylim=lims,
-       col=mycols[proportions$Diagnosis])
+       col=mycols[proportions$Diagnosis], pch=mypch[proportions$Diagnosis])
   lines(c(0,1),c(0,1),lty=2,col="darkgray")
@@
 
-  plot(proportions[,"z_cp"]~proportions[,"l_cp"], pch=20,xlab=prettynames["l_cp"],ylab=prettynames["z_cp"],
+  plot(proportions[,"z_cp"]~proportions[,"l_cp"], xlab=prettynames["l_cp"],ylab=prettynames["z_cp"],
        xlim=lims,ylim=lims,
-       col=mycols[proportions$Diagnosis])
+       col=mycols[proportions$Diagnosis], pch=mypch[proportions$Diagnosis])
   lines(c(0,1),c(0,1),lty=2,col="darkgray")
@@
 Although the classification at the clip level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of childrens’ individual development.
-Looking at all 20 children together, we found a strong positive correlation (r (18) = 0.811, CI 0.58-0.92, p<0.001) between Linguistic proportion by child from the Zooniverse and the lab annotators' data.
-We notice a trend for Linguistic Proportions from Zooniverse data to be lower than that from laboratory data (see Fig.
-3).
-When we split by participant group, correlations remain high [Angelman Syndrome r(8) = .88, CI .55-.97, p<0.001; low-risk control r(8) =.82, CI .4 - .96, p<0.01].
+Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman Syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
 
-Similarly, a strong positive correlation (r=0.937, 0.84 0.97, p<0.001) is found in the Canonical Proportion.
-Unlike the linguistic proportion, the error for Canonical Ratio is very small and does not suggest a systematic over- or under-estimation (see Fig.
-XX).
-When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman Syndrome, who are also older than the low-risk control children [Angelman Syndrome r(8) = .86, CI 0.5-.96, p<0.01; low-risk control r(8) = .96, CI 0.83-0.99, p<0.001].
+Similarly, a strong positive correlation is found in the Canonical Proportion r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`].
+When we split by participant group, correlations remain high although we do note they are somewhat smaller for the older children with Angelman Syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
 
@@
 
-```{r regmod}
-lin_mod=lm(z_lp~l_lp*Diagnosis*Age,data=proportions)
-summary(lin_mod)
-plot(lin_mod)
+```{r glmer-agediag}
+# one outlier in ling prop is removed
+
+ling_all=glmer(lab_ling~zoo_ling*Diagnosis*Age + (1|ChildID),data=data_all,family="binomial",subset=c(ChildID!="3681_1"))
+#plot(lin_mod)
+# doesn't converge
+
+ling_AS=glmer(lab_ling~zoo_ling*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis=="AngelmanSyndrome"))
+#only effect of zoo
+
+ling_LR=glmer(lab_ling~zoo_ling*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis!="AngelmanSyndrome" & ChildID!="3681_1"))
+#both main and interaction are sig
+
+can_all=glmer(lab_can~zoo_can*Diagnosis*Age + (1|ChildID),data=data_all,family="binomial")
+#doesn't converge either
+
+can_AS=glmer(lab_can~zoo_can*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis=="AngelmanSyndrome"))
+#only effect of zoo
+
+can_LR=glmer(lab_can~zoo_can*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis!="AngelmanSyndrome"))
+#both main and interaction are sig
+
 ```
@@
 
-We additionally explored under what conditions Zooniverse judgments more closely aligned with laboratory judgments:
+We additionally explored under what conditions Zooniverse judgments more closely aligned with laboratory judgments.
+In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as "Junk" (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora [@cychosz].
+A high proportion of "Junk" may indicate that automated segmentation was errorful for those children, and may be a sign that the rest of the data could be compromised as well.
+
+```{r prelim-junk}
+junkXdiag=t.test(proportions$z_junk ~ proportions$Diagnosis)
+
+```
+
+```{r cor-junk}
+
+mycols=c("blue","darkgreen")
+mypch=c(10,18)
+names(mycols)=names(mypch)=c("lower_junk","higher_junk")
+
+layout(matrix(c(1:2), 1, 2, byrow = F))
+par(mar = c(4, 4, 1, 1))
+
+#create matrix to hold r results
+myr=matrix(NA,nrow=2,ncol=2)
+rownames(myr)<-c("lp","cp")
+colnames(myr)<-c("lower_junk","higher_junk")
+mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
+
+lims=range(c(proportions[,"z_lp"],proportions[,"l_lp"]))
+
+  plot(proportions[,"z_lp"]~proportions[,"l_lp"], xlab=prettynames["l_lp"],ylab=prettynames["z_lp"],
+       xlim=lims,ylim=lims,
+       col=mycols[proportions$median_junk], pch=mypch[proportions$median_junk])
+  lines(c(0,1),c(0,1),lty=2,col="darkgray")
+
+  abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$median_junk=="lower_junk")),col="blue")
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$median_junk=="lower_junk","z_lp"],proportions[proportions$median_junk=="lower_junk","l_lp"])
+    myr[1,"lower_junk"]=round(thiscor$estimate,3)
+    mylb[1,"lower_junk"]=round(thiscor$conf.int[1],3)
+    myhb[1,"lower_junk"]=round(thiscor$conf.int[2],3)
+    myn[1,"lower_junk"]=round(thiscor$parameter,3)
+    myp[1,"lower_junk"]=round(thiscor$p.value,3)
+
+  #repeat for LR  
+    abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$median_junk!="lower_junk")),col="darkgreen")
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$median_junk!="lower_junk","l_lp"],proportions[proportions$median_junk!="lower_junk","z_lp"])
+    myr[1,"higher_junk"]=round(thiscor$estimate,3)
+    mylb[1,"higher_junk"]=round(thiscor$conf.int[1],3)
+    myhb[1,"higher_junk"]=round(thiscor$conf.int[2],3)
+    myn[1,"higher_junk"]=round(thiscor$parameter,3)
+    myp[1,"higher_junk"]=round(thiscor$p.value,3)
+
+    # repeat whole for cp
+
+    lims=range(c(proportions[,"z_cp"],proportions[,"l_cp"]))
+
+  plot(proportions[,"z_cp"]~proportions[,"l_cp"], xlab=prettynames["l_cp"],ylab=prettynames["z_cp"],
+       xlim=lims,ylim=lims,
+       col=mycols[proportions$median_junk], pch=mypch[proportions$median_junk])
+  lines(c(0,1),c(0,1),lty=2,col="darkgray")
+
+  abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$median_junk=="lower_junk")),col="blue")
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$median_junk=="lower_junk","z_cp"],proportions[proportions$median_junk=="lower_junk","l_cp"])
+    myr[2,"lower_junk"]=round(thiscor$estimate,3)
+    mylb[2,"lower_junk"]=round(thiscor$conf.int[1],3)
+    myhb[2,"lower_junk"]=round(thiscor$conf.int[2],3)
+    myn[2,"lower_junk"]=round(thiscor$parameter,3)
+    myp[2,"lower_junk"]=round(thiscor$p.value,3)
+
+  #repeat for higher_junk  
+    abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$median_junk!="lower_junk")),col="darkgreen")
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$median_junk!="lower_junk","l_cp"],proportions[proportions$median_junk!="lower_junk","z_cp"])
+    myr[2,"higher_junk"]=round(thiscor$estimate,3)
+    mylb[2,"higher_junk"]=round(thiscor$conf.int[1],3)
+    myhb[2,"higher_junk"]=round(thiscor$conf.int[2],3)
+    myn[2,"higher_junk"]=round(thiscor$parameter,3)
+    myp[2,"higher_junk"]=round(thiscor$p.value,3)
+
+```
+
+We investigated this hypothesis by calculating the proportion of their data labeled as "Junk" for each individual child.
+There was no significant difference in the proportions of their data labeled as "Junk" for older children with Angelman Syndrome (M = `r round(junkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the younger low risk children (M = `r round(junkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r junkXdiag$parameter`) = `r round(junkXdiag$statistic,3)`, p = `r round(junkXdiag$p.value,3)`.
+We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk".
+Results were similar across these two post-hoc subgroups.
+For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower junk group was r(`r myn["lp","lower_junk"]`) = `r myr["lp","lower_junk"]`, CI [`r mylb["lp","lower_junk"]`,`r myhb["lp","lower_junk"]`], p=`r myp["lp","lower_junk"]`]; and for the higher junk group it was r(`r myn["lp","higher_junk"]`) = `r myr["lp","higher_junk"]`, CI [`r mylb["lp","higher_junk"]`,`r myhb["lp","higher_junk"]`], p=`r myp["lp","higher_junk"]`.
+For Canonical Proportion, the correlation across lab and Zooniverse data for the lower junk group was r(`r myn["cp","lower_junk"]`) = `r myr["cp","lower_junk"]`, CI [`r mylb["cp","lower_junk"]`,`r myhb["cp","lower_junk"]`], p=`r myp["cp","lower_junk"]`]; and for the higher junk group it was r(`r myn["cp","higher_junk"]`) = `r myr["cp","higher_junk"]`, CI [`r mylb["cp","higher_junk"]`,`r myhb["cp","higher_junk"]`], p=`r myp["cp","higher_junk"]`.
+Thus, it does not seem that a higher proportion of "Junk" judgments is an index of low quality data.
+
+<!-- - look at whether kids who have a high "junk" proportion according to Zooniverse  -->
+<!--      - have lower levels of concordance across Zoo coders -->
 
 - take all 5 Zoo judgments or 3 or 1 (may need to run the latter two a couple of times to check for potential variance -- if variance is small, we'll just say so, if it's large then we'll have to write a loop and report CIs)
-- look at whether kids who have a high "junk" proportion according to Zooniverse 
-     - lead to lower Zoo-PU correlations
-     - have lower levels of concordance across Zoo coders
 - look at what happens to TD Zoo-PU correlation when you undersample 100 vocs to 80 and 60 and to the same number of non-junk vocs as in AS

### 2020-10-27 202010_jslhr/analyses.Rmd
coded chunk-based analyses

@@
 Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman Syndrome r(`r myn["z_cp","AS"]`) = `r myr["z_cp","AS"]`, CI [`r mylb["z_cp","AS"]`,`r myhb["z_cp","AS"]`], p=`r myp["z_cp","AS"]`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r myr["z_cp","LR"]`, CI [`r mylb["z_cp","LR"]`,`r myhb["z_cp","LR"]`], p=`r myp["z_cp","LR"]`].
+```{r bychunk-age}
+read.csv("../Derived_Data/chunk_majority_agreement_1751.csv")->chunk
+bychild=table(chunk$Answer,chunk$ChildID)
+cans=bychild["Canonical",]
+noncans=bychild["Non-Canonical",]
+junk=bychild["Junk",]
+sum_bychild=table(chunk$ChildID)
+
+ling_prop=(cans+noncans)/(sum_bychild-junk)
+can_prop=cans/(cans+noncans)
+
+demo_data$ChildID==names(ling_prop)
+demo_data$ChildID==names(can_prop)
+
+props_bychild_chunk=cbind(ling_prop,can_prop,demo_data)
+colnames(props_bychild_chunk)[1:2]<-c("chi","ling_prop")
+
+z_lp_as=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
+
+z_lp_lr=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
+
+z_cp_as=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
+
+z_cp_lr=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
+```
 
@@
 We therefore repeated our analyses but deriving our proportions for the Zooniverse data not from the segment-level composite, but rather the individual chunk-level annotations.
-We found that the Linguistic proportion was not strongly associated with age in the Angelman Syndrome group [r(8) = -.04, CI -.6 .6, p=0.91], but increased with age  in the low-risk control group [r(8) = .71, CI .15 .92, p=.02].
-Similarly, the Canonical proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman Syndrome r(8) = -.35, CI -.80 .35, p=.32; and developmental increases among low-risk control r(8) = .57, CI -.08 0.88, p=.83.
+We found that the Linguistic proportion was not strongly associated with age in the Angelman Syndrome group: r(`r z_lp_as$parameter`) = `r z_lp_as$estimate`, CI [`r z_lp_as$conf.int[1]`,`r z_lp_as$conf.int[2]`], p=`r z_lp_as$p.value`], but increased with age  in the low-risk control group: r(`r z_lp_lr$parameter`) = `r z_lp_lr$estimate`, CI [`r z_lp_lr$conf.int[1]`,`r z_lp_lr$conf.int[2]`], p=`r z_lp_lr$p.value`.
+Similarly, the Canonical proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman Syndrome r(`r z_cp_as$parameter`) = `r z_cp_as$estimate`, CI [`r z_cp_as$conf.int[1]`,`r z_cp_as$conf.int[2]`], p=`r z_cp_as$p.value`]; and developmental increases among low-risk control r(`r z_cp_lr$parameter`) = `r z_cp_lr$estimate`, CI [`r z_cp_lr$conf.int[1]`,`r z_cp_lr$conf.int[2]`], p=`r z_cp_lr$p.value`.
 
@@
 
-```{r}
+```{r mygac-conf}
 mygac=gac(data = sample_data, kat = 5, weight = c("unweighted"),
@@
 
-When we use chunk-level judgments to derive the our child-level metrics, we observe very similar levels of correlation: linguistic proportion overall r (18) = .84, CI .63 .93, p<0.001 [Angelman Syndrome r(8) = .82, CI .41 .95, p<0.01; low-risk control r(8) = .84, CI .44 0.96, p<0.01]; canonical proportion overall r (18) = .96, CI .90 .98, p<0.001 [Angelman Syndrome r(8) = .85, CI .49 .85, p<0.01; low-risk control r(8) = .97, CI .89 .99, p<0.001].
+```{r zoo-lab-chunk}
+
+#create matrix to hold r results
+myr=matrix(NA,nrow=2,ncol=3)
+rownames(myr)<-c("lp","cp")
+colnames(myr)<-c("AS","LR","all")
+mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
+
+proportions=proportions[order(proportions$ChildID),]
+
+  #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
+    myr[1,"AS"]=round(thiscor$estimate,3)
+    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
+    myn[1,"AS"]=round(thiscor$parameter,3)
+    myp[1,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"])
+    myr[1,"LR"]=round(thiscor$estimate,3)
+    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
+    myn[1,"LR"]=round(thiscor$parameter,3)
+    myp[1,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(props_bychild_chunk$ling_prop,proportions$l_lp)
+    myr[1,"all"]=round(thiscor$estimate,3)
+    mylb[1,"all"]=round(thiscor$conf.int[1],3)
+    myhb[1,"all"]=round(thiscor$conf.int[2],3)
+    myn[1,"all"]=round(thiscor$parameter,3)
+    myp[1,"all"]=round(thiscor$p.value,3)
+
+    # repeat whole for cp
+
+  #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
+    myr[2,"AS"]=round(thiscor$estimate,3)
+    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
+    myn[2,"AS"]=round(thiscor$parameter,3)
+    myp[2,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"])
+    myr[2,"LR"]=round(thiscor$estimate,3)
+    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
+    myn[2,"LR"]=round(thiscor$parameter,3)
+    myp[2,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(props_bychild_chunk$can_prop,proportions$l_cp)
+    myr[2,"all"]=round(thiscor$estimate,3)
+    mylb[2,"all"]=round(thiscor$conf.int[1],3)
+    myhb[2,"all"]=round(thiscor$conf.int[2],3)
+    myn[2,"all"]=round(thiscor$parameter,3)
+    myp[2,"all"]=round(thiscor$p.value,3)
+```
+
+When we use chunk-level judgments to derive the our child-level metrics, we observe very similar levels of correlation: linguistic proportion overall r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman Syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
+As for canonical proportion overall r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`; Angelman Syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`];  low-risk control infants r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
 

### 2020-10-27 202010_jslhr/analyses.Rmd
added explo ana - first half

@@
 
-```{r}
+```{r, echo=F}
 # remove non-majority labels
@@
 
-```{r prepare data by child}
+```{r prepare data by child, echo=F}
 
@@
 
-```{r corage}
+```{r corage, echo=F}
 
@@
 Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman Syndrome r(`r myn["z_cp","AS"]`) = `r myr["z_cp","AS"]`, CI [`r mylb["z_cp","AS"]`,`r myhb["z_cp","AS"]`], p=`r myp["z_cp","AS"]`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r myr["z_cp","LR"]`, CI [`r mylb["z_cp","LR"]`,`r myhb["z_cp","LR"]`], p=`r myp["z_cp","LR"]`].
-```{r bychunk-age}
-read.csv("../Derived_Data/chunk_majority_agreement_1751.csv")->chunk
-bychild=table(chunk$Answer,chunk$ChildID)
-cans=bychild["Canonical",]
-noncans=bychild["Non-Canonical",]
-junk=bychild["Junk",]
-sum_bychild=table(chunk$ChildID)
-
-ling_prop=(cans+noncans)/(sum_bychild-junk)
-can_prop=cans/(cans+noncans)
-
-demo_data$ChildID==names(ling_prop)
-demo_data$ChildID==names(can_prop)
-
-props_bychild_chunk=cbind(ling_prop,can_prop,demo_data)
-colnames(props_bychild_chunk)[1:2]<-c("chi","ling_prop")
-
-z_lp_as=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
-
-z_lp_lr=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
-
-z_cp_as=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
-
-z_cp_lr=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
-```
-
-Given that these derived metrics are defined at the child level, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
-We therefore repeated our analyses but deriving our proportions for the Zooniverse data not from the segment-level composite, but rather the individual chunk-level annotations.
-We found that the Linguistic proportion was not strongly associated with age in the Angelman Syndrome group: r(`r z_lp_as$parameter`) = `r z_lp_as$estimate`, CI [`r z_lp_as$conf.int[1]`,`r z_lp_as$conf.int[2]`], p=`r z_lp_as$p.value`], but increased with age  in the low-risk control group: r(`r z_lp_lr$parameter`) = `r z_lp_lr$estimate`, CI [`r z_lp_lr$conf.int[1]`,`r z_lp_lr$conf.int[2]`], p=`r z_lp_lr$p.value`.
-Similarly, the Canonical proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman Syndrome r(`r z_cp_as$parameter`) = `r z_cp_as$estimate`, CI [`r z_cp_as$conf.int[1]`,`r z_cp_as$conf.int[2]`], p=`r z_cp_as$p.value`]; and developmental increases among low-risk control r(`r z_cp_lr$parameter`) = `r z_cp_lr$estimate`, CI [`r z_cp_lr$conf.int[1]`,`r z_cp_lr$conf.int[2]`], p=`r z_cp_lr$p.value`.
 
@@
 
-```{r mygac-conf}
+```{r mygac-conf, echo=F}
 mygac=gac(data = sample_data, kat = 5, weight = c("unweighted"),
@@
 
-```{r figure-prec-rec}
+```{r figure-prec-rec, echo=F}
 
@@
 for(i in 1:dim(conf_tab)[2]) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
-colSums(my_conf_tab)
+#colSums(my_conf_tab) #check
 prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
@@
 
-```{r corlab-zoo}
+```{r corlab-zoo, echo=F}
 
@@
     myp[2,"all"]=round(thiscor$p.value,3)
+
+mycortab=t(myr)
+rownames(mycortab)<-paste("All seg",rownames(mycortab))
 ```
@@
 
-```{r zoo-lab-chunk}
-
-#create matrix to hold r results
-myr=matrix(NA,nrow=2,ncol=3)
-rownames(myr)<-c("lp","cp")
-colnames(myr)<-c("AS","LR","all")
-mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
-
-proportions=proportions[order(proportions$ChildID),]
-
-  #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
-    myr[1,"AS"]=round(thiscor$estimate,3)
-    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
-    myn[1,"AS"]=round(thiscor$parameter,3)
-    myp[1,"AS"]=round(thiscor$p.value,3)
-
-  #repeat for LR
-
-      #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"])
-    myr[1,"LR"]=round(thiscor$estimate,3)
-    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
-    myn[1,"LR"]=round(thiscor$parameter,3)
-    myp[1,"LR"]=round(thiscor$p.value,3)
-
-#repeat for all cor
-
-      thiscor=cor.test(props_bychild_chunk$ling_prop,proportions$l_lp)
-    myr[1,"all"]=round(thiscor$estimate,3)
-    mylb[1,"all"]=round(thiscor$conf.int[1],3)
-    myhb[1,"all"]=round(thiscor$conf.int[2],3)
-    myn[1,"all"]=round(thiscor$parameter,3)
-    myp[1,"all"]=round(thiscor$p.value,3)
-
-    # repeat whole for cp
-
-  #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
-    myr[2,"AS"]=round(thiscor$estimate,3)
-    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
-    myn[2,"AS"]=round(thiscor$parameter,3)
-    myp[2,"AS"]=round(thiscor$p.value,3)
-
-  #repeat for LR
-
-      #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"])
-    myr[2,"LR"]=round(thiscor$estimate,3)
-    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
-    myn[2,"LR"]=round(thiscor$parameter,3)
-    myp[2,"LR"]=round(thiscor$p.value,3)
-
-#repeat for all cor
-
-      thiscor=cor.test(props_bychild_chunk$can_prop,proportions$l_cp)
-    myr[2,"all"]=round(thiscor$estimate,3)
-    mylb[2,"all"]=round(thiscor$conf.int[1],3)
-    myhb[2,"all"]=round(thiscor$conf.int[2],3)
-    myn[2,"all"]=round(thiscor$parameter,3)
-    myp[2,"all"]=round(thiscor$p.value,3)
-```
-
-When we use chunk-level judgments to derive the our child-level metrics, we observe very similar levels of correlation: linguistic proportion overall r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman Syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
-As for canonical proportion overall r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`; Angelman Syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`];  low-risk control infants r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
-
-```{r glmer-agediag}
+```{r glmer-agediag, echo=F}
 # one outlier in ling prop is removed
@@
 
-```{r prelim-junk}
+```{r prelim-junk, echo=F}
 junkXdiag=t.test(proportions$z_junk ~ proportions$Diagnosis)
@@
 
-```{r cor-junk}
+```{r cor-junk, echo=F}
 
@@
 
+Above, we concluded that derived metrics seem more promising than segment-level data.
+Notice that derived metrics do not require matching of clips (500 ms presented to Zooniverse participants) to segments (the original LENA segments presented to laboratory participants).
+As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
+We therefore repeated our analyses but deriving our proportions for the Zooniverse data not from the segment-level composite, but rather the individual chunk-level annotations.
+
+```{r bychunk-age, echo=F}
+read.csv("../Derived_Data/chunk_majority_agreement_1751.csv")->chunk
+bychild=table(chunk$Answer,chunk$ChildID)
+cans=bychild["Canonical",]
+noncans=bychild["Non-Canonical",]
+junk=bychild["Junk",]
+sum_bychild=table(chunk$ChildID)
+
+ling_prop=(cans+noncans)/(sum_bychild-junk)
+can_prop=cans/(cans+noncans)
+
+#check that they match before merging
+#demo_data$ChildID==names(ling_prop)
+#demo_data$ChildID==names(can_prop)
+
+props_bychild_chunk=cbind(ling_prop,can_prop,demo_data)
+colnames(props_bychild_chunk)[1:2]<-c("chi","ling_prop")
+
+z_lp_as=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
+
+z_lp_lr=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
+
+z_cp_as=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
+
+z_cp_lr=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
+```
+
+Regarding correlations with age, we found that the Linguistic proportion was not strongly associated with age in the Angelman Syndrome group: r(`r z_lp_as$parameter`) = `r z_lp_as$estimate`, CI [`r z_lp_as$conf.int[1]`,`r z_lp_as$conf.int[2]`], p=`r z_lp_as$p.value`], but increased with age  in the low-risk control group: r(`r z_lp_lr$parameter`) = `r z_lp_lr$estimate`, CI [`r z_lp_lr$conf.int[1]`,`r z_lp_lr$conf.int[2]`], p=`r z_lp_lr$p.value`.
+Similarly, the Canonical proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman Syndrome r(`r z_cp_as$parameter`) = `r z_cp_as$estimate`, CI [`r z_cp_as$conf.int[1]`,`r z_cp_as$conf.int[2]`], p=`r z_cp_as$p.value`]; and developmental increases among low-risk control r(`r z_cp_lr$parameter`) = `r z_cp_lr$estimate`, CI [`r z_cp_lr$conf.int[1]`,`r z_cp_lr$conf.int[2]`], p=`r z_cp_lr$p.value`.
+
+```{r zoo-lab-chunk, echo=F}
+
+#create matrix to hold r results
+myr=matrix(NA,nrow=2,ncol=3)
+rownames(myr)<-c("lp","cp")
+colnames(myr)<-c("AS","LR","all")
+mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
+
+proportions=proportions[order(proportions$ChildID),]
+
+  #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
+    myr[1,"AS"]=round(thiscor$estimate,3)
+    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
+    myn[1,"AS"]=round(thiscor$parameter,3)
+    myp[1,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"])
+    myr[1,"LR"]=round(thiscor$estimate,3)
+    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
+    myn[1,"LR"]=round(thiscor$parameter,3)
+    myp[1,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(props_bychild_chunk$ling_prop,proportions$l_lp)
+    myr[1,"all"]=round(thiscor$estimate,3)
+    mylb[1,"all"]=round(thiscor$conf.int[1],3)
+    myhb[1,"all"]=round(thiscor$conf.int[2],3)
+    myn[1,"all"]=round(thiscor$parameter,3)
+    myp[1,"all"]=round(thiscor$p.value,3)
+
+    # repeat whole for cp
+
+  #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
+    myr[2,"AS"]=round(thiscor$estimate,3)
+    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
+    myn[2,"AS"]=round(thiscor$parameter,3)
+    myp[2,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"])
+    myr[2,"LR"]=round(thiscor$estimate,3)
+    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
+    myn[2,"LR"]=round(thiscor$parameter,3)
+    myp[2,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(props_bychild_chunk$can_prop,proportions$l_cp)
+    myr[2,"all"]=round(thiscor$estimate,3)
+    mylb[2,"all"]=round(thiscor$conf.int[1],3)
+    myhb[2,"all"]=round(thiscor$conf.int[2],3)
+    myn[2,"all"]=round(thiscor$parameter,3)
+    myp[2,"all"]=round(thiscor$p.value,3)
+
+ mycortab=rbind(mycortab,t(myr))
+rownames(mycortab)[4:6]<-paste("Chunks",rownames(t(myr)))
+
+```
+
+As for correlations between Zooniverse and laboratory-derived metrics, we observe very similar levels of correlation: linguistic proportion overall r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman Syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
+As for canonical proportion overall r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`; Angelman Syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`];  low-risk control infants r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
+
+Since all of these results are very similar to those obtained when first matching clips to segments, we conclude that in the future this step may not be necessary.
+Instead, researchers can derive linguistic and canonical proportions directly from citizen scientists' clip level judgements [which was in fact what @cyhosz did].
+
+```{r downsample}
+
+#we start from the data base that has all the info
+nsegs=data.frame(table(data_all$ChildID))
+colnames(nsegs)<-c("ChildID","nsegs")
+
+merge(nsegs,demo_data)->nsegs
+nseg_t=t.test(nsegs$nsegs~nsegs$Diagnosis)
+
+resampling=matrix(NA,nrow=50,ncol=6)
+colnames(resampling)<-c("lp_r_all","lp_r_as","lp_r_lr","cp_r_all","cp_r_as","cp_r_lr")
+for(i in 1:50){
+  sel_seg=NULL
+  for(thischi in levels(data_all$ChildID)) sel_seg<-c(sel_seg,as.character(data_all$segmentId_DB[sample(data_all$segmentId_DB[data_all$ChildID==thischi],100)]))
+
+  sampling=data_all[as.character(data_all$segmentId_DB) %in% as.character(sel_seg),]
+
+  #get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
+  ztab=table(sampling$ChildID,sampling$Zoon_classif)
+  z_lp=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
+  z_cp=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
+
+  ltab=table(sampling$ChildID,sampling$lab)
+  l_lp=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
+  l_cp=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
+
+  thistab=cbind(demo_data,z_lp,z_cp,l_lp,l_cp)
+
+  resampling[i,"lp_r_all"]=cor.test(thistab$z_lp,thistab$l_lp)$estimate
+  resampling[i,"lp_r_as"]=cor.test(thistab$z_lp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
+  resampling[i,"lp_r_lr"]=cor.test(thistab$z_lp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate
+
+   resampling[i,"cp_r_all"]=cor.test(thistab$z_cp,thistab$l_cp)$estimate
+  resampling[i,"cp_r_as"]=cor.test(thistab$z_cp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
+  resampling[i,"cp_r_lr"]=cor.test(thistab$z_cp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate
+ }
+myr=colMeans(resampling)
+mycortab=rbind(mycortab,cbind(myr[c("lp_r_as" ,"lp_r_lr","lp_r_all")],myr[c("cp_r_as" ,"cp_r_lr","cp_r_all")]))
+rownames(mycortab)[7:9]<-c("100 seg AS","100 seg LR","100 seg all")
+round(mycortab,3)->mycortab
+colnames(mycortab)<-c("Ling. Prop.","Can. Prop.")
+resampling=data.frame(resampling)
+```
+
+Next, we looked at whether having more segments from each child may lead to more reliable metrics.
+We observed a non-significant trend [t(`r nseg_t$parameter`)=`r nseg_t$estimate`] for lower number of segments in the Angelman Syndrome group (mean `r nseg_t$estimate[1]`, range `r min(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`) than among low-risk control infants (mean `r nseg_t$estimate[2]`, range `r min(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`), likely because of the lower volubility of the former children.
+Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications [CITE MEG].
+We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children [following CITE MEG].
+We randomly sampled 100 segments from each child's data, and recalculated the linguistic and canonical proportions based only on these associated judgments.
+We repeated this process 50 times, to assess the extent to which the association between laboratory- and Zooniverse-derived metrics varied in each random sample.
+The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table XX), particularly for the low-risk younger infants, and for linguistic proportion.
+This may indicate that, particularly in some groups of infants, 100 segments may not be sufficient to capture the child's linguistic and canonical proportions.
+We also observed quite a bit of variance across the 50 runs.
+For linguistic proportion, the range of correlations found for all infants was [`r min(resampling$lp_r_all)`,`r max(resampling$lp_r_all)`]; for the Angelman Syndrome group [`r min(resampling$lp_r_as)`,`r max(resampling$lp_r_as)`]; for the low-risk group [`r min(resampling$lp_r_lr)`,`r max(resampling$lp_r_lr)`].
+For canonical proportion, the range of correlations found for all infants was [`r min(resampling$cp_r_all)`,`r max(resampling$cp_r_all)`]; for the Angelman Syndrome group [`r min(resampling$cp_r_as)`,`r max(resampling$cp_r_as)`]; for the low-risk group [`r min(resampling$cp_r_lr)`,`r max(resampling$cp_r_lr)`].
+
+```{r tab-main, results="as.is"}
+kable(mycortab)
+```
+
 <!-- - look at whether kids who have a high "junk" proportion according to Zooniverse  -->
@@
 
-- take all 5 Zoo judgments or 3 or 1 (may need to run the latter two a couple of times to check for potential variance -- if variance is small, we'll just say so, if it's large then we'll have to write a loop and report CIs)
-- look at what happens to TD Zoo-PU correlation when you undersample 100 vocs to 80 and 60 and to the same number of non-junk vocs as in AS
+<!-- - take all 5 Zoo judgments or 3 or 1 (may need to run the latter two a couple of times to check for potential variance -- if variance is small, we'll just say so, if it's large then we'll have to write a loop and report CIs) -->

### 2020-10-28 202010_jslhr/paper.Rmd
paper first full version and it knits

---
title: "Describing vocalizations in young children: A big data approach through citizen science annotations"
shorttitle        : "Children's vocalizations and citizen science"

author: 
  - name          : "Chiara Semenzin"
    affiliation   : "1"
    corresponding : no
  - name          : "Lisa Hamrick"
    affiliation   : "2"
    corresponding : no
  - name          : "Amanda Seidl"
    affiliation   : "2"
    corresponding : no
  - name          : "Bridgette Kelleher"
    affiliation   : "2"
    corresponding : no
  - name          : "Alejandrina Cristia"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "29 rue d'Ulm, 75005, Paris, France"
    email         : "alecristia@gmail.com"

affiliation:
  - id            : "1"
    institution   : "Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University"
  - id            : "2"
    institution   : "Purdue University"

author_note: >
       AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017); and the J. S. McDonnell Foundation Understanding Human Cognition Scholar Award. LH acknowledges the National Institute of Deafness and Other Communication Disorders (F31DC018219). BK acknowledges the National Institute of Mental Health (K23MH111955) and the Kinley Trust. The funders had no impact on this study.  All authors approved the final manuscript as submitted and agree to be accountable for all aspects of the work and have no conflict of interests to disclose.

# Abbreviations LENA^TM^: Language Environment Analysis.
# ASHA recommends that abstracts be 150–250 words. The size limit for what can be included in your submission is set above 300 words,
#currently 259 words

abstract: >
  **Purpose:** Recording young children's vocalizations through wearables is a promising method. However, accurately and rapidly annotating these files remains challenging. Online crowdsourcing with the collaboration of citizen scientists could be a feasible solution. In this paper, we assess the extent to which citizen scientists' annotations align with those gathered in the lab for recordings collected from young children. 
  \newline

  **Method:** Segments identified by LENA^TM^ as produced by the key child were extracted from one daylong recording for each of 20 participants: 10  low-risk control children and 10 children diagnosed  with  Angelman  syndrome, a  neurogenetic  syndrome  characterized  by  severe  language  impairments. Speech samples were annotated by trained annotators in the laboratory as well as by citizen scientists on Zooniverse. All annotators assigned one of five labels to each sample: Canonical, Non-Canonical, Crying, Laughing, and Junk. This allowed the derivation of two child-level vocalization metrics: the Linguistic Proportion, and the Canonical Proportion.

  \newline

  **Results:** At the segment level, Zooniverse classifications had moderate precision and recall. More importantly, the Linguistic Proportion and the Canonical Proportion derived from Zooniverse annotations were highly correlated with those derived from laboratory annotations.

  \newline

  **Conclusion:** Annotations obtained through a citizen science platform can help us overcome challenges posed by the process of annotating daylong speech recordings. Particularly when used in composites or derived metrics, such annotations can be used to investigate early markers of language delays in non-typically developing children.

# type: speech; areas of interest: assessment, infants, meta-analyses, prevention, speech recognition, screening, language, early intervention, cultural and ling diversity, dev lang disorders, autism, bilingualism --> choose 5: infants, prevention, dev lang disorders, early interv, diagnostics
# Clinical practice research that evaluates methods of prevention, screening, diagnosis, treatment or models of service delivery that may improve the function and quality of life for individuals with or at risk of developing disorders of speech, language, cognition, swallowing, audition, and/or balance.
# #up to 3 keywords
# keywords          : " reliability; validation; daylong audiorecordings"
# wordcount         : "7,113"

bibliography      : ["zoo.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

header-includes:
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\singlespacing}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1.5}}

class             : "man"
output            : papaja::apa6_pdf

---

```{r setup, include = FALSE}
library("papaja")
library(tidyverse)
library(dplyr)
library(ggplot2)
library(caret)
library(scales)
library(kableExtra)
library(rel)
library(ggpubr)
library(lme4)
#knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = FALSE, fig.pos = "T")
knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = FALSE)
```

```{r analysis-preferences, include = FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r read-in}
# read datasets

data_ang <- read.csv("../Derived_Data/classifications_PU_zoon_final17.csv",header=T,sep=",")
data_td <- read.csv("../Derived_Data/classifications_PU_zoon_final.csv")
data_all<-rbind(data_ang, data_td)

#remove the word mixed that takes up space and is unnecessary
data_all$Zoon_classif=factor(gsub("Mixed_","",as.character(data_all$Zoon_classif),fixed=T))
#relevel the factor so that it's easier to read
data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=c("Canonical","Non-Canonical",                   "Crying","Laughing","Junk",levels(data_all$Zoon_classif)[grep("_",levels(data_all$Zoon_classif))]))
# create lab column with easier to read correspondance
data_all$lab<-as.character(data_all$Major_Choice)
data_all$lab[data_all$lab=="Non-canonical syllables"]<-"Non-Canonical"
data_all$lab[data_all$lab=="Canonical syllables"]<-"Canonical"
data_all$lab[data_all$lab %in% c("Don't mark","None")]<-"Junk"
data_all$lab=factor(data_all$lab,levels=levels(data_all$Zoon_classif))
#apply same factor levels as zooniverse so that we can do symmetrical confusion matrices

#add binomials for Linguistic Proportion
data_all$lab_ling=ifelse(data_all$lab %in% c("Canonical","Non-Canonical"),1,0)
data_all$zoo_ling=ifelse(data_all$Zoon_classif %in% c("Canonical","Non-Canonical"),1,0)
data_all$lab_ling[data_all$lab=="Junk"]<-NA
data_all$zoo_ling[data_all$lab=="Junk"]<-NA

#add binomials for Canonical Proportion
data_all$lab_can=data_all$zoo_can=NA
data_all$lab_can[data_all$lab=="Canonical"]<-1
data_all$lab_can[data_all$lab=="Non-Canonical"]<-0
data_all$zoo_can[data_all$Zoon_classif=="Canonical"]<-1
data_all$zoo_can[data_all$Zoon_classif=="Non-Canonical"]<-0

demo_data=read.csv("../Derived_Data/demo-data.tsv",sep="\t")
#add filenames to demo data, to be used later
demo_data_fn <- demo_data %>% 
     left_join(select(data_all, filename, ChildID), by = c("ChildID"))
demo_data_fn<-unique(demo_data_fn)

```

TODO

- create OSF for this paper specifically
- references!
- merge noncan/can for LENA^TM^ in Table 1
- appropriately merge col names in tab 1
- add 2 explo analyses that are still missing
- Redo pipeline so that it's transparent & integrated

Language production at 3 years of age is the best predictor of later academic and economic achievement (Law et al., 2003). Early language delays can adversely affect children’s literacy, behavior, social interaction, and scholastic achievement extending well into adulthood, with early interventions being a better societal investment than later ones (Heckman, 2006). Advancements in the field of wearable technologies, such as LENA^TM^ recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally. Wearable recorders allow data collection to happen in the child’s natural environment, and at a large scale, which may be particularly helpful for children whose speech is not easily elicited. Although such long-form recordings are increasingly common REF, challenges remain with respect to how these data are handled, annotated, and analyzed. The present work reports on the validity of labels and child-level descriptors of children's vocalizations gathered from citizen scientists, in a comparison to expert lab annotators. In the rest of the Introduction, we first briefly summarize two metrics that can be used to describe individual children's vocalizations in the context of long-form recordings. We then introduce crowdsourcing in general, and crowdsourcing by citizen scientists in particular, as a potential avenue for more rapidly deriving these key metrics from vocalization recordings.

###  Describing children's vocalizations in long-form recordings

A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis, and it is beyond the scope of this paper to provide a full summary of this work. We instead focus on two descriptors -- Linguistic Proportion and Canonical Proportion -- that are particularly amenable to the kind of big data approaches long-form recordings require.

To understand why these two descriptors are relevant, it is worthwhile to first consider how the way in which vocal development has been studied in the past may not easily translate to long-form recordings. As mentioned earlier, long-form recordings have several advantages, including capturing the child's vocal patterns in their natural environment and being able to accumulate a great amount of data easily, which is particularly useful for diagnoses and pathologies characterized by low levels of vocal production. However, this also means that there is a great amount of data, and that the audio quality is compromised, as children go about their day transitioning from small to large rooms, with fewer or more people around them, thus providing analysts with an unstable and challenging long audio. In fact, human annotation of such audio is estimated to take about 30 times the audio length, with these estimates being greater the more precise one wants or needs to be (Seidl et al., 2019). It is for this reason that most users of long-form recordings have turned to automatized software for at least a first-pass analysis.

The most commonly used software was created by the LENA^TM^ Foundation, and it returns a diarization of the audio signal split into key talkers, including the child wearing the device, as well as a split of these audio attributed to the child into three categories: crying, other fixed and vegetative signals, and speech vocalizations. These subcategories have not been widely validated (Cristia et al., 2020). To provide an idea of their accuracy, we classified the same segments that are used in the present study using LENA^TM^'s subtypes[^1] to compared these against the laboratory annotations used in the present study. Table 1 shows that LENA^TM^ has a good recall for speech-like vocalizations and crying but not laughing. Moreover, LENA^TM^ lacks an important distinction between more and less advanced vocalizations.

<!-- AS said about not laughing: might be worth noting how low frequency this is and how that might be part of the problem in addition to being difficult to distinguish from crying. -->

[^1]: If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as "junk" or not categorized.

```{r tab-LENA-challenge, comment = F, message = F, hide = T, warning = F, echo = F, results = "asis"}
mycaption="Comparison of recall percentages obtained with the baseline algorithm created by the ComParE team, in the context of the ComParE 2019 BabySounds subchallenge (Schuller et al., 2019), LENATM labels (note LENATM does not distinguish between Canonical and Non-canonical), and those obtained in this study through Zooniverse annotations. Label frequency indicates the prevalence of the relevant label (Crying, Laughing, etc.) in each dataset (the same Lab dataset was used for both LENATM and Zooniverse). UAR stands for unweighted average recall, WAR for weighted average recall (which takes into account label frequency)."

read.table("tab-lena-challenge.txt",header=T,sep="\t")->tab1
colnames(tab1)<-c("","Recall","","","Label","Frequency")
kable(tab1,cap=mycaption)
```

Other automatized algorithms have been developed in recent years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical (Schuller et al., 2019). The split of speech vocalizations into canonical and non-canonical, and the overt inclusion of laughing, allow a somewhat finer-grained picture of children's vocal activity. An advantage of these algorithms is that there exists a challenge where performance has been benchmarked, so unlike LENA^TM^'s algorithms, we know which algorithms perform better than others. The ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55% (see Table 1). The team who won the challenge improved this by about 2%, primarily through gains in the laughing class obtained by adding training data (REF). Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are considered, changes very much.

One possibility that has only recently begun to be explored is the use of derived metrics. So instead of evaluating algorithms and other annotation procedures on their accuracy at individual labels, one can derive a metric that more closely relates to vocalization development.

For example, most work using LENA^TM^ software reports on children's vocalization counts. Although this has been sometimes criticized as being more about quantity than quality REF, this is a promising metric because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA^TM^ REF. Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r ~ .3 with standardized language scores in a meta-analysis REF. Thus despite representing a composite of skills, vocalization count may provide a useful estimate of vocalization development.

<!-- AS said about voc count not sure I agree: https://pubs.asha.org/doi/full/10.1044/2020_JSLHR-19-00281?casa_token=fBByMmPlxbMAAAAA%3AbS2O5Mrymn30_aQWybitwMecENkDYLTjzexCqZcnX2Rx1VJ7OaT9qkIGNAAM95xRBGW_D3XX7mMEWHo -->

Here, we focus on two alternative composite metrics. An additional metric that can be derived once children's vocalizations are split into crying, laughing, canonical, and non-canonical is the Linguistic Proportion: the proportion of vocalizations that are linguistic (canonical and non-canonical) out of all vocalizations. To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age REF.

<!-- BT said about LP May want to put this 2nd as more exploatory given more lit on can. prop. -->

Yet another metric that could be extracted to estimate relative linguistic complexity is the Canonical Proportion: the proportion of vocalizations that are canonical out of all linguistic (canonical and non-canonical) vocalizations. One recent study has established that the Canonical Proportion extracted from daylong recordings is significantly correlated with age in a multicultural and multilingual sample REF. Importantly, this association was established in a sample of children going up to 3 years of age, and thus well beyond the babbling period, and into first words and word combinations.

There is a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this work focuses on babbling and thus typically on infants under one year of age. A critical milestone involves the increasingly common production of Canonical Syllables, consonant-vowel (CV) or vowel-consonant (VC) sequences that resemble those found in adult speech (Oller et al., 1998). Given its adult-like CV or VC structure, canonical babble is considered to be a starting point on the path to recognizable speech. Canonical syllables show a higher complexity given the smooth articulatory transition between a consonant and a vowel (or vice versa), when compared to more primitive sounds such as squeals, or isolated vowels. Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development (Oller et al., 1999). In addition, delayed onset of canonical babbling has been documented in a wide variety of disorders and delays (Belardi et al., 2017; Fasolo et al., 2008; Lang et al., 2019; Patten et al., 2014; Stoel-Gammon, 2016).

<!-- AS said: maybe name the disorders & delays here? -->

In sum, there is a growing literature attempting to use data from daylong recordings to describe young children's vocalizations, but there are two outstanding challenges. The first pertains to how the data are analyzed, with human annotation being costly. The second relates to how to extract useful descriptors, both in terms of individual vocalizations and at the child level.

### Crowdsourcing: A potential solution for annotation

Crowdsourcing refers to the process whereby a task is solved by a crowd, rather than an individual. A number of fields, particularly in the data-driven sciences, have already engaged in the collection of data (including annotations) through crowdsourcing, thanks to its low cost and ecological value (Crump, McDonnell, & Gureckis, 2013; Sescleifer, Francoisse, & Lin, 2018;  Sprouse, 2011). For example, a systematic review on crowd-sourced ratings of speech (Sescleifer et al., 2018) found that "lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology"; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81. On the other hand, the systematic review returned only 8 studies (of which only four were published in peer-reviewed journals), suggesting that there is considerable need for further research on this topic.

Mechanical Turk (MTurk) MTurk is an online labor market created by Amazon to assist "requesters"''" in hiring and paying "workers" for the completion of computerized tasks. Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable (Berinsky et al., 2012), some question marks are raised as some "workers'" turn out to be bots, or poorly motivated (and potentially exploited).

A promising crowd-sourcing alternative has arisen in recent years: citizen science, a research technique that engages the public in the collection of scientific information (ref). As citizen scientists do not receive compensation, this alternative to platforms such as MTurk can overcome the limitations posed by potential exploitation of workforce and/or the use of bots. Volunteers are entirely motivated by the desire to contribute to research advancements as well as the pleasure they derive from the task itself.

One of the most successful platforms hosting citizen science projects is Zooniverse (zooniverse.org). The website hosts a multitude of interdisciplinary projects that have allowed the public to take part in cutting-edge scientific research, from marine biology to papyrology. Zooniverse has proven extremely useful in those fields where the complexity of the data collected is too high to be automatically interpreted using computer algorithms (ref) At the same time, the tasks that human volunteers are asked to complete are sufficiently simple that citizens can carry them out without a background in science or any extensive training (ref).

Citizen science may be particularly helpful when analyzing infants' data from wearables, which remain challenging to annotate as mentioned above. There is one previous study that attempted this approach. @cychosz2019canonical drew child vocalization data from a diverse set of corpora centered on children learning one of four languages: English, Tseltal, Tsimane', and Yélî. For the English and Tsimane' corpora, vocalizations were automatically identified using LENA^TM^, whereas the other two were extracted through manual segmentation. Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received 3 classifications into crying, laughing, canonical, non-canonical, or "junk" (with the latter tag used for clips that did not contain clear child voice). They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgement of canonical or non-canonical. They found that this canonical ratio increased with infant age in an analysis collapsing across corpora, which provides a first proof of principle that citizen scientists' annotations could be an appropriate solution to the problem of annotating child vocalizations from wearables. The authors also studied how their canonical ratio related to age within each corpus that had a sufficient number of children varying in age. They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the "junk" category was very high. Additionally, they found that the correlation between canonical ratio and age was much weaker in two corpora where vocalizations were segmented with LENA^TM^ than through manual segmentation. In sum, these analyses within corpora reveal that data may not be uniformly useful, with potential variation across corpora, children, and methods.

### The present work

Our work seeks to broaden the already promising results that have emerged from previous work summarized above. We hope that this methodology will open the road to larger scale analyses of children's vocalizations as captured by wearables. In addition, we went beyond @cychosz2019canonical's study in two important ways. First, we relied on the largest and best established citizen science platform: Zooniverse hosts more than 1.6 million users from diverse walks of life, and it offers a completely automatized API system to more easily scale tasks in a transparent and cumulative science fashion. In contrast, the authors of that paper told us that, in order to complete their data annotation, they recruited the help of students in their classes and labs, and sometimes the researchers themselves took on a considerable annotation load. This means that the process draws on expert resources, whose time is expensive. Moreover, the people these researchers recruit tend to be more accustomed to child vocalization data, which may improve their annotation performance, and thus provide an overestimate of the quality of the annotations that can be realistically done within a truly citizen science framework. Those authors also had to rely on manual extraction of the resulting data from the iHearUplay team, which added to the administrative load. Second, a core goal of the present work was to determine how citizen scientists' annotations fare compared to the current gold standard, laboratory annotations. @cychosz2019canonical did not have a gold standard for their data.

Thus, the present study aimed to contribute a key piece of evidence missing in this discussion: To what extent do laboratory and citizen science annotations agree when describing young children's vocalizations? We examined the extent to which such classifications agree, by quantifying the correspondence across these two modes of annotations at the level of individual clips (vocalizations), and at the level of individual children. It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels. We therefore included children in a wide age range, with some of the children being diagnosed with Angelman Syndrome, as well as low-risk controls.

At the segment level, we checked the extent to which laboratory annotations made by experts agreed with judgments made by citizen scientists. We used confusion matrices to describe annotation convergence and divergence patterns, and overall accuracy, kappa, and Gwet's AC1 coefficient as statistical descriptors. Given previous results from machine learning classifiers REF, we expected agreement to be moderate at this level, with potentially more erratic results for categories that are rarer.

When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (canonical+non-canonical)/(all vocalizations); and the Canonical Proportion (canonical)/(all linguistic vocalizations). Previous work suggests that children with Angelman Syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the canonical ratio was expected to increase with age [@hamrick2019capturing]. Although we could not rely on previous work to make predictions regarding the Linguistic Proportion, we reasoned that we should observe an increase in the Linguistic Proportion for both populations.

More specific to our research aims, we checked the degree to which laboratory and citizen science annotations converged at the child level by using correlations across the two. We additionally explored under what conditions Zooniverse judgments are more closely aligned with laboratory judgments using linear mixed models.

## Methods

All analyses and key data are available for reproduction on OSF (https://osf.io/57yha/). Computer code used to collect the Zooniverse judgments is available from GitHub (Semenzin, 2020).

### Participants

The full data set includes data from 20 children: 10 English-speaking children (6 males, 4 females; age range 11-53 months, mean=41.5 months) diagnosed with Angelman syndrome (AS) and 10 low-risk control (LRC) children (6 males, 4 females; age range 4-18 months, mean=11.7 months). As children diagnosed with Angelman syndrome typically have severe cognitive and language delays and most do not produce more than 1-2 words consistently, we compare their linguistic production data against that of younger children with theoretically similar language profiles.

### Equipment and data collection procedure

Recordings were obtained with the Language ENvironment Analysis (LENA^TM^) Digital Language Processor,  LENA^TM^ is a lightweight recorder (< 60 g, 5.5 cm x 8.5 cm x 1.5 cm) designed to be worn inside a breast-pocket of purpose-made clothing. In our data, every child contributed at least one daylong recording, with one child contributing two recordings (duration 11.58-16.00 hours, mean 15.05 hours).

### Data preprocessing (LENA^TM^)

The first step we took before beginning the annotation process was the identification of children's vocalizations within day-long recordings. We did this automatically, using the LENA^TM^ proprietary software. Specifically, the LENA^TM^ software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child’s environment (e.g., Female-Adult, TV). Key child labels have a precision of about XX% and a recall of about XX% (ref). These stretches of audio assigned to the key child can contain cries, vegetative sounds, and linguistic vocalizations, as well as interstitial pauses. We will refer to them as segments, to highlight the fact that they are stretches of audio segmented out as belonging to the key child, but they may not be individual vocalizations.

<!-- BT said Lisa add in how portions of recording were selected (high vol + random) -->

### Annotation by lab experts

Annotation was carried out at the segment level during lab coding. A subset of segments LENA^TM^ identified as the key child were selected for human annotation.  Segments were selected from 30 five-minute sections of the child’s recording: 10 sections during which the child was vocalizing at their highest rate, and 20 sections drawn randomly from the remainder of the recording. To ensure generalizability of our data, we included in this subset a number of segments of audio coming from both periods of high child volubility, as well as randomly selected periods of the day.

Annotations were carried out by 20 undergraduate students working in either Speech, Language, and Hearing Sciences or clinical psychology laboratories. Before annotation, coders completed an ethical conduct of research course and received HIPAA compliance training. After this, they complete a brief training on vocal maturity that involves 1) reading relevant literature, 2) reviewing examples of various annotation categories and 3) completing a brief quiz to test their annotation accuracy. Coders then proceed to the classification, where they make decisions on whether a segment is a cry, laugh, non-canonical syllable(s), canonical syllable(s), or a word. Coders are instructed to classify each segment based on the highest level of vocal maturity it contains (e.g., a segment with both canonical and non-canonical syllables would be classified as canonical; a segment with both laughing and non-canonical syllables would be classified as non-canonical). For canonical, non-canonical, and word segments, coders indicated how many syllables of each type of vocal maturity they hear (e.g., "ah ma ba" would be 1 non-canonical and 2 canonical syllables). Coders are also given a "Don’t Mark" option, which they are instructed to use if the segment does not sound like a segment made by the target child, or if there is any overlapping speech or other noise in the background which could affect acoustic analyses. For the purposes of the present study, we considered a "word" judgment to be equivalent to a "canonical" judgment. This may be false sometimes (e.g., "oh" may be coded as a real word by laboratory experts, although it is a non-canonical syllable), but we could not have a "word" judgment in our Zooniverse annotations because that requires context (which compromises privacy -- see next section). This process resulted in high levels of agreement: 97.22% of the lab-annotated segments had either 2 or 3 coders agree.

<!-- Reliability is built into the annotation routine such that each segment is coded by 3 separate coders, and only segments where 2 of 3 coders agree are used in subsequent phases of this study.  -->
<!-- Lisa aC removed this because this is actually false -- all segments were put up, including those with fewer than 2 agreement -->

<!-- Lisa, What happens if there are multiple syllables or multiple words? also, what if there is speech and non-speech, eg cry, in the same segment? -->

### Annotation on Zooniverse

A substantial concern that emerges when considering citizen science annotation for spontaneous speech data is the risk of a privacy breach: even short clips can contain personal information that can expose the identity of the speaker, such as addresses or names, or contain sensitive and private information. Fortunately, a study by Seidl et al. (2019) determined that when segments are divided into shorter chunks (400-600ms), human annotators can code our categories of interest (canonical, laughing, crying, etc.) with a classification quality comparable to the one carried out on full segments. We therefore used very short chunks (500ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.

Specifically, the segments extracted were automatically cut into clips of exactly 500ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform. To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms clips and their corresponding segment. Clips were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling (https://github.com/zooniverse/panoptes-cli).

Citizen scientists could access the project by logging in, or as anonymous participants. Before starting the annotation, participants were given a quick tutorial which walked them through the steps of the annotation workflow (i.e., how to play a sound, how to make a selection; see Figure 1). In this tutorial, we included one audio example for each category in order to make the classification task as smooth as possible. Further clarifications on what constitutes Canonical and Non-canonical utterances, as well as five audio examples for each of the categories, were available through the Field Guide, which users could access by clicking on the right side of the screen. The Field Guide could be consulted at any point of the classification without interrupting the task.

Users were asked to assign each 500ms clip to one out of five possible categories: (1) canonical, (2) non-canonical, (3) crying, (4) laughing, and (5) junk (overlapping speech, non-infant speech, silence, external sounds). Notice that the latter is narrower than the "Don't mark" category of the lab annotation routine, which was used any time there was noise or overlap that could affect acoustic analyses.

Following recommendations from the Zooniverse board, we collected 5 judgments per clip, rather than 3 as used in the laboratory. Majority agreement is thus achieved when 3 out of 5 judgements agree (rather than 2 out of 3). In exploratory analyses, we assess the impact of collecting a smaller number of judgments per clip.

``` {r fig-zoo, comment=F, message=F, hide = T, warning = F, echo = F,  fig.cap="Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needs to make; as well as the field guide, which remains accessible on the right of the screen throughout the session, and which allows users to revise examples of the different categories."}
knitr::include_graphics("zooniverse-pufig.pdf")
```

### Data post-processing

An impressive total of 4,825 individual Zooniverse users provided labels for the Maturity of Baby Sounds project, of which the present data set is one part. For this project, we collected a total of 169,767 judgments provided for 33,731 500-ms chunks, corresponding to 11,980 LENA^TM^ segments. Nearly a fifth of chunks did not have at least 3 labels in agreement out of the 5 Zooniverse labels (N = 6,585, 19% of all chunks). Of the chunks without a majority agreement, 4341 (66%) contained one or two Junk judgements (out of 5), 6523 (99,9%) had at least two matching judgements (the threshold used for lab-annotated segments), and only 61 (0,01%) had 5 different judgements. Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the 81% of chunks that did have at least 3 labels in agreement; this represented 135,725 labels for 27,145 chunks, corresponding to 11,593 LENA^TM^ segments. As the segments average 1.12 seconds in length, this means about 3.8 hours of audio data were annotated by 8 different annotators (3 in the laboratory, 5 on Zooniverse).

Lab annotators provided judgments at the level of LENA^TM^ segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments. Therefore, we needed to combine chunk-level judgments to reconstruct segment-level judgments.

We did this by considering the majority label extracted from Zooniverse judgements for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were junk, then the segment was labelled as junk; then, following a hierarchy (Canonical>Non-canonical>Laughing>Crying), segments with at least one instance of each judgement were labelled as such: e.g. if one canonical judgement was present, then the segment was labelled as canonical, otherwise if a non-canonical judgement was present, the segment was labelled as non-canonical, and so on. This is essentially the same thought process expert annotators followed when labeling segment-level audio, as each segment can represent multiple classification subtypes). LENA^TM^ segments are not equivalent to child vocalizations: A segment may contain 1 or more vocalizations, which may be of the same type or not.

<!-- AS said I feel like a figure might help here. -->

```{r, echo=F}
# remove non-majority labels
#TODO

# we map the mixed
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical"]<-"Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical_Crying"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Crying"]<-"Crying"
data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Crying"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing_Crying"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Canonical"]<-"Canonical"
# +
data_all$Zoon_classif[data_all$Zoon_classif=="Canonical_Crying"]<-"Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Canonical_Laughing"]<-"Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical_Crying"]<-"Non-Canonical"

data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Laughing"]<-"Crying"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Canonical_Laughing"]<-"Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Laughing_Non-Canonical"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Non-Canonical"]<-"Non-Canonical"
data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Non-Canonical_Laughing"]<-"Non-Canonical" 
data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing"]<-"Non-Canonical"

#and reset the factors for cleanliness
data_all$Zoon_classif=factor(data_all$Zoon_classif)
data_all$lab=factor(data_all$lab)
sample_data<-cbind(data_all$lab,data_all$Zoon_classif)

table_zoon=round(table(data_all$Zoon)/sum(table(data_all$Zoon))*100)
table_lab=round(table(data_all$lab)/sum(table(data_all$lab))*100)
```

```{r prepare data by child, echo=F}

#get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
ztab=table(data_all$filename,data_all$Zoon_classif)
z_lp=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
z_cp=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])

ltab=table(data_all$filename,data_all$lab)
l_lp=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
l_cp=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])

#and also get junk from zooniverse
z_junk=ztab[,c("Junk")]/rowSums(ztab)

#put all the proportions together
if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) proportions=cbind(rownames(ztab),z_lp,z_cp,l_lp,l_cp,z_junk) else print("oops this code needs to be more complex because we don't have the same kids for the two proportions")
colnames(proportions)[1]<-"filename"

#ages=aggregate(data_all$Age,by=list(data_all$ChildID),mean) #this is a weird way of adding ages, since all of the ages for a given child should be the same if there is only one recording, and if there are multiple recordings, then we should not get the mean
#improvement: now we merge with a demo data tab, but note this is merged with child id, so the problem of multiple recs per child is still there

# Created demo_data with filenames. Use filenames instead of childIDs to merge proportions and demo data.

merge(proportions,demo_data_fn,by="filename")->proportions
colnames(proportions)[dim(proportions)[2]]<-"Age"

#cbinding results in text, so we numerize the proportions
for(thisvar in c("z_lp","z_cp","l_lp","l_cp","z_junk")) proportions[,thisvar]=as.numeric(as.character(proportions[,thisvar]))

#add median junk
proportions$median_junk<-ifelse(proportions$z_junk<median(proportions$z_junk),"lower_junk","higher_junk")

#summary(proportions)

merge(data_all,proportions,all=T)->data_all
```
## Results

### Descriptive analyses

<!-- lh said I think you mention above that individuals with AS rarely use more than a handful of meaningful words - if we decide to go with just looking at these analyses broken down by group, that might be helpful to reiterate here -->
<!-- ac said in response: why do you talk about meaningful words? These analyses do not separate words from the rest... -->

In this section, we provide descriptive analyses of our dataset. According to lab annotators, `r table_lab["Canonical"]`% of segments were canonical, `r table_lab["Non-Canonical"]`% non-canonical, `r table_lab["Laughing"]`% laughing, and `r table_lab["Crying"]`% crying, with the remaining `r table_lab["Junk"]`% being categorized as "Don't mark". Zooniverse data revealed a similar distribution:  `r table_zoon["Canonical"]`% canonical, `r table_zoon["Non-Canonical"]`% non-canonical, `r table_zoon["Laughing"]`% laughing, `r table_zoon["Crying"]`% crying, `r table_zoon["Junk"]`% junk.
Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = ("Canonical"+"Non-Canonical")/"All vocalizations" (i.e., we remove, junk), and ii) Canonical Proportion = "Canonical"/("Canonical"+"Non-Canonical") (i.e., we remove junk + non-linguistic). See Figure 2 for results.

```{r fig-corage, echo=F,fig.cap="Correlations between child-level descriptors and age as a function of metric (linguistic ratio in the top row, canonical ratio in the bottom row), annotation method, and child group (red=low-risk, and black Angelman Syndrome)"}

prettynames=c("Linguistic Proportion (Zooniverse)","Canonical Proportion (Zooniverse)",
             "Linguistic Proportion (Lab)","Canonical Proportion (Lab)" )
names(prettynames)<-c("z_lp","z_cp","l_lp","l_cp")
mycols=c("black","red")
names(mycols)<-c("Low-RiskControl","AngelmanSyndrome")
mypch=c(4,20)
names(mypch)<-c("Low-RiskControl","AngelmanSyndrome")

lp_range=range(proportions[,c("z_lp","l_lp")])
cp_range=range(proportions[,c("z_cp","l_cp")])

#jpeg("../Results/corage.jpg",width=20,height=20,units="cm",res=300)
layout(matrix(c(1:4), 2, 2, byrow = F))
par(mar = c(4, 4, 1, 1))

#create matrix to hold r results
myr=matrix(NA,nrow=4,ncol=2)
rownames(myr)<-c("z_lp","z_cp","l_lp","l_cp")
colnames(myr)<-c("AS","LR")
mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices

for(thisvar in c("z_lp","z_cp","l_lp","l_cp")) {

  if(thisvar %in% c("z_lp","l_lp")) myrange=lp_range else myrange=cp_range

  plot(proportions[,thisvar]~proportions$Age, pch=mypch[proportions$Diagnosis],xlab="Age (months)",ylab=prettynames[thisvar],
       col=mycols[proportions$Diagnosis],
         ylim=myrange)
  abline(lm(proportions[,thisvar]~proportions$Age,subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black")

  #get cor properties
  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"])
    myr[thisvar,"AS"]=round(thiscor$estimate,3)
    mylb[thisvar,"AS"]=round(thiscor$conf.int[1],3)
    myhb[thisvar,"AS"]=round(thiscor$conf.int[2],3)
    myn[thisvar,"AS"]=round(thiscor$parameter,3)
    myp[thisvar,"AS"]=round(thiscor$p.value,3)

    #add r to figure
  text(mean(proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"]),mean(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar]),paste0("r=",myr[thisvar,"AS"]),col="black")

  #repeat for LR  
    abline(lm(proportions[,thisvar]~proportions$Age,subset=c(proportions$Diagnosis!="AngelmanSyndrome")),col="red")

      #get cor properties
  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"])
    myr[thisvar,"LR"]=round(thiscor$estimate,3)
    mylb[thisvar,"LR"]=round(thiscor$conf.int[1],3)
    myhb[thisvar,"LR"]=round(thiscor$conf.int[2],3)
    myn[thisvar,"LR"]=round(thiscor$parameter,3)
    myp[thisvar,"LR"]=round(thiscor$p.value,3)

    #add r to figure

    text(mean(proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"]),mean(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar]),paste0("r=",myr[thisvar,"LR"]),col="red")
}
#dev.off()
```

Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups. There was a near-zero relationship among the older children diagnosed with Angelman Syndrome r(`r myn["l_lp","AS"]`) = `r myr["l_lp","AS"]`, CI [`r mylb["l_lp","AS"]`,`r myhb["l_lp","AS"]`], p=`r myp["l_lp","AS"]`]; and a significant association among younger low-risk control children r(`r myn["l_lp","LR"]`) = `r myr["l_lp","LR"]`, CI [`r mylb["l_lp","LR"]`,`r myhb["l_lp","LR"]`], p=`r myp["l_lp","LR"]`]. The Canonical Proportion exhibited non-significant developmental decreases among older children diagnosed with Angelman Syndrome r(`r myn["l_cp","AS"]`) = `r myr["l_cp","AS"]`, CI [`r mylb["l_cp","AS"]`,`r myhb["l_cp","AS"]`], p=`r myp["l_cp","AS"]`]; and marginal developmental increases among low-risk control r(`r myn["l_cp","LR"]`) = `r myr["l_cp","LR"]`, CI [`r mylb["l_cp","LR"]`,`r myhb["l_cp","LR"]`], p=`r myp["l_cp","LR"]`].

Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman Syndrome r(`r myn["z_lp","AS"]`) = `r myr["z_lp","AS"]`, CI [`r mylb["z_lp","AS"]`,`r myhb["z_lp","AS"]`], p=`r myp["z_lp","AS"]`]; whereas younger low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r myr["z_lp","LR"]`, CI [`r mylb["z_lp","LR"]`,`r myhb["z_lp","LR"]`], p=`r myp["z_lp","LR"]`]. Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman Syndrome r(`r myn["z_cp","AS"]`) = `r myr["z_cp","AS"]`, CI [`r mylb["z_cp","AS"]`,`r myhb["z_cp","AS"]`], p=`r myp["z_cp","AS"]`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r myr["z_cp","LR"]`, CI [`r mylb["z_cp","LR"]`,`r myhb["z_cp","LR"]`], p=`r myp["z_cp","LR"]`].

## Main analyses

Next, we discuss the correspondence between citizen science classifications and the laboratory gold standard, at the level of individual clips. Results were visualized with a confusion matrix showing precision and recall (Figure 3): the diagonal elements show the number of correct segment-level classifications for each class while the off-diagonal elements show non-matching classifications.

```{r mygac-conf, echo=F}
mygac=gac(data = sample_data, kat = 5, weight = c("unweighted"),
    conf.level = 0.95)

mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
# this package uses sensitivity & specificity
#Sensitivity=recall
#Specificity=precision
#mycf
```

```{r fig-prec-rec, echo=F}

colsums=colSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:dim(conf_tab)[2]) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
#colSums(my_conf_tab) #check
prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"pr"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","pr")])
prec_plot = ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(pr)), colour = "white") +
  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=4) +
  geom_text(aes(label = Freq), vjust = 1,size=4) +
  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# repeat for recall
rowsums=rowSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:dim(conf_tab)[2]) my_conf_tab[,i]=my_conf_tab[,i]/rowsums[i]
#rowSums(my_conf_tab)
prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"rec"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","rec")])
rec_plot = ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(rec)), colour = "white") +
  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=4) +
  geom_text(aes(label = Freq), vjust = 1,size=4) +
  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Recall")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggarrange(prec_plot, rec_plot, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)

```

These visualizations suggest that performance is moderate to good, which was confirmed via statistical analyses. The overall (weighted) accuracy is `r round(mycf$overall["Accuracy"]*100)`%, CI = [`r round(mycf$overall["AccuracyLower"]*100)`,`r round(mycf$overall["AccuracyUpper"]*100)`],  kappa is `r round(mycf$overall["Kappa"],3)`, and the Gwet's AC1 coefficient is `r round(mygac$est,3)`, CI = [`r round(mygac$lb,3)`,`r round(mygac$ub,3)`].

### Child level descriptors

```{r corlab-zoo, echo=F}

layout(matrix(c(1:2), 1, 2, byrow = F))
par(mar = c(4, 4, 1, 1))

#create matrix to hold r results
myr=matrix(NA,nrow=2,ncol=3)
rownames(myr)<-c("lp","cp")
colnames(myr)<-c("AS","LR","all")
mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices

lims=range(c(proportions[,"z_lp"],proportions[,"l_lp"]))

  plot(proportions[,"z_lp"]~proportions[,"l_lp"], xlab=prettynames["l_lp"],ylab=prettynames["z_lp"],
       xlim=lims,ylim=lims,
       col=mycols[proportions$Diagnosis], pch=mypch[proportions$Diagnosis])
  lines(c(0,1),c(0,1),lty=2,col="darkgray")

  abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black")

  #get cor properties
  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_lp"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
    myr[1,"AS"]=round(thiscor$estimate,3)
    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
    myn[1,"AS"]=round(thiscor$parameter,3)
    myp[1,"AS"]=round(thiscor$p.value,3)

  #repeat for LR  
    abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$Diagnosis!="AngelmanSyndrome")),col="red")

      #get cor properties
  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"],proportions[proportions$Diagnosis!="AngelmanSyndrome","z_lp"])
    myr[1,"LR"]=round(thiscor$estimate,3)
    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
    myn[1,"LR"]=round(thiscor$parameter,3)
    myp[1,"LR"]=round(thiscor$p.value,3)

#repeat for all cor

      thiscor=cor.test(proportions$z_lp,proportions$l_lp)
    myr[1,"all"]=round(thiscor$estimate,3)
    mylb[1,"all"]=round(thiscor$conf.int[1],3)
    myhb[1,"all"]=round(thiscor$conf.int[2],3)
    myn[1,"all"]=round(thiscor$parameter,3)
    myp[1,"all"]=round(thiscor$p.value,3)

    # repeat whole for cp

    lims=range(c(proportions[,"z_cp"],proportions[,"l_cp"]))

  plot(proportions[,"z_cp"]~proportions[,"l_cp"], xlab=prettynames["l_cp"],ylab=prettynames["z_cp"],
       xlim=lims,ylim=lims,
       col=mycols[proportions$Diagnosis], pch=mypch[proportions$Diagnosis])
  lines(c(0,1),c(0,1),lty=2,col="darkgray")

  abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black")

  #get cor properties
  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_cp"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
    myr[2,"AS"]=round(thiscor$estimate,3)
    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
    myn[2,"AS"]=round(thiscor$parameter,3)
    myp[2,"AS"]=round(thiscor$p.value,3)

  #repeat for LR  
    abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$Diagnosis!="AngelmanSyndrome")),col="red")

      #get cor properties
  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"],proportions[proportions$Diagnosis!="AngelmanSyndrome","z_cp"])
    myr[2,"LR"]=round(thiscor$estimate,3)
    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
    myn[2,"LR"]=round(thiscor$parameter,3)
    myp[2,"LR"]=round(thiscor$p.value,3)

#repeat for all cor

      thiscor=cor.test(proportions$z_cp,proportions$l_cp)
    myr[2,"all"]=round(thiscor$estimate,3)
    mylb[2,"all"]=round(thiscor$conf.int[1],3)
    myhb[2,"all"]=round(thiscor$conf.int[2],3)
    myn[2,"all"]=round(thiscor$parameter,3)
    myp[2,"all"]=round(thiscor$p.value,3)

mycortab=t(myr)
rownames(mycortab)<-paste("All seg",rownames(mycortab))
```

Although the classification at the clip level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of childrens’ individual development. Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.  When we split by participant group, correlations remain high: for Angelman Syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].

Similarly, a strong positive correlation is found in the Canonical Proportion r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`]. When we split by participant group, correlations remain high although we do note they are somewhat smaller for the older children with Angelman Syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.

```{r glmer-agediag, echo=F}
# one outlier in ling prop is removed

ling_all=glmer(lab_ling~zoo_ling*Diagnosis*Age + (1|ChildID),data=data_all,family="binomial",subset=c(ChildID!="3681_1"))
#plot(lin_mod)
# doesn't converge

ling_AS=glmer(lab_ling~zoo_ling*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis=="AngelmanSyndrome"))
#only effect of zoo

ling_LR=glmer(lab_ling~zoo_ling*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis!="AngelmanSyndrome" & ChildID!="3681_1"))
#both main and interaction are sig

can_all=glmer(lab_can~zoo_can*Diagnosis*Age + (1|ChildID),data=data_all,family="binomial")
#doesn't converge either

can_AS=glmer(lab_can~zoo_can*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis=="AngelmanSyndrome"))
#only effect of zoo

can_LR=glmer(lab_can~zoo_can*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis!="AngelmanSyndrome"))
#both main and interaction are sig

```

## Additional analyses

We additionally explored under what conditions Zooniverse judgments more closely aligned with laboratory judgments. In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as "Junk" (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora [@cychosz2019canonical]. A high proportion of "Junk" may indicate that automated segmentation was errorful for those children, and may be a sign that the rest of the data could be compromised as well.

```{r prelim-junk, echo=F}
junkXdiag=t.test(proportions$z_junk ~ proportions$Diagnosis)

```

```{r cor-junk, echo=F}

mycols=c("blue","darkgreen")
mypch=c(10,18)
names(mycols)=names(mypch)=c("lower_junk","higher_junk")

layout(matrix(c(1:2), 1, 2, byrow = F))
par(mar = c(4, 4, 1, 1))

#create matrix to hold r results
myr=matrix(NA,nrow=2,ncol=2)
rownames(myr)<-c("lp","cp")
colnames(myr)<-c("lower_junk","higher_junk")
mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices

lims=range(c(proportions[,"z_lp"],proportions[,"l_lp"]))

  plot(proportions[,"z_lp"]~proportions[,"l_lp"], xlab=prettynames["l_lp"],ylab=prettynames["z_lp"],
       xlim=lims,ylim=lims,
       col=mycols[proportions$median_junk], pch=mypch[proportions$median_junk])
  lines(c(0,1),c(0,1),lty=2,col="darkgray")

  abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$median_junk=="lower_junk")),col="blue")

  #get cor properties
  thiscor=cor.test(proportions[proportions$median_junk=="lower_junk","z_lp"],proportions[proportions$median_junk=="lower_junk","l_lp"])
    myr[1,"lower_junk"]=round(thiscor$estimate,3)
    mylb[1,"lower_junk"]=round(thiscor$conf.int[1],3)
    myhb[1,"lower_junk"]=round(thiscor$conf.int[2],3)
    myn[1,"lower_junk"]=round(thiscor$parameter,3)
    myp[1,"lower_junk"]=round(thiscor$p.value,3)

  #repeat for LR  
    abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$median_junk!="lower_junk")),col="darkgreen")

      #get cor properties
  thiscor=cor.test(proportions[proportions$median_junk!="lower_junk","l_lp"],proportions[proportions$median_junk!="lower_junk","z_lp"])
    myr[1,"higher_junk"]=round(thiscor$estimate,3)
    mylb[1,"higher_junk"]=round(thiscor$conf.int[1],3)
    myhb[1,"higher_junk"]=round(thiscor$conf.int[2],3)
    myn[1,"higher_junk"]=round(thiscor$parameter,3)
    myp[1,"higher_junk"]=round(thiscor$p.value,3)

    # repeat whole for cp

    lims=range(c(proportions[,"z_cp"],proportions[,"l_cp"]))

  plot(proportions[,"z_cp"]~proportions[,"l_cp"], xlab=prettynames["l_cp"],ylab=prettynames["z_cp"],
       xlim=lims,ylim=lims,
       col=mycols[proportions$median_junk], pch=mypch[proportions$median_junk])
  lines(c(0,1),c(0,1),lty=2,col="darkgray")

  abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$median_junk=="lower_junk")),col="blue")

  #get cor properties
  thiscor=cor.test(proportions[proportions$median_junk=="lower_junk","z_cp"],proportions[proportions$median_junk=="lower_junk","l_cp"])
    myr[2,"lower_junk"]=round(thiscor$estimate,3)
    mylb[2,"lower_junk"]=round(thiscor$conf.int[1],3)
    myhb[2,"lower_junk"]=round(thiscor$conf.int[2],3)
    myn[2,"lower_junk"]=round(thiscor$parameter,3)
    myp[2,"lower_junk"]=round(thiscor$p.value,3)

  #repeat for higher_junk  
    abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$median_junk!="lower_junk")),col="darkgreen")

      #get cor properties
  thiscor=cor.test(proportions[proportions$median_junk!="lower_junk","l_cp"],proportions[proportions$median_junk!="lower_junk","z_cp"])
    myr[2,"higher_junk"]=round(thiscor$estimate,3)
    mylb[2,"higher_junk"]=round(thiscor$conf.int[1],3)
    myhb[2,"higher_junk"]=round(thiscor$conf.int[2],3)
    myn[2,"higher_junk"]=round(thiscor$parameter,3)
    myp[2,"higher_junk"]=round(thiscor$p.value,3)

```

We investigated this hypothesis by calculating the proportion of their data labeled as "Junk" for each individual child.  There was no significant difference in the proportions of their data labeled as "Junk" for older children with Angelman Syndrome (M = `r round(junkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the younger low risk children (M = `r round(junkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r junkXdiag$parameter`) = `r round(junkXdiag$statistic,3)`, p = `r round(junkXdiag$p.value,3)`. We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk". Results were similar across these two post-hoc subgroups. For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower junk group was r(`r myn["lp","lower_junk"]`) = `r myr["lp","lower_junk"]`, CI [`r mylb["lp","lower_junk"]`,`r myhb["lp","lower_junk"]`], p=`r myp["lp","lower_junk"]`]; and for the higher junk group it was r(`r myn["lp","higher_junk"]`) = `r myr["lp","higher_junk"]`, CI [`r mylb["lp","higher_junk"]`,`r myhb["lp","higher_junk"]`], p=`r myp["lp","higher_junk"]`. For Canonical Proportion, the correlation across lab and Zooniverse data for the lower junk group was r(`r myn["cp","lower_junk"]`) = `r myr["cp","lower_junk"]`, CI [`r mylb["cp","lower_junk"]`,`r myhb["cp","lower_junk"]`], p=`r myp["cp","lower_junk"]`]; and for the higher junk group it was r(`r myn["cp","higher_junk"]`) = `r myr["cp","higher_junk"]`, CI [`r mylb["cp","higher_junk"]`,`r myhb["cp","higher_junk"]`], p=`r myp["cp","higher_junk"]`. Thus, it does not seem that a higher proportion of "Junk" judgments is an index of low quality data.

Above, we concluded that derived metrics seem more promising than segment-level data. Notice that derived metrics do not require matching of clips (500 ms presented to Zooniverse participants) to segments (the original LENA^TM^ segments presented to laboratory participants). As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment. We therefore repeated our analyses but deriving our proportions for the Zooniverse data not from the segment-level composite, but rather the individual chunk-level annotations.

```{r bychunk-age, echo=F}
read.csv("../Derived_Data/chunk_majority_agreement_1751.csv")->chunk
bychild=table(chunk$Answer,chunk$ChildID)
cans=bychild["Canonical",]
noncans=bychild["Non-Canonical",]
junk=bychild["Junk",]
sum_bychild=table(chunk$ChildID)

ling_prop=(cans+noncans)/(sum_bychild-junk)
can_prop=cans/(cans+noncans)

#check that they match before merging
#demo_data$ChildID==names(ling_prop)
#demo_data$ChildID==names(can_prop)

props_bychild_chunk=cbind(ling_prop,can_prop,demo_data)
colnames(props_bychild_chunk)[1:2]<-c("chi","ling_prop")

z_lp_as=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])

z_lp_lr=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])

z_cp_as=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])

z_cp_lr=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
```

Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman Syndrome group: r(`r z_lp_as$parameter`) = `r z_lp_as$estimate`, CI [`r z_lp_as$conf.int[1]`,`r z_lp_as$conf.int[2]`], p=`r z_lp_as$p.value`], but increased with age  in the low-risk control group: r(`r z_lp_lr$parameter`) = `r z_lp_lr$estimate`, CI [`r z_lp_lr$conf.int[1]`,`r z_lp_lr$conf.int[2]`], p=`r z_lp_lr$p.value`. Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman Syndrome r(`r z_cp_as$parameter`) = `r z_cp_as$estimate`, CI [`r z_cp_as$conf.int[1]`,`r z_cp_as$conf.int[2]`], p=`r z_cp_as$p.value`]; and developmental increases among low-risk control r(`r z_cp_lr$parameter`) = `r z_cp_lr$estimate`, CI [`r z_cp_lr$conf.int[1]`,`r z_cp_lr$conf.int[2]`], p=`r z_cp_lr$p.value`.

```{r zoo-lab-chunk, echo=F}

#create matrix to hold r results
myr=matrix(NA,nrow=2,ncol=3)
rownames(myr)<-c("lp","cp")
colnames(myr)<-c("AS","LR","all")
mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices

proportions=proportions[order(proportions$ChildID),]

  #get cor properties
  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
    myr[1,"AS"]=round(thiscor$estimate,3)
    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
    myn[1,"AS"]=round(thiscor$parameter,3)
    myp[1,"AS"]=round(thiscor$p.value,3)

  #repeat for LR

      #get cor properties
  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"])
    myr[1,"LR"]=round(thiscor$estimate,3)
    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
    myn[1,"LR"]=round(thiscor$parameter,3)
    myp[1,"LR"]=round(thiscor$p.value,3)

#repeat for all cor

      thiscor=cor.test(props_bychild_chunk$ling_prop,proportions$l_lp)
    myr[1,"all"]=round(thiscor$estimate,3)
    mylb[1,"all"]=round(thiscor$conf.int[1],3)
    myhb[1,"all"]=round(thiscor$conf.int[2],3)
    myn[1,"all"]=round(thiscor$parameter,3)
    myp[1,"all"]=round(thiscor$p.value,3)

    # repeat whole for cp

  #get cor properties
  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
    myr[2,"AS"]=round(thiscor$estimate,3)
    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
    myn[2,"AS"]=round(thiscor$parameter,3)
    myp[2,"AS"]=round(thiscor$p.value,3)

  #repeat for LR

      #get cor properties
  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"])
    myr[2,"LR"]=round(thiscor$estimate,3)
    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
    myn[2,"LR"]=round(thiscor$parameter,3)
    myp[2,"LR"]=round(thiscor$p.value,3)

#repeat for all cor

      thiscor=cor.test(props_bychild_chunk$can_prop,proportions$l_cp)
    myr[2,"all"]=round(thiscor$estimate,3)
    mylb[2,"all"]=round(thiscor$conf.int[1],3)
    myhb[2,"all"]=round(thiscor$conf.int[2],3)
    myn[2,"all"]=round(thiscor$parameter,3)
    myp[2,"all"]=round(thiscor$p.value,3)

 mycortab=rbind(mycortab,t(myr))
rownames(mycortab)[4:6]<-paste("Chunks",rownames(t(myr)))

```

As for correlations between Zooniverse and laboratory-derived metrics, we observe very similar levels of correlation: Linguistic Proportion overall r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.  When we split by participant group, correlations remain high: for Angelman Syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`]. As for Canonical Proportion overall r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`; Angelman Syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`];  low-risk control infants r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.

Since all of these results are very similar to those obtained when first matching clips to segments, we conclude that in the future this step may not be necessary. Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' clip level judgements [which was in fact what @cychosz2019canonical did].

```{r downsample}

#we start from the data base that has all the info
nsegs=data.frame(table(data_all$ChildID))
colnames(nsegs)<-c("ChildID","nsegs")

merge(nsegs,demo_data)->nsegs
nseg_t=t.test(nsegs$nsegs~nsegs$Diagnosis)

resampling=matrix(NA,nrow=50,ncol=6)
colnames(resampling)<-c("lp_r_all","lp_r_as","lp_r_lr","cp_r_all","cp_r_as","cp_r_lr")
for(i in 1:50){
  sel_seg=NULL
  for(thischi in levels(data_all$ChildID)) sel_seg<-c(sel_seg,as.character(data_all$segmentId_DB[sample(data_all$segmentId_DB[data_all$ChildID==thischi],100)]))

  sampling=data_all[as.character(data_all$segmentId_DB) %in% as.character(sel_seg),]

  #get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
  ztab=table(sampling$ChildID,sampling$Zoon_classif)
  z_lp=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
  z_cp=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])

  ltab=table(sampling$ChildID,sampling$lab)
  l_lp=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
  l_cp=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])

  thistab=cbind(demo_data,z_lp,z_cp,l_lp,l_cp)

  resampling[i,"lp_r_all"]=cor.test(thistab$z_lp,thistab$l_lp)$estimate
  resampling[i,"lp_r_as"]=cor.test(thistab$z_lp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
  resampling[i,"lp_r_lr"]=cor.test(thistab$z_lp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate

   resampling[i,"cp_r_all"]=cor.test(thistab$z_cp,thistab$l_cp)$estimate
  resampling[i,"cp_r_as"]=cor.test(thistab$z_cp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
  resampling[i,"cp_r_lr"]=cor.test(thistab$z_cp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate
 }
myr=colMeans(resampling)
mycortab=rbind(mycortab,cbind(myr[c("lp_r_as" ,"lp_r_lr","lp_r_all")],myr[c("cp_r_as" ,"cp_r_lr","cp_r_all")]))
rownames(mycortab)[7:9]<-c("100 seg AS","100 seg LR","100 seg all")
round(mycortab,3)->mycortab

resampling=data.frame(resampling)
```

Next, we looked at whether having more segments from each child may lead to more reliable metrics.  We observed a non-significant trend [t(`r nseg_t$parameter`)=`r nseg_t$estimate`] for lower number of segments in the Angelman Syndrome group (mean `r nseg_t$estimate[1]`, range `r min(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`) than among low-risk control infants (mean `r nseg_t$estimate[2]`, range `r min(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`), likely because of the lower volubility of the former children. Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications [CITE MEG]. We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children [following CITE MEG]. We randomly sampled 100 segments from each child's data, and recalculated the linguistic and Canonical Proportions based only on these associated judgments. We repeated this process 50 times, to assess the extent to which the association between laboratory- and Zooniverse-derived metrics varied in each random sample. The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table XX), particularly for the low-risk younger infants, and for Linguistic Proportion. This may indicate that, particularly in some groups of infants, 100 segments may not be sufficient to capture the child's linguistic and Canonical Proportions. We also observed quite a bit of variance across the 50 runs. For Linguistic Proportion, the range of correlations found for all infants was [`r min(resampling$lp_r_all)`,`r max(resampling$lp_r_all)`]; for the Angelman Syndrome group [`r min(resampling$lp_r_as)`,`r max(resampling$lp_r_as)`]; for the low-risk group [`r min(resampling$lp_r_lr)`,`r max(resampling$lp_r_lr)`].  For Canonical Proportion, the range of correlations found for all infants was [`r min(resampling$cp_r_all)`,`r max(resampling$cp_r_all)`]; for the Angelman Syndrome group [`r min(resampling$cp_r_as)`,`r max(resampling$cp_r_as)`]; for the low-risk group [`r min(resampling$cp_r_lr)`,`r max(resampling$cp_r_lr)`].

```{r tab-cors, results="as.is"}
colnames(mycortab)<-c("Linguistic Proportion","Canonical Proportion")
kable(mycortab,cap="Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman Syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways: All seg indicates that Zooniverse annotations at the clip level were first combined at the segment level; Chunks that they were analyzed directly at the clip level. Both of these are based on all the data. In contrast, rows labeled 100 seg indicate that laboratory- and Zooniverse-derived metrics were based on only 100 segments (median over 50 runs in which 100 segments were randomly selected from each child).")
```

<!-- - look at whether kids who have a high "junk" proportion according to Zooniverse  -->
<!--      - have lower levels of concordance across Zoo coders -->

<!-- - take all 5 Zoo judgments or 3 or 1 (may need to run the latter two a couple of times to check for potential variance -- if variance is small, we'll just say so, if it's large then we'll have to write a loop and report CIs) -->

## Discussion

In the present study we assessed the extent to which child vocalizations from LENA^TM^ daylong recordings can be accurately described based on classifications collected using the citizen science platform Zooniverse. Our recordings came from children diagnosed with a language disorder, as well as low-risk controls, and included a range of different ages. Excerpts of these recordings had previously been annotated by trained experts in the lab, and served as the "gold standard".

In descriptive terms, we found that category frequency was similar across laboratory and citizen science annotations. Correlations between age and our two child-level metrics revealed different developmental patterns between our sample of older children with a diagnosis of Angelman Syndrome and younger low-risk controls. Given that the two groups of children differ on both of these features (age and diagnosis), we cannot empirically tease apart these two differences. Nonetheless, inspection of Figure 2 suggests different interpretations for the two derived metrics. Overall, the pattern for Linguistic Proportion is compatible with a non-linear trajectory common to the two groups, whereby the Linguistic Proportion increases rapidly up to 20 months, and hovers around 90-100% thereafter. In this case, the lack of correlation with age in the Angelman Syndrome group suggests that there is little systematic variance in this metric after 20 months of age. In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and slower decreases among children in the Angelman Syndrome group.

Next, we looked at agreement between laboratory and Zooniverse annotation at the level of individual segments. We found that some categories had high precision and recall -- notably Non-canonical segments were detected quite accurately, and to a lesser extent the same was true for Canonical segments. This is interesting because it suggests that Zooniverse annotations may be trusted particularly for detecting these two types of speech-like segments.

In other cases, precision was fairly high but recall was lower. This was the case of the "Junk/Don't mark" category. Higher precision and low recall indicates that when Zooniverse annotations return a "Junk" judgement, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as "Don't mark" in the laboratory. Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the "junk" category less) than laboratory annotators, who have access to the whole vocalization.  It is also plausible, on the other hand, that lab annotators are more "picky" with the data, and tend to reject clips (tag them as "Don't mark"). Indeed, laboratory coders were instructed to be  sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use "Don't mark" if there was any noise or verlapping speech that would potentially affect acoustic analyses.

Both Crying and Laughing showed moderate to good recall, but low precision. This means that a high proportion of segments detected as these two categories in the lab can be classified in the appropriate category by Zooniverse annotators, but Zooniverse annotators also put into these categories many segments that lab annotators classified in a different manner. Most saliently, there is a problematic level of confusion between Crying and Non-Canonical, whereby lab-detected Crying was almost as likely to be classified as such or Non-canonical in Zooniverse (the recall proportion was 49 and 47% respectively), and even worse, 55% of the segments Zooniverse annotators classified as Crying were actually Non-Canonical according to lab annotations. This is likely due to the fact that lab annotators had access to the whole segment, and could even listen to the context of that segment, so they could accurately interpret a segment as crying. Thus, it is possible that additional analyses, using for instance pitch or duration characteristics at the level of the whole segment rather than the clip could improve classification performance, but for the time being, we cannot recommend Zooniverse annotations to identify crying bouts.

As for laughing, the recall is quite good but precision is low with confusion involving Junk and Non-Canonical to similar extents. We are tentative in our interpretation of this result because the category was very sparsely represented, with barely 200 segments classified as Laughing in the whole data set. We thus recommend additional work, although we expect that perhaps there will be a similar problem (and solution) as indicated for Crying, whereby determination of Laughing versus Non-Canonical will require inspection of segment-level properties.

Turning now to the child-level analyses, we were particularly interested in derived metrics because this is more typically the goal of such analyses, for instance to study potential individual or group differences, or investigate the impact of an intervention. Our results show that, despite agreement variation at the level of the segment, derived metrics were highly correlated across laboratory and Zooniverse data, meaning that variability in precision and recall did not substantially impact key outcome variables.  Errors only seemed systematic for the Linguistic Proportion (and not for Canonical Proportion), and they were due to lower Linguistic Proportions from Zooniverse than laboratory data. This is uncertainly due to the confusion between Crying and Non-Canonical, with a substantial part of Crying segments being classified as Non-Canonical by Zooniverse annotators. Nonetheless, since these errors were systematic across infants, they do not affect the study of group or individual variation.

Our exploratory analyses then dug into some methodological aspects of our data. We found that, although individuals vary widely in terms of what proportion of their data is classified as Junk by Zooniverse annotators, this is not a sign that their data as a whole cannot be trusted, with similar levels of correlations between laboratory- and Zooniverse-derived metrics across median groups based on proportion of their data classified as Junk.

Next, we found that more data led to more reliable measurements: Specifically, our correlations based on 300-500 segments per child were markedly higher than those based on 100 segments per child, particularly when looking at our low-risk, younger participants. It is important to note that an advantage of Zooniverse, compared to laboratory annotations, is that it is easier to scale up: More data from each individual child can be annotated this way, which will help countering noise.

In sum, we believe these data show that young children's vocalizations can be faithfully described by relying on the judgments of Zooniverse's citizen scientists, particularly when the goal is to describe vocalizations at a broader level than individual clips. It is important to note that from a methodological point of view, the use of crowdsourcing is not merely a question of being an ecological technique, but it can also offer benefits in the area of scientific rigor and reproducibility. As previously mentioned in the Introduction section, Zooniverse offers both an automated system of data handling as well as an enormous pool of users: access to a large-scale group of blinded participants can in itself help researchers overcome the methodological compromises that sometimes lead to biased results. Additionally, crowd-sourcing and open-science frameworks can come together to encourage replication of existing work: open-source software, shared data and citizen science platforms would make it possible, and actually easy, to run identical studies and evaluate research results independently.

### Further research directions

These findings open up a new, exciting avenue for future research. Our pipeline to upload data on Zooniverse is scalable and quickly adapted to new datasets, including data on infants of different ages, learning different languages, or living in environments with different acoustic properties (e.g., rural versus urban environments). This means large existing datasets of daylong recordings can be reliably annotated in an ecological way using citizen science (ref on existing datasets).

We believe additional research would be welcome on the two measures we used to describe children’s linguistic productions: Linguistic Proportion and Canonical Proportion. Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion. We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant for other disorders that are related to emotional disorders (REF REF). Additionally, while Canonical Proportion has been studied in much previous work, we believe further research is necessary using larger scales (e.g., sampling from daylong recordings to have a more ecological picture of the child, REF; increasing the sheer number of children studied, to have sufficient power to detect effects) and with broader coverage (notably including children growing up in a broad range of languages and cultures, REF), to make sure that assumptions regarding cross-linguistic and maturational universality are justified (REF).

Assuming these child-level metrics continue to be supported by such additional data, it would be interesting for future investigations to expand the analysis of Linguistic Proportion and Canonical Proportion to recover differences and similarities in language development between low-risk children and children with disorders other than Angelman Syndrome. An extensive body of work has found that children with autism spectrum disorders experience language delays (Fasolo, Majorano, & D’Odorico, 2008; Lang et al., 2019; Patten et al., 2014; Stoel-Gammon, 1989). One caveat is that these studies used CBO, which requires recordings to be gathered periodically in order to be determined.

It is also important to note that work on non-typically developing children poses in itself some challenges. Our data contained a lower number of vocalizations for children diagnosed with Angelman Syndrome, as they tend to produce fewer vocalizations than low-risk controls REF. Notice this was the case even in our data, where we selected much younger low-risk control childrens as the comparison point.

<!-- AC removed: Additionally, we found an imbalance between the quantity of clips assigned to each category. Further research should look at validating the results of our approach by oversampling data from non-typically developing children, so as to overcome both the risk of lack of data in some categories, and the imbalance in the quantity of vocalizations between groups. -->

The lack of age effects in AS seems like the bigger "challenge" - if the goal is to use this information to inform how children are developing, are these the right metrics for that population? Longitudinal work within AS may be needed to track the "natural history" of vocalizations to better contextualize how these metrics might be relevant or useful. For example, there is interest in using naturalistic recordings as outcome measures for clinical/drug trials. Are we seeing enough variability in AS to assume that these metrics could "budge" in response to a treatment? It seems we are seeing variability across kids, which is promising. We are also seeing agreement across coders in AS files, which is also important. Next questions would be whether these metrics change within individuals, and whether rates of change may be a candidate for tracking treatment. Graphs that show the two key variables across groups and age may help inform these questions.

Finally, broader generalizations of our technique could be of interest to readers of this paper. Any researcher is able to undergo the procedure to create Zooniverse projects. For instance, a separate pipeline could be set up to annotate potentially disordered speech by older adults at risk of, or diagnosed with, neurodegenerative diseases. Although it would be important to similarly validate such annotations, we believe there is strong promise for such tasks, because many neurodegenerative disorders affect local aspects of the speech, which are detected at the syllable level. In contrast, tasks that require more speech or audio context may not be well suited to citizen science platforms, because they would require playing longer clips, which may reveal identifying or sensitive information. Although citizen scientists themselves are well-intentioned, the platform could be used by others who want to exploit the system for other means.

## Conclusion

In this study we validated the quality of annotations obtained through a citizen science platform, Zooniverse, as compared to a gold standard of human expert annotators. We analyzed the correspondence between annotations at the individual clip level, and the individual-child level, using Canonical Proportion and Linguistic Proportion as descriptors. We found a moderate to good accuracy in the first, and strong positive correlations in the latter. We can conclude that citizen scientists are a reliable source of fast and ecological annotation of speech data, particularly when results are combined into child-level descriptors. Our study suggests that the same methodology may be applied to several research questions in the study of language acquisition and language disorders. This finding is particularly welcome in an era when wearables open new avenues for studying human behavior and development in an ecological manner.

\newpage

# Acknowledgements

We are grateful to ...

\newpage

MOVE FIGURE 1 HERE

\newpage

MOVE FIGURE 2 HERE

etc

# References

<!-- These lines ensure references are set with hanging indents in PDF documents; they are ignored in Word. -->
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

### 2020-10-28 202010_jslhr/paper.tex
paper first full version and it knits

--- 2020-10-28 202010_jslhr/paper.Rmd
@@
----
-title: "Describing vocalizations in young children: A big data approach through citizen science annotations"
-shorttitle        : "Children's vocalizations and citizen science"
+\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[english,,man,floatsintext]{apa6}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provides euro and other symbols
+\else % if luatex or xelatex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage[]{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\usepackage{hyperref}
+\hypersetup{
+            pdftitle={Describing vocalizations in young children: A big data approach through citizen science annotations},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
 
-author: 
-  - name          : "Chiara Semenzin"
-    affiliation   : "1"
-    corresponding : no
-  - name          : "Lisa Hamrick"
-    affiliation   : "2"
-    corresponding : no
-  - name          : "Amanda Seidl"
-    affiliation   : "2"
-    corresponding : no
-  - name          : "Bridgette Kelleher"
-    affiliation   : "2"
-    corresponding : no
-  - name          : "Alejandrina Cristia"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "29 rue d'Ulm, 75005, Paris, France"
-    email         : "alecristia@gmail.com"
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 
-affiliation:
-  - id            : "1"
-    institution   : "Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University"
-  - id            : "2"
-    institution   : "Purdue University"
+% Manuscript styling
+\usepackage{upgreek}
+\captionsetup{font=singlespacing,justification=justified}
 
-author_note: >
-       AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017); and the J.
+% Table formatting
+\usepackage{longtable}
+\usepackage{lscape}
+% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
+\usepackage{multirow}		% Table styling
+\usepackage{tabularx}		% Control Column width
+\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
+\usepackage{threeparttablex}            % Lets threeparttable work with longtable
+
+% Create new environments so endfloat can handle them
+% \newenvironment{ltable}
+%   {\begin{landscape}\begin{center}\begin{threeparttable}}
+%   {\end{threeparttable}\end{center}\end{landscape}}
+\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
+
+% Enables adjusting longtable caption width to table width
+% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
+\makeatletter
+\newcommand\LastLTentrywidth{1em}
+\newlength\longtablewidth
+\setlength{\longtablewidth}{1in}
+\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
+
+% \setlength{\parindent}{0.5in}
+% \setlength{\parskip}{0pt plus 0pt minus 0pt}
+
+% \usepackage{etoolbox}
+\makeatletter
+\patchcmd{\HyOrg@maketitle}
+  {\section{\normalfont\normalsize\abstractname}}
+  {\section*{\normalfont\normalsize\abstractname}}
+  {}{\typeout{Failed to patch abstract.}}
+\makeatother
+\shorttitle{Children's vocalizations and citizen science}
+\author{Chiara Semenzin\textsuperscript{1}, Lisa Hamrick\textsuperscript{2}, Amanda Seidl\textsuperscript{2}, Bridgette Kelleher\textsuperscript{2}, \& Alejandrina Cristia\textsuperscript{1}}
+\affiliation{
+\vspace{0.5cm}
+\textsuperscript{1} Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University\\\textsuperscript{2} Purdue University}
+\authornote{AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017); and the J.
 S.
@@
 
-# Abbreviations LENA^TM^: Language Environment Analysis.
-# ASHA recommends that abstracts be 150–250 words. The size limit for what can be included in your submission is set above 300 words,
-#currently 259 words
+Correspondence concerning this article should be addressed to Alejandrina Cristia, 29 rue d'Ulm, 75005, Paris, France. E-mail: alecristia@gmail.com}
+\usepackage{lineno}
 
-abstract: >
-  **Purpose:** Recording young children's vocalizations through wearables is a promising method.
+\linenumbers
+\usepackage{csquotes}
+\usepackage{setspace}
+\AtBeginEnvironment{tabular}{\singlespacing}
+\AtBeginEnvironment{lltable}{\singlespacing}
+\AtBeginEnvironment{tablenotes}{\singlespacing}
+\captionsetup[table]{font={stretch=1.5}}
+\captionsetup[figure]{font={stretch=1.5}}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[shorthands=off,main=english]{babel}
+\else
+  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
+  \usepackage{polyglossia}
+  \setmainlanguage[]{english}
+\fi
+
+\title{Describing vocalizations in young children: A big data approach through citizen science annotations}
+
+\date{}
+
+\abstract{
+\textbf{Purpose:} Recording young children's vocalizations through wearables is a promising method.
 However, accurately and rapidly annotating these files remains challenging.
@@
 
-  **Method:** Segments identified by LENA^TM^ as produced by the key child were extracted from one daylong recording for each of 20 participants: 10  low-risk control children and 10 children diagnosed  with  Angelman  syndrome, a  neurogenetic  syndrome  characterized  by  severe  language  impairments.
+\textbf{Method:} Segments identified by LENA\textsuperscript{TM} as produced by the key child were extracted from one daylong recording for each of 20 participants: 10 low-risk control children and 10 children diagnosed with Angelman syndrome, a neurogenetic syndrome characterized by severe language impairments.
 Speech samples were annotated by trained annotators in the laboratory as well as by citizen scientists on Zooniverse.
@@
 This allowed the derivation of two child-level vocalization metrics: the Linguistic Proportion, and the Canonical Proportion.
-
 \newline
 
-  **Results:** At the segment level, Zooniverse classifications had moderate precision and recall.
+\textbf{Results:} At the segment level, Zooniverse classifications had moderate precision and recall.
 More importantly, the Linguistic Proportion and the Canonical Proportion derived from Zooniverse annotations were highly correlated with those derived from laboratory annotations.
-
 \newline
 
-  **Conclusion:** Annotations obtained through a citizen science platform can help us overcome challenges posed by the process of annotating daylong speech recordings.
+\textbf{Conclusion:} Annotations obtained through a citizen science platform can help us overcome challenges posed by the process of annotating daylong speech recordings.
 Particularly when used in composites or derived metrics, such annotations can be used to investigate early markers of language delays in non-typically developing children.
+}
 
-# type: speech; areas of interest: assessment, infants, meta-analyses, prevention, speech recognition, screening, language, early intervention, cultural and ling diversity, dev lang disorders, autism, bilingualism --> choose 5: infants, prevention, dev lang disorders, early interv, diagnostics
-# Clinical practice research that evaluates methods of prevention, screening, diagnosis, treatment or models of service delivery that may improve the function and quality of life for individuals with or at risk of developing disorders of speech, language, cognition, swallowing, audition, and/or balance.
-# #up to 3 keywords
-# keywords          : " reliability; validation; daylong audiorecordings"
-# wordcount         : "7,113"
-
-bibliography      : ["zoo.bib"]
-
-figsintext        : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-lineno            : yes
-
-header-includes:
-  - \usepackage{setspace}
-  - \AtBeginEnvironment{tabular}{\singlespacing}
-  - \AtBeginEnvironment{lltable}{\singlespacing}
-  - \AtBeginEnvironment{tablenotes}{\singlespacing}
-  - \captionsetup[table]{font={stretch=1.5}}
-  - \captionsetup[figure]{font={stretch=1.5}}
-
-class             : "man"
-output            : papaja::apa6_pdf
-
----
-
-```{r setup, include = FALSE}
-library("papaja")
-library(tidyverse)
-library(dplyr)
-library(ggplot2)
-library(caret)
-library(scales)
-library(kableExtra)
-library(rel)
-library(ggpubr)
-library(lme4)
-#knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = FALSE, fig.pos = "T")
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = FALSE)
-```
-
-```{r analysis-preferences, include = FALSE}
-# Seed for random number generation
-set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
-```
-
-```{r read-in}
-# read datasets
-
-data_ang <- read.csv("../Derived_Data/classifications_PU_zoon_final17.csv",header=T,sep=",")
-data_td <- read.csv("../Derived_Data/classifications_PU_zoon_final.csv")
-data_all<-rbind(data_ang, data_td)
-
-#remove the word mixed that takes up space and is unnecessary
-data_all$Zoon_classif=factor(gsub("Mixed_","",as.character(data_all$Zoon_classif),fixed=T))
-#relevel the factor so that it's easier to read
-data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=c("Canonical","Non-Canonical",                   "Crying","Laughing","Junk",levels(data_all$Zoon_classif)[grep("_",levels(data_all$Zoon_classif))]))
-# create lab column with easier to read correspondance
-data_all$lab<-as.character(data_all$Major_Choice)
-data_all$lab[data_all$lab=="Non-canonical syllables"]<-"Non-Canonical"
-data_all$lab[data_all$lab=="Canonical syllables"]<-"Canonical"
-data_all$lab[data_all$lab %in% c("Don't mark","None")]<-"Junk"
-data_all$lab=factor(data_all$lab,levels=levels(data_all$Zoon_classif))
-#apply same factor levels as zooniverse so that we can do symmetrical confusion matrices
-
-#add binomials for Linguistic Proportion
-data_all$lab_ling=ifelse(data_all$lab %in% c("Canonical","Non-Canonical"),1,0)
-data_all$zoo_ling=ifelse(data_all$Zoon_classif %in% c("Canonical","Non-Canonical"),1,0)
-data_all$lab_ling[data_all$lab=="Junk"]<-NA
-data_all$zoo_ling[data_all$lab=="Junk"]<-NA
-
-#add binomials for Canonical Proportion
-data_all$lab_can=data_all$zoo_can=NA
-data_all$lab_can[data_all$lab=="Canonical"]<-1
-data_all$lab_can[data_all$lab=="Non-Canonical"]<-0
-data_all$zoo_can[data_all$Zoon_classif=="Canonical"]<-1
-data_all$zoo_can[data_all$Zoon_classif=="Non-Canonical"]<-0
-
-demo_data=read.csv("../Derived_Data/demo-data.tsv",sep="\t")
-#add filenames to demo data, to be used later
-demo_data_fn <- demo_data %>% 
-     left_join(select(data_all, filename, ChildID), by = c("ChildID"))
-demo_data_fn<-unique(demo_data_fn)
-
-```
+\begin{document}
+\maketitle
 
@@
 
-- create OSF for this paper specifically
-- references!
-- merge noncan/can for LENA^TM^ in Table 1
-- appropriately merge col names in tab 1
-- add 2 explo analyses that are still missing
-- Redo pipeline so that it's transparent & integrated
+\begin{itemize}
+\tightlist
+\item
+  create OSF for this paper specifically
+\item
+  references!
+\item
+  merge noncan/can for LENA\textsuperscript{TM} in Table 1
+\item
+  appropriately merge col names in tab 1
+\item
+  add 2 explo analyses that are still missing
+\item
+  Redo pipeline so that it's transparent \& integrated
+\end{itemize}
 
 Language production at 3 years of age is the best predictor of later academic and economic achievement (Law et al., 2003).
-Early language delays can adversely affect children’s literacy, behavior, social interaction, and scholastic achievement extending well into adulthood, with early interventions being a better societal investment than later ones (Heckman, 2006).
-Advancements in the field of wearable technologies, such as LENA^TM^ recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally.
-Wearable recorders allow data collection to happen in the child’s natural environment, and at a large scale, which may be particularly helpful for children whose speech is not easily elicited.
+Early language delays can adversely affect children's literacy, behavior, social interaction, and scholastic achievement extending well into adulthood, with early interventions being a better societal investment than later ones (Heckman, 2006).
+Advancements in the field of wearable technologies, such as LENA\textsuperscript{TM} recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally.
+Wearable recorders allow data collection to happen in the child's natural environment, and at a large scale, which may be particularly helpful for children whose speech is not easily elicited.
 Although such long-form recordings are increasingly common REF, challenges remain with respect to how these data are handled, annotated, and analyzed.
@@
 
-###  Describing children's vocalizations in long-form recordings
+\hypertarget{describing-childrens-vocalizations-in-long-form-recordings}{%
+\subsubsection{Describing children's vocalizations in long-form recordings}\label{describing-childrens-vocalizations-in-long-form-recordings}}
 
@@
 
-The most commonly used software was created by the LENA^TM^ Foundation, and it returns a diarization of the audio signal split into key talkers, including the child wearing the device, as well as a split of these audio attributed to the child into three categories: crying, other fixed and vegetative signals, and speech vocalizations.
+The most commonly used software was created by the LENA\textsuperscript{TM} Foundation, and it returns a diarization of the audio signal split into key talkers, including the child wearing the device, as well as a split of these audio attributed to the child into three categories: crying, other fixed and vegetative signals, and speech vocalizations.
 These subcategories have not been widely validated (Cristia et al., 2020).
-To provide an idea of their accuracy, we classified the same segments that are used in the present study using LENA^TM^'s subtypes[^1] to compared these against the laboratory annotations used in the present study.
-Table 1 shows that LENA^TM^ has a good recall for speech-like vocalizations and crying but not laughing.
-Moreover, LENA^TM^ lacks an important distinction between more and less advanced vocalizations.
+To provide an idea of their accuracy, we classified the same segments that are used in the present study using LENA\textsuperscript{TM}'s subtypes\footnote{If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as \enquote{junk} or not categorized.} to compared these against the laboratory annotations used in the present study.
+Table 1 shows that LENA\textsuperscript{TM} has a good recall for speech-like vocalizations and crying but not laughing.
+Moreover, LENA\textsuperscript{TM} lacks an important distinction between more and less advanced vocalizations.
 
-<!-- AS said about not laughing: might be worth noting how low frequency this is and how that might be part of the problem in addition to being difficult to distinguish from crying. -->
+\begin{table}
 
-[^1]: If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as "junk" or not categorized.
-
-```{r tab-LENA-challenge, comment = F, message = F, hide = T, warning = F, echo = F, results = "asis"}
-mycaption="Comparison of recall percentages obtained with the baseline algorithm created by the ComParE team, in the context of the ComParE 2019 BabySounds subchallenge (Schuller et al., 2019), LENATM labels (note LENATM does not distinguish between Canonical and Non-canonical), and those obtained in this study through Zooniverse annotations.
+\caption{\label{tab:tab-LENA-challenge}Comparison of recall percentages obtained with the baseline algorithm created by the ComParE team, in the context of the ComParE 2019 BabySounds subchallenge (Schuller et al., 2019), LENATM labels (note LENATM does not distinguish between Canonical and Non-canonical), and those obtained in this study through Zooniverse annotations.
 Label frequency indicates the prevalence of the relevant label (Crying, Laughing, etc.) in each dataset (the same Lab dataset was used for both LENATM and Zooniverse).
-UAR stands for unweighted average recall, WAR for weighted average recall (which takes into account label frequency)."
-
-read.table("tab-lena-challenge.txt",header=T,sep="\t")->tab1
-colnames(tab1)<-c("","Recall","","","Label","Frequency")
-kable(tab1,cap=mycaption)
-```
+UAR stands for unweighted average recall, WAR for weighted average recall (which takes into account label frequency).}
+\centering
+\begin{tabular}[t]{l|l|l|l|l|l}
+\hline
+ & Recall &  &  & Label & Frequency\\
+\hline
+ & C2019B & LENA & Zoon. & C2019B & Lab\\
+\hline
+Crying & 70.6 & 78.0 & 49.0 & 669 & 581\\
+\hline
+Laughing & 41.5 & 12.0 & 56.0 & 149 & 137\\
+\hline
+Non-can. & 24.1 & 73.0 & 77.0 & 4485 & 5512\\
+\hline
+Canonical & 66.4 &  & 59.0 & 1426 & 1715\\
+\hline
+Junk & 67.3 & 1.0 & 44.0 & 4575 & 2169\\
+\hline
+UAR & 54.0 & 41.0 & 57.0 &  & \\
+\hline
+WAR & 49.9 & 57.0 & 65.0 &  & \\
+\hline
+\end{tabular}
+\end{table}
 
@@
 The split of speech vocalizations into canonical and non-canonical, and the overt inclusion of laughing, allow a somewhat finer-grained picture of children's vocal activity.
-An advantage of these algorithms is that there exists a challenge where performance has been benchmarked, so unlike LENA^TM^'s algorithms, we know which algorithms perform better than others.
-The ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55% (see Table 1).
-The team who won the challenge improved this by about 2%, primarily through gains in the laughing class obtained by adding training data (REF).
+An advantage of these algorithms is that there exists a challenge where performance has been benchmarked, so unlike LENA\textsuperscript{TM}'s algorithms, we know which algorithms perform better than others.
+The ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55\% (see Table 1).
+The team who won the challenge improved this by about 2\%, primarily through gains in the laughing class obtained by adding training data (REF).
 Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are considered, changes very much.
@@
 
-For example, most work using LENA^TM^ software reports on children's vocalization counts.
-Although this has been sometimes criticized as being more about quantity than quality REF, this is a promising metric because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA^TM^ REF.
-Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r ~ .3 with standardized language scores in a meta-analysis REF.
+For example, most work using LENA\textsuperscript{TM} software reports on children's vocalization counts.
+Although this has been sometimes criticized as being more about quantity than quality REF, this is a promising metric because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA\textsuperscript{TM} REF.
+Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r \textasciitilde{} .3 with standardized language scores in a meta-analysis REF.
 Thus despite representing a composite of skills, vocalization count may provide a useful estimate of vocalization development.
-
-<!-- AS said about voc count not sure I agree: https://pubs.asha.org/doi/full/10.1044/2020_JSLHR-19-00281?casa_token=fBByMmPlxbMAAAAA%3AbS2O5Mrymn30_aQWybitwMecENkDYLTjzexCqZcnX2Rx1VJ7OaT9qkIGNAAM95xRBGW_D3XX7mMEWHo -->
 
@@
 To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age REF.
-
-<!-- BT said about LP May want to put this 2nd as more exploatory given more lit on can. prop. -->
 
@@
 
-<!-- AS said: maybe name the disorders & delays here? -->
-
 In sum, there is a growing literature attempting to use data from daylong recordings to describe young children's vocalizations, but there are two outstanding challenges.
@@
 
-### Crowdsourcing: A potential solution for annotation
+\hypertarget{crowdsourcing-a-potential-solution-for-annotation}{%
+\subsubsection{Crowdsourcing: A potential solution for annotation}\label{crowdsourcing-a-potential-solution-for-annotation}}
 
 Crowdsourcing refers to the process whereby a task is solved by a crowd, rather than an individual.
-A number of fields, particularly in the data-driven sciences, have already engaged in the collection of data (including annotations) through crowdsourcing, thanks to its low cost and ecological value (Crump, McDonnell, & Gureckis, 2013; Sescleifer, Francoisse, & Lin, 2018;  Sprouse, 2011).
-For example, a systematic review on crowd-sourced ratings of speech (Sescleifer et al., 2018) found that "lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology"; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81.
+A number of fields, particularly in the data-driven sciences, have already engaged in the collection of data (including annotations) through crowdsourcing, thanks to its low cost and ecological value (Crump, McDonnell, \& Gureckis, 2013; Sescleifer, Francoisse, \& Lin, 2018; Sprouse, 2011).
+For example, a systematic review on crowd-sourced ratings of speech (Sescleifer et al., 2018) found that \enquote{lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology}; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81.
 On the other hand, the systematic review returned only 8 studies (of which only four were published in peer-reviewed journals), suggesting that there is considerable need for further research on this topic.
 
-Mechanical Turk (MTurk) MTurk is an online labor market created by Amazon to assist "requesters"''" in hiring and paying "workers" for the completion of computerized tasks.
-Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable (Berinsky et al., 2012), some question marks are raised as some "workers'" turn out to be bots, or poorly motivated (and potentially exploited).
+Mechanical Turk (MTurk) MTurk is an online labor market created by Amazon to assist \enquote{requesters}'\enquote{" in hiring and paying \enquote{workers} for the completion of computerized tasks.
+Although it is a leading crowdsourcing service, and some evidence suggests MTurkers} annotations can be quite reliable (Berinsky et al., 2012), some question marks are raised as some \enquote{workers'} turn out to be bots, or poorly motivated (and potentially exploited).
 
@@
 There is one previous study that attempted this approach.
-@cychosz2019canonical drew child vocalization data from a diverse set of corpora centered on children learning one of four languages: English, Tseltal, Tsimane', and Yélî.
-For the English and Tsimane' corpora, vocalizations were automatically identified using LENA^TM^, whereas the other two were extracted through manual segmentation.
-Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received 3 classifications into crying, laughing, canonical, non-canonical, or "junk" (with the latter tag used for clips that did not contain clear child voice).
+Cychosz et al.
+(2019) drew child vocalization data from a diverse set of corpora centered on children learning one of four languages: English, Tseltal, Tsimane', and Yélî.
+For the English and Tsimane' corpora, vocalizations were automatically identified using LENA\textsuperscript{TM}, whereas the other two were extracted through manual segmentation.
+Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received 3 classifications into crying, laughing, canonical, non-canonical, or \enquote{junk} (with the latter tag used for clips that did not contain clear child voice).
 They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgement of canonical or non-canonical.
@@
 The authors also studied how their canonical ratio related to age within each corpus that had a sufficient number of children varying in age.
-They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the "junk" category was very high.
-Additionally, they found that the correlation between canonical ratio and age was much weaker in two corpora where vocalizations were segmented with LENA^TM^ than through manual segmentation.
+They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the \enquote{junk} category was very high.
+Additionally, they found that the correlation between canonical ratio and age was much weaker in two corpora where vocalizations were segmented with LENA\textsuperscript{TM} than through manual segmentation.
 In sum, these analyses within corpora reveal that data may not be uniformly useful, with potential variation across corpora, children, and methods.
 
-### The present work
+\hypertarget{the-present-work}{%
+\subsubsection{The present work}\label{the-present-work}}
 
@@
 We hope that this methodology will open the road to larger scale analyses of children's vocalizations as captured by wearables.
-In addition, we went beyond @cychosz2019canonical's study in two important ways.
+In addition, we went beyond Cychosz et al.
+(2019)\enquote{s study in two important ways.
 First, we relied on the largest and best established citizen science platform: Zooniverse hosts more than 1.6 million users from diverse walks of life, and it offers a completely automatized API system to more easily scale tasks in a transparent and cumulative science fashion.
@@
 Those authors also had to rely on manual extraction of the resulting data from the iHearUplay team, which added to the administrative load.
-Second, a core goal of the present work was to determine how citizen scientists' annotations fare compared to the current gold standard, laboratory annotations.
-@cychosz2019canonical did not have a gold standard for their data.
+Second, a core goal of the present work was to determine how citizen scientists} annotations fare compared to the current gold standard, laboratory annotations.
+Cychosz et al.
+(2019) did not have a gold standard for their data.
 
@@
 When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (canonical+non-canonical)/(all vocalizations); and the Canonical Proportion (canonical)/(all linguistic vocalizations).
-Previous work suggests that children with Angelman Syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the canonical ratio was expected to increase with age [@hamrick2019capturing].
+Previous work suggests that children with Angelman Syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the canonical ratio was expected to increase with age (Hamrick, Seidl, \& Kelleher, n.d.).
 Although we could not rely on previous work to make predictions regarding the Linguistic Proportion, we reasoned that we should observe an increase in the Linguistic Proportion for both populations.
@@
 
-## Methods
+\hypertarget{methods}{%
+\subsection{Methods}\label{methods}}
 
-All analyses and key data are available for reproduction on OSF (https://osf.io/57yha/). Computer code used to collect the Zooniverse judgments is available from GitHub (Semenzin, 2020).
+All analyses and key data are available for reproduction on OSF (\url{https://osf.io/57yha/}). Computer code used to collect the Zooniverse judgments is available from GitHub (Semenzin, 2020).
 
-### Participants
+\hypertarget{participants}{%
+\subsubsection{Participants}\label{participants}}
 
@@
 
-### Equipment and data collection procedure
+\hypertarget{equipment-and-data-collection-procedure}{%
+\subsubsection{Equipment and data collection procedure}\label{equipment-and-data-collection-procedure}}
 
-Recordings were obtained with the Language ENvironment Analysis (LENA^TM^) Digital Language Processor,  LENA^TM^ is a lightweight recorder (< 60 g, 5.5 cm x 8.5 cm x 1.5 cm) designed to be worn inside a breast-pocket of purpose-made clothing.
+Recordings were obtained with the Language ENvironment Analysis (LENA\textsuperscript{TM}) Digital Language Processor, LENA\textsuperscript{TM} is a lightweight recorder (\textless{} 60 g, 5.5 cm x 8.5 cm x 1.5 cm) designed to be worn inside a breast-pocket of purpose-made clothing.
 In our data, every child contributed at least one daylong recording, with one child contributing two recordings (duration 11.58-16.00 hours, mean 15.05 hours).
 
-### Data preprocessing (LENA^TM^)
+\hypertarget{data-preprocessing-lenatm}{%
+\subsubsection{\texorpdfstring{Data preprocessing (LENA\textsuperscript{TM})}{Data preprocessing (LENATM)}}\label{data-preprocessing-lenatm}}
 
 The first step we took before beginning the annotation process was the identification of children's vocalizations within day-long recordings.
-We did this automatically, using the LENA^TM^ proprietary software.
-Specifically, the LENA^TM^ software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child’s environment (e.g., Female-Adult, TV).
-Key child labels have a precision of about XX% and a recall of about XX% (ref).
+We did this automatically, using the LENA\textsuperscript{TM} proprietary software.
+Specifically, the LENA\textsuperscript{TM} software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child's environment (e.g., Female-Adult, TV).
+Key child labels have a precision of about XX\% and a recall of about XX\% (ref).
 These stretches of audio assigned to the key child can contain cries, vegetative sounds, and linguistic vocalizations, as well as interstitial pauses.
@@
 
-<!-- BT said Lisa add in how portions of recording were selected (high vol + random) -->
-
-### Annotation by lab experts
+\hypertarget{annotation-by-lab-experts}{%
+\subsubsection{Annotation by lab experts}\label{annotation-by-lab-experts}}
 
 Annotation was carried out at the segment level during lab coding.
-A subset of segments LENA^TM^ identified as the key child were selected for human annotation.
-Segments were selected from 30 five-minute sections of the child’s recording: 10 sections during which the child was vocalizing at their highest rate, and 20 sections drawn randomly from the remainder of the recording.
+A subset of segments LENA\textsuperscript{TM} identified as the key child were selected for human annotation.
+Segments were selected from 30 five-minute sections of the child's recording: 10 sections during which the child was vocalizing at their highest rate, and 20 sections drawn randomly from the remainder of the recording.
 To ensure generalizability of our data, we included in this subset a number of segments of audio coming from both periods of high child volubility, as well as randomly selected periods of the day.
@@
 Coders are instructed to classify each segment based on the highest level of vocal maturity it contains (e.g., a segment with both canonical and non-canonical syllables would be classified as canonical; a segment with both laughing and non-canonical syllables would be classified as non-canonical).
-For canonical, non-canonical, and word segments, coders indicated how many syllables of each type of vocal maturity they hear (e.g., "ah ma ba" would be 1 non-canonical and 2 canonical syllables).
-Coders are also given a "Don’t Mark" option, which they are instructed to use if the segment does not sound like a segment made by the target child, or if there is any overlapping speech or other noise in the background which could affect acoustic analyses.
-For the purposes of the present study, we considered a "word" judgment to be equivalent to a "canonical" judgment.
-This may be false sometimes (e.g., "oh" may be coded as a real word by laboratory experts, although it is a non-canonical syllable), but we could not have a "word" judgment in our Zooniverse annotations because that requires context (which compromises privacy -- see next section).
-This process resulted in high levels of agreement: 97.22% of the lab-annotated segments had either 2 or 3 coders agree.
+For canonical, non-canonical, and word segments, coders indicated how many syllables of each type of vocal maturity they hear (e.g., \enquote{ah ma ba} would be 1 non-canonical and 2 canonical syllables).
+Coders are also given a \enquote{Don't Mark} option, which they are instructed to use if the segment does not sound like a segment made by the target child, or if there is any overlapping speech or other noise in the background which could affect acoustic analyses.
+For the purposes of the present study, we considered a \enquote{word} judgment to be equivalent to a \enquote{canonical} judgment.
+This may be false sometimes (e.g., \enquote{oh} may be coded as a real word by laboratory experts, although it is a non-canonical syllable), but we could not have a \enquote{word} judgment in our Zooniverse annotations because that requires context (which compromises privacy -- see next section).
+This process resulted in high levels of agreement: 97.22\% of the lab-annotated segments had either 2 or 3 coders agree.
 
-<!-- Reliability is built into the annotation routine such that each segment is coded by 3 separate coders, and only segments where 2 of 3 coders agree are used in subsequent phases of this study.
--->
-<!-- Lisa aC removed this because this is actually false -- all segments were put up, including those with fewer than 2 agreement -->
-
-<!-- Lisa, What happens if there are multiple syllables or multiple words? also, what if there is speech and non-speech, eg cry, in the same segment? -->
-
-### Annotation on Zooniverse
+\hypertarget{annotation-on-zooniverse}{%
+\subsubsection{Annotation on Zooniverse}\label{annotation-on-zooniverse}}
 
@@
 To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms clips and their corresponding segment.
-Clips were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling (https://github.com/zooniverse/panoptes-cli).
+Clips were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling (\url{https://github.com/zooniverse/panoptes-cli}).
 
@@
 Users were asked to assign each 500ms clip to one out of five possible categories: (1) canonical, (2) non-canonical, (3) crying, (4) laughing, and (5) junk (overlapping speech, non-infant speech, silence, external sounds).
-Notice that the latter is narrower than the "Don't mark" category of the lab annotation routine, which was used any time there was noise or overlap that could affect acoustic analyses.
+Notice that the latter is narrower than the \enquote{Don't mark} category of the lab annotation routine, which was used any time there was noise or overlap that could affect acoustic analyses.
 
@@
 
-``` {r fig-zoo, comment=F, message=F, hide = T, warning = F, echo = F,  fig.cap="Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needs to make; as well as the field guide, which remains accessible on the right of the screen throughout the session, and which allows users to revise examples of the different categories."}
-knitr::include_graphics("zooniverse-pufig.pdf")
-```
+\begin{figure}
+\centering
+\includegraphics{zooniverse-pufig.pdf}
+\caption{\label{fig:fig-zoo}Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needs to make; as well as the field guide, which remains accessible on the right of the screen throughout the session, and which allows users to revise examples of the different categories.}
+\end{figure}
 
-### Data post-processing
+\hypertarget{data-post-processing}{%
+\subsubsection{Data post-processing}\label{data-post-processing}}
 
 An impressive total of 4,825 individual Zooniverse users provided labels for the Maturity of Baby Sounds project, of which the present data set is one part.
-For this project, we collected a total of 169,767 judgments provided for 33,731 500-ms chunks, corresponding to 11,980 LENA^TM^ segments.
-Nearly a fifth of chunks did not have at least 3 labels in agreement out of the 5 Zooniverse labels (N = 6,585, 19% of all chunks).
-Of the chunks without a majority agreement, 4341 (66%) contained one or two Junk judgements (out of 5), 6523 (99,9%) had at least two matching judgements (the threshold used for lab-annotated segments), and only 61 (0,01%) had 5 different judgements.
-Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the 81% of chunks that did have at least 3 labels in agreement; this represented 135,725 labels for 27,145 chunks, corresponding to 11,593 LENA^TM^ segments.
+For this project, we collected a total of 169,767 judgments provided for 33,731 500-ms chunks, corresponding to 11,980 LENA\textsuperscript{TM} segments.
+Nearly a fifth of chunks did not have at least 3 labels in agreement out of the 5 Zooniverse labels (N = 6,585, 19\% of all chunks).
+Of the chunks without a majority agreement, 4341 (66\%) contained one or two Junk judgements (out of 5), 6523 (99,9\%) had at least two matching judgements (the threshold used for lab-annotated segments), and only 61 (0,01\%) had 5 different judgements.
+Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the 81\% of chunks that did have at least 3 labels in agreement; this represented 135,725 labels for 27,145 chunks, corresponding to 11,593 LENA\textsuperscript{TM} segments.
 As the segments average 1.12 seconds in length, this means about 3.8 hours of audio data were annotated by 8 different annotators (3 in the laboratory, 5 on Zooniverse).
 
-Lab annotators provided judgments at the level of LENA^TM^ segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments.
+Lab annotators provided judgments at the level of LENA\textsuperscript{TM} segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments.
 Therefore, we needed to combine chunk-level judgments to reconstruct segment-level judgments.
 
-We did this by considering the majority label extracted from Zooniverse judgements for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were junk, then the segment was labelled as junk; then, following a hierarchy (Canonical>Non-canonical>Laughing>Crying), segments with at least one instance of each judgement were labelled as such: e.g.
-if one canonical judgement was present, then the segment was labelled as canonical, otherwise if a non-canonical judgement was present, the segment was labelled as non-canonical, and so on.
+We did this by considering the majority label extracted from Zooniverse judgements for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were junk, then the segment was labelled as junk; then, following a hierarchy (Canonical\textgreater{}Non-canonical\textgreater{}Laughing\textgreater{}Crying), segments with at least one instance of each judgement were labelled as such: e.g.~if one canonical judgement was present, then the segment was labelled as canonical, otherwise if a non-canonical judgement was present, the segment was labelled as non-canonical, and so on.
 This is essentially the same thought process expert annotators followed when labeling segment-level audio, as each segment can represent multiple classification subtypes).
-LENA^TM^ segments are not equivalent to child vocalizations: A segment may contain 1 or more vocalizations, which may be of the same type or not.
+LENA\textsuperscript{TM} segments are not equivalent to child vocalizations: A segment may contain 1 or more vocalizations, which may be of the same type or not.
 
-<!-- AS said I feel like a figure might help here. -->
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
 
-```{r, echo=F}
-# remove non-majority labels
-#TODO
-
-# we map the mixed
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical"]<-"Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical_Crying"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Crying"]<-"Crying"
-data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Crying"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing_Crying"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Canonical"]<-"Canonical"
-# +
-data_all$Zoon_classif[data_all$Zoon_classif=="Canonical_Crying"]<-"Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Canonical_Laughing"]<-"Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical_Crying"]<-"Non-Canonical"
-
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Laughing"]<-"Crying"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Canonical_Laughing"]<-"Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Laughing_Non-Canonical"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Non-Canonical"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Non-Canonical_Laughing"]<-"Non-Canonical" 
-data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing"]<-"Non-Canonical"
-
-#and reset the factors for cleanliness
-data_all$Zoon_classif=factor(data_all$Zoon_classif)
-data_all$lab=factor(data_all$lab)
-sample_data<-cbind(data_all$lab,data_all$Zoon_classif)
-
-table_zoon=round(table(data_all$Zoon)/sum(table(data_all$Zoon))*100)
-table_lab=round(table(data_all$lab)/sum(table(data_all$lab))*100)
-```
-
-```{r prepare data by child, echo=F}
-
-#get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
-ztab=table(data_all$filename,data_all$Zoon_classif)
-z_lp=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
-z_cp=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
-
-ltab=table(data_all$filename,data_all$lab)
-l_lp=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
-l_cp=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
-
-#and also get junk from zooniverse
-z_junk=ztab[,c("Junk")]/rowSums(ztab)
-
-#put all the proportions together
-if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) proportions=cbind(rownames(ztab),z_lp,z_cp,l_lp,l_cp,z_junk) else print("oops this code needs to be more complex because we don't have the same kids for the two proportions")
-colnames(proportions)[1]<-"filename"
-
-#ages=aggregate(data_all$Age,by=list(data_all$ChildID),mean) #this is a weird way of adding ages, since all of the ages for a given child should be the same if there is only one recording, and if there are multiple recordings, then we should not get the mean
-#improvement: now we merge with a demo data tab, but note this is merged with child id, so the problem of multiple recs per child is still there
-
-# Created demo_data with filenames. Use filenames instead of childIDs to merge proportions and demo data.
-
-merge(proportions,demo_data_fn,by="filename")->proportions
-colnames(proportions)[dim(proportions)[2]]<-"Age"
-
-#cbinding results in text, so we numerize the proportions
-for(thisvar in c("z_lp","z_cp","l_lp","l_cp","z_junk")) proportions[,thisvar]=as.numeric(as.character(proportions[,thisvar]))
-
-#add median junk
-proportions$median_junk<-ifelse(proportions$z_junk<median(proportions$z_junk),"lower_junk","higher_junk")
-
-#summary(proportions)
-
-merge(data_all,proportions,all=T)->data_all
-```
-## Results
-
-### Descriptive analyses
-
-<!-- lh said I think you mention above that individuals with AS rarely use more than a handful of meaningful words - if we decide to go with just looking at these analyses broken down by group, that might be helpful to reiterate here -->
-<!-- ac said in response: why do you talk about meaningful words? These analyses do not separate words from the rest... -->
+\hypertarget{descriptive-analyses}{%
+\subsubsection{Descriptive analyses}\label{descriptive-analyses}}
 
 In this section, we provide descriptive analyses of our dataset.
-According to lab annotators, `r table_lab["Canonical"]`% of segments were canonical, `r table_lab["Non-Canonical"]`% non-canonical, `r table_lab["Laughing"]`% laughing, and `r table_lab["Crying"]`% crying, with the remaining `r table_lab["Junk"]`% being categorized as "Don't mark".
-Zooniverse data revealed a similar distribution:  `r table_zoon["Canonical"]`% canonical, `r table_zoon["Non-Canonical"]`% non-canonical, `r table_zoon["Laughing"]`% laughing, `r table_zoon["Crying"]`% crying, `r table_zoon["Junk"]`% junk.
-Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = ("Canonical"+"Non-Canonical")/"All vocalizations" (i.e., we remove, junk), and ii) Canonical Proportion = "Canonical"/("Canonical"+"Non-Canonical") (i.e., we remove junk + non-linguistic).
+According to lab annotators, 15\% of segments were canonical, 56\% non-canonical, 2\% laughing, and 5\% crying, with the remaining 22\% being categorized as \enquote{Don't mark}.
+Zooniverse data revealed a similar distribution: 15\% canonical, 60\% non-canonical, 4\% laughing, 9\% crying, 13\% junk.
+Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (\enquote{Canonical}+\enquote{Non-Canonical})/\enquote{All vocalizations} (i.e., we remove, junk), and ii) Canonical Proportion = \enquote{Canonical}/(\enquote{Canonical}+\enquote{Non-Canonical}) (i.e., we remove junk + non-linguistic).
 See Figure 2 for results.
 
-```{r fig-corage, echo=F,fig.cap="Correlations between child-level descriptors and age as a function of metric (linguistic ratio in the top row, canonical ratio in the bottom row), annotation method, and child group (red=low-risk, and black Angelman Syndrome)"}
-
-prettynames=c("Linguistic Proportion (Zooniverse)","Canonical Proportion (Zooniverse)",
-             "Linguistic Proportion (Lab)","Canonical Proportion (Lab)" )
-names(prettynames)<-c("z_lp","z_cp","l_lp","l_cp")
-mycols=c("black","red")
-names(mycols)<-c("Low-RiskControl","AngelmanSyndrome")
-mypch=c(4,20)
-names(mypch)<-c("Low-RiskControl","AngelmanSyndrome")
-
-lp_range=range(proportions[,c("z_lp","l_lp")])
-cp_range=range(proportions[,c("z_cp","l_cp")])
-
-#jpeg("../Results/corage.jpg",width=20,height=20,units="cm",res=300)
-layout(matrix(c(1:4), 2, 2, byrow = F))
-par(mar = c(4, 4, 1, 1))
-
-#create matrix to hold r results
-myr=matrix(NA,nrow=4,ncol=2)
-rownames(myr)<-c("z_lp","z_cp","l_lp","l_cp")
-colnames(myr)<-c("AS","LR")
-mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
-
-for(thisvar in c("z_lp","z_cp","l_lp","l_cp")) {
-
-  if(thisvar %in% c("z_lp","l_lp")) myrange=lp_range else myrange=cp_range
-
-  plot(proportions[,thisvar]~proportions$Age, pch=mypch[proportions$Diagnosis],xlab="Age (months)",ylab=prettynames[thisvar],
-       col=mycols[proportions$Diagnosis],
-         ylim=myrange)
-  abline(lm(proportions[,thisvar]~proportions$Age,subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black")
-
-  #get cor properties
-  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"])
-    myr[thisvar,"AS"]=round(thiscor$estimate,3)
-    mylb[thisvar,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[thisvar,"AS"]=round(thiscor$conf.int[2],3)
-    myn[thisvar,"AS"]=round(thiscor$parameter,3)
-    myp[thisvar,"AS"]=round(thiscor$p.value,3)
-
-    #add r to figure
-  text(mean(proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"]),mean(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar]),paste0("r=",myr[thisvar,"AS"]),col="black")
-
-  #repeat for LR  
-    abline(lm(proportions[,thisvar]~proportions$Age,subset=c(proportions$Diagnosis!="AngelmanSyndrome")),col="red")
-
-      #get cor properties
-  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"])
-    myr[thisvar,"LR"]=round(thiscor$estimate,3)
-    mylb[thisvar,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[thisvar,"LR"]=round(thiscor$conf.int[2],3)
-    myn[thisvar,"LR"]=round(thiscor$parameter,3)
-    myp[thisvar,"LR"]=round(thiscor$p.value,3)
-
-    #add r to figure
-
-    text(mean(proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"]),mean(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar]),paste0("r=",myr[thisvar,"LR"]),col="red")
-}
-#dev.off()
-```
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/fig-corage-1.pdf}
+\caption{\label{fig:fig-corage}Correlations between child-level descriptors and age as a function of metric (linguistic ratio in the top row, canonical ratio in the bottom row), annotation method, and child group (red=low-risk, and black Angelman Syndrome)}
+\end{figure}
 
 Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups.
-There was a near-zero relationship among the older children diagnosed with Angelman Syndrome r(`r myn["l_lp","AS"]`) = `r myr["l_lp","AS"]`, CI [`r mylb["l_lp","AS"]`,`r myhb["l_lp","AS"]`], p=`r myp["l_lp","AS"]`]; and a significant association among younger low-risk control children r(`r myn["l_lp","LR"]`) = `r myr["l_lp","LR"]`, CI [`r mylb["l_lp","LR"]`,`r myhb["l_lp","LR"]`], p=`r myp["l_lp","LR"]`].
-The Canonical Proportion exhibited non-significant developmental decreases among older children diagnosed with Angelman Syndrome r(`r myn["l_cp","AS"]`) = `r myr["l_cp","AS"]`, CI [`r mylb["l_cp","AS"]`,`r myhb["l_cp","AS"]`], p=`r myp["l_cp","AS"]`]; and marginal developmental increases among low-risk control r(`r myn["l_cp","LR"]`) = `r myr["l_cp","LR"]`, CI [`r mylb["l_cp","LR"]`,`r myhb["l_cp","LR"]`], p=`r myp["l_cp","LR"]`].
+There was a near-zero relationship among the older children diagnosed with Angelman Syndrome r(8) = 0.02, CI {[}-0.62,0.64{]}, p=0.97{]}; and a significant association among younger low-risk control children r(8) = 0.74, CI {[}0.21,0.94{]}, p=0.01{]}.
+The Canonical Proportion exhibited non-significant developmental decreases among older children diagnosed with Angelman Syndrome r(8) = -0.56, CI {[}-0.88,0.11{]}, p=0.09{]}; and marginal developmental increases among low-risk control r(8) = 0.61, CI {[}-0.04,0.90{]}, p=0.06{]}.
 
-Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman Syndrome r(`r myn["z_lp","AS"]`) = `r myr["z_lp","AS"]`, CI [`r mylb["z_lp","AS"]`,`r myhb["z_lp","AS"]`], p=`r myp["z_lp","AS"]`]; whereas younger low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r myr["z_lp","LR"]`, CI [`r mylb["z_lp","LR"]`,`r myhb["z_lp","LR"]`], p=`r myp["z_lp","LR"]`].
-Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman Syndrome r(`r myn["z_cp","AS"]`) = `r myr["z_cp","AS"]`, CI [`r mylb["z_cp","AS"]`,`r myhb["z_cp","AS"]`], p=`r myp["z_cp","AS"]`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r myr["z_cp","LR"]`, CI [`r mylb["z_cp","LR"]`,`r myhb["z_cp","LR"]`], p=`r myp["z_cp","LR"]`].
+Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman Syndrome r(8) = -0.10, CI {[}-0.68,0.57{]}, p=0.79{]}; whereas younger low-risk control children showed a significant increase with age r(8) = 0.68, CI {[}0.08,0.92{]}, p=0.03{]}.
+Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman Syndrome r(8) = -0.42, CI {[}-0.83,0.29{]}, p=0.23{]}; and marginal developmental increases among low-risk control children, r(8) = 0.60, CI {[}-0.04,0.89{]}, p=0.06{]}.
 
-## Main analyses
+\hypertarget{main-analyses}{%
+\subsection{Main analyses}\label{main-analyses}}
 
@@
 
-```{r mygac-conf, echo=F}
-mygac=gac(data = sample_data, kat = 5, weight = c("unweighted"),
-    conf.level = 0.95)
-
-mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
-conf_tab=mycf$table
-# this package uses sensitivity & specificity
-#Sensitivity=recall
-#Specificity=precision
-#mycf
-```
-
-```{r fig-prec-rec, echo=F}
-
-colsums=colSums(conf_tab)
-my_conf_tab=conf_tab
-for(i in 1:dim(conf_tab)[2]) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
-#colSums(my_conf_tab) #check
-prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
-prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
-colnames(prop_cat)[3]<-"pr"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","pr")])
-prec_plot = ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(pr)), colour = "white") +
-  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=4) +
-  geom_text(aes(label = Freq), vjust = 1,size=4) +
-  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
-     theme(legend.position = "none") +
-  xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-
-# repeat for recall
-rowsums=rowSums(conf_tab)
-my_conf_tab=conf_tab
-for(i in 1:dim(conf_tab)[2]) my_conf_tab[,i]=my_conf_tab[,i]/rowsums[i]
-#rowSums(my_conf_tab)
-prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
-prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
-colnames(prop_cat)[3]<-"rec"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","rec")])
-rec_plot = ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(rec)), colour = "white") +
-  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=4) +
-  geom_text(aes(label = Freq), vjust = 1,size=4) +
-  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
-     theme(legend.position = "none") +
-  xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Recall")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-
-ggarrange(prec_plot, rec_plot, 
-          labels = c("A", "B"),
-          ncol = 2, nrow = 1)
-
-```
+\includegraphics{paper_files/figure-latex/fig-prec-rec-1.pdf}
 
 These visualizations suggest that performance is moderate to good, which was confirmed via statistical analyses.
-The overall (weighted) accuracy is `r round(mycf$overall["Accuracy"]*100)`%, CI = [`r round(mycf$overall["AccuracyLower"]*100)`,`r round(mycf$overall["AccuracyUpper"]*100)`],  kappa is `r round(mycf$overall["Kappa"],3)`, and the Gwet's AC1 coefficient is `r round(mygac$est,3)`, CI = [`r round(mygac$lb,3)`,`r round(mygac$ub,3)`].
+The overall (weighted) accuracy is 65\%, CI = {[}64,66{]}, kappa is 0.43, and the Gwet's AC1 coefficient is 0.59, CI = {[}0.58,0.60{]}.
 
-### Child level descriptors
+\hypertarget{child-level-descriptors}{%
+\subsubsection{Child level descriptors}\label{child-level-descriptors}}
 
-```{r corlab-zoo, echo=F}
+\includegraphics{paper_files/figure-latex/corlab-zoo-1.pdf}
 
-layout(matrix(c(1:2), 1, 2, byrow = F))
-par(mar = c(4, 4, 1, 1))
+Although the classification at the clip level is only moderately accurate, what we are ultimately interested in is whether citizen scientists' classifications are able to provide a reliable snapshot of childrens' individual development.
+Looking at all 20 children together, we found a strong positive correlation r(18) = 0.81, CI {[}0.57,0.92{]}, p=0{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman Syndrome r(8) = 0.88, CI {[}0.55,0.97{]}, p=0.00{]}; low-risk control r(8) = 0.82, CI {[}0.40,0.96{]}, p=0.00{]}.
 
-#create matrix to hold r results
-myr=matrix(NA,nrow=2,ncol=3)
-rownames(myr)<-c("lp","cp")
-colnames(myr)<-c("AS","LR","all")
-mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
+Similarly, a strong positive correlation is found in the Canonical Proportion r(18) = 0.94, CI {[}0.84,0.98{]}, p=0{]}.
+When we split by participant group, correlations remain high although we do note they are somewhat smaller for the older children with Angelman Syndrome: r(8) = 0.86, CI {[}0.50,0.97{]}, p=0.00{]}; than the low-risk control children r(8) = 0.96, CI {[}0.83,0.99{]}, p=0.
 
-lims=range(c(proportions[,"z_lp"],proportions[,"l_lp"]))
-
-  plot(proportions[,"z_lp"]~proportions[,"l_lp"], xlab=prettynames["l_lp"],ylab=prettynames["z_lp"],
-       xlim=lims,ylim=lims,
-       col=mycols[proportions$Diagnosis], pch=mypch[proportions$Diagnosis])
-  lines(c(0,1),c(0,1),lty=2,col="darkgray")
-
-  abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black")
-
-  #get cor properties
-  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_lp"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
-    myr[1,"AS"]=round(thiscor$estimate,3)
-    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
-    myn[1,"AS"]=round(thiscor$parameter,3)
-    myp[1,"AS"]=round(thiscor$p.value,3)
-
-  #repeat for LR  
-    abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$Diagnosis!="AngelmanSyndrome")),col="red")
-
-      #get cor properties
-  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"],proportions[proportions$Diagnosis!="AngelmanSyndrome","z_lp"])
-    myr[1,"LR"]=round(thiscor$estimate,3)
-    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
-    myn[1,"LR"]=round(thiscor$parameter,3)
-    myp[1,"LR"]=round(thiscor$p.value,3)
-
-#repeat for all cor
-
-      thiscor=cor.test(proportions$z_lp,proportions$l_lp)
-    myr[1,"all"]=round(thiscor$estimate,3)
-    mylb[1,"all"]=round(thiscor$conf.int[1],3)
-    myhb[1,"all"]=round(thiscor$conf.int[2],3)
-    myn[1,"all"]=round(thiscor$parameter,3)
-    myp[1,"all"]=round(thiscor$p.value,3)
-
-    # repeat whole for cp
-
-    lims=range(c(proportions[,"z_cp"],proportions[,"l_cp"]))
-
-  plot(proportions[,"z_cp"]~proportions[,"l_cp"], xlab=prettynames["l_cp"],ylab=prettynames["z_cp"],
-       xlim=lims,ylim=lims,
-       col=mycols[proportions$Diagnosis], pch=mypch[proportions$Diagnosis])
-  lines(c(0,1),c(0,1),lty=2,col="darkgray")
-
-  abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black")
-
-  #get cor properties
-  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_cp"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
-    myr[2,"AS"]=round(thiscor$estimate,3)
-    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
-    myn[2,"AS"]=round(thiscor$parameter,3)
-    myp[2,"AS"]=round(thiscor$p.value,3)
-
-  #repeat for LR  
-    abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$Diagnosis!="AngelmanSyndrome")),col="red")
-
-      #get cor properties
-  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"],proportions[proportions$Diagnosis!="AngelmanSyndrome","z_cp"])
-    myr[2,"LR"]=round(thiscor$estimate,3)
-    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
-    myn[2,"LR"]=round(thiscor$parameter,3)
-    myp[2,"LR"]=round(thiscor$p.value,3)
-
-#repeat for all cor
-
-      thiscor=cor.test(proportions$z_cp,proportions$l_cp)
-    myr[2,"all"]=round(thiscor$estimate,3)
-    mylb[2,"all"]=round(thiscor$conf.int[1],3)
-    myhb[2,"all"]=round(thiscor$conf.int[2],3)
-    myn[2,"all"]=round(thiscor$parameter,3)
-    myp[2,"all"]=round(thiscor$p.value,3)
-
-mycortab=t(myr)
-rownames(mycortab)<-paste("All seg",rownames(mycortab))
-```
-
-Although the classification at the clip level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of childrens’ individual development.
-Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman Syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
-
-Similarly, a strong positive correlation is found in the Canonical Proportion r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`].
-When we split by participant group, correlations remain high although we do note they are somewhat smaller for the older children with Angelman Syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
-
-```{r glmer-agediag, echo=F}
-# one outlier in ling prop is removed
-
-ling_all=glmer(lab_ling~zoo_ling*Diagnosis*Age + (1|ChildID),data=data_all,family="binomial",subset=c(ChildID!="3681_1"))
-#plot(lin_mod)
-# doesn't converge
-
-ling_AS=glmer(lab_ling~zoo_ling*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis=="AngelmanSyndrome"))
-#only effect of zoo
-
-ling_LR=glmer(lab_ling~zoo_ling*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis!="AngelmanSyndrome" & ChildID!="3681_1"))
-#both main and interaction are sig
-
-can_all=glmer(lab_can~zoo_can*Diagnosis*Age + (1|ChildID),data=data_all,family="binomial")
-#doesn't converge either
-
-can_AS=glmer(lab_can~zoo_can*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis=="AngelmanSyndrome"))
-#only effect of zoo
-
-can_LR=glmer(lab_can~zoo_can*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis!="AngelmanSyndrome"))
-#both main and interaction are sig
-
-```
-
-## Additional analyses
+\hypertarget{additional-analyses}{%
+\subsection{Additional analyses}\label{additional-analyses}}
 
 We additionally explored under what conditions Zooniverse judgments more closely aligned with laboratory judgments.
-In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as "Junk" (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora [@cychosz2019canonical].
-A high proportion of "Junk" may indicate that automated segmentation was errorful for those children, and may be a sign that the rest of the data could be compromised as well.
+In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as \enquote{Junk} (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora (Cychosz et al., 2019).
+A high proportion of \enquote{Junk} may indicate that automated segmentation was errorful for those children, and may be a sign that the rest of the data could be compromised as well.
 
-```{r prelim-junk, echo=F}
-junkXdiag=t.test(proportions$z_junk ~ proportions$Diagnosis)
+\includegraphics{paper_files/figure-latex/cor-junk-1.pdf}
 
-```
-
-```{r cor-junk, echo=F}
-
-mycols=c("blue","darkgreen")
-mypch=c(10,18)
-names(mycols)=names(mypch)=c("lower_junk","higher_junk")
-
-layout(matrix(c(1:2), 1, 2, byrow = F))
-par(mar = c(4, 4, 1, 1))
-
-#create matrix to hold r results
-myr=matrix(NA,nrow=2,ncol=2)
-rownames(myr)<-c("lp","cp")
-colnames(myr)<-c("lower_junk","higher_junk")
-mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
-
-lims=range(c(proportions[,"z_lp"],proportions[,"l_lp"]))
-
-  plot(proportions[,"z_lp"]~proportions[,"l_lp"], xlab=prettynames["l_lp"],ylab=prettynames["z_lp"],
-       xlim=lims,ylim=lims,
-       col=mycols[proportions$median_junk], pch=mypch[proportions$median_junk])
-  lines(c(0,1),c(0,1),lty=2,col="darkgray")
-
-  abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$median_junk=="lower_junk")),col="blue")
-
-  #get cor properties
-  thiscor=cor.test(proportions[proportions$median_junk=="lower_junk","z_lp"],proportions[proportions$median_junk=="lower_junk","l_lp"])
-    myr[1,"lower_junk"]=round(thiscor$estimate,3)
-    mylb[1,"lower_junk"]=round(thiscor$conf.int[1],3)
-    myhb[1,"lower_junk"]=round(thiscor$conf.int[2],3)
-    myn[1,"lower_junk"]=round(thiscor$parameter,3)
-    myp[1,"lower_junk"]=round(thiscor$p.value,3)
-
-  #repeat for LR  
-    abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$median_junk!="lower_junk")),col="darkgreen")
-
-      #get cor properties
-  thiscor=cor.test(proportions[proportions$median_junk!="lower_junk","l_lp"],proportions[proportions$median_junk!="lower_junk","z_lp"])
-    myr[1,"higher_junk"]=round(thiscor$estimate,3)
-    mylb[1,"higher_junk"]=round(thiscor$conf.int[1],3)
-    myhb[1,"higher_junk"]=round(thiscor$conf.int[2],3)
-    myn[1,"higher_junk"]=round(thiscor$parameter,3)
-    myp[1,"higher_junk"]=round(thiscor$p.value,3)
-
-    # repeat whole for cp
-
-    lims=range(c(proportions[,"z_cp"],proportions[,"l_cp"]))
-
-  plot(proportions[,"z_cp"]~proportions[,"l_cp"], xlab=prettynames["l_cp"],ylab=prettynames["z_cp"],
-       xlim=lims,ylim=lims,
-       col=mycols[proportions$median_junk], pch=mypch[proportions$median_junk])
-  lines(c(0,1),c(0,1),lty=2,col="darkgray")
-
-  abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$median_junk=="lower_junk")),col="blue")
-
-  #get cor properties
-  thiscor=cor.test(proportions[proportions$median_junk=="lower_junk","z_cp"],proportions[proportions$median_junk=="lower_junk","l_cp"])
-    myr[2,"lower_junk"]=round(thiscor$estimate,3)
-    mylb[2,"lower_junk"]=round(thiscor$conf.int[1],3)
-    myhb[2,"lower_junk"]=round(thiscor$conf.int[2],3)
-    myn[2,"lower_junk"]=round(thiscor$parameter,3)
-    myp[2,"lower_junk"]=round(thiscor$p.value,3)
-
-  #repeat for higher_junk  
-    abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$median_junk!="lower_junk")),col="darkgreen")
-
-      #get cor properties
-  thiscor=cor.test(proportions[proportions$median_junk!="lower_junk","l_cp"],proportions[proportions$median_junk!="lower_junk","z_cp"])
-    myr[2,"higher_junk"]=round(thiscor$estimate,3)
-    mylb[2,"higher_junk"]=round(thiscor$conf.int[1],3)
-    myhb[2,"higher_junk"]=round(thiscor$conf.int[2],3)
-    myn[2,"higher_junk"]=round(thiscor$parameter,3)
-    myp[2,"higher_junk"]=round(thiscor$p.value,3)
-
-```
-
-We investigated this hypothesis by calculating the proportion of their data labeled as "Junk" for each individual child.
-There was no significant difference in the proportions of their data labeled as "Junk" for older children with Angelman Syndrome (M = `r round(junkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the younger low risk children (M = `r round(junkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r junkXdiag$parameter`) = `r round(junkXdiag$statistic,3)`, p = `r round(junkXdiag$p.value,3)`.
-We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk".
+We investigated this hypothesis by calculating the proportion of their data labeled as \enquote{Junk} for each individual child.
+There was no significant difference in the proportions of their data labeled as \enquote{Junk} for older children with Angelman Syndrome (M = 0.13) compared to the younger low risk children (M = 0.14): Welch's t(13.37) = -0.52, p = 0.61.
+We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as \enquote{Junk}.
 Results were similar across these two post-hoc subgroups.
-For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower junk group was r(`r myn["lp","lower_junk"]`) = `r myr["lp","lower_junk"]`, CI [`r mylb["lp","lower_junk"]`,`r myhb["lp","lower_junk"]`], p=`r myp["lp","lower_junk"]`]; and for the higher junk group it was r(`r myn["lp","higher_junk"]`) = `r myr["lp","higher_junk"]`, CI [`r mylb["lp","higher_junk"]`,`r myhb["lp","higher_junk"]`], p=`r myp["lp","higher_junk"]`.
-For Canonical Proportion, the correlation across lab and Zooniverse data for the lower junk group was r(`r myn["cp","lower_junk"]`) = `r myr["cp","lower_junk"]`, CI [`r mylb["cp","lower_junk"]`,`r myhb["cp","lower_junk"]`], p=`r myp["cp","lower_junk"]`]; and for the higher junk group it was r(`r myn["cp","higher_junk"]`) = `r myr["cp","higher_junk"]`, CI [`r mylb["cp","higher_junk"]`,`r myhb["cp","higher_junk"]`], p=`r myp["cp","higher_junk"]`.
-Thus, it does not seem that a higher proportion of "Junk" judgments is an index of low quality data.
+For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower junk group was r(8) = 0.87, CI {[}0.53,0.97{]}, p=0.00{]}; and for the higher junk group it was r(8) = 0.87, CI {[}0.54,0.97{]}, p=0.00.
+For Canonical Proportion, the correlation across lab and Zooniverse data for the lower junk group was r(8) = 0.96, CI {[}0.82,0.99{]}, p=0{]}; and for the higher junk group it was r(8) = 0.93, CI {[}0.73,0.98{]}, p=0.
+Thus, it does not seem that a higher proportion of \enquote{Junk} judgments is an index of low quality data.
 
 Above, we concluded that derived metrics seem more promising than segment-level data.
-Notice that derived metrics do not require matching of clips (500 ms presented to Zooniverse participants) to segments (the original LENA^TM^ segments presented to laboratory participants).
+Notice that derived metrics do not require matching of clips (500 ms presented to Zooniverse participants) to segments (the original LENA\textsuperscript{TM} segments presented to laboratory participants).
 As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
@@
 
-```{r bychunk-age, echo=F}
-read.csv("../Derived_Data/chunk_majority_agreement_1751.csv")->chunk
-bychild=table(chunk$Answer,chunk$ChildID)
-cans=bychild["Canonical",]
-noncans=bychild["Non-Canonical",]
-junk=bychild["Junk",]
-sum_bychild=table(chunk$ChildID)
+Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman Syndrome group: r(8) = -0.04, CI {[}-0.65,0.61{]}, p=0.92{]}, but increased with age in the low-risk control group: r(8) = 0.72, CI {[}0.16,0.93{]}, p=0.02.
+Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman Syndrome r(8) = -0.38, CI {[}-0.81,0.33{]}, p=0.28{]}; and developmental increases among low-risk control r(8) = 0.57, CI {[}-0.09,0.88{]}, p=0.08.
 
-ling_prop=(cans+noncans)/(sum_bychild-junk)
-can_prop=cans/(cans+noncans)
-
-#check that they match before merging
-#demo_data$ChildID==names(ling_prop)
-#demo_data$ChildID==names(can_prop)
-
-props_bychild_chunk=cbind(ling_prop,can_prop,demo_data)
-colnames(props_bychild_chunk)[1:2]<-c("chi","ling_prop")
-
-z_lp_as=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
-
-z_lp_lr=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
-
-z_cp_as=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
-
-z_cp_lr=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
-```
-
-Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman Syndrome group: r(`r z_lp_as$parameter`) = `r z_lp_as$estimate`, CI [`r z_lp_as$conf.int[1]`,`r z_lp_as$conf.int[2]`], p=`r z_lp_as$p.value`], but increased with age  in the low-risk control group: r(`r z_lp_lr$parameter`) = `r z_lp_lr$estimate`, CI [`r z_lp_lr$conf.int[1]`,`r z_lp_lr$conf.int[2]`], p=`r z_lp_lr$p.value`.
-Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman Syndrome r(`r z_cp_as$parameter`) = `r z_cp_as$estimate`, CI [`r z_cp_as$conf.int[1]`,`r z_cp_as$conf.int[2]`], p=`r z_cp_as$p.value`]; and developmental increases among low-risk control r(`r z_cp_lr$parameter`) = `r z_cp_lr$estimate`, CI [`r z_cp_lr$conf.int[1]`,`r z_cp_lr$conf.int[2]`], p=`r z_cp_lr$p.value`.
-
-```{r zoo-lab-chunk, echo=F}
-
-#create matrix to hold r results
-myr=matrix(NA,nrow=2,ncol=3)
-rownames(myr)<-c("lp","cp")
-colnames(myr)<-c("AS","LR","all")
-mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
-
-proportions=proportions[order(proportions$ChildID),]
-
-  #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
-    myr[1,"AS"]=round(thiscor$estimate,3)
-    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
-    myn[1,"AS"]=round(thiscor$parameter,3)
-    myp[1,"AS"]=round(thiscor$p.value,3)
-
-  #repeat for LR
-
-      #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"])
-    myr[1,"LR"]=round(thiscor$estimate,3)
-    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
-    myn[1,"LR"]=round(thiscor$parameter,3)
-    myp[1,"LR"]=round(thiscor$p.value,3)
-
-#repeat for all cor
-
-      thiscor=cor.test(props_bychild_chunk$ling_prop,proportions$l_lp)
-    myr[1,"all"]=round(thiscor$estimate,3)
-    mylb[1,"all"]=round(thiscor$conf.int[1],3)
-    myhb[1,"all"]=round(thiscor$conf.int[2],3)
-    myn[1,"all"]=round(thiscor$parameter,3)
-    myp[1,"all"]=round(thiscor$p.value,3)
-
-    # repeat whole for cp
-
-  #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
-    myr[2,"AS"]=round(thiscor$estimate,3)
-    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
-    myn[2,"AS"]=round(thiscor$parameter,3)
-    myp[2,"AS"]=round(thiscor$p.value,3)
-
-  #repeat for LR
-
-      #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"])
-    myr[2,"LR"]=round(thiscor$estimate,3)
-    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
-    myn[2,"LR"]=round(thiscor$parameter,3)
-    myp[2,"LR"]=round(thiscor$p.value,3)
-
-#repeat for all cor
-
-      thiscor=cor.test(props_bychild_chunk$can_prop,proportions$l_cp)
-    myr[2,"all"]=round(thiscor$estimate,3)
-    mylb[2,"all"]=round(thiscor$conf.int[1],3)
-    myhb[2,"all"]=round(thiscor$conf.int[2],3)
-    myn[2,"all"]=round(thiscor$parameter,3)
-    myp[2,"all"]=round(thiscor$p.value,3)
-
- mycortab=rbind(mycortab,t(myr))
-rownames(mycortab)[4:6]<-paste("Chunks",rownames(t(myr)))
-
-```
-
-As for correlations between Zooniverse and laboratory-derived metrics, we observe very similar levels of correlation: Linguistic Proportion overall r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman Syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
-As for Canonical Proportion overall r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`; Angelman Syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`];  low-risk control infants r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
+As for correlations between Zooniverse and laboratory-derived metrics, we observe very similar levels of correlation: Linguistic Proportion overall r(18) = 0.84, CI {[}0.64,0.94{]}, p=0{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman Syndrome r(8) = 0.84, CI {[}0.46,0.96{]}, p=0.00{]}; low-risk control r(8) = 0.84, CI {[}0.45,0.96{]}, p=0.00{]}.
+As for Canonical Proportion overall r(18) = 0.96, CI {[}0.91,0.98{]}, p=0; Angelman Syndrome: r(8) = 0.87, CI {[}0.52,0.97{]}, p=0.00{]}; low-risk control infants r(8) = 0.98, CI {[}0.89,0.99{]}, p=0.
 
 Since all of these results are very similar to those obtained when first matching clips to segments, we conclude that in the future this step may not be necessary.
-Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' clip level judgements [which was in fact what @cychosz2019canonical did].
-
-```{r downsample}
-
-#we start from the data base that has all the info
-nsegs=data.frame(table(data_all$ChildID))
-colnames(nsegs)<-c("ChildID","nsegs")
-
-merge(nsegs,demo_data)->nsegs
-nseg_t=t.test(nsegs$nsegs~nsegs$Diagnosis)
-
-resampling=matrix(NA,nrow=50,ncol=6)
-colnames(resampling)<-c("lp_r_all","lp_r_as","lp_r_lr","cp_r_all","cp_r_as","cp_r_lr")
-for(i in 1:50){
-  sel_seg=NULL
-  for(thischi in levels(data_all$ChildID)) sel_seg<-c(sel_seg,as.character(data_all$segmentId_DB[sample(data_all$segmentId_DB[data_all$ChildID==thischi],100)]))
-
-  sampling=data_all[as.character(data_all$segmentId_DB) %in% as.character(sel_seg),]
-
-  #get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
-  ztab=table(sampling$ChildID,sampling$Zoon_classif)
-  z_lp=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
-  z_cp=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
-
-  ltab=table(sampling$ChildID,sampling$lab)
-  l_lp=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
-  l_cp=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
-
-  thistab=cbind(demo_data,z_lp,z_cp,l_lp,l_cp)
-
-  resampling[i,"lp_r_all"]=cor.test(thistab$z_lp,thistab$l_lp)$estimate
-  resampling[i,"lp_r_as"]=cor.test(thistab$z_lp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
-  resampling[i,"lp_r_lr"]=cor.test(thistab$z_lp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate
-
-   resampling[i,"cp_r_all"]=cor.test(thistab$z_cp,thistab$l_cp)$estimate
-  resampling[i,"cp_r_as"]=cor.test(thistab$z_cp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
-  resampling[i,"cp_r_lr"]=cor.test(thistab$z_cp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate
- }
-myr=colMeans(resampling)
-mycortab=rbind(mycortab,cbind(myr[c("lp_r_as" ,"lp_r_lr","lp_r_all")],myr[c("cp_r_as" ,"cp_r_lr","cp_r_all")]))
-rownames(mycortab)[7:9]<-c("100 seg AS","100 seg LR","100 seg all")
-round(mycortab,3)->mycortab
-
-resampling=data.frame(resampling)
-```
+Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' clip level judgements (which was in fact what Cychosz et al., 2019 did).
 
 Next, we looked at whether having more segments from each child may lead to more reliable metrics.
-We observed a non-significant trend [t(`r nseg_t$parameter`)=`r nseg_t$estimate`] for lower number of segments in the Angelman Syndrome group (mean `r nseg_t$estimate[1]`, range `r min(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`) than among low-risk control infants (mean `r nseg_t$estimate[2]`, range `r min(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`), likely because of the lower volubility of the former children.
-Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications [CITE MEG].
-We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children [following CITE MEG].
+We observed a non-significant trend {[}t(17.92)=527.40 and 631.90{]} for lower number of segments in the Angelman Syndrome group (mean 527.40, range 326-924) than among low-risk control infants (mean 631.90, range 351-1074), likely because of the lower volubility of the former children.
+Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications {[}CITE MEG{]}.
+We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children {[}following CITE MEG{]}.
 We randomly sampled 100 segments from each child's data, and recalculated the linguistic and Canonical Proportions based only on these associated judgments.
@@
 We also observed quite a bit of variance across the 50 runs.
-For Linguistic Proportion, the range of correlations found for all infants was [`r min(resampling$lp_r_all)`,`r max(resampling$lp_r_all)`]; for the Angelman Syndrome group [`r min(resampling$lp_r_as)`,`r max(resampling$lp_r_as)`]; for the low-risk group [`r min(resampling$lp_r_lr)`,`r max(resampling$lp_r_lr)`].
-For Canonical Proportion, the range of correlations found for all infants was [`r min(resampling$cp_r_all)`,`r max(resampling$cp_r_all)`]; for the Angelman Syndrome group [`r min(resampling$cp_r_as)`,`r max(resampling$cp_r_as)`]; for the low-risk group [`r min(resampling$cp_r_lr)`,`r max(resampling$cp_r_lr)`].
+For Linguistic Proportion, the range of correlations found for all infants was {[}0.64,0.88{]}; for the Angelman Syndrome group {[}0.76,0.94{]}; for the low-risk group {[}0.45,0.95{]}.
+For Canonical Proportion, the range of correlations found for all infants was {[}0.82,0.96{]}; for the Angelman Syndrome group {[}0.40,0.97{]}; for the low-risk group {[}0.75,0.97{]}.
 
-```{r tab-cors, results="as.is"}
-colnames(mycortab)<-c("Linguistic Proportion","Canonical Proportion")
-kable(mycortab,cap="Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman Syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways: All seg indicates that Zooniverse annotations at the clip level were first combined at the segment level; Chunks that they were analyzed directly at the clip level.
+\begin{table}
+
+\caption{\label{tab:tab-cors}Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman Syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways: All seg indicates that Zooniverse annotations at the clip level were first combined at the segment level; Chunks that they were analyzed directly at the clip level.
 Both of these are based on all the data.
-In contrast, rows labeled 100 seg indicate that laboratory- and Zooniverse-derived metrics were based on only 100 segments (median over 50 runs in which 100 segments were randomly selected from each child).")
-```
+In contrast, rows labeled 100 seg indicate that laboratory- and Zooniverse-derived metrics were based on only 100 segments (median over 50 runs in which 100 segments were randomly selected from each child).}
+\centering
+\begin{tabular}[t]{l|r|r}
+\hline
+  & Linguistic Proportion & Canonical Proportion\\
+\hline
+All seg AS & 0.877 & 0.858\\
+\hline
+All seg LR & 0.823 & 0.960\\
+\hline
+All seg all & 0.811 & 0.937\\
+\hline
+Chunks AS & 0.845 & 0.866\\
+\hline
+Chunks LR & 0.841 & 0.975\\
+\hline
+Chunks all & 0.845 & 0.963\\
+\hline
+100 seg AS & 0.857 & 0.807\\
+\hline
+100 seg LR & 0.806 & 0.892\\
+\hline
+100 seg all & 0.783 & 0.903\\
+\hline
+\end{tabular}
+\end{table}
 
-<!-- - look at whether kids who have a high "junk" proportion according to Zooniverse  -->
-<!--      - have lower levels of concordance across Zoo coders -->
+\hypertarget{discussion}{%
+\subsection{Discussion}\label{discussion}}
 
-<!-- - take all 5 Zoo judgments or 3 or 1 (may need to run the latter two a couple of times to check for potential variance -- if variance is small, we'll just say so, if it's large then we'll have to write a loop and report CIs) -->
-
-## Discussion
-
-In the present study we assessed the extent to which child vocalizations from LENA^TM^ daylong recordings can be accurately described based on classifications collected using the citizen science platform Zooniverse.
+In the present study we assessed the extent to which child vocalizations from LENA\textsuperscript{TM} daylong recordings can be accurately described based on classifications collected using the citizen science platform Zooniverse.
 Our recordings came from children diagnosed with a language disorder, as well as low-risk controls, and included a range of different ages.
-Excerpts of these recordings had previously been annotated by trained experts in the lab, and served as the "gold standard".
+Excerpts of these recordings had previously been annotated by trained experts in the lab, and served as the \enquote{gold standard}.
 
@@
 Nonetheless, inspection of Figure 2 suggests different interpretations for the two derived metrics.
-Overall, the pattern for Linguistic Proportion is compatible with a non-linear trajectory common to the two groups, whereby the Linguistic Proportion increases rapidly up to 20 months, and hovers around 90-100% thereafter.
+Overall, the pattern for Linguistic Proportion is compatible with a non-linear trajectory common to the two groups, whereby the Linguistic Proportion increases rapidly up to 20 months, and hovers around 90-100\% thereafter.
 In this case, the lack of correlation with age in the Angelman Syndrome group suggests that there is little systematic variance in this metric after 20 months of age.
@@
 In other cases, precision was fairly high but recall was lower.
-This was the case of the "Junk/Don't mark" category.
-Higher precision and low recall indicates that when Zooniverse annotations return a "Junk" judgement, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as "Don't mark" in the laboratory.
-Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the "junk" category less) than laboratory annotators, who have access to the whole vocalization.
-It is also plausible, on the other hand, that lab annotators are more "picky" with the data, and tend to reject clips (tag them as "Don't mark").
-Indeed, laboratory coders were instructed to be  sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use "Don't mark" if there was any noise or verlapping speech that would potentially affect acoustic analyses.
+This was the case of the \enquote{Junk/Don't mark} category.
+Higher precision and low recall indicates that when Zooniverse annotations return a \enquote{Junk} judgement, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as \enquote{Don't mark} in the laboratory.
+Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the \enquote{junk} category less) than laboratory annotators, who have access to the whole vocalization.
+It is also plausible, on the other hand, that lab annotators are more \enquote{picky} with the data, and tend to reject clips (tag them as \enquote{Don't mark}).
+Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use \enquote{Don't mark} if there was any noise or verlapping speech that would potentially affect acoustic analyses.
 
@@
 This means that a high proportion of segments detected as these two categories in the lab can be classified in the appropriate category by Zooniverse annotators, but Zooniverse annotators also put into these categories many segments that lab annotators classified in a different manner.
-Most saliently, there is a problematic level of confusion between Crying and Non-Canonical, whereby lab-detected Crying was almost as likely to be classified as such or Non-canonical in Zooniverse (the recall proportion was 49 and 47% respectively), and even worse, 55% of the segments Zooniverse annotators classified as Crying were actually Non-Canonical according to lab annotations.
+Most saliently, there is a problematic level of confusion between Crying and Non-Canonical, whereby lab-detected Crying was almost as likely to be classified as such or Non-canonical in Zooniverse (the recall proportion was 49 and 47\% respectively), and even worse, 55\% of the segments Zooniverse annotators classified as Crying were actually Non-Canonical according to lab annotations.
 This is likely due to the fact that lab annotators had access to the whole segment, and could even listen to the context of that segment, so they could accurately interpret a segment as crying.
@@
 
-### Further research directions
+\hypertarget{further-research-directions}{%
+\subsubsection{Further research directions}\label{further-research-directions}}
 
@@
 
-We believe additional research would be welcome on the two measures we used to describe children’s linguistic productions: Linguistic Proportion and Canonical Proportion.
+We believe additional research would be welcome on the two measures we used to describe children's linguistic productions: Linguistic Proportion and Canonical Proportion.
 Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion.
@@
 Assuming these child-level metrics continue to be supported by such additional data, it would be interesting for future investigations to expand the analysis of Linguistic Proportion and Canonical Proportion to recover differences and similarities in language development between low-risk children and children with disorders other than Angelman Syndrome.
-An extensive body of work has found that children with autism spectrum disorders experience language delays (Fasolo, Majorano, & D’Odorico, 2008; Lang et al., 2019; Patten et al., 2014; Stoel-Gammon, 1989).
+An extensive body of work has found that children with autism spectrum disorders experience language delays (Fasolo, Majorano, \& D'Odorico, 2008; Lang et al., 2019; Patten et al., 2014; Stoel-Gammon, 1989).
 One caveat is that these studies used CBO, which requires recordings to be gathered periodically in order to be determined.
@@
 
-<!-- AC removed: Additionally, we found an imbalance between the quantity of clips assigned to each category.
-Further research should look at validating the results of our approach by oversampling data from non-typically developing children, so as to overcome both the risk of lack of data in some categories, and the imbalance in the quantity of vocalizations between groups.
--->
-
-The lack of age effects in AS seems like the bigger "challenge" - if the goal is to use this information to inform how children are developing, are these the right metrics for that population?
-Longitudinal work within AS may be needed to track the "natural history" of vocalizations to better contextualize how these metrics might be relevant or useful.
+The lack of age effects in AS seems like the bigger \enquote{challenge} - if the goal is to use this information to inform how children are developing, are these the right metrics for that population?
+Longitudinal work within AS may be needed to track the \enquote{natural history} of vocalizations to better contextualize how these metrics might be relevant or useful.
 For example, there is interest in using naturalistic recordings as outcome measures for clinical/drug trials.
-Are we seeing enough variability in AS to assume that these metrics could "budge" in response to a treatment?
+Are we seeing enough variability in AS to assume that these metrics could \enquote{budge} in response to a treatment?
 It seems we are seeing variability across kids, which is promising.
@@
 
-## Conclusion
+\hypertarget{conclusion}{%
+\subsection{Conclusion}\label{conclusion}}
 
@@
 
-# Acknowledgements
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
 
-We are grateful to ...
+We are grateful to \ldots{}
 
@@
 
-# References
+\hypertarget{references}{%
+\section{References}\label{references}}
 
-<!-- These lines ensure references are set with hanging indents in PDF documents; they are ignored in Word. -->
 \setlength{\parindent}{-0.5in}
 \setlength{\leftskip}{0.5in}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-cychosz2019canonical}{}%
+Cychosz, M., Cristia, A., Bergelson, E., Casillas, M., Baudet, G., Warlaumont, A. S., \ldots{} Seidl, A. (2019). Canonical babble development in a large-scale crosslinguistic corpus.
+
+\leavevmode\hypertarget{ref-hamrick2019capturing}{}%
+Hamrick, L., Seidl, A., \& Kelleher, B.
+(n.d.).
+Capturing variability in early language skills among children with angelman syndrome: A novel semi-automated approach.
+\emph{Presented at the Angelman Biomarkers and Outcome Measures Alliance Scientific Meeting}.
+
+\end{document}

### 2020-10-29 202010_jslhr/paper.Rmd
clean manuscript

@@
 BK acknowledges the National Institute of Mental Health (K23MH111955) and the Kinley Trust.
+This publication uses data generated via the Zooniverse.org platform, development of which is funded by generous support, including a Global Impact Award from Google, and by a grant from the Alfred P.
+Sloan Foundation.
 The funders had no impact on this study.
@@
 library(kableExtra)
+library(huxtable) #merging cells in tab1
 library(rel)
@@
 
-TODO
-
-- create OSF for this paper specifically
-- references!
-- merge noncan/can for LENA^TM^ in Table 1
-- appropriately merge col names in tab 1
-- add 2 explo analyses that are still missing
-- Redo pipeline so that it's transparent & integrated
-
-Language production at 3 years of age is the best predictor of later academic and economic achievement (Law et al., 2003).
-Early language delays can adversely affect children’s literacy, behavior, social interaction, and scholastic achievement extending well into adulthood, with early interventions being a better societal investment than later ones (Heckman, 2006).
+Since early language delays can adversely affect children’s literacy, behavior, social interaction, and scholastic achievement extending well into adulthood, early interventions have been described as a better societal investment than later ones [@heckman2006skill].
 Advancements in the field of wearable technologies, such as LENA^TM^ recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally.
 Wearable recorders allow data collection to happen in the child’s natural environment, and at a large scale, which may be particularly helpful for children whose speech is not easily elicited.
-Although such long-form recordings are increasingly common REF, challenges remain with respect to how these data are handled, annotated, and analyzed.
+Although such long-form recordings are increasingly common [@ganek2018language], challenges remain with respect to how these data are handled, annotated, and analyzed [@casillas2019step].
 The present work reports on the validity of labels and child-level descriptors of children's vocalizations gathered from citizen scientists, in a comparison to expert lab annotators.
@@
 However, this also means that there is a great amount of data, and that the audio quality is compromised, as children go about their day transitioning from small to large rooms, with fewer or more people around them, thus providing analysts with an unstable and challenging long audio.
-In fact, human annotation of such audio is estimated to take about 30 times the audio length, with these estimates being greater the more precise one wants or needs to be (Seidl et al., 2019).
+In fact, human annotation of such audio is estimated to take about 30 times the audio length, with these estimates being greater the more precise one wants or needs to be [@casillas2017new].
 It is for this reason that most users of long-form recordings have turned to automatized software for at least a first-pass analysis.
@@
 The most commonly used software was created by the LENA^TM^ Foundation, and it returns a diarization of the audio signal split into key talkers, including the child wearing the device, as well as a split of these audio attributed to the child into three categories: crying, other fixed and vegetative signals, and speech vocalizations.
-These subcategories have not been widely validated (Cristia et al., 2020).
+These subcategories have not been widely validated [@cristia2020accuracy].
 To provide an idea of their accuracy, we classified the same segments that are used in the present study using LENA^TM^'s subtypes[^1] to compared these against the laboratory annotations used in the present study.
@@
 <!-- AS said about not laughing: might be worth noting how low frequency this is and how that might be part of the problem in addition to being difficult to distinguish from crying. -->
-
-[^1]: If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as "junk" or not categorized.
+<!-- ac2as I'm not sure this is needed here -- we do mention that laughing is rare in the next paragraph -->
+
+[^1]: If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as "Junk" or not categorized.
 
@@
 
-read.table("tab-lena-challenge.txt",header=T,sep="\t")->tab1
-colnames(tab1)<-c("","Recall","","","Label","Frequency")
-kable(tab1,cap=mycaption)
-```
-
-Other automatized algorithms have been developed in recent years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical (Schuller et al., 2019).
+read.table("tab-lena-challenge.txt",header=F,sep="\t",skip=1)->tab1
+#olnames(tab1)<-c("","Recall","Recall","Recall","Label Frequency","Label Frequency")
+
+# as_hux(tab1) %>% 
+#     insert_row("","Recall","","","Label Frequency","", fill = "") %>% 
+#     merge_cells(1, 2:4) %>% 
+#     merge_cells(1, 5:6) %>%  
+#     merge_cells(2, 5:6) %>%  
+#     set_header_rows(1, TRUE) %>% 
+#    style_headers(bold = TRUE)
+
+kable(tab1,cap=mycaption, col.names = NULL)%>%
+  collapse_rows(columns = 1:6, valign = "middle") %>%
+add_header_above(c(" " = 1, "Recall"= 3, "Label Frequency" = 2))
+```
+
+Other automatized algorithms have been developed in recent years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical [@schuller2019interspeech].
 The split of speech vocalizations into canonical and non-canonical, and the overt inclusion of laughing, allow a somewhat finer-grained picture of children's vocal activity.
@@
 The ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55% (see Table 1).
-The team who won the challenge improved this by about 2%, primarily through gains in the laughing class obtained by adding training data (REF).
-Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are considered, changes very much.
+The team who won the challenge improved this by about 2%, primarily through gains in the laughing class obtained by adding training data [@yeh2019using].
+Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are weighted based on their frequency of occurrence, changes very much.
 
@@
 For example, most work using LENA^TM^ software reports on children's vocalization counts.
-Although this has been sometimes criticized as being more about quantity than quality REF, this is a promising metric because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA^TM^ REF.
-Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r ~ .3 with standardized language scores in a meta-analysis REF.
+Although this has been sometimes criticized as being more about quantity than quality [@mcdaniel2020predicting], this is a promising metric because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA^TM^ [@cristia2020accuracy].
+Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r ~ .3 with standardized language scores in a meta-analysis [@wang2020meta].
 Thus despite representing a composite of skills, vocalization count may provide a useful estimate of vocalization development.
@@
 <!-- AS said about voc count not sure I agree: https://pubs.asha.org/doi/full/10.1044/2020_JSLHR-19-00281?casa_token=fBByMmPlxbMAAAAA%3AbS2O5Mrymn30_aQWybitwMecENkDYLTjzexCqZcnX2Rx1VJ7OaT9qkIGNAAM95xRBGW_D3XX7mMEWHo -->
+<!-- ac2as fair point -- I'll mention this paper below, when explaining why we may want to go beyond CVC -->
 
@@
 An additional metric that can be derived once children's vocalizations are split into crying, laughing, canonical, and non-canonical is the Linguistic Proportion: the proportion of vocalizations that are linguistic (canonical and non-canonical) out of all vocalizations.
-To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age REF.
+To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age.
 
 <!-- BT said about LP May want to put this 2nd as more exploatory given more lit on can. prop. -->
-
-Yet another metric that could be extracted to estimate relative linguistic complexity is the Canonical Proportion: the proportion of vocalizations that are canonical out of all linguistic (canonical and non-canonical) vocalizations.
-One recent study has established that the Canonical Proportion extracted from daylong recordings is significantly correlated with age in a multicultural and multilingual sample REF.
+<!-- I see the point, but doesn't it make more logical sense to discuss it first? -->
+
+Yet another metric that could be extracted to estimate relative linguistic complexity is the Canonical Proportion: the proportion of vocalizations that contain a canonical transition or syllable out of all linguistic (canonical and non-canonical) vocalizations.
+One recent study has established that the Canonical Proportion extracted from daylong recordings is significantly correlated with age in a multicultural and multilingual sample [@cychosz2019canonical].
 Importantly, this association was established in a sample of children going up to 3 years of age, and thus well beyond the babbling period, and into first words and word combinations.
@@
 There is a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this work focuses on babbling and thus typically on infants under one year of age.
-A critical milestone involves the increasingly common production of Canonical Syllables, consonant-vowel (CV) or vowel-consonant (VC) sequences that resemble those found in adult speech (Oller et al., 1998).
-Given its adult-like CV or VC structure, canonical babble is considered to be a starting point on the path to recognizable speech.
+A critical milestone involves the increasingly common production of Canonical Syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech [@oller1998late].
+Given its adult-like consonant-vowel or vowel-consonant structure, canonical babble is considered to be a starting point on the path to recognizable speech.
 Canonical syllables show a higher complexity given the smooth articulatory transition between a consonant and a vowel (or vice versa), when compared to more primitive sounds such as squeals, or isolated vowels.
-Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development (Oller et al., 1999).
-In addition, delayed onset of canonical babbling has been documented in a wide variety of disorders and delays (Belardi et al., 2017; Fasolo et al., 2008; Lang et al., 2019; Patten et al., 2014; Stoel-Gammon, 2016).
+Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development [@oller2000emergence].
+In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization count in a sample of children with an Autism Spectrum Disorder [@mcdaniel2020predicting].
 
 <!-- AS said: maybe name the disorders & delays here? -->
+<!-- ac2as I removed this because I haven't read the papers, and I'm not sure Chiara has either, so I cannot be sure that the citations are correct -- they were taken from Meg's paper.
+In any case, since this is about onset, which we don't measure here, I don't think this section is needed: In addition, delayed onset of canonical babbling has been documented in a wide variety of disorders and delays (Belardi et al., 2017; Fasolo et al., 2008; Lang et al., 2019; Patten et al., 2014; Stoel-Gammon, 2016).
+-->
 
@@
 The first pertains to how the data are analyzed, with human annotation being costly.
-The second relates to how to extract useful descriptors, both in terms of individual vocalizations and at the child level.
+The second relates to how useful descriptors are extracted, both in terms of individual vocalizations and at the child level.
 
@@
 Crowdsourcing refers to the process whereby a task is solved by a crowd, rather than an individual.
-A number of fields, particularly in the data-driven sciences, have already engaged in the collection of data (including annotations) through crowdsourcing, thanks to its low cost and ecological value (Crump, McDonnell, & Gureckis, 2013; Sescleifer, Francoisse, & Lin, 2018;  Sprouse, 2011).
-For example, a systematic review on crowd-sourced ratings of speech (Sescleifer et al., 2018) found that "lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology"; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81.
+A number of fields, particularly in the data-driven sciences, have already engaged in the collection of data (including annotations) through crowdsourcing, thanks to its low cost and ecological value [@crump2013evaluating;@sescleifer2018systematic].
+For example, a systematic review on crowd-sourced ratings of speech found that "lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology"; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81 [@sescleifer2018systematic].
 On the other hand, the systematic review returned only 8 studies (of which only four were published in peer-reviewed journals), suggesting that there is considerable need for further research on this topic.
 
-Mechanical Turk (MTurk) MTurk is an online labor market created by Amazon to assist "requesters"''" in hiring and paying "workers" for the completion of computerized tasks.
-Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable (Berinsky et al., 2012), some question marks are raised as some "workers'" turn out to be bots, or poorly motivated (and potentially exploited).
-
-A promising crowd-sourcing alternative has arisen in recent years: citizen science, a research technique that engages the public in the collection of scientific information (ref).
+Mechanical Turk (MTurk) MTurk is an online labor market created by Amazon to assist "requesters" in hiring and paying "workers" for the completion of computerized tasks.
+Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable [@berinsky2012evaluating], some question marks are raised as some "workers" turn out to be bots, or are poorly motivated (and potentially exploited) humans.
+
+A promising crowd-sourcing alternative has arisen in recent years: citizen science, a research technique that engages the public in the collection of scientific information.
 As citizen scientists do not receive compensation, this alternative to platforms such as MTurk can overcome the limitations posed by potential exploitation of workforce and/or the use of bots.
@@
 
-One of the most successful platforms hosting citizen science projects is Zooniverse (zooniverse.org).
+One of the most successful platforms hosting citizen science projects is Zooniverse [zooniverse.org; @borne2011zooniverse].
 The website hosts a multitude of interdisciplinary projects that have allowed the public to take part in cutting-edge scientific research, from marine biology to papyrology.
-Zooniverse has proven extremely useful in those fields where the complexity of the data collected is too high to be automatically interpreted using computer algorithms (ref) At the same time, the tasks that human volunteers are asked to complete are sufficiently simple that citizens can carry them out without a background in science or any extensive training (ref).
+Zooniverse has proven extremely useful in those fields where the complexity of the data collected is too high to be automatically interpreted using computer algorithms.
+At the same time, the tasks that human volunteers are asked to complete are sufficiently simple that citizens can carry them out without a background in science or any extensive training.
 
@@
 For the English and Tsimane' corpora, vocalizations were automatically identified using LENA^TM^, whereas the other two were extracted through manual segmentation.
-Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received 3 classifications into crying, laughing, canonical, non-canonical, or "junk" (with the latter tag used for clips that did not contain clear child voice).
+Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received 3 classifications into crying, laughing, canonical, non-canonical, or "Junk" (with the latter tag used for clips that did not contain clear child voice).
 They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgement of canonical or non-canonical.
@@
 The authors also studied how their canonical ratio related to age within each corpus that had a sufficient number of children varying in age.
-They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the "junk" category was very high.
+They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the "Junk" category was very high.
 Additionally, they found that the correlation between canonical ratio and age was much weaker in two corpora where vocalizations were segmented with LENA^TM^ than through manual segmentation.
@@
 We used confusion matrices to describe annotation convergence and divergence patterns, and overall accuracy, kappa, and Gwet's AC1 coefficient as statistical descriptors.
-Given previous results from machine learning classifiers REF, we expected agreement to be moderate at this level, with potentially more erratic results for categories that are rarer.
+Given previous results suggesting that performing this classification based on local acoustic cues is challenging [@Seidl19;@schuller2019interspeech], we expected agreement to be only moderate at this level.
 
@@
 Although we could not rely on previous work to make predictions regarding the Linguistic Proportion, we reasoned that we should observe an increase in the Linguistic Proportion for both populations.
-
 More specific to our research aims, we checked the degree to which laboratory and citizen science annotations converged at the child level by using correlations across the two.
-We additionally explored under what conditions Zooniverse judgments are more closely aligned with laboratory judgments using linear mixed models.
+
+<!-- We additionally explored under what conditions Zooniverse judgments are more closely aligned with laboratory judgments using linear mixed models. -->
 
@@
 
-All analyses and key data are available for reproduction on OSF (https://osf.io/57yha/). Computer code used to collect the Zooniverse judgments is available from GitHub (Semenzin, 2020).
+All analyses and key data are available for reproduction on OSF (https://osf.io/57yha/).
+Computer code used to collect the Zooniverse judgments is available from GitHub [@semenzin2020zooniverse].
+This manuscript is reproducible thanks to the use of RMd [@baumer2015r] and Papaja [@aust2017papaja]  on R [@team2013r].
 
@@
 
-The full data set includes data from 20 children: 10 English-speaking children (6 males, 4 females; age range 11-53 months, mean=41.5 months) diagnosed with Angelman syndrome (AS) and 10 low-risk control (LRC) children (6 males, 4 females; age range 4-18 months, mean=11.7 months).
+The full data set includes data from 20 children: 10 English-speaking children (6 males, 4 females; age range 11-53 months, mean=41.5 months) diagnosed with Angelman syndrome and 10 low-risk control children (6 males, 4 females; age range 4-18 months, mean=11.7 months).
 As children diagnosed with Angelman syndrome typically have severe cognitive and language delays and most do not produce more than 1-2 words consistently, we compare their linguistic production data against that of younger children with theoretically similar language profiles.
@@
 
-Recordings were obtained with the Language ENvironment Analysis (LENA^TM^) Digital Language Processor,  LENA^TM^ is a lightweight recorder (< 60 g, 5.5 cm x 8.5 cm x 1.5 cm) designed to be worn inside a breast-pocket of purpose-made clothing.
+Recordings were obtained with the Language ENvironment Analysis (LENA^TM^) Digital Language Processor,  a lightweight recorder (< 60 g, 5.5 cm x 8.5 cm x 1.5 cm) designed to be worn inside a breast-pocket of purpose-made clothing.
 In our data, every child contributed at least one daylong recording, with one child contributing two recordings (duration 11.58-16.00 hours, mean 15.05 hours).
@@
 The first step we took before beginning the annotation process was the identification of children's vocalizations within day-long recordings.
-We did this automatically, using the LENA^TM^ proprietary software.
+We did this automatically, using the LENA^TM^ proprietary software [@xu2008signal].
 Specifically, the LENA^TM^ software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child’s environment (e.g., Female-Adult, TV).
-Key child labels have a precision of about XX% and a recall of about XX% (ref).
+Key child labels have a precision of about 60% and a recall of about 50% [@cristia2020thorough].
 These stretches of audio assigned to the key child can contain cries, vegetative sounds, and linguistic vocalizations, as well as interstitial pauses.
-We will refer to them as segments, to highlight the fact that they are stretches of audio segmented out as belonging to the key child, but they may not be individual vocalizations.
-
-<!-- BT said Lisa add in how portions of recording were selected (high vol + random) -->
+We will refer to them as *segments*, to highlight the fact that they are stretches of audio segmented out as belonging to the key child, but they may not be individual vocalizations.
 
@@
 A substantial concern that emerges when considering citizen science annotation for spontaneous speech data is the risk of a privacy breach: even short clips can contain personal information that can expose the identity of the speaker, such as addresses or names, or contain sensitive and private information.
-Fortunately, a study by Seidl et al.
-(2019) determined that when segments are divided into shorter chunks (400-600ms), human annotators can code our categories of interest (canonical, laughing, crying, etc.) with a classification quality comparable to the one carried out on full segments.
+Fortunately, @Seidl19 determined that when segments are divided into shorter chunks (400-600ms), human annotators with little training can code our categories of interest (canonical, laughing, crying, etc.) with a classification quality comparable to the one carried out on full segments.
 We therefore used very short chunks (500ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.
@@
 To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms clips and their corresponding segment.
-Clips were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling (https://github.com/zooniverse/panoptes-cli).
+Clips were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling [@bowyer2015panoptes].
 
@@
 
-Users were asked to assign each 500ms clip to one out of five possible categories: (1) canonical, (2) non-canonical, (3) crying, (4) laughing, and (5) junk (overlapping speech, non-infant speech, silence, external sounds).
+Users were asked to assign each 500ms clip to one out of five possible categories: (1) canonical, (2) non-canonical, (3) crying, (4) laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
 Notice that the latter is narrower than the "Don't mark" category of the lab annotation routine, which was used any time there was noise or overlap that could affect acoustic analyses.
@@
 Majority agreement is thus achieved when 3 out of 5 judgements agree (rather than 2 out of 3).
-In exploratory analyses, we assess the impact of collecting a smaller number of judgments per clip.
+
+<!-- In exploratory analyses, we assess the impact of collecting a smaller number of judgments per clip. -->
 
@@
 ### Data post-processing
+<!-- ac this section still has numbers typed in  -->
 
@@
 
-We did this by considering the majority label extracted from Zooniverse judgements for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were junk, then the segment was labelled as junk; then, following a hierarchy (Canonical>Non-canonical>Laughing>Crying), segments with at least one instance of each judgement were labelled as such: e.g.
-if one canonical judgement was present, then the segment was labelled as canonical, otherwise if a non-canonical judgement was present, the segment was labelled as non-canonical, and so on.
+We did this by considering the majority label extracted from Zooniverse judgements for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were Junk, then the segment was labelled as Junk; then, following a hierarchy (Canonical>Non-canonical>Crying>Laughing), segments with at least one instance of each judgement were labelled as such:
+
+- if one canonical judgement was present, then the segment was labelled as Canonical, 
+- else, if a Non-Canonical judgement was present, the segment was labelled as Non-Canonical, 
+- else, if a Crying judgment was present, the segment was labelled as Crying,
+- else, the segment was labelled as Laughing.
+
 This is essentially the same thought process expert annotators followed when labeling segment-level audio, as each segment can represent multiple classification subtypes).
@@
 
-<!-- AS said I feel like a figure might help here. -->
-
-```{r, echo=F}
+```{r clean up, echo=F}
 # remove non-majority labels
@@
 
-#and also get junk from zooniverse
-z_junk=ztab[,c("Junk")]/rowSums(ztab)
+#and also get Junk from zooniverse
+z_Junk=ztab[,c("Junk")]/rowSums(ztab)
 
 #put all the proportions together
-if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) proportions=cbind(rownames(ztab),z_lp,z_cp,l_lp,l_cp,z_junk) else print("oops this code needs to be more complex because we don't have the same kids for the two proportions")
+if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) proportions=cbind(rownames(ztab),z_lp,z_cp,l_lp,l_cp,z_Junk) else print("oops this code needs to be more complex because we don't have the same kids for the two proportions")
 colnames(proportions)[1]<-"filename"
@@
 #cbinding results in text, so we numerize the proportions
-for(thisvar in c("z_lp","z_cp","l_lp","l_cp","z_junk")) proportions[,thisvar]=as.numeric(as.character(proportions[,thisvar]))
-
-#add median junk
-proportions$median_junk<-ifelse(proportions$z_junk<median(proportions$z_junk),"lower_junk","higher_junk")
+for(thisvar in c("z_lp","z_cp","l_lp","l_cp","z_Junk")) proportions[,thisvar]=as.numeric(as.character(proportions[,thisvar]))
+
+#add median Junk
+proportions$median_Junk<-ifelse(proportions$z_Junk<median(proportions$z_Junk),"lower_Junk","higher_Junk")
 
@@
 ### Descriptive analyses
-
-<!-- lh said I think you mention above that individuals with AS rarely use more than a handful of meaningful words - if we decide to go with just looking at these analyses broken down by group, that might be helpful to reiterate here -->
-<!-- ac said in response: why do you talk about meaningful words? These analyses do not separate words from the rest... -->
 
@@
 According to lab annotators, `r table_lab["Canonical"]`% of segments were canonical, `r table_lab["Non-Canonical"]`% non-canonical, `r table_lab["Laughing"]`% laughing, and `r table_lab["Crying"]`% crying, with the remaining `r table_lab["Junk"]`% being categorized as "Don't mark".
-Zooniverse data revealed a similar distribution:  `r table_zoon["Canonical"]`% canonical, `r table_zoon["Non-Canonical"]`% non-canonical, `r table_zoon["Laughing"]`% laughing, `r table_zoon["Crying"]`% crying, `r table_zoon["Junk"]`% junk.
-Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = ("Canonical"+"Non-Canonical")/"All vocalizations" (i.e., we remove, junk), and ii) Canonical Proportion = "Canonical"/("Canonical"+"Non-Canonical") (i.e., we remove junk + non-linguistic).
+Zooniverse data revealed a similar distribution:  `r table_zoon["Canonical"]`% canonical, `r table_zoon["Non-Canonical"]`% non-canonical, `r table_zoon["Laughing"]`% laughing, `r table_zoon["Crying"]`% crying, `r table_zoon["Junk"]`% Junk.
+Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = ("Canonical"+"Non-Canonical")/"All vocalizations" (i.e., we remove, Junk), and ii) Canonical Proportion = "Canonical"/("Canonical"+"Non-Canonical") (i.e., we remove Junk + non-linguistic).
 See Figure 2 for results.
@@
 
-We additionally explored under what conditions Zooniverse judgments more closely aligned with laboratory judgments.
+In this section, we report on explorations of under what conditions Zooniverse judgments more closely aligned with laboratory judgments.
 In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as "Junk" (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora [@cychosz2019canonical].
@@
 
-```{r prelim-junk, echo=F}
-junkXdiag=t.test(proportions$z_junk ~ proportions$Diagnosis)
-
-```
-
-```{r cor-junk, echo=F}
+```{r prelim-Junk, echo=F}
+JunkXdiag=t.test(proportions$z_Junk ~ proportions$Diagnosis)
+
+```
+
+```{r cor-Junk, echo=F}
 
@@
 mypch=c(10,18)
-names(mycols)=names(mypch)=c("lower_junk","higher_junk")
+names(mycols)=names(mypch)=c("lower_Junk","higher_Junk")
 
@@
 rownames(myr)<-c("lp","cp")
-colnames(myr)<-c("lower_junk","higher_junk")
+colnames(myr)<-c("lower_Junk","higher_Junk")
 mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
@@
        xlim=lims,ylim=lims,
-       col=mycols[proportions$median_junk], pch=mypch[proportions$median_junk])
+       col=mycols[proportions$median_Junk], pch=mypch[proportions$median_Junk])
   lines(c(0,1),c(0,1),lty=2,col="darkgray")
 
-  abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$median_junk=="lower_junk")),col="blue")
-
-  #get cor properties
-  thiscor=cor.test(proportions[proportions$median_junk=="lower_junk","z_lp"],proportions[proportions$median_junk=="lower_junk","l_lp"])
-    myr[1,"lower_junk"]=round(thiscor$estimate,3)
-    mylb[1,"lower_junk"]=round(thiscor$conf.int[1],3)
-    myhb[1,"lower_junk"]=round(thiscor$conf.int[2],3)
-    myn[1,"lower_junk"]=round(thiscor$parameter,3)
-    myp[1,"lower_junk"]=round(thiscor$p.value,3)
+  abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$median_Junk=="lower_Junk")),col="blue")
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$median_Junk=="lower_Junk","z_lp"],proportions[proportions$median_Junk=="lower_Junk","l_lp"])
+    myr[1,"lower_Junk"]=round(thiscor$estimate,3)
+    mylb[1,"lower_Junk"]=round(thiscor$conf.int[1],3)
+    myhb[1,"lower_Junk"]=round(thiscor$conf.int[2],3)
+    myn[1,"lower_Junk"]=round(thiscor$parameter,3)
+    myp[1,"lower_Junk"]=round(thiscor$p.value,3)
 
   #repeat for LR  
-    abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$median_junk!="lower_junk")),col="darkgreen")
-
-      #get cor properties
-  thiscor=cor.test(proportions[proportions$median_junk!="lower_junk","l_lp"],proportions[proportions$median_junk!="lower_junk","z_lp"])
-    myr[1,"higher_junk"]=round(thiscor$estimate,3)
-    mylb[1,"higher_junk"]=round(thiscor$conf.int[1],3)
-    myhb[1,"higher_junk"]=round(thiscor$conf.int[2],3)
-    myn[1,"higher_junk"]=round(thiscor$parameter,3)
-    myp[1,"higher_junk"]=round(thiscor$p.value,3)
+    abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$median_Junk!="lower_Junk")),col="darkgreen")
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$median_Junk!="lower_Junk","l_lp"],proportions[proportions$median_Junk!="lower_Junk","z_lp"])
+    myr[1,"higher_Junk"]=round(thiscor$estimate,3)
+    mylb[1,"higher_Junk"]=round(thiscor$conf.int[1],3)
+    myhb[1,"higher_Junk"]=round(thiscor$conf.int[2],3)
+    myn[1,"higher_Junk"]=round(thiscor$parameter,3)
+    myp[1,"higher_Junk"]=round(thiscor$p.value,3)
 
@@
        xlim=lims,ylim=lims,
-       col=mycols[proportions$median_junk], pch=mypch[proportions$median_junk])
+       col=mycols[proportions$median_Junk], pch=mypch[proportions$median_Junk])
   lines(c(0,1),c(0,1),lty=2,col="darkgray")
 
-  abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$median_junk=="lower_junk")),col="blue")
-
-  #get cor properties
-  thiscor=cor.test(proportions[proportions$median_junk=="lower_junk","z_cp"],proportions[proportions$median_junk=="lower_junk","l_cp"])
-    myr[2,"lower_junk"]=round(thiscor$estimate,3)
-    mylb[2,"lower_junk"]=round(thiscor$conf.int[1],3)
-    myhb[2,"lower_junk"]=round(thiscor$conf.int[2],3)
-    myn[2,"lower_junk"]=round(thiscor$parameter,3)
-    myp[2,"lower_junk"]=round(thiscor$p.value,3)
-
-  #repeat for higher_junk  
-    abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$median_junk!="lower_junk")),col="darkgreen")
-
-      #get cor properties
-  thiscor=cor.test(proportions[proportions$median_junk!="lower_junk","l_cp"],proportions[proportions$median_junk!="lower_junk","z_cp"])
-    myr[2,"higher_junk"]=round(thiscor$estimate,3)
-    mylb[2,"higher_junk"]=round(thiscor$conf.int[1],3)
-    myhb[2,"higher_junk"]=round(thiscor$conf.int[2],3)
-    myn[2,"higher_junk"]=round(thiscor$parameter,3)
-    myp[2,"higher_junk"]=round(thiscor$p.value,3)
+  abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$median_Junk=="lower_Junk")),col="blue")
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$median_Junk=="lower_Junk","z_cp"],proportions[proportions$median_Junk=="lower_Junk","l_cp"])
+    myr[2,"lower_Junk"]=round(thiscor$estimate,3)
+    mylb[2,"lower_Junk"]=round(thiscor$conf.int[1],3)
+    myhb[2,"lower_Junk"]=round(thiscor$conf.int[2],3)
+    myn[2,"lower_Junk"]=round(thiscor$parameter,3)
+    myp[2,"lower_Junk"]=round(thiscor$p.value,3)
+
+  #repeat for higher_Junk  
+    abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$median_Junk!="lower_Junk")),col="darkgreen")
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$median_Junk!="lower_Junk","l_cp"],proportions[proportions$median_Junk!="lower_Junk","z_cp"])
+    myr[2,"higher_Junk"]=round(thiscor$estimate,3)
+    mylb[2,"higher_Junk"]=round(thiscor$conf.int[1],3)
+    myhb[2,"higher_Junk"]=round(thiscor$conf.int[2],3)
+    myn[2,"higher_Junk"]=round(thiscor$parameter,3)
+    myp[2,"higher_Junk"]=round(thiscor$p.value,3)
 
@@
 We investigated this hypothesis by calculating the proportion of their data labeled as "Junk" for each individual child.
-There was no significant difference in the proportions of their data labeled as "Junk" for older children with Angelman Syndrome (M = `r round(junkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the younger low risk children (M = `r round(junkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r junkXdiag$parameter`) = `r round(junkXdiag$statistic,3)`, p = `r round(junkXdiag$p.value,3)`.
+There was no significant difference in the proportions of their data labeled as "Junk" for older children with Angelman Syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the younger low risk children (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, p = `r round(JunkXdiag$p.value,3)`.
 We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk".
 Results were similar across these two post-hoc subgroups.
-For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower junk group was r(`r myn["lp","lower_junk"]`) = `r myr["lp","lower_junk"]`, CI [`r mylb["lp","lower_junk"]`,`r myhb["lp","lower_junk"]`], p=`r myp["lp","lower_junk"]`]; and for the higher junk group it was r(`r myn["lp","higher_junk"]`) = `r myr["lp","higher_junk"]`, CI [`r mylb["lp","higher_junk"]`,`r myhb["lp","higher_junk"]`], p=`r myp["lp","higher_junk"]`.
-For Canonical Proportion, the correlation across lab and Zooniverse data for the lower junk group was r(`r myn["cp","lower_junk"]`) = `r myr["cp","lower_junk"]`, CI [`r mylb["cp","lower_junk"]`,`r myhb["cp","lower_junk"]`], p=`r myp["cp","lower_junk"]`]; and for the higher junk group it was r(`r myn["cp","higher_junk"]`) = `r myr["cp","higher_junk"]`, CI [`r mylb["cp","higher_junk"]`,`r myhb["cp","higher_junk"]`], p=`r myp["cp","higher_junk"]`.
+For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["lp","lower_Junk"]`) = `r myr["lp","lower_Junk"]`, CI [`r mylb["lp","lower_Junk"]`,`r myhb["lp","lower_Junk"]`], p=`r myp["lp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["lp","higher_Junk"]`) = `r myr["lp","higher_Junk"]`, CI [`r mylb["lp","higher_Junk"]`,`r myhb["lp","higher_Junk"]`], p=`r myp["lp","higher_Junk"]`.
+For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["cp","lower_Junk"]`) = `r myr["cp","lower_Junk"]`, CI [`r mylb["cp","lower_Junk"]`,`r myhb["cp","lower_Junk"]`], p=`r myp["cp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["cp","higher_Junk"]`) = `r myr["cp","higher_Junk"]`, CI [`r mylb["cp","higher_Junk"]`,`r myhb["cp","higher_Junk"]`], p=`r myp["cp","higher_Junk"]`.
 Thus, it does not seem that a higher proportion of "Junk" judgments is an index of low quality data.
@@
 noncans=bychild["Non-Canonical",]
-junk=bychild["Junk",]
+Junk=bychild["Junk",]
 sum_bychild=table(chunk$ChildID)
 
-ling_prop=(cans+noncans)/(sum_bychild-junk)
+ling_prop=(cans+noncans)/(sum_bychild-Junk)
 can_prop=cans/(cans+noncans)
@@
 We observed a non-significant trend [t(`r nseg_t$parameter`)=`r nseg_t$estimate`] for lower number of segments in the Angelman Syndrome group (mean `r nseg_t$estimate[1]`, range `r min(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`) than among low-risk control infants (mean `r nseg_t$estimate[2]`, range `r min(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`), likely because of the lower volubility of the former children.
-Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications [CITE MEG].
-We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children [following CITE MEG].
+Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications [@cychosz2019canonical].
+We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children [following @cychosz2019canonical].
 We randomly sampled 100 segments from each child's data, and recalculated the linguistic and Canonical Proportions based only on these associated judgments.
@@
 
-<!-- - look at whether kids who have a high "junk" proportion according to Zooniverse  -->
+<!-- - look at whether kids who have a high "Junk" proportion according to Zooniverse  -->
 <!--      - have lower levels of concordance across Zoo coders -->
@@
 Higher precision and low recall indicates that when Zooniverse annotations return a "Junk" judgement, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as "Don't mark" in the laboratory.
-Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the "junk" category less) than laboratory annotators, who have access to the whole vocalization.
+Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the "Junk" category less) than laboratory annotators, who have access to the whole vocalization.
 It is also plausible, on the other hand, that lab annotators are more "picky" with the data, and tend to reject clips (tag them as "Don't mark").
@@
 Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion.
-We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant for other disorders that are related to emotional disorders (REF REF).
-Additionally, while Canonical Proportion has been studied in much previous work, we believe further research is necessary using larger scales (e.g., sampling from daylong recordings to have a more ecological picture of the child, REF; increasing the sheer number of children studied, to have sufficient power to detect effects) and with broader coverage (notably including children growing up in a broad range of languages and cultures, REF), to make sure that assumptions regarding cross-linguistic and maturational universality are justified (REF).
+We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant for other disorders that are related to emotional disorders [a review in @halpern2016excessive].
+Additionally, while Canonical Proportion has been studied in much previous work, we believe further research is necessary using larger scales (saliently increasing the sheer number of children studied, to have sufficient power to detect group effects) and with broader coverage [notably including children growing up in a broad range of languages and cultures, as in @cychosz2019canonical], to make sure that assumptions regarding cross-linguistic and maturational universality are amply justified and not merely a result of having low power to detect differences, or simply sampling from similar populations.
 
 Assuming these child-level metrics continue to be supported by such additional data, it would be interesting for future investigations to expand the analysis of Linguistic Proportion and Canonical Proportion to recover differences and similarities in language development between low-risk children and children with disorders other than Angelman Syndrome.
-An extensive body of work has found that children with autism spectrum disorders experience language delays (Fasolo, Majorano, & D’Odorico, 2008; Lang et al., 2019; Patten et al., 2014; Stoel-Gammon, 1989).
-One caveat is that these studies used CBO, which requires recordings to be gathered periodically in order to be determined.
-
-It is also important to note that work on non-typically developing children poses in itself some challenges.
-Our data contained a lower number of vocalizations for children diagnosed with Angelman Syndrome, as they tend to produce fewer vocalizations than low-risk controls REF.
-Notice this was the case even in our data, where we selected much younger low-risk control childrens as the comparison point.
+That is, it is at present unclear whether relatively lower Canonical Proportions will also be found in other populations, or whether this is more specific to Angelman Syndrome.
+For instance, some previous work has found differences in babbling even when comparing late talkers with toddlers with no specific diagnosis  [@fasolo2008babbling].
+We thus encourage further research inspecting Canonical Proportion in other populations, as well as methodological work that attempts to generate metrics that are similarly easy to gather at scale.
+
+It is also important to note that work on children with non-normative development poses in itself some challenges.
+Our data contained a lower number of vocalizations for children diagnosed with Angelman Syndrome, even though we selected much younger low-risk control childrens as the comparison group.
+However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
+Now that this method has been validated, researchers can extract all segments that LENA^TM^ identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
 
@@
 
-The lack of age effects in AS seems like the bigger "challenge" - if the goal is to use this information to inform how children are developing, are these the right metrics for that population?
-Longitudinal work within AS may be needed to track the "natural history" of vocalizations to better contextualize how these metrics might be relevant or useful.
-For example, there is interest in using naturalistic recordings as outcome measures for clinical/drug trials.
-Are we seeing enough variability in AS to assume that these metrics could "budge" in response to a treatment?
-It seems we are seeing variability across kids, which is promising.
-We are also seeing agreement across coders in AS files, which is also important.
-Next questions would be whether these metrics change within individuals, and whether rates of change may be a candidate for tracking treatment.
-Graphs that show the two key variables across groups and age may help inform these questions.
+Another open question we hope future work looks into is the assessment of within-participant changes.
+Our dataset only contained one daylong recording per child (with one exception), so other datasets in which a greater number of recordings are collected longitudinally can bring novel insights.
+Is the age-related decline we observed among children in the Angelman Syndrome group truly a developmental decline that happens by necessity in this population, or might it reflect partially other aspects of children's experience and behavior?
+Additionally, longitudinal work within the same infant may be needed to track the "natural history" of vocalizations to better contextualize how these metrics should be interpreted.
+It is also an open question how large must effects be, at the individual level, in order to be detected with this method.
+For example, there is interest in using naturalistic recordings as outcome measures for clinical (behavioral or pharmacological) trials.
+The fact that agreement across annotators is high even in the data coming from children diagnosed with Angelman Syndrome is encouraging, but it would be important to establish whether absolute levels of Canonical Proportion, and/or rates of change over longer time spans, may be good candidates for tracking the effects of any treatment.
 
 Finally, broader generalizations of our technique could be of interest to readers of this paper.
-Any researcher is able to undergo the procedure to create Zooniverse projects.
-For instance, a separate pipeline could be set up to annotate potentially disordered speech by older adults at risk of, or diagnosed with, neurodegenerative diseases.
+Any researcher is able to undergo the procedure to create Zooniverse projects (https://www.zooniverse.org/about/faq).
+For instance, a project could be set up to annotate potentially disordered speech by older adults at risk of, or diagnosed with, neurodegenerative diseases.
 Although it would be important to similarly validate such annotations, we believe there is strong promise for such tasks, because many neurodegenerative disorders affect local aspects of the speech, which are detected at the syllable level.

### 2020-10-29 202010_jslhr/paper.tex
clean manuscript

@@
 
-TODO
-
-\begin{itemize}
-\tightlist
-\item
-  create OSF for this paper specifically
-\item
-  references!
-\item
-  merge noncan/can for LENA\textsuperscript{TM} in Table 1
-\item
-  appropriately merge col names in tab 1
-\item
-  add 2 explo analyses that are still missing
-\item
-  Redo pipeline so that it's transparent \& integrated
-\end{itemize}
-
-Language production at 3 years of age is the best predictor of later academic and economic achievement (Law et al., 2003).
-Early language delays can adversely affect children's literacy, behavior, social interaction, and scholastic achievement extending well into adulthood, with early interventions being a better societal investment than later ones (Heckman, 2006).
+Since early language delays can adversely affect children's literacy, behavior, social interaction, and scholastic achievement extending well into adulthood, early interventions have been described as a better societal investment than later ones (Heckman, 2006).
 Advancements in the field of wearable technologies, such as LENA\textsuperscript{TM} recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally.
 Wearable recorders allow data collection to happen in the child's natural environment, and at a large scale, which may be particularly helpful for children whose speech is not easily elicited.
-Although such long-form recordings are increasingly common REF, challenges remain with respect to how these data are handled, annotated, and analyzed.
+Although such long-form recordings are increasingly common (Ganek \& Eriks-Brophy, 2018), challenges remain with respect to how these data are handled, annotated, and analyzed (Casillas \& Cristia, 2019).
 The present work reports on the validity of labels and child-level descriptors of children's vocalizations gathered from citizen scientists, in a comparison to expert lab annotators.
@@
 However, this also means that there is a great amount of data, and that the audio quality is compromised, as children go about their day transitioning from small to large rooms, with fewer or more people around them, thus providing analysts with an unstable and challenging long audio.
-In fact, human annotation of such audio is estimated to take about 30 times the audio length, with these estimates being greater the more precise one wants or needs to be (Seidl et al., 2019).
+In fact, human annotation of such audio is estimated to take about 30 times the audio length, with these estimates being greater the more precise one wants or needs to be (Casillas et al., 2017).
 It is for this reason that most users of long-form recordings have turned to automatized software for at least a first-pass analysis.
@@
 These subcategories have not been widely validated (Cristia et al., 2020).
-To provide an idea of their accuracy, we classified the same segments that are used in the present study using LENA\textsuperscript{TM}'s subtypes\footnote{If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as \enquote{junk} or not categorized.} to compared these against the laboratory annotations used in the present study.
+To provide an idea of their accuracy, we classified the same segments that are used in the present study using LENA\textsuperscript{TM}'s subtypes\footnote{If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as \enquote{Junk} or not categorized.} to compared these against the laboratory annotations used in the present study.
 Table 1 shows that LENA\textsuperscript{TM} has a good recall for speech-like vocalizations and crying but not laughing.
@@
 \hline
- & Recall &  &  & Label & Frequency\\
-\hline
+\multicolumn{1}{c|}{ } & \multicolumn{3}{c|}{Recall} & \multicolumn{2}{c}{Label Frequency} \\
+\cline{2-4} \cline{5-6}
  & C2019B & LENA & Zoon. & C2019B & Lab\\
@@
 Crying & 70.6 & 78.0 & 49.0 & 669 & 581\\
-\hline
+\cline{1-6}
 Laughing & 41.5 & 12.0 & 56.0 & 149 & 137\\
-\hline
-Non-can. & 24.1 & 73.0 & 77.0 & 4485 & 5512\\
-\hline
-Canonical & 66.4 &  & 59.0 & 1426 & 1715\\
-\hline
+\cline{1-6}
+Non-can. & 24.1 &  & 77.0 & 4485 & 5512\\
+\cline{1-2}
+\cline{4-6}
+Canonical & 66.4 & \multirow{-2}{*}{\raggedright\arraybackslash 73.0} & 59.0 & 1426 & 1715\\
+\cline{1-6}
 Junk & 67.3 & 1.0 & 44.0 & 4575 & 2169\\
-\hline
+\cline{1-6}
 UAR & 54.0 & 41.0 & 57.0 &  & \\
-\hline
-WAR & 49.9 & 57.0 & 65.0 &  & \\
+\cline{1-4}
+WAR & 49.9 & 57.0 & 65.0 & \multirow{-2}{*}{\raggedright\arraybackslash } & \multirow{-2}{*}{\raggedright\arraybackslash }\\
 \hline
@@
 The ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55\% (see Table 1).
-The team who won the challenge improved this by about 2\%, primarily through gains in the laughing class obtained by adding training data (REF).
-Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are considered, changes very much.
+The team who won the challenge improved this by about 2\%, primarily through gains in the laughing class obtained by adding training data (Yeh et al., 2019).
+Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are weighted based on their frequency of occurrence, changes very much.
 
@@
 For example, most work using LENA\textsuperscript{TM} software reports on children's vocalization counts.
-Although this has been sometimes criticized as being more about quantity than quality REF, this is a promising metric because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA\textsuperscript{TM} REF.
-Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r \textasciitilde{} .3 with standardized language scores in a meta-analysis REF.
+Although this has been sometimes criticized as being more about quantity than quality (McDaniel, Yoder, Estes, \& Rogers, 2020), this is a promising metric because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA\textsuperscript{TM} (Cristia et al., 2020).
+Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r \textasciitilde{} .3 with standardized language scores in a meta-analysis (Wang, Williams, Dilley, \& Houston, 2020).
 Thus despite representing a composite of skills, vocalization count may provide a useful estimate of vocalization development.
@@
 An additional metric that can be derived once children's vocalizations are split into crying, laughing, canonical, and non-canonical is the Linguistic Proportion: the proportion of vocalizations that are linguistic (canonical and non-canonical) out of all vocalizations.
-To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age REF.
+To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age.
 
-Yet another metric that could be extracted to estimate relative linguistic complexity is the Canonical Proportion: the proportion of vocalizations that are canonical out of all linguistic (canonical and non-canonical) vocalizations.
-One recent study has established that the Canonical Proportion extracted from daylong recordings is significantly correlated with age in a multicultural and multilingual sample REF.
+Yet another metric that could be extracted to estimate relative linguistic complexity is the Canonical Proportion: the proportion of vocalizations that contain a canonical transition or syllable out of all linguistic (canonical and non-canonical) vocalizations.
+One recent study has established that the Canonical Proportion extracted from daylong recordings is significantly correlated with age in a multicultural and multilingual sample (Cychosz et al., 2019).
 Importantly, this association was established in a sample of children going up to 3 years of age, and thus well beyond the babbling period, and into first words and word combinations.
@@
 There is a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this work focuses on babbling and thus typically on infants under one year of age.
-A critical milestone involves the increasingly common production of Canonical Syllables, consonant-vowel (CV) or vowel-consonant (VC) sequences that resemble those found in adult speech (Oller et al., 1998).
-Given its adult-like CV or VC structure, canonical babble is considered to be a starting point on the path to recognizable speech.
+A critical milestone involves the increasingly common production of Canonical Syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech (Oller, Eilers, Neal, \& Cobo-Lewis, 1998).
+Given its adult-like consonant-vowel or vowel-consonant structure, canonical babble is considered to be a starting point on the path to recognizable speech.
 Canonical syllables show a higher complexity given the smooth articulatory transition between a consonant and a vowel (or vice versa), when compared to more primitive sounds such as squeals, or isolated vowels.
-Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development (Oller et al., 1999).
-In addition, delayed onset of canonical babbling has been documented in a wide variety of disorders and delays (Belardi et al., 2017; Fasolo et al., 2008; Lang et al., 2019; Patten et al., 2014; Stoel-Gammon, 2016).
+Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development (Oller, 2000).
+In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization count in a sample of children with an Autism Spectrum Disorder (McDaniel et al., 2020).
 
@@
 Crowdsourcing refers to the process whereby a task is solved by a crowd, rather than an individual.
-A number of fields, particularly in the data-driven sciences, have already engaged in the collection of data (including annotations) through crowdsourcing, thanks to its low cost and ecological value (Crump, McDonnell, \& Gureckis, 2013; Sescleifer, Francoisse, \& Lin, 2018; Sprouse, 2011).
-For example, a systematic review on crowd-sourced ratings of speech (Sescleifer et al., 2018) found that \enquote{lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology}; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81.
+A number of fields, particularly in the data-driven sciences, have already engaged in the collection of data (including annotations) through crowdsourcing, thanks to its low cost and ecological value (Crump, McDonnell, \& Gureckis, 2013; Sescleifer, Francoisse, \& Lin, 2018).
+For example, a systematic review on crowd-sourced ratings of speech found that \enquote{lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology}; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81 (Sescleifer et al., 2018).
 On the other hand, the systematic review returned only 8 studies (of which only four were published in peer-reviewed journals), suggesting that there is considerable need for further research on this topic.
 
-Mechanical Turk (MTurk) MTurk is an online labor market created by Amazon to assist \enquote{requesters}'\enquote{" in hiring and paying \enquote{workers} for the completion of computerized tasks.
-Although it is a leading crowdsourcing service, and some evidence suggests MTurkers} annotations can be quite reliable (Berinsky et al., 2012), some question marks are raised as some \enquote{workers'} turn out to be bots, or poorly motivated (and potentially exploited).
+Mechanical Turk (MTurk) MTurk is an online labor market created by Amazon to assist \enquote{requesters} in hiring and paying \enquote{workers} for the completion of computerized tasks.
+Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable (Berinsky, Huber, \& Lenz, 2012), some question marks are raised as some \enquote{workers} turn out to be bots, or are poorly motivated (and potentially exploited) humans.
 
-A promising crowd-sourcing alternative has arisen in recent years: citizen science, a research technique that engages the public in the collection of scientific information (ref).
+A promising crowd-sourcing alternative has arisen in recent years: citizen science, a research technique that engages the public in the collection of scientific information.
 As citizen scientists do not receive compensation, this alternative to platforms such as MTurk can overcome the limitations posed by potential exploitation of workforce and/or the use of bots.
@@
 
-One of the most successful platforms hosting citizen science projects is Zooniverse (zooniverse.org).
+One of the most successful platforms hosting citizen science projects is Zooniverse (zooniverse.org; Borne \& Zooniverse Team, 2011).
 The website hosts a multitude of interdisciplinary projects that have allowed the public to take part in cutting-edge scientific research, from marine biology to papyrology.
-Zooniverse has proven extremely useful in those fields where the complexity of the data collected is too high to be automatically interpreted using computer algorithms (ref) At the same time, the tasks that human volunteers are asked to complete are sufficiently simple that citizens can carry them out without a background in science or any extensive training (ref).
+Zooniverse has proven extremely useful in those fields where the complexity of the data collected is too high to be automatically interpreted using computer algorithms.
+At the same time, the tasks that human volunteers are asked to complete are sufficiently simple that citizens can carry them out without a background in science or any extensive training.
 
@@
 For the English and Tsimane' corpora, vocalizations were automatically identified using LENA\textsuperscript{TM}, whereas the other two were extracted through manual segmentation.
-Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received 3 classifications into crying, laughing, canonical, non-canonical, or \enquote{junk} (with the latter tag used for clips that did not contain clear child voice).
+Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received 3 classifications into crying, laughing, canonical, non-canonical, or \enquote{Junk} (with the latter tag used for clips that did not contain clear child voice).
 They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgement of canonical or non-canonical.
@@
 The authors also studied how their canonical ratio related to age within each corpus that had a sufficient number of children varying in age.
-They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the \enquote{junk} category was very high.
+They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the \enquote{Junk} category was very high.
 Additionally, they found that the correlation between canonical ratio and age was much weaker in two corpora where vocalizations were segmented with LENA\textsuperscript{TM} than through manual segmentation.
@@
 We used confusion matrices to describe annotation convergence and divergence patterns, and overall accuracy, kappa, and Gwet's AC1 coefficient as statistical descriptors.
-Given previous results from machine learning classifiers REF, we expected agreement to be moderate at this level, with potentially more erratic results for categories that are rarer.
+Given previous results suggesting that performing this classification based on local acoustic cues is challenging (Schuller et al., 2019; Seidl, Warlaumont, \& Cristia, 2019), we expected agreement to be only moderate at this level.
 
@@
 Although we could not rely on previous work to make predictions regarding the Linguistic Proportion, we reasoned that we should observe an increase in the Linguistic Proportion for both populations.
-
 More specific to our research aims, we checked the degree to which laboratory and citizen science annotations converged at the child level by using correlations across the two.
-We additionally explored under what conditions Zooniverse judgments are more closely aligned with laboratory judgments using linear mixed models.
 
@@
 
-All analyses and key data are available for reproduction on OSF (\url{https://osf.io/57yha/}). Computer code used to collect the Zooniverse judgments is available from GitHub (Semenzin, 2020).
+All analyses and key data are available for reproduction on OSF (\url{https://osf.io/57yha/}).
+Computer code used to collect the Zooniverse judgments is available from GitHub (Semenzin \& Cristia, 2020).
+This manuscript is reproducible thanks to the use of RMd (Baumer \& Udwin, 2015) and Papaja (Aust \& Barth, 2017) on R (Team \& others, 2013).
 
@@
 
-Recordings were obtained with the Language ENvironment Analysis (LENA\textsuperscript{TM}) Digital Language Processor, LENA\textsuperscript{TM} is a lightweight recorder (\textless{} 60 g, 5.5 cm x 8.5 cm x 1.5 cm) designed to be worn inside a breast-pocket of purpose-made clothing.
+Recordings were obtained with the Language ENvironment Analysis (LENA\textsuperscript{TM}) Digital Language Processor, a lightweight recorder (\textless{} 60 g, 5.5 cm x 8.5 cm x 1.5 cm) designed to be worn inside a breast-pocket of purpose-made clothing.
 In our data, every child contributed at least one daylong recording, with one child contributing two recordings (duration 11.58-16.00 hours, mean 15.05 hours).
@@
 The first step we took before beginning the annotation process was the identification of children's vocalizations within day-long recordings.
-We did this automatically, using the LENA\textsuperscript{TM} proprietary software.
+We did this automatically, using the LENA\textsuperscript{TM} proprietary software (Xu et al., 2008).
 Specifically, the LENA\textsuperscript{TM} software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child's environment (e.g., Female-Adult, TV).
-Key child labels have a precision of about XX\% and a recall of about XX\% (ref).
+Key child labels have a precision of about 60\% and a recall of about 50\% (Cristia, Lavechin, et al., 2020).
 These stretches of audio assigned to the key child can contain cries, vegetative sounds, and linguistic vocalizations, as well as interstitial pauses.
-We will refer to them as segments, to highlight the fact that they are stretches of audio segmented out as belonging to the key child, but they may not be individual vocalizations.
+We will refer to them as \emph{segments}, to highlight the fact that they are stretches of audio segmented out as belonging to the key child, but they may not be individual vocalizations.
 
@@
 A substantial concern that emerges when considering citizen science annotation for spontaneous speech data is the risk of a privacy breach: even short clips can contain personal information that can expose the identity of the speaker, such as addresses or names, or contain sensitive and private information.
-Fortunately, a study by Seidl et al.
-(2019) determined that when segments are divided into shorter chunks (400-600ms), human annotators can code our categories of interest (canonical, laughing, crying, etc.) with a classification quality comparable to the one carried out on full segments.
+Fortunately, Seidl et al.
+(2019) determined that when segments are divided into shorter chunks (400-600ms), human annotators with little training can code our categories of interest (canonical, laughing, crying, etc.) with a classification quality comparable to the one carried out on full segments.
 We therefore used very short chunks (500ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.
@@
 To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms clips and their corresponding segment.
-Clips were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling (\url{https://github.com/zooniverse/panoptes-cli}).
+Clips were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling (Bowyer, Lintott, Hines, Allan, \& Paget, 2015).
 
@@
 
-Users were asked to assign each 500ms clip to one out of five possible categories: (1) canonical, (2) non-canonical, (3) crying, (4) laughing, and (5) junk (overlapping speech, non-infant speech, silence, external sounds).
+Users were asked to assign each 500ms clip to one out of five possible categories: (1) canonical, (2) non-canonical, (3) crying, (4) laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
 Notice that the latter is narrower than the \enquote{Don't mark} category of the lab annotation routine, which was used any time there was noise or overlap that could affect acoustic analyses.
@@
 Majority agreement is thus achieved when 3 out of 5 judgements agree (rather than 2 out of 3).
-In exploratory analyses, we assess the impact of collecting a smaller number of judgments per clip.
 
@@
 
-We did this by considering the majority label extracted from Zooniverse judgements for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were junk, then the segment was labelled as junk; then, following a hierarchy (Canonical\textgreater{}Non-canonical\textgreater{}Laughing\textgreater{}Crying), segments with at least one instance of each judgement were labelled as such: e.g.~if one canonical judgement was present, then the segment was labelled as canonical, otherwise if a non-canonical judgement was present, the segment was labelled as non-canonical, and so on.
+We did this by considering the majority label extracted from Zooniverse judgements for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were Junk, then the segment was labelled as Junk; then, following a hierarchy (Canonical\textgreater{}Non-canonical\textgreater{}Crying\textgreater{}Laughing), segments with at least one instance of each judgement were labelled as such:
+
+\begin{itemize}
+\tightlist
+\item
+  if one canonical judgement was present, then the segment was labelled as Canonical,
+\item
+  else, if a Non-Canonical judgement was present, the segment was labelled as Non-Canonical,
+\item
+  else, if a Crying judgment was present, the segment was labelled as Crying,
+\item
+  else, the segment was labelled as Laughing.
+\end{itemize}
+
 This is essentially the same thought process expert annotators followed when labeling segment-level audio, as each segment can represent multiple classification subtypes).
@@
 According to lab annotators, 15\% of segments were canonical, 56\% non-canonical, 2\% laughing, and 5\% crying, with the remaining 22\% being categorized as \enquote{Don't mark}.
-Zooniverse data revealed a similar distribution: 15\% canonical, 60\% non-canonical, 4\% laughing, 9\% crying, 13\% junk.
-Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (\enquote{Canonical}+\enquote{Non-Canonical})/\enquote{All vocalizations} (i.e., we remove, junk), and ii) Canonical Proportion = \enquote{Canonical}/(\enquote{Canonical}+\enquote{Non-Canonical}) (i.e., we remove junk + non-linguistic).
+Zooniverse data revealed a similar distribution: 15\% canonical, 60\% non-canonical, 4\% laughing, 9\% crying, 13\% Junk.
+Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (\enquote{Canonical}+\enquote{Non-Canonical})/\enquote{All vocalizations} (i.e., we remove, Junk), and ii) Canonical Proportion = \enquote{Canonical}/(\enquote{Canonical}+\enquote{Non-Canonical}) (i.e., we remove Junk + non-linguistic).
 See Figure 2 for results.
@@
 
-We additionally explored under what conditions Zooniverse judgments more closely aligned with laboratory judgments.
+In this section, we report on explorations of under what conditions Zooniverse judgments more closely aligned with laboratory judgments.
 In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as \enquote{Junk} (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora (Cychosz et al., 2019).
@@
 
-\includegraphics{paper_files/figure-latex/cor-junk-1.pdf}
+\includegraphics{paper_files/figure-latex/cor-Junk-1.pdf}
 
@@
 Results were similar across these two post-hoc subgroups.
-For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower junk group was r(8) = 0.87, CI {[}0.53,0.97{]}, p=0.00{]}; and for the higher junk group it was r(8) = 0.87, CI {[}0.54,0.97{]}, p=0.00.
-For Canonical Proportion, the correlation across lab and Zooniverse data for the lower junk group was r(8) = 0.96, CI {[}0.82,0.99{]}, p=0{]}; and for the higher junk group it was r(8) = 0.93, CI {[}0.73,0.98{]}, p=0.
+For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = 0.87, CI {[}0.53,0.97{]}, p=0.00{]}; and for the higher Junk group it was r(8) = 0.87, CI {[}0.54,0.97{]}, p=0.00.
+For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = 0.96, CI {[}0.82,0.99{]}, p=0{]}; and for the higher Junk group it was r(8) = 0.93, CI {[}0.73,0.98{]}, p=0.
 Thus, it does not seem that a higher proportion of \enquote{Junk} judgments is an index of low quality data.
@@
 We observed a non-significant trend {[}t(17.92)=527.40 and 631.90{]} for lower number of segments in the Angelman Syndrome group (mean 527.40, range 326-924) than among low-risk control infants (mean 631.90, range 351-1074), likely because of the lower volubility of the former children.
-Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications {[}CITE MEG{]}.
-We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children {[}following CITE MEG{]}.
+Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications (Cychosz et al., 2019).
+We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children (following Cychosz et al., 2019).
 We randomly sampled 100 segments from each child's data, and recalculated the linguistic and Canonical Proportions based only on these associated judgments.
@@
 Higher precision and low recall indicates that when Zooniverse annotations return a \enquote{Junk} judgement, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as \enquote{Don't mark} in the laboratory.
-Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the \enquote{junk} category less) than laboratory annotators, who have access to the whole vocalization.
+Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the \enquote{Junk} category less) than laboratory annotators, who have access to the whole vocalization.
 It is also plausible, on the other hand, that lab annotators are more \enquote{picky} with the data, and tend to reject clips (tag them as \enquote{Don't mark}).
@@
 Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion.
-We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant for other disorders that are related to emotional disorders (REF REF).
-Additionally, while Canonical Proportion has been studied in much previous work, we believe further research is necessary using larger scales (e.g., sampling from daylong recordings to have a more ecological picture of the child, REF; increasing the sheer number of children studied, to have sufficient power to detect effects) and with broader coverage (notably including children growing up in a broad range of languages and cultures, REF), to make sure that assumptions regarding cross-linguistic and maturational universality are justified (REF).
+We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant for other disorders that are related to emotional disorders (a review in Halpern \& Coelho, 2016).
+Additionally, while Canonical Proportion has been studied in much previous work, we believe further research is necessary using larger scales (saliently increasing the sheer number of children studied, to have sufficient power to detect group effects) and with broader coverage (notably including children growing up in a broad range of languages and cultures, as in Cychosz et al., 2019), to make sure that assumptions regarding cross-linguistic and maturational universality are amply justified and not merely a result of having low power to detect differences, or simply sampling from similar populations.
 
 Assuming these child-level metrics continue to be supported by such additional data, it would be interesting for future investigations to expand the analysis of Linguistic Proportion and Canonical Proportion to recover differences and similarities in language development between low-risk children and children with disorders other than Angelman Syndrome.
-An extensive body of work has found that children with autism spectrum disorders experience language delays (Fasolo, Majorano, \& D'Odorico, 2008; Lang et al., 2019; Patten et al., 2014; Stoel-Gammon, 1989).
-One caveat is that these studies used CBO, which requires recordings to be gathered periodically in order to be determined.
+That is, it is at present unclear whether relatively lower Canonical Proportions will also be found in other populations, or whether this is more specific to Angelman Syndrome.
+For instance, some previous work has found differences in babbling even when comparing late talkers with toddlers with no specific diagnosis (Fasolo, Majorano, \& D'Odorico, 2008).
+We thus encourage further research inspecting Canonical Proportion in other populations, as well as methodological work that attempts to generate metrics that are similarly easy to gather at scale.
 
-It is also important to note that work on non-typically developing children poses in itself some challenges.
-Our data contained a lower number of vocalizations for children diagnosed with Angelman Syndrome, as they tend to produce fewer vocalizations than low-risk controls REF.
-Notice this was the case even in our data, where we selected much younger low-risk control childrens as the comparison point.
+It is also important to note that work on children with non-normative development poses in itself some challenges.
+Our data contained a lower number of vocalizations for children diagnosed with Angelman Syndrome, even though we selected much younger low-risk control childrens as the comparison group.
+However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
+Now that this method has been validated, researchers can extract all segments that LENA\textsuperscript{TM} identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
 
-The lack of age effects in AS seems like the bigger \enquote{challenge} - if the goal is to use this information to inform how children are developing, are these the right metrics for that population?
-Longitudinal work within AS may be needed to track the \enquote{natural history} of vocalizations to better contextualize how these metrics might be relevant or useful.
-For example, there is interest in using naturalistic recordings as outcome measures for clinical/drug trials.
-Are we seeing enough variability in AS to assume that these metrics could \enquote{budge} in response to a treatment?
-It seems we are seeing variability across kids, which is promising.
-We are also seeing agreement across coders in AS files, which is also important.
-Next questions would be whether these metrics change within individuals, and whether rates of change may be a candidate for tracking treatment.
-Graphs that show the two key variables across groups and age may help inform these questions.
+Another open question we hope future work looks into is the assessment of within-participant changes.
+Our dataset only contained one daylong recording per child (with one exception), so other datasets in which a greater number of recordings are collected longitudinally can bring novel insights.
+Is the age-related decline we observed among children in the Angelman Syndrome group truly a developmental decline that happens by necessity in this population, or might it reflect partially other aspects of children's experience and behavior?
+Additionally, longitudinal work within the same infant may be needed to track the \enquote{natural history} of vocalizations to better contextualize how these metrics should be interpreted.
+It is also an open question how large must effects be, at the individual level, in order to be detected with this method.
+For example, there is interest in using naturalistic recordings as outcome measures for clinical (behavioral or pharmacological) trials.
+The fact that agreement across annotators is high even in the data coming from children diagnosed with Angelman Syndrome is encouraging, but it would be important to establish whether absolute levels of Canonical Proportion, and/or rates of change over longer time spans, may be good candidates for tracking the effects of any treatment.
 
 Finally, broader generalizations of our technique could be of interest to readers of this paper.
-Any researcher is able to undergo the procedure to create Zooniverse projects.
-For instance, a separate pipeline could be set up to annotate potentially disordered speech by older adults at risk of, or diagnosed with, neurodegenerative diseases.
+Any researcher is able to undergo the procedure to create Zooniverse projects (\url{https://www.zooniverse.org/about/faq}).
+For instance, a project could be set up to annotate potentially disordered speech by older adults at risk of, or diagnosed with, neurodegenerative diseases.
 Although it would be important to similarly validate such annotations, we believe there is strong promise for such tasks, because many neurodegenerative disorders affect local aspects of the speech, which are detected at the syllable level.
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-aust2017papaja}{}%
+Aust, F., \& Barth, M. (2017). Papaja: Create apa manuscripts with r markdown.
+
+\leavevmode\hypertarget{ref-baumer2015r}{}%
+Baumer, B., \& Udwin, D. (2015). R markdown. \emph{Wiley Interdisciplinary Reviews: Computational Statistics}, \emph{7}(3), 167--177.
+
+\leavevmode\hypertarget{ref-berinsky2012evaluating}{}%
+Berinsky, A. J., Huber, G. A., \& Lenz, G. S. (2012). Evaluating online labor markets for experimental research: Amazon. Com's mechanical turk. \emph{Political Analysis}, \emph{20}(3), 351--368.
+
+\leavevmode\hypertarget{ref-borne2011zooniverse}{}%
+Borne, K., \& Zooniverse Team. (2011). The zooniverse: A framework for knowledge discovery from citizen science data. \emph{AGUFM}, \emph{2011}, ED23C--0650.
+
+\leavevmode\hypertarget{ref-bowyer2015panoptes}{}%
+Bowyer, A., Lintott, C., Hines, G., Allan, C., \& Paget, E.
+(2015).
+Panoptes, a project building tool for citizen science.
+In \emph{Proceedings of the aaai conference on human computation and crowdsourcing (hcomp'15).
+AAAI, san diego, ca, usa}.
+
+\leavevmode\hypertarget{ref-casillas2017new}{}%
+Casillas, M., Bergelson, E., Warlaumont, A.
+S., Cristia, A., Soderstrom, M., VanDam, M., \& Sloetjes, H.
+(2017).
+A new workflow for semi-automatized annotations: Tests with long-form naturalistic recordings of childrens language environments.
+In \emph{Interspeech 2017} (pp.
+2098--2102).
+
+\leavevmode\hypertarget{ref-casillas2019step}{}%
+Casillas, M., \& Cristia, A. (2019). A step-by-step guide to collecting and analyzing long-format speech environment (lfse) recordings. \emph{Collabra}, \emph{5}(1).
+
+\leavevmode\hypertarget{ref-cristia2020accuracy}{}%
+Cristia, A., Bulgarelli, F., \& Bergelson, E.
+(2020).
+Accuracy of the language environment analysis system segmentation and metrics: A systematic review.
+\emph{Journal of Speech, Language, and Hearing Research}, \emph{63}(4), 1093--1105.
+
+\leavevmode\hypertarget{ref-cristia2020thorough}{}%
+Cristia, A., Lavechin, M., Scaff, C., Soderstrom, M., Rowland, C., Räsänen, O., \ldots{} Bergelson, E.
+(2020).
+A thorough evaluation of the language environment analysis (lena) system.
+\emph{Behavior Research Methods}, 1--20.
+
+\leavevmode\hypertarget{ref-crump2013evaluating}{}%
+Crump, M. J., McDonnell, J. V., \& Gureckis, T. M. (2013). Evaluating amazon's mechanical turk as a tool for experimental behavioral research. \emph{PloS One}, \emph{8}(3), e57410.
+
 \leavevmode\hypertarget{ref-cychosz2019canonical}{}%
 Cychosz, M., Cristia, A., Bergelson, E., Casillas, M., Baudet, G., Warlaumont, A. S., \ldots{} Seidl, A. (2019). Canonical babble development in a large-scale crosslinguistic corpus.
+
+\leavevmode\hypertarget{ref-fasolo2008babbling}{}%
+Fasolo, M., Majorano, M., \& D'Odorico, L. (2008). Babbling and first words in children with slow expressive development. \emph{Clinical Linguistics \& Phonetics}, \emph{22}(2), 83--94.
+
+\leavevmode\hypertarget{ref-ganek2018language}{}%
+Ganek, H., \& Eriks-Brophy, A.
+(2018).
+Language environment analysis (lena) system investigation of day long recordings in children: A literature review.
+\emph{Journal of Communication Disorders}, \emph{72}, 77--85.
+
+\leavevmode\hypertarget{ref-halpern2016excessive}{}%
+Halpern, R., \& Coelho, R. (2016). Excessive crying in infants. \emph{Jornal de Pediatria}, \emph{92}(3), S40--S45.
 
@@
 
+\leavevmode\hypertarget{ref-heckman2006skill}{}%
+Heckman, J. J. (2006). Skill formation and the economics of investing in disadvantaged children. \emph{Science}, \emph{312}(5782), 1900--1902.
+
+\leavevmode\hypertarget{ref-mcdaniel2020predicting}{}%
+McDaniel, J., Yoder, P., Estes, A., \& Rogers, S.
+J.
+(2020).
+Predicting expressive language from early vocalizations in young children with autism spectrum disorder: Which vocal measure is best?
+\emph{Journal of Speech, Language, and Hearing Research}, \emph{63}(5), 1509--1520.
+
+\leavevmode\hypertarget{ref-oller2000emergence}{}%
+Oller, D. (2000). The emergence of the capacity for speech. Erlbaum.{[}aJLL, DKO{]}(2004) Underpinnings for a theory of communicative~\ldots{}.
+
+\leavevmode\hypertarget{ref-oller1998late}{}%
+Oller, D.
+K., Eilers, R.
+E., Neal, A.
+R., \& Cobo-Lewis, A.
+B.
+(1998).
+Late onset canonical babbling: A possible early marker of abnormal development.
+\emph{American Journal on Mental Retardation}, \emph{103}(3), 249--263.
+
+\leavevmode\hypertarget{ref-schuller2019interspeech}{}%
+Schuller, B.
+W., Batliner, A., Bergler, C., Pokorny, F.
+B., Krajewski, J., Cychosz, M., \ldots{} others.
+(2019).
+The interspeech 2019 computational paralinguistics challenge: Styrian dialects, continuous sleepiness, baby sounds \& orca activity.
+In \emph{Interspeech} (pp.
+2378--2382).
+
+\leavevmode\hypertarget{ref-Seidl19}{}%
+Seidl, A., Warlaumont, A., \& Cristia, A.
+(2019).
+Towards detection of canonical babbling by citizen scientists: Performance as a function of clip length.
+In (pp.
+3579--3583).
+\url{https://doi.org/10.21437/Interspeech.2019-1773}
+
+\leavevmode\hypertarget{ref-semenzin2020zooniverse}{}%
+Semenzin, C., \& Cristia, A. (2020). Zooniverse pipeline for long format recordings.
+
+\leavevmode\hypertarget{ref-sescleifer2018systematic}{}%
+Sescleifer, A. M., Francoisse, C. A., \& Lin, A. Y. (2018). Systematic review: Online crowdsourcing to assess perceptual speech outcomes. \emph{Journal of Surgical Research}, \emph{232}, 351--364.
+
+\leavevmode\hypertarget{ref-team2013r}{}%
+Team, R. C., \& others. (2013). R: A language and environment for statistical computing.
+
+\leavevmode\hypertarget{ref-wang2020meta}{}%
+Wang, Y., Williams, R., Dilley, L., \& Houston, D.
+M.
+(2020).
+A meta-analysis of the predictability of lena™ automated measures for child language development.
+\emph{Developmental Review}, \emph{57}, 100921.
+
+\leavevmode\hypertarget{ref-xu2008signal}{}%
+Xu, D., Yapanel, U., Gray, S., Gilkerson, J., Richards, J., \& Hansen, J.
+(2008).
+Signal processing for young child speech language development.
+In \emph{First workshop on child, computer and interaction}.
+
+\leavevmode\hypertarget{ref-yeh2019using}{}%
+Yeh, S.-L., Chao, G.-Y., Su, B.-H., Huang, Y.-L., Lin, M.-H., Tsai, Y.-C., \ldots{} others.
+(2019).
+Using attention networks and adversarial augmentation for styrian dialect continuous sleepiness and baby sound recognition.
+In \emph{INTERSPEECH} (pp.
+2398--2402).
+
 \end{document}

### 2020-11-03 202010_jslhr/paper.Rmd
integrated amanda and bridgette comments

@@
 As mentioned earlier, long-form recordings have several advantages, including capturing the child's vocal patterns in their natural environment and being able to accumulate a great amount of data easily, which is particularly useful for diagnoses and pathologies characterized by low levels of vocal production.
-However, this also means that there is a great amount of data, and that the audio quality is compromised, as children go about their day transitioning from small to large rooms, with fewer or more people around them, thus providing analysts with an unstable and challenging long audio.
+However, this also means that the audiorecording is harder to process than an audiorecording gathered in more manicured and stable conditions: In a typical daylong recording, the child will sometimes be in a quiet room with just her primary caregiver, but later in a noisy supermarket or having dinner with family and friends.
 In fact, human annotation of such audio is estimated to take about 30 times the audio length, with these estimates being greater the more precise one wants or needs to be [@casillas2017new].
@@
 Here, we focus on two alternative composite metrics.
-An additional metric that can be derived once children's vocalizations are split into crying, laughing, canonical, and non-canonical is the Linguistic Proportion: the proportion of vocalizations that are linguistic (canonical and non-canonical) out of all vocalizations.
+One additional metric that can be derived once children's vocalizations are split into crying, laughing, canonical, and non-canonical is the Linguistic Proportion: the proportion of vocalizations that are linguistic (canonical and non-canonical) out of all vocalizations.
 To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age.
@@
 First, we relied on the largest and best established citizen science platform: Zooniverse hosts more than 1.6 million users from diverse walks of life, and it offers a completely automatized API system to more easily scale tasks in a transparent and cumulative science fashion.
-In contrast, the authors of that paper told us that, in order to complete their data annotation, they recruited the help of students in their classes and labs, and sometimes the researchers themselves took on a considerable annotation load.
-This means that the process draws on expert resources, whose time is expensive.
+In contrast, @cychosz2019canonical actively recruited people to provide annotations, stating that "[a]nnotators included language and speech researchers, undergraduate students, research assistants, and other interested parties, totaling 136 unique annotators".
+A process that draws from research assistants relies on expert resources, whose time is expensive.
 Moreover, the people these researchers recruit tend to be more accustomed to child vocalization data, which may improve their annotation performance, and thus provide an overestimate of the quality of the annotations that can be realistically done within a truly citizen science framework.
@@
 It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels.
-We therefore included children in a wide age range, with some of the children being diagnosed with Angelman Syndrome, as well as low-risk controls.
+We therefore included children in a wide age range, with some of the children being diagnosed with Angelman syndrome, as well as low-risk controls.
 
@@
 When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (canonical+non-canonical)/(all vocalizations); and the Canonical Proportion (canonical)/(all linguistic vocalizations).
-Previous work suggests that children with Angelman Syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the canonical ratio was expected to increase with age [@hamrick2019capturing].
+Previous work suggests that children with Angelman syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the canonical ratio was expected to increase with age [@hamrick2019capturing].
 Although we could not rely on previous work to make predictions regarding the Linguistic Proportion, we reasoned that we should observe an increase in the Linguistic Proportion for both populations.
@@
 
-Annotations were carried out by 20 undergraduate students working in either Speech, Language, and Hearing Sciences or clinical psychology laboratories.
+Annotations were carried out by 20 undergraduate students working in either Speech, Language, and Hearing Sciences or Psychological Science laboratories.
 Before annotation, coders completed an ethical conduct of research course and received HIPAA compliance training.
@@
 For canonical, non-canonical, and word segments, coders indicated how many syllables of each type of vocal maturity they hear (e.g., "ah ma ba" would be 1 non-canonical and 2 canonical syllables).
-Coders are also given a "Don’t Mark" option, which they are instructed to use if the segment does not sound like a segment made by the target child, or if there is any overlapping speech or other noise in the background which could affect acoustic analyses.
 For the purposes of the present study, we considered a "word" judgment to be equivalent to a "canonical" judgment.
 This may be false sometimes (e.g., "oh" may be coded as a real word by laboratory experts, although it is a non-canonical syllable), but we could not have a "word" judgment in our Zooniverse annotations because that requires context (which compromises privacy -- see next section).
+
+Coders are also given a "Don’t Mark" option, which they are instructed to use if the segment does not sound like a segment made by the target child, or if there is any overlapping speech or other noise in the background which could affect acoustic analyses.
+Notice, therefore, that this category is broader than the "Junk" category used by Zooniverse annotators.
+
 This process resulted in high levels of agreement: 97.22% of the lab-annotated segments had either 2 or 3 coders agree.
@@
 
-```{r fig-corage, echo=F,fig.cap="Correlations between child-level descriptors and age as a function of metric (linguistic ratio in the top row, canonical ratio in the bottom row), annotation method, and child group (red=low-risk, and black Angelman Syndrome)"}
+```{r fig-corage, echo=F,fig.cap="Correlations between child-level descriptors and age as a function of metric (linguistic ratio in the top row, canonical ratio in the bottom row), annotation method, and child group (red=low-risk, and black Angelman syndrome)."}
 
@@
 Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups.
-There was a near-zero relationship among the older children diagnosed with Angelman Syndrome r(`r myn["l_lp","AS"]`) = `r myr["l_lp","AS"]`, CI [`r mylb["l_lp","AS"]`,`r myhb["l_lp","AS"]`], p=`r myp["l_lp","AS"]`]; and a significant association among younger low-risk control children r(`r myn["l_lp","LR"]`) = `r myr["l_lp","LR"]`, CI [`r mylb["l_lp","LR"]`,`r myhb["l_lp","LR"]`], p=`r myp["l_lp","LR"]`].
-The Canonical Proportion exhibited non-significant developmental decreases among older children diagnosed with Angelman Syndrome r(`r myn["l_cp","AS"]`) = `r myr["l_cp","AS"]`, CI [`r mylb["l_cp","AS"]`,`r myhb["l_cp","AS"]`], p=`r myp["l_cp","AS"]`]; and marginal developmental increases among low-risk control r(`r myn["l_cp","LR"]`) = `r myr["l_cp","LR"]`, CI [`r mylb["l_cp","LR"]`,`r myhb["l_cp","LR"]`], p=`r myp["l_cp","LR"]`].
-
-Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman Syndrome r(`r myn["z_lp","AS"]`) = `r myr["z_lp","AS"]`, CI [`r mylb["z_lp","AS"]`,`r myhb["z_lp","AS"]`], p=`r myp["z_lp","AS"]`]; whereas younger low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r myr["z_lp","LR"]`, CI [`r mylb["z_lp","LR"]`,`r myhb["z_lp","LR"]`], p=`r myp["z_lp","LR"]`].
-Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman Syndrome r(`r myn["z_cp","AS"]`) = `r myr["z_cp","AS"]`, CI [`r mylb["z_cp","AS"]`,`r myhb["z_cp","AS"]`], p=`r myp["z_cp","AS"]`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r myr["z_cp","LR"]`, CI [`r mylb["z_cp","LR"]`,`r myhb["z_cp","LR"]`], p=`r myp["z_cp","LR"]`].
+There was a near-zero relationship among the older children diagnosed with Angelman syndrome r(`r myn["l_lp","AS"]`) = `r myr["l_lp","AS"]`, CI [`r mylb["l_lp","AS"]`,`r myhb["l_lp","AS"]`], p=`r myp["l_lp","AS"]`]; and a significant association among younger low-risk control children r(`r myn["l_lp","LR"]`) = `r myr["l_lp","LR"]`, CI [`r mylb["l_lp","LR"]`,`r myhb["l_lp","LR"]`], p=`r myp["l_lp","LR"]`].
+The Canonical Proportion exhibited non-significant developmental decreases among older children diagnosed with Angelman syndrome r(`r myn["l_cp","AS"]`) = `r myr["l_cp","AS"]`, CI [`r mylb["l_cp","AS"]`,`r myhb["l_cp","AS"]`], p=`r myp["l_cp","AS"]`]; and marginal developmental increases among low-risk control r(`r myn["l_cp","LR"]`) = `r myr["l_cp","LR"]`, CI [`r mylb["l_cp","LR"]`,`r myhb["l_cp","LR"]`], p=`r myp["l_cp","LR"]`].
+
+Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(`r myn["z_lp","AS"]`) = `r myr["z_lp","AS"]`, CI [`r mylb["z_lp","AS"]`,`r myhb["z_lp","AS"]`], p=`r myp["z_lp","AS"]`]; whereas younger low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r myr["z_lp","LR"]`, CI [`r mylb["z_lp","LR"]`,`r myhb["z_lp","LR"]`], p=`r myp["z_lp","LR"]`].
+Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman syndrome r(`r myn["z_cp","AS"]`) = `r myr["z_cp","AS"]`, CI [`r mylb["z_cp","AS"]`,`r myhb["z_cp","AS"]`], p=`r myp["z_cp","AS"]`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r myr["z_cp","LR"]`, CI [`r mylb["z_cp","LR"]`,`r myhb["z_cp","LR"]`], p=`r myp["z_cp","LR"]`].
 
@@
 
-```{r fig-prec-rec, echo=F}
+```{r fig-prec-rec, echo=F,fig.cap="Number of segments as a function of their majority label from laboratory annotations (rows) versus Zooniverse annotations (columns).
+Percentages (and shading) indicate Precision (left) and Recall (right).
+For precision, the percentage indicates what percentage of the segments that have a given majority label according to Zooniverse data are attributed a given label in Lab data (i.e., columns add up to 100%).
+For recall, the percentage indicates what percentage of the segments that have a given majority label according to Lab data are attributed a given label in Zooniverse data (i.e., rows add up to 100%)."}
 
@@
 Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman Syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
+When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
 
 Similarly, a strong positive correlation is found in the Canonical Proportion r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`].
-When we split by participant group, correlations remain high although we do note they are somewhat smaller for the older children with Angelman Syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
+When we split by participant group, correlations remain high although we do note they are somewhat smaller for the older children with Angelman syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
 
@@
 We investigated this hypothesis by calculating the proportion of their data labeled as "Junk" for each individual child.
-There was no significant difference in the proportions of their data labeled as "Junk" for older children with Angelman Syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the younger low risk children (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, p = `r round(JunkXdiag$p.value,3)`.
+There was no significant difference in the proportions of their data labeled as "Junk" for older children with Angelman syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the younger low risk children (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, p = `r round(JunkXdiag$p.value,3)`.
 We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk".
@@
 
-Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman Syndrome group: r(`r z_lp_as$parameter`) = `r z_lp_as$estimate`, CI [`r z_lp_as$conf.int[1]`,`r z_lp_as$conf.int[2]`], p=`r z_lp_as$p.value`], but increased with age  in the low-risk control group: r(`r z_lp_lr$parameter`) = `r z_lp_lr$estimate`, CI [`r z_lp_lr$conf.int[1]`,`r z_lp_lr$conf.int[2]`], p=`r z_lp_lr$p.value`.
-Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman Syndrome r(`r z_cp_as$parameter`) = `r z_cp_as$estimate`, CI [`r z_cp_as$conf.int[1]`,`r z_cp_as$conf.int[2]`], p=`r z_cp_as$p.value`]; and developmental increases among low-risk control r(`r z_cp_lr$parameter`) = `r z_cp_lr$estimate`, CI [`r z_cp_lr$conf.int[1]`,`r z_cp_lr$conf.int[2]`], p=`r z_cp_lr$p.value`.
+Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman syndrome group: r(`r z_lp_as$parameter`) = `r z_lp_as$estimate`, CI [`r z_lp_as$conf.int[1]`,`r z_lp_as$conf.int[2]`], p=`r z_lp_as$p.value`], but increased with age  in the low-risk control group: r(`r z_lp_lr$parameter`) = `r z_lp_lr$estimate`, CI [`r z_lp_lr$conf.int[1]`,`r z_lp_lr$conf.int[2]`], p=`r z_lp_lr$p.value`.
+Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman syndrome r(`r z_cp_as$parameter`) = `r z_cp_as$estimate`, CI [`r z_cp_as$conf.int[1]`,`r z_cp_as$conf.int[2]`], p=`r z_cp_as$p.value`]; and developmental increases among low-risk control r(`r z_cp_lr$parameter`) = `r z_cp_lr$estimate`, CI [`r z_cp_lr$conf.int[1]`,`r z_cp_lr$conf.int[2]`], p=`r z_cp_lr$p.value`.
 
@@
 As for correlations between Zooniverse and laboratory-derived metrics, we observe very similar levels of correlation: Linguistic Proportion overall r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman Syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
-As for Canonical Proportion overall r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`; Angelman Syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`];  low-risk control infants r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
+When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
+As for Canonical Proportion overall r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`; Angelman syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`];  low-risk control infants r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
 
@@
 Next, we looked at whether having more segments from each child may lead to more reliable metrics.
-We observed a non-significant trend [t(`r nseg_t$parameter`)=`r nseg_t$estimate`] for lower number of segments in the Angelman Syndrome group (mean `r nseg_t$estimate[1]`, range `r min(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`) than among low-risk control infants (mean `r nseg_t$estimate[2]`, range `r min(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`), likely because of the lower volubility of the former children.
+We observed a non-significant trend [t(`r nseg_t$parameter`)=`r nseg_t$estimate`] for lower number of segments in the Angelman syndrome group (mean `r nseg_t$estimate[1]`, range `r min(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`) than among low-risk control infants (mean `r nseg_t$estimate[2]`, range `r min(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`), likely because of the lower volubility of the former children.
 Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications [@cychosz2019canonical].
@@
 We also observed quite a bit of variance across the 50 runs.
-For Linguistic Proportion, the range of correlations found for all infants was [`r min(resampling$lp_r_all)`,`r max(resampling$lp_r_all)`]; for the Angelman Syndrome group [`r min(resampling$lp_r_as)`,`r max(resampling$lp_r_as)`]; for the low-risk group [`r min(resampling$lp_r_lr)`,`r max(resampling$lp_r_lr)`].
-For Canonical Proportion, the range of correlations found for all infants was [`r min(resampling$cp_r_all)`,`r max(resampling$cp_r_all)`]; for the Angelman Syndrome group [`r min(resampling$cp_r_as)`,`r max(resampling$cp_r_as)`]; for the low-risk group [`r min(resampling$cp_r_lr)`,`r max(resampling$cp_r_lr)`].
+For Linguistic Proportion, the range of correlations found for all infants was [`r min(resampling$lp_r_all)`,`r max(resampling$lp_r_all)`]; for the Angelman syndrome group [`r min(resampling$lp_r_as)`,`r max(resampling$lp_r_as)`]; for the low-risk group [`r min(resampling$lp_r_lr)`,`r max(resampling$lp_r_lr)`].
+For Canonical Proportion, the range of correlations found for all infants was [`r min(resampling$cp_r_all)`,`r max(resampling$cp_r_all)`]; for the Angelman syndrome group [`r min(resampling$cp_r_as)`,`r max(resampling$cp_r_as)`]; for the low-risk group [`r min(resampling$cp_r_lr)`,`r max(resampling$cp_r_lr)`].
 
@@
 colnames(mycortab)<-c("Linguistic Proportion","Canonical Proportion")
-kable(mycortab,cap="Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman Syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways: All seg indicates that Zooniverse annotations at the clip level were first combined at the segment level; Chunks that they were analyzed directly at the clip level.
+kable(mycortab,cap="Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways: All seg indicates that Zooniverse annotations at the clip level were first combined at the segment level; Chunks that they were analyzed directly at the clip level.
 Both of these are based on all the data.
@@
 In descriptive terms, we found that category frequency was similar across laboratory and citizen science annotations.
-Correlations between age and our two child-level metrics revealed different developmental patterns between our sample of older children with a diagnosis of Angelman Syndrome and younger low-risk controls.
+Correlations between age and our two child-level metrics revealed different developmental patterns between our sample of older children with a diagnosis of Angelman syndrome and younger low-risk controls.
 Given that the two groups of children differ on both of these features (age and diagnosis), we cannot empirically tease apart these two differences.
@@
 Overall, the pattern for Linguistic Proportion is compatible with a non-linear trajectory common to the two groups, whereby the Linguistic Proportion increases rapidly up to 20 months, and hovers around 90-100% thereafter.
-In this case, the lack of correlation with age in the Angelman Syndrome group suggests that there is little systematic variance in this metric after 20 months of age.
-In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and slower decreases among children in the Angelman Syndrome group.
+In this case, the lack of correlation with age in the Angelman syndrome group suggests that there is little systematic variance in this metric after 20 months of age.
+In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and slower decreases among children in the Angelman syndrome group.
 
@@
 
-As for laughing, the recall is quite good but precision is low with confusion involving Junk and Non-Canonical to similar extents.
+As for laughing, the recall is quite good but precision is low, with confusion involving Junk and Non-Canonical to similar extents.
 We are tentative in our interpretation of this result because the category was very sparsely represented, with barely 200 segments classified as Laughing in the whole data set.
@@
 Errors only seemed systematic for the Linguistic Proportion (and not for Canonical Proportion), and they were due to lower Linguistic Proportions from Zooniverse than laboratory data.
-This is uncertainly due to the confusion between Crying and Non-Canonical, with a substantial part of Crying segments being classified as Non-Canonical by Zooniverse annotators.
+This is due to the fact that a substantial number of Crying segments were classified as Non-Canonical by Zooniverse annotators.
 Nonetheless, since these errors were systematic across infants, they do not affect the study of group or individual variation.
@@
 Our exploratory analyses then dug into some methodological aspects of our data.
-We found that, although individuals vary widely in terms of what proportion of their data is classified as Junk by Zooniverse annotators, this is not a sign that their data as a whole cannot be trusted, with similar levels of correlations between laboratory- and Zooniverse-derived metrics across median groups based on proportion of their data classified as Junk.
+We found that, although individuals vary widely in terms of what proportion of their data is classified as Junk by Zooniverse annotators, this is not a sign that their data as a whole is compromised, with similar levels of correlations between laboratory- and Zooniverse-derived metrics across median groups based on proportion of their data classified as Junk.
 
@@
 
-In sum, we believe these data show that young children's vocalizations can be faithfully described by relying on the judgments of Zooniverse's citizen scientists, particularly when the goal is to describe vocalizations at a broader level than individual clips.
+In sum, we believe these data show that young children's vocalizations can be accurately described by relying on the judgments of Zooniverse's citizen scientists, particularly when the goal is to describe vocalizations at a broader level than individual clips.
 It is important to note that from a methodological point of view, the use of crowdsourcing is not merely a question of being an ecological technique, but it can also offer benefits in the area of scientific rigor and reproducibility.
@@
 
-We believe additional research would be welcome on the two measures we used to describe children’s linguistic productions: Linguistic Proportion and Canonical Proportion.
+Our results warrant additional research on two metrics we used to describe children’s linguistic productions: Linguistic Proportion and Canonical Proportion.
 Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion.
@@
 
-Assuming these child-level metrics continue to be supported by such additional data, it would be interesting for future investigations to expand the analysis of Linguistic Proportion and Canonical Proportion to recover differences and similarities in language development between low-risk children and children with disorders other than Angelman Syndrome.
-That is, it is at present unclear whether relatively lower Canonical Proportions will also be found in other populations, or whether this is more specific to Angelman Syndrome.
+Assuming these child-level metrics continue to be supported by such additional data, it would be interesting for future investigations to expand the analysis of Linguistic Proportion and Canonical Proportion to recover differences and similarities in language development between low-risk children and children with disorders other than Angelman syndrome.
+That is, it is at present unclear whether relatively lower Canonical Proportions will also be found in other populations, or whether this is more specific to Angelman syndrome.
 For instance, some previous work has found differences in babbling even when comparing late talkers with toddlers with no specific diagnosis  [@fasolo2008babbling].
@@
 It is also important to note that work on children with non-normative development poses in itself some challenges.
-Our data contained a lower number of vocalizations for children diagnosed with Angelman Syndrome, even though we selected much younger low-risk control childrens as the comparison group.
+Our data contained a lower number of vocalizations for children diagnosed with Angelman syndrome, even though we selected much younger low-risk control childrens as the comparison group.
 However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
@@
 Our dataset only contained one daylong recording per child (with one exception), so other datasets in which a greater number of recordings are collected longitudinally can bring novel insights.
-Is the age-related decline we observed among children in the Angelman Syndrome group truly a developmental decline that happens by necessity in this population, or might it reflect partially other aspects of children's experience and behavior?
+Is the age-related decline we observed among children in the Angelman syndrome group truly a developmental decline that happens by necessity in this population, or might it reflect partially other aspects of children's experience and behavior?
 Additionally, longitudinal work within the same infant may be needed to track the "natural history" of vocalizations to better contextualize how these metrics should be interpreted.
@@
 For example, there is interest in using naturalistic recordings as outcome measures for clinical (behavioral or pharmacological) trials.
-The fact that agreement across annotators is high even in the data coming from children diagnosed with Angelman Syndrome is encouraging, but it would be important to establish whether absolute levels of Canonical Proportion, and/or rates of change over longer time spans, may be good candidates for tracking the effects of any treatment.
+The fact that agreement across annotators is high even in the data coming from children diagnosed with Angelman syndrome is encouraging, but it would be important to establish whether absolute levels of Canonical Proportion, and/or rates of change over longer time spans, may be good candidates for tracking the effects of any treatment.
 

### 2020-11-03 202010_jslhr/paper.tex
integrated amanda and bridgette comments

@@
 As mentioned earlier, long-form recordings have several advantages, including capturing the child's vocal patterns in their natural environment and being able to accumulate a great amount of data easily, which is particularly useful for diagnoses and pathologies characterized by low levels of vocal production.
-However, this also means that there is a great amount of data, and that the audio quality is compromised, as children go about their day transitioning from small to large rooms, with fewer or more people around them, thus providing analysts with an unstable and challenging long audio.
+However, this also means that the audiorecording is harder to process than an audiorecording gathered in more manicured and stable conditions: In a typical daylong recording, the child will sometimes be in a quiet room with just her primary caregiver, but later in a noisy supermarket or having dinner with family and friends.
 In fact, human annotation of such audio is estimated to take about 30 times the audio length, with these estimates being greater the more precise one wants or needs to be (Casillas et al., 2017).
@@
 First, we relied on the largest and best established citizen science platform: Zooniverse hosts more than 1.6 million users from diverse walks of life, and it offers a completely automatized API system to more easily scale tasks in a transparent and cumulative science fashion.
-In contrast, the authors of that paper told us that, in order to complete their data annotation, they recruited the help of students in their classes and labs, and sometimes the researchers themselves took on a considerable annotation load.
-This means that the process draws on expert resources, whose time is expensive.
+In contrast, Cychosz et al.
+(2019) actively recruited people to provide annotations, stating that \enquote{{[}a{]}nnotators included language and speech researchers, undergraduate students, research assistants, and other interested parties, totaling 136 unique annotators}.
+A process that draws from research assistants relies on expert resources, whose time is expensive.
 Moreover, the people these researchers recruit tend to be more accustomed to child vocalization data, which may improve their annotation performance, and thus provide an overestimate of the quality of the annotations that can be realistically done within a truly citizen science framework.
@@
 When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (canonical+non-canonical)/(all vocalizations); and the Canonical Proportion (canonical)/(all linguistic vocalizations).
-Previous work suggests that children with Angelman Syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the canonical ratio was expected to increase with age (Hamrick, Seidl, \& Kelleher, n.d.).
+Previous work suggests that children with Angelman syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the canonical ratio was expected to increase with age (Hamrick, Seidl, \& Kelleher, n.d.).
 Although we could not rely on previous work to make predictions regarding the Linguistic Proportion, we reasoned that we should observe an increase in the Linguistic Proportion for both populations.
@@
 For canonical, non-canonical, and word segments, coders indicated how many syllables of each type of vocal maturity they hear (e.g., \enquote{ah ma ba} would be 1 non-canonical and 2 canonical syllables).
-Coders are also given a \enquote{Don't Mark} option, which they are instructed to use if the segment does not sound like a segment made by the target child, or if there is any overlapping speech or other noise in the background which could affect acoustic analyses.
 For the purposes of the present study, we considered a \enquote{word} judgment to be equivalent to a \enquote{canonical} judgment.
 This may be false sometimes (e.g., \enquote{oh} may be coded as a real word by laboratory experts, although it is a non-canonical syllable), but we could not have a \enquote{word} judgment in our Zooniverse annotations because that requires context (which compromises privacy -- see next section).
+
+Coders are also given a \enquote{Don't Mark} option, which they are instructed to use if the segment does not sound like a segment made by the target child, or if there is any overlapping speech or other noise in the background which could affect acoustic analyses.
+Notice, therefore, that this category is broader than the \enquote{Junk} category used by Zooniverse annotators.
+
 This process resulted in high levels of agreement: 97.22\% of the lab-annotated segments had either 2 or 3 coders agree.
@@
 \includegraphics{paper_files/figure-latex/fig-corage-1.pdf}
-\caption{\label{fig:fig-corage}Correlations between child-level descriptors and age as a function of metric (linguistic ratio in the top row, canonical ratio in the bottom row), annotation method, and child group (red=low-risk, and black Angelman Syndrome)}
+\caption{\label{fig:fig-corage}Correlations between child-level descriptors and age as a function of metric (linguistic ratio in the top row, canonical ratio in the bottom row), annotation method, and child group (red=low-risk, and black Angelman syndrome).}
 \end{figure}
@@
 Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups.
-There was a near-zero relationship among the older children diagnosed with Angelman Syndrome r(8) = 0.02, CI {[}-0.62,0.64{]}, p=0.97{]}; and a significant association among younger low-risk control children r(8) = 0.74, CI {[}0.21,0.94{]}, p=0.01{]}.
-The Canonical Proportion exhibited non-significant developmental decreases among older children diagnosed with Angelman Syndrome r(8) = -0.56, CI {[}-0.88,0.11{]}, p=0.09{]}; and marginal developmental increases among low-risk control r(8) = 0.61, CI {[}-0.04,0.90{]}, p=0.06{]}.
+There was a near-zero relationship among the older children diagnosed with Angelman syndrome r(8) = 0.02, CI {[}-0.62,0.64{]}, p=0.97{]}; and a significant association among younger low-risk control children r(8) = 0.74, CI {[}0.21,0.94{]}, p=0.01{]}.
+The Canonical Proportion exhibited non-significant developmental decreases among older children diagnosed with Angelman syndrome r(8) = -0.56, CI {[}-0.88,0.11{]}, p=0.09{]}; and marginal developmental increases among low-risk control r(8) = 0.61, CI {[}-0.04,0.90{]}, p=0.06{]}.
 
-Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman Syndrome r(8) = -0.10, CI {[}-0.68,0.57{]}, p=0.79{]}; whereas younger low-risk control children showed a significant increase with age r(8) = 0.68, CI {[}0.08,0.92{]}, p=0.03{]}.
-Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman Syndrome r(8) = -0.42, CI {[}-0.83,0.29{]}, p=0.23{]}; and marginal developmental increases among low-risk control children, r(8) = 0.60, CI {[}-0.04,0.89{]}, p=0.06{]}.
+Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(8) = -0.10, CI {[}-0.68,0.57{]}, p=0.79{]}; whereas younger low-risk control children showed a significant increase with age r(8) = 0.68, CI {[}0.08,0.92{]}, p=0.03{]}.
+Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman syndrome r(8) = -0.42, CI {[}-0.83,0.29{]}, p=0.23{]}; and marginal developmental increases among low-risk control children, r(8) = 0.60, CI {[}-0.04,0.89{]}, p=0.06{]}.
 
@@
 
+\begin{figure}
+\centering
 \includegraphics{paper_files/figure-latex/fig-prec-rec-1.pdf}
+\caption{\label{fig:fig-prec-rec}Number of segments as a function of their majority label from laboratory annotations (rows) versus Zooniverse annotations (columns).
+Percentages (and shading) indicate Precision (left) and Recall (right).
+For precision, the percentage indicates what percentage of the segments that have a given majority label according to Zooniverse data are attributed a given label in Lab data (i.e., columns add up to 100\%).
+For recall, the percentage indicates what percentage of the segments that have a given majority label according to Lab data are attributed a given label in Zooniverse data (i.e., rows add up to 100\%).}
+\end{figure}
 
@@
 Looking at all 20 children together, we found a strong positive correlation r(18) = 0.81, CI {[}0.57,0.92{]}, p=0{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman Syndrome r(8) = 0.88, CI {[}0.55,0.97{]}, p=0.00{]}; low-risk control r(8) = 0.82, CI {[}0.40,0.96{]}, p=0.00{]}.
+When we split by participant group, correlations remain high: for Angelman syndrome r(8) = 0.88, CI {[}0.55,0.97{]}, p=0.00{]}; low-risk control r(8) = 0.82, CI {[}0.40,0.96{]}, p=0.00{]}.
 
 Similarly, a strong positive correlation is found in the Canonical Proportion r(18) = 0.94, CI {[}0.84,0.98{]}, p=0{]}.
-When we split by participant group, correlations remain high although we do note they are somewhat smaller for the older children with Angelman Syndrome: r(8) = 0.86, CI {[}0.50,0.97{]}, p=0.00{]}; than the low-risk control children r(8) = 0.96, CI {[}0.83,0.99{]}, p=0.
+When we split by participant group, correlations remain high although we do note they are somewhat smaller for the older children with Angelman syndrome: r(8) = 0.86, CI {[}0.50,0.97{]}, p=0.00{]}; than the low-risk control children r(8) = 0.96, CI {[}0.83,0.99{]}, p=0.
 
@@
 We investigated this hypothesis by calculating the proportion of their data labeled as \enquote{Junk} for each individual child.
-There was no significant difference in the proportions of their data labeled as \enquote{Junk} for older children with Angelman Syndrome (M = 0.13) compared to the younger low risk children (M = 0.14): Welch's t(13.37) = -0.52, p = 0.61.
+There was no significant difference in the proportions of their data labeled as \enquote{Junk} for older children with Angelman syndrome (M = 0.13) compared to the younger low risk children (M = 0.14): Welch's t(13.37) = -0.52, p = 0.61.
 We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as \enquote{Junk}.
@@
 
-Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman Syndrome group: r(8) = -0.04, CI {[}-0.65,0.61{]}, p=0.92{]}, but increased with age in the low-risk control group: r(8) = 0.72, CI {[}0.16,0.93{]}, p=0.02.
-Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman Syndrome r(8) = -0.38, CI {[}-0.81,0.33{]}, p=0.28{]}; and developmental increases among low-risk control r(8) = 0.57, CI {[}-0.09,0.88{]}, p=0.08.
+Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman syndrome group: r(8) = -0.04, CI {[}-0.65,0.61{]}, p=0.92{]}, but increased with age in the low-risk control group: r(8) = 0.72, CI {[}0.16,0.93{]}, p=0.02.
+Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman syndrome r(8) = -0.38, CI {[}-0.81,0.33{]}, p=0.28{]}; and developmental increases among low-risk control r(8) = 0.57, CI {[}-0.09,0.88{]}, p=0.08.
 
 As for correlations between Zooniverse and laboratory-derived metrics, we observe very similar levels of correlation: Linguistic Proportion overall r(18) = 0.84, CI {[}0.64,0.94{]}, p=0{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman Syndrome r(8) = 0.84, CI {[}0.46,0.96{]}, p=0.00{]}; low-risk control r(8) = 0.84, CI {[}0.45,0.96{]}, p=0.00{]}.
-As for Canonical Proportion overall r(18) = 0.96, CI {[}0.91,0.98{]}, p=0; Angelman Syndrome: r(8) = 0.87, CI {[}0.52,0.97{]}, p=0.00{]}; low-risk control infants r(8) = 0.98, CI {[}0.89,0.99{]}, p=0.
+When we split by participant group, correlations remain high: for Angelman syndrome r(8) = 0.84, CI {[}0.46,0.96{]}, p=0.00{]}; low-risk control r(8) = 0.84, CI {[}0.45,0.96{]}, p=0.00{]}.
+As for Canonical Proportion overall r(18) = 0.96, CI {[}0.91,0.98{]}, p=0; Angelman syndrome: r(8) = 0.87, CI {[}0.52,0.97{]}, p=0.00{]}; low-risk control infants r(8) = 0.98, CI {[}0.89,0.99{]}, p=0.
 
@@
 Next, we looked at whether having more segments from each child may lead to more reliable metrics.
-We observed a non-significant trend {[}t(17.92)=527.40 and 631.90{]} for lower number of segments in the Angelman Syndrome group (mean 527.40, range 326-924) than among low-risk control infants (mean 631.90, range 351-1074), likely because of the lower volubility of the former children.
+We observed a non-significant trend {[}t(17.92)=527.40 and 631.90{]} for lower number of segments in the Angelman syndrome group (mean 527.40, range 326-924) than among low-risk control infants (mean 631.90, range 351-1074), likely because of the lower volubility of the former children.
 Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications (Cychosz et al., 2019).
@@
 We also observed quite a bit of variance across the 50 runs.
-For Linguistic Proportion, the range of correlations found for all infants was {[}0.64,0.88{]}; for the Angelman Syndrome group {[}0.76,0.94{]}; for the low-risk group {[}0.45,0.95{]}.
-For Canonical Proportion, the range of correlations found for all infants was {[}0.82,0.96{]}; for the Angelman Syndrome group {[}0.40,0.97{]}; for the low-risk group {[}0.75,0.97{]}.
+For Linguistic Proportion, the range of correlations found for all infants was {[}0.64,0.88{]}; for the Angelman syndrome group {[}0.76,0.94{]}; for the low-risk group {[}0.45,0.95{]}.
+For Canonical Proportion, the range of correlations found for all infants was {[}0.82,0.96{]}; for the Angelman syndrome group {[}0.40,0.97{]}; for the low-risk group {[}0.75,0.97{]}.
 
@@
 
-\caption{\label{tab:tab-cors}Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman Syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways: All seg indicates that Zooniverse annotations at the clip level were first combined at the segment level; Chunks that they were analyzed directly at the clip level.
+\caption{\label{tab:tab-cors}Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways: All seg indicates that Zooniverse annotations at the clip level were first combined at the segment level; Chunks that they were analyzed directly at the clip level.
 Both of these are based on all the data.
@@
 Overall, the pattern for Linguistic Proportion is compatible with a non-linear trajectory common to the two groups, whereby the Linguistic Proportion increases rapidly up to 20 months, and hovers around 90-100\% thereafter.
-In this case, the lack of correlation with age in the Angelman Syndrome group suggests that there is little systematic variance in this metric after 20 months of age.
-In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and slower decreases among children in the Angelman Syndrome group.
+In this case, the lack of correlation with age in the Angelman syndrome group suggests that there is little systematic variance in this metric after 20 months of age.
+In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and slower decreases among children in the Angelman syndrome group.
 
@@
 
-We believe additional research would be welcome on the two measures we used to describe children's linguistic productions: Linguistic Proportion and Canonical Proportion.
+Our results warrant additional research on two metrics we used to describe children's linguistic productions: Linguistic Proportion and Canonical Proportion.
 Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion.
@@
 
-Assuming these child-level metrics continue to be supported by such additional data, it would be interesting for future investigations to expand the analysis of Linguistic Proportion and Canonical Proportion to recover differences and similarities in language development between low-risk children and children with disorders other than Angelman Syndrome.
-That is, it is at present unclear whether relatively lower Canonical Proportions will also be found in other populations, or whether this is more specific to Angelman Syndrome.
+Assuming these child-level metrics continue to be supported by such additional data, it would be interesting for future investigations to expand the analysis of Linguistic Proportion and Canonical Proportion to recover differences and similarities in language development between low-risk children and children with disorders other than Angelman syndrome.
+That is, it is at present unclear whether relatively lower Canonical Proportions will also be found in other populations, or whether this is more specific to Angelman syndrome.
 For instance, some previous work has found differences in babbling even when comparing late talkers with toddlers with no specific diagnosis (Fasolo, Majorano, \& D'Odorico, 2008).
@@
 Our dataset only contained one daylong recording per child (with one exception), so other datasets in which a greater number of recordings are collected longitudinally can bring novel insights.
-Is the age-related decline we observed among children in the Angelman Syndrome group truly a developmental decline that happens by necessity in this population, or might it reflect partially other aspects of children's experience and behavior?
+Is the age-related decline we observed among children in the Angelman syndrome group truly a developmental decline that happens by necessity in this population, or might it reflect partially other aspects of children's experience and behavior?
 Additionally, longitudinal work within the same infant may be needed to track the \enquote{natural history} of vocalizations to better contextualize how these metrics should be interpreted.

### 2020-11-03 202010_jslhr/sup_mat.Rmd
knitting version of supplementary materials

@@
 ---
-title: "Sup mat for JSLHR version"
-date: "2020-10-17"
+title: "Supplementary materials to: Describing vocalizations in young children: A big data approach through citizen science annotations"
 output:
@@
     toc_depth: 3
-  html_document:
-    df_print: paged
-    toc: yes
-    toc_depth: '3'
 ---
@@
 
-- 2020-10-17 first version
-
-## TODO:
-
-- so much!
+- 2020-11-03 first version
 
@@
 
-Here we look at to what extent zooniverse and lab annotations match at the level of individual segments. Each data point is one segment (one "vocalization").
+Here we look at to what extent zooniverse and lab annotations match at the level of individual segments.
+Each data point is one segment (using LENA segmentation).
+Unlike in the main paper, here we will show results before applying the ordered rules that give prevalence to canonical, non-canonical, laughing, crying (in that order).
 
@@
 
-## repeat collapsing
+## Collapse across "mixed"
 
@@
 sample_data<-cbind(data_all$lab,data_all$Zoon_classif)
-gac(data = sample_data, kat = 5, weight = c("unweighted"),
-    conf.level = 0.95)
-mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
+
+```
+
+## Separate confusion matrices for Angelman syndrome children
+
+```{r}
+data_as_td<-left_join(data_all,demo_data,on="ChildID")
+# CM with just AS kids
+data_AS<-subset(data_as_td, Diagnosis=="AngelmanSyndrome")
+mycf=confusionMatrix(data_AS$lab, data_AS$Zoon_classif, dnn = c("Lab","Zooniverse"))
 conf_tab=mycf$table
-# this package uses sensitivity & specificity
-#Sensitivity=recall
-#Specificity=precision
 mycf
@@
 
-```{r prec2}
-pdf("../Results/precision_final.pdf",height=10,width=10)
+```{r prec2AS}
 colsums=colSums(conf_tab)
@@
         axis.text.x = element_text(angle=90, hjust=1))
-dev.off()
+
 ```
 
-```{r recall2}
-pdf("../Results/recall_final.pdf",height=10,width=10)
+```{r recall2_AS}
 prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
@@
         axis.text.x = element_text(angle=90, hjust=1))
-dev.off()
+
 ```
 
+## Separate confusion matrices with just the low risk controls
+
 ```{r}
-data_as_td<-left_join(data_all,demo_data,on="ChildID")
-# CM with just AS kids
-data_AS<-subset(data_as_td, Diagnosis=="AngelmanSyndrome")
-mycf=confusionMatrix(data_AS$lab, data_AS$Zoon_classif, dnn = c("Lab","Zooniverse"))
+# CM with just TD kids
+data_TD<-subset(data_as_td, Diagnosis=="Low-RiskControl")
+mycf=confusionMatrix(data_TD$lab, data_TD$Zoon_classif, dnn = c("Lab","Zooniverse"))
 conf_tab=mycf$table
@@
 ```
-```{r prec2AS}
-pdf("../Results/precision_AS.pdf",height=10,width=10)
+
+```{r prec3AS}
 colsums=colSums(conf_tab)
@@
         axis.text.x = element_text(angle=90, hjust=1))
-dev.off()
+
 ```
 
-```{r recall2_AS}
-pdf("../Results/recall_AS.pdf",height=10,width=10)
+```{r recall3}
 prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
@@
         axis.text.x = element_text(angle=90, hjust=1))
-dev.off()
-```
-
-```{r}
-# CM with just TD kids
-data_TD<-subset(data_as_td, Diagnosis=="Low-RiskControl")
-mycf=confusionMatrix(data_TD$lab, data_TD$Zoon_classif, dnn = c("Lab","Zooniverse"))
-conf_tab=mycf$table
-mycf
-```
-
-```{r prec3AS}
-pdf("../Results/precision_TD.pdf",height=10,width=10)
-colsums=colSums(conf_tab)
-my_conf_tab=conf_tab
-for(i in 1:5) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
-colSums(my_conf_tab)
-prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
-prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
-colnames(prop_cat)[3]<-"pr"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","pr")])
-ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(pr)), colour = "white") +
-  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=8) +
-  geom_text(aes(label = Freq), vjust = 1,size=8) +
-  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
-     theme(legend.position = "none") +
-  xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Precision")+theme(text = element_text(size=20),
-        axis.text.x = element_text(angle=90, hjust=1))
-dev.off()
-```
-
-```{r recall3}
-pdf("../Results/recall_TD.pdf",height=10,width=10)
-prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
-prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
-colnames(prop_cat)[3]<-"rec"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","rec")])
-ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(rec)), colour = "white") +
-  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=8) +
-  geom_text(aes(label = Freq), vjust = 1,size=8) +
-  scale_fill_gradient(low = "white", high = "red", name = "Proportion") +
-     theme(legend.position = "none") +
-  xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Recall")+theme(text = element_text(size=20),
-        axis.text.x = element_text(angle=90, hjust=1))
-dev.off()
-```
-
-## Child level descriptors
-
-Although there may be errors at the level of the segment, what we really care about is whether Zooniverse annotations give a reliable image of the child's individual development.
-This is what we look at in this section.
-In all of these graphs, red points correspond to children diagnosed with Angelman Syndrome, black for low-risk control.
-
-```{r prepare data by child}
-#get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
-ztab=table(data_all$filename,data_all$Zoon_classif)
-z_lr=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
-z_cr=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
-ltab=table(data_all$filename,data_all$lab)
-l_lr=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
-l_cr=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
-#put all the ratios together
-if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) ratios=cbind(rownames(ztab),z_lr,z_cr,l_lr,l_cr) else print("oops this code needs to be more complex because we don't have the same kids for the two ratios")
-colnames(ratios)[1]<-"filename"
-
-#ages=aggregate(data_all$Age,by=list(data_all$ChildID),mean) #this is a weird way of adding ages, since all of the ages for a given child should be the same if there is only one recording, and if there are multiple recordings, then we should not get the mean
-#improvement: now we merge with a demo data tab, but note this is merged with child id, so the problem of multiple recs per child is still there
-
-# Created demo_data with filenames. Use filenames instead of childIDs to merge ratios and demo data.
-
-merge(ratios,demo_data_fn,by="filename")->ratios
-colnames(ratios)[dim(ratios)[2]]<-"Age"
-#cbinding results in text, so we numerize the ratios
-for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) ratios[,thisvar]=as.numeric(as.character(ratios[,thisvar]))
-summary(ratios)
-```
-
-## Correlations with age
-
-We first look generally at two measures that have been found to relate to age:
-
-- linguistic ratio = ("Canonical"+"Non-Canonical")/"All vocalizations" (i.e. we remove junk)
-- canonical ratio = "Canonical"/("Canonical"+"Non-Canonical") (i.e. we remove junk + non-linguistic vocalizations)
-
-TODO
-
-- make margins smaller
-
-- remove title repetition (age only at the bottom)
-
-```{r corage}
-
-prettynames=c("Linguistic Ratio (Zooniverse)","Canonical Ratio (Zooniverse)",
-             "Linguistic Ratio (Lab)","Canonical Ratio (Lab)" )
-names(prettynames)<-c("z_lr","z_cr","l_lr","l_cr")
-mycols=c("black","red")
-names(mycols)<-c("Low-RiskControl","AngelmanSyndrome")
-mypch=c(4,20)
-names(mypch)<-c("Low-RiskControl","AngelmanSyndrome")
-
-jpeg("../Results/corage.jpg",width=20,height=20,units="cm",res=300)
-layout(matrix(c(1:4), 2, 2, byrow = F))
-for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
-
-  plot(ratios[,thisvar]~ratios$Age, pch=mypch[ratios$Diagnosis],xlab="Age (months)",ylab=prettynames[thisvar],
-       col=mycols[ratios$Diagnosis])
-  abline(lm(ratios[,thisvar]~ratios$Age,subset=c(ratios$Diagnosis=="AngelmanSyndrome")),col="black")
-  myr=paste0("r=",round(cor.test(ratios[ratios$Diagnosis=="AngelmanSyndrome",thisvar],ratios$Age[ratios$Diagnosis=="AngelmanSyndrome"])$estimate,3))
-  text(mean(ratios$Age[ratios$Diagnosis=="AngelmanSyndrome"]),mean(ratios[ratios$Diagnosis=="AngelmanSyndrome",thisvar]),myr,col="black")
-    abline(lm(ratios[,thisvar]~ratios$Age,subset=c(ratios$Diagnosis!="AngelmanSyndrome")),col="red")
-    myr=paste0("r=",round(cor.test(ratios[ratios$Diagnosis!="AngelmanSyndrome",thisvar],ratios$Age[ratios$Diagnosis!="AngelmanSyndrome"])$estimate,3))
-    text(mean(ratios$Age[ratios$Diagnosis!="AngelmanSyndrome"]),mean(ratios[ratios$Diagnosis!="AngelmanSyndrome",thisvar]),myr,col="red")
-}
-dev.off()
-```
-
-## Correlations across zooniverse and lab
-
-But the key thing for us: Are Zooniverse annotations describing children similar to lab annotations? The answer is clearly yes.
-
-TODO:
-
-- add code to print out the results paragraphs
-- change figure generation to have lines fitted to each group
-
-```{r corlab-zoo}
-#Ling ratio
-pdf("../Results/ling_rat_z_vs_l_final.pdf",height=5,width=5)
-lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]))
-  myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
-  plot(ratios[,"z_lr"]~ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims,
-       col=mycols[ratios$Diagnosis])
-  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]))
-  lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
-  #CR
-pdf("../Results/can_rat_z_vs_l_final.pdf",height=5,width=5)
-lims=range(c(ratios[,"z_cr"],ratios[,"l_cr"]))
-    myr=round(cor.test(ratios[,"z_cr"],ratios[,"l_cr"])$estimate,3)
-  plot(ratios[,"z_cr"]~ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims,
-       col=mycols[ratios$Diagnosis])
-  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]),col="darkgray")
-    lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
 
 ```
-
-```{r regmod}
-lin_mod=lm(z_lr~l_lr*Diagnosis*Age,data=ratios)
-summary(lin_mod)
-plot(lin_mod)
-```
-
-## Code junkyard
-
-This is redundant code
-
-Different pdfs for diff groups
-
-```{r corage2, eval=F}
-prettynames=c("AS Linguistic Ratio (Zooniverse)","AS Canonical Ratio (Zooniverse)",
-             "AS Linguistic Ratio (Lab)","AS Canonical Ratio (Lab)" )
-names(prettynames)<-c("z_lr","z_cr","l_lr","l_cr")
-
-AS_ratios<-subset(ratios, Diagnosis=="AngelmanSyndrome")
-TD_ratios<-subset(ratios, Diagnosis=="Low-RiskControl")
-
-for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
-  myr=round(cor.test(AS_ratios[,thisvar],AS_ratios$Age)$estimate,3)
-  plot(AS_ratios[,thisvar]~AS_ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr))
-  abline(lm(AS_ratios[,thisvar]~AS_ratios$Age))
-  }
-```
-
-```{r corage3, eval=F}
-prettynames=c("TD Linguistic Ratio (Zooniverse)","TD Canonical Ratio (Zooniverse)",
-             "TD Linguistic Ratio (Lab)","TD Canonical Ratio (Lab)" )
-for(thisvar in c("z_lr","z_cr","l_lr","l_cr")) {
-  myr=round(cor.test(TD_ratios[,thisvar],TD_ratios$Age)$estimate,3)
-  plot(TD_ratios[,thisvar]~TD_ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr))
-  abline(lm(TD_ratios[,thisvar]~TD_ratios$Age))
-  }
-```
-
-```{r corlab-zoo2, eval=F}
-#Ling ratio
-pdf("../Results/td_ling_rat_z_vs_l_final.pdf",height=5,width=5)
-lims=range(c(TD_ratios[,"z_lr"],TD_ratios[,"l_lr"]))
-  myr=round(cor.test(TD_ratios[,"z_lr"],TD_ratios[,"l_lr"])$estimate,3)
-  plot(TD_ratios[,"z_lr"]~TD_ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims,
-       col=mycols[TD_ratios$Diagnosis])
-  abline(lm(TD_ratios[,"z_lr"]~TD_ratios[,"l_lr"]))
-  lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
-  #CR
-pdf("../Results/td_can_rat_z_vs_l_final.pdf",height=5,width=5)
-lims=range(c(TD_ratios[,"z_cr"],TD_ratios[,"l_cr"]))
-    myr=round(cor.test(TD_ratios[,"z_cr"],TD_ratios[,"l_cr"])$estimate,3)
-  plot(TD_ratios[,"z_cr"]~TD_ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims,
-       col=mycols[TD_ratios$Diagnosis])
-  abline(lm(TD_ratios[,"z_cr"]~TD_ratios[,"l_cr"]),col="darkgray")
-    lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
-```
-
-```{r corlab-zoo3, eval=F}
-#Ling ratio
-pdf("../Results/AS_ling_rat_z_vs_l_final.pdf",height=5,width=5)
-lims=range(c(AS_ratios[,"z_lr"],AS_ratios[,"l_lr"]))
-  myr=round(cor.test(AS_ratios[,"z_lr"],AS_ratios[,"l_lr"])$estimate,3)
-  plot(AS_ratios[,"z_lr"]~AS_ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims,
-       col=mycols[AS_ratios$Diagnosis])
-  abline(lm(AS_ratios[,"z_lr"]~AS_ratios[,"l_lr"]))
-  lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
-  #CR
-pdf("../Results/AS_can_rat_z_vs_l_final.pdf",height=5,width=5)
-lims=range(c(AS_ratios[,"z_cr"],AS_ratios[,"l_cr"]))
-    myr=round(cor.test(AS_ratios[,"z_cr"],AS_ratios[,"l_cr"])$estimate,3)
-  plot(AS_ratios[,"z_cr"]~AS_ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
-       xlim=lims,ylim=lims,
-       col=mycols[AS_ratios$Diagnosis])
-  abline(lm(AS_ratios[,"z_cr"]~AS_ratios[,"l_cr"]),col="darkgray")
-    lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
-```
-
-combined pdf
-
-```{r , eval=F}
-#COMBINED to save space
-pdf("../Results/combined_final.pdf",height=5,width=5)
-lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]),c(ratios[,"z_cr"],ratios[,"l_cr"]))
-  #myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
-  plot(ratios[,"z_lr"]~ratios[,"l_lr"],xlab="Laboratory annotations",ylab="Zooniverse annotations",
-       xlim=lims,ylim=lims,
-       pch=20,col=mycols[ratios$Diagnosis])
-    points(ratios[,"z_cr"]~ratios[,"l_cr"], pch=2, col=mycols[ratios$Diagnosis])
-  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]))
-  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]),lty=3)
- # lines(c(0,1),c(0,1),lty=2,col="darkgray")
-dev.off()
-```

### 2020-11-05 202010_jslhr/paper.Rmd
incorporated Lisa comments

@@
 ---
-title: "Describing vocalizations in young children: A big data approach through citizen science annotations"
+title: "Describing vocalizations in young children: A big data approach through citizen science annotation"
 shorttitle        : "Children's vocalizations and citizen science"
@@
 Since early language delays can adversely affect children’s literacy, behavior, social interaction, and scholastic achievement extending well into adulthood, early interventions have been described as a better societal investment than later ones [@heckman2006skill].
-Advancements in the field of wearable technologies, such as LENA^TM^ recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally.
+Some research suggests that spontaneous behavior, captured for instance via home videos, could provide important indices to development [@belardi2017retrospective;@overby2020retrospective].
+Advancements in the field of wearable technologies, such as LENA^TM^ recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally [@rankine2017language;@vandam2019use;@oller2010automated].
 Wearable recorders allow data collection to happen in the child’s natural environment, and at a large scale, which may be particularly helpful for children whose speech is not easily elicited.
@@
 There is a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this work focuses on babbling and thus typically on infants under one year of age.
-A critical milestone involves the increasingly common production of Canonical Syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech [@oller1998late].
+A critical milestone involves the increasingly common production of canonical syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech [@oller1998late].
 Given its adult-like consonant-vowel or vowel-consonant structure, canonical babble is considered to be a starting point on the path to recognizable speech.
@@
 Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development [@oller2000emergence].
-In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization count in a sample of children with an Autism Spectrum Disorder [@mcdaniel2020predicting].
+In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization count in a sample of children diagnosed with an autism spectrum disorder [@mcdaniel2020predicting].
 
@@
 For the English and Tsimane' corpora, vocalizations were automatically identified using LENA^TM^, whereas the other two were extracted through manual segmentation.
-Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received 3 classifications into crying, laughing, canonical, non-canonical, or "Junk" (with the latter tag used for clips that did not contain clear child voice).
+Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received three classifications into Crying, Laughing, Canonical, Non-Canonical, or Junk (with the latter tag used for clips that did not contain a child voice).
 They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgement of canonical or non-canonical.
-They found that this canonical ratio increased with infant age in an analysis collapsing across corpora, which provides a first proof of principle that citizen scientists' annotations could be an appropriate solution to the problem of annotating child vocalizations from wearables.
-The authors also studied how their canonical ratio related to age within each corpus that had a sufficient number of children varying in age.
+They found that this Canonical Proportion increased with infant age in an analysis collapsing across corpora, which provides a first proof of principle that citizen scientists' annotations could be an appropriate solution to the problem of annotating child vocalizations from wearables.
+The authors also studied how their Canonical Proportion related to age within each corpus that had a sufficient number of children varying in age.
 They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the "Junk" category was very high.
-Additionally, they found that the correlation between canonical ratio and age was much weaker in two corpora where vocalizations were segmented with LENA^TM^ than through manual segmentation.
+Additionally, they found that the correlation between Canonical Proportion and age was much weaker in two corpora where vocalizations were segmented with LENA^TM^ than through manual segmentation.
 In sum, these analyses within corpora reveal that data may not be uniformly useful, with potential variation across corpora, children, and methods.
@@
 
-When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (canonical+non-canonical)/(all vocalizations); and the Canonical Proportion (canonical)/(all linguistic vocalizations).
-Previous work suggests that children with Angelman syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the canonical ratio was expected to increase with age [@hamrick2019capturing].
+When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk vocalizations); and the Canonical Proportion (Canonical)/(all linguistic vocalizations).
+Previous work suggests that children with Angelman syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the Canonical Proportion was expected to increase with age [@hamrick2019capturing;@hamrick2019measurement].
 Although we could not rely on previous work to make predictions regarding the Linguistic Proportion, we reasoned that we should observe an increase in the Linguistic Proportion for both populations.
@@
 As children diagnosed with Angelman syndrome typically have severe cognitive and language delays and most do not produce more than 1-2 words consistently, we compare their linguistic production data against that of younger children with theoretically similar language profiles.
+This study was approved by the local institutional review board under the project name Neurodevelopmental Natu- ral History Study, Purdue University, IRB-1811021381.
 
@@
 Before annotation, coders completed an ethical conduct of research course and received HIPAA compliance training.
-After this, they complete a brief training on vocal maturity that involves 1) reading relevant literature, 2) reviewing examples of various annotation categories and 3) completing a brief quiz to test their annotation accuracy.
-Coders then proceed to the classification, where they make decisions on whether a segment is a cry, laugh, non-canonical syllable(s), canonical syllable(s), or a word.
-Coders are instructed to classify each segment based on the highest level of vocal maturity it contains (e.g., a segment with both canonical and non-canonical syllables would be classified as canonical; a segment with both laughing and non-canonical syllables would be classified as non-canonical).
+After this, they completed a brief training on vocal maturity that involved 1) reading relevant literature, 2) reviewing examples of various annotation categories and 3) submitting answers to a brief quiz to test their annotation accuracy.
+Coders then proceeded to the classification, where they made decisions on whether a segment is a cry, laugh, non-canonical syllable(s), canonical syllable(s), or a word.
+Coders were instructed to classify each segment based on the highest level of vocal maturity it contains (e.g., a segment with both canonical and non-canonical syllables would be classified as canonical; a segment with both laughing and non-canonical syllables would be classified as non-canonical).
 For canonical, non-canonical, and word segments, coders indicated how many syllables of each type of vocal maturity they hear (e.g., "ah ma ba" would be 1 non-canonical and 2 canonical syllables).
@@
 
-This process resulted in high levels of agreement: 97.22% of the lab-annotated segments had either 2 or 3 coders agree.
+This process resulted in high levels of inter-rater agreement: Two or more of the three total coders who annotated that segment agreed on 97.22% of the lab-annotated segments.
 
@@
 A substantial concern that emerges when considering citizen science annotation for spontaneous speech data is the risk of a privacy breach: even short clips can contain personal information that can expose the identity of the speaker, such as addresses or names, or contain sensitive and private information.
-Fortunately, @Seidl19 determined that when segments are divided into shorter chunks (400-600ms), human annotators with little training can code our categories of interest (canonical, laughing, crying, etc.) with a classification quality comparable to the one carried out on full segments.
+Fortunately, @Seidl19 determined that when segments are divided into shorter chunks (400-600ms), human annotators with little training can code our categories of interest (Canonical, Laughing, Crying, etc.) with a classification quality comparable to the one carried out on full segments.
 We therefore used very short chunks (500ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.
@@
 
-Citizen scientists could access the project by logging in, or as anonymous participants.
+Citizen scientists could access our project, "Maturity of Baby Sounds", by logging in, or as anonymous participants.
 Before starting the annotation, participants were given a quick tutorial which walked them through the steps of the annotation workflow (i.e., how to play a sound, how to make a selection; see Figure 1).
 In this tutorial, we included one audio example for each category in order to make the classification task as smooth as possible.
-Further clarifications on what constitutes Canonical and Non-canonical utterances, as well as five audio examples for each of the categories, were available through the Field Guide, which users could access by clicking on the right side of the screen.
+Further clarifications on what constitutes canonical and non-canonical sequences, as well as five audio examples for each of the categories, were available through the Field Guide, which users could access by clicking on the right side of the screen.
 The Field Guide could be consulted at any point of the classification without interrupting the task.
 
-Users were asked to assign each 500ms clip to one out of five possible categories: (1) canonical, (2) non-canonical, (3) crying, (4) laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
+Users were asked to assign each 500ms clip to one out of five possible categories: (1) Canonical, (2) Non-Canonical, (3) Crying, (4) Laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
 Notice that the latter is narrower than the "Don't mark" category of the lab annotation routine, which was used any time there was noise or overlap that could affect acoustic analyses.
 
-Following recommendations from the Zooniverse board, we collected 5 judgments per clip, rather than 3 as used in the laboratory.
-Majority agreement is thus achieved when 3 out of 5 judgements agree (rather than 2 out of 3).
+Following recommendations from the Zooniverse board, we collected five judgments per clip, rather than three as used in the laboratory.
+Majority agreement is thus achieved when three out of five judgements agree (rather than two out of three).
 
@@
 
-``` {r fig-zoo, comment=F, message=F, hide = T, warning = F, echo = F,  fig.cap="Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needs to make; as well as the field guide, which remains accessible on the right of the screen throughout the session, and which allows users to revise examples of the different categories."}
+``` {r fig-zoo, comment=F, message=F, hide = T, warning = F, echo = F,  fig.cap="Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needed to make; as well as the Field Guide, which remained accessible on the right of the screen throughout the session, and which allowed users to revise examples of the different categories."}
 knitr::include_graphics("zooniverse-pufig.pdf")
@@
 
+<!-- lh said: I'm also wondering what the breakdown of lab annotations was for the Zooniverse segments without majority agreement -->
+
 Lab annotators provided judgments at the level of LENA^TM^ segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments.
@@
 
-- if one canonical judgement was present, then the segment was labelled as Canonical, 
+- if one Canonical judgement was present, then the segment was labelled as Canonical, 
 - else, if a Non-Canonical judgement was present, the segment was labelled as Non-Canonical, 
@@
 
-This is essentially the same thought process expert annotators followed when labeling segment-level audio, as each segment can represent multiple classification subtypes).
+This is essentially the same thought process expert annotators followed when labeling segment-level audio, as each segment can represent multiple classification subtypes.
 LENA^TM^ segments are not equivalent to child vocalizations: A segment may contain 1 or more vocalizations, which may be of the same type or not.
@@
 ```{r clean up, echo=F}
-# remove non-majority labels
-#TODO
+# remove non-majority labels from lab
+data_all[data_all$Num_Agreement>2,]->data_all
 
@@
 
-#get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
+#get the ns by child, then calculate the linguistic proportion & canonical proportion, separately for zooniverse & lab
 ztab=table(data_all$filename,data_all$Zoon_classif)
@@
 In this section, we provide descriptive analyses of our dataset.
-According to lab annotators, `r table_lab["Canonical"]`% of segments were canonical, `r table_lab["Non-Canonical"]`% non-canonical, `r table_lab["Laughing"]`% laughing, and `r table_lab["Crying"]`% crying, with the remaining `r table_lab["Junk"]`% being categorized as "Don't mark".
-Zooniverse data revealed a similar distribution:  `r table_zoon["Canonical"]`% canonical, `r table_zoon["Non-Canonical"]`% non-canonical, `r table_zoon["Laughing"]`% laughing, `r table_zoon["Crying"]`% crying, `r table_zoon["Junk"]`% Junk.
-Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = ("Canonical"+"Non-Canonical")/"All vocalizations" (i.e., we remove, Junk), and ii) Canonical Proportion = "Canonical"/("Canonical"+"Non-Canonical") (i.e., we remove Junk + non-linguistic).
+According to lab annotators, `r table_lab["Canonical"]`% of segments were Canonical, `r table_lab["Non-Canonical"]`% Non-Canonical, `r table_lab["Laughing"]`% Laughing, and `r table_lab["Crying"]`% Crying, with the remaining `r table_lab["Junk"]`% being categorized as "Don't mark".
+Zooniverse data revealed a similar distribution:  `r table_zoon["Canonical"]`% Canonical, `r table_zoon["Non-Canonical"]`% Non-Canonical, `r table_zoon["Laughing"]`% Laughing, `r table_zoon["Crying"]`% Crying, `r table_zoon["Junk"]`% Junk.
+Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical+Non-Canonical)/All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical/(Canonical+Non-Canonical) (i.e., we remove Junk, Crying, and Laughing).
 See Figure 2 for results.
 
-```{r fig-corage, echo=F,fig.cap="Correlations between child-level descriptors and age as a function of metric (linguistic ratio in the top row, canonical ratio in the bottom row), annotation method, and child group (red=low-risk, and black Angelman syndrome)."}
-
-prettynames=c("Linguistic Proportion (Zooniverse)","Canonical Proportion (Zooniverse)",
+```{r fig-corage, echo=F,fig.cap="Correlations between child-level descriptors and age as a function of metric (Linguistic Proportion in the top row, Canonical Proportion in the bottom row), annotation method, and child group (red=low-risk, and black= Angelman syndrome)."}
+
+prettynames=c("Linguistic Proportion (Zoon.)","Canonical Proportion (Zoon.)",
              "Linguistic Proportion (Lab)","Canonical Proportion (Lab)" )
@@
 Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups.
-There was a near-zero relationship among the older children diagnosed with Angelman syndrome r(`r myn["l_lp","AS"]`) = `r myr["l_lp","AS"]`, CI [`r mylb["l_lp","AS"]`,`r myhb["l_lp","AS"]`], p=`r myp["l_lp","AS"]`]; and a significant association among younger low-risk control children r(`r myn["l_lp","LR"]`) = `r myr["l_lp","LR"]`, CI [`r mylb["l_lp","LR"]`,`r myhb["l_lp","LR"]`], p=`r myp["l_lp","LR"]`].
-The Canonical Proportion exhibited non-significant developmental decreases among older children diagnosed with Angelman syndrome r(`r myn["l_cp","AS"]`) = `r myr["l_cp","AS"]`, CI [`r mylb["l_cp","AS"]`,`r myhb["l_cp","AS"]`], p=`r myp["l_cp","AS"]`]; and marginal developmental increases among low-risk control r(`r myn["l_cp","LR"]`) = `r myr["l_cp","LR"]`, CI [`r mylb["l_cp","LR"]`,`r myhb["l_cp","LR"]`], p=`r myp["l_cp","LR"]`].
-
-Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(`r myn["z_lp","AS"]`) = `r myr["z_lp","AS"]`, CI [`r mylb["z_lp","AS"]`,`r myhb["z_lp","AS"]`], p=`r myp["z_lp","AS"]`]; whereas younger low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r myr["z_lp","LR"]`, CI [`r mylb["z_lp","LR"]`,`r myhb["z_lp","LR"]`], p=`r myp["z_lp","LR"]`].
-Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman syndrome r(`r myn["z_cp","AS"]`) = `r myr["z_cp","AS"]`, CI [`r mylb["z_cp","AS"]`,`r myhb["z_cp","AS"]`], p=`r myp["z_cp","AS"]`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r myr["z_cp","LR"]`, CI [`r mylb["z_cp","LR"]`,`r myhb["z_cp","LR"]`], p=`r myp["z_cp","LR"]`].
+There was a near-zero relationship among the children diagnosed with Angelman syndrome r(`r myn["l_lp","AS"]`) = `r myr["l_lp","AS"]`, CI [`r mylb["l_lp","AS"]`,`r myhb["l_lp","AS"]`], p=`r myp["l_lp","AS"]`]; and a significant association among low-risk control children r(`r myn["l_lp","LR"]`) = `r myr["l_lp","LR"]`, CI [`r mylb["l_lp","LR"]`,`r myhb["l_lp","LR"]`], p=`r myp["l_lp","LR"]`].
+The Canonical Proportion exhibited non-significant developmental decreases among children diagnosed with Angelman syndrome r(`r myn["l_cp","AS"]`) = `r myr["l_cp","AS"]`, CI [`r mylb["l_cp","AS"]`,`r myhb["l_cp","AS"]`], p=`r myp["l_cp","AS"]`]; and marginal developmental increases among low-risk control r(`r myn["l_cp","LR"]`) = `r myr["l_cp","LR"]`, CI [`r mylb["l_cp","LR"]`,`r myhb["l_cp","LR"]`], p=`r myp["l_cp","LR"]`].
+
+Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(`r myn["z_lp","AS"]`) = `r myr["z_lp","AS"]`, CI [`r mylb["z_lp","AS"]`,`r myhb["z_lp","AS"]`], p=`r myp["z_lp","AS"]`]; whereas low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r myr["z_lp","LR"]`, CI [`r mylb["z_lp","LR"]`,`r myhb["z_lp","LR"]`], p=`r myp["z_lp","LR"]`].
+Similarly, there were non-significant developmental decreases in the Canonical Proportion among children with Angelman syndrome r(`r myn["z_cp","AS"]`) = `r myr["z_cp","AS"]`, CI [`r mylb["z_cp","AS"]`,`r myhb["z_cp","AS"]`], p=`r myp["z_cp","AS"]`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r myr["z_cp","LR"]`, CI [`r mylb["z_cp","LR"]`,`r myhb["z_cp","LR"]`], p=`r myp["z_cp","LR"]`].
 
@@
 ```{r fig-prec-rec, echo=F,fig.cap="Number of segments as a function of their majority label from laboratory annotations (rows) versus Zooniverse annotations (columns).
-Percentages (and shading) indicate Precision (left) and Recall (right).
+Percentages (and shading) indicate precision (left) and recall (right).
 For precision, the percentage indicates what percentage of the segments that have a given majority label according to Zooniverse data are attributed a given label in Lab data (i.e., columns add up to 100%).
@@
 
-```{r corlab-zoo, echo=F}
+```{r fig-corlab-zoo, echo=F,fig.cap="Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of child group (red=low-risk, and black=Angelman syndrome)."}
 
@@
 Although the classification at the clip level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of childrens’ individual development.
+We illustrate this association in Figure 3.
 Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
@@
 Similarly, a strong positive correlation is found in the Canonical Proportion r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`].
-When we split by participant group, correlations remain high although we do note they are somewhat smaller for the older children with Angelman syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
-
-```{r glmer-agediag, echo=F}
-# one outlier in ling prop is removed
-
-ling_all=glmer(lab_ling~zoo_ling*Diagnosis*Age + (1|ChildID),data=data_all,family="binomial",subset=c(ChildID!="3681_1"))
-#plot(lin_mod)
-# doesn't converge
-
-ling_AS=glmer(lab_ling~zoo_ling*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis=="AngelmanSyndrome"))
-#only effect of zoo
-
-ling_LR=glmer(lab_ling~zoo_ling*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis!="AngelmanSyndrome" & ChildID!="3681_1"))
-#both main and interaction are sig
-
-can_all=glmer(lab_can~zoo_can*Diagnosis*Age + (1|ChildID),data=data_all,family="binomial")
-#doesn't converge either
-
-can_AS=glmer(lab_can~zoo_can*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis=="AngelmanSyndrome"))
-#only effect of zoo
-
-can_LR=glmer(lab_can~zoo_can*Age + (1|ChildID),data=data_all,family="binomial",subset=c(Diagnosis!="AngelmanSyndrome"))
-#both main and interaction are sig
-
-```
+When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
 
@@
 
-```{r cor-Junk, echo=F}
+```{r fig-cor-Junk, echo=F,fig.cap="Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of whether children's Junk proportion was higher or lower than the median for all children (blue=lower, and green=higher)."}
 
@@
 We investigated this hypothesis by calculating the proportion of their data labeled as "Junk" for each individual child.
-There was no significant difference in the proportions of their data labeled as "Junk" for older children with Angelman syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the younger low risk children (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, p = `r round(JunkXdiag$p.value,3)`.
+There was no significant difference in the proportions of their data labeled as "Junk" for children with Angelman syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the low-risk group (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, p = `r round(JunkXdiag$p.value,3)`.
 We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk".
-Results were similar across these two post-hoc subgroups.
+Results were similar across these two post-hoc subgroups (see Figure 4).
 For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["lp","lower_Junk"]`) = `r myr["lp","lower_Junk"]`, CI [`r mylb["lp","lower_Junk"]`,`r myhb["lp","lower_Junk"]`], p=`r myp["lp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["lp","higher_Junk"]`) = `r myr["lp","higher_Junk"]`, CI [`r mylb["lp","higher_Junk"]`,`r myhb["lp","higher_Junk"]`], p=`r myp["lp","higher_Junk"]`.
@@
 As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
-We therefore repeated our analyses but deriving our proportions for the Zooniverse data not from the segment-level composite, but rather the individual chunk-level annotations.
+We therefore repeated our analyses but instead of deriving our proportions for the Zooniverse data from the segment-level composite, we did it based on the individual chunk-level annotations.
 
@@
 
-  #get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
+  #get the ns by child, then calculate the linguistic proportion & canonical proportion, separately for zooniverse & lab
   ztab=table(sampling$ChildID,sampling$Zoon_classif)
@@
 We repeated this process 50 times, to assess the extent to which the association between laboratory- and Zooniverse-derived metrics varied in each random sample.
-The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table XX), particularly for the low-risk younger infants, and for Linguistic Proportion.
-This may indicate that, particularly in some groups of infants, 100 segments may not be sufficient to capture the child's linguistic and Canonical Proportions.
+The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table 2), particularly for the low-risk younger infants, and for Linguistic Proportion.
+This may indicate that, particularly in some groups of infants, 100 segments may not be sufficient to capture a stable estimate of the child's Linguistic and Canonical Proportions.
 We also observed quite a bit of variance across the 50 runs.
@@
 colnames(mycortab)<-c("Linguistic Proportion","Canonical Proportion")
-kable(mycortab,cap="Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways: All seg indicates that Zooniverse annotations at the clip level were first combined at the segment level; Chunks that they were analyzed directly at the clip level.
+
+mycortab_sorted= rbind(mycortab[grep("AS",rownames(mycortab)),],
+                       mycortab[grep("LR",rownames(mycortab)),],
+                       mycortab[grep("all",rownames(mycortab)),]
+)
+
+kable(mycortab_sorted,cap="Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways.
+First, All seg indicates that Zooniverse annotations at the clip level were first combined at the segment level.
+Second, Chunks indicates that they were analyzed directly at the clip level.
 Both of these are based on all the data.
-In contrast, rows labeled 100 seg indicate that laboratory- and Zooniverse-derived metrics were based on only 100 segments (median over 50 runs in which 100 segments were randomly selected from each child).")
+Third, rows labeled 100 seg indicate that laboratory- and Zooniverse-derived metrics were based on only 100 segments (median over 50 runs in which 100 segments were randomly selected from each child).")
 ```
@@
 In the present study we assessed the extent to which child vocalizations from LENA^TM^ daylong recordings can be accurately described based on classifications collected using the citizen science platform Zooniverse.
-Our recordings came from children diagnosed with a language disorder, as well as low-risk controls, and included a range of different ages.
+Our recordings came from children diagnosed with a genetic syndrome associated with severe language disorder, as well as low-risk controls, and included a range of different ages.
 Excerpts of these recordings had previously been annotated by trained experts in the lab, and served as the "gold standard".
@@
 Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the "Junk" category less) than laboratory annotators, who have access to the whole vocalization.
-It is also plausible, on the other hand, that lab annotators are more "picky" with the data, and tend to reject clips (tag them as "Don't mark").
-Indeed, laboratory coders were instructed to be  sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use "Don't mark" if there was any noise or verlapping speech that would potentially affect acoustic analyses.
+This is consistent with the fact that our lab definition of "Junk" was more stringent than the Zooniverse definition.
+Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use "Don't mark" if there was *any* noise or verlapping speech that would potentially affect acoustic analyses.
 
@@
 
-As for laughing, the recall is quite good but precision is low, with confusion involving Junk and Non-Canonical to similar extents.
+As for Laughing, the recall is quite good but precision is low, with confusion involving "Junk" and Non-Canonical to similar extents.
 We are tentative in our interpretation of this result because the category was very sparsely represented, with barely 200 segments classified as Laughing in the whole data set.
@@
 Our pipeline to upload data on Zooniverse is scalable and quickly adapted to new datasets, including data on infants of different ages, learning different languages, or living in environments with different acoustic properties (e.g., rural versus urban environments).
-This means large existing datasets of daylong recordings can be reliably annotated in an ecological way using citizen science (ref on existing datasets).
+This means large existing datasets of daylong recordings, notably those hosted on Homebank [@vandam2016homebank], can be reliably annotated in an ecological way using citizen science.
 
@@
 Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion.
-We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant for other disorders that are related to emotional disorders [a review in @halpern2016excessive].
+We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant for other disorders that are related to emotional disorders [see a review in @halpern2016excessive].
 Additionally, while Canonical Proportion has been studied in much previous work, we believe further research is necessary using larger scales (saliently increasing the sheer number of children studied, to have sufficient power to detect group effects) and with broader coverage [notably including children growing up in a broad range of languages and cultures, as in @cychosz2019canonical], to make sure that assumptions regarding cross-linguistic and maturational universality are amply justified and not merely a result of having low power to detect differences, or simply sampling from similar populations.
@@
 We analyzed the correspondence between annotations at the individual clip level, and the individual-child level, using Canonical Proportion and Linguistic Proportion as descriptors.
-We found a moderate to good accuracy in the first, and strong positive correlations in the latter.
+We found moderate to good accuracy in the first, and strong positive correlations in the latter.
 We can conclude that citizen scientists are a reliable source of fast and ecological annotation of speech data, particularly when results are combined into child-level descriptors.
-Our study suggests that the same methodology may be applied to several research questions in the study of language acquisition and language disorders.
+The same methodology may be applied to several research questions in the study of language acquisition and language disorders.
 This finding is particularly welcome in an era when wearables open new avenues for studying human behavior and development in an ecological manner.

### 2020-11-05 202010_jslhr/paper.tex
incorporated Lisa comments

@@
 \hypersetup{
-            pdftitle={Describing vocalizations in young children: A big data approach through citizen science annotations},
+            pdftitle={Describing vocalizations in young children: A big data approach through citizen science annotation},
             pdfborder={0 0 0},
@@
 
-\title{Describing vocalizations in young children: A big data approach through citizen science annotations}
+\title{Describing vocalizations in young children: A big data approach through citizen science annotation}
 
@@
 Since early language delays can adversely affect children's literacy, behavior, social interaction, and scholastic achievement extending well into adulthood, early interventions have been described as a better societal investment than later ones (Heckman, 2006).
-Advancements in the field of wearable technologies, such as LENA\textsuperscript{TM} recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally.
+Some research suggests that spontaneous behavior, captured for instance via home videos, could provide important indices to development (Belardi et al., 2017; Overby, Belardi, \& Schreiber, 2020).
+Advancements in the field of wearable technologies, such as LENA\textsuperscript{TM} recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally (Oller et al., 2010; Rankine et al., 2017; VanDam \& Yoshinaga-Itano, 2019).
 Wearable recorders allow data collection to happen in the child's natural environment, and at a large scale, which may be particularly helpful for children whose speech is not easily elicited.
@@
 There is a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this work focuses on babbling and thus typically on infants under one year of age.
-A critical milestone involves the increasingly common production of Canonical Syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech (Oller, Eilers, Neal, \& Cobo-Lewis, 1998).
+A critical milestone involves the increasingly common production of canonical syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech (Oller, Eilers, Neal, \& Cobo-Lewis, 1998).
 Given its adult-like consonant-vowel or vowel-consonant structure, canonical babble is considered to be a starting point on the path to recognizable speech.
@@
 Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development (Oller, 2000).
-In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization count in a sample of children with an Autism Spectrum Disorder (McDaniel et al., 2020).
+In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization count in a sample of children diagnosed with an autism spectrum disorder (McDaniel et al., 2020).
 
@@
 For the English and Tsimane' corpora, vocalizations were automatically identified using LENA\textsuperscript{TM}, whereas the other two were extracted through manual segmentation.
-Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received 3 classifications into crying, laughing, canonical, non-canonical, or \enquote{Junk} (with the latter tag used for clips that did not contain clear child voice).
+Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received three classifications into Crying, Laughing, Canonical, Non-Canonical, or Junk (with the latter tag used for clips that did not contain a child voice).
 They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgement of canonical or non-canonical.
-They found that this canonical ratio increased with infant age in an analysis collapsing across corpora, which provides a first proof of principle that citizen scientists' annotations could be an appropriate solution to the problem of annotating child vocalizations from wearables.
-The authors also studied how their canonical ratio related to age within each corpus that had a sufficient number of children varying in age.
+They found that this Canonical Proportion increased with infant age in an analysis collapsing across corpora, which provides a first proof of principle that citizen scientists' annotations could be an appropriate solution to the problem of annotating child vocalizations from wearables.
+The authors also studied how their Canonical Proportion related to age within each corpus that had a sufficient number of children varying in age.
 They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the \enquote{Junk} category was very high.
-Additionally, they found that the correlation between canonical ratio and age was much weaker in two corpora where vocalizations were segmented with LENA\textsuperscript{TM} than through manual segmentation.
+Additionally, they found that the correlation between Canonical Proportion and age was much weaker in two corpora where vocalizations were segmented with LENA\textsuperscript{TM} than through manual segmentation.
 In sum, these analyses within corpora reveal that data may not be uniformly useful, with potential variation across corpora, children, and methods.
@@
 
-When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (canonical+non-canonical)/(all vocalizations); and the Canonical Proportion (canonical)/(all linguistic vocalizations).
-Previous work suggests that children with Angelman syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the canonical ratio was expected to increase with age (Hamrick, Seidl, \& Kelleher, n.d.).
+When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk vocalizations); and the Canonical Proportion (Canonical)/(all linguistic vocalizations).
+Previous work suggests that children with Angelman syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the Canonical Proportion was expected to increase with age (Hamrick et al., n.d.; Hamrick \& Tonnsen, n.d.).
 Although we could not rely on previous work to make predictions regarding the Linguistic Proportion, we reasoned that we should observe an increase in the Linguistic Proportion for both populations.
@@
 Before annotation, coders completed an ethical conduct of research course and received HIPAA compliance training.
-After this, they complete a brief training on vocal maturity that involves 1) reading relevant literature, 2) reviewing examples of various annotation categories and 3) completing a brief quiz to test their annotation accuracy.
-Coders then proceed to the classification, where they make decisions on whether a segment is a cry, laugh, non-canonical syllable(s), canonical syllable(s), or a word.
-Coders are instructed to classify each segment based on the highest level of vocal maturity it contains (e.g., a segment with both canonical and non-canonical syllables would be classified as canonical; a segment with both laughing and non-canonical syllables would be classified as non-canonical).
+After this, they completed a brief training on vocal maturity that involved 1) reading relevant literature, 2) reviewing examples of various annotation categories and 3) submitting answers to a brief quiz to test their annotation accuracy.
+Coders then proceeded to the classification, where they made decisions on whether a segment is a cry, laugh, non-canonical syllable(s), canonical syllable(s), or a word.
+Coders were instructed to classify each segment based on the highest level of vocal maturity it contains (e.g., a segment with both canonical and non-canonical syllables would be classified as canonical; a segment with both laughing and non-canonical syllables would be classified as non-canonical).
 For canonical, non-canonical, and word segments, coders indicated how many syllables of each type of vocal maturity they hear (e.g., \enquote{ah ma ba} would be 1 non-canonical and 2 canonical syllables).
@@
 
-This process resulted in high levels of agreement: 97.22\% of the lab-annotated segments had either 2 or 3 coders agree.
+This process resulted in high levels of inter-rater agreement: Two or more of the three total coders who annotated that segment agreed on 97.22\% of the lab-annotated segments.
 
@@
 Fortunately, Seidl et al.
-(2019) determined that when segments are divided into shorter chunks (400-600ms), human annotators with little training can code our categories of interest (canonical, laughing, crying, etc.) with a classification quality comparable to the one carried out on full segments.
+(2019) determined that when segments are divided into shorter chunks (400-600ms), human annotators with little training can code our categories of interest (Canonical, Laughing, Crying, etc.) with a classification quality comparable to the one carried out on full segments.
 We therefore used very short chunks (500ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.
@@
 
-Citizen scientists could access the project by logging in, or as anonymous participants.
+Citizen scientists could access our project, \enquote{Maturity of Baby Sounds}, by logging in, or as anonymous participants.
 Before starting the annotation, participants were given a quick tutorial which walked them through the steps of the annotation workflow (i.e., how to play a sound, how to make a selection; see Figure 1).
 In this tutorial, we included one audio example for each category in order to make the classification task as smooth as possible.
-Further clarifications on what constitutes Canonical and Non-canonical utterances, as well as five audio examples for each of the categories, were available through the Field Guide, which users could access by clicking on the right side of the screen.
+Further clarifications on what constitutes canonical and non-canonical sequences, as well as five audio examples for each of the categories, were available through the Field Guide, which users could access by clicking on the right side of the screen.
 The Field Guide could be consulted at any point of the classification without interrupting the task.
 
-Users were asked to assign each 500ms clip to one out of five possible categories: (1) canonical, (2) non-canonical, (3) crying, (4) laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
+Users were asked to assign each 500ms clip to one out of five possible categories: (1) Canonical, (2) Non-Canonical, (3) Crying, (4) Laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
 Notice that the latter is narrower than the \enquote{Don't mark} category of the lab annotation routine, which was used any time there was noise or overlap that could affect acoustic analyses.
 
-Following recommendations from the Zooniverse board, we collected 5 judgments per clip, rather than 3 as used in the laboratory.
-Majority agreement is thus achieved when 3 out of 5 judgements agree (rather than 2 out of 3).
+Following recommendations from the Zooniverse board, we collected five judgments per clip, rather than three as used in the laboratory.
+Majority agreement is thus achieved when three out of five judgements agree (rather than two out of three).
 
@@
 \includegraphics{zooniverse-pufig.pdf}
-\caption{\label{fig:fig-zoo}Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needs to make; as well as the field guide, which remains accessible on the right of the screen throughout the session, and which allows users to revise examples of the different categories.}
+\caption{\label{fig:fig-zoo}Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needed to make; as well as the Field Guide, which remained accessible on the right of the screen throughout the session, and which allowed users to revise examples of the different categories.}
 \end{figure}
@@
 \item
-  if one canonical judgement was present, then the segment was labelled as Canonical,
+  if one Canonical judgement was present, then the segment was labelled as Canonical,
 \item
@@
 
-This is essentially the same thought process expert annotators followed when labeling segment-level audio, as each segment can represent multiple classification subtypes).
+This is essentially the same thought process expert annotators followed when labeling segment-level audio, as each segment can represent multiple classification subtypes.
 LENA\textsuperscript{TM} segments are not equivalent to child vocalizations: A segment may contain 1 or more vocalizations, which may be of the same type or not.
@@
 In this section, we provide descriptive analyses of our dataset.
-According to lab annotators, 15\% of segments were canonical, 56\% non-canonical, 2\% laughing, and 5\% crying, with the remaining 22\% being categorized as \enquote{Don't mark}.
-Zooniverse data revealed a similar distribution: 15\% canonical, 60\% non-canonical, 4\% laughing, 9\% crying, 13\% Junk.
-Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (\enquote{Canonical}+\enquote{Non-Canonical})/\enquote{All vocalizations} (i.e., we remove, Junk), and ii) Canonical Proportion = \enquote{Canonical}/(\enquote{Canonical}+\enquote{Non-Canonical}) (i.e., we remove Junk + non-linguistic).
+According to lab annotators, 14\% of segments were Canonical, 60\% Non-Canonical, 1\% Laughing, and 4\% Crying, with the remaining 21\% being categorized as \enquote{Don't mark}.
+Zooniverse data revealed a similar distribution: 15\% Canonical, 60\% Non-Canonical, 3\% Laughing, 7\% Crying, 15\% Junk.
+Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical+Non-Canonical)/All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical/(Canonical+Non-Canonical) (i.e., we remove Junk, Crying, and Laughing).
 See Figure 2 for results.
@@
 \includegraphics{paper_files/figure-latex/fig-corage-1.pdf}
-\caption{\label{fig:fig-corage}Correlations between child-level descriptors and age as a function of metric (linguistic ratio in the top row, canonical ratio in the bottom row), annotation method, and child group (red=low-risk, and black Angelman syndrome).}
+\caption{\label{fig:fig-corage}Correlations between child-level descriptors and age as a function of metric (Linguistic Proportion in the top row, Canonical Proportion in the bottom row), annotation method, and child group (red=low-risk, and black= Angelman syndrome).}
 \end{figure}
@@
 Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups.
-There was a near-zero relationship among the older children diagnosed with Angelman syndrome r(8) = 0.02, CI {[}-0.62,0.64{]}, p=0.97{]}; and a significant association among younger low-risk control children r(8) = 0.74, CI {[}0.21,0.94{]}, p=0.01{]}.
-The Canonical Proportion exhibited non-significant developmental decreases among older children diagnosed with Angelman syndrome r(8) = -0.56, CI {[}-0.88,0.11{]}, p=0.09{]}; and marginal developmental increases among low-risk control r(8) = 0.61, CI {[}-0.04,0.90{]}, p=0.06{]}.
+There was a near-zero relationship among the children diagnosed with Angelman syndrome r(8) = -0.09, CI {[}-0.68,0.57{]}, p=0.80{]}; and a significant association among low-risk control children r(8) = 0.70, CI {[}0.13,0.92{]}, p=0.02{]}.
+The Canonical Proportion exhibited non-significant developmental decreases among children diagnosed with Angelman syndrome r(8) = -0.39, CI {[}-0.82,0.31{]}, p=0.26{]}; and marginal developmental increases among low-risk control r(8) = 0.52, CI {[}-0.17,0.86{]}, p=0.13{]}.
 
-Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(8) = -0.10, CI {[}-0.68,0.57{]}, p=0.79{]}; whereas younger low-risk control children showed a significant increase with age r(8) = 0.68, CI {[}0.08,0.92{]}, p=0.03{]}.
-Similarly, there were non-significant developmental decreases in the Canonical among children with Angelman syndrome r(8) = -0.42, CI {[}-0.83,0.29{]}, p=0.23{]}; and marginal developmental increases among low-risk control children, r(8) = 0.60, CI {[}-0.04,0.89{]}, p=0.06{]}.
+Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(8) = -0.17, CI {[}-0.72,0.51{]}, p=0.63{]}; whereas low-risk control children showed a significant increase with age r(8) = 0.63, CI {[}0.00,0.90{]}, p=0.05{]}.
+Similarly, there were non-significant developmental decreases in the Canonical Proportion among children with Angelman syndrome r(8) = -0.41, CI {[}-0.82,0.30{]}, p=0.24{]}; and marginal developmental increases among low-risk control children, r(8) = 0.51, CI {[}-0.17,0.86{]}, p=0.13{]}.
 
@@
 \caption{\label{fig:fig-prec-rec}Number of segments as a function of their majority label from laboratory annotations (rows) versus Zooniverse annotations (columns).
-Percentages (and shading) indicate Precision (left) and Recall (right).
+Percentages (and shading) indicate precision (left) and recall (right).
 For precision, the percentage indicates what percentage of the segments that have a given majority label according to Zooniverse data are attributed a given label in Lab data (i.e., columns add up to 100\%).
@@
 These visualizations suggest that performance is moderate to good, which was confirmed via statistical analyses.
-The overall (weighted) accuracy is 65\%, CI = {[}64,66{]}, kappa is 0.43, and the Gwet's AC1 coefficient is 0.59, CI = {[}0.58,0.60{]}.
+The overall (weighted) accuracy is 73\%, CI = {[}72,74{]}, kappa is 0.53, and the Gwet's AC1 coefficient is 0.68, CI = {[}0.67,0.69{]}.
 
@@
 
-\includegraphics{paper_files/figure-latex/corlab-zoo-1.pdf}
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/fig-corlab-zoo-1.pdf}
+\caption{\label{fig:fig-corlab-zoo}Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of child group (red=low-risk, and black=Angelman syndrome).}
+\end{figure}
 
 Although the classification at the clip level is only moderately accurate, what we are ultimately interested in is whether citizen scientists' classifications are able to provide a reliable snapshot of childrens' individual development.
-Looking at all 20 children together, we found a strong positive correlation r(18) = 0.81, CI {[}0.57,0.92{]}, p=0{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman syndrome r(8) = 0.88, CI {[}0.55,0.97{]}, p=0.00{]}; low-risk control r(8) = 0.82, CI {[}0.40,0.96{]}, p=0.00{]}.
+We illustrate this association in Figure 3.
+Looking at all 20 children together, we found a strong positive correlation r(18) = 0.79, CI {[}0.54,0.92{]}, p=0{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman syndrome r(8) = 0.88, CI {[}0.56,0.97{]}, p=0.00{]}; low-risk control r(8) = 0.83, CI {[}0.42,0.96{]}, p=0.00{]}.
 
-Similarly, a strong positive correlation is found in the Canonical Proportion r(18) = 0.94, CI {[}0.84,0.98{]}, p=0{]}.
-When we split by participant group, correlations remain high although we do note they are somewhat smaller for the older children with Angelman syndrome: r(8) = 0.86, CI {[}0.50,0.97{]}, p=0.00{]}; than the low-risk control children r(8) = 0.96, CI {[}0.83,0.99{]}, p=0.
+Similarly, a strong positive correlation is found in the Canonical Proportion r(18) = 0.92, CI {[}0.81,0.97{]}, p=0{]}.
+When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman syndrome: r(8) = 0.68, CI {[}0.09,0.92{]}, p=0.03{]}; than the low-risk control children r(8) = 0.98, CI {[}0.93,1.00{]}, p=0.
 
@@
 
-\includegraphics{paper_files/figure-latex/cor-Junk-1.pdf}
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/fig-cor-Junk-1.pdf}
+\caption{\label{fig:fig-cor-Junk}Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of whether children's Junk proportion was higher or lower than the median for all children (blue=lower, and green=higher).}
+\end{figure}
 
 We investigated this hypothesis by calculating the proportion of their data labeled as \enquote{Junk} for each individual child.
-There was no significant difference in the proportions of their data labeled as \enquote{Junk} for older children with Angelman syndrome (M = 0.13) compared to the younger low risk children (M = 0.14): Welch's t(13.37) = -0.52, p = 0.61.
+There was no significant difference in the proportions of their data labeled as \enquote{Junk} for children with Angelman syndrome (M = 0.14) compared to the low-risk group (M = 0.16): Welch's t(13.35) = -0.58, p = 0.57.
 We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as \enquote{Junk}.
-Results were similar across these two post-hoc subgroups.
-For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = 0.87, CI {[}0.53,0.97{]}, p=0.00{]}; and for the higher Junk group it was r(8) = 0.87, CI {[}0.54,0.97{]}, p=0.00.
-For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = 0.96, CI {[}0.82,0.99{]}, p=0{]}; and for the higher Junk group it was r(8) = 0.93, CI {[}0.73,0.98{]}, p=0.
+Results were similar across these two post-hoc subgroups (see Figure 4).
+For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = 0.86, CI {[}0.51,0.97{]}, p=0.00{]}; and for the higher Junk group it was r(8) = 0.90, CI {[}0.61,0.98{]}, p=0.
+For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = 0.96, CI {[}0.82,0.99{]}, p=0{]}; and for the higher Junk group it was r(8) = 0.90, CI {[}0.61,0.98{]}, p=0.
 Thus, it does not seem that a higher proportion of \enquote{Junk} judgments is an index of low quality data.
@@
 
-As for correlations between Zooniverse and laboratory-derived metrics, we observe very similar levels of correlation: Linguistic Proportion overall r(18) = 0.84, CI {[}0.64,0.94{]}, p=0{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman syndrome r(8) = 0.84, CI {[}0.46,0.96{]}, p=0.00{]}; low-risk control r(8) = 0.84, CI {[}0.45,0.96{]}, p=0.00{]}.
-As for Canonical Proportion overall r(18) = 0.96, CI {[}0.91,0.98{]}, p=0; Angelman syndrome: r(8) = 0.87, CI {[}0.52,0.97{]}, p=0.00{]}; low-risk control infants r(8) = 0.98, CI {[}0.89,0.99{]}, p=0.
+As for correlations between Zooniverse and laboratory-derived metrics, we observe very similar levels of correlation: Linguistic Proportion overall r(18) = 0.80, CI {[}0.55,0.92{]}, p=0{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman syndrome r(8) = 0.84, CI {[}0.46,0.96{]}, p=0.00{]}; low-risk control r(8) = 0.79, CI {[}0.33,0.95{]}, p=0.01{]}.
+As for Canonical Proportion overall r(18) = 0.95, CI {[}0.87,0.98{]}, p=0; Angelman syndrome: r(8) = 0.72, CI {[}0.17,0.93{]}, p=0.02{]}; low-risk control infants r(8) = 0.97, CI {[}0.88,0.99{]}, p=0.
 
@@
 Next, we looked at whether having more segments from each child may lead to more reliable metrics.
-We observed a non-significant trend {[}t(17.92)=527.40 and 631.90{]} for lower number of segments in the Angelman syndrome group (mean 527.40, range 326-924) than among low-risk control infants (mean 631.90, range 351-1074), likely because of the lower volubility of the former children.
+We observed a non-significant trend {[}t(17.88)=348.70 and 375.50{]} for lower number of segments in the Angelman syndrome group (mean 348.70, range 163-599) than among low-risk control infants (mean 375.50, range 224-627), likely because of the lower volubility of the former children.
 Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications (Cychosz et al., 2019).
@@
 We repeated this process 50 times, to assess the extent to which the association between laboratory- and Zooniverse-derived metrics varied in each random sample.
-The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table XX), particularly for the low-risk younger infants, and for Linguistic Proportion.
-This may indicate that, particularly in some groups of infants, 100 segments may not be sufficient to capture the child's linguistic and Canonical Proportions.
+The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table 2), particularly for the low-risk younger infants, and for Linguistic Proportion.
+This may indicate that, particularly in some groups of infants, 100 segments may not be sufficient to capture a stable estimate of the child's Linguistic and Canonical Proportions.
 We also observed quite a bit of variance across the 50 runs.
-For Linguistic Proportion, the range of correlations found for all infants was {[}0.64,0.88{]}; for the Angelman syndrome group {[}0.76,0.94{]}; for the low-risk group {[}0.45,0.95{]}.
-For Canonical Proportion, the range of correlations found for all infants was {[}0.82,0.96{]}; for the Angelman syndrome group {[}0.40,0.97{]}; for the low-risk group {[}0.75,0.97{]}.
+For Linguistic Proportion, the range of correlations found for all infants was {[}0.51,0.91{]}; for the Angelman syndrome group {[}0.60,0.96{]}; for the low-risk group {[}0.05,0.96{]}.
+For Canonical Proportion, the range of correlations found for all infants was {[}0.82,0.96{]}; for the Angelman syndrome group {[}0.59,0.95{]}; for the low-risk group {[}0.77,0.98{]}.
 
@@
 
-\caption{\label{tab:tab-cors}Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways: All seg indicates that Zooniverse annotations at the clip level were first combined at the segment level; Chunks that they were analyzed directly at the clip level.
+\caption{\label{tab:tab-cors}Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways.
+First, All seg indicates that Zooniverse annotations at the clip level were first combined at the segment level.
+Second, Chunks indicates that they were analyzed directly at the clip level.
 Both of these are based on all the data.
-In contrast, rows labeled 100 seg indicate that laboratory- and Zooniverse-derived metrics were based on only 100 segments (median over 50 runs in which 100 segments were randomly selected from each child).}
+Third, rows labeled 100 seg indicate that laboratory- and Zooniverse-derived metrics were based on only 100 segments (median over 50 runs in which 100 segments were randomly selected from each child).}
 \centering
@@
 \hline
-All seg AS & 0.877 & 0.858\\
+All seg AS & 0.878 & 0.681\\
 \hline
-All seg LR & 0.823 & 0.960\\
+Chunks AS & 0.845 & 0.724\\
 \hline
-All seg all & 0.811 & 0.937\\
+100 seg AS & 0.855 & 0.809\\
 \hline
-Chunks AS & 0.845 & 0.866\\
+All seg LR & 0.828 & 0.984\\
 \hline
-Chunks LR & 0.841 & 0.975\\
+Chunks LR & 0.794 & 0.972\\
 \hline
-Chunks all & 0.845 & 0.963\\
+100 seg LR & 0.609 & 0.907\\
 \hline
-100 seg AS & 0.857 & 0.807\\
+All seg all & 0.794 & 0.920\\
 \hline
-100 seg LR & 0.806 & 0.892\\
+Chunks all & 0.798 & 0.949\\
 \hline
-100 seg all & 0.783 & 0.903\\
+100 seg all & 0.761 & 0.889\\
 \hline
@@
 In the present study we assessed the extent to which child vocalizations from LENA\textsuperscript{TM} daylong recordings can be accurately described based on classifications collected using the citizen science platform Zooniverse.
-Our recordings came from children diagnosed with a language disorder, as well as low-risk controls, and included a range of different ages.
+Our recordings came from children diagnosed with a genetic syndrome associated with severe language disorder, as well as low-risk controls, and included a range of different ages.
 Excerpts of these recordings had previously been annotated by trained experts in the lab, and served as the \enquote{gold standard}.
@@
 Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the \enquote{Junk} category less) than laboratory annotators, who have access to the whole vocalization.
-It is also plausible, on the other hand, that lab annotators are more \enquote{picky} with the data, and tend to reject clips (tag them as \enquote{Don't mark}).
-Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use \enquote{Don't mark} if there was any noise or verlapping speech that would potentially affect acoustic analyses.
+This is consistent with the fact that our lab definition of \enquote{Junk} was more stringent than the Zooniverse definition.
+Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use \enquote{Don't mark} if there was \emph{any} noise or verlapping speech that would potentially affect acoustic analyses.
 
@@
 
-As for laughing, the recall is quite good but precision is low, with confusion involving Junk and Non-Canonical to similar extents.
+As for Laughing, the recall is quite good but precision is low, with confusion involving \enquote{Junk} and Non-Canonical to similar extents.
 We are tentative in our interpretation of this result because the category was very sparsely represented, with barely 200 segments classified as Laughing in the whole data set.
@@
 Our pipeline to upload data on Zooniverse is scalable and quickly adapted to new datasets, including data on infants of different ages, learning different languages, or living in environments with different acoustic properties (e.g., rural versus urban environments).
-This means large existing datasets of daylong recordings can be reliably annotated in an ecological way using citizen science (ref on existing datasets).
+This means large existing datasets of daylong recordings, notably those hosted on Homebank (Vandam et al., 2016), can be reliably annotated in an ecological way using citizen science.
 
@@
 Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion.
-We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant for other disorders that are related to emotional disorders (a review in Halpern \& Coelho, 2016).
+We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant for other disorders that are related to emotional disorders (see a review in Halpern \& Coelho, 2016).
 Additionally, while Canonical Proportion has been studied in much previous work, we believe further research is necessary using larger scales (saliently increasing the sheer number of children studied, to have sufficient power to detect group effects) and with broader coverage (notably including children growing up in a broad range of languages and cultures, as in Cychosz et al., 2019), to make sure that assumptions regarding cross-linguistic and maturational universality are amply justified and not merely a result of having low power to detect differences, or simply sampling from similar populations.
@@
 Baumer, B., \& Udwin, D. (2015). R markdown. \emph{Wiley Interdisciplinary Reviews: Computational Statistics}, \emph{7}(3), 167--177.
+
+\leavevmode\hypertarget{ref-belardi2017retrospective}{}%
+Belardi, K., Watson, L.
+R., Faldowski, R.
+A., Hazlett, H., Crais, E., Baranek, G.
+T., \ldots{} Oller, D.
+K.
+(2017).
+A retrospective video analysis of canonical babbling and volubility in infants with fragile x syndrome at 9--12 months of age.
+\emph{Journal of Autism and Developmental Disorders}, \emph{47}(4), 1193--1206.
 
@@
 (n.d.).
-Capturing variability in early language skills among children with angelman syndrome: A novel semi-automated approach.
+Capturing variability in early language skills among children with Angelman syndrome: A novel semi-automated approach.
 \emph{Presented at the Angelman Biomarkers and Outcome Measures Alliance Scientific Meeting}.
+
+\leavevmode\hypertarget{ref-hamrick2019measurement}{}%
+Hamrick, L., \& Tonnsen, B.
+(n.d.).
+Measurement and characterization of early social communication in children with Angelman syndrome.
+In A.
+Sadhwani \& K.
+Okoniewski (Eds.), \emph{Advances in characterizing developmental trajectories in angelman syndrome across the lifespan, gatlinburg conference}.
+San Antonio, TX.
 
@@
 \leavevmode\hypertarget{ref-oller2000emergence}{}%
-Oller, D. (2000). The emergence of the capacity for speech. Erlbaum.{[}aJLL, DKO{]}(2004) Underpinnings for a theory of communicative~\ldots{}.
+Oller, D. (2000). The emergence of the capacity for speech. Erlbaum.
 
@@
 \emph{American Journal on Mental Retardation}, \emph{103}(3), 249--263.
+
+\leavevmode\hypertarget{ref-oller2010automated}{}%
+Oller, D.
+K., Niyogi, P., Gray, S., Richards, J.
+A., Gilkerson, J., Xu, D., \ldots{} Warren, S.
+F.
+(2010).
+Automated vocal analysis of naturalistic recordings from children with autism, language delay, and typical development.
+\emph{Proceedings of the National Academy of Sciences}, \emph{107}(30), 13354--13359.
+
+\leavevmode\hypertarget{ref-overby2020retrospective}{}%
+Overby, M., Belardi, K., \& Schreiber, J.
+(2020).
+A retrospective video analysis of canonical babbling and volubility in infants later diagnosed with childhood apraxia of speech.
+\emph{Clinical Linguistics \& Phonetics}, \emph{34}(7), 634--651.
+
+\leavevmode\hypertarget{ref-rankine2017language}{}%
+Rankine, J., Li, E., Lurie, S., Rieger, H., Fourie, E., Siper, P.
+M., \ldots{} Kolevzon, A.
+(2017).
+Language environment analysis (lena) in phelan-mcdermid syndrome: Validity and suggestions for use in minimally verbal children with autism spectrum disorder.
+\emph{Journal of Autism and Developmental Disorders}, \emph{47}(6), 1605--1617.
 
@@
 
+\leavevmode\hypertarget{ref-vandam2016homebank}{}%
+Vandam, M., Warlaumont, A.
+S., Bergelson, E., Cristia, A., Soderstrom, M., De Palma, P., \& Macwhinney, B.
+(2016).
+Homebank: An online repository of daylong child-centered audio recordings.
+\emph{Seminars in Speech and Language}, \emph{37}(2), 128--142.
+\url{https://doi.org/10.1055/s-0036-1580745}
+
+\leavevmode\hypertarget{ref-vandam2019use}{}%
+VanDam, M., \& Yoshinaga-Itano, C. (2019). Use of the lena autism screen with children who are deaf or hard of hearing. \emph{Medicina}, \emph{55}(8), 495.
+
 \leavevmode\hypertarget{ref-wang2020meta}{}%

### 2020-11-10 data-analyses/code/zoon_chunk_level.Rmd
updating scripts - cannot reproduce

---
title: "Chunk_level_jslhr"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
library(tidyverse)
library(dplyr)
library(ggplot2)
library(caret)
library(scales)
dodiv=function(x) x/sum(x, na.rm=T)
```

```{r cars}
zoon_chunks=read.csv("/Users/chiarasemenzin/Desktop/chunk_majority_agreement.csv",sep=",")
demo_data=read.csv("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Data/demo-data.tsv",sep="\t")
demo_data$Child_ID<-demo_data$ChildID
data_all<-left_join(zoon_chunks,demo_data,by="Child_ID")
ratios_seg<-read.csv("/Users/chiarasemenzin/ratios_lb_z.csv")
```

```{r prepare data by child}
#get the ns by child, then calculate the linguistic ratio & canonical ratio, separately for zooniverse & lab
ztab=table(data_all$Child_ID,data_all$Answer)
z_lr=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
z_cr=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
ratios=as.data.frame(cbind(rownames(ztab),z_lr,z_cr))
colnames(ratios)[1]<-"Child_ID"
merge(ratios,demo_data,by="Child_ID")->ratios
ratios$z_lr<-as.numeric(ratios$z_lr)
ratios$z_cr<-as.numeric(ratios$z_cr)
```

```{r corage}
prettynames=c("Linguistic Ratio (Zooniverse)","Canonical Ratio (Zooniverse)",
             "Linguistic Ratio (Lab)","Canonical Ratio (Lab)" )
names(prettynames)<-c("z_lr","z_cr")
#!!! this is working the opposite than it should! but note that to get angsynd kids to come out in red, I have to do it like this -- it doesn't make sense...
# Now it works alright
mycols=c("black","red")
names(mycols)<-c("Low-RiskControl","AngelmanSyndrome")
for(thisvar in c("z_lr","z_cr")) {
  myr=round(cor.test(ratios[,thisvar],ratios$Age)$estimate,3)
  print(cor.test(ratios[,thisvar],ratios$Age))
  plot(ratios[,thisvar]~ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr),
       col=mycols[ratios$Diagnosis])
  abline(lm(ratios[,thisvar]~ratios$Age))
  }
```
``````{r corage}
prettynames=c("Linguistic Ratio (Zooniverse)","Canonical Ratio (Zooniverse)")
names(prettynames)<-c("z_lr","z_cr")
AS_ratios<-subset(ratios, Diagnosis=="AngelmanSyndrome")
TD_ratios<-subset(ratios, Diagnosis=="Low-RiskControl")

mycols=c("black","red")
names(mycols)<-c("Low-RiskControl","AngelmanSyndrome")
for(thisvar in c("z_lr","z_cr")) {
  myr=round(cor.test(AS_ratios[,thisvar],AS_ratios$Age)$estimate,3)
  plot(AS_ratios[,thisvar]~AS_ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr))
  abline(lm(AS_ratios[,thisvar]~AS_ratios$Age))
  print(thisvar)
  print(cor.test(AS_ratios[,thisvar],AS_ratios$Age))
}

for(thisvar in c("z_lr","z_cr")) {
  myr=round(cor.test(TD_ratios[,thisvar],TD_ratios$Age)$estimate,3)
  plot(TD_ratios[,thisvar]~TD_ratios$Age, pch=20,xlab="Age (months)",ylab=prettynames[thisvar],main=paste0("r=",myr))
  abline(lm(TD_ratios[,thisvar]~TD_ratios$Age))
  print(thisvar)
  print(cor.test(TD_ratios[,thisvar],TD_ratios$Age))
  }
```

```{r corlab zoo}
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/ling_rat_z_vs_l_chunk.pdf",height=5,width=5)

ratios<-left_join(ratios,ratios_seg[c("l_lr","l_cr","ChildID")],by="ChildID")

lims=range(c(ratios[,"z_lr"],ratios[,"l_lr"]))
  myr=round(cor.test(ratios[,"z_lr"],ratios[,"l_lr"])$estimate,3)
  plot(ratios[,"z_lr"]~ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[ratios$Diagnosis])
  abline(lm(ratios[,"z_lr"]~ratios[,"l_lr"]))
  lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
print(cor.test(ratios[,"z_lr"],ratios[,"l_lr"]))

  #CR
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/can_rat_z_vs_l_chunk.pdf",height=5,width=5)
lims=range(c(ratios[,"z_cr"],ratios[,"l_cr"]))
    myr=round(cor.test(ratios[,"z_cr"],ratios[,"l_cr"])$estimate,3)
  plot(ratios[,"z_cr"]~ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[ratios$Diagnosis])
  abline(lm(ratios[,"z_cr"]~ratios[,"l_cr"]),col="darkgray")
    lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
print(cor.test(ratios[,"z_cr"],ratios[,"l_cr"]))
```

``` {r subsetdata}
prettynames=c("AS Linguistic Ratio (Zooniverse)","AS Canonical Ratio (Zooniverse)",
             "AS Linguistic Ratio (Lab)","AS Canonical Ratio (Lab)" )
names(prettynames)<-c("z_lr","z_cr","l_lr","l_cr")

AS_ratios<-subset(ratios, Diagnosis=="AngelmanSyndrome")
TD_ratios<-subset(ratios, Diagnosis=="Low-RiskControl")
```

``` {r cor pred}
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/td_ling_rat_z_vs_l_chunk.pdf",height=5,width=5)
lims=range(c(TD_ratios[,"z_lr"],TD_ratios[,"l_lr"]))
  myr=round(cor.test(TD_ratios[,"z_lr"],TD_ratios[,"l_lr"])$estimate,3)
  plot(TD_ratios[,"z_lr"]~TD_ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[TD_ratios$Diagnosis])
  abline(lm(TD_ratios[,"z_lr"]~TD_ratios[,"l_lr"]))
  lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
print(cor.test(TD_ratios[,"z_lr"],TD_ratios[,"l_lr"]))

  #CR
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/td_can_rat_z_vs_l_final.pdf",height=5,width=5)
lims=range(c(TD_ratios[,"z_cr"],TD_ratios[,"l_cr"]))
    myr=round(cor.test(TD_ratios[,"z_cr"],TD_ratios[,"l_cr"])$estimate,3)
  plot(TD_ratios[,"z_cr"]~TD_ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[TD_ratios$Diagnosis])
  abline(lm(TD_ratios[,"z_cr"]~TD_ratios[,"l_cr"]),col="darkgray")
    lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
print(cor.test(TD_ratios[,"z_cr"],TD_ratios[,"l_cr"]))
```
```{r corlab-zoo}
#Ling ratio
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/AS_ling_rat_z_vs_l_chunks.pdf",height=5,width=5)
lims=range(c(AS_ratios[,"z_lr"],AS_ratios[,"l_lr"]))
  myr=round(cor.test(AS_ratios[,"z_lr"],AS_ratios[,"l_lr"])$estimate,3)
  plot(AS_ratios[,"z_lr"]~AS_ratios[,"l_lr"], pch=20,xlab=prettynames["l_lr"],ylab=prettynames["z_lr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[AS_ratios$Diagnosis])
  abline(lm(AS_ratios[,"z_lr"]~AS_ratios[,"l_lr"]))
  lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
print(cor.test(AS_ratios[,"z_lr"],AS_ratios[,"l_lr"]))
  #CR
pdf("/Users/chiarasemenzin/Documents/GitHub/zoo-babble-validation/Results/AS_can_rat_z_vs_l_chunks.pdf",height=5,width=5)
lims=range(c(AS_ratios[,"z_cr"],AS_ratios[,"l_cr"]))
    myr=round(cor.test(AS_ratios[,"z_cr"],AS_ratios[,"l_cr"])$estimate,3)
  plot(AS_ratios[,"z_cr"]~AS_ratios[,"l_cr"], pch=20,xlab=prettynames["l_cr"],ylab=prettynames["z_cr"],main=paste0("r=",myr),
       xlim=lims,ylim=lims,
       col=mycols[AS_ratios$Diagnosis])
  abline(lm(AS_ratios[,"z_cr"]~AS_ratios[,"l_cr"]),col="darkgray")
    lines(c(0,1),c(0,1),lty=2,col="darkgray")
dev.off()
print(cor.test(AS_ratios[,"z_cr"],AS_ratios[,"l_cr"]))
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-11-10 Scripts/chunk_to_segment.py

@@ -1,87 +1,0 @@
-import numpy as np
-from collections import defaultdict
-import json
-
-# ----------------------------------------------------------------------------------------
-
-#This script creates a dictionary file in json format with mapping from 500ms ->chunks<- to Lab annotation ->segments<-
-
-
-result_final="/Users/chiarasemenzin/Desktop/LSCP/SLT/result_final_lisa.csv"  #file with segment info and lab classifications
-result_zoon="/Users/chiarasemenzin/Desktop/LSCP/SLT/merged_metadata.csv"  #metadata for zooniverse clips
-
-
-# Segment to timestamp
-def seg_to_ts(result_final):
-    seg2ts={}
-    with open(result_final) as final:
-        for e,line in enumerate(final):
-            line=line.split(",")
-            if e > 0:
-                seg2ts[line[0]]=(float(line[4]),float(line[5]),line[1])
-    return seg2ts
-
-# Timestamp to segment
-def chunk_to_ts(result_zoon):
-    chunk2ts={}
-    with open(result_zoon) as zoon:
-        for e,line in enumerate(zoon):
-            line=line.split(",")
-            if e > 0:
-                #chunk2ts[line[0]]=[float(i) for i in line[4].split("_")]
-                chunk2ts[line[1]]=[float(line[4]),float(line[5])] #get timestamps
-                chunk2ts[line[1]].append(line[2]) #get name of chunk
-    return chunk2ts
-
-# Calculate average duration of segment
-def avg_duration(result_final):
-    seg_ts=[]
-    with open(result_final) as final:
-        for e,line in enumerate(final):
-            line=line.split(",")
-            if e > 0:
-                seg_ts.append(float(line[5])-float(line[4]))
-    avg_dur=sum(seg_ts)/len(seg_ts)
-    return avg_duration
-
-def check(segments,chunks):
-    """" Check which chunks belong to which segment, by file"""
-
-    segments_final = defaultdict(list)
-
-    for name,chunk in chunks.items(): 
-        diffs_onset=[]
-        diffs_offeset=[]
-        candidate_seg=dict([i for i in segments.items() if i[1][2]==chunk[2]])
-        for k,seg in candidate_seg.items():
-            if seg[2] == chunk[2]:
-                print("Looking at ",k,seg[2])
-                diffs_onset.append([abs(chunk[0]-seg[0]),k])
-                diffs_offeset.append([abs(chunk[1]-seg[1]),k])
-        min_on=min([i[0] for i in diffs_onset])
[…]

### 2020-11-10 Scripts/merge_datasets_zoon.py

@@ -1,117 +1,0 @@
-import pandas as pd
-from collections import OrderedDict
-import csv
-
-print("Started.")
-
-# define paths to your existing zooniverse classifications and metadata locations
-# These are the files you get as output from the Jupyter Notebook convert_zooniverse.ipynb
-
-classifs=pd.read_csv("/Users/chiarasemenzin/Desktop/LSCP/SLT/zooniverse_data_all_final.csv")
-metadata=pd.read_csv("/Users/chiarasemenzin/Desktop/zoon_metadata/Alex_metadata/MetaData_LAAC_20200504_ac.csv",sep='\t')
-filename_links=pd.read_csv("/Users/chiarasemenzin/filename_links.csv")
-#first_class=pd.read_csv("/Users/chiarasemenzin/Desktop/LSCP/experiment/results/lwl_results/first_classifications_last.csv")
-#classifs=pd.read_csv("/Users/chiarasemenzin/Desktop/zoon_metadata/classifs/zooniverse_class_june.csv")
-#metadata_phonses = pd.read_csv("/Users/chiarasemenzin/Desktop/zoon_metadata/metadata-phonses-merged.csv")
-
-# Clean dataset from first pilot files
-patternDel=".wav.mp3"
-filter = classifs['AudioData'].str.contains(patternDel)
-classifs_filter=classifs[~filter]
-
-
-def merge_dataset_simple(classifs,metadata,get_stats=True):
-    """Match classifications with metadata file without info other than clip name (e.g. rttms) """
-    df=OrderedDict({"AudioData":[], "Answer":[]})
-    print("Creating merged dataset...")
-    counter=0
-    clips=0
-    for i,j in metadata.iterrows():
-        matches=classifs[classifs.AudioData.str.contains(str(j.AudioData))]
-        counter+=matches.shape[0]
-        print("found {} matches".format(matches.shape[0]))
-        if matches.shape[0] != 0:
-            clips+=1
-        for answer in matches.Answer.tolist():
-            df["AudioData"].append(j.AudioData)
-            df["Answer"].append(answer)
-#    print("Count "+str(counter))
-    print("writing to CSV...")
-    df_new = pd.DataFrame(df, columns=df.keys())
-    df_new.to_csv("zoon_classifs_sep15.csv")
-    if get_stats:
-        log = open("/Users/chiarasemenzin/Desktop/SLT/logs/stats_log.txt", "w")
-        log.write("Total labels: {} \nTotal clips: {}".format(counter,clips))
[…]

### 2020-11-10 Scripts/write_seg_data.py

@@ -1,90 +1,0 @@
-import pandas as pd
-import numpy as np
-from collections import defaultdict, Counter
-import json
-
-# ----------------------------------------------------------------------------------------
-# Load data
-
-result_zoon="/Users/chiarasemenzin/Desktop/LSCP/SLT/classifications_17sept.csv" #classifs from zooniverse merged with md
-zoon=pd.read_csv(result_zoon)
-lisa=pd.read_csv("/Users/chiarasemenzin/Desktop/LSCP/SLT/result_final_lisa.csv") #file with metadata and PU classifications
-stats_log="stats_log.txt" #give a name to the log 
-
-with open("/Users/chiarasemenzin/Desktop/LSCP/SLT/dict_3.json", "r") as read_file: # Put here path to json file 
-    print("Converting JSON encoded data into Python dictionary")
-    d = json.load(read_file)
-
-
-# ----------------------------------------------------------------------------------------
-
-
-def agreement_by_chunk(d,classifs):
-    """Determine majority agreement within chunks"""
-    answers = defaultdict(list)
-    no_maj_count=0
-    no_3_labels=0
-    valid_count=0
-    total_labels=0
-    for k,v in d.items():
-        for chunk in v:
-            c=Counter(list(classifs.loc[classifs["AudioData"] == int(chunk)]["Answer"]))
-            total_labels+=len(c)
-            if len(c)!= 0:
-                value, count = c.most_common()[0]
-                print(sum([i for i in c.values()]))
-                if count >= 3 and sum([i for i in c.values()])>=3: #at least 3 classifs agree
-                    answers[k].append(value)
-                    valid_count+=1
-                    #print("Okay")
-                elif count < 3 and sum([i for i in c.values()])>=3:
-                    #print("No majority agreement")
-                    no_maj_count+=1
-                else:
-                    #print("Not enough data")
-                    no_3_labels+=1
-    log = open(stats_log, "w")
-    log.write(" Valid clips: {}\n No majority agreement: {}\n Less than 3 labels: {}, total labels: {}".format(valid_count,no_maj_count,no_3_labels,total_labels))
-    log.close()
-    return answers
-
-
-def agreement_by_segment(answers):
-    """Determine classification within segment """
-#    TODO
-    seg_dict=[[],[]]
-    for k,v in answers.items():
-        if set(v) != {"Junk"}: # Check Junk is not the only answer
[…]

### 2020-11-10 data-analyses/code/chunk_to_segment.py

@@ -1,0 +1,87 @@
+import numpy as np
+from collections import defaultdict
+import json
+
+# ----------------------------------------------------------------------------------------
+
+#This script creates a dictionary file in json format with mapping from 500ms chunks to Lab annotation segments
+
+
+result_final="/Users/chiarasemenzin/Desktop/LSCP/SLT/result_final_lisa.csv"  #file with segment info and lab classifications
+result_zoon="/Users/chiarasemenzin/Desktop/LSCP/SLT/merged_metadata.csv"  #metadata for zooniverse clips
+
+
+# Segment to timestamp
+def seg_to_ts(result_final):
+    seg2ts={}
+    with open(result_final) as final:
+        for e,line in enumerate(final):
+            line=line.split(",")
+            if e > 0:
+                seg2ts[line[0]]=(float(line[4]),float(line[5]),line[1])
+    return seg2ts
+
+# Timestamp to segment
+def chunk_to_ts(result_zoon):
+    chunk2ts={}
+    with open(result_zoon) as zoon:
+        for e,line in enumerate(zoon):
+            line=line.split(",")
+            if e > 0:
+                #chunk2ts[line[0]]=[float(i) for i in line[4].split("_")]
+                chunk2ts[line[1]]=[float(line[4]),float(line[5])] #get timestamps
+                chunk2ts[line[1]].append(line[2]) #get name of chunk
+    return chunk2ts
+
+# Calculate average duration of segment
+def avg_duration(result_final):
+    seg_ts=[]
+    with open(result_final) as final:
+        for e,line in enumerate(final):
+            line=line.split(",")
+            if e > 0:
+                seg_ts.append(float(line[5])-float(line[4]))
+    avg_dur=sum(seg_ts)/len(seg_ts)
+    return avg_duration
+
+def check(segments,chunks):
+    """" Check which chunks belong to which segment, by file"""
+
+    segments_final = defaultdict(list)
+
+    for name,chunk in chunks.items(): 
+        diffs_onset=[]
+        diffs_offeset=[]
+        candidate_seg=dict([i for i in segments.items() if i[1][2]==chunk[2]])
+        for k,seg in candidate_seg.items():
+            if seg[2] == chunk[2]:
+                print("Looking at ",k,seg[2])
+                diffs_onset.append([abs(chunk[0]-seg[0]),k])
+                diffs_offeset.append([abs(chunk[1]-seg[1]),k])
+        min_on=min([i[0] for i in diffs_onset])
[…]

### 2020-11-10 data-analyses/code/merge_datasets_zoon.py

@@ -1,0 +1,117 @@
+import pandas as pd
+from collections import OrderedDict
+import csv
+
+print("Started.")
+
+# define paths to your existing zooniverse classifications and metadata locations
+# These are the files you get as output from the Jupyter Notebook convert_zooniverse.ipynb
+
+classifs=pd.read_csv("/metadata/zooniverse_data_all_final.csv")
+metadata=pd.read_csv("/metadata/metadata_all_PU.csv",sep=',')
+filename_links=pd.read_csv("/metadata/filename_links.csv")
+#first_class=pd.read_csv("/Users/chiarasemenzin/Desktop/LSCP/experiment/results/lwl_results/first_classifications_last.csv")
+#classifs=pd.read_csv("/Users/chiarasemenzin/Desktop/zoon_metadata/classifs/zooniverse_class_june.csv")
+#metadata_phonses = pd.read_csv("/Users/chiarasemenzin/Desktop/zoon_metadata/metadata-phonses-merged.csv")
+
+# Clean dataset from first pilot files
+patternDel=".wav.mp3"
+filter = classifs['AudioData'].str.contains(patternDel)
+classifs_filter=classifs[~filter]
+
+
+def merge_dataset_simple(classifs,metadata,get_stats=True):
+    """Match classifications with metadata file without info other than clip name (e.g. rttms) """
+    df=OrderedDict({"AudioData":[], "Answer":[]})
+    print("Creating merged dataset...")
+    counter=0
+    clips=0
+    for i,j in metadata.iterrows():
+        matches=classifs[classifs.AudioData.str.contains(str(j.AudioData))]
+        counter+=matches.shape[0]
+        print("found {} matches".format(matches.shape[0]))
+        if matches.shape[0] != 0:
+            clips+=1
+        for answer in matches.Answer.tolist():
+            df["AudioData"].append(j.AudioData)
+            df["Answer"].append(answer)
+#    print("Count "+str(counter))
+    print("writing to CSV...")
+    df_new = pd.DataFrame(df, columns=df.keys())
+    df_new.to_csv("zoon_classifs_sep15.csv")
+    if get_stats:
+        log = open("/Users/chiarasemenzin/Desktop/SLT/logs/stats_log.txt", "w")
+        log.write("Total labels: {} \nTotal clips: {}".format(counter,clips))
+        log.close()
+    print("done.")
+
+
+
+def merge_dataset(classifs, metadata):
+    """Merge classifications with Metadata Containing:
[…]

### 2020-11-10 data-analyses/code/write_seg_data.py

@@ -1,0 +1,91 @@
+import pandas as pd
+import numpy as np
+from collections import defaultdict, Counter
+import json
+
+# ----------------------------------------------------------------------------------------
+# Load data
+
+result_final=pd.read_csv("/Users/chiarasemenzin/Desktop/LSCP/SLT/result_final_lisa.csv")  #file with segment info and lab classifications
+result_zoon="/Users/chiarasemenzin/Desktop/Final_JSLHR/zooniverse_classifications/merged_classifs_1751.csv"#classifs from zooniverse merged with md
+
+zoon=pd.read_csv(result_zoon)
+stats_log="stats_log_19oct.txt" #give a name to the log
+
+with open("/Users/chiarasemenzin/Desktop/Final_JSLHR/metadata/dict_4.json", "r") as read_file: # Put here path to json file
+    print("Converting JSON encoded data into Python dictionary")
+    d = json.load(read_file)
+
+
+# ----------------------------------------------------------------------------------------
+
+
+def agreement_by_chunk(d,classifs):
+    """Determine majority agreement within chunks"""
+    answers = defaultdict(list)
+    no_maj_count=0
+    no_3_labels=0
+    valid_count=0
+    total_labels=0
+    for k,v in d.items():
+        for chunk in v:
+            c=Counter(list(classifs.loc[classifs["AudioData"] == int(chunk)]["Answer"]))
+            total_labels+=len(c)
+            if len(c)!= 0:
+                value, count = c.most_common()[0]
+                print(sum([i for i in c.values()]))
+                if count >= 3 and sum([i for i in c.values()])>=3: #at least 3 classifs agree
+                    answers[k].append(value)
+                    valid_count+=1
+                    #print("Okay")
+                elif count < 3 and sum([i for i in c.values()])>=3:
+                    #print("No majority agreement")
+                    no_maj_count+=1
+                else:
+                    #print("Not enough data")
+                    no_3_labels+=1
+    log = open(stats_log, "w")
+    log.write(" Valid clips: {}\n No majority agreement: {}\n Less than 3 labels: {}, total labels: {}".format(valid_count,no_maj_count,no_3_labels,total_labels))
+    log.close()
+    return answers
+
+
+def agreement_by_segment(answers):
+    """Determine classification within segment """
+#    TODO
+    seg_dict=[[],[]]
+    for k,v in answers.items():
[…]

### 2020-11-10 data_analyses/code/chunk_to_segment.py
reproducibility not tested

@@ -7,8 +7,8 @@ import json
 #This script creates a dictionary file in json format with mapping from 500ms chunks to Lab annotation segments
 
 
-result_final="/Users/chiarasemenzin/Desktop/LSCP/SLT/result_final_lisa.csv"  #file with segment info and lab classifications
-result_zoon="/Users/chiarasemenzin/Desktop/LSCP/SLT/merged_metadata.csv"  #metadata for zooniverse clips
+result_final="../metadata/result_final_lisa.csv"  #file with segment info and lab classifications
+result_zoon="../metadata/metadata_all_PU.csv"  #metadata for zooniverse clips
 
 
 # Segment to timestamp

### 2020-11-10 data_analyses/code/merge_datasets_zoon.py

@@ -7,9 +7,11 @@ print("Started.")
 # define paths to your existing zooniverse classifications and metadata locations
 # These are the files you get as output from the Jupyter Notebook convert_zooniverse.ipynb
 
-classifs=pd.read_csv("/metadata/zooniverse_data_all_final.csv")
-metadata=pd.read_csv("/metadata/metadata_all_PU.csv",sep=',')
-filename_links=pd.read_csv("/metadata/filename_links.csv")
+classifs=pd.read_csv("../zooniverse_classifications/zooniverse_data_all_final.csv")
+metadata=pd.read_csv("../metadata/metadata_all_PU.csv",sep=',')
+filename_links=pd.read_csv("../metadata/filename_links.csv")
+
+
 #first_class=pd.read_csv("/Users/chiarasemenzin/Desktop/LSCP/experiment/results/lwl_results/first_classifications_last.csv")
 #classifs=pd.read_csv("/Users/chiarasemenzin/Desktop/zoon_metadata/classifs/zooniverse_class_june.csv")
 #metadata_phonses = pd.read_csv("/Users/chiarasemenzin/Desktop/zoon_metadata/metadata-phonses-merged.csv")

### 2020-11-17 202010_jslhr/lena-ana/add-lena-ana.Rmd
almost ready to submit

--- 2020-10-17 202010_jslhr/add-lena-ana.Rmd
@@
 
+NOTE: This section cannot be reproduced because it requires access to data that haven't been shared.
+
 ```{r read-data,eval=F}
-read.csv("../Derived_Data/result_final_lisa.csv")->metadata
+read.csv("result_final_lisa.csv")->metadata
 metadata[1,]
@@
 }
-write.csv(all_segments,"../Derived_Data/LENA_type_lab.csv",row.names = F)
+write.csv(all_segments,"LENA_type_lab.csv",row.names = F)
 ```
@@
 
-read.csv("../Derived_Data/LENA_type_lab.csv")->all_segments
+read.csv("LENA_type_lab.csv")->all_segments
 

### 2020-11-17 202010_jslhr/paper.Rmd
almost ready to submit

@@
 # #up to 3 keywords
-# keywords          : " reliability; validation; daylong audiorecordings"
+# keywords          : "reliability; validation; daylong audiorecordings"
 # wordcount         : "7,113"
@@
 
-figsintext        : yes
+figsintext        : no
 figurelist        : no
@@
 class             : "man"
-output            : papaja::apa6_pdf
+output            : papaja::apa6_word
 
@@
 
-data_ang <- read.csv("../Derived_Data/classifications_PU_zoon_final17.csv",header=T,sep=",")
-data_td <- read.csv("../Derived_Data/classifications_PU_zoon_final.csv")
+data_ang <- read.csv("segment_AS.csv",header=T,sep=",")
+data_td <- read.csv("segment_LRC.csv")
 data_all<-rbind(data_ang, data_td)
@@
 
-demo_data=read.csv("../Derived_Data/demo-data.tsv",sep="\t")
+demo_data=read.csv("demo-data.tsv",sep="\t")
 #add filenames to demo data, to be used later
@@
 
-kable(tab1,cap=mycaption, col.names = NULL)%>%
-  collapse_rows(columns = 1:6, valign = "middle") %>%
-add_header_above(c(" " = 1, "Recall"= 3, "Label Frequency" = 2))
+kable(tab1,cap=mycaption, col.names = NULL) %>%
+  collapse_rows(columns = 1:6, valign = "middle") #%>%
+#add_header_above(c(" " = 1, "Recall"= 3, "Label Frequency" = 2))
 ```
@@
 ```{r bychunk-age, echo=F}
-read.csv("../Derived_Data/chunk_majority_agreement_1751.csv")->chunk
+read.csv("zoo_chunk_majority_agreement.csv")->chunk
 bychild=table(chunk$Answer,chunk$ChildID)
@@
 
-We are grateful to ...
-
-\newpage
-
-MOVE FIGURE 1 HERE
-
-\newpage
-
-MOVE FIGURE 2 HERE
-
-etc
+We are grateful to the families who contributed their data; to the Zooniverse volunteers that made this work possible; and the audience and technical committee of SLT 2020 for helpful feedback on preliminary analyses.
 

### 2020-11-17 202010_jslhr/paper.tex
almost ready to submit

@@
 %
-\documentclass[english,,man,floatsintext]{apa6}
+\documentclass[english,,man]{apa6}
 \usepackage{lmodern}
@@
 Correspondence concerning this article should be addressed to Alejandrina Cristia, 29 rue d'Ulm, 75005, Paris, France. E-mail: alecristia@gmail.com}
+\DeclareDelayedFloatFlavor{ThreePartTable}{table}
+\DeclareDelayedFloatFlavor{lltable}{table}
+\DeclareDelayedFloatFlavor*{longtable}{table}
+\makeatletter
+\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
+\makeatother
 \usepackage{lineno}
@@
 We also observed quite a bit of variance across the 50 runs.
-For Linguistic Proportion, the range of correlations found for all infants was {[}0.51,0.91{]}; for the Angelman syndrome group {[}0.60,0.96{]}; for the low-risk group {[}0.05,0.96{]}.
-For Canonical Proportion, the range of correlations found for all infants was {[}0.82,0.96{]}; for the Angelman syndrome group {[}0.59,0.95{]}; for the low-risk group {[}0.77,0.98{]}.
+For Linguistic Proportion, the range of correlations found for all infants was {[}0.44,0.89{]}; for the Angelman syndrome group {[}0.49,0.97{]}; for the low-risk group {[}0.29,0.99{]}.
+For Canonical Proportion, the range of correlations found for all infants was {[}0.75,0.94{]}; for the Angelman syndrome group {[}0.79,0.99{]}; for the low-risk group {[}0.55,0.94{]}.
 
@@
 \hline
-100 seg AS & 0.855 & 0.809\\
+100 seg AS & 0.770 & 0.912\\
 \hline
@@
 \hline
-100 seg LR & 0.609 & 0.907\\
+100 seg LR & 0.841 & 0.808\\
 \hline
@@
 \hline
-100 seg all & 0.761 & 0.889\\
+100 seg all & 0.719 & 0.873\\
 \hline

### 2020-11-17 202010_jslhr/paper.Rmd
ready to submit

@@
 Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development [@oller2000emergence].
-In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization count in a sample of children diagnosed with an autism spectrum disorder [@mcdaniel2020predicting].
+In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization counts in a sample of children diagnosed with an autism spectrum disorder [@mcdaniel2020predicting].
 
@@
 Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received three classifications into Crying, Laughing, Canonical, Non-Canonical, or Junk (with the latter tag used for clips that did not contain a child voice).
-They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgement of canonical or non-canonical.
+They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgment of canonical or non-canonical.
 They found that this Canonical Proportion increased with infant age in an analysis collapsing across corpora, which provides a first proof of principle that citizen scientists' annotations could be an appropriate solution to the problem of annotating child vocalizations from wearables.
@@
 As children diagnosed with Angelman syndrome typically have severe cognitive and language delays and most do not produce more than 1-2 words consistently, we compare their linguistic production data against that of younger children with theoretically similar language profiles.
-This study was approved by the local institutional review board under the project name Neurodevelopmental Natu- ral History Study, Purdue University, IRB-1811021381.
+This study was approved by the local institutional review board under the project name Neurodevelopmental Natural History Study, Purdue University, IRB-1811021381.
 
@@
 Following recommendations from the Zooniverse board, we collected five judgments per clip, rather than three as used in the laboratory.
-Majority agreement is thus achieved when three out of five judgements agree (rather than two out of three).
+Majority agreement is thus achieved when three out of five judgments agree (rather than two out of three).
 
@@
 Nearly a fifth of chunks did not have at least 3 labels in agreement out of the 5 Zooniverse labels (N = 6,585, 19% of all chunks).
-Of the chunks without a majority agreement, 4341 (66%) contained one or two Junk judgements (out of 5), 6523 (99,9%) had at least two matching judgements (the threshold used for lab-annotated segments), and only 61 (0,01%) had 5 different judgements.
+Of the chunks without a majority agreement, 4341 (66%) contained one or two Junk judgments (out of 5), 6523 (99,9%) had at least two matching judgments (the threshold used for lab-annotated segments), and only 61 (0,01%) had 5 different judgments.
 Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the 81% of chunks that did have at least 3 labels in agreement; this represented 135,725 labels for 27,145 chunks, corresponding to 11,593 LENA^TM^ segments.
@@
 
-We did this by considering the majority label extracted from Zooniverse judgements for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were Junk, then the segment was labelled as Junk; then, following a hierarchy (Canonical>Non-canonical>Crying>Laughing), segments with at least one instance of each judgement were labelled as such:
-
-- if one Canonical judgement was present, then the segment was labelled as Canonical, 
-- else, if a Non-Canonical judgement was present, the segment was labelled as Non-Canonical, 
+We did this by considering the majority label extracted from Zooniverse judgments for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were Junk, then the segment was labelled as Junk; then, following a hierarchy (Canonical>Non-canonical>Crying>Laughing), segments with at least one instance of each judgment were labelled as such:
+
+- if one Canonical judgment was present, then the segment was labelled as Canonical, 
+- else, if a Non-Canonical judgment was present, the segment was labelled as Non-Canonical, 
 - else, if a Crying judgment was present, the segment was labelled as Crying,
@@
 
-Although the classification at the clip level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of childrens’ individual development.
+Although the classification at the clip level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of children's individual development.
 We illustrate this association in Figure 3.
@@
 Since all of these results are very similar to those obtained when first matching clips to segments, we conclude that in the future this step may not be necessary.
-Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' clip level judgements [which was in fact what @cychosz2019canonical did].
+Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' clip level judgments [which was in fact what @cychosz2019canonical did].
 
@@
 This was the case of the "Junk/Don't mark" category.
-Higher precision and low recall indicates that when Zooniverse annotations return a "Junk" judgement, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as "Don't mark" in the laboratory.
+Higher precision and low recall indicates that when Zooniverse annotations return a "Junk" judgment, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as "Don't mark" in the laboratory.
 Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the "Junk" category less) than laboratory annotators, who have access to the whole vocalization.
 This is consistent with the fact that our lab definition of "Junk" was more stringent than the Zooniverse definition.
-Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use "Don't mark" if there was *any* noise or verlapping speech that would potentially affect acoustic analyses.
+Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use "Don't mark" if there was *any* noise or overlapping speech that would potentially affect acoustic analyses.
 
@@
 It is also important to note that work on children with non-normative development poses in itself some challenges.
-Our data contained a lower number of vocalizations for children diagnosed with Angelman syndrome, even though we selected much younger low-risk control childrens as the comparison group.
+Our data contained a lower number of vocalizations for children diagnosed with Angelman syndrome, even though we selected much younger low-risk control children as the comparison group.
 However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.

### 2020-11-17 202010_jslhr/paper.Rmd
minor

--- 2020-11-17 202010_jslhr/paper.Rmd
@@
 Third, rows labeled 100 seg indicate that laboratory- and Zooniverse-derived metrics were based on only 100 segments (median over 50 runs in which 100 segments were randomly selected from each child).")
+write.csv(mycortab_sorted,"mycortab_sorted.csv")
 ```

### 2021-01-04 202010_jslhr/paper.Rmd
preparing sub2, not all comments have been integrated

@@
 
-# type: speech; areas of interest: assessment, infants, meta-analyses, prevention, speech recognition, screening, language, early intervention, cultural and ling diversity, dev lang disorders, autism, bilingualism --> choose 5: infants, prevention, dev lang disorders, early interv, diagnostics
-# Clinical practice research that evaluates methods of prevention, screening, diagnosis, treatment or models of service delivery that may improve the function and quality of life for individuals with or at risk of developing disorders of speech, language, cognition, swallowing, audition, and/or balance.
 # #up to 3 keywords
-# keywords          : "reliability; validation; daylong audiorecordings"
+# keywords          : "reliability; validation; daylong audiorecordings; citizen science; Zooniverse; LENA; long-form recordings"
 # wordcount         : "7,113"
@@
 The present work reports on the validity of labels and child-level descriptors of children's vocalizations gathered from citizen scientists, in a comparison to expert lab annotators.
+
 In the rest of the Introduction, we first briefly summarize two metrics that can be used to describe individual children's vocalizations in the context of long-form recordings.
@@
 
-A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis, and it is beyond the scope of this paper to provide a full summary of this work.
-We instead focus on two descriptors -- Linguistic Proportion and Canonical Proportion -- that are particularly amenable to the kind of big data approaches long-form recordings require.
-
-To understand why these two descriptors are relevant, it is worthwhile to first consider how the way in which vocal development has been studied in the past may not easily translate to long-form recordings.
+A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis, and it is beyond the scope of this paper to provide a full summary of this work [but see **XX ADD REFERENCE** for a recent summary].
+Moreover, the way in which vocal development has been studied in the past may not easily translate to long-form recordings.
 As mentioned earlier, long-form recordings have several advantages, including capturing the child's vocal patterns in their natural environment and being able to accumulate a great amount of data easily, which is particularly useful for diagnoses and pathologies characterized by low levels of vocal production.
@@
 
+#### Automated first-pass analyses
+
 The most commonly used software was created by the LENA^TM^ Foundation, and it returns a diarization of the audio signal split into key talkers, including the child wearing the device, as well as a split of these audio attributed to the child into three categories: crying, other fixed and vegetative signals, and speech vocalizations.
 These subcategories have not been widely validated [@cristia2020accuracy].
-To provide an idea of their accuracy, we classified the same segments that are used in the present study using LENA^TM^'s subtypes[^1] to compared these against the laboratory annotations used in the present study.
-Table 1 shows that LENA^TM^ has a good recall for speech-like vocalizations and crying but not laughing.
 Moreover, LENA^TM^ lacks an important distinction between more and less advanced vocalizations.
-
-<!-- AS said about not laughing: might be worth noting how low frequency this is and how that might be part of the problem in addition to being difficult to distinguish from crying. -->
-<!-- ac2as I'm not sure this is needed here -- we do mention that laughing is rare in the next paragraph -->
-
-[^1]: If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as "Junk" or not categorized.
 
@@
 
-Other automatized algorithms have been developed in recent years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical [@schuller2019interspeech].
+**CHECK FLOW IN THIS PARA** Other automatized algorithms have been developed in recent years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical [@schuller2019interspeech].
 The split of speech vocalizations into canonical and non-canonical, and the overt inclusion of laughing, allow a somewhat finer-grained picture of children's vocal activity.
 An advantage of these algorithms is that there exists a challenge where performance has been benchmarked, so unlike LENA^TM^'s algorithms, we know which algorithms perform better than others.
-The ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55% (see Table 1).
-The team who won the challenge improved this by about 2%, primarily through gains in the laughing class obtained by adding training data [@yeh2019using].
+The ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55%.
+To provide an idea of the accuracy across these different methods, we classified the same segments that are used in the present study using LENA^TM^'s subtypes[^1] to compare these against the laboratory annotations used in the present study; and we also show the results of ComParE 2019 BabySounds sub-challenge (although note it is not based on the same data as the other two columns).
+Table 1 shows that LENA^TM^ has a good recall for speech-like vocalizations and crying but not laughing.
+**ADD MORE INFORMATION SO READERS UNDERSTAND THE POINT OF THIS TABLE.**  The team who won the challenge improved this by about 2%, primarily through gains in the laughing class obtained by adding training data [@yeh2019using].
 Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are weighted based on their frequency of occurrence, changes very much.
+
+[^1]: If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as "Junk" or not categorized.
+
+#### Derived metrics
 
@@
 
-For example, most work using LENA^TM^ software reports on children's vocalization counts.
-Although this has been sometimes criticized as being more about quantity than quality [@mcdaniel2020predicting], this is a promising metric because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA^TM^ [@cristia2020accuracy].
+For example, most work using LENA^TM^ software reports on children's vocalization counts -- this is a derived metric because it is not simply a description of sections of the audio, but instead it integrates over a certain period of time.
+Although this has been sometimes criticized as being more about quantity than quality [@mcdaniel2020predicting], this is a promising metric of individuals' vocalization development because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA^TM^ [@cristia2020accuracy].
 Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r ~ .3 with standardized language scores in a meta-analysis [@wang2020meta].
@@
 
-<!-- AS said about voc count not sure I agree: https://pubs.asha.org/doi/full/10.1044/2020_JSLHR-19-00281?casa_token=fBByMmPlxbMAAAAA%3AbS2O5Mrymn30_aQWybitwMecENkDYLTjzexCqZcnX2Rx1VJ7OaT9qkIGNAAM95xRBGW_D3XX7mMEWHo -->
-<!-- ac2as fair point -- I'll mention this paper below, when explaining why we may want to go beyond CVC -->
-
 Here, we focus on two alternative composite metrics.
-One additional metric that can be derived once children's vocalizations are split into crying, laughing, canonical, and non-canonical is the Linguistic Proportion: the proportion of vocalizations that are linguistic (canonical and non-canonical) out of all vocalizations.
+One metric that can be derived once children's vocalizations are split into crying, laughing, canonical, and non-canonical is the Linguistic Proportion: the proportion of vocalizations that are linguistic (canonical and non-canonical) out of all vocalizations.
 To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age.
-
-<!-- BT said about LP May want to put this 2nd as more exploatory given more lit on can. prop. -->
-<!-- I see the point, but doesn't it make more logical sense to discuss it first? -->
 
@@
 
+### Crowdsourcing: A potential solution for annotation
+
 In sum, there is a growing literature attempting to use data from daylong recordings to describe young children's vocalizations, but there are two outstanding challenges.
-The first pertains to how the data are analyzed, with human annotation being costly.
-The second relates to how useful descriptors are extracted, both in terms of individual vocalizations and at the child level.
-
-### Crowdsourcing: A potential solution for annotation
+The first pertains to how the data are annotated, with human annotation being costly.
+The second relates to how descriptors of vocal development are generated, i.e., how annotations of audio sections are integrated into vocal development metrics.
 
@@
 Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable [@berinsky2012evaluating], some question marks are raised as some "workers" turn out to be bots, or are poorly motivated (and potentially exploited) humans.
+
+#### Citizen scientists to the rescue
 
@@
 Thus, the present study aimed to contribute a key piece of evidence missing in this discussion: To what extent do laboratory and citizen science annotations agree when describing young children's vocalizations?
-We examined the extent to which such classifications agree, by quantifying the correspondence across these two modes of annotations at the level of individual clips (vocalizations), and at the level of individual children.
-It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels.
-We therefore included children in a wide age range, with some of the children being diagnosed with Angelman syndrome, as well as low-risk controls.
+We examined the extent to which such classifications agree, by quantifying the correspondence across these two modes of annotations at the level of individual segments, and at the level of individual children.
 
@@
 
-<!-- We additionally explored under what conditions Zooniverse judgments are more closely aligned with laboratory judgments using linear mixed models. -->
+It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels.
+We therefore included children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range, as well as a group of low-risk control infants.
+**Some additional information about Angelman syndrome would be helpful to understand the findings.
+Were the different patterns between the canonical and linguistic proportions predicted by the already-known characteristics of this syndrome?
+Has other work included this population?
+Why was this population selected?
+Are there hypotheses about the validity between groups or is this more exploratory?** **relative dearth of LENA data with atypical populations **
 
@@
 Specifically, the segments extracted were automatically cut into clips of exactly 500ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform.
-To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms clips and their corresponding segment.
-Clips were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling [@bowyer2015panoptes].
+To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms chunks and their corresponding segment.
+Chunks were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling [@bowyer2015panoptes].
 
@@
 Majority agreement is thus achieved when three out of five judgments agree (rather than two out of three).
-
-<!-- In exploratory analyses, we assess the impact of collecting a smaller number of judgments per clip. -->
 
@@
 See Figure 2 for results.
+**the children with Angelman syndrome scored much lower than the typically-developing infants on the canonical measure but higher on the linguistic measure.**
 
@@
 
-Above, we concluded that derived metrics seem more promising than segment-level data.
-Notice that derived metrics do not require matching of clips (500 ms presented to Zooniverse participants) to segments (the original LENA^TM^ segments presented to laboratory participants).
+Above, we concluded that derived metrics integrating information across audio chunks or clips (linguistic and canonical proportion) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of clips (500 ms presented to Zooniverse participants) to segments (the original LENA^TM^ segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
 As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
@@
 
+**remaining methodological problems with the LENA pipeline (p. XX);
+the relatively small sample sizes of the two groups we studied (p. XX);
+the kinds of populations that we cannot be certain this work generalizes to (p. XX).
+relative dearth of LENA data with atypical populations **
+
 It is also important to note that work on children with non-normative development poses in itself some challenges.

### 2021-01-05 202010_jslhr/paper.Rmd
revision in progress, no longer knits to docx

@@
 
-<!-- AS said: maybe name the disorders & delays here? -->
-<!-- ac2as I removed this because I haven't read the papers, and I'm not sure Chiara has either, so I cannot be sure that the citations are correct -- they were taken from Meg's paper.
-In any case, since this is about onset, which we don't measure here, I don't think this section is needed: In addition, delayed onset of canonical babbling has been documented in a wide variety of disorders and delays (Belardi et al., 2017; Fasolo et al., 2008; Lang et al., 2019; Patten et al., 2014; Stoel-Gammon, 2016).
--->
-
 ### Crowdsourcing: A potential solution for annotation
@@
 For the English and Tsimane' corpora, vocalizations were automatically identified using LENA^TM^, whereas the other two were extracted through manual segmentation.
-Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received three classifications into Crying, Laughing, Canonical, Non-Canonical, or Junk (with the latter tag used for clips that did not contain a child voice).
+Segments were then split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received three classifications into Crying, Laughing, Canonical, Non-Canonical, or Junk (with the latter tag used for clips that did not contain a child voice).
 They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgment of canonical or non-canonical.
@@
 
-When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk vocalizations); and the Canonical Proportion (Canonical)/(all linguistic vocalizations).
+When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk judgments); and the Canonical Proportion (Canonical)/(all linguistic judgments).
 Previous work suggests that children with Angelman syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the Canonical Proportion was expected to increase with age [@hamrick2019capturing;@hamrick2019measurement].
@@
 
-<!-- Reliability is built into the annotation routine such that each segment is coded by 3 separate coders, and only segments where 2 of 3 coders agree are used in subsequent phases of this study.
--->
-<!-- Lisa aC removed this because this is actually false -- all segments were put up, including those with fewer than 2 agreement -->
-
-<!-- Lisa, What happens if there are multiple syllables or multiple words? also, what if there is speech and non-speech, eg cry, in the same segment? -->
-
 ### Annotation on Zooniverse
@@
 A substantial concern that emerges when considering citizen science annotation for spontaneous speech data is the risk of a privacy breach: even short clips can contain personal information that can expose the identity of the speaker, such as addresses or names, or contain sensitive and private information.
-Fortunately, @Seidl19 determined that when segments are divided into shorter chunks (400-600ms), human annotators with little training can code our categories of interest (Canonical, Laughing, Crying, etc.) with a classification quality comparable to the one carried out on full segments.
-We therefore used very short chunks (500ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.
+Fortunately, @Seidl19 determined that when segments are divided into shorter clips (400-600ms), human annotators with little training can code our categories of interest (Canonical, Laughing, Crying, etc.) with a classification quality comparable to the one carried out on full segments.
+We therefore used very short clips (500ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.
 
 Specifically, the segments extracted were automatically cut into clips of exactly 500ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform.
+We will call this level *chunk*, to convey that they are part of a bigger unit (the LENA^TM^-identified "segment").
 To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms chunks and their corresponding segment.
@@
 
-Users were asked to assign each 500ms clip to one out of five possible categories: (1) Canonical, (2) Non-Canonical, (3) Crying, (4) Laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
-Notice that the latter is narrower than the "Don't mark" category of the lab annotation routine, which was used any time there was noise or overlap that could affect acoustic analyses.
-
-Following recommendations from the Zooniverse board, we collected five judgments per clip, rather than three as used in the laboratory.
-Majority agreement is thus achieved when three out of five judgments agree (rather than two out of three).
-
 ``` {r fig-zoo, comment=F, message=F, hide = T, warning = F, echo = F,  fig.cap="Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needed to make; as well as the Field Guide, which remained accessible on the right of the screen throughout the session, and which allowed users to revise examples of the different categories."}
@@
 ```
+
+Users were asked to assign each 500ms chunk to one out of five possible categories: (1) Canonical, (2) Non-Canonical, (3) Crying, (4) Laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
+Notice that the latter is narrower than the "Don't mark" category of the lab annotation routine, which was used any time there was noise or overlap that could affect acoustic analyses.
+
+Following recommendations from the Zooniverse board, we collected five judgments per chunk, rather than three as used in the laboratory.
+Majority agreement is thus achieved when three out of five judgments agree (rather than two out of three).
 
@@
 LENA^TM^ segments are not equivalent to child vocalizations: A segment may contain 1 or more vocalizations, which may be of the same type or not.
+Figure 2 conveys the relationship between chunks and segments, and how judgments are combined into labels.
+
+```{r fig-process, comment=F, message=F, hide = T, warning = F, echo = F, fig.cap = "Correspondance between LENA-identified segments (which can contain one or more vocalizations) and chunks.
+Notice that both segments and chunks are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority.
+To compare laboratory and Zooniverse annotations at the level of segments, chunk-level judgments are combined using the hierarchy Canonical>Non-canonical>Crying>Laughing."}
+knitr::include_graphics("fig_levels.key.pdf")
+
+```
 
@@
 
-Next, we discuss the correspondence between citizen science classifications and the laboratory gold standard, at the level of individual clips.
+Next, we discuss the correspondence between citizen science classifications and the laboratory gold standard, at the level of individual segments.
 Results were visualized with a confusion matrix showing precision and recall (Figure 3): the diagonal elements show the number of correct segment-level classifications for each class while the off-diagonal elements show non-matching classifications.
@@
 
-Although the classification at the clip level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of children's individual development.
+Although the classification at the segment level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of children's individual development.
 We illustrate this association in Figure 3.
@@
 
-Above, we concluded that derived metrics integrating information across audio chunks or clips (linguistic and canonical proportion) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of clips (500 ms presented to Zooniverse participants) to segments (the original LENA^TM^ segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
+Above, we concluded that derived metrics integrating information across audio clips (linguistic and canonical proportion, which can be derived from segment- or chunk-level judgments) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA^TM^ segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
 As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
@@
 
-Since all of these results are very similar to those obtained when first matching clips to segments, we conclude that in the future this step may not be necessary.
-Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' clip level judgments [which was in fact what @cychosz2019canonical did].
+Since all of these results are very similar to those obtained when first matching chunks to segments, we conclude that in the future this step may not be necessary.
+Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' chunk level judgments [which was in fact what @cychosz2019canonical did].
 
@@
 kable(mycortab_sorted,cap="Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways.
-First, All seg indicates that Zooniverse annotations at the clip level were first combined at the segment level.
-Second, Chunks indicates that they were analyzed directly at the clip level.
+First, All seg indicates that Zooniverse annotations at the chunk level were first combined at the segment level.
+Second, Chunks indicates that they were analyzed directly at the chunk level.
 Both of these are based on all the data.
@@
 This is likely due to the fact that lab annotators had access to the whole segment, and could even listen to the context of that segment, so they could accurately interpret a segment as crying.
-Thus, it is possible that additional analyses, using for instance pitch or duration characteristics at the level of the whole segment rather than the clip could improve classification performance, but for the time being, we cannot recommend Zooniverse annotations to identify crying bouts.
+Thus, it is possible that additional analyses, using for instance pitch or duration characteristics at the level of the whole segment rather than the chunk could improve classification performance, but for the time being, we cannot recommend Zooniverse annotations to identify crying bouts.
 
@@
 
-In sum, we believe these data show that young children's vocalizations can be accurately described by relying on the judgments of Zooniverse's citizen scientists, particularly when the goal is to describe vocalizations at a broader level than individual clips.
+In sum, we believe these data show that young children's vocalizations can be accurately described by relying on the judgments of Zooniverse's citizen scientists, particularly when the goal is to describe vocalizations at a broader level than individual chunks or segments.
 It is important to note that from a methodological point of view, the use of crowdsourcing is not merely a question of being an ecological technique, but it can also offer benefits in the area of scientific rigor and reproducibility.
@@
 Now that this method has been validated, researchers can extract all segments that LENA^TM^ identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
-
-<!-- AC removed: Additionally, we found an imbalance between the quantity of clips assigned to each category.
-Further research should look at validating the results of our approach by oversampling data from non-typically developing children, so as to overcome both the risk of lack of data in some categories, and the imbalance in the quantity of vocalizations between groups.
--->
 
@@
 In this study we validated the quality of annotations obtained through a citizen science platform, Zooniverse, as compared to a gold standard of human expert annotators.
-We analyzed the correspondence between annotations at the individual clip level, and the individual-child level, using Canonical Proportion and Linguistic Proportion as descriptors.
+We analyzed the correspondence between annotations at the individual segment level, and the individual-child level, using Canonical Proportion and Linguistic Proportion as descriptors.
 We found moderate to good accuracy in the first, and strong positive correlations in the latter.

### 2021-01-05 202010_jslhr/paper.tex
revision in progress, no longer knits to docx

@@
 
-A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis, and it is beyond the scope of this paper to provide a full summary of this work.
-We instead focus on two descriptors -- Linguistic Proportion and Canonical Proportion -- that are particularly amenable to the kind of big data approaches long-form recordings require.
-
-To understand why these two descriptors are relevant, it is worthwhile to first consider how the way in which vocal development has been studied in the past may not easily translate to long-form recordings.
+A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis, and it is beyond the scope of this paper to provide a full summary of this work {[}but see \textbf{XX ADD REFERENCE} for a recent summary{]}.
+Moreover, the way in which vocal development has been studied in the past may not easily translate to long-form recordings.
 As mentioned earlier, long-form recordings have several advantages, including capturing the child's vocal patterns in their natural environment and being able to accumulate a great amount of data easily, which is particularly useful for diagnoses and pathologies characterized by low levels of vocal production.
@@
 
+\hypertarget{automated-first-pass-analyses}{%
+\paragraph{Automated first-pass analyses}\label{automated-first-pass-analyses}}
+
 The most commonly used software was created by the LENA\textsuperscript{TM} Foundation, and it returns a diarization of the audio signal split into key talkers, including the child wearing the device, as well as a split of these audio attributed to the child into three categories: crying, other fixed and vegetative signals, and speech vocalizations.
 These subcategories have not been widely validated (Cristia et al., 2020).
-To provide an idea of their accuracy, we classified the same segments that are used in the present study using LENA\textsuperscript{TM}'s subtypes\footnote{If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as \enquote{Junk} or not categorized.} to compared these against the laboratory annotations used in the present study.
-Table 1 shows that LENA\textsuperscript{TM} has a good recall for speech-like vocalizations and crying but not laughing.
 Moreover, LENA\textsuperscript{TM} lacks an important distinction between more and less advanced vocalizations.
@@
 \hline
-\multicolumn{1}{c|}{ } & \multicolumn{3}{c|}{Recall} & \multicolumn{2}{c}{Label Frequency} \\
-\cline{2-4} \cline{5-6}
  & C2019B & LENA & Zoon. & C2019B & Lab\\
@@
 
-Other automatized algorithms have been developed in recent years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical (Schuller et al., 2019).
+\textbf{CHECK FLOW IN THIS PARA} Other automatized algorithms have been developed in recent years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical (Schuller et al., 2019).
 The split of speech vocalizations into canonical and non-canonical, and the overt inclusion of laughing, allow a somewhat finer-grained picture of children's vocal activity.
 An advantage of these algorithms is that there exists a challenge where performance has been benchmarked, so unlike LENA\textsuperscript{TM}'s algorithms, we know which algorithms perform better than others.
-The ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55\% (see Table 1).
-The team who won the challenge improved this by about 2\%, primarily through gains in the laughing class obtained by adding training data (Yeh et al., 2019).
+The ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55\%.
+To provide an idea of the accuracy across these different methods, we classified the same segments that are used in the present study using LENA\textsuperscript{TM}'s subtypes\footnote{If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as \enquote{Junk} or not categorized.} to compare these against the laboratory annotations used in the present study; and we also show the results of ComParE 2019 BabySounds sub-challenge (although note it is not based on the same data as the other two columns).
+Table 1 shows that LENA\textsuperscript{TM} has a good recall for speech-like vocalizations and crying but not laughing.
+\textbf{ADD MORE INFORMATION SO READERS UNDERSTAND THE POINT OF THIS TABLE.} The team who won the challenge improved this by about 2\%, primarily through gains in the laughing class obtained by adding training data (Yeh et al., 2019).
 Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are weighted based on their frequency of occurrence, changes very much.
+
+\hypertarget{derived-metrics}{%
+\paragraph{Derived metrics}\label{derived-metrics}}
 
@@
 
-For example, most work using LENA\textsuperscript{TM} software reports on children's vocalization counts.
-Although this has been sometimes criticized as being more about quantity than quality (McDaniel, Yoder, Estes, \& Rogers, 2020), this is a promising metric because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA\textsuperscript{TM} (Cristia et al., 2020).
+For example, most work using LENA\textsuperscript{TM} software reports on children's vocalization counts -- this is a derived metric because it is not simply a description of sections of the audio, but instead it integrates over a certain period of time.
+Although this has been sometimes criticized as being more about quantity than quality (McDaniel, Yoder, Estes, \& Rogers, 2020), this is a promising metric of individuals' vocalization development because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA\textsuperscript{TM} (Cristia et al., 2020).
 Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r \textasciitilde{} .3 with standardized language scores in a meta-analysis (Wang, Williams, Dilley, \& Houston, 2020).
@@
 Here, we focus on two alternative composite metrics.
-One additional metric that can be derived once children's vocalizations are split into crying, laughing, canonical, and non-canonical is the Linguistic Proportion: the proportion of vocalizations that are linguistic (canonical and non-canonical) out of all vocalizations.
+One metric that can be derived once children's vocalizations are split into crying, laughing, canonical, and non-canonical is the Linguistic Proportion: the proportion of vocalizations that are linguistic (canonical and non-canonical) out of all vocalizations.
 To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age.
@@
 Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development (Oller, 2000).
-In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization count in a sample of children diagnosed with an autism spectrum disorder (McDaniel et al., 2020).
-
-In sum, there is a growing literature attempting to use data from daylong recordings to describe young children's vocalizations, but there are two outstanding challenges.
-The first pertains to how the data are analyzed, with human annotation being costly.
-The second relates to how useful descriptors are extracted, both in terms of individual vocalizations and at the child level.
+In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization counts in a sample of children diagnosed with an autism spectrum disorder (McDaniel et al., 2020).
 
@@
 \subsubsection{Crowdsourcing: A potential solution for annotation}\label{crowdsourcing-a-potential-solution-for-annotation}}
+
+In sum, there is a growing literature attempting to use data from daylong recordings to describe young children's vocalizations, but there are two outstanding challenges.
+The first pertains to how the data are annotated, with human annotation being costly.
+The second relates to how descriptors of vocal development are generated, i.e., how annotations of audio sections are integrated into vocal development metrics.
 
@@
 Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable (Berinsky, Huber, \& Lenz, 2012), some question marks are raised as some \enquote{workers} turn out to be bots, or are poorly motivated (and potentially exploited) humans.
+
+\hypertarget{citizen-scientists-to-the-rescue}{%
+\paragraph{Citizen scientists to the rescue}\label{citizen-scientists-to-the-rescue}}
 
@@
 For the English and Tsimane' corpora, vocalizations were automatically identified using LENA\textsuperscript{TM}, whereas the other two were extracted through manual segmentation.
-Vocalizations were split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received three classifications into Crying, Laughing, Canonical, Non-Canonical, or Junk (with the latter tag used for clips that did not contain a child voice).
-They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgement of canonical or non-canonical.
+Segments were then split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received three classifications into Crying, Laughing, Canonical, Non-Canonical, or Junk (with the latter tag used for clips that did not contain a child voice).
+They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgment of canonical or non-canonical.
 They found that this Canonical Proportion increased with infant age in an analysis collapsing across corpora, which provides a first proof of principle that citizen scientists' annotations could be an appropriate solution to the problem of annotating child vocalizations from wearables.
@@
 
-When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk vocalizations); and the Canonical Proportion (Canonical)/(all linguistic vocalizations).
-Previous work suggests that children with Angelman syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the Canonical Proportion was expected to increase with age (Hamrick et al., n.d.; Hamrick \& Tonnsen, n.d.).
+When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk judgments); and the Canonical Proportion (Canonical)/(all linguistic judgments).
+Previous work suggests that children with Angelman syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the Canonical Proportion was expected to increase with age (Hamrick et al., 2019; Hamrick \& Tonnsen, 2019).
 Although we could not rely on previous work to make predictions regarding the Linguistic Proportion, we reasoned that we should observe an increase in the Linguistic Proportion for both populations.
 More specific to our research aims, we checked the degree to which laboratory and citizen science annotations converged at the child level by using correlations across the two.
+
+It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels.
+We therefore included children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range, as well as a group of low-risk control infants.
+\textbf{Some additional information about Angelman syndrome would be helpful to understand the findings.
+Were the different patterns between the canonical and linguistic proportions predicted by the already-known characteristics of this syndrome?
+Has other work included this population?
+Why was this population selected?
+Are there hypotheses about the validity between groups or is this more exploratory?} \textbf{relative dearth of LENA data with atypical populations }
 
@@
 Fortunately, Seidl et al.
-(2019) determined that when segments are divided into shorter chunks (400-600ms), human annotators with little training can code our categories of interest (Canonical, Laughing, Crying, etc.) with a classification quality comparable to the one carried out on full segments.
-We therefore used very short chunks (500ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.
+(2019) determined that when segments are divided into shorter clips (400-600ms), human annotators with little training can code our categories of interest (Canonical, Laughing, Crying, etc.) with a classification quality comparable to the one carried out on full segments.
+We therefore used very short clips (500ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.
 
 Specifically, the segments extracted were automatically cut into clips of exactly 500ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform.
-To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms clips and their corresponding segment.
-Clips were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling (Bowyer, Lintott, Hines, Allan, \& Paget, 2015).
+We will call this level \emph{chunk}, to convey that they are part of a bigger unit (the LENA\textsuperscript{TM}-identified \enquote{segment}).
+To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms chunks and their corresponding segment.
+Chunks were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling (Bowyer, Lintott, Hines, Allan, \& Paget, 2015).
 
@@
 
-Users were asked to assign each 500ms clip to one out of five possible categories: (1) Canonical, (2) Non-Canonical, (3) Crying, (4) Laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
-Notice that the latter is narrower than the \enquote{Don't mark} category of the lab annotation routine, which was used any time there was noise or overlap that could affect acoustic analyses.
-
-Following recommendations from the Zooniverse board, we collected five judgments per clip, rather than three as used in the laboratory.
-Majority agreement is thus achieved when three out of five judgements agree (rather than two out of three).
-
 \begin{figure}
@@
 \end{figure}
+
+Users were asked to assign each 500ms chunk to one out of five possible categories: (1) Canonical, (2) Non-Canonical, (3) Crying, (4) Laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
+Notice that the latter is narrower than the \enquote{Don't mark} category of the lab annotation routine, which was used any time there was noise or overlap that could affect acoustic analyses.
+
+Following recommendations from the Zooniverse board, we collected five judgments per chunk, rather than three as used in the laboratory.
+Majority agreement is thus achieved when three out of five judgments agree (rather than two out of three).
 
@@
 Nearly a fifth of chunks did not have at least 3 labels in agreement out of the 5 Zooniverse labels (N = 6,585, 19\% of all chunks).
-Of the chunks without a majority agreement, 4341 (66\%) contained one or two Junk judgements (out of 5), 6523 (99,9\%) had at least two matching judgements (the threshold used for lab-annotated segments), and only 61 (0,01\%) had 5 different judgements.
+Of the chunks without a majority agreement, 4341 (66\%) contained one or two Junk judgments (out of 5), 6523 (99,9\%) had at least two matching judgments (the threshold used for lab-annotated segments), and only 61 (0,01\%) had 5 different judgments.
 Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the 81\% of chunks that did have at least 3 labels in agreement; this represented 135,725 labels for 27,145 chunks, corresponding to 11,593 LENA\textsuperscript{TM} segments.
@@
 
-We did this by considering the majority label extracted from Zooniverse judgements for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were Junk, then the segment was labelled as Junk; then, following a hierarchy (Canonical\textgreater{}Non-canonical\textgreater{}Crying\textgreater{}Laughing), segments with at least one instance of each judgement were labelled as such:
+We did this by considering the majority label extracted from Zooniverse judgments for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were Junk, then the segment was labelled as Junk; then, following a hierarchy (Canonical\textgreater{}Non-canonical\textgreater{}Crying\textgreater{}Laughing), segments with at least one instance of each judgment were labelled as such:
 
@@
 \item
-  if one Canonical judgement was present, then the segment was labelled as Canonical,
+  if one Canonical judgment was present, then the segment was labelled as Canonical,
 \item
-  else, if a Non-Canonical judgement was present, the segment was labelled as Non-Canonical,
+  else, if a Non-Canonical judgment was present, the segment was labelled as Non-Canonical,
 \item
@@
 LENA\textsuperscript{TM} segments are not equivalent to child vocalizations: A segment may contain 1 or more vocalizations, which may be of the same type or not.
+Figure 2 conveys the relationship between chunks and segments, and how judgments are combined into labels.
+
+\begin{figure}
+\centering
+\includegraphics{fig_levels.key.pdf}
+\caption{(\#fig:fig:process)Correspondance between LENA-identified segments (which can contain one or more vocalizations) and chunks.
+Notice that both segments and chunks are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority.
+To compare laboratory and Zooniverse annotations at the level of segments, chunk-level judgments are combined using the hierarchy Canonical\textgreater{}Non-canonical\textgreater{}Crying\textgreater{}Laughing.}
+\end{figure}
 
@@
 See Figure 2 for results.
+\textbf{the children with Angelman syndrome scored much lower than the typically-developing infants on the canonical measure but higher on the linguistic measure.}
 
@@
 
-Although the classification at the clip level is only moderately accurate, what we are ultimately interested in is whether citizen scientists' classifications are able to provide a reliable snapshot of childrens' individual development.
+Although the classification at the segment level is only moderately accurate, what we are ultimately interested in is whether citizen scientists' classifications are able to provide a reliable snapshot of children's individual development.
 We illustrate this association in Figure 3.
@@
 
-Above, we concluded that derived metrics seem more promising than segment-level data.
-Notice that derived metrics do not require matching of clips (500 ms presented to Zooniverse participants) to segments (the original LENA\textsuperscript{TM} segments presented to laboratory participants).
+Above, we concluded that derived metrics integrating information across audio clips (linguistic and canonical proportion, which can be derived from segment- or chunk-level judgments) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA\textsuperscript{TM} segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
 As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
@@
 
-Since all of these results are very similar to those obtained when first matching clips to segments, we conclude that in the future this step may not be necessary.
-Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' clip level judgements (which was in fact what Cychosz et al., 2019 did).
+Since all of these results are very similar to those obtained when first matching chunks to segments, we conclude that in the future this step may not be necessary.
+Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' chunk level judgments (which was in fact what Cychosz et al., 2019 did).
 
@@
 \caption{\label{tab:tab-cors}Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways.
-First, All seg indicates that Zooniverse annotations at the clip level were first combined at the segment level.
-Second, Chunks indicates that they were analyzed directly at the clip level.
+First, All seg indicates that Zooniverse annotations at the chunk level were first combined at the segment level.
+Second, Chunks indicates that they were analyzed directly at the chunk level.
 Both of these are based on all the data.
@@
 This was the case of the \enquote{Junk/Don't mark} category.
-Higher precision and low recall indicates that when Zooniverse annotations return a \enquote{Junk} judgement, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as \enquote{Don't mark} in the laboratory.
+Higher precision and low recall indicates that when Zooniverse annotations return a \enquote{Junk} judgment, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as \enquote{Don't mark} in the laboratory.
 Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the \enquote{Junk} category less) than laboratory annotators, who have access to the whole vocalization.
 This is consistent with the fact that our lab definition of \enquote{Junk} was more stringent than the Zooniverse definition.
-Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use \enquote{Don't mark} if there was \emph{any} noise or verlapping speech that would potentially affect acoustic analyses.
+Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use \enquote{Don't mark} if there was \emph{any} noise or overlapping speech that would potentially affect acoustic analyses.
 
@@
 
+\textbf{remaining methodological problems with the LENA pipeline (p.~XX);
+the relatively small sample sizes of the two groups we studied (p.~XX);
+the kinds of populations that we cannot be certain this work generalizes to (p.~XX).
+relative dearth of LENA data with atypical populations }
+
 It is also important to note that work on children with non-normative development poses in itself some challenges.
-Our data contained a lower number of vocalizations for children diagnosed with Angelman syndrome, even though we selected much younger low-risk control childrens as the comparison group.
+Our data contained a lower number of vocalizations for children diagnosed with Angelman syndrome, even though we selected much younger low-risk control children as the comparison group.
 However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
@@
 
-We are grateful to \ldots{}
-
-\newpage
-
-MOVE FIGURE 1 HERE
-
-\newpage
-
-MOVE FIGURE 2 HERE
-
-etc
+We are grateful to the families who contributed their data; to the Zooniverse volunteers that made this work possible; and the audience and technical committee of SLT 2020 for helpful feedback on preliminary analyses.
 
@@
 \leavevmode\hypertarget{ref-aust2017papaja}{}%
-Aust, F., \& Barth, M. (2017). Papaja: Create apa manuscripts with r markdown.
+Aust, F., \& Barth, M. (2017). Papaja: Create APA manuscripts with R Markdown.
 
@@
 (2017).
-A retrospective video analysis of canonical babbling and volubility in infants with fragile x syndrome at 9--12 months of age.
+A retrospective video analysis of canonical babbling and volubility in infants with Fragile X Syndrome at 9--12 months of age.
 \emph{Journal of Autism and Developmental Disorders}, \emph{47}(4), 1193--1206.
@@
 \leavevmode\hypertarget{ref-berinsky2012evaluating}{}%
-Berinsky, A. J., Huber, G. A., \& Lenz, G. S. (2012). Evaluating online labor markets for experimental research: Amazon. Com's mechanical turk. \emph{Political Analysis}, \emph{20}(3), 351--368.
+Berinsky, A. J., Huber, G. A., \& Lenz, G. S. (2012). Evaluating online labor markets for experimental research: Amazon.com's Mechanical Turk. \emph{Political Analysis}, \emph{20}(3), 351--368.
 
@@
 Panoptes, a project building tool for citizen science.
-In \emph{Proceedings of the aaai conference on human computation and crowdsourcing (hcomp'15).
-AAAI, san diego, ca, usa}.
+In \emph{Proceedings of the AAAI Conference on Human Computation and Crowdsourcing (HCOMP'15)}.
+San Diego, CA.
 
@@
 (2017).
-A new workflow for semi-automatized annotations: Tests with long-form naturalistic recordings of childrens language environments.
+A new workflow for semi-automatized annotations: Tests with long-form naturalistic recordings of children's language environments.
 In \emph{Interspeech 2017} (pp.
@@
 \leavevmode\hypertarget{ref-casillas2019step}{}%
-Casillas, M., \& Cristia, A. (2019). A step-by-step guide to collecting and analyzing long-format speech environment (lfse) recordings. \emph{Collabra}, \emph{5}(1).
+Casillas, M., \& Cristia, A. (2019). A step-by-step guide to collecting and analyzing long-format speech environment (LFSE) recordings. \emph{Collabra}, \emph{5}(1).
 
@@
 (2020).
-Accuracy of the language environment analysis system segmentation and metrics: A systematic review.
+Accuracy of the Language Environment Analysis system segmentation and metrics: A systematic review.
 \emph{Journal of Speech, Language, and Hearing Research}, \emph{63}(4), 1093--1105.
@@
 (2020).
-A thorough evaluation of the language environment analysis (lena) system.
+A thorough evaluation of the Language Environment Analysis (LENA) system.
 \emph{Behavior Research Methods}, 1--20.
@@
 (2018).
-Language environment analysis (lena) system investigation of day long recordings in children: A literature review.
+Language ENvironment analysis (LENA) system investigation of day long recordings in children: A literature review.
 \emph{Journal of Communication Disorders}, \emph{72}, 77--85.
@@
 Hamrick, L., Seidl, A., \& Kelleher, B.
-(n.d.).
+(2019).
 Capturing variability in early language skills among children with Angelman syndrome: A novel semi-automated approach.
@@
 Hamrick, L., \& Tonnsen, B.
-(n.d.).
+(2019).
 Measurement and characterization of early social communication in children with Angelman syndrome.
@@
 Sadhwani \& K.
-Okoniewski (Eds.), \emph{Advances in characterizing developmental trajectories in angelman syndrome across the lifespan, gatlinburg conference}.
+Okoniewski (Eds.), \emph{Advances in characterizing developmental trajectories in Angelman syndrome across the lifespan, Gatlinburg conference}.
 San Antonio, TX.
@@
 \leavevmode\hypertarget{ref-oller2000emergence}{}%
-Oller, D. (2000). The emergence of the capacity for speech. Erlbaum.
+Oller, D. (2000). The emergence of the capacity for speech. Mahwah, NJ: Erlbaum.
 
@@
 (2017).
-Language environment analysis (lena) in phelan-mcdermid syndrome: Validity and suggestions for use in minimally verbal children with autism spectrum disorder.
+Language ENvironment Analysis (LENA) in Phelan-McDermid syndrome: Validity and suggestions for use in minimally verbal children with Autism Spectrum Disorder.
 \emph{Journal of Autism and Developmental Disorders}, \emph{47}(6), 1605--1617.
@@
 (2019).
-The interspeech 2019 computational paralinguistics challenge: Styrian dialects, continuous sleepiness, baby sounds \& orca activity.
+The INTERSPEECH 2019 Computational Paralinguistics Challenge: Styrian Dialects, Continuous Sleepiness, Baby Sounds \& Orca Activity.
 In \emph{Interspeech} (pp.
@@
 Towards detection of canonical babbling by citizen scientists: Performance as a function of clip length.
-In (pp.
+In \emph{Interspeech} (pp.
 3579--3583).
@@
 \leavevmode\hypertarget{ref-semenzin2020zooniverse}{}%
-Semenzin, C., \& Cristia, A. (2020). Zooniverse pipeline for long format recordings.
+Semenzin, C., \& Cristia, A. (2020). Zooniverse pipeline for long format recordings. Retrieved from \url{https://github.com/LAAC-LSCP/Zooniverse}
 
@@
 (2016).
-Homebank: An online repository of daylong child-centered audio recordings.
 \emph{Seminars in Speech and Language}, \emph{37}(2), 128--142.
@@
 \leavevmode\hypertarget{ref-vandam2019use}{}%
-VanDam, M., \& Yoshinaga-Itano, C. (2019). Use of the lena autism screen with children who are deaf or hard of hearing. \emph{Medicina}, \emph{55}(8), 495.
+VanDam, M., \& Yoshinaga-Itano, C. (2019). Use of the LENA Autism Screen with children who are deaf or hard of hearing. \emph{Medicina}, \emph{55}(8), 495.
 
@@
 (2020).
-A meta-analysis of the predictability of lena™ automated measures for child language development.
+A meta-analysis of the predictability of LENA™ automated measures for child language development.
 \emph{Developmental Review}, \emph{57}, 100921.
@@
 Signal processing for young child speech language development.
-In \emph{First workshop on child, computer and interaction}.
+In \emph{First workshop on Child, Computer and Interaction}.
 
@@
 (2019).
-Using attention networks and adversarial augmentation for styrian dialect continuous sleepiness and baby sound recognition.
+Using Attention Networks and Adversarial Augmentation for Styrian Dialect, Continuous Sleepiness and Baby Sound Recognition.
 In \emph{INTERSPEECH} (pp.

### 2021-01-05 202010_jslhr/paper.Rmd
revision in progress, knits to docx

@@
 figsintext        : no
-figurelist        : no
+figurelist        : yes
 tablelist         : no
@@
 lineno            : yes
+
+always_allow_html: true
 
@@
 We will call this level *chunk*, to convey that they are part of a bigger unit (the LENA^TM^-identified "segment").
+Note that segments and chunks are collectively referred to as clips because they are parts of the longer audio recording.
 To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms chunks and their corresponding segment.
@@
 
-Following recommendations from the Zooniverse board, we collected five judgments per chunk, rather than three as used in the laboratory.
+Following recommendations from the Zooniverse board, we collected five judgments per clip, rather than three as used in the laboratory.
 Majority agreement is thus achieved when three out of five judgments agree (rather than two out of three).
@@
 
-```{r fig-process, comment=F, message=F, hide = T, warning = F, echo = F, fig.cap = "Correspondance between LENA-identified segments (which can contain one or more vocalizations) and chunks.
-Notice that both segments and chunks are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority.
+```{r fig-process, comment=F, message=F, hide = T, warning = F, echo = F, fig.cap = "Correspondence between LENA-identified segments (which can contain one or more vocalizations) and chunks.
+Segments and chunks are collectively referred to as clips because they are parts of the longer audio recording.
+Notice that clips are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority.
 To compare laboratory and Zooniverse annotations at the level of segments, chunk-level judgments are combined using the hierarchy Canonical>Non-canonical>Crying>Laughing."}
@@
 Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical+Non-Canonical)/All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical/(Canonical+Non-Canonical) (i.e., we remove Junk, Crying, and Laughing).
-See Figure 2 for results.
+See Figure 3 for results.
 **the children with Angelman syndrome scored much lower than the typically-developing infants on the canonical measure but higher on the linguistic measure.**
@@
 Next, we discuss the correspondence between citizen science classifications and the laboratory gold standard, at the level of individual segments.
-Results were visualized with a confusion matrix showing precision and recall (Figure 3): the diagonal elements show the number of correct segment-level classifications for each class while the off-diagonal elements show non-matching classifications.
+Results were visualized with a confusion matrix showing precision and recall (Figure 4): the diagonal elements show the number of correct segment-level classifications for each class while the off-diagonal elements show non-matching classifications.
 
@@
 Although the classification at the segment level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of children's individual development.
-We illustrate this association in Figure 3.
+We illustrate this association in Figure 5.
 Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
@@
 We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk".
-Results were similar across these two post-hoc subgroups (see Figure 4).
+Results were similar across these two post-hoc subgroups (see Figure 6).
 For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["lp","lower_Junk"]`) = `r myr["lp","lower_Junk"]`, CI [`r mylb["lp","lower_Junk"]`,`r myhb["lp","lower_Junk"]`], p=`r myp["lp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["lp","higher_Junk"]`) = `r myr["lp","higher_Junk"]`, CI [`r mylb["lp","higher_Junk"]`,`r myhb["lp","higher_Junk"]`], p=`r myp["lp","higher_Junk"]`.
@@
 
-Since all of these results are very similar to those obtained when first matching chunks to segments, we conclude that in the future this step may not be necessary.
-Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' chunk level judgments [which was in fact what @cychosz2019canonical did].
+Since all of these results are very similar to those obtained when collapsing chunk-level judgments to generate segment-level judgments, we conclude that in the future this step may not be necessary.
+Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' chunk-level judgments [which was in fact what @cychosz2019canonical did].
 
@@
 Given that the two groups of children differ on both of these features (age and diagnosis), we cannot empirically tease apart these two differences.
-Nonetheless, inspection of Figure 2 suggests different interpretations for the two derived metrics.
+Nonetheless, inspection of Figure 3 suggests different interpretations for the two derived metrics.
 Overall, the pattern for Linguistic Proportion is compatible with a non-linear trajectory common to the two groups, whereby the Linguistic Proportion increases rapidly up to 20 months, and hovers around 90-100% thereafter.

### 2021-01-05 202010_jslhr/paper.Rmd
finished first pass

@@
 class             : "man"
-output            : papaja::apa6_word
+output            : papaja::apa6_pdf
 
@@
 
-```{r tab-LENA-challenge, comment = F, message = F, hide = T, warning = F, echo = F, results = "asis"}
-mycaption="Comparison of recall percentages obtained with the baseline algorithm created by the ComParE team, in the context of the ComParE 2019 BabySounds subchallenge (Schuller et al., 2019), LENATM labels (note LENATM does not distinguish between Canonical and Non-canonical), and those obtained in this study through Zooniverse annotations.
-Label frequency indicates the prevalence of the relevant label (Crying, Laughing, etc.) in each dataset (the same Lab dataset was used for both LENATM and Zooniverse).
-UAR stands for unweighted average recall, WAR for weighted average recall (which takes into account label frequency)."
-
-read.table("tab-lena-challenge.txt",header=F,sep="\t",skip=1)->tab1
-#olnames(tab1)<-c("","Recall","Recall","Recall","Label Frequency","Label Frequency")
-
-# as_hux(tab1) %>% 
-#     insert_row("","Recall","","","Label Frequency","", fill = "") %>% 
-#     merge_cells(1, 2:4) %>% 
-#     merge_cells(1, 5:6) %>%  
-#     merge_cells(2, 5:6) %>%  
-#     set_header_rows(1, TRUE) %>% 
-#    style_headers(bold = TRUE)
-
-kable(tab1,cap=mycaption, col.names = NULL) %>%
-  collapse_rows(columns = 1:6, valign = "middle") #%>%
-#add_header_above(c(" " = 1, "Recall"= 3, "Label Frequency" = 2))
-```
+----- INSERT TABLE 1 HERE -----
 
@@
 
-``` {r fig-zoo, comment=F, message=F, hide = T, warning = F, echo = F,  fig.cap="Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needed to make; as well as the Field Guide, which remained accessible on the right of the screen throughout the session, and which allowed users to revise examples of the different categories."}
-knitr::include_graphics("zooniverse-pufig.pdf")
-```
+----- INSERT FIGURE 1 HERE -----
 
@@
 
-```{r fig-process, comment=F, message=F, hide = T, warning = F, echo = F, fig.cap = "Correspondence between LENA-identified segments (which can contain one or more vocalizations) and chunks.
-Segments and chunks are collectively referred to as clips because they are parts of the longer audio recording.
-Notice that clips are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority.
-To compare laboratory and Zooniverse annotations at the level of segments, chunk-level judgments are combined using the hierarchy Canonical>Non-canonical>Crying>Laughing."}
-knitr::include_graphics("fig_levels.key.pdf")
-
-```
+----- INSERT FIGURE 2 HERE -----
 
@@
 
+----- INSERT FIGURE 3 HERE -----
+
+Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups.
+There was a near-zero relationship among the children diagnosed with Angelman syndrome r(`r myn["l_lp","AS"]`) = `r myr["l_lp","AS"]`, CI [`r mylb["l_lp","AS"]`,`r myhb["l_lp","AS"]`], p=`r myp["l_lp","AS"]`]; and a significant association among low-risk control children r(`r myn["l_lp","LR"]`) = `r myr["l_lp","LR"]`, CI [`r mylb["l_lp","LR"]`,`r myhb["l_lp","LR"]`], p=`r myp["l_lp","LR"]`].
+The Canonical Proportion exhibited non-significant developmental decreases among children diagnosed with Angelman syndrome r(`r myn["l_cp","AS"]`) = `r myr["l_cp","AS"]`, CI [`r mylb["l_cp","AS"]`,`r myhb["l_cp","AS"]`], p=`r myp["l_cp","AS"]`]; and marginal developmental increases among low-risk control r(`r myn["l_cp","LR"]`) = `r myr["l_cp","LR"]`, CI [`r mylb["l_cp","LR"]`,`r myhb["l_cp","LR"]`], p=`r myp["l_cp","LR"]`].
+
+Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(`r myn["z_lp","AS"]`) = `r myr["z_lp","AS"]`, CI [`r mylb["z_lp","AS"]`,`r myhb["z_lp","AS"]`], p=`r myp["z_lp","AS"]`]; whereas low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r myr["z_lp","LR"]`, CI [`r mylb["z_lp","LR"]`,`r myhb["z_lp","LR"]`], p=`r myp["z_lp","LR"]`].
+Similarly, there were non-significant developmental decreases in the Canonical Proportion among children with Angelman syndrome r(`r myn["z_cp","AS"]`) = `r myr["z_cp","AS"]`, CI [`r mylb["z_cp","AS"]`,`r myhb["z_cp","AS"]`], p=`r myp["z_cp","AS"]`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r myr["z_cp","LR"]`, CI [`r mylb["z_cp","LR"]`,`r myhb["z_cp","LR"]`], p=`r myp["z_cp","LR"]`].
+
+## Main analyses
+
+Next, we discuss the correspondence between citizen science classifications and the laboratory gold standard, at the level of individual segments.
+Results were visualized with a confusion matrix showing precision and recall (Figure 4): the diagonal elements show the number of correct segment-level classifications for each class while the off-diagonal elements show non-matching classifications.
+
+```{r mygac-conf, echo=F}
+mygac=gac(data = sample_data, kat = 5, weight = c("unweighted"),
+    conf.level = 0.95)
+
+mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
+conf_tab=mycf$table
+# this package uses sensitivity & specificity
+#Sensitivity=recall
+#Specificity=precision
+#mycf
+```
+
+----- INSERT FIGURE 4 HERE -----
+
+These visualizations suggest that performance is moderate to good, which was confirmed via statistical analyses.
+The overall (weighted) accuracy is `r round(mycf$overall["Accuracy"]*100)`%, CI = [`r round(mycf$overall["AccuracyLower"]*100)`,`r round(mycf$overall["AccuracyUpper"]*100)`],  kappa is `r round(mycf$overall["Kappa"],3)`, and the Gwet's AC1 coefficient is `r round(mygac$est,3)`, CI = [`r round(mygac$lb,3)`,`r round(mygac$ub,3)`].
+
+### Child level descriptors
+
+----- INSERT FIGURE 5 HERE -----
+
+Although the classification at the segment level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of children's individual development.
+We illustrate this association in Figure 5.
+Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
+
+Similarly, a strong positive correlation is found in the Canonical Proportion r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`].
+When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
+
+## Additional analyses
+
+In this section, we report on explorations of under what conditions Zooniverse judgments more closely aligned with laboratory judgments.
+In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as "Junk" (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora [@cychosz2019canonical].
+A high proportion of "Junk" may indicate that automated segmentation was errorful for those children, and may be a sign that the rest of the data could be compromised as well.
+
+```{r prelim-Junk, echo=F}
+JunkXdiag=t.test(proportions$z_Junk ~ proportions$Diagnosis)
+
+```
+
+----- INSERT FIGURE 6 HERE -----
+
+We investigated this hypothesis by calculating the proportion of their data labeled as "Junk" for each individual child.
+There was no significant difference in the proportions of their data labeled as "Junk" for children with Angelman syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the low-risk group (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, p = `r round(JunkXdiag$p.value,3)`.
+We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk".
+Results were similar across these two post-hoc subgroups (see Figure 6).
+For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["lp","lower_Junk"]`) = `r myr["lp","lower_Junk"]`, CI [`r mylb["lp","lower_Junk"]`,`r myhb["lp","lower_Junk"]`], p=`r myp["lp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["lp","higher_Junk"]`) = `r myr["lp","higher_Junk"]`, CI [`r mylb["lp","higher_Junk"]`,`r myhb["lp","higher_Junk"]`], p=`r myp["lp","higher_Junk"]`.
+For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["cp","lower_Junk"]`) = `r myr["cp","lower_Junk"]`, CI [`r mylb["cp","lower_Junk"]`,`r myhb["cp","lower_Junk"]`], p=`r myp["cp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["cp","higher_Junk"]`) = `r myr["cp","higher_Junk"]`, CI [`r mylb["cp","higher_Junk"]`,`r myhb["cp","higher_Junk"]`], p=`r myp["cp","higher_Junk"]`.
+Thus, it does not seem that a higher proportion of "Junk" judgments is an index of low quality data.
+
+Above, we concluded that derived metrics integrating information across audio clips (linguistic and canonical proportion, which can be derived from segment- or chunk-level judgments) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA^TM^ segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
+As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
+We therefore repeated our analyses but instead of deriving our proportions for the Zooniverse data from the segment-level composite, we did it based on the individual chunk-level annotations.
+To facilitate comparison, correlations across laboratory- and Zooniverse-derived proportions are provided in Table 2.
+
+----- INSERT TABLE 2 HERE -----
+
+```{r bychunk-age, echo=F}
+read.csv("zoo_chunk_majority_agreement.csv")->chunk
+bychild=table(chunk$Answer,chunk$ChildID)
+cans=bychild["Canonical",]
+noncans=bychild["Non-Canonical",]
+Junk=bychild["Junk",]
+sum_bychild=table(chunk$ChildID)
+
+ling_prop=(cans+noncans)/(sum_bychild-Junk)
+can_prop=cans/(cans+noncans)
+
+#check that they match before merging
+#demo_data$ChildID==names(ling_prop)
+#demo_data$ChildID==names(can_prop)
+
+props_bychild_chunk=cbind(ling_prop,can_prop,demo_data)
+colnames(props_bychild_chunk)[1:2]<-c("chi","ling_prop")
+
+z_lp_as=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
+
+z_lp_lr=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
+
+z_cp_as=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
+
+z_cp_lr=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
+```
+
+Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman syndrome group: r(`r z_lp_as$parameter`) = `r z_lp_as$estimate`, CI [`r z_lp_as$conf.int[1]`,`r z_lp_as$conf.int[2]`], p=`r z_lp_as$p.value`], but increased with age  in the low-risk control group: r(`r z_lp_lr$parameter`) = `r z_lp_lr$estimate`, CI [`r z_lp_lr$conf.int[1]`,`r z_lp_lr$conf.int[2]`], p=`r z_lp_lr$p.value`.
+Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman syndrome r(`r z_cp_as$parameter`) = `r z_cp_as$estimate`, CI [`r z_cp_as$conf.int[1]`,`r z_cp_as$conf.int[2]`], p=`r z_cp_as$p.value`]; and developmental increases among low-risk control r(`r z_cp_lr$parameter`) = `r z_cp_lr$estimate`, CI [`r z_cp_lr$conf.int[1]`,`r z_cp_lr$conf.int[2]`], p=`r z_cp_lr$p.value`.
+
+```{r zoo-lab-chunk, echo=F}
+
+#create matrix to hold r results
+myr=matrix(NA,nrow=2,ncol=3)
+rownames(myr)<-c("lp","cp")
+colnames(myr)<-c("AS","LR","all")
+mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
+
+proportions=proportions[order(proportions$ChildID),]
+
+  #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
+    myr[1,"AS"]=round(thiscor$estimate,3)
+    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
+    myn[1,"AS"]=round(thiscor$parameter,3)
+    myp[1,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"])
+    myr[1,"LR"]=round(thiscor$estimate,3)
+    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
+    myn[1,"LR"]=round(thiscor$parameter,3)
+    myp[1,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(props_bychild_chunk$ling_prop,proportions$l_lp)
+    myr[1,"all"]=round(thiscor$estimate,3)
+    mylb[1,"all"]=round(thiscor$conf.int[1],3)
+    myhb[1,"all"]=round(thiscor$conf.int[2],3)
+    myn[1,"all"]=round(thiscor$parameter,3)
+    myp[1,"all"]=round(thiscor$p.value,3)
+
+    # repeat whole for cp
+
+  #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
+    myr[2,"AS"]=round(thiscor$estimate,3)
+    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
+    myn[2,"AS"]=round(thiscor$parameter,3)
+    myp[2,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"])
+    myr[2,"LR"]=round(thiscor$estimate,3)
+    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
+    myn[2,"LR"]=round(thiscor$parameter,3)
+    myp[2,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(props_bychild_chunk$can_prop,proportions$l_cp)
+    myr[2,"all"]=round(thiscor$estimate,3)
+    mylb[2,"all"]=round(thiscor$conf.int[1],3)
+    myhb[2,"all"]=round(thiscor$conf.int[2],3)
+    myn[2,"all"]=round(thiscor$parameter,3)
+    myp[2,"all"]=round(thiscor$p.value,3)
+
+ mycortab=rbind(mycortab,t(myr))
+rownames(mycortab)[4:6]<-paste("Chunks",rownames(t(myr)))
+
+```
+
+As for correlations between Zooniverse- and laboratory-derived metrics (e.g., linguistic proportion as estimated using Zooniverse coding, versus laboratory coding), we observe very similar levels of correlation: Linguistic Proportion overall r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
+As for Canonical Proportion overall r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`; Angelman syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`];  low-risk control infants r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
+
+Since all of these results are very similar to those obtained when collapsing chunk-level judgments to generate segment-level judgments, we conclude that in the future this step may not be necessary.
+Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' chunk-level judgments [which was in fact what @cychosz2019canonical did].
+
+```{r downsample}
+
+#we start from the data base that has all the info
+nsegs=data.frame(table(data_all$ChildID))
+colnames(nsegs)<-c("ChildID","nsegs")
+
+merge(nsegs,demo_data)->nsegs
+nseg_t=t.test(nsegs$nsegs~nsegs$Diagnosis)
+
+resampling=matrix(NA,nrow=50,ncol=6)
+colnames(resampling)<-c("lp_r_all","lp_r_as","lp_r_lr","cp_r_all","cp_r_as","cp_r_lr")
+for(i in 1:50){
+  sel_seg=NULL
+  for(thischi in levels(data_all$ChildID)) sel_seg<-c(sel_seg,as.character(data_all$segmentId_DB[sample(data_all$segmentId_DB[data_all$ChildID==thischi],100)]))
+
+  sampling=data_all[as.character(data_all$segmentId_DB) %in% as.character(sel_seg),]
+
+  #get the ns by child, then calculate the linguistic proportion & canonical proportion, separately for zooniverse & lab
+  ztab=table(sampling$ChildID,sampling$Zoon_classif)
+  z_lp=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
+  z_cp=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
+
+  ltab=table(sampling$ChildID,sampling$lab)
+  l_lp=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
+  l_cp=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
+
+  thistab=cbind(demo_data,z_lp,z_cp,l_lp,l_cp)
+
+  resampling[i,"lp_r_all"]=cor.test(thistab$z_lp,thistab$l_lp)$estimate
+  resampling[i,"lp_r_as"]=cor.test(thistab$z_lp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
+  resampling[i,"lp_r_lr"]=cor.test(thistab$z_lp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate
+
+   resampling[i,"cp_r_all"]=cor.test(thistab$z_cp,thistab$l_cp)$estimate
+  resampling[i,"cp_r_as"]=cor.test(thistab$z_cp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
+  resampling[i,"cp_r_lr"]=cor.test(thistab$z_cp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate
+ }
+myr=colMeans(resampling)
+mycortab=rbind(mycortab,cbind(myr[c("lp_r_as" ,"lp_r_lr","lp_r_all")],myr[c("cp_r_as" ,"cp_r_lr","cp_r_all")]))
+rownames(mycortab)[7:9]<-c("100 seg AS","100 seg LR","100 seg all")
+round(mycortab,3)->mycortab
+
+resampling=data.frame(resampling)
+```
+
+Next, we looked at whether having more segments from each child may lead to more reliable metrics.
+We observed a non-significant trend [t(`r nseg_t$parameter`)=`r nseg_t$estimate`] for lower number of segments in the Angelman syndrome group (mean `r nseg_t$estimate[1]`, range `r min(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`) than among low-risk control infants (mean `r nseg_t$estimate[2]`, range `r min(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`), likely because of the lower volubility of the former children.
+Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications [@cychosz2019canonical].
+We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children [following @cychosz2019canonical].
+We randomly sampled 100 segments from each child's data, and recalculated the Linguistic and Canonical Proportions based only on these associated judgments.
+We repeated this process 50 times, to assess the extent to which the association between laboratory- and Zooniverse-derived metrics varied in each random sample (e.g., the correlation between Linguistic Proportion as estimated using Zooniverse coding on the one hand, and Linguistic Proportion as estimated using laboratory coding on the other).
+The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table 2), particularly for the low-risk younger infants, and for Linguistic Proportion.
+This may indicate that, especially in some groups of infants, 100 segments may not be sufficient to capture a stable estimate of the child's Linguistic and Canonical Proportions.
+We also observed quite a bit of variance across the 50 runs.
+For Linguistic Proportion, the range of correlations found for all infants was [`r min(resampling$lp_r_all)`,`r max(resampling$lp_r_all)`]; for the Angelman syndrome group [`r min(resampling$lp_r_as)`,`r max(resampling$lp_r_as)`]; for the low-risk group [`r min(resampling$lp_r_lr)`,`r max(resampling$lp_r_lr)`].
+For Canonical Proportion, the range of correlations found for all infants was [`r min(resampling$cp_r_all)`,`r max(resampling$cp_r_all)`]; for the Angelman syndrome group [`r min(resampling$cp_r_as)`,`r max(resampling$cp_r_as)`]; for the low-risk group [`r min(resampling$cp_r_lr)`,`r max(resampling$cp_r_lr)`].
+
+<!-- - look at whether kids who have a high "Junk" proportion according to Zooniverse  -->
+<!--      - have lower levels of concordance across Zoo coders -->
+
+<!-- - take all 5 Zoo judgments or 3 or 1 (may need to run the latter two a couple of times to check for potential variance -- if variance is small, we'll just say so, if it's large then we'll have to write a loop and report CIs) -->
+
+## Discussion
+
+In the present study we assessed the extent to which child vocalizations from LENA^TM^ daylong recordings can be accurately described based on classifications collected using the citizen science platform Zooniverse.
+Our recordings came from children diagnosed with a genetic syndrome associated with severe language disorder, as well as low-risk controls, and included a range of different ages.
+Excerpts of these recordings had previously been annotated by trained experts in the lab, and served as the "gold standard".
+
+In descriptive terms, we found that category frequency was similar across laboratory and citizen science annotations.
+Correlations between age and our two child-level metrics revealed different developmental patterns between our sample of older children with a diagnosis of Angelman syndrome and younger low-risk controls.
+Given that the two groups of children differ on both of these features (age and diagnosis), we cannot empirically tease apart these two differences.
+Nonetheless, inspection of Figure 3 suggests different interpretations for the two derived metrics.
+Overall, the pattern for Linguistic Proportion is compatible with a non-linear trajectory common to the two groups, whereby the Linguistic Proportion increases rapidly up to 20 months, and hovers around 90-100% thereafter.
+In this case, the lack of correlation with age in the Angelman syndrome group suggests that there is little systematic variance in this metric after 20 months of age.
+In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and relatively stable levels of Canonical Proportion among the children in the Angelman syndrome group.
+
+Next, we looked at agreement between laboratory and Zooniverse annotation at the level of individual segments.
+We found that some categories had high precision and recall -- notably Non-canonical segments were detected quite accurately, and to a lesser extent the same was true for Canonical segments.
+This is interesting because it suggests that Zooniverse annotations may be trusted particularly for detecting these two types of speech-like segments.
+
+In other cases, precision was fairly high but recall was lower.
+This was the case of the "Junk/Don't mark" category.
+Higher precision and low recall indicates that when Zooniverse annotations return a "Junk" judgment, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as "Don't mark" in the laboratory.
+Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the "Junk" category less) than laboratory annotators, who have access to the whole vocalization.
+This is consistent with the fact that our lab definition of "Junk" was more stringent than the Zooniverse definition.
+Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use "Don't mark" if there was *any* noise or overlapping speech that would potentially affect acoustic analyses.
+
+Both Crying and Laughing showed moderate to good recall, but low precision.
+This means that a high proportion of segments detected as these two categories in the lab can be classified in the appropriate category by Zooniverse annotators, but Zooniverse annotators also put into these categories many segments that lab annotators classified in a different manner.
+Most saliently, there is a problematic level of confusion between Crying and Non-Canonical, whereby lab-detected Crying was almost as likely to be classified as such or Non-canonical in Zooniverse (the recall proportion was 49 and 47% respectively), and even worse, 55% of the segments Zooniverse annotators classified as Crying were actually Non-Canonical according to lab annotations.
+This is likely due to the fact that lab annotators had access to the whole segment, and could even listen to the context of that segment, so they could accurately interpret a segment as crying.
+Thus, it is possible that additional analyses, using for instance pitch or duration characteristics at the level of the whole segment rather than the chunk could improve classification performance, but for the time being, we cannot recommend Zooniverse annotations to identify crying bouts.
+
+As for Laughing, the recall is quite good but precision is low, with confusion involving "Junk" and Non-Canonical to similar extents.
+We are tentative in our interpretation of this result because the category was very sparsely represented, with barely 200 segments classified as Laughing in the whole data set.
+We thus recommend additional work, although we expect that perhaps there will be a similar problem (and solution) as indicated for Crying, whereby determination of Laughing versus Non-Canonical will require inspection of segment-level properties.
+
+Turning now to the child-level analyses, we were particularly interested in derived metrics because this is more typically the goal of such analyses, for instance to study potential individual or group differences, or investigate the impact of an intervention.
+Our results show that, despite agreement variation at the level of the segment, derived metrics were highly correlated across laboratory and Zooniverse data, meaning that variability in precision and recall did not substantially impact key outcome variables.
+Errors only seemed systematic for the Linguistic Proportion (and not for Canonical Proportion), and they were due to lower Linguistic Proportions from Zooniverse than laboratory data.
+This is due to the fact that a substantial number of Crying segments were classified as Non-Canonical by Zooniverse annotators.
+Nonetheless, since these errors were systematic across infants, they do not affect the study of group or individual variation.
+
+Our exploratory analyses then dug into some methodological aspects of our data.
+We found that, although individuals vary widely in terms of what proportion of their data is classified as Junk by Zooniverse annotators, this is not a sign that their data as a whole is compromised, with similar levels of correlations between laboratory- and Zooniverse-derived metrics across median groups based on proportion of their data classified as Junk.
+
+Next, we found that more data led to more reliable measurements: Specifically, our correlations based on 300-500 segments per child were markedly higher than those based on 100 segments per child, particularly when looking at our low-risk, younger participants.
+It is important to note that an advantage of Zooniverse, compared to laboratory annotations, is that it is easier to scale up: More data from each individual child can be annotated this way, which will help countering noise.
+
+In sum, we believe these data show that young children's vocalizations can be accurately described by relying on the judgments of Zooniverse's citizen scientists, particularly when the goal is to describe vocalizations at a broader level than individual chunks or segments.
+It is important to note that from a methodological point of view, the use of crowdsourcing is not merely a question of being an ecological technique, but it can also offer benefits in the area of scientific rigor and reproducibility.
+As previously mentioned in the Introduction section, Zooniverse offers both an automated system of data handling as well as an enormous pool of users: access to a large-scale group of blinded participants can in itself help researchers overcome the methodological compromises that sometimes lead to biased results.
+Additionally, crowd-sourcing and open-science frameworks can come together to encourage replication of existing work: open-source software, shared data and citizen science platforms would make it possible, and actually easy, to run identical studies and evaluate research results independently.
+
+### Further research directions
+
+These findings open up a new, exciting avenue for future research.
+Our pipeline to upload data on Zooniverse is scalable and quickly adapted to new datasets, including data on infants of different ages, learning different languages, or living in environments with different acoustic properties (e.g., rural versus urban environments).
+This means large existing datasets of daylong recordings, notably those hosted on Homebank [@vandam2016homebank], can be reliably annotated in an ecological way using citizen science.
+
+Our results warrant additional research on two metrics we used to describe children’s linguistic productions: Linguistic Proportion and Canonical Proportion.
+Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion.
+We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant for other disorders that are related to emotional disorders [see a review in @halpern2016excessive].
+Additionally, while Canonical Proportion has been studied in much previous work, we believe further research is necessary using larger scales (saliently increasing the sheer number of children studied, to have sufficient power to detect group effects) and with broader coverage [notably including children growing up in a broad range of languages and cultures, as in @cychosz2019canonical], to make sure that assumptions regarding cross-linguistic and maturational universality are amply justified and not merely a result of having low power to detect differences, or simply sampling from similar populations.
+
+Assuming these child-level metrics continue to be supported by such additional data, it would be interesting for future investigations to expand the analysis of Linguistic Proportion and Canonical Proportion to recover differences and similarities in language development between low-risk children and children with disorders other than Angelman syndrome.
+That is, it is at present unclear whether relatively lower Canonical Proportions will also be found in other populations, or whether this is more specific to Angelman syndrome.
+For instance, some previous work has found differences in babbling even when comparing late talkers with toddlers with no specific diagnosis  [@fasolo2008babbling].
+We thus encourage further research inspecting Canonical Proportion in other populations, as well as methodological work that attempts to generate metrics that are similarly easy to gather at scale.
+
+**remaining methodological problems with the LENA pipeline (p. XX);
+the relatively small sample sizes of the two groups we studied (p. XX);
+the kinds of populations that we cannot be certain this work generalizes to (p. XX).
+relative dearth of LENA data with atypical populations **
+
+It is also important to note that work on children with non-normative development poses in itself some challenges.
+Our data contained a lower number of vocalizations for children diagnosed with Angelman syndrome, even though we selected much younger low-risk control children as the comparison group.
+However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
+Now that this method has been validated, researchers can extract all segments that LENA^TM^ identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
+
+Another open question we hope future work looks into is the assessment of within-participant changes.
+Our dataset only contained one daylong recording per child (with one exception), so other datasets in which a greater number of recordings are collected longitudinally can bring novel insights.
+Is the age-related decline we observed among children in the Angelman syndrome group truly a developmental decline that happens by necessity in this population, or might it reflect partially other aspects of children's experience and behavior?
+Additionally, longitudinal work within the same infant may be needed to track the "natural history" of vocalizations to better contextualize how these metrics should be interpreted.
+It is also an open question how large must effects be, at the individual level, in order to be detected with this method.
+For example, there is interest in using naturalistic recordings as outcome measures for clinical (behavioral or pharmacological) trials.
+The fact that agreement across annotators is high even in the data coming from children diagnosed with Angelman syndrome is encouraging, but it would be important to establish whether absolute levels of Canonical Proportion, and/or rates of change over longer time spans, may be good candidates for tracking the effects of any treatment.
+
+Finally, broader generalizations of our technique could be of interest to readers of this paper.
+Any researcher is able to undergo the procedure to create Zooniverse projects (https://www.zooniverse.org/about/faq).
+For instance, a project could be set up to annotate potentially disordered speech by older adults at risk of, or diagnosed with, neurodegenerative diseases.
+Although it would be important to similarly validate such annotations, we believe there is strong promise for such tasks, because many neurodegenerative disorders affect local aspects of the speech, which are detected at the syllable level.
+In contrast, tasks that require more speech or audio context may not be well suited to citizen science platforms, because they would require playing longer clips, which may reveal identifying or sensitive information.
+Although citizen scientists themselves are well-intentioned, the platform could be used by others who want to exploit the system for other means.
+
+## Conclusion
+
+In this study we validated the quality of annotations obtained through a citizen science platform, Zooniverse, as compared to a gold standard of human expert annotators.
+We analyzed the correspondence between annotations at the individual segment level, and the individual-child level, using Canonical Proportion and Linguistic Proportion as descriptors.
+We found moderate to good accuracy in the first, and strong positive correlations in the latter.
+We can conclude that citizen scientists are a reliable source of fast and ecological annotation of speech data, particularly when results are combined into child-level descriptors.
+The same methodology may be applied to several research questions in the study of language acquisition and language disorders.
+This finding is particularly welcome in an era when wearables open new avenues for studying human behavior and development in an ecological manner.
+
+\newpage
+
+# Acknowledgements
+
+We are grateful to the families who contributed their data; to the Zooniverse volunteers that made this work possible; and the audience and technical committee of SLT 2020 for helpful feedback on preliminary analyses.
+
+# References
+
+<!-- These lines ensure references are set with hanging indents in PDF documents; they are ignored in Word. -->
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+
+``` {r fig-zoo, comment=F, message=F, hide = T, warning = F, echo = F,  fig.cap="Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needed to make; as well as the Field Guide, which remained accessible on the right of the screen throughout the session, and which allowed users to revise examples of the different categories."}
+knitr::include_graphics("zooniverse-pufig.pdf")
+```
+
+```{r fig-process, comment=F, message=F, hide = T, warning = F, echo = F, fig.cap = "Correspondence between LENA-identified segments (which can contain one or more vocalizations) and chunks.
+Segments and chunks are collectively referred to as clips because they are parts of the longer audio recording.
+Notice that clips are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority.
+To compare laboratory and Zooniverse annotations at the level of segments, chunk-level judgments are combined using the hierarchy Canonical>Non-canonical>Crying>Laughing."}
+knitr::include_graphics("fig_levels.key.pdf")
+
+```
+
 ```{r fig-corage, echo=F,fig.cap="Correlations between child-level descriptors and age as a function of metric (Linguistic Proportion in the top row, Canonical Proportion in the bottom row), annotation method, and child group (red=low-risk, and black= Angelman syndrome)."}
@@
 #dev.off()
-```
-
-Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups.
-There was a near-zero relationship among the children diagnosed with Angelman syndrome r(`r myn["l_lp","AS"]`) = `r myr["l_lp","AS"]`, CI [`r mylb["l_lp","AS"]`,`r myhb["l_lp","AS"]`], p=`r myp["l_lp","AS"]`]; and a significant association among low-risk control children r(`r myn["l_lp","LR"]`) = `r myr["l_lp","LR"]`, CI [`r mylb["l_lp","LR"]`,`r myhb["l_lp","LR"]`], p=`r myp["l_lp","LR"]`].
-The Canonical Proportion exhibited non-significant developmental decreases among children diagnosed with Angelman syndrome r(`r myn["l_cp","AS"]`) = `r myr["l_cp","AS"]`, CI [`r mylb["l_cp","AS"]`,`r myhb["l_cp","AS"]`], p=`r myp["l_cp","AS"]`]; and marginal developmental increases among low-risk control r(`r myn["l_cp","LR"]`) = `r myr["l_cp","LR"]`, CI [`r mylb["l_cp","LR"]`,`r myhb["l_cp","LR"]`], p=`r myp["l_cp","LR"]`].
-
-Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(`r myn["z_lp","AS"]`) = `r myr["z_lp","AS"]`, CI [`r mylb["z_lp","AS"]`,`r myhb["z_lp","AS"]`], p=`r myp["z_lp","AS"]`]; whereas low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r myr["z_lp","LR"]`, CI [`r mylb["z_lp","LR"]`,`r myhb["z_lp","LR"]`], p=`r myp["z_lp","LR"]`].
-Similarly, there were non-significant developmental decreases in the Canonical Proportion among children with Angelman syndrome r(`r myn["z_cp","AS"]`) = `r myr["z_cp","AS"]`, CI [`r mylb["z_cp","AS"]`,`r myhb["z_cp","AS"]`], p=`r myp["z_cp","AS"]`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r myr["z_cp","LR"]`, CI [`r mylb["z_cp","LR"]`,`r myhb["z_cp","LR"]`], p=`r myp["z_cp","LR"]`].
-
-## Main analyses
-
-Next, we discuss the correspondence between citizen science classifications and the laboratory gold standard, at the level of individual segments.
-Results were visualized with a confusion matrix showing precision and recall (Figure 4): the diagonal elements show the number of correct segment-level classifications for each class while the off-diagonal elements show non-matching classifications.
-
-```{r mygac-conf, echo=F}
-mygac=gac(data = sample_data, kat = 5, weight = c("unweighted"),
-    conf.level = 0.95)
-
-mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
-conf_tab=mycf$table
-# this package uses sensitivity & specificity
-#Sensitivity=recall
-#Specificity=precision
-#mycf
 ```
@@
 
-These visualizations suggest that performance is moderate to good, which was confirmed via statistical analyses.
-The overall (weighted) accuracy is `r round(mycf$overall["Accuracy"]*100)`%, CI = [`r round(mycf$overall["AccuracyLower"]*100)`,`r round(mycf$overall["AccuracyUpper"]*100)`],  kappa is `r round(mycf$overall["Kappa"],3)`, and the Gwet's AC1 coefficient is `r round(mygac$est,3)`, CI = [`r round(mygac$lb,3)`,`r round(mygac$ub,3)`].
-
-### Child level descriptors
-
 ```{r fig-corlab-zoo, echo=F,fig.cap="Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of child group (red=low-risk, and black=Angelman syndrome)."}
@@
 
-Although the classification at the segment level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of children's individual development.
-We illustrate this association in Figure 5.
-Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
-
-Similarly, a strong positive correlation is found in the Canonical Proportion r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`].
-When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
-
-## Additional analyses
-
-In this section, we report on explorations of under what conditions Zooniverse judgments more closely aligned with laboratory judgments.
-In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as "Junk" (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora [@cychosz2019canonical].
-A high proportion of "Junk" may indicate that automated segmentation was errorful for those children, and may be a sign that the rest of the data could be compromised as well.
-
-```{r prelim-Junk, echo=F}
-JunkXdiag=t.test(proportions$z_Junk ~ proportions$Diagnosis)
-
-```
-
 ```{r fig-cor-Junk, echo=F,fig.cap="Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of whether children's Junk proportion was higher or lower than the median for all children (blue=lower, and green=higher)."}
@@
 
-We investigated this hypothesis by calculating the proportion of their data labeled as "Junk" for each individual child.
-There was no significant difference in the proportions of their data labeled as "Junk" for children with Angelman syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the low-risk group (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, p = `r round(JunkXdiag$p.value,3)`.
-We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk".
-Results were similar across these two post-hoc subgroups (see Figure 6).
-For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["lp","lower_Junk"]`) = `r myr["lp","lower_Junk"]`, CI [`r mylb["lp","lower_Junk"]`,`r myhb["lp","lower_Junk"]`], p=`r myp["lp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["lp","higher_Junk"]`) = `r myr["lp","higher_Junk"]`, CI [`r mylb["lp","higher_Junk"]`,`r myhb["lp","higher_Junk"]`], p=`r myp["lp","higher_Junk"]`.
-For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["cp","lower_Junk"]`) = `r myr["cp","lower_Junk"]`, CI [`r mylb["cp","lower_Junk"]`,`r myhb["cp","lower_Junk"]`], p=`r myp["cp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["cp","higher_Junk"]`) = `r myr["cp","higher_Junk"]`, CI [`r mylb["cp","higher_Junk"]`,`r myhb["cp","higher_Junk"]`], p=`r myp["cp","higher_Junk"]`.
-Thus, it does not seem that a higher proportion of "Junk" judgments is an index of low quality data.
-
-Above, we concluded that derived metrics integrating information across audio clips (linguistic and canonical proportion, which can be derived from segment- or chunk-level judgments) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA^TM^ segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
-As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
-We therefore repeated our analyses but instead of deriving our proportions for the Zooniverse data from the segment-level composite, we did it based on the individual chunk-level annotations.
-
-```{r bychunk-age, echo=F}
-read.csv("zoo_chunk_majority_agreement.csv")->chunk
-bychild=table(chunk$Answer,chunk$ChildID)
-cans=bychild["Canonical",]
-noncans=bychild["Non-Canonical",]
-Junk=bychild["Junk",]
-sum_bychild=table(chunk$ChildID)
-
-ling_prop=(cans+noncans)/(sum_bychild-Junk)
-can_prop=cans/(cans+noncans)
-
-#check that they match before merging
-#demo_data$ChildID==names(ling_prop)
-#demo_data$ChildID==names(can_prop)
-
-props_bychild_chunk=cbind(ling_prop,can_prop,demo_data)
-colnames(props_bychild_chunk)[1:2]<-c("chi","ling_prop")
-
-z_lp_as=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
-
-z_lp_lr=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
-
-z_cp_as=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
-
-z_cp_lr=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
-```
-
-Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman syndrome group: r(`r z_lp_as$parameter`) = `r z_lp_as$estimate`, CI [`r z_lp_as$conf.int[1]`,`r z_lp_as$conf.int[2]`], p=`r z_lp_as$p.value`], but increased with age  in the low-risk control group: r(`r z_lp_lr$parameter`) = `r z_lp_lr$estimate`, CI [`r z_lp_lr$conf.int[1]`,`r z_lp_lr$conf.int[2]`], p=`r z_lp_lr$p.value`.
-Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman syndrome r(`r z_cp_as$parameter`) = `r z_cp_as$estimate`, CI [`r z_cp_as$conf.int[1]`,`r z_cp_as$conf.int[2]`], p=`r z_cp_as$p.value`]; and developmental increases among low-risk control r(`r z_cp_lr$parameter`) = `r z_cp_lr$estimate`, CI [`r z_cp_lr$conf.int[1]`,`r z_cp_lr$conf.int[2]`], p=`r z_cp_lr$p.value`.
-
-```{r zoo-lab-chunk, echo=F}
-
-#create matrix to hold r results
-myr=matrix(NA,nrow=2,ncol=3)
-rownames(myr)<-c("lp","cp")
-colnames(myr)<-c("AS","LR","all")
-mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
-
-proportions=proportions[order(proportions$ChildID),]
-
-  #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
-    myr[1,"AS"]=round(thiscor$estimate,3)
-    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
-    myn[1,"AS"]=round(thiscor$parameter,3)
-    myp[1,"AS"]=round(thiscor$p.value,3)
-
-  #repeat for LR
-
-      #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"])
-    myr[1,"LR"]=round(thiscor$estimate,3)
-    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
-    myn[1,"LR"]=round(thiscor$parameter,3)
-    myp[1,"LR"]=round(thiscor$p.value,3)
-
-#repeat for all cor
-
-      thiscor=cor.test(props_bychild_chunk$ling_prop,proportions$l_lp)
-    myr[1,"all"]=round(thiscor$estimate,3)
-    mylb[1,"all"]=round(thiscor$conf.int[1],3)
-    myhb[1,"all"]=round(thiscor$conf.int[2],3)
-    myn[1,"all"]=round(thiscor$parameter,3)
-    myp[1,"all"]=round(thiscor$p.value,3)
-
-    # repeat whole for cp
-
-  #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
-    myr[2,"AS"]=round(thiscor$estimate,3)
-    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
-    myn[2,"AS"]=round(thiscor$parameter,3)
-    myp[2,"AS"]=round(thiscor$p.value,3)
-
-  #repeat for LR
-
-      #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"])
-    myr[2,"LR"]=round(thiscor$estimate,3)
-    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
-    myn[2,"LR"]=round(thiscor$parameter,3)
-    myp[2,"LR"]=round(thiscor$p.value,3)
-
-#repeat for all cor
-
-      thiscor=cor.test(props_bychild_chunk$can_prop,proportions$l_cp)
-    myr[2,"all"]=round(thiscor$estimate,3)
-    mylb[2,"all"]=round(thiscor$conf.int[1],3)
-    myhb[2,"all"]=round(thiscor$conf.int[2],3)
-    myn[2,"all"]=round(thiscor$parameter,3)
-    myp[2,"all"]=round(thiscor$p.value,3)
-
- mycortab=rbind(mycortab,t(myr))
-rownames(mycortab)[4:6]<-paste("Chunks",rownames(t(myr)))
-
-```
-
-As for correlations between Zooniverse and laboratory-derived metrics, we observe very similar levels of correlation: Linguistic Proportion overall r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
-As for Canonical Proportion overall r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`; Angelman syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`];  low-risk control infants r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
-
-Since all of these results are very similar to those obtained when collapsing chunk-level judgments to generate segment-level judgments, we conclude that in the future this step may not be necessary.
-Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' chunk-level judgments [which was in fact what @cychosz2019canonical did].
-
-```{r downsample}
-
-#we start from the data base that has all the info
-nsegs=data.frame(table(data_all$ChildID))
-colnames(nsegs)<-c("ChildID","nsegs")
-
-merge(nsegs,demo_data)->nsegs
-nseg_t=t.test(nsegs$nsegs~nsegs$Diagnosis)
-
-resampling=matrix(NA,nrow=50,ncol=6)
-colnames(resampling)<-c("lp_r_all","lp_r_as","lp_r_lr","cp_r_all","cp_r_as","cp_r_lr")
-for(i in 1:50){
-  sel_seg=NULL
-  for(thischi in levels(data_all$ChildID)) sel_seg<-c(sel_seg,as.character(data_all$segmentId_DB[sample(data_all$segmentId_DB[data_all$ChildID==thischi],100)]))
-
-  sampling=data_all[as.character(data_all$segmentId_DB) %in% as.character(sel_seg),]
-
-  #get the ns by child, then calculate the linguistic proportion & canonical proportion, separately for zooniverse & lab
-  ztab=table(sampling$ChildID,sampling$Zoon_classif)
-  z_lp=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
-  z_cp=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
-
-  ltab=table(sampling$ChildID,sampling$lab)
-  l_lp=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
-  l_cp=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
-
-  thistab=cbind(demo_data,z_lp,z_cp,l_lp,l_cp)
-
-  resampling[i,"lp_r_all"]=cor.test(thistab$z_lp,thistab$l_lp)$estimate
-  resampling[i,"lp_r_as"]=cor.test(thistab$z_lp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
-  resampling[i,"lp_r_lr"]=cor.test(thistab$z_lp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate
-
-   resampling[i,"cp_r_all"]=cor.test(thistab$z_cp,thistab$l_cp)$estimate
-  resampling[i,"cp_r_as"]=cor.test(thistab$z_cp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
-  resampling[i,"cp_r_lr"]=cor.test(thistab$z_cp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate
- }
-myr=colMeans(resampling)
-mycortab=rbind(mycortab,cbind(myr[c("lp_r_as" ,"lp_r_lr","lp_r_all")],myr[c("cp_r_as" ,"cp_r_lr","cp_r_all")]))
-rownames(mycortab)[7:9]<-c("100 seg AS","100 seg LR","100 seg all")
-round(mycortab,3)->mycortab
-
-resampling=data.frame(resampling)
-```
-
-Next, we looked at whether having more segments from each child may lead to more reliable metrics.
-We observed a non-significant trend [t(`r nseg_t$parameter`)=`r nseg_t$estimate`] for lower number of segments in the Angelman syndrome group (mean `r nseg_t$estimate[1]`, range `r min(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`) than among low-risk control infants (mean `r nseg_t$estimate[2]`, range `r min(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`), likely because of the lower volubility of the former children.
-Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications [@cychosz2019canonical].
-We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children [following @cychosz2019canonical].
-We randomly sampled 100 segments from each child's data, and recalculated the linguistic and Canonical Proportions based only on these associated judgments.
-We repeated this process 50 times, to assess the extent to which the association between laboratory- and Zooniverse-derived metrics varied in each random sample.
-The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table 2), particularly for the low-risk younger infants, and for Linguistic Proportion.
-This may indicate that, particularly in some groups of infants, 100 segments may not be sufficient to capture a stable estimate of the child's Linguistic and Canonical Proportions.
-We also observed quite a bit of variance across the 50 runs.
-For Linguistic Proportion, the range of correlations found for all infants was [`r min(resampling$lp_r_all)`,`r max(resampling$lp_r_all)`]; for the Angelman syndrome group [`r min(resampling$lp_r_as)`,`r max(resampling$lp_r_as)`]; for the low-risk group [`r min(resampling$lp_r_lr)`,`r max(resampling$lp_r_lr)`].
-For Canonical Proportion, the range of correlations found for all infants was [`r min(resampling$cp_r_all)`,`r max(resampling$cp_r_all)`]; for the Angelman syndrome group [`r min(resampling$cp_r_as)`,`r max(resampling$cp_r_as)`]; for the low-risk group [`r min(resampling$cp_r_lr)`,`r max(resampling$cp_r_lr)`].
+```{r tab-LENA-challenge, comment = F, message = F, hide = T, warning = F, echo = F, results = "asis"}
+mycaption="Comparison of recall percentages obtained with the baseline algorithm created by the ComParE team, in the context of the ComParE 2019 BabySounds subchallenge (Schuller et al., 2019), LENATM labels (note LENATM does not distinguish between Canonical and Non-canonical), and those obtained in this study through Zooniverse annotations.
+Label frequency indicates the prevalence of the relevant label (Crying, Laughing, etc.) in each dataset (the same Lab dataset was used for both LENATM and Zooniverse).
+UAR stands for unweighted average recall, WAR for weighted average recall (which takes into account label frequency)."
+
+read.table("tab-lena-challenge.txt",header=F,sep="\t",skip=1)->tab1
+#olnames(tab1)<-c("","Recall","Recall","Recall","Label Frequency","Label Frequency")
+
+# as_hux(tab1) %>% 
+#     insert_row("","Recall","","","Label Frequency","", fill = "") %>% 
+#     merge_cells(1, 2:4) %>% 
+#     merge_cells(1, 5:6) %>%  
+#     merge_cells(2, 5:6) %>%  
+#     set_header_rows(1, TRUE) %>% 
+#    style_headers(bold = TRUE)
+
+kable(tab1,cap=mycaption, col.names = NULL) %>%
+  collapse_rows(columns = 1:6, valign = "middle") #%>%
+#add_header_above(c(" " = 1, "Recall"= 3, "Label Frequency" = 2))
+```
 
@@
 ```
-
-<!-- - look at whether kids who have a high "Junk" proportion according to Zooniverse  -->
-<!--      - have lower levels of concordance across Zoo coders -->
-
-<!-- - take all 5 Zoo judgments or 3 or 1 (may need to run the latter two a couple of times to check for potential variance -- if variance is small, we'll just say so, if it's large then we'll have to write a loop and report CIs) -->
-
-## Discussion
-
-In the present study we assessed the extent to which child vocalizations from LENA^TM^ daylong recordings can be accurately described based on classifications collected using the citizen science platform Zooniverse.
-Our recordings came from children diagnosed with a genetic syndrome associated with severe language disorder, as well as low-risk controls, and included a range of different ages.
-Excerpts of these recordings had previously been annotated by trained experts in the lab, and served as the "gold standard".
-
-In descriptive terms, we found that category frequency was similar across laboratory and citizen science annotations.
-Correlations between age and our two child-level metrics revealed different developmental patterns between our sample of older children with a diagnosis of Angelman syndrome and younger low-risk controls.
-Given that the two groups of children differ on both of these features (age and diagnosis), we cannot empirically tease apart these two differences.
-Nonetheless, inspection of Figure 3 suggests different interpretations for the two derived metrics.
-Overall, the pattern for Linguistic Proportion is compatible with a non-linear trajectory common to the two groups, whereby the Linguistic Proportion increases rapidly up to 20 months, and hovers around 90-100% thereafter.
-In this case, the lack of correlation with age in the Angelman syndrome group suggests that there is little systematic variance in this metric after 20 months of age.
-In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and slower decreases among children in the Angelman syndrome group.
-
-Next, we looked at agreement between laboratory and Zooniverse annotation at the level of individual segments.
-We found that some categories had high precision and recall -- notably Non-canonical segments were detected quite accurately, and to a lesser extent the same was true for Canonical segments.
-This is interesting because it suggests that Zooniverse annotations may be trusted particularly for detecting these two types of speech-like segments.
-
-In other cases, precision was fairly high but recall was lower.
-This was the case of the "Junk/Don't mark" category.
-Higher precision and low recall indicates that when Zooniverse annotations return a "Junk" judgment, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as "Don't mark" in the laboratory.
-Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the "Junk" category less) than laboratory annotators, who have access to the whole vocalization.
-This is consistent with the fact that our lab definition of "Junk" was more stringent than the Zooniverse definition.
-Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use "Don't mark" if there was *any* noise or overlapping speech that would potentially affect acoustic analyses.
-
-Both Crying and Laughing showed moderate to good recall, but low precision.
-This means that a high proportion of segments detected as these two categories in the lab can be classified in the appropriate category by Zooniverse annotators, but Zooniverse annotators also put into these categories many segments that lab annotators classified in a different manner.
-Most saliently, there is a problematic level of confusion between Crying and Non-Canonical, whereby lab-detected Crying was almost as likely to be classified as such or Non-canonical in Zooniverse (the recall proportion was 49 and 47% respectively), and even worse, 55% of the segments Zooniverse annotators classified as Crying were actually Non-Canonical according to lab annotations.
-This is likely due to the fact that lab annotators had access to the whole segment, and could even listen to the context of that segment, so they could accurately interpret a segment as crying.
-Thus, it is possible that additional analyses, using for instance pitch or duration characteristics at the level of the whole segment rather than the chunk could improve classification performance, but for the time being, we cannot recommend Zooniverse annotations to identify crying bouts.
-
-As for Laughing, the recall is quite good but precision is low, with confusion involving "Junk" and Non-Canonical to similar extents.
-We are tentative in our interpretation of this result because the category was very sparsely represented, with barely 200 segments classified as Laughing in the whole data set.
-We thus recommend additional work, although we expect that perhaps there will be a similar problem (and solution) as indicated for Crying, whereby determination of Laughing versus Non-Canonical will require inspection of segment-level properties.
-
-Turning now to the child-level analyses, we were particularly interested in derived metrics because this is more typically the goal of such analyses, for instance to study potential individual or group differences, or investigate the impact of an intervention.
-Our results show that, despite agreement variation at the level of the segment, derived metrics were highly correlated across laboratory and Zooniverse data, meaning that variability in precision and recall did not substantially impact key outcome variables.
-Errors only seemed systematic for the Linguistic Proportion (and not for Canonical Proportion), and they were due to lower Linguistic Proportions from Zooniverse than laboratory data.
-This is due to the fact that a substantial number of Crying segments were classified as Non-Canonical by Zooniverse annotators.
-Nonetheless, since these errors were systematic across infants, they do not affect the study of group or individual variation.
-
-Our exploratory analyses then dug into some methodological aspects of our data.
-We found that, although individuals vary widely in terms of what proportion of their data is classified as Junk by Zooniverse annotators, this is not a sign that their data as a whole is compromised, with similar levels of correlations between laboratory- and Zooniverse-derived metrics across median groups based on proportion of their data classified as Junk.
-
-Next, we found that more data led to more reliable measurements: Specifically, our correlations based on 300-500 segments per child were markedly higher than those based on 100 segments per child, particularly when looking at our low-risk, younger participants.
-It is important to note that an advantage of Zooniverse, compared to laboratory annotations, is that it is easier to scale up: More data from each individual child can be annotated this way, which will help countering noise.
-
-In sum, we believe these data show that young children's vocalizations can be accurately described by relying on the judgments of Zooniverse's citizen scientists, particularly when the goal is to describe vocalizations at a broader level than individual chunks or segments.
-It is important to note that from a methodological point of view, the use of crowdsourcing is not merely a question of being an ecological technique, but it can also offer benefits in the area of scientific rigor and reproducibility.
-As previously mentioned in the Introduction section, Zooniverse offers both an automated system of data handling as well as an enormous pool of users: access to a large-scale group of blinded participants can in itself help researchers overcome the methodological compromises that sometimes lead to biased results.
-Additionally, crowd-sourcing and open-science frameworks can come together to encourage replication of existing work: open-source software, shared data and citizen science platforms would make it possible, and actually easy, to run identical studies and evaluate research results independently.
-
-### Further research directions
-
-These findings open up a new, exciting avenue for future research.
-Our pipeline to upload data on Zooniverse is scalable and quickly adapted to new datasets, including data on infants of different ages, learning different languages, or living in environments with different acoustic properties (e.g., rural versus urban environments).
-This means large existing datasets of daylong recordings, notably those hosted on Homebank [@vandam2016homebank], can be reliably annotated in an ecological way using citizen science.
-
-Our results warrant additional research on two metrics we used to describe children’s linguistic productions: Linguistic Proportion and Canonical Proportion.
-Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion.
-We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant for other disorders that are related to emotional disorders [see a review in @halpern2016excessive].
-Additionally, while Canonical Proportion has been studied in much previous work, we believe further research is necessary using larger scales (saliently increasing the sheer number of children studied, to have sufficient power to detect group effects) and with broader coverage [notably including children growing up in a broad range of languages and cultures, as in @cychosz2019canonical], to make sure that assumptions regarding cross-linguistic and maturational universality are amply justified and not merely a result of having low power to detect differences, or simply sampling from similar populations.
-
-Assuming these child-level metrics continue to be supported by such additional data, it would be interesting for future investigations to expand the analysis of Linguistic Proportion and Canonical Proportion to recover differences and similarities in language development between low-risk children and children with disorders other than Angelman syndrome.
-That is, it is at present unclear whether relatively lower Canonical Proportions will also be found in other populations, or whether this is more specific to Angelman syndrome.
-For instance, some previous work has found differences in babbling even when comparing late talkers with toddlers with no specific diagnosis  [@fasolo2008babbling].
-We thus encourage further research inspecting Canonical Proportion in other populations, as well as methodological work that attempts to generate metrics that are similarly easy to gather at scale.
-
-**remaining methodological problems with the LENA pipeline (p. XX);
-the relatively small sample sizes of the two groups we studied (p. XX);
-the kinds of populations that we cannot be certain this work generalizes to (p. XX).
-relative dearth of LENA data with atypical populations **
-
-It is also important to note that work on children with non-normative development poses in itself some challenges.
-Our data contained a lower number of vocalizations for children diagnosed with Angelman syndrome, even though we selected much younger low-risk control children as the comparison group.
-However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
-Now that this method has been validated, researchers can extract all segments that LENA^TM^ identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
-
-Another open question we hope future work looks into is the assessment of within-participant changes.
-Our dataset only contained one daylong recording per child (with one exception), so other datasets in which a greater number of recordings are collected longitudinally can bring novel insights.
-Is the age-related decline we observed among children in the Angelman syndrome group truly a developmental decline that happens by necessity in this population, or might it reflect partially other aspects of children's experience and behavior?
-Additionally, longitudinal work within the same infant may be needed to track the "natural history" of vocalizations to better contextualize how these metrics should be interpreted.
-It is also an open question how large must effects be, at the individual level, in order to be detected with this method.
-For example, there is interest in using naturalistic recordings as outcome measures for clinical (behavioral or pharmacological) trials.
-The fact that agreement across annotators is high even in the data coming from children diagnosed with Angelman syndrome is encouraging, but it would be important to establish whether absolute levels of Canonical Proportion, and/or rates of change over longer time spans, may be good candidates for tracking the effects of any treatment.
-
-Finally, broader generalizations of our technique could be of interest to readers of this paper.
-Any researcher is able to undergo the procedure to create Zooniverse projects (https://www.zooniverse.org/about/faq).
-For instance, a project could be set up to annotate potentially disordered speech by older adults at risk of, or diagnosed with, neurodegenerative diseases.
-Although it would be important to similarly validate such annotations, we believe there is strong promise for such tasks, because many neurodegenerative disorders affect local aspects of the speech, which are detected at the syllable level.
-In contrast, tasks that require more speech or audio context may not be well suited to citizen science platforms, because they would require playing longer clips, which may reveal identifying or sensitive information.
-Although citizen scientists themselves are well-intentioned, the platform could be used by others who want to exploit the system for other means.
-
-## Conclusion
-
-In this study we validated the quality of annotations obtained through a citizen science platform, Zooniverse, as compared to a gold standard of human expert annotators.
-We analyzed the correspondence between annotations at the individual segment level, and the individual-child level, using Canonical Proportion and Linguistic Proportion as descriptors.
-We found moderate to good accuracy in the first, and strong positive correlations in the latter.
-We can conclude that citizen scientists are a reliable source of fast and ecological annotation of speech data, particularly when results are combined into child-level descriptors.
-The same methodology may be applied to several research questions in the study of language acquisition and language disorders.
-This finding is particularly welcome in an era when wearables open new avenues for studying human behavior and development in an ecological manner.
-
-\newpage
-
-# Acknowledgements
-
-We are grateful to the families who contributed their data; to the Zooniverse volunteers that made this work possible; and the audience and technical committee of SLT 2020 for helpful feedback on preliminary analyses.
-
-# References
-
-<!-- These lines ensure references are set with hanging indents in PDF documents; they are ignored in Word. -->
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}

### 2021-01-06 202010_jslhr/paper.Rmd
implemented up to Discussion but 2-3 para remain to do, knits

@@
 
+Other automatized algorithms have been developed in the last two years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical [@schuller2019interspeech].
+The split of speech vocalizations into canonical and non-canonical, and the overt inclusion of laughing, allow a somewhat finer-grained picture of children's vocal activity than what the LENA^TM^ algorithm currently offers.
+Another advantage of this new line of research is that there exists a challenge where performance has been benchmarked, so unlike LENA^TM^'s algorithms, we know which algorithms perform better than others.
+Indeed, the ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55%.
+By re-using the same test set, researchers can develop new algorithms and prove that they can outperform this baseline (or any subsequent best-performing algorithm benchmarked in the same test set).
+
+Table 1 provides an idea of the accuracy across different methods.
+Specifically, we calculated accuracy of the LENA^TM^'s subtypes[^1] based on the laboratory annotations used in the present study; and we also show the results of ComParE 2019 BabySounds sub-challenge, although note it is not based on the same data as the other two columns.
+Table 1 shows that LENA^TM^ has a good recall for speech-like vocalizations and crying but not laughing, and by design it does not distinguish between canonical and non-canonical vocalizations.
+The ComParE 2019 BabySounds sub-challenge's baseline model achieved a more moderate recall on crying than LENA, but clearly outperforms the LENA^TM^ algorithm in the Laughing category, with the additional advantage of distinguishing canonical and non-canonical -- although note that performance in non-canonical is barely above a chance level of 20% (based on 5 categories, assuming equal probabilities).
+Notice that the baseline model also included a category called "Junk" for sections that turned out not to be the child's vocalizations after all, but LENA^TM^ here has a low recall because only sections that LENA^TM^ classified as being child vocalizations are considered in the present study, and thus this class's category cannot but be low-performing for LENA^TM^.
+That is, since we selected only segments attributed to the key child, any segment that turns out not to be a child vocalization is a LENA^TM^ error.
+We return to the comparison between LENA^TM^, the ComParE 2019 BabySounds sub-challenge baseline, and our Zooniverse-based method in the Discussion, but for now, we simply point out that there are new algorithms being developed, but their performance --although competitive when compared to LENA^TM^-- is underwhelming.
+
 ----- INSERT TABLE 1 HERE -----
-
-**CHECK FLOW IN THIS PARA** Other automatized algorithms have been developed in recent years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical [@schuller2019interspeech].
-The split of speech vocalizations into canonical and non-canonical, and the overt inclusion of laughing, allow a somewhat finer-grained picture of children's vocal activity.
-An advantage of these algorithms is that there exists a challenge where performance has been benchmarked, so unlike LENA^TM^'s algorithms, we know which algorithms perform better than others.
-The ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55%.
-To provide an idea of the accuracy across these different methods, we classified the same segments that are used in the present study using LENA^TM^'s subtypes[^1] to compare these against the laboratory annotations used in the present study; and we also show the results of ComParE 2019 BabySounds sub-challenge (although note it is not based on the same data as the other two columns).
-Table 1 shows that LENA^TM^ has a good recall for speech-like vocalizations and crying but not laughing.
-**ADD MORE INFORMATION SO READERS UNDERSTAND THE POINT OF THIS TABLE.**  The team who won the challenge improved this by about 2%, primarily through gains in the laughing class obtained by adding training data [@yeh2019using].
-Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are weighted based on their frequency of occurrence, changes very much.
 
@@
 When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk judgments); and the Canonical Proportion (Canonical)/(all linguistic judgments).
+
+It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels.
+We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA^TM [see meta-analyses in @wang2020meta;@cristia2020accuracy]; as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range.
+Some previous work had shown that vocal development proceeded differently when compared to a normative sample [@hamrick2019capturing;@hamrick2019measurement].
+Generally, studies on vocal development among children with genetic disorders reveal lower volubility and lower prevalence of canonical syllables [@rankine2017language;@belardi2017retrospective; see also @roche2018early for a review of retrospective video analysis of children diagnosed with Autism Spectrum Disorder, Rett Syndrome, or Fragile X].
+Low volubility is one of the contexts in which daylong recordings may be particularly advantageous over alternative recording methods [see @rankine2017language for discussion], since it is challenging to measure vocal development accurately based on shorter recordings (as the child may not vocalize at all in a 1-2h recording).
+Additionally, the ideal method will be sensitive not only to the large changes that occur over normative development, but also to the potentially smaller individual differences found among children.
+Such a level of sensitivity is important if these methods are ever to be used to detect potential effects of intervention.
+
+Returning to our overall predictions with these goals in mind, we were particularly interested in describing potential differences in accuracy of derived metrics as a function of our two participant groups.
 Previous work suggests that children with Angelman syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the Canonical Proportion was expected to increase with age [@hamrick2019capturing;@hamrick2019measurement].
@@
 More specific to our research aims, we checked the degree to which laboratory and citizen science annotations converged at the child level by using correlations across the two.
-
-It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels.
-We therefore included children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range, as well as a group of low-risk control infants.
-**Some additional information about Angelman syndrome would be helpful to understand the findings.
-Were the different patterns between the canonical and linguistic proportions predicted by the already-known characteristics of this syndrome?
-Has other work included this population?
-Why was this population selected?
-Are there hypotheses about the validity between groups or is this more exploratory?** **relative dearth of LENA data with atypical populations **
 
@@
 Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical+Non-Canonical)/All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical/(Canonical+Non-Canonical) (i.e., we remove Junk, Crying, and Laughing).
-See Figure 3 for results.
-**the children with Angelman syndrome scored much lower than the typically-developing infants on the canonical measure but higher on the linguistic measure.**
+Figure 3 shows results of plotting our Proportions against child age.
+We draw the readers' attention to the fact that the children diagnosed with Angelman syndrome had overall higher levels of Linguistic Proportion than the typically-developing infants, but the opposite was true for Canonical Proportion.
+
+```{r matr}
+#create matrix to hold r results
+myr=matrix(NA,nrow=4,ncol=2)
+rownames(myr)<-c("z_lp","z_cp","l_lp","l_cp")
+colnames(myr)<-c("AS","LR")
+mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
+
+for(thisvar in c("z_lp","z_cp","l_lp","l_cp")) {
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"])
+    myr[thisvar,"AS"]=round(thiscor$estimate,3)
+    mylb[thisvar,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[thisvar,"AS"]=round(thiscor$conf.int[2],3)
+    myn[thisvar,"AS"]=round(thiscor$parameter,3)
+    myp[thisvar,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"])
+    myr[thisvar,"LR"]=round(thiscor$estimate,3)
+    mylb[thisvar,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[thisvar,"LR"]=round(thiscor$conf.int[2],3)
+    myn[thisvar,"LR"]=round(thiscor$parameter,3)
+    myp[thisvar,"LR"]=round(thiscor$p.value,3)
+
+    #add r to figure
+
+}
+```
 
@@
 
------ INSERT FIGURE 5 HERE -----
+```{r matr2}
+#create matrix to hold r results
+myr=matrix(NA,nrow=2,ncol=3)
+rownames(myr)<-c("lp","cp")
+colnames(myr)<-c("AS","LR","all")
+mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_lp"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
+    myr[1,"AS"]=round(thiscor$estimate,3)
+    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
+    myn[1,"AS"]=round(thiscor$parameter,3)
+    myp[1,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"],proportions[proportions$Diagnosis!="AngelmanSyndrome","z_lp"])
+    myr[1,"LR"]=round(thiscor$estimate,3)
+    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
+    myn[1,"LR"]=round(thiscor$parameter,3)
+    myp[1,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(proportions$z_lp,proportions$l_lp)
+    myr[1,"all"]=round(thiscor$estimate,3)
+    mylb[1,"all"]=round(thiscor$conf.int[1],3)
+    myhb[1,"all"]=round(thiscor$conf.int[2],3)
+    myn[1,"all"]=round(thiscor$parameter,3)
+    myp[1,"all"]=round(thiscor$p.value,3)
+
+    # repeat whole for cp
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_cp"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
+    myr[2,"AS"]=round(thiscor$estimate,3)
+    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
+    myn[2,"AS"]=round(thiscor$parameter,3)
+    myp[2,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"],proportions[proportions$Diagnosis!="AngelmanSyndrome","z_cp"])
+    myr[2,"LR"]=round(thiscor$estimate,3)
+    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
+    myn[2,"LR"]=round(thiscor$parameter,3)
+    myp[2,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(proportions$z_cp,proportions$l_cp)
+    myr[2,"all"]=round(thiscor$estimate,3)
+    mylb[2,"all"]=round(thiscor$conf.int[1],3)
+    myhb[2,"all"]=round(thiscor$conf.int[2],3)
+    myn[2,"all"]=round(thiscor$parameter,3)
+    myp[2,"all"]=round(thiscor$p.value,3)
+
+mycortab=t(myr)
+rownames(mycortab)<-paste("All seg",rownames(mycortab))
+```
 
@@
 
+----- INSERT FIGURE 5 HERE -----
+
 Similarly, a strong positive correlation is found in the Canonical Proportion r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`].
@@
 
-```
-
------ INSERT FIGURE 6 HERE -----
+#create matrix to hold r results
+myr=matrix(NA,nrow=2,ncol=2)
+rownames(myr)<-c("lp","cp")
+colnames(myr)<-c("lower_Junk","higher_Junk")
+mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$median_Junk=="lower_Junk","z_lp"],proportions[proportions$median_Junk=="lower_Junk","l_lp"])
+    myr[1,"lower_Junk"]=round(thiscor$estimate,3)
+    mylb[1,"lower_Junk"]=round(thiscor$conf.int[1],3)
+    myhb[1,"lower_Junk"]=round(thiscor$conf.int[2],3)
+    myn[1,"lower_Junk"]=round(thiscor$parameter,3)
+    myp[1,"lower_Junk"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$median_Junk!="lower_Junk","l_lp"],proportions[proportions$median_Junk!="lower_Junk","z_lp"])
+    myr[1,"higher_Junk"]=round(thiscor$estimate,3)
+    mylb[1,"higher_Junk"]=round(thiscor$conf.int[1],3)
+    myhb[1,"higher_Junk"]=round(thiscor$conf.int[2],3)
+    myn[1,"higher_Junk"]=round(thiscor$parameter,3)
+    myp[1,"higher_Junk"]=round(thiscor$p.value,3)
+
+    # repeat whole for cp
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$median_Junk=="lower_Junk","z_cp"],proportions[proportions$median_Junk=="lower_Junk","l_cp"])
+    myr[2,"lower_Junk"]=round(thiscor$estimate,3)
+    mylb[2,"lower_Junk"]=round(thiscor$conf.int[1],3)
+    myhb[2,"lower_Junk"]=round(thiscor$conf.int[2],3)
+    myn[2,"lower_Junk"]=round(thiscor$parameter,3)
+    myp[2,"lower_Junk"]=round(thiscor$p.value,3)
+
+  #repeat for higher_Junk
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$median_Junk!="lower_Junk","l_cp"],proportions[proportions$median_Junk!="lower_Junk","z_cp"])
+    myr[2,"higher_Junk"]=round(thiscor$estimate,3)
+    mylb[2,"higher_Junk"]=round(thiscor$conf.int[1],3)
+    myhb[2,"higher_Junk"]=round(thiscor$conf.int[2],3)
+    myn[2,"higher_Junk"]=round(thiscor$parameter,3)
+    myp[2,"higher_Junk"]=round(thiscor$p.value,3)
+```
 
@@
 Thus, it does not seem that a higher proportion of "Junk" judgments is an index of low quality data.
+
+----- INSERT FIGURE 6 HERE -----
 
@@
 
+Notwithstanding our enthusiasm, we would like to point out some limitations of our results that can be addressed with more research.
+The first area in which more research would be welcome involves the pre-processing.
+We used LENA^TM^ algorithms to detect vocalizations by the key child, but there are newer, better-performing, open-sourced alternatives that could be used instead [@lavechin2020open].
+Still, performance is most often established on children whose development is thought to be normative, being raised in small households located in urban environments, whereas it would be important to check for accuracy in more diverse samples, where the children have a diagnosis, multiple siblings, and/or are growing up in a rural environment.
+Speech technology developments would also be welcome to complement citizen scientists' classifications, particularly for categories that may be hard to classify based on short samples.
+**remaining methodological problems with the LENA pipeline (p.
+XX); The team who won the challenge improved this by about 2%, primarily through gains in the laughing class obtained by adding training data [@yeh2019using].
+Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are weighted based on their frequency of occurrence, changes very much.
+**
+
 Our results warrant additional research on two metrics we used to describe children’s linguistic productions: Linguistic Proportion and Canonical Proportion.
@@
 
-**remaining methodological problems with the LENA pipeline (p. XX);
-the relatively small sample sizes of the two groups we studied (p. XX);
-the kinds of populations that we cannot be certain this work generalizes to (p. XX).
-relative dearth of LENA data with atypical populations **
+**relative dearth of LENA data with atypical populations **
 
@@
 However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
-Now that this method has been validated, researchers can extract all segments that LENA^TM^ identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
+Now that this method has been validated, researchers can extract all segments that LENA^TM^ identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.**the relatively small sample sizes of the two groups we studied (p.
+XX);**
+**the kinds of populations that we cannot be certain this work generalizes to (p. XX).**
 

### 2021-01-06 202010_jslhr/paper.tex
implemented up to Discussion but 2-3 para remain to do, knits

@@
 \usepackage{csquotes}
+\usepackage[titles]{tocloft}
+\cftpagenumbersoff{figure}
+\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
+\renewcommand{\cftfigaftersnum}{.\space}
+\setlength{\cftfigindent}{0pt}
+\setlength{\cftafterloftitleskip}{0pt}
+\settowidth{\cftfignumwidth}{Figure 10.\qquad}
 \usepackage{setspace}
@@
 
-\begin{table}
+Other automatized algorithms have been developed in the last two years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical (Schuller et al., 2019).
+The split of speech vocalizations into canonical and non-canonical, and the overt inclusion of laughing, allow a somewhat finer-grained picture of children's vocal activity than what the LENA\textsuperscript{TM} algorithm currently offers.
+Another advantage of this new line of research is that there exists a challenge where performance has been benchmarked, so unlike LENA\textsuperscript{TM}'s algorithms, we know which algorithms perform better than others.
+Indeed, the ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55\%.
+By re-using the same test set, researchers can develop new algorithms and prove that they can outperform this baseline (or any subsequent best-performing algorithm benchmarked in the same test set).
 
-\caption{\label{tab:tab-LENA-challenge}Comparison of recall percentages obtained with the baseline algorithm created by the ComParE team, in the context of the ComParE 2019 BabySounds subchallenge (Schuller et al., 2019), LENATM labels (note LENATM does not distinguish between Canonical and Non-canonical), and those obtained in this study through Zooniverse annotations.
-Label frequency indicates the prevalence of the relevant label (Crying, Laughing, etc.) in each dataset (the same Lab dataset was used for both LENATM and Zooniverse).
-UAR stands for unweighted average recall, WAR for weighted average recall (which takes into account label frequency).}
-\centering
-\begin{tabular}[t]{l|l|l|l|l|l}
-\hline
- & C2019B & LENA & Zoon. & C2019B & Lab\\
-\hline
-Crying & 70.6 & 78.0 & 49.0 & 669 & 581\\
-\cline{1-6}
-Laughing & 41.5 & 12.0 & 56.0 & 149 & 137\\
-\cline{1-6}
-Non-can. & 24.1 &  & 77.0 & 4485 & 5512\\
-\cline{1-2}
-\cline{4-6}
-Canonical & 66.4 & \multirow{-2}{*}{\raggedright\arraybackslash 73.0} & 59.0 & 1426 & 1715\\
-\cline{1-6}
-Junk & 67.3 & 1.0 & 44.0 & 4575 & 2169\\
-\cline{1-6}
-UAR & 54.0 & 41.0 & 57.0 &  & \\
-\cline{1-4}
-WAR & 49.9 & 57.0 & 65.0 & \multirow{-2}{*}{\raggedright\arraybackslash } & \multirow{-2}{*}{\raggedright\arraybackslash }\\
-\hline
-\end{tabular}
-\end{table}
+Table 1 provides an idea of the accuracy across different methods.
+Specifically, we calculated accuracy of the LENA\textsuperscript{TM}'s subtypes\footnote{If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as \enquote{Junk} or not categorized.} based on the laboratory annotations used in the present study; and we also show the results of ComParE 2019 BabySounds sub-challenge, although note it is not based on the same data as the other two columns.
+Table 1 shows that LENA\textsuperscript{TM} has a good recall for speech-like vocalizations and crying but not laughing, and by design it does not distinguish between canonical and non-canonical vocalizations.
+The ComParE 2019 BabySounds sub-challenge's baseline model achieved a more moderate recall on crying than LENA, but clearly outperforms the LENA\textsuperscript{TM} algorithm in the Laughing category, with the additional advantage of distinguishing canonical and non-canonical -- although note that performance in non-canonical is barely above a chance level of 20\% (based on 5 categories, assuming equal probabilities).
+Notice that the baseline model also included a category called \enquote{Junk} for sections that turned out not to be the child's vocalizations after all, but LENA\textsuperscript{TM} here has a low recall because only sections that LENA\textsuperscript{TM} classified as being child vocalizations are considered in the present study, and thus this class's category cannot but be low-performing for LENA\textsuperscript{TM}.
+That is, since we selected only segments attributed to the key child, any segment that turns out not to be a child vocalization is a LENA\textsuperscript{TM} error.
+We return to the comparison between LENA\textsuperscript{TM}, the ComParE 2019 BabySounds sub-challenge baseline, and our Zooniverse-based method in the Discussion, but for now, we simply point out that there are new algorithms being developed, but their performance --although competitive when compared to LENA\textsuperscript{TM}-- is underwhelming.
 
-\textbf{CHECK FLOW IN THIS PARA} Other automatized algorithms have been developed in recent years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical (Schuller et al., 2019).
-The split of speech vocalizations into canonical and non-canonical, and the overt inclusion of laughing, allow a somewhat finer-grained picture of children's vocal activity.
-An advantage of these algorithms is that there exists a challenge where performance has been benchmarked, so unlike LENA\textsuperscript{TM}'s algorithms, we know which algorithms perform better than others.
-The ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55\%.
-To provide an idea of the accuracy across these different methods, we classified the same segments that are used in the present study using LENA\textsuperscript{TM}'s subtypes\footnote{If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as \enquote{Junk} or not categorized.} to compare these against the laboratory annotations used in the present study; and we also show the results of ComParE 2019 BabySounds sub-challenge (although note it is not based on the same data as the other two columns).
-Table 1 shows that LENA\textsuperscript{TM} has a good recall for speech-like vocalizations and crying but not laughing.
-\textbf{ADD MORE INFORMATION SO READERS UNDERSTAND THE POINT OF THIS TABLE.} The team who won the challenge improved this by about 2\%, primarily through gains in the laughing class obtained by adding training data (Yeh et al., 2019).
-Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are weighted based on their frequency of occurrence, changes very much.
+----- INSERT TABLE 1 HERE -----
 
@@
 When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk judgments); and the Canonical Proportion (Canonical)/(all linguistic judgments).
+
+It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels.
+We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA\^{}TM (see meta-analyses in Wang et al., 2020; Cristia et al., 2020); as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range.
+Some previous work had shown that vocal development proceeded differently when compared to a normative sample (Hamrick et al., 2019; Hamrick \& Tonnsen, 2019).
+Generally, studies on vocal development among children with genetic disorders reveal lower volubility and lower prevalence of canonical syllables (Belardi et al., 2017; Rankine et al., 2017; see also {\textbf{???}} for a review of retrospective video analysis of children diagnosed with Autism Spectrum Disorder, Rett Syndrome, or Fragile X).
+Low volubility is one of the contexts in which daylong recordings may be particularly advantageous over alternative recording methods (see Rankine et al., 2017 for discussion), since it is challenging to measure vocal development accurately based on shorter recordings (as the child may not vocalize at all in a 1-2h recording).
+Additionally, the ideal method will be sensitive not only to the large changes that occur over normative development, but also to the potentially smaller individual differences found among children.
+Such a level of sensitivity is important if these methods are ever to be used to detect potential effects of intervention.
+
+Returning to our overall predictions with these goals in mind, we were particularly interested in describing potential differences in accuracy of derived metrics as a function of our two participant groups.
 Previous work suggests that children with Angelman syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the Canonical Proportion was expected to increase with age (Hamrick et al., 2019; Hamrick \& Tonnsen, 2019).
@@
 More specific to our research aims, we checked the degree to which laboratory and citizen science annotations converged at the child level by using correlations across the two.
-
-It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels.
-We therefore included children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range, as well as a group of low-risk control infants.
-\textbf{Some additional information about Angelman syndrome would be helpful to understand the findings.
-Were the different patterns between the canonical and linguistic proportions predicted by the already-known characteristics of this syndrome?
-Has other work included this population?
-Why was this population selected?
-Are there hypotheses about the validity between groups or is this more exploratory?} \textbf{relative dearth of LENA data with atypical populations }
 
@@
 We will call this level \emph{chunk}, to convey that they are part of a bigger unit (the LENA\textsuperscript{TM}-identified \enquote{segment}).
+Note that segments and chunks are collectively referred to as clips because they are parts of the longer audio recording.
 To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms chunks and their corresponding segment.
@@
 
-\begin{figure}
-\centering
-\includegraphics{zooniverse-pufig.pdf}
-\caption{\label{fig:fig-zoo}Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needed to make; as well as the Field Guide, which remained accessible on the right of the screen throughout the session, and which allowed users to revise examples of the different categories.}
-\end{figure}
+----- INSERT FIGURE 1 HERE -----
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{fig_levels.key.pdf}
-\caption{(\#fig:fig:process)Correspondance between LENA-identified segments (which can contain one or more vocalizations) and chunks.
-Notice that both segments and chunks are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority.
-To compare laboratory and Zooniverse annotations at the level of segments, chunk-level judgments are combined using the hierarchy Canonical\textgreater{}Non-canonical\textgreater{}Crying\textgreater{}Laughing.}
-\end{figure}
+----- INSERT FIGURE 2 HERE -----
 
@@
 Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical+Non-Canonical)/All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical/(Canonical+Non-Canonical) (i.e., we remove Junk, Crying, and Laughing).
-See Figure 2 for results.
-\textbf{the children with Angelman syndrome scored much lower than the typically-developing infants on the canonical measure but higher on the linguistic measure.}
+Figure 3 shows results of plotting our Proportions against child age.
+We draw the readers' attention to the fact that the children diagnosed with Angelman syndrome had overall higher levels of Linguistic Proportion than the typically-developing infants, but the opposite was true for Canonical Proportion.
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/fig-corage-1.pdf}
-\caption{\label{fig:fig-corage}Correlations between child-level descriptors and age as a function of metric (Linguistic Proportion in the top row, Canonical Proportion in the bottom row), annotation method, and child group (red=low-risk, and black= Angelman syndrome).}
-\end{figure}
+----- INSERT FIGURE 3 HERE -----
 
@@
 Next, we discuss the correspondence between citizen science classifications and the laboratory gold standard, at the level of individual segments.
-Results were visualized with a confusion matrix showing precision and recall (Figure 3): the diagonal elements show the number of correct segment-level classifications for each class while the off-diagonal elements show non-matching classifications.
+Results were visualized with a confusion matrix showing precision and recall (Figure 4): the diagonal elements show the number of correct segment-level classifications for each class while the off-diagonal elements show non-matching classifications.
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/fig-prec-rec-1.pdf}
-\caption{\label{fig:fig-prec-rec}Number of segments as a function of their majority label from laboratory annotations (rows) versus Zooniverse annotations (columns).
-Percentages (and shading) indicate precision (left) and recall (right).
-For precision, the percentage indicates what percentage of the segments that have a given majority label according to Zooniverse data are attributed a given label in Lab data (i.e., columns add up to 100\%).
-For recall, the percentage indicates what percentage of the segments that have a given majority label according to Lab data are attributed a given label in Zooniverse data (i.e., rows add up to 100\%).}
-\end{figure}
+----- INSERT FIGURE 4 HERE -----
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/fig-corlab-zoo-1.pdf}
-\caption{\label{fig:fig-corlab-zoo}Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of child group (red=low-risk, and black=Angelman syndrome).}
-\end{figure}
-
 Although the classification at the segment level is only moderately accurate, what we are ultimately interested in is whether citizen scientists' classifications are able to provide a reliable snapshot of children's individual development.
-We illustrate this association in Figure 3.
+We illustrate this association in Figure 5.
 Looking at all 20 children together, we found a strong positive correlation r(18) = 0.79, CI {[}0.54,0.92{]}, p=0{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
 When we split by participant group, correlations remain high: for Angelman syndrome r(8) = 0.88, CI {[}0.56,0.97{]}, p=0.00{]}; low-risk control r(8) = 0.83, CI {[}0.42,0.96{]}, p=0.00{]}.
+
+----- INSERT FIGURE 5 HERE -----
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/fig-cor-Junk-1.pdf}
-\caption{\label{fig:fig-cor-Junk}Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of whether children's Junk proportion was higher or lower than the median for all children (blue=lower, and green=higher).}
-\end{figure}
-
 We investigated this hypothesis by calculating the proportion of their data labeled as \enquote{Junk} for each individual child.
@@
 We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as \enquote{Junk}.
-Results were similar across these two post-hoc subgroups (see Figure 4).
+Results were similar across these two post-hoc subgroups (see Figure 6).
 For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = 0.86, CI {[}0.51,0.97{]}, p=0.00{]}; and for the higher Junk group it was r(8) = 0.90, CI {[}0.61,0.98{]}, p=0.
@@
 
+----- INSERT FIGURE 6 HERE -----
+
 Above, we concluded that derived metrics integrating information across audio clips (linguistic and canonical proportion, which can be derived from segment- or chunk-level judgments) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA\textsuperscript{TM} segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
@@
 We therefore repeated our analyses but instead of deriving our proportions for the Zooniverse data from the segment-level composite, we did it based on the individual chunk-level annotations.
+To facilitate comparison, correlations across laboratory- and Zooniverse-derived proportions are provided in Table 2.
+
+----- INSERT TABLE 2 HERE -----
 
@@
 
-As for correlations between Zooniverse and laboratory-derived metrics, we observe very similar levels of correlation: Linguistic Proportion overall r(18) = 0.80, CI {[}0.55,0.92{]}, p=0{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+As for correlations between Zooniverse- and laboratory-derived metrics (e.g., linguistic proportion as estimated using Zooniverse coding, versus laboratory coding), we observe very similar levels of correlation: Linguistic Proportion overall r(18) = 0.80, CI {[}0.55,0.92{]}, p=0{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
 When we split by participant group, correlations remain high: for Angelman syndrome r(8) = 0.84, CI {[}0.46,0.96{]}, p=0.00{]}; low-risk control r(8) = 0.79, CI {[}0.33,0.95{]}, p=0.01{]}.
@@
 
-Since all of these results are very similar to those obtained when first matching chunks to segments, we conclude that in the future this step may not be necessary.
-Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' chunk level judgments (which was in fact what Cychosz et al., 2019 did).
+Since all of these results are very similar to those obtained when collapsing chunk-level judgments to generate segment-level judgments, we conclude that in the future this step may not be necessary.
+Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' chunk-level judgments (which was in fact what Cychosz et al., 2019 did).
 
@@
 We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children (following Cychosz et al., 2019).
-We randomly sampled 100 segments from each child's data, and recalculated the linguistic and Canonical Proportions based only on these associated judgments.
-We repeated this process 50 times, to assess the extent to which the association between laboratory- and Zooniverse-derived metrics varied in each random sample.
+We randomly sampled 100 segments from each child's data, and recalculated the Linguistic and Canonical Proportions based only on these associated judgments.
+We repeated this process 50 times, to assess the extent to which the association between laboratory- and Zooniverse-derived metrics varied in each random sample (e.g., the correlation between Linguistic Proportion as estimated using Zooniverse coding on the one hand, and Linguistic Proportion as estimated using laboratory coding on the other).
 The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table 2), particularly for the low-risk younger infants, and for Linguistic Proportion.
-This may indicate that, particularly in some groups of infants, 100 segments may not be sufficient to capture a stable estimate of the child's Linguistic and Canonical Proportions.
+This may indicate that, especially in some groups of infants, 100 segments may not be sufficient to capture a stable estimate of the child's Linguistic and Canonical Proportions.
 We also observed quite a bit of variance across the 50 runs.
@@
 For Canonical Proportion, the range of correlations found for all infants was {[}0.75,0.94{]}; for the Angelman syndrome group {[}0.79,0.99{]}; for the low-risk group {[}0.55,0.94{]}.
-
-\begin{table}
-
-\caption{\label{tab:tab-cors}Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways.
-First, All seg indicates that Zooniverse annotations at the chunk level were first combined at the segment level.
-Second, Chunks indicates that they were analyzed directly at the chunk level.
-Both of these are based on all the data.
-Third, rows labeled 100 seg indicate that laboratory- and Zooniverse-derived metrics were based on only 100 segments (median over 50 runs in which 100 segments were randomly selected from each child).}
-\centering
-\begin{tabular}[t]{l|r|r}
-\hline
-  & Linguistic Proportion & Canonical Proportion\\
-\hline
-All seg AS & 0.878 & 0.681\\
-\hline
-Chunks AS & 0.845 & 0.724\\
-\hline
-100 seg AS & 0.770 & 0.912\\
-\hline
-All seg LR & 0.828 & 0.984\\
-\hline
-Chunks LR & 0.794 & 0.972\\
-\hline
-100 seg LR & 0.841 & 0.808\\
-\hline
-All seg all & 0.794 & 0.920\\
-\hline
-Chunks all & 0.798 & 0.949\\
-\hline
-100 seg all & 0.719 & 0.873\\
-\hline
-\end{tabular}
-\end{table}
 
@@
 Given that the two groups of children differ on both of these features (age and diagnosis), we cannot empirically tease apart these two differences.
-Nonetheless, inspection of Figure 2 suggests different interpretations for the two derived metrics.
+Nonetheless, inspection of Figure 3 suggests different interpretations for the two derived metrics.
 Overall, the pattern for Linguistic Proportion is compatible with a non-linear trajectory common to the two groups, whereby the Linguistic Proportion increases rapidly up to 20 months, and hovers around 90-100\% thereafter.
 In this case, the lack of correlation with age in the Angelman syndrome group suggests that there is little systematic variance in this metric after 20 months of age.
-In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and slower decreases among children in the Angelman syndrome group.
+In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and relatively stable levels of Canonical Proportion among the children in the Angelman syndrome group.
 
@@
 
+Notwithstanding our enthusiasm, we would like to point out some limitations of our results that can be addressed with more research.
+The first area in which more research would be welcome involves the pre-processing.
+We used LENA\textsuperscript{TM} algorithms to detect vocalizations by the key child, but there are newer, better-performing, open-sourced alternatives that could be used instead ({\textbf{???}}).
+Still, performance is most often established on children whose development is thought to be normative, being raised in small households located in urban environments, whereas it would be important to check for accuracy in more diverse samples, where the children have a diagnosis, multiple siblings, and/or are growing up in a rural environment.
+Speech technology developments would also be welcome to complement citizen scientists' classifications, particularly for categories that may be hard to classify based on short samples.
+\textbf{remaining methodological problems with the LENA pipeline (p.~XX); The team who won the challenge improved this by about 2\%, primarily through gains in the laughing class obtained by adding training data (Yeh et al., 2019).
+Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are weighted based on their frequency of occurrence, changes very much.
+}
+
 Our results warrant additional research on two metrics we used to describe children's linguistic productions: Linguistic Proportion and Canonical Proportion.
@@
 
-\textbf{remaining methodological problems with the LENA pipeline (p.~XX);
-the relatively small sample sizes of the two groups we studied (p.~XX);
-the kinds of populations that we cannot be certain this work generalizes to (p.~XX).
-relative dearth of LENA data with atypical populations }
+\textbf{relative dearth of LENA data with atypical populations }
 
@@
 However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
-Now that this method has been validated, researchers can extract all segments that LENA\textsuperscript{TM} identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
+Now that this method has been validated, researchers can extract all segments that LENA\textsuperscript{TM} identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.\textbf{the relatively small sample sizes of the two groups we studied (p.~XX);}
+\textbf{the kinds of populations that we cannot be certain this work generalizes to (p.~XX).}
 
@@
 \setlength{\leftskip}{0.5in}
+
+\begin{figure}
+\centering
+\includegraphics{zooniverse-pufig.pdf}
+\caption{\label{fig:fig-zoo}Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needed to make; as well as the Field Guide, which remained accessible on the right of the screen throughout the session, and which allowed users to revise examples of the different categories.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{fig_levels.key.pdf}
+\caption{\label{fig:fig-process}Correspondence between LENA-identified segments (which can contain one or more vocalizations) and chunks.
+Segments and chunks are collectively referred to as clips because they are parts of the longer audio recording.
+Notice that clips are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority.
+To compare laboratory and Zooniverse annotations at the level of segments, chunk-level judgments are combined using the hierarchy Canonical\textgreater{}Non-canonical\textgreater{}Crying\textgreater{}Laughing.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/fig-corage-1.pdf}
+\caption{\label{fig:fig-corage}Correlations between child-level descriptors and age as a function of metric (Linguistic Proportion in the top row, Canonical Proportion in the bottom row), annotation method, and child group (red=low-risk, and black= Angelman syndrome).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/fig-prec-rec-1.pdf}
+\caption{\label{fig:fig-prec-rec}Number of segments as a function of their majority label from laboratory annotations (rows) versus Zooniverse annotations (columns).
+Percentages (and shading) indicate precision (left) and recall (right).
+For precision, the percentage indicates what percentage of the segments that have a given majority label according to Zooniverse data are attributed a given label in Lab data (i.e., columns add up to 100\%).
+For recall, the percentage indicates what percentage of the segments that have a given majority label according to Lab data are attributed a given label in Zooniverse data (i.e., rows add up to 100\%).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/fig-corlab-zoo-1.pdf}
+\caption{\label{fig:fig-corlab-zoo}Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of child group (red=low-risk, and black=Angelman syndrome).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/fig-cor-Junk-1.pdf}
+\caption{\label{fig:fig-cor-Junk}Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of whether children's Junk proportion was higher or lower than the median for all children (blue=lower, and green=higher).}
+\end{figure}
+
+\begin{table}
+
+\caption{\label{tab:tab-LENA-challenge}Comparison of recall percentages obtained with the baseline algorithm created by the ComParE team, in the context of the ComParE 2019 BabySounds subchallenge (Schuller et al., 2019), LENATM labels (note LENATM does not distinguish between Canonical and Non-canonical), and those obtained in this study through Zooniverse annotations.
+Label frequency indicates the prevalence of the relevant label (Crying, Laughing, etc.) in each dataset (the same Lab dataset was used for both LENATM and Zooniverse).
+UAR stands for unweighted average recall, WAR for weighted average recall (which takes into account label frequency).}
+\centering
+\begin{tabular}[t]{l|l|l|l|l|l}
+\hline
+ & C2019B & LENA & Zoon. & C2019B & Lab\\
+\hline
+Crying & 70.6 & 78.0 & 49.0 & 669 & 581\\
+\cline{1-6}
+Laughing & 41.5 & 12.0 & 56.0 & 149 & 137\\
+\cline{1-6}
+Non-can. & 24.1 &  & 77.0 & 4485 & 5512\\
+\cline{1-2}
+\cline{4-6}
+Canonical & 66.4 & \multirow{-2}{*}{\raggedright\arraybackslash 73.0} & 59.0 & 1426 & 1715\\
+\cline{1-6}
+Junk & 67.3 & 1.0 & 44.0 & 4575 & 2169\\
+\cline{1-6}
+UAR & 54.0 & 41.0 & 57.0 &  & \\
+\cline{1-4}
+WAR & 49.9 & 57.0 & 65.0 & \multirow{-2}{*}{\raggedright\arraybackslash } & \multirow{-2}{*}{\raggedright\arraybackslash }\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:tab-cors}Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways.
+First, All seg indicates that Zooniverse annotations at the chunk level were first combined at the segment level.
+Second, Chunks indicates that they were analyzed directly at the chunk level.
+Both of these are based on all the data.
+Third, rows labeled 100 seg indicate that laboratory- and Zooniverse-derived metrics were based on only 100 segments (median over 50 runs in which 100 segments were randomly selected from each child).}
+\centering
+\begin{tabular}[t]{r|r}
+\hline
+Linguistic Proportion & Canonical Proportion\\
+\hline
+0.878 & 0.681\\
+\hline
+0.828 & 0.984\\
+\hline
+0.794 & 0.920\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 \leavevmode\hypertarget{ref-crump2013evaluating}{}%
-Crump, M. J., McDonnell, J. V., \& Gureckis, T. M. (2013). Evaluating amazon's mechanical turk as a tool for experimental behavioral research. \emph{PloS One}, \emph{8}(3), e57410.
+Crump, M. J., McDonnell, J. V., \& Gureckis, T. M. (2013). Evaluating Amazon's Mechanical Turk as a tool for experimental behavioral research. \emph{PloS One}, \emph{8}(3), e57410.
 
@@
 
+\clearpage
+\renewcommand{\listfigurename}{Figure captions}
+
 \end{document}

### 2021-01-07 202010_jslhr/paper.Rmd
ready for coauthor feedback

@@
 knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+
+printr=function(x) sub("0.", ".", round(x,3))
 ```
@@
 
-data_ang <- read.csv("segment_AS.csv",header=T,sep=",")
+data_ang <- read.csv("segment_AS.csv",header=T,sep =",")
 data_td <- read.csv("segment_LRC.csv")
@@
 
-demo_data=read.csv("demo-data.tsv",sep="\t")
+demo_data=read.csv("demo-data.tsv",sep ="\t")
 #add filenames to demo data, to be used later
@@
 It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels.
-We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA^TM [see meta-analyses in @wang2020meta;@cristia2020accuracy]; as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range.
+We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA^TM^ [see meta-analyses in @wang2020meta;@cristia2020accuracy]; as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range.
 Some previous work had shown that vocal development proceeded differently when compared to a normative sample [@hamrick2019capturing;@hamrick2019measurement].
-Generally, studies on vocal development among children with genetic disorders reveal lower volubility and lower prevalence of canonical syllables [@rankine2017language;@belardi2017retrospective; see also @roche2018early for a review of retrospective video analysis of children diagnosed with Autism Spectrum Disorder, Rett Syndrome, or Fragile X].
+Generally, studies on vocal development among children with genetic disorders reveal lower volubility and lower prevalence of canonical syllables [@rankine2017language;@belardi2017retrospective; see also @roche2018early for a review of retrospective video analysis of children diagnosed with Autism Spectrum Disorder, Rett Syndrome, or Fragile X, documenting these and other vocal characteristics].
 Low volubility is one of the contexts in which daylong recordings may be particularly advantageous over alternative recording methods [see @rankine2017language for discussion], since it is challenging to measure vocal development accurately based on shorter recordings (as the child may not vocalize at all in a 1-2h recording).
@@
 
-The full data set includes data from 20 children: 10 English-speaking children (6 males, 4 females; age range 11-53 months, mean=41.5 months) diagnosed with Angelman syndrome and 10 low-risk control children (6 males, 4 females; age range 4-18 months, mean=11.7 months).
+The full data set includes data from 20 children: 10 English-speaking children (6 males, 4 females; age M = 41.5 months, range 11-53 months) diagnosed with Angelman syndrome and 10 low-risk control children (6 males, 4 females; age M = 11.7 months, range 4-18 months).
 As children diagnosed with Angelman syndrome typically have severe cognitive and language delays and most do not produce more than 1-2 words consistently, we compare their linguistic production data against that of younger children with theoretically similar language profiles.
@@
 ztab=table(data_all$filename,data_all$Zoon_classif)
-z_lp=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
-z_cp=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
+z_lp =rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
+z_cp =ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
 
 ltab=table(data_all$filename,data_all$lab)
-l_lp=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
-l_cp=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
+l_lp =rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
+l_cp =ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
 
@@
 Zooniverse data revealed a similar distribution:  `r table_zoon["Canonical"]`% Canonical, `r table_zoon["Non-Canonical"]`% Non-Canonical, `r table_zoon["Laughing"]`% Laughing, `r table_zoon["Crying"]`% Crying, `r table_zoon["Junk"]`% Junk.
-Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical+Non-Canonical)/All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical/(Canonical+Non-Canonical) (i.e., we remove Junk, Crying, and Laughing).
+Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical + Non-Canonical) / All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical / (Canonical + Non-Canonical) (i.e., we remove Junk, Crying, and Laughing).
 Figure 3 shows results of plotting our Proportions against child age.
@@
 colnames(myr)<-c("AS","LR")
-mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
+mylb=myhb=myn=myp =myr #this is an ugly way of creating 5 matrices
 
@@
 Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups.
-There was a near-zero relationship among the children diagnosed with Angelman syndrome r(`r myn["l_lp","AS"]`) = `r myr["l_lp","AS"]`, CI [`r mylb["l_lp","AS"]`,`r myhb["l_lp","AS"]`], p=`r myp["l_lp","AS"]`]; and a significant association among low-risk control children r(`r myn["l_lp","LR"]`) = `r myr["l_lp","LR"]`, CI [`r mylb["l_lp","LR"]`,`r myhb["l_lp","LR"]`], p=`r myp["l_lp","LR"]`].
-The Canonical Proportion exhibited non-significant developmental decreases among children diagnosed with Angelman syndrome r(`r myn["l_cp","AS"]`) = `r myr["l_cp","AS"]`, CI [`r mylb["l_cp","AS"]`,`r myhb["l_cp","AS"]`], p=`r myp["l_cp","AS"]`]; and marginal developmental increases among low-risk control r(`r myn["l_cp","LR"]`) = `r myr["l_cp","LR"]`, CI [`r mylb["l_cp","LR"]`,`r myhb["l_cp","LR"]`], p=`r myp["l_cp","LR"]`].
-
-Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(`r myn["z_lp","AS"]`) = `r myr["z_lp","AS"]`, CI [`r mylb["z_lp","AS"]`,`r myhb["z_lp","AS"]`], p=`r myp["z_lp","AS"]`]; whereas low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r myr["z_lp","LR"]`, CI [`r mylb["z_lp","LR"]`,`r myhb["z_lp","LR"]`], p=`r myp["z_lp","LR"]`].
-Similarly, there were non-significant developmental decreases in the Canonical Proportion among children with Angelman syndrome r(`r myn["z_cp","AS"]`) = `r myr["z_cp","AS"]`, CI [`r mylb["z_cp","AS"]`,`r myhb["z_cp","AS"]`], p=`r myp["z_cp","AS"]`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r myr["z_cp","LR"]`, CI [`r mylb["z_cp","LR"]`,`r myhb["z_cp","LR"]`], p=`r myp["z_cp","LR"]`].
+There was a near-zero relationship among the children diagnosed with Angelman syndrome r(`r myn["l_lp","AS"]`) = `r printr(myr["l_lp","AS"])`, CI [`r printr(mylb["l_lp","AS"])`,`r printr(myhb["l_lp","AS"])`], *p* = `r printp(myp["l_lp","AS"])`]; and a significant association among low-risk control children r(`r myn["l_lp","LR"]`) = `r printr(myr["l_lp","LR"])`, CI [`r printr(mylb["l_lp","LR"])`,`r printr(myhb["l_lp","LR"])`], *p* = `r printp(myp["l_lp","LR"])`].
+The Canonical Proportion exhibited non-significant developmental decreases among children diagnosed with Angelman syndrome r(`r myn["l_cp","AS"]`) = `r printr(myr["l_cp","AS"])`, CI [`r printr(mylb["l_cp","AS"])`,`r printr(myhb["l_cp","AS"])`], *p* = `r printp(myp["l_cp","AS"])`]; and marginal developmental increases among low-risk control r(`r myn["l_cp","LR"]`) = `r printr(myr["l_cp","LR"])`, CI [`r printr(mylb["l_cp","LR"])`,`r printr(myhb["l_cp","LR"])`], *p* = `r printp(myp["l_cp","LR"])`].
+
+Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(`r myn["z_lp","AS"]`) = `r printr(myr["z_lp","AS"])`, CI [`r printr(mylb["z_lp","AS"])`,`r printr(myhb["z_lp","AS"])`], *p* = `r printp(myp["z_lp","AS"])`]; whereas low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r printr(myr["z_lp","LR"])`, CI [`r printr(mylb["z_lp","LR"])`,`r printr(myhb["z_lp","LR"])`], *p* = `r printp(myp["z_lp","LR"])`].
+Similarly, there were non-significant developmental decreases in the Canonical Proportion among children with Angelman syndrome r(`r myn["z_cp","AS"]`) = `r printr(myr["z_cp","AS"])`, CI [`r printr(mylb["z_cp","AS"])`,`r printr(myhb["z_cp","AS"])`], *p* = `r printp(myp["z_cp","AS"])`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r printr(myr["z_cp","LR"])`, CI [`r printr(mylb["z_cp","LR"])`,`r printr(myhb["z_cp","LR"])`], *p* = `r printp(myp["z_cp","LR"])`].
 
@@
 These visualizations suggest that performance is moderate to good, which was confirmed via statistical analyses.
-The overall (weighted) accuracy is `r round(mycf$overall["Accuracy"]*100)`%, CI = [`r round(mycf$overall["AccuracyLower"]*100)`,`r round(mycf$overall["AccuracyUpper"]*100)`],  kappa is `r round(mycf$overall["Kappa"],3)`, and the Gwet's AC1 coefficient is `r round(mygac$est,3)`, CI = [`r round(mygac$lb,3)`,`r round(mygac$ub,3)`].
+The overall (weighted) accuracy is `r round(mycf$overall["Accuracy"]*100)`%, CI = [`r round(mycf$overall["AccuracyLower"]*100)`,`r round(mycf$overall["AccuracyUpper"]*100)`],  kappa is `r printr(mycf$overall["Kappa"])`, and the Gwet's AC1 coefficient is `r printr(mygac$est)`, CI = [`r printr(mygac$lb)`,`r printr(mygac$ub)`].
 
@@
 colnames(myr)<-c("AS","LR","all")
-mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
+mylb=myhb=myn=myp =myr #this is an ugly way of creating 5 matrices
 
@@
 We illustrate this association in Figure 5.
-Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
+Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r printr(myr["lp","all"])`, CI [`r printr(mylb["lp","all"])`,`r printr(myhb["lp","all"])`], *p* `r printp(myp["lp","all"])`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r printr(myr["lp","AS"])`, CI [`r printr(mylb["lp","AS"])`,`r printr(myhb["lp","AS"])`], *p* = `r printp(myp["lp","AS"])`]; low-risk control r(`r myn["lp","LR"]`) = `r printr(myr["lp","LR"])`, CI [`r printr(mylb["lp","LR"])`,`r printr(myhb["lp","LR"])`], *p* = `r printp(myp["lp","LR"])`].
 
@@
 
-Similarly, a strong positive correlation is found in the Canonical Proportion r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`].
-When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
+Similarly, a strong positive correlation is found in the Canonical Proportion r(`r myn["cp","all"]`) = `r printr(myr["cp","all"])`, CI [`r printr(mylb["cp","all"])`,`r printr(myhb["cp","all"])`], *p* `r printp(myp["cp","all"])`].
+When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman syndrome: r(`r myn["cp","AS"]`) = `r printr(myr["cp","AS"])`, CI [`r printr(mylb["cp","AS"])`,`r printr(myhb["cp","AS"])`], *p* = `r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r printr(myr["cp","LR"])`, CI [`r printr(mylb["cp","LR"])`,`r printr(myhb["cp","LR"])`], *p* `r printp(myp["cp","LR"])`.
 
@@
 colnames(myr)<-c("lower_Junk","higher_Junk")
-mylb=myhb=myn=myp=myr #this is an ugly way of creating 5 matrices
+mylb=myhb=myn=myp =myr #this is an ugly way of creating 5 matrices
 
@@
 We investigated this hypothesis by calculating the proportion of their data labeled as "Junk" for each individual child.
-There was no significant difference in the proportions of their data labeled as "Junk" for children with Angelman syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the low-risk group (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, p = `r round(JunkXdiag$p.value,3)`.
+There was no significant difference in the proportions of their data labeled as "Junk" for children with Angelman syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the low-risk group (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, *p* = `r printp(JunkXdiag$p.value)`.
 We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk".
 Results were similar across these two post-hoc subgroups (see Figure 6).
-For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["lp","lower_Junk"]`) = `r myr["lp","lower_Junk"]`, CI [`r mylb["lp","lower_Junk"]`,`r myhb["lp","lower_Junk"]`], p=`r myp["lp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["lp","higher_Junk"]`) = `r myr["lp","higher_Junk"]`, CI [`r mylb["lp","higher_Junk"]`,`r myhb["lp","higher_Junk"]`], p=`r myp["lp","higher_Junk"]`.
-For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["cp","lower_Junk"]`) = `r myr["cp","lower_Junk"]`, CI [`r mylb["cp","lower_Junk"]`,`r myhb["cp","lower_Junk"]`], p=`r myp["cp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["cp","higher_Junk"]`) = `r myr["cp","higher_Junk"]`, CI [`r mylb["cp","higher_Junk"]`,`r myhb["cp","higher_Junk"]`], p=`r myp["cp","higher_Junk"]`.
+For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["lp","lower_Junk"]`) = `r printr(myr["lp","lower_Junk"])`, CI [`r printr(mylb["lp","lower_Junk"])`,`r printr(myhb["lp","lower_Junk"])`], *p* = `r printp(myp["lp","lower_Junk"])`]; and for the higher Junk group it was r(`r myn["lp","higher_Junk"]`) = `r printr(myr["lp","higher_Junk"])`, CI [`r printr(mylb["lp","higher_Junk"])`,`r printr(myhb["lp","higher_Junk"])`], *p* `r printp(myp["lp","higher_Junk"])`.
+For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["cp","lower_Junk"]`) = `r printr(myr["cp","lower_Junk"])`, CI [`r printr(mylb["cp","lower_Junk"])`,`r printr(myhb["cp","lower_Junk"])`], *p* `r printp(myp["cp","lower_Junk"])`]; and for the higher Junk group it was r(`r myn["cp","higher_Junk"]`) = `r printr(myr["cp","higher_Junk"])`, CI [`r printr(mylb["cp","higher_Junk"])`,`r printr(myhb["cp","higher_Junk"])`], *p* `r printp(myp["cp","higher_Junk"])`.
 Thus, it does not seem that a higher proportion of "Junk" judgments is an index of low quality data.
@@
 
-Above, we concluded that derived metrics integrating information across audio clips (linguistic and canonical proportion, which can be derived from segment- or chunk-level judgments) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA^TM^ segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
+Above, we concluded that derived metrics integrating information across audio clips (Linguistic and Canonical Proportion, which can be derived from segment- or chunk-level judgments) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA^TM^ segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
 As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
@@
 
-ling_prop=(cans+noncans)/(sum_bychild-Junk)
-can_prop=cans/(cans+noncans)
+ling_prop =(cans+noncans)/(sum_bychild-Junk)
+can_prop =cans/(cans+noncans)
 
@@
 
-Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman syndrome group: r(`r z_lp_as$parameter`) = `r z_lp_as$estimate`, CI [`r z_lp_as$conf.int[1]`,`r z_lp_as$conf.int[2]`], p=`r z_lp_as$p.value`], but increased with age  in the low-risk control group: r(`r z_lp_lr$parameter`) = `r z_lp_lr$estimate`, CI [`r z_lp_lr$conf.int[1]`,`r z_lp_lr$conf.int[2]`], p=`r z_lp_lr$p.value`.
-Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman syndrome r(`r z_cp_as$parameter`) = `r z_cp_as$estimate`, CI [`r z_cp_as$conf.int[1]`,`r z_cp_as$conf.int[2]`], p=`r z_cp_as$p.value`]; and developmental increases among low-risk control r(`r z_cp_lr$parameter`) = `r z_cp_lr$estimate`, CI [`r z_cp_lr$conf.int[1]`,`r z_cp_lr$conf.int[2]`], p=`r z_cp_lr$p.value`.
+Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman syndrome group: r(`r z_lp_as$parameter`) = `r printr(z_lp_as$estimate)`, CI [`r printr(z_lp_as$conf.int[1])`,`r printr(z_lp_as$conf.int[2])`], *p* = `r printp(z_lp_as$p.value)`], but increased with age  in the low-risk control group: r(`r z_lp_lr$parameter`) = `r printr(z_lp_lr$estimate)`, CI [`r printr(z_lp_lr$conf.int[1])`,`r printr(z_lp_lr$conf.int[2])`], *p* = `r printp(z_lp_lr$p.value)`.
+Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman syndrome r(`r z_cp_as$parameter`) = `r printr(z_cp_as$estimate)`, CI [`r printr(z_cp_as$conf.int[1])`,`r printr(z_cp_as$conf.int[2])`], *p* = `r printp(z_cp_as$p.value)`]; and developmental increases among low-risk control r(`r z_cp_lr$parameter`) = `r printr(z_cp_lr$estimate)`, CI [`r printr(z_cp_lr$conf.int[1])`,`r printr(z_cp_lr$conf.int[2])`], *p* = `r printp(z_cp_lr$p.value)`.
 
@@
 
-As for correlations between Zooniverse- and laboratory-derived metrics (e.g., linguistic proportion as estimated using Zooniverse coding, versus laboratory coding), we observe very similar levels of correlation: Linguistic Proportion overall r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].
-As for Canonical Proportion overall r(`r myn["cp","all"]`) = `r myr["cp","all"]`, CI [`r mylb["cp","all"]`,`r myhb["cp","all"]`], p=`r myp["cp","all"]`; Angelman syndrome: r(`r myn["cp","AS"]`) = `r myr["cp","AS"]`, CI [`r mylb["cp","AS"]`,`r myhb["cp","AS"]`], p=`r myp["cp","AS"]`];  low-risk control infants r(`r myn["cp","LR"]`) = `r myr["cp","LR"]`, CI [`r mylb["cp","LR"]`,`r myhb["cp","LR"]`], p=`r myp["cp","LR"]`.
-
-Since all of these results are very similar to those obtained when collapsing chunk-level judgments to generate segment-level judgments, we conclude that in the future this step may not be necessary.
-Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' chunk-level judgments [which was in fact what @cychosz2019canonical did].
+As for correlations between Zooniverse- and laboratory-derived metrics (e.g., linguistic proportion as estimated using Zooniverse coding, versus laboratory coding), we observe very similar levels of correlation: Linguistic Proportion overall r(`r myn["lp","all"]`) = `r printr(myr["lp","all"])`, CI [`r printr(mylb["lp","all"])`,`r printr(myhb["lp","all"])`], *p* `r printp(myp["lp","all"])`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r printr(myr["lp","AS"])`, CI [`r printr(mylb["lp","AS"])`,`r printr(myhb["lp","AS"])`], *p* = `r printp(myp["lp","AS"])`]; low-risk control r(`r myn["lp","LR"]`) = `r printr(myr["lp","LR"])`, CI [`r printr(mylb["lp","LR"])`,`r printr(myhb["lp","LR"])`], *p* = `r printp(myp["lp","LR"])`].
+As for Canonical Proportion overall r(`r myn["cp","all"]`) = `r printr(myr["cp","all"])`, CI [`r printr(mylb["cp","all"])`,`r printr(myhb["cp","all"])`], *p* `r printp(myp["cp","all"])`; Angelman syndrome: r(`r myn["cp","AS"]`) = `r printr(myr["cp","AS"])`, CI [`r printr(mylb["cp","AS"])`,`r printr(myhb["cp","AS"])`], *p* = `r printp(myp["cp","AS"])`];  low-risk control infants r(`r myn["cp","LR"]`) = `r printr(myr["cp","LR"])`, CI [`r printr(mylb["cp","LR"])`,`r printr(myhb["cp","LR"])`], *p* `r printp(myp["cp","LR"])`.
+
+Since all of these results are very similar to those obtained when collapsing chunk-level judgments to generate segment-level judgments, we conclude that in the future this step can be skipped.
+Instead, researchers can derive Linguistic and Canonical Proportions directly from citizen scientists' chunk-level judgments [which was in fact what @cychosz2019canonical did].
 
@@
   ztab=table(sampling$ChildID,sampling$Zoon_classif)
-  z_lp=rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
-  z_cp=ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
+  z_lp =rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
+  z_cp =ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
 
   ltab=table(sampling$ChildID,sampling$lab)
-  l_lp=rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
-  l_cp=ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
+  l_lp =rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
+  l_cp =ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
 
@@
 Next, we looked at whether having more segments from each child may lead to more reliable metrics.
-We observed a non-significant trend [t(`r nseg_t$parameter`)=`r nseg_t$estimate`] for lower number of segments in the Angelman syndrome group (mean `r nseg_t$estimate[1]`, range `r min(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`) than among low-risk control infants (mean `r nseg_t$estimate[2]`, range `r min(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`), likely because of the lower volubility of the former children.
+We observed a non-significant trend [Welch t(`r nseg_t$parameter`)=`r nseg_t$statistic`] for lower number of segments in the Angelman syndrome group (M = `r nseg_t$estimate[1]`, range `r min(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`) than among low-risk control infants (M = `r nseg_t$estimate[2]`, range `r min(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`), likely because of the lower volubility of the former children.
 Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications [@cychosz2019canonical].
@@
 We repeated this process 50 times, to assess the extent to which the association between laboratory- and Zooniverse-derived metrics varied in each random sample (e.g., the correlation between Linguistic Proportion as estimated using Zooniverse coding on the one hand, and Linguistic Proportion as estimated using laboratory coding on the other).
+
 The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table 2), particularly for the low-risk younger infants, and for Linguistic Proportion.
@@
 We also observed quite a bit of variance across the 50 runs.
-For Linguistic Proportion, the range of correlations found for all infants was [`r min(resampling$lp_r_all)`,`r max(resampling$lp_r_all)`]; for the Angelman syndrome group [`r min(resampling$lp_r_as)`,`r max(resampling$lp_r_as)`]; for the low-risk group [`r min(resampling$lp_r_lr)`,`r max(resampling$lp_r_lr)`].
-For Canonical Proportion, the range of correlations found for all infants was [`r min(resampling$cp_r_all)`,`r max(resampling$cp_r_all)`]; for the Angelman syndrome group [`r min(resampling$cp_r_as)`,`r max(resampling$cp_r_as)`]; for the low-risk group [`r min(resampling$cp_r_lr)`,`r max(resampling$cp_r_lr)`].
+For Linguistic Proportion, the range of correlations found for all infants was [`r printp(min(resampling$lp_r_all))`,`r printp(max(resampling$lp_r_all))`]; for the Angelman syndrome group [`r printp(min(resampling$lp_r_as))`,`r printp(max(resampling$lp_r_as))`]; for the low-risk group [`r printp(min(resampling$lp_r_lr))`,`r printp(max(resampling$lp_r_lr))`].
+For Canonical Proportion, the range of correlations found for all infants was [`r printp(min(resampling$cp_r_all))`,`r printp(max(resampling$cp_r_all))`]; for the Angelman syndrome group [`r printp(min(resampling$cp_r_as))`,`r printp(max(resampling$cp_r_as))`]; for the low-risk group [`r printp(min(resampling$cp_r_lr))`,`r printp(max(resampling$cp_r_lr))`].
 
@@
 
-Notwithstanding our enthusiasm, we would like to point out some limitations of our results that can be addressed with more research.
-The first area in which more research would be welcome involves the pre-processing.
+Notwithstanding our enthusiasm, we would like to point out some limitations of our results that should be addressed in future research.
+The first area in which more work would be welcome involves the pre-processing.
 We used LENA^TM^ algorithms to detect vocalizations by the key child, but there are newer, better-performing, open-sourced alternatives that could be used instead [@lavechin2020open].
-Still, performance is most often established on children whose development is thought to be normative, being raised in small households located in urban environments, whereas it would be important to check for accuracy in more diverse samples, where the children have a diagnosis, multiple siblings, and/or are growing up in a rural environment.
+Still, performance is most often established on children whose development is thought to be normative, being raised in small households located in urban environments, whereas it would be important to check for accuracy in more diverse samples, where the children have a diagnosis [e.g., @Dykstra2013], multiple siblings [e.g., @Elo], and/or are growing up in a rural environment [e.g., @cristia2020thorough].
 Speech technology developments would also be welcome to complement citizen scientists' classifications, particularly for categories that may be hard to classify based on short samples.
-**remaining methodological problems with the LENA pipeline (p.
-XX); The team who won the challenge improved this by about 2%, primarily through gains in the laughing class obtained by adding training data [@yeh2019using].
-Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are weighted based on their frequency of occurrence, changes very much.
-**
+There is a great deal of work to be done in this sense.
+The Introduction mentioned a challenge that established a baseline performance with unweighted average recall (UAR) of 54% (when recall is averaged across the 5 classes, Laughing, Crying, Canonical, Non-canonical, and Junk, giving equal weight to each of them rather than weighing them based on their frequency of occurrence).
+As reported in Table 1, our own method leads to a UAR of 57%, so one may think there is not much to gain with the automated method.
+However, notice that the automated method does lead to much better performance than Zooniverse classifications for Crying, and to some improvements in Canonical and Junk.
+For instance, the team who won the challenge in 2019 improved UAR by about 2%, primarily through gains in the laughing class obtained by adding training data [@yeh2019using].
+That state of the art was challenged by @kaya2020combining, who obtained a UAR of 61% on the same data as the challenge, thanks to improvements in all of the classes but for Laughing.
+Thus, there may be some gains to be obtained in specific classes.
 
@@
 Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion.
-We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant for other disorders that are related to emotional disorders [see a review in @halpern2016excessive].
+We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant to emotional disorders [see a review in @halpern2016excessive].
 Additionally, while Canonical Proportion has been studied in much previous work, we believe further research is necessary using larger scales (saliently increasing the sheer number of children studied, to have sufficient power to detect group effects) and with broader coverage [notably including children growing up in a broad range of languages and cultures, as in @cychosz2019canonical], to make sure that assumptions regarding cross-linguistic and maturational universality are amply justified and not merely a result of having low power to detect differences, or simply sampling from similar populations.
@@
 
-**relative dearth of LENA data with atypical populations **
-
 It is also important to note that work on children with non-normative development poses in itself some challenges.
@@
 However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
-Now that this method has been validated, researchers can extract all segments that LENA^TM^ identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.**the relatively small sample sizes of the two groups we studied (p.
-XX);**
-**the kinds of populations that we cannot be certain this work generalizes to (p. XX).**
+Now that this method has been validated, researchers can extract all segments that LENA^TM^ identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
+This approach may be particularly useful in the case of data from children with a diagnosis, whose data tends to be under-represented in LENA^TM^ research, compared to data from normative children.
+We also hope future work makes use of the scalability of the method by employing larger sample sizes than those used in the present study, where data from only 10 children was used.
 
@@
 
-``` {r fig-zoo, comment=F, message=F, hide = T, warning = F, echo = F,  fig.cap="Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needed to make; as well as the Field Guide, which remained accessible on the right of the screen throughout the session, and which allowed users to revise examples of the different categories."}
+``` {r fig-zoo, comment=F, message=F, hide = T, warning = F, echo = F,  fig.cap ="Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needed to make; as well as the Field Guide, which remained accessible on the right of the screen throughout the session, and which allowed users to revise examples of the different categories."}
 knitr::include_graphics("zooniverse-pufig.pdf")
@@
 
-```{r fig-corage, echo=F,fig.cap="Correlations between child-level descriptors and age as a function of metric (Linguistic Proportion in the top row, Canonical Proportion in the bottom row), annotation method, and child group (red=low-risk, and black= Angelman syndrome)."}
+```{r fig-corage, echo=F,fig.cap ="Correlations between child-level descriptors and age as a function of metric (Linguistic Proportion in the top row, Canonical Proportion in the bottom row), annotation method, and child group (red=low-risk, and black= Angelman syndrome)."}
 
@@
 
-```{r fig-prec-rec, echo=F,fig.cap="Number of segments as a function of their majority label from laboratory annotations (rows) versus Zooniverse annotations (columns).
+```{r fig-prec-rec, echo=F,fig.cap ="Number of segments as a function of their majority label from laboratory annotations (rows) versus Zooniverse annotations (columns).
 Percentages (and shading) indicate precision (left) and recall (right).
@@
 
-```{r fig-corlab-zoo, echo=F,fig.cap="Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of child group (red=low-risk, and black=Angelman syndrome)."}
+```{r fig-corlab-zoo, echo=F,fig.cap ="Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of child group (red=low-risk, and black=Angelman syndrome)."}
 
@@
 
-```{r fig-cor-Junk, echo=F,fig.cap="Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of whether children's Junk proportion was higher or lower than the median for all children (blue=lower, and green=higher)."}
+```{r fig-cor-Junk, echo=F,fig.cap ="Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of whether children's Junk proportion was higher or lower than the median for all children (blue=lower, and green=higher)."}
 
@@
 
-read.table("tab-lena-challenge.txt",header=F,sep="\t",skip=1)->tab1
+read.table("tab-lena-challenge.txt",header=F,sep ="\t",skip =1)->tab1
 #olnames(tab1)<-c("","Recall","Recall","Recall","Label Frequency","Label Frequency")
@@
 
-kable(tab1,cap=mycaption, col.names = NULL) %>%
+kable(tab1,cap =mycaption, col.names = NULL) %>%
   collapse_rows(columns = 1:6, valign = "middle") #%>%
@@
 
-kable(mycortab_sorted,cap="Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways.
+kable(mycortab_sorted,cap ="Pearson correlation coefficients across metrics derived from laboratory and Zooniverse annotations in the Angelman syndrome (AS) group data, low-risk (LR) group data, or for all children together (all) in three ways.
 First, All seg indicates that Zooniverse annotations at the chunk level were first combined at the segment level.

### 2021-01-07 202010_jslhr/paper.tex
ready for coauthor feedback

@@
 It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels.
-We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA\^{}TM (see meta-analyses in Wang et al., 2020; Cristia et al., 2020); as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range.
+We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA\textsuperscript{TM} (see meta-analyses in Wang et al., 2020; Cristia et al., 2020); as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range.
 Some previous work had shown that vocal development proceeded differently when compared to a normative sample (Hamrick et al., 2019; Hamrick \& Tonnsen, 2019).
-Generally, studies on vocal development among children with genetic disorders reveal lower volubility and lower prevalence of canonical syllables (Belardi et al., 2017; Rankine et al., 2017; see also {\textbf{???}} for a review of retrospective video analysis of children diagnosed with Autism Spectrum Disorder, Rett Syndrome, or Fragile X).
+Generally, studies on vocal development among children with genetic disorders reveal lower volubility and lower prevalence of canonical syllables (Belardi et al., 2017; Rankine et al., 2017; see also Roche et al., 2018 for a review of retrospective video analysis of children diagnosed with Autism Spectrum Disorder, Rett Syndrome, or Fragile X, documenting these and other vocal characteristics).
 Low volubility is one of the contexts in which daylong recordings may be particularly advantageous over alternative recording methods (see Rankine et al., 2017 for discussion), since it is challenging to measure vocal development accurately based on shorter recordings (as the child may not vocalize at all in a 1-2h recording).
@@
 Zooniverse data revealed a similar distribution: 15\% Canonical, 60\% Non-Canonical, 3\% Laughing, 7\% Crying, 15\% Junk.
-Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical+Non-Canonical)/All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical/(Canonical+Non-Canonical) (i.e., we remove Junk, Crying, and Laughing).
+Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical + Non-Canonical) / All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical / (Canonical + Non-Canonical) (i.e., we remove Junk, Crying, and Laughing).
 Figure 3 shows results of plotting our Proportions against child age.
@@
 Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups.
-There was a near-zero relationship among the children diagnosed with Angelman syndrome r(8) = -0.09, CI {[}-0.68,0.57{]}, p=0.80{]}; and a significant association among low-risk control children r(8) = 0.70, CI {[}0.13,0.92{]}, p=0.02{]}.
-The Canonical Proportion exhibited non-significant developmental decreases among children diagnosed with Angelman syndrome r(8) = -0.39, CI {[}-0.82,0.31{]}, p=0.26{]}; and marginal developmental increases among low-risk control r(8) = 0.52, CI {[}-0.17,0.86{]}, p=0.13{]}.
+There was a near-zero relationship among the children diagnosed with Angelman syndrome r(8) = -.091, CI {[}-.682,.571{]}, \emph{p} = .802{]}; and a significant association among low-risk control children r(8) = .7, CI {[}.126,.923{]}, \emph{p} = .024{]}.
+The Canonical Proportion exhibited non-significant developmental decreases among children diagnosed with Angelman syndrome r(8) = -.393, CI {[}-.82,.314{]}, \emph{p} = .261{]}; and marginal developmental increases among low-risk control r(8) = .516, CI {[}-.168,.865{]}, \emph{p} = .126{]}.
 
-Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(8) = -0.17, CI {[}-0.72,0.51{]}, p=0.63{]}; whereas low-risk control children showed a significant increase with age r(8) = 0.63, CI {[}0.00,0.90{]}, p=0.05{]}.
-Similarly, there were non-significant developmental decreases in the Canonical Proportion among children with Angelman syndrome r(8) = -0.41, CI {[}-0.82,0.30{]}, p=0.24{]}; and marginal developmental increases among low-risk control children, r(8) = 0.51, CI {[}-0.17,0.86{]}, p=0.13{]}.
+Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(8) = -.174, CI {[}-.724,.512{]}, \emph{p} = .630{]}; whereas low-risk control children showed a significant increase with age r(8) = .629, CI {[}-.001,.902{]}, \emph{p} = .051{]}.
+Similarly, there were non-significant developmental decreases in the Canonical Proportion among children with Angelman syndrome r(8) = -.407, CI {[}-.825,.299{]}, \emph{p} = .243{]}; and marginal developmental increases among low-risk control children, r(8) = .514, CI {[}-.171,.864{]}, \emph{p} = .129{]}.
 
@@
 These visualizations suggest that performance is moderate to good, which was confirmed via statistical analyses.
-The overall (weighted) accuracy is 73\%, CI = {[}72,74{]}, kappa is 0.53, and the Gwet's AC1 coefficient is 0.68, CI = {[}0.67,0.69{]}.
+The overall (weighted) accuracy is 73\%, CI = {[}72,74{]}, kappa is .531, and the Gwet's AC1 coefficient is .682, CI = {[}.669,.694{]}.
 
@@
 We illustrate this association in Figure 5.
-Looking at all 20 children together, we found a strong positive correlation r(18) = 0.79, CI {[}0.54,0.92{]}, p=0{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman syndrome r(8) = 0.88, CI {[}0.56,0.97{]}, p=0.00{]}; low-risk control r(8) = 0.83, CI {[}0.42,0.96{]}, p=0.00{]}.
+Looking at all 20 children together, we found a strong positive correlation r(18) = .794, CI {[}.543,.915{]}, \emph{p} \textless{} .001{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman syndrome r(8) = .878, CI {[}.555,.971{]}, \emph{p} = .001{]}; low-risk control r(8) = .828, CI {[}.415,.958{]}, \emph{p} = .003{]}.
 
@@
 
-Similarly, a strong positive correlation is found in the Canonical Proportion r(18) = 0.92, CI {[}0.81,0.97{]}, p=0{]}.
-When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman syndrome: r(8) = 0.68, CI {[}0.09,0.92{]}, p=0.03{]}; than the low-risk control children r(8) = 0.98, CI {[}0.93,1.00{]}, p=0.
+Similarly, a strong positive correlation is found in the Canonical Proportion r(18) = .92, CI {[}.806,.968{]}, \emph{p} \textless{} .001{]}.
+When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman syndrome: r(8) = .681, CI {[}.091,.917{]}, \emph{p} = 0.03{]}; than the low-risk control children r(8) = .984, CI {[}.932,.996{]}, \emph{p} \textless{} .001.
 
@@
 We investigated this hypothesis by calculating the proportion of their data labeled as \enquote{Junk} for each individual child.
-There was no significant difference in the proportions of their data labeled as \enquote{Junk} for children with Angelman syndrome (M = 0.14) compared to the low-risk group (M = 0.16): Welch's t(13.35) = -0.58, p = 0.57.
+There was no significant difference in the proportions of their data labeled as \enquote{Junk} for children with Angelman syndrome (M = 0.14) compared to the low-risk group (M = 0.16): Welch's t(13.35) = -0.58, \emph{p} = .569.
 We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as \enquote{Junk}.
 Results were similar across these two post-hoc subgroups (see Figure 6).
-For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = 0.86, CI {[}0.51,0.97{]}, p=0.00{]}; and for the higher Junk group it was r(8) = 0.90, CI {[}0.61,0.98{]}, p=0.
-For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = 0.96, CI {[}0.82,0.99{]}, p=0{]}; and for the higher Junk group it was r(8) = 0.90, CI {[}0.61,0.98{]}, p=0.
+For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = .863, CI {[}.511,.967{]}, \emph{p} = .001{]}; and for the higher Junk group it was r(8) = .895, CI {[}.608,.975{]}, \emph{p} \textless{} .001.
+For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = .957, CI {[}.824,.99{]}, \emph{p} \textless{} .001{]}; and for the higher Junk group it was r(8) = .895, CI {[}.608,.975{]}, \emph{p} \textless{} .001.
 Thus, it does not seem that a higher proportion of \enquote{Junk} judgments is an index of low quality data.
@@
 
-Above, we concluded that derived metrics integrating information across audio clips (linguistic and canonical proportion, which can be derived from segment- or chunk-level judgments) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA\textsuperscript{TM} segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
+Above, we concluded that derived metrics integrating information across audio clips (Linguistic and Canonical Proportion, which can be derived from segment- or chunk-level judgments) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA\textsuperscript{TM} segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
 As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
@@
 
-Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman syndrome group: r(8) = -0.04, CI {[}-0.65,0.61{]}, p=0.92{]}, but increased with age in the low-risk control group: r(8) = 0.72, CI {[}0.16,0.93{]}, p=0.02.
-Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman syndrome r(8) = -0.38, CI {[}-0.81,0.33{]}, p=0.28{]}; and developmental increases among low-risk control r(8) = 0.57, CI {[}-0.09,0.88{]}, p=0.08.
+Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman syndrome group: r(8) = -.038, CI {[}-.652,.606{]}, \emph{p} = .918{]}, but increased with age in the low-risk control group: r(8) = .716, CI {[}.157,.927{]}, \emph{p} = .020.
+Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman syndrome r(8) = -.38, CI {[}-.815,.328{]}, \emph{p} = .278{]}; and developmental increases among low-risk control r(8) = .573, CI {[}-.088,.884{]}, \emph{p} = .083.
 
-As for correlations between Zooniverse- and laboratory-derived metrics (e.g., linguistic proportion as estimated using Zooniverse coding, versus laboratory coding), we observe very similar levels of correlation: Linguistic Proportion overall r(18) = 0.80, CI {[}0.55,0.92{]}, p=0{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman syndrome r(8) = 0.84, CI {[}0.46,0.96{]}, p=0.00{]}; low-risk control r(8) = 0.79, CI {[}0.33,0.95{]}, p=0.01{]}.
-As for Canonical Proportion overall r(18) = 0.95, CI {[}0.87,0.98{]}, p=0; Angelman syndrome: r(8) = 0.72, CI {[}0.17,0.93{]}, p=0.02{]}; low-risk control infants r(8) = 0.97, CI {[}0.88,0.99{]}, p=0.
+As for correlations between Zooniverse- and laboratory-derived metrics (e.g., linguistic proportion as estimated using Zooniverse coding, versus laboratory coding), we observe very similar levels of correlation: Linguistic Proportion overall r(18) = .798, CI {[}.549,.917{]}, \emph{p} \textless{} .001{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman syndrome r(8) = .845, CI {[}.46,.963{]}, \emph{p} = .002{]}; low-risk control r(8) = .794, CI {[}.328,.949{]}, \emph{p} = .006{]}.
+As for Canonical Proportion overall r(18) = .949, CI {[}.873,.98{]}, \emph{p} \textless{} .001; Angelman syndrome: r(8) = .724, CI {[}.173,.93{]}, \emph{p} = .018{]}; low-risk control infants r(8) = .972, CI {[}.881,.993{]}, \emph{p} \textless{} .001.
 
-Since all of these results are very similar to those obtained when collapsing chunk-level judgments to generate segment-level judgments, we conclude that in the future this step may not be necessary.
-Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' chunk-level judgments (which was in fact what Cychosz et al., 2019 did).
+Since all of these results are very similar to those obtained when collapsing chunk-level judgments to generate segment-level judgments, we conclude that in the future this step can be skipped.
+Instead, researchers can derive Linguistic and Canonical Proportions directly from citizen scientists' chunk-level judgments (which was in fact what Cychosz et al., 2019 did).
 
 Next, we looked at whether having more segments from each child may lead to more reliable metrics.
-We observed a non-significant trend {[}t(17.88)=348.70 and 375.50{]} for lower number of segments in the Angelman syndrome group (mean 348.70, range 163-599) than among low-risk control infants (mean 375.50, range 224-627), likely because of the lower volubility of the former children.
+We observed a non-significant trend {[}Welch t(17.88)=-0.53{]} for lower number of segments in the Angelman syndrome group (M = 348.70, range 163-599) than among low-risk control infants (M = 375.50, range 224-627), likely because of the lower volubility of the former children.
 Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications (Cychosz et al., 2019).
@@
 We also observed quite a bit of variance across the 50 runs.
-For Linguistic Proportion, the range of correlations found for all infants was {[}0.44,0.89{]}; for the Angelman syndrome group {[}0.49,0.97{]}; for the low-risk group {[}0.29,0.99{]}.
-For Canonical Proportion, the range of correlations found for all infants was {[}0.75,0.94{]}; for the Angelman syndrome group {[}0.79,0.99{]}; for the low-risk group {[}0.55,0.94{]}.
+For Linguistic Proportion, the range of correlations found for all infants was {[}.442,.893{]}; for the Angelman syndrome group {[}.490,.969{]}; for the low-risk group {[}.289,.989{]}.
+For Canonical Proportion, the range of correlations found for all infants was {[}.750,.943{]}; for the Angelman syndrome group {[}.790,.994{]}; for the low-risk group {[}.547,.939{]}.
 
@@
 
-Notwithstanding our enthusiasm, we would like to point out some limitations of our results that can be addressed with more research.
-The first area in which more research would be welcome involves the pre-processing.
-We used LENA\textsuperscript{TM} algorithms to detect vocalizations by the key child, but there are newer, better-performing, open-sourced alternatives that could be used instead ({\textbf{???}}).
-Still, performance is most often established on children whose development is thought to be normative, being raised in small households located in urban environments, whereas it would be important to check for accuracy in more diverse samples, where the children have a diagnosis, multiple siblings, and/or are growing up in a rural environment.
+Notwithstanding our enthusiasm, we would like to point out some limitations of our results that should be addressed in future research.
+The first area in which more work would be welcome involves the pre-processing.
+We used LENA\textsuperscript{TM} algorithms to detect vocalizations by the key child, but there are newer, better-performing, open-sourced alternatives that could be used instead (Lavechin, Bousbib, Bredin, Dupoux, \& Cristia, 2020).
+Still, performance is most often established on children whose development is thought to be normative, being raised in small households located in urban environments, whereas it would be important to check for accuracy in more diverse samples, where the children have a diagnosis (e.g., Dykstra et al., 2013), multiple siblings (e.g., Elo, 2016), and/or are growing up in a rural environment (e.g., Cristia, Lavechin, et al., 2020).
 Speech technology developments would also be welcome to complement citizen scientists' classifications, particularly for categories that may be hard to classify based on short samples.
-\textbf{remaining methodological problems with the LENA pipeline (p.~XX); The team who won the challenge improved this by about 2\%, primarily through gains in the laughing class obtained by adding training data (Yeh et al., 2019).
-Given that the laughing class is very rare, improvements in this class do not mean that performance as a whole, when all classes are weighted based on their frequency of occurrence, changes very much.
-}
+There is a great deal of work to be done in this sense.
+The Introduction mentioned a challenge that established a baseline performance with unweighted average recall (UAR) of 54\% (when recall is averaged across the 5 classes, Laughing, Crying, Canonical, Non-canonical, and Junk, giving equal weight to each of them rather than weighing them based on their frequency of occurrence).
+As reported in Table 1, our own method leads to a UAR of 57\%, so one may think there is not much to gain with the automated method.
+However, notice that the automated method does lead to much better performance than Zooniverse classifications for Crying, and to some improvements in Canonical and Junk.
+For instance, the team who won the challenge in 2019 improved UAR by about 2\%, primarily through gains in the laughing class obtained by adding training data (Yeh et al., 2019).
+That state of the art was challenged by Kaya, Verkholyak, Markitantov, and Karpov (2020), who obtained a UAR of 61\% on the same data as the challenge, thanks to improvements in all of the classes but for Laughing.
+Thus, there may be some gains to be obtained in specific classes.
 
@@
 Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion.
-We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant for other disorders that are related to emotional disorders (see a review in Halpern \& Coelho, 2016).
+We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant to emotional disorders (see a review in Halpern \& Coelho, 2016).
 Additionally, while Canonical Proportion has been studied in much previous work, we believe further research is necessary using larger scales (saliently increasing the sheer number of children studied, to have sufficient power to detect group effects) and with broader coverage (notably including children growing up in a broad range of languages and cultures, as in Cychosz et al., 2019), to make sure that assumptions regarding cross-linguistic and maturational universality are amply justified and not merely a result of having low power to detect differences, or simply sampling from similar populations.
@@
 
-\textbf{relative dearth of LENA data with atypical populations }
-
 It is also important to note that work on children with non-normative development poses in itself some challenges.
@@
 However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
-Now that this method has been validated, researchers can extract all segments that LENA\textsuperscript{TM} identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.\textbf{the relatively small sample sizes of the two groups we studied (p.~XX);}
-\textbf{the kinds of populations that we cannot be certain this work generalizes to (p.~XX).}
+Now that this method has been validated, researchers can extract all segments that LENA\textsuperscript{TM} identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
+This approach may be particularly useful in the case of data from children with a diagnosis, whose data tends to be under-represented in LENA\textsuperscript{TM} research, compared to data from normative children.
+We also hope future work makes use of the scalability of the method by employing larger sample sizes than those used in the present study, where data from only 10 children was used.
 
@@
 
+\leavevmode\hypertarget{ref-Dykstra2013}{}%
+Dykstra, J.
+R., Sabatos-Devito, M.
+G., Irvin, D.
+W., Boyd, B.
+A., Hume, K.
+A., \& Odom, S.
+L.
+(2013).
+Using the Language Environment Analysis (LENA) system in preschool classrooms with children with autism spectrum disorders.
+\emph{Autism}, \emph{17}(5), 582--594.
+\url{https://doi.org/https://doi.org/10.1177/1362361312446206}
+
+\leavevmode\hypertarget{ref-Elo}{}%
+Elo, H. (2016). \emph{Acquiring language as a twin}. Tampere, Finland: Tampere University Press.
+
 \leavevmode\hypertarget{ref-fasolo2008babbling}{}%
@@
 Heckman, J. J. (2006). Skill formation and the economics of investing in disadvantaged children. \emph{Science}, \emph{312}(5782), 1900--1902.
+
+\leavevmode\hypertarget{ref-kaya2020combining}{}%
+Kaya, H., Verkholyak, O., Markitantov, M., \& Karpov, A.
+(2020).
+Combining clustering and functionals based acoustic feature representations for classification of baby sounds.
+In \emph{Companion publication of the 2020 international conference on multimodal interaction} (pp.
+509--513).
+
+\leavevmode\hypertarget{ref-lavechin2020open}{}%
+Lavechin, M., Bousbib, R., Bredin, H., Dupoux, E., \& Cristia, A.
+(2020).
+An open-source voice type classifier for child-centered daylong recordings.
+\emph{Proceedings of Interspeech}.
+\url{https://doi.org/https://arxiv.org/abs/2005.12656v2}
 
@@
 \emph{Journal of Autism and Developmental Disorders}, \emph{47}(6), 1605--1617.
+
+\leavevmode\hypertarget{ref-roche2018early}{}%
+Roche, L., Zhang, D., Bartl-Pokorny, K.
+D., Pokorny, F.
+B., Schuller, B.
+W., Esposito, G., \ldots{} others.
+(2018).
+Early vocal development in autism spectrum disorder, rett syndrome, and fragile x syndrome: Insights from studies using retrospective video analysis.
+\emph{Advances in Neurodevelopmental Disorders}, \emph{2}(1), 49--61.
 

### 2021-01-07 data_analyses/code/recover_zooniverse_metadata.py
convert metadata generated with Chiara's script to the format generated by child-project

@@ -1,0 +1,39 @@
+import pandas as pd 
+import os 
+import sys
+import json
+
+source = sys.argv[1]
+destination = sys.argv[2]
+
+subjects = pd.read_csv('data_analyses/maturity-of-baby-sounds-subjects.csv')
+links = pd.read_csv('data_analyses/metadata/filename_links.csv')
+links.rename(columns = {'FileName': 'recording'}, inplace = True)
+
+def get_name(s):
+    try:
+        return os.path.splitext(json.loads(s)['Name'])[0]
+    except:
+        return 'NA'
+
+subjects['AudioData'] = subjects['metadata'].apply(get_name).astype(str)
+
+df = pd.read_csv(source, dtype = {'AudioData': str})
+df = df.merge(subjects, left_on = 'AudioData', right_on = 'AudioData')
+df = df.merge(links, left_on = 'ChildID', right_on = 'ChildID')
+
+df['onset'] = (df['onset']*1000).astype(int)
+df['offset'] = (df['offset']*1000).astype(int)
+
+df['onset'] = df['onset'] + 500 * df['chunk_pos']
+df['offset'] = df['onset'] + 500
+
+df['speaker_type'] = 'CHI'
+df['wav'] = 'NA'
+df['mp3'] = 'NA'
+df['date_extracted'] = 'NA'
+df['uploaded'] = True
+
+df.rename(columns = {'subject_id': 'zooniverse_id'}, inplace = True)
+
+df.to_csv(destination, index = False)

### 2021-01-28 202010_jslhr/paper.Rmd
cleaned up pipeline in progress

@@
 printr=function(x) sub("0.", ".", round(x,3))
+
+#document session info
+
+sink("lastknit_session_info.txt")
+sessionInfo()
+sink()
 ```
@@
 ```{r read-in}
+x <- read.csv("../data_analyses/output/metadata_all_PU.csv",header=T,sep =",")
+y <- read.csv("../data_analyses/files_from_zooniverse/zooniverse_data_all_final.csv",header=T,sep =",")
+x$AudioData=paste0(x$AudioData,".mp3")
+merge(x,y,by="AudioData")->chunks
+
+chunks$chunkID=paste(chunks$ChildID,chunks$onset)
+
+#merge with children's info
+demo_data <- read.csv("../data_analyses/files_from_elsewhere/demo-data.tsv",header=T,sep ="\t")
+merge(demo_data,chunks,by="ChildID")->chunks
+
+write.csv(chunks,"../data_analyses/output/chunks_individual_judgments.csv")
+
+# next version collapses across chunks to get the nb of different classifications
+table(chunks$chunkID,chunks$Answer)->mytab
+mytab_mj <- mytab>=3  #set as True the answer types that have 3 or more votes
+mymj <- t(t(apply(mytab_mj, 1, function(u) paste( names(which(u)), collapse=", " )))) #get the answer that has majority vote
+nvotes<- t(t(apply(mytab, 1, max))) #get the number of votes for that 
+maj_jud=cbind(row.names(mytab),mymj,nvotes)
+colnames(maj_jud)<-c("chunkID","Answer","nvotes")
+maj_jud2=as.data.frame(maj_jud)
+maj_jud2$ChildID = gsub(" .*","",maj_jud2$chunkID)
+
+#add participant information
+merge(demo_data,maj_jud2,by="ChildID")->maj_jud
+
+write.csv(maj_jud,"../data_analyses/output/chunks_maj_judgments.csv")
+
+#add segment-level info from the lab
+#TODO !!!
+
 # read datasets
-
 data_ang <- read.csv("segment_AS.csv",header=T,sep =",")

### 2021-01-28 data_analyses/code/recover_zooniverse_metadata.py

@@ -7,8 +7,8 @@ import argparse
 parser = argparse.ArgumentParser(description = '')
 parser.add_argument('source', help = 'path to the source metadata to convert')
 parser.add_argument('destination', help = 'destination path')
-parser.add_argument('--subjects', help = 'path to the Zooniverse .csv subjects export', default = 'data_analyses/maturity-of-baby-sounds-subjects.csv')
-parser.add_argument('--links', help = 'path to the dataframe linking ChildID to FileName', default = 'data_analyses/metadata/filename_links.csv')
+parser.add_argument('--subjects', help = 'path to the Zooniverse .csv subjects export', default = 'data_analyses/files_from_zooniverse/maturity-of-baby-sounds-subjects.csv')
+parser.add_argument('--links', help = 'path to the dataframe linking ChildID to FileName', default = 'data_analyses/files_from_elsewhere/filename_links.csv')
 
 args = parser.parse_args()
 

### 2021-01-28 202010_jslhr/paper.Rmd
nearly done with new pipeline

--- 2021-01-07 202010_jslhr/paper.Rmd
@@
 ```{r read-in}
+
 # read datasets
-
 data_ang <- read.csv("segment_AS.csv",header=T,sep =",")

### 2021-01-28 202010_jslhr/generate_jslhr_data.R

@@ -1,0 +1,83 @@
+#TODO rather than removing them, take the first 5 judgments
+
+x <- read.csv("../data_analyses/output/metadata_all_PU.csv",header=T,sep =",")
+y <- read.csv("../data_analyses/files_from_zooniverse/zooniverse_data_all_final.csv",header=T,sep =",")
+
+# Remove weird answers, which do not belong here
+y$Answer<-factor(y$Answer,levels=c("Canonical","Non-Canonical","Crying","Laughing","Junk"))
+summary(y$Answer)
+y=y[!is.na(y$Answer),]
+
+# Remove additional votes when there are more than 5? I'm not sure it's the case, perhaps these are repeated names?
+table(y$AudioData)->sumvotes
+sum(sumvotes>5)
+#rownames(sumvotes)[sumvotes>5]
+
+x$AudioData=paste0(x$AudioData,".mp3")
+merge(x,y,by="AudioData")->chunks
+
+chunks$chunkID=paste(chunks$ChildID,chunks$onset)
+
+#merge with children's info
+demo_data <- read.csv("../data_analyses/files_from_elsewhere/demo-data.tsv",header=T,sep ="\t")
+merge(demo_data,chunks,by="ChildID")->chunks
+
+write.csv(chunks,"../data_analyses/output/chunks_individual_judgments.csv")
+
+# next version collapses across chunks to get the nb of different classifications
+table(chunks$chunkID,chunks$Answer)->mytab
+sumvotes<- t(t(apply(mytab, 1, sum))) #get the total number of votes
+
+#TODO rather than removing them, take the first 5 judgments
+#some clips have been voted too many times --> remove them from consideration
+row.names(mytab)[sumvotes>5]->exclude
+chunks[!(chunks$chunkID %in% exclude),]->chunks
+
+#redo the table & the stats
+table(chunks$chunkID,chunks$Answer)->mytab
+nvotes<- t(t(apply(mytab, 1, max))) #get the number of votes for that 
+
+mytab_mj <- mytab >= 3  #set as True the answer types that have more than 3 votes
+mymj <- t(t(apply(mytab_mj, 1, function(u) paste( names(which(u)), collapse=", " )))) #get the answer that has majority vote
+
+maj_jud=data.frame(cbind(row.names(mytab),unlist(mymj),nvotes))
[…]

### 2021-01-29 data_analyses/code/recover_zooniverse_metadata.py
filename_links update

@@ -12,9 +12,12 @@ parser.add_argument('--links', help = 'path to the dataframe linking ChildID to 
 
 args = parser.parse_args()
 
+print("Loading Zooniverse subjects")
 subjects = pd.read_csv(args.subjects)
+
+print("Loading ChildID <-> recording map")
 links = pd.read_csv(args.links)
-links.rename(columns = {'FileName': 'recording'}, inplace = True)
+links.rename(columns = {'filename': 'recording'}, inplace = True)
 
 def get_name(s):
     try:
@@ -24,9 +27,10 @@ def get_name(s):
 
 subjects['AudioData'] = subjects['metadata'].apply(get_name).astype(str)
 
-df = pd.read_csv(args.source, dtype = {'AudioData': str})
+print("Loading local metadata export")
+df = pd.read_csv(args.source, dtype = {'AudioData': str, 'Age': str})
 df = df.merge(subjects, left_on = 'AudioData', right_on = 'AudioData')
-df = df.merge(links, left_on = 'ChildID', right_on = 'ChildID')
+df = df.merge(links[['ChildID', 'recording']], left_on = 'ChildID', right_on = 'ChildID')
 
 df['onset'] = (df['onset']*1000).astype(int)
 df['offset'] = (df['offset']*1000).astype(int)

### 2021-02-09 202010_jslhr/generate_jslhr_data.R
r pipeline with notes for real

@@ -1,107 +1,167 @@
+## PHASE 1 -- RESTRICT ZOONIVERSE DATA TO PU DATA
+
 library(rjson)
-#file that has "subject" information -- in Zooniverse terms, that is info about the file that was classified
-zoosj <- read.csv("../data_analyses/output/metadata_all_PU.csv",header=T,sep =",")
+getname2=function(x) fromJSON(x)[[1]]
+
+#original chunk metadata
+origmetadata <- read.csv("../data_analyses/files_from_elsewhere/metadata_all_PU.csv",colClasses="factor",header=T,sep =",")
+dim(origmetadata)
+origmetadata$filename=paste0(origmetadata$AudioData,".mp3")
+length(levels(factor(origmetadata$filename)))
+#33730
+
+# #file that has "subject" information -- in Zooniverse terms, that is info about the file that was classified
+zoosj <- read.csv("../data_analyses/files_from_zooniverse/maturity-of-baby-sounds-subjects.csv",header=T,sep =",")
+zoosj$filename=unlist(lapply(as.character(zoosj[,"metadata"]), getname2))
+length(levels(factor(zoosj$filename)))
+#154063
+sum(levels(factor(zoosj$filename)) %in% levels(factor(origmetadata$filename)))
+#33730
+
+merge(origmetadata,zoosj,by="filename")->zoosj_PU
+dim(zoosj_PU)
+#33881 because some of the names are repeated
+mytab=table(zoosj_PU$filename)
+length(names(mytab[mytab==2]))
+#exploration revealed that one batch of 151 was uploaded twice -- removing the second upload
+for(mych in names(mytab[mytab==2])){
+  second_upload_id=max(zoosj_PU[zoosj_PU$filename==mych,"subject_id"])
+  zoosj_PU=zoosj_PU[zoosj_PU$subject_id!=second_upload_id,]
+}
+dim(zoosj_PU)
+#33730
+
+write.csv(zoosj_PU,"../data_analyses/output/zoo_subj_info_pu.csv")
+
+read.csv("../data_analyses/output/zoo_subj_info_pu.csv",colClasses="factor",header=T,sep =",")->zoosj
 
 #file that has the classifications (ie cit sci answers)
-classifs <- read.csv("../data_analyses/files_from_zooniverse/zooniverse_data_all_final.csv",header=T,sep =",")
[…]

### 2021-02-11 202010_jslhr/paper.Rmd
ms broken (bad practice alert)

@@
 
-# read datasets
-data_ang <- read.csv("segment_AS.csv",header=T,sep =",")
-data_td <- read.csv("segment_LRC.csv")
-data_all<-rbind(data_ang, data_td)
-
-#remove the word mixed that takes up space and is unnecessary
-data_all$Zoon_classif=factor(gsub("Mixed_","",as.character(data_all$Zoon_classif),fixed=T))
-#relevel the factor so that it's easier to read
-data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=c("Canonical","Non-Canonical",                   "Crying","Laughing","Junk",levels(data_all$Zoon_classif)[grep("_",levels(data_all$Zoon_classif))]))
+read.csv("../data_analyses/output/key_info.csv")->x
+rownames(x)<-x$X
+
+read.csv("../data_analyses/output/chunks_maj_judgments.csv")->chunks
+read.csv("../data_analyses/output/zoo_lab_maj_judgments.csv")->data_all
+
+#use better names
+data_all$Zoon_classif=data_all$Answer
+
 # create lab column with easier to read correspondance
@@
 
-A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis, and it is beyond the scope of this paper to provide a full summary of this work [but see **XX ADD REFERENCE** for a recent summary].
+A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis, and it is beyond the scope of this paper to provide a full summary of this work.
 Moreover, the way in which vocal development has been studied in the past may not easily translate to long-form recordings.
@@
 
-An impressive total of 4,825 individual Zooniverse users provided labels for the Maturity of Baby Sounds project, of which the present data set is one part.
-For this project, we collected a total of 169,767 judgments provided for 33,731 500-ms chunks, corresponding to 11,980 LENA^TM^ segments.
-Nearly a fifth of chunks did not have at least 3 labels in agreement out of the 5 Zooniverse labels (N = 6,585, 19% of all chunks).
-Of the chunks without a majority agreement, 4341 (66%) contained one or two Junk judgments (out of 5), 6523 (99,9%) had at least two matching judgments (the threshold used for lab-annotated segments), and only 61 (0,01%) had 5 different judgments.
-Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the 81% of chunks that did have at least 3 labels in agreement; this represented 135,725 labels for 27,145 chunks, corresponding to 11,593 LENA^TM^ segments.
-As the segments average 1.12 seconds in length, this means about 3.8 hours of audio data were annotated by 8 different annotators (3 in the laboratory, 5 on Zooniverse).
+An impressive total of `r x["nusers","x"]` individual Zooniverse users provided `r x["njudgments","x"]` judgments on `r x["nchunks","x"]` 500-ms chunks, corresponding to `r dim(data_all)[1]` 11,980 LENA^TM^ segments.
+For a variety of technical mishaps, XX chunks had to be excluded, leading to `r length(levels(factor(chunks$subject_id)))` chunks and `r length(levels(factor(data_all$segmentId_DB)))` LENA^TM^ segments for which we had five Zooniverse labels.
+Among these, `r round(sum(chunks$nvotes<2)/dim(chunks)[1]*100,1)`% had no matching labels, `r round(sum(chunks$nvotes==2)/dim(chunks)[1]*100,1)`% had two matching labels (the threshold used for lab-annotated segments), and `r round(sum(chunks$nvotes>=3)/dim(chunks)[1]*100,1)`% had three or more matching labels, or majority agreement.
+Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the `r round(sum(chunks$nvotes>=3)/dim(chunks)[1]*100,1)`%  of chunks that *did* have at least three labels in agreement (i.e., majority agreement).
+This represented `r sum(chunks$nvotes>=3)*5` 135,725 labels for `r sum(chunks$nvotes>=3)` 27,145 chunks, corresponding to `r dim(data_all)[1]` LENA^TM^ segments.
+Given that each chunk is 500 ms in length, this means about `r (sum(chunks$nvotes>=3)*.5)/60/60` hours of audio data were annotated by 8 different annotators (3 in the laboratory, 5 on Zooniverse).
 
@@
 data_all[data_all$Num_Agreement>2,]->data_all
-
-# we map the mixed
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical"]<-"Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical_Crying"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Crying"]<-"Crying"
-data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Crying"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing_Crying"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Canonical"]<-"Canonical"
-# +
-data_all$Zoon_classif[data_all$Zoon_classif=="Canonical_Crying"]<-"Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Canonical_Laughing"]<-"Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical_Crying"]<-"Non-Canonical"
-
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Laughing"]<-"Crying"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Canonical_Laughing"]<-"Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Laughing_Non-Canonical"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Non-Canonical"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Non-Canonical_Laughing"]<-"Non-Canonical" 
-data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing"]<-"Non-Canonical"
 

### 2021-02-11 202010_jslhr/generate_jslhr_data.R

@@ -1,183 +1,0 @@
-## PHASE 1 -- RESTRICT ZOONIVERSE DATA TO PU DATA
-
-library(rjson)
-getname2=function(x) fromJSON(x)[[1]]
-
-#original chunk metadata
-origmetadata <- read.csv("../data_analyses/files_from_elsewhere/metadata_all_PU.csv",colClasses="factor",header=T,sep =",")
-dim(origmetadata)
-origmetadata$filename=paste0(origmetadata$AudioData,".mp3")
-length(levels(factor(origmetadata$filename)))
-#33730
-
-# #file that has "subject" information -- in Zooniverse terms, that is info about the file that was classified
-zoosj <- read.csv("../data_analyses/files_from_zooniverse/maturity-of-baby-sounds-subjects.csv",header=T,sep =",")
-zoosj$filename=unlist(lapply(as.character(zoosj[,"metadata"]), getname2))
-length(levels(factor(zoosj$filename)))
-#154063
-sum(levels(factor(zoosj$filename)) %in% levels(factor(origmetadata$filename)))
-#33730
-
-merge(origmetadata,zoosj,by="filename")->zoosj_PU
-dim(zoosj_PU)
-#33881 because some of the names are repeated
-mytab=table(zoosj_PU$filename)
-length(names(mytab[mytab==2]))
-#exploration revealed that one batch of 151 was uploaded twice -- removing the second upload
-for(mych in names(mytab[mytab==2])){
-  second_upload_id=max(zoosj_PU[zoosj_PU$filename==mych,"subject_id"])
-  zoosj_PU=zoosj_PU[zoosj_PU$subject_id!=second_upload_id,]
-}
-dim(zoosj_PU)
-#33730
-
-write.csv(zoosj_PU,"../data_analyses/output/zoo_subj_info_pu.csv")
-
-read.csv("../data_analyses/output/zoo_subj_info_pu.csv",colClasses="factor",header=T,sep =",")->zoosj
-
-#file that has the classifications (ie cit sci answers)
-classifs <- read.csv("../data_analyses/files_from_zooniverse/maturity-of-baby-sounds-classifications.csv",header=T,sep =",")
-classifs$subject_id=fromJSON(as.character(classifs[,"subject_data"]))[[1]]$retired$subject_id
-length(levels(factor(classifs$subject_id)))
-#79162
-classifs -> full_classifs
[…]

### 2021-02-14 202010_jslhr/preprocess.R
missing files from previous commit

@@ -1,0 +1,137 @@
+## PHASE 1 -- clean classifications
+
+read.csv("../data_analyses/output/zoo_anno_info_pu.csv",colClasses="factor",header=T,sep =",")->classifs
+
+library(rjson)
+getval=function(x) fromJSON(x)[[1]]$value
+getname=function(x) fromJSON(x)[[1]]$Name
+
+
+## Cleaning stuff up ##
+classifs$Answer=unlist(lapply(as.character(classifs[,"annotations"]), getval))
+table(classifs$Answer) #no weird answers
+
+classifs$filename=unlist(lapply(as.character(classifs[,"subject_data"]), getname))
+length(levels(factor(classifs$filename)))
+#33730 chunks
+
+# Remove additional votes when there are more than 5, 
+# and remove clips with fewer than 5
+table(classifs$filename)->sumvotes
+toomany=rownames(sumvotes)[sumvotes>5]
+length(toomany) #271 have too many votes
+justright=rownames(sumvotes)[sumvotes==5]
+length(justright) #33458 have the right amount
+toofew=rownames(sumvotes)[sumvotes<5]
+length(toofew)#1 chunk didn't receive 5 votes
+
+classifs$nclasschunk=NA
+clean=classifs[classifs$filename %in% justright,] #start by adding the normal ones
+for(thischunk in toomany){
+  classifs$nclasschunk[classifs$filename==thischunk]<-1:sum(classifs$filename==thischunk)
+  clean=rbind(clean,classifs[classifs$filename==thischunk & classifs$nclasschunk<=5 & !is.na(classifs$nclasschunk),])
+}
+
+#check that the procedure worked
+clean$filename=as.factor(as.character(clean$filename))
+table(clean$filename)->sumvotes
+sum(sumvotes>5)
+sum(sumvotes==5) #33729 chunks
+sum(sumvotes<5)
+
+rm(classifs) #clean up to make space
+
+merge(zoosj,clean,by="filename")->chunks
+
+rm(clean) #clean up to make space
+
+write.csv(chunks,"../data_analyses/output/clean_classifications.csv")
+
+## PHASE 2 -- Generate views on the data: majority judgment on the chunk
+
+#file that contains subject info & annotation info from PU, 
+#filtered down to have 5 classifications per subject ID (= chunk)
[…]

### 2021-02-14 data_analyses/code/generate_jslhr_data.R

@@ -1,0 +1,64 @@
+## PHASE 1 -- RESTRICT ZOONIVERSE DATA TO PU DATA
+
+library(rjson)
+getname2=function(x) fromJSON(x)[[1]]
+
+#original chunk metadata
+origmetadata <- read.csv("../files_from_elsewhere/metadata_all_PU.csv",colClasses="factor",header=T,sep =",")
+dim(origmetadata)
+origmetadata$filename=paste0(origmetadata$AudioData,".mp3")
+length(levels(factor(origmetadata$filename)))
+#33730
+
+# #file that has "subject" information -- in Zooniverse terms, that is info about the file that was classified
+zoosj <- read.csv("../files_from_zooniverse/maturity-of-baby-sounds-subjects.csv",header=T,sep =",")
+zoosj$filename=unlist(lapply(as.character(zoosj[,"metadata"]), getname2))
+length(levels(factor(zoosj$filename)))
+#154063
+sum(levels(factor(zoosj$filename)) %in% levels(factor(origmetadata$filename)))
+#33730
+
+merge(origmetadata,zoosj,by="filename")->zoosj_PU
+dim(zoosj_PU)
+#33881 because some of the names are repeated
+mytab=table(zoosj_PU$filename)
+length(names(mytab[mytab==2]))
+#exploration revealed that one batch of 151 was uploaded twice -- removing the second upload
+for(mych in names(mytab[mytab==2])){
+  second_upload_id=max(zoosj_PU[zoosj_PU$filename==mych,"subject_id"])
+  zoosj_PU=zoosj_PU[zoosj_PU$subject_id!=second_upload_id,]
+}
+dim(zoosj_PU)
+#33730
+
+write.csv(zoosj_PU,"../output/zoo_subj_info_pu.csv")
+
+read.csv("../output/zoo_subj_info_pu.csv",colClasses="factor",header=T,sep =",")->zoosj
+
+#file that has the classifications (ie cit sci answers)
+classifs <- read.csv("../files_from_zooniverse/maturity-of-baby-sounds-classifications.csv",header=T,sep =",")
+classifs$subject_id=fromJSON(as.character(classifs[,"subject_data"]))[[1]]$retired$subject_id
+length(levels(factor(classifs$subject_id)))
+#151991
+classifs -> full_classifs
+classifs=subset(classifs, classifs$subject_id %in% levels(factor(zoosj$subject_id)))
+length(levels(factor(classifs$subject_id)))
[…]

### 2021-02-16 202010_jslhr/paper.Rmd
for proofing

@@
 # keywords          : "reliability; validation; daylong audiorecordings; citizen science; Zooniverse; LENA; long-form recordings"
-# wordcount         : "7,113"
+# wordcount         : "9375"
 
@@
 footnotelist      : no
-lineno            : yes
+lineno            : no
 
@@
 
+#document session info
+capture.output(sessionInfo(),file="lastknit_session_info.txt")
+
 printr=function(x) sub("0.", ".", round(x,3))
 
-#document session info
-
-sink("lastknit_session_info.txt")
-sessionInfo()
-sink()
 ```
@@
 read.csv("../data_analyses/output/zoo_lab_maj_judgments.csv")->data_all
+
+label_options=c("Canonical"   ,     "Crying"  ,        "Junk"  ,    "Laughing", "Non-Canonical")
 
@@
 data_all$lab[data_all$lab %in% c("Don't mark","None")]<-"Junk"
-data_all$lab=factor(data_all$lab,levels=levels(data_all$Zoon_classif))
+data_all$lab=factor(data_all$lab,levels=label_options)
 #apply same factor levels as zooniverse so that we can do symmetrical confusion matrices
@@
 ### Data post-processing
-<!-- ac this section still has numbers typed in  -->
-
-An impressive total of `r x["nusers","x"]` individual Zooniverse users provided `r x["njudgments","x"]` judgments on `r x["nchunks","x"]` 500-ms chunks, corresponding to `r dim(data_all)[1]` 11,980 LENA^TM^ segments.
-For a variety of technical mishaps, XX chunks had to be excluded, leading to `r length(levels(factor(chunks$subject_id)))` chunks and `r length(levels(factor(data_all$segmentId_DB)))` LENA^TM^ segments for which we had five Zooniverse labels.
-Among these, `r round(sum(chunks$nvotes<2)/dim(chunks)[1]*100,1)`% had no matching labels, `r round(sum(chunks$nvotes==2)/dim(chunks)[1]*100,1)`% had two matching labels (the threshold used for lab-annotated segments), and `r round(sum(chunks$nvotes>=3)/dim(chunks)[1]*100,1)`% had three or more matching labels, or majority agreement.
-Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the `r round(sum(chunks$nvotes>=3)/dim(chunks)[1]*100,1)`%  of chunks that *did* have at least three labels in agreement (i.e., majority agreement).
-This represented `r sum(chunks$nvotes>=3)*5` 135,725 labels for `r sum(chunks$nvotes>=3)` 27,145 chunks, corresponding to `r dim(data_all)[1]` LENA^TM^ segments.
+
+The following description uses several key concepts: A "judgment" or "label" is a single classification provided by a Zooniverse user in response to a single chunk; a "chunk" is a 500-ms clip extracted from the daylong audio to represent part or all of a segment; a segment is a section of the daylong audio automatically classified as having been uttered by the key child according to the LENA^TM^ algorithm.
+Bearing this terminology in mind, we provide an overview of the data and how it was processed before the analyses below.
+
+An impressive total of `r prettyNum(x["nusers","x"],big.mark=",")` individual Zooniverse users provided `r prettyNum(x["njudgments","x"],big.mark=",")` judgments on `r prettyNum(x["nchunks","x"],big.mark=",")` 500-ms chunks, corresponding to `r prettyNum(dim(data_all)[1],big.mark=",")` LENA^TM^ segments.
+For a variety of technical mishaps, `r prettyNum((x["njudgments","x"] - dim(chunks)[1]*5),big.mark=",")` judgments had to be excluded, leaving `r prettyNum(length(levels(factor(chunks$filename))),big.mark=",")` chunks (from `r prettyNum(length(levels(factor(data_all$segmentId_DB))),big.mark=",")` LENA^TM^ segments) for which we had five Zooniverse labels (for a total of `r prettyNum(dim(chunks)[1]*5,big.mark=",")` judgments included).
+Among these `r prettyNum(length(levels(factor(chunks$filename))),big.mark=",")` chunks, `r round(sum(chunks$nvotes<2)/dim(chunks)[1]*100,2)`% had no matching labels, `r round(sum(chunks$nvotes==2)/dim(chunks)[1]*100,2)`% had two matching labels (the threshold used for lab-annotated segments), and `r round(sum(chunks$nvotes>=3)/dim(chunks)[1]*100,2)`% had three or more matching labels, or majority agreement within Zooniverse labels.
+Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the `r round(sum(chunks$nvotes>=3)/dim(chunks)[1]*100,2)`%  of chunks that *did* have at least three labels in agreement (i.e., majority agreement).
+This represented `r prettyNum(sum(chunks$nvotes>=3)*5,big.mark=",")` labels for `r prettyNum(sum(chunks$nvotes>=3),big.mark=",")` chunks, corresponding to `r prettyNum(dim(data_all)[1],big.mark=",")` LENA^TM^ segments.
 Given that each chunk is 500 ms in length, this means about `r (sum(chunks$nvotes>=3)*.5)/60/60` hours of audio data were annotated by 8 different annotators (3 in the laboratory, 5 on Zooniverse).
@@
 
-Lab annotators provided judgments at the level of LENA^TM^ segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments.
+There was one more layer of pre-processing that was necessary for our analyses: Lab annotators provided judgments at the level of LENA^TM^ segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments.
 Therefore, we needed to combine chunk-level judgments to reconstruct segment-level judgments.
-
 We did this by considering the majority label extracted from Zooniverse judgments for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were Junk, then the segment was labelled as Junk; then, following a hierarchy (Canonical>Non-canonical>Crying>Laughing), segments with at least one instance of each judgment were labelled as such:
@@
 
+# remove non-majority labels from zooniverse
+data_all[data_all$Zoon_classif %in% label_options,]->data_all
+
 #and reset the factors for cleanliness
-data_all$Zoon_classif=factor(data_all$Zoon_classif)
-data_all$lab=factor(data_all$lab)
+data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=label_options)
+data_all$lab=factor(data_all$lab, levels=label_options)
 sample_data<-cbind(data_all$lab,data_all$Zoon_classif)
@@
 We also observed quite a bit of variance across the 50 runs.
-For Linguistic Proportion, the range of correlations found for all infants was [`r printp(min(resampling$lp_r_all))`,`r printp(max(resampling$lp_r_all))`]; for the Angelman syndrome group [`r printp(min(resampling$lp_r_as))`,`r printp(max(resampling$lp_r_as))`]; for the low-risk group [`r printp(min(resampling$lp_r_lr))`,`r printp(max(resampling$lp_r_lr))`].
-For Canonical Proportion, the range of correlations found for all infants was [`r printp(min(resampling$cp_r_all))`,`r printp(max(resampling$cp_r_all))`]; for the Angelman syndrome group [`r printp(min(resampling$cp_r_as))`,`r printp(max(resampling$cp_r_as))`]; for the low-risk group [`r printp(min(resampling$cp_r_lr))`,`r printp(max(resampling$cp_r_lr))`].
+For Linguistic Proportion, the range of correlations found for all infants was [`r printr(min(resampling$lp_r_all))`,`r printr(max(resampling$lp_r_all))`]; for the Angelman syndrome group [`r printr(min(resampling$lp_r_as))`,`r printr(max(resampling$lp_r_as))`]; for the low-risk group [`r printr(min(resampling$lp_r_lr))`,`r printr(max(resampling$lp_r_lr))`].
+For Canonical Proportion, the range of correlations found for all infants was [`r printr(min(resampling$cp_r_all))`,`r printr(max(resampling$cp_r_all))`]; for the Angelman syndrome group [`r printr(min(resampling$cp_r_as))`,`r printr(max(resampling$cp_r_as))`]; for the low-risk group [`r printr(min(resampling$cp_r_lr))`,`r printr(max(resampling$cp_r_lr))`].
 

### 2021-02-16 202010_jslhr/paper.tex
for proofing

@@
 \makeatother
-\usepackage{lineno}
-
-\linenumbers
 \usepackage{csquotes}
@@
 
-A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis, and it is beyond the scope of this paper to provide a full summary of this work {[}but see \textbf{XX ADD REFERENCE} for a recent summary{]}.
+A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis, and it is beyond the scope of this paper to provide a full summary of this work.
 Moreover, the way in which vocal development has been studied in the past may not easily translate to long-form recordings.
@@
 
-An impressive total of 4,825 individual Zooniverse users provided labels for the Maturity of Baby Sounds project, of which the present data set is one part.
-For this project, we collected a total of 169,767 judgments provided for 33,731 500-ms chunks, corresponding to 11,980 LENA\textsuperscript{TM} segments.
-Nearly a fifth of chunks did not have at least 3 labels in agreement out of the 5 Zooniverse labels (N = 6,585, 19\% of all chunks).
-Of the chunks without a majority agreement, 4341 (66\%) contained one or two Junk judgments (out of 5), 6523 (99,9\%) had at least two matching judgments (the threshold used for lab-annotated segments), and only 61 (0,01\%) had 5 different judgments.
-Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the 81\% of chunks that did have at least 3 labels in agreement; this represented 135,725 labels for 27,145 chunks, corresponding to 11,593 LENA\textsuperscript{TM} segments.
-As the segments average 1.12 seconds in length, this means about 3.8 hours of audio data were annotated by 8 different annotators (3 in the laboratory, 5 on Zooniverse).
+The following description uses several key concepts: A \enquote{judgment} or \enquote{label} is a single classification provided by a Zooniverse user in response to a single chunk; a \enquote{chunk} is a 500-ms clip extracted from the daylong audio to represent part or all of a segment; a segment is a section of the daylong audio automatically classified as having been uttered by the key child according to the LENA\textsuperscript{TM} algorithm.
+Bearing this terminology in mind, we provide an overview of the data and how it was processed before the analyses below.
 
-Lab annotators provided judgments at the level of LENA\textsuperscript{TM} segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments.
+An impressive total of 1,532 individual Zooniverse users provided 169,803 judgments on 33,730 500-ms chunks, corresponding to 11,980 LENA\textsuperscript{TM} segments.
+For a variety of technical mishaps, 1,158 judgments had to be excluded, leaving 33,729 chunks (from 11,980 LENA\textsuperscript{TM} segments) for which we had five Zooniverse labels (for a total of 168,645 judgments included).
+Among these 33,729 chunks, 0.18\% had no matching labels, 19.52\% had two matching labels (the threshold used for lab-annotated segments), and 80.30\% had three or more matching labels, or majority agreement within Zooniverse labels.
+Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the 80.30\% of chunks that \emph{did} have at least three labels in agreement (i.e., majority agreement).
+This represented 135,420 labels for 27,084 chunks, corresponding to 11,980 LENA\textsuperscript{TM} segments.
+Given that each chunk is 500 ms in length, this means about 3.76 hours of audio data were annotated by 8 different annotators (3 in the laboratory, 5 on Zooniverse).
+
+There was one more layer of pre-processing that was necessary for our analyses: Lab annotators provided judgments at the level of LENA\textsuperscript{TM} segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments.
 Therefore, we needed to combine chunk-level judgments to reconstruct segment-level judgments.
-
 We did this by considering the majority label extracted from Zooniverse judgments for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were Junk, then the segment was labelled as Junk; then, following a hierarchy (Canonical\textgreater{}Non-canonical\textgreater{}Crying\textgreater{}Laughing), segments with at least one instance of each judgment were labelled as such:
@@
 According to lab annotators, 14\% of segments were Canonical, 60\% Non-Canonical, 1\% Laughing, and 4\% Crying, with the remaining 21\% being categorized as \enquote{Don't mark}.
-Zooniverse data revealed a similar distribution: 15\% Canonical, 60\% Non-Canonical, 3\% Laughing, 7\% Crying, 15\% Junk.
+Zooniverse data revealed a similar distribution: 15\% Canonical, 60\% Non-Canonical, 3\% Laughing, 8\% Crying, 15\% Junk.
 Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical + Non-Canonical) / All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical / (Canonical + Non-Canonical) (i.e., we remove Junk, Crying, and Laughing).
@@
 Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups.
-There was a near-zero relationship among the children diagnosed with Angelman syndrome r(8) = -.091, CI {[}-.682,.571{]}, \emph{p} = .802{]}; and a significant association among low-risk control children r(8) = .7, CI {[}.126,.923{]}, \emph{p} = .024{]}.
-The Canonical Proportion exhibited non-significant developmental decreases among children diagnosed with Angelman syndrome r(8) = -.393, CI {[}-.82,.314{]}, \emph{p} = .261{]}; and marginal developmental increases among low-risk control r(8) = .516, CI {[}-.168,.865{]}, \emph{p} = .126{]}.
+There was a near-zero relationship among the children diagnosed with Angelman syndrome r(8) = -.092, CI {[}-.682,.571{]}, \emph{p} = .801{]}; and a significant association among low-risk control children r(8) = .7, CI {[}.126,.923{]}, \emph{p} = .024{]}.
+The Canonical Proportion exhibited non-significant developmental decreases among children diagnosed with Angelman syndrome r(8) = -.393, CI {[}-.82,.315{]}, \emph{p} = .261{]}; and marginal developmental increases among low-risk control r(8) = .516, CI {[}-.169,.865{]}, \emph{p} = .127{]}.
 
-Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(8) = -.174, CI {[}-.724,.512{]}, \emph{p} = .630{]}; whereas low-risk control children showed a significant increase with age r(8) = .629, CI {[}-.001,.902{]}, \emph{p} = .051{]}.
-Similarly, there were non-significant developmental decreases in the Canonical Proportion among children with Angelman syndrome r(8) = -.407, CI {[}-.825,.299{]}, \emph{p} = .243{]}; and marginal developmental increases among low-risk control children, r(8) = .514, CI {[}-.171,.864{]}, \emph{p} = .129{]}.
+Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(8) = -.179, CI {[}-.727,.508{]}, \emph{p} = .621{]}; whereas low-risk control children showed a significant increase with age r(8) = .632, CI {[}.004,.903{]}, \emph{p} = .050{]}.
+Similarly, there were non-significant developmental decreases in the Canonical Proportion among children with Angelman syndrome r(8) = -.404, CI {[}-.824,.303{]}, \emph{p} = .248{]}; and marginal developmental increases among low-risk control children, r(8) = .518, CI {[}-.165,.866{]}, \emph{p} = .125{]}.
 
@@
 These visualizations suggest that performance is moderate to good, which was confirmed via statistical analyses.
-The overall (weighted) accuracy is 73\%, CI = {[}72,74{]}, kappa is .531, and the Gwet's AC1 coefficient is .682, CI = {[}.669,.694{]}.
+The overall (weighted) accuracy is 73\%, CI = {[}72,74{]}, kappa is .534, and the Gwet's AC1 coefficient is .683, CI = {[}.671,.695{]}.
 
@@
 We illustrate this association in Figure 5.
-Looking at all 20 children together, we found a strong positive correlation r(18) = .794, CI {[}.543,.915{]}, \emph{p} \textless{} .001{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman syndrome r(8) = .878, CI {[}.555,.971{]}, \emph{p} = .001{]}; low-risk control r(8) = .828, CI {[}.415,.958{]}, \emph{p} = .003{]}.
+Looking at all 20 children together, we found a strong positive correlation r(18) = .8, CI {[}.554,.918{]}, \emph{p} \textless{} .001{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remain high: for Angelman syndrome r(8) = .873, CI {[}.539,.97{]}, \emph{p} = .001{]}; low-risk control r(8) = .831, CI {[}.421,.959{]}, \emph{p} = .003{]}.
 
@@
 
-Similarly, a strong positive correlation is found in the Canonical Proportion r(18) = .92, CI {[}.806,.968{]}, \emph{p} \textless{} .001{]}.
-When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman syndrome: r(8) = .681, CI {[}.091,.917{]}, \emph{p} = 0.03{]}; than the low-risk control children r(8) = .984, CI {[}.932,.996{]}, \emph{p} \textless{} .001.
+Similarly, a strong positive correlation is found in the Canonical Proportion r(18) = .921, CI {[}.807,.969{]}, \emph{p} \textless{} .001{]}.
+When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman syndrome: r(8) = .681, CI {[}.089,.917{]}, \emph{p} = 0.03{]}; than the low-risk control children r(8) = .985, CI {[}.937,.997{]}, \emph{p} \textless{} .001.
 
@@
 We investigated this hypothesis by calculating the proportion of their data labeled as \enquote{Junk} for each individual child.
-There was no significant difference in the proportions of their data labeled as \enquote{Junk} for children with Angelman syndrome (M = 0.14) compared to the low-risk group (M = 0.16): Welch's t(13.35) = -0.58, \emph{p} = .569.
+There was no significant difference in the proportions of their data labeled as \enquote{Junk} for children with Angelman syndrome (M = 0.14) compared to the low-risk group (M = 0.17): Welch's t(13.33) = -0.56, \emph{p} = .588.
 We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as \enquote{Junk}.
 Results were similar across these two post-hoc subgroups (see Figure 6).
-For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = .863, CI {[}.511,.967{]}, \emph{p} = .001{]}; and for the higher Junk group it was r(8) = .895, CI {[}.608,.975{]}, \emph{p} \textless{} .001.
-For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = .957, CI {[}.824,.99{]}, \emph{p} \textless{} .001{]}; and for the higher Junk group it was r(8) = .895, CI {[}.608,.975{]}, \emph{p} \textless{} .001.
+For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = .866, CI {[}.521,.968{]}, \emph{p} = .001{]}; and for the higher Junk group it was r(8) = .893, CI {[}.603,.975{]}, \emph{p} \textless{} .001.
+For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = .958, CI {[}.827,.99{]}, \emph{p} \textless{} .001{]}; and for the higher Junk group it was r(8) = .896, CI {[}.61,.975{]}, \emph{p} \textless{} .001.
 Thus, it does not seem that a higher proportion of \enquote{Junk} judgments is an index of low quality data.
@@
 When we split by participant group, correlations remain high: for Angelman syndrome r(8) = .845, CI {[}.46,.963{]}, \emph{p} = .002{]}; low-risk control r(8) = .794, CI {[}.328,.949{]}, \emph{p} = .006{]}.
-As for Canonical Proportion overall r(18) = .949, CI {[}.873,.98{]}, \emph{p} \textless{} .001; Angelman syndrome: r(8) = .724, CI {[}.173,.93{]}, \emph{p} = .018{]}; low-risk control infants r(8) = .972, CI {[}.881,.993{]}, \emph{p} \textless{} .001.
+As for Canonical Proportion overall r(18) = .949, CI {[}.873,.98{]}, \emph{p} \textless{} .001; Angelman syndrome: r(8) = .724, CI {[}.173,.93{]}, \emph{p} = .018{]}; low-risk control infants r(8) = .971, CI {[}.88,.993{]}, \emph{p} \textless{} .001.
 
@@
 Next, we looked at whether having more segments from each child may lead to more reliable metrics.
-We observed a non-significant trend {[}Welch t(17.88)=-0.53{]} for lower number of segments in the Angelman syndrome group (M = 348.70, range 163-599) than among low-risk control infants (M = 375.50, range 224-627), likely because of the lower volubility of the former children.
+We observed a non-significant trend {[}Welch t(17.88)=-0.53{]} for lower number of segments in the Angelman syndrome group (M = 348.40, range 163-599) than among low-risk control infants (M = 375.30, range 224-627), likely because of the lower volubility of the former children.
 Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications (Cychosz et al., 2019).
@@
 We also observed quite a bit of variance across the 50 runs.
-For Linguistic Proportion, the range of correlations found for all infants was {[}.442,.893{]}; for the Angelman syndrome group {[}.490,.969{]}; for the low-risk group {[}.289,.989{]}.
-For Canonical Proportion, the range of correlations found for all infants was {[}.750,.943{]}; for the Angelman syndrome group {[}.790,.994{]}; for the low-risk group {[}.547,.939{]}.
+For Linguistic Proportion, the range of correlations found for all infants was {[}.614,.877{]}; for the Angelman syndrome group {[}.426,.951{]}; for the low-risk group {[}.637,.937{]}.
+For Canonical Proportion, the range of correlations found for all infants was {[}.776,.955{]}; for the Angelman syndrome group {[}-.063,.915{]}; for the low-risk group {[}.868,.991{]}.
 
@@
 \hline
-0.878 & 0.681\\
+0.873 & 0.681\\
 \hline
-0.828 & 0.984\\
+0.831 & 0.985\\
 \hline
-0.794 & 0.920\\
+0.800 & 0.921\\
 \hline

### 2021-02-16 202010_jslhr/preprocess.R

@@ -1,6 +1,10 @@
 ## PHASE 1 -- clean classifications
+zoosj <- read.csv("../data_analyses/output/zoo_subj_info_pu.csv",header=T,sep =",")
 
 read.csv("../data_analyses/output/zoo_anno_info_pu.csv",colClasses="factor",header=T,sep =",")->classifs
+
+dim(classifs)
+#[1] 168852     15
 
 library(rjson)
 getval=function(x) fromJSON(x)[[1]]$value
@@ -19,9 +23,9 @@ length(levels(factor(classifs$filename)))
 # and remove clips with fewer than 5
 table(classifs$filename)->sumvotes
 toomany=rownames(sumvotes)[sumvotes>5]
-length(toomany) #271 have too many votes
+length(toomany) #271 have too many votes -- now 194
 justright=rownames(sumvotes)[sumvotes==5]
-length(justright) #33458 have the right amount
+length(justright) #33458 have the right amount -- now 33535
 toofew=rownames(sumvotes)[sumvotes<5]
 length(toofew)#1 chunk didn't receive 5 votes
 
@@ -42,6 +46,8 @@ sum(sumvotes<5)
 rm(classifs) #clean up to make space
 
 merge(zoosj,clean,by="filename")->chunks
+dim(chunks)
+#[1] 168645     36
 
 rm(clean) #clean up to make space
 
@@ -116,7 +122,7 @@ sum(!(dict.simple %in% maj_jud$AudioData)) # 1 not found
 
 rownames(dict)<-dict$chunk
 maj_jud$segmentId_DB <- dict[maj_jud$AudioData,"segmentId_DB"]
-length(levels(factor(maj_jud$segmentId_DB))) #12170
+length(levels(factor(maj_jud$segmentId_DB))) #12170 segments
 
 #generate majority at the segment level using our rule:
 # canonical > non-canonical > crying > laughing > junk

### 2021-02-17 202010_jslhr/paper.Rmd
proofing in progress

@@
 S.
-McDonnell Foundation Understanding Human Cognition Scholar Award.
+McDonnell Foundation (Understanding Human Cognition Scholar Award).
 LH acknowledges the National Institute of Deafness and Other Communication Disorders (F31DC018219).
@@
 The funders had no impact on this study.
-All authors approved the final manuscript as submitted and agree to be accountable for all aspects of the work and have no conflict of interests to disclose.
-
-# Abbreviations LENA^TM^: Language Environment Analysis.
+All authors approved the final manuscript as submitted, agree to be accountable for all aspects of the work and have no conflict of interests to disclose.
+
+# Abbreviations LENA&reg;: Language Environment Analysis.
 # ASHA recommends that abstracts be 150–250 words. The size limit for what can be included in your submission is set above 300 words,
@@
 abstract: >
-  **Purpose:** Recording young children's vocalizations through wearables is a promising method.
+  **Purpose:** Recording young children's vocalizations through wearables is a promising method to assess language development.
 However, accurately and rapidly annotating these files remains challenging.
@@
 
-  **Method:** Segments identified by LENA^TM^ as produced by the key child were extracted from one daylong recording for each of 20 participants: 10  low-risk control children and 10 children diagnosed  with  Angelman  syndrome, a  neurogenetic  syndrome  characterized  by  severe  language  impairments.
+  **Method:** Segments identified by LENA&reg; as produced by the key child were extracted from one daylong recording for each of 20 participants: 10  low-risk control children and 10 children diagnosed  with  Angelman  syndrome, a  neurogenetic  syndrome  characterized  by  severe  language  impairments.
 Speech samples were annotated by trained annotators in the laboratory as well as by citizen scientists on Zooniverse.
@@
   **Conclusion:** Annotations obtained through a citizen science platform can help us overcome challenges posed by the process of annotating daylong speech recordings.
-Particularly when used in composites or derived metrics, such annotations can be used to investigate early markers of language delays in non-typically developing children.
+Particularly when used in composites or derived metrics, such annotations can be used to investigate early markers of language delays.
 
@@
 Some research suggests that spontaneous behavior, captured for instance via home videos, could provide important indices to development [@belardi2017retrospective;@overby2020retrospective].
-Advancements in the field of wearable technologies, such as LENA^TM^ recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally [@rankine2017language;@vandam2019use;@oller2010automated].
+Advancements in the field of wearable technologies, such as LENA&reg; recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally [@rankine2017language;@vandam2019use;@oller2010automated].
 Wearable recorders allow data collection to happen in the child’s natural environment, and at a large scale, which may be particularly helpful for children whose speech is not easily elicited.
@@
 
-A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis, and it is beyond the scope of this paper to provide a full summary of this work.
-Moreover, the way in which vocal development has been studied in the past may not easily translate to long-form recordings.
-As mentioned earlier, long-form recordings have several advantages, including capturing the child's vocal patterns in their natural environment and being able to accumulate a great amount of data easily, which is particularly useful for diagnoses and pathologies characterized by low levels of vocal production.
+A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis.
+It is beyond the scope of this paper to provide a full summary of this extensive body work, mainly because the way in which vocal development has been studied in the past may not easily translate to long-form recordings.
+Long-form recordings have several advantages, including capturing the child's vocal patterns in their natural environment and being able to accumulate a great amount of data easily, which is particularly useful for diagnoses and pathologies characterized by low levels of vocal production.
 However, this also means that the audiorecording is harder to process than an audiorecording gathered in more manicured and stable conditions: In a typical daylong recording, the child will sometimes be in a quiet room with just her primary caregiver, but later in a noisy supermarket or having dinner with family and friends.
@@
 
-The most commonly used software was created by the LENA^TM^ Foundation, and it returns a diarization of the audio signal split into key talkers, including the child wearing the device, as well as a split of these audio attributed to the child into three categories: crying, other fixed and vegetative signals, and speech vocalizations.
+The most commonly used software was created by the LENA&reg; Foundation, and it returns a diarization of the audio signal split into key talkers, including the child wearing the device, as well as a split of sections attributed to the child into three categories: crying, other fixed and vegetative signals, and speech vocalizations.
 These subcategories have not been widely validated [@cristia2020accuracy].
-Moreover, LENA^TM^ lacks an important distinction between more and less advanced vocalizations.
+Moreover, until recently LENA&reg; lacked an important distinction between more and less advanced vocalizations.[^1]
+
+[^1]: The latest version of the LENA&reg; software does tag canonical utterances separately from non-canonical ones.
 
 Other automatized algorithms have been developed in the last two years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical [@schuller2019interspeech].
-The split of speech vocalizations into canonical and non-canonical, and the overt inclusion of laughing, allow a somewhat finer-grained picture of children's vocal activity than what the LENA^TM^ algorithm currently offers.
-Another advantage of this new line of research is that there exists a challenge where performance has been benchmarked, so unlike LENA^TM^'s algorithms, we know which algorithms perform better than others.
-Indeed, the ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55%.
+The overt inclusion of laughing allows a somewhat finer-grained picture of children's vocal activity than what the LENA&reg; algorithm currently offers.
+Another advantage of this new line of research is that there exists a challenge where performance has been benchmarked, so unlike LENA&reg;'s algorithms, we know which algorithms perform better than others.
+Indeed, the ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with an unweighted average recall of around 55% on the test set.
 By re-using the same test set, researchers can develop new algorithms and prove that they can outperform this baseline (or any subsequent best-performing algorithm benchmarked in the same test set).
@@
 Table 1 provides an idea of the accuracy across different methods.
-Specifically, we calculated accuracy of the LENA^TM^'s subtypes[^1] based on the laboratory annotations used in the present study; and we also show the results of ComParE 2019 BabySounds sub-challenge, although note it is not based on the same data as the other two columns.
-Table 1 shows that LENA^TM^ has a good recall for speech-like vocalizations and crying but not laughing, and by design it does not distinguish between canonical and non-canonical vocalizations.
-The ComParE 2019 BabySounds sub-challenge's baseline model achieved a more moderate recall on crying than LENA, but clearly outperforms the LENA^TM^ algorithm in the Laughing category, with the additional advantage of distinguishing canonical and non-canonical -- although note that performance in non-canonical is barely above a chance level of 20% (based on 5 categories, assuming equal probabilities).
-Notice that the baseline model also included a category called "Junk" for sections that turned out not to be the child's vocalizations after all, but LENA^TM^ here has a low recall because only sections that LENA^TM^ classified as being child vocalizations are considered in the present study, and thus this class's category cannot but be low-performing for LENA^TM^.
-That is, since we selected only segments attributed to the key child, any segment that turns out not to be a child vocalization is a LENA^TM^ error.
-We return to the comparison between LENA^TM^, the ComParE 2019 BabySounds sub-challenge baseline, and our Zooniverse-based method in the Discussion, but for now, we simply point out that there are new algorithms being developed, but their performance --although competitive when compared to LENA^TM^-- is underwhelming.
+Specifically, we calculated accuracy of the LENA&reg;'s subtypes[^2] based on the laboratory annotations used in the present study; and we also show the results of ComParE 2019 BabySounds sub-challenge, although note it is not based on the same data as the other two columns.
+Table 1 shows that LENA&reg; has a good recall for speech-like vocalizations and crying but not laughing, and at the time it did not distinguish between canonical and non-canonical vocalizations.
+The ComParE 2019 BabySounds sub-challenge's baseline model achieved a more moderate recall on crying than LENA, but clearly outperforms the LENA&reg; algorithm in the Laughing category, with the additional advantage of distinguishing canonical and non-canonical -- although note that performance in non-canonical is barely above a chance level of 20% (based on 5 categories, assuming equal probabilities).
+Notice that the baseline model also included a category called "Junk" for sections that turned out not to be the child's vocalizations after all.
+LENA&reg; has a low recall for the Junk category because only sections that LENA&reg; classified as being child vocalizations are considered in the present analysis.
+As a result, any segment that turns out not to be a child vocalization is a LENA&reg; error.
+We return to the comparison between LENA&reg;, the ComParE 2019 BabySounds sub-challenge baseline, and our Zooniverse-based method in the Discussion; for now, we simply point out that there are new algorithms being developed, but their performance --although competitive when compared to LENA&reg;-- is underwhelming.
 
@@
 
-[^1]: If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as "Junk" or not categorized.
+[^2]: If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as "Junk" or not categorized.
+When these data were analyzed, the LENA&reg; software did not distinguish canonical and non-canonical.
 
@@
 
-One possibility that has only recently begun to be explored is the use of derived metrics.
-So instead of evaluating algorithms and other annotation procedures on their accuracy at individual labels, one can derive a metric that more closely relates to vocalization development.
-
-For example, most work using LENA^TM^ software reports on children's vocalization counts -- this is a derived metric because it is not simply a description of sections of the audio, but instead it integrates over a certain period of time.
-Although this has been sometimes criticized as being more about quantity than quality [@mcdaniel2020predicting], this is a promising metric of individuals' vocalization development because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA^TM^ [@cristia2020accuracy].
+One possibility that has only recently begun to be explored is the use of *derived metrics*.
+So instead of evaluating algorithms and other annotation procedures on  accuracy of individual sections of the audio, one can integrate over time (e.g., over the whole day-long audio) to derive a metric that more closely relates to vocalization development.
+
+For example, most work using LENA&reg; software reports on children's vocalization counts -- a derived metric because it is not simply a description of sections of the audio, but instead it integrates over the whole recording length.
+Although linguistic vocalization count has been criticized as being more about quantity than quality [@mcdaniel2020predicting], it is a promising metric of individuals' vocalization development because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA&reg; [@cristia2020accuracy].
 Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r ~ .3 with standardized language scores in a meta-analysis [@wang2020meta].
-Thus despite representing a composite of skills, vocalization count may provide a useful estimate of vocalization development.
+Thus despite representing a composite of skills, vocalization count may provide a useful estimate of vocalization development.[^3]
+
+[^3]: LENA&reg;'s child vocalization counts (CVC) aggregates all key child vocalizations detected as being *linguistic*.
+As a result, an increase in CVC could indicate an actual increase in how many linguistic vocalizations the child produces, or a change in how well they are detected as such by the algorithm.
+Either way, they seem to correlate with vocal development, at least based on their correlation with age (in e.g.) and language scores in [@wang2020meta].
 
@@
 @cychosz2019canonical drew child vocalization data from a diverse set of corpora centered on children learning one of four languages: English, Tseltal, Tsimane', and Yélî.
-For the English and Tsimane' corpora, vocalizations were automatically identified using LENA^TM^, whereas the other two were extracted through manual segmentation.
+For the English and Tsimane' corpora, vocalizations were automatically identified using LENA&reg;, whereas the other two were extracted through manual segmentation.
 Segments were then split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received three classifications into Crying, Laughing, Canonical, Non-Canonical, or Junk (with the latter tag used for clips that did not contain a child voice).
@@
 They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the "Junk" category was very high.
-Additionally, they found that the correlation between Canonical Proportion and age was much weaker in two corpora where vocalizations were segmented with LENA^TM^ than through manual segmentation.
+Additionally, they found that the correlation between Canonical Proportion and age was much weaker in two corpora where vocalizations were segmented with LENA&reg; than through manual segmentation.
 In sum, these analyses within corpora reveal that data may not be uniformly useful, with potential variation across corpora, children, and methods.
@@
 It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels.
-We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA^TM^ [see meta-analyses in @wang2020meta;@cristia2020accuracy]; as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range.
+We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA&reg; [see meta-analyses in @wang2020meta;@cristia2020accuracy]; as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range.
 Some previous work had shown that vocal development proceeded differently when compared to a normative sample [@hamrick2019capturing;@hamrick2019measurement].
@@
 
-Recordings were obtained with the Language ENvironment Analysis (LENA^TM^) Digital Language Processor,  a lightweight recorder (< 60 g, 5.5 cm x 8.5 cm x 1.5 cm) designed to be worn inside a breast-pocket of purpose-made clothing.
+Recordings were obtained with the Language ENvironment Analysis (LENA&reg;) Digital Language Processor,  a lightweight recorder (< 60 g, 5.5 cm x 8.5 cm x 1.5 cm) designed to be worn inside a breast-pocket of purpose-made clothing.
 In our data, every child contributed at least one daylong recording, with one child contributing two recordings (duration 11.58-16.00 hours, mean 15.05 hours).
 
-### Data preprocessing (LENA^TM^)
+### Data preprocessing (LENA&reg;)
 
 The first step we took before beginning the annotation process was the identification of children's vocalizations within day-long recordings.
-We did this automatically, using the LENA^TM^ proprietary software [@xu2008signal].
-Specifically, the LENA^TM^ software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child’s environment (e.g., Female-Adult, TV).
+We did this automatically, using the LENA&reg; proprietary software [@xu2008signal].
+Specifically, the LENA&reg; software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child’s environment (e.g., Female-Adult, TV).
 Key child labels have a precision of about 60% and a recall of about 50% [@cristia2020thorough].
@@
 Annotation was carried out at the segment level during lab coding.
-A subset of segments LENA^TM^ identified as the key child were selected for human annotation.
+A subset of segments LENA&reg; identified as the key child were selected for human annotation.
 Segments were selected from 30 five-minute sections of the child’s recording: 10 sections during which the child was vocalizing at their highest rate, and 20 sections drawn randomly from the remainder of the recording.
@@
 Specifically, the segments extracted were automatically cut into clips of exactly 500ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform.
-We will call this level *chunk*, to convey that they are part of a bigger unit (the LENA^TM^-identified "segment").
+We will call this level *chunk*, to convey that they are part of a bigger unit (the LENA&reg;-identified "segment").
 Note that segments and chunks are collectively referred to as clips because they are parts of the longer audio recording.
@@
 
-The following description uses several key concepts: A "judgment" or "label" is a single classification provided by a Zooniverse user in response to a single chunk; a "chunk" is a 500-ms clip extracted from the daylong audio to represent part or all of a segment; a segment is a section of the daylong audio automatically classified as having been uttered by the key child according to the LENA^TM^ algorithm.
+The following description uses several key concepts: A "judgment" or "label" is a single classification provided by a Zooniverse user in response to a single chunk; a "chunk" is a 500-ms clip extracted from the daylong audio to represent part or all of a segment; a segment is a section of the daylong audio automatically classified as having been uttered by the key child according to the LENA&reg; algorithm.
 Bearing this terminology in mind, we provide an overview of the data and how it was processed before the analyses below.
 
-An impressive total of `r prettyNum(x["nusers","x"],big.mark=",")` individual Zooniverse users provided `r prettyNum(x["njudgments","x"],big.mark=",")` judgments on `r prettyNum(x["nchunks","x"],big.mark=",")` 500-ms chunks, corresponding to `r prettyNum(dim(data_all)[1],big.mark=",")` LENA^TM^ segments.
-For a variety of technical mishaps, `r prettyNum((x["njudgments","x"] - dim(chunks)[1]*5),big.mark=",")` judgments had to be excluded, leaving `r prettyNum(length(levels(factor(chunks$filename))),big.mark=",")` chunks (from `r prettyNum(length(levels(factor(data_all$segmentId_DB))),big.mark=",")` LENA^TM^ segments) for which we had five Zooniverse labels (for a total of `r prettyNum(dim(chunks)[1]*5,big.mark=",")` judgments included).
+An impressive total of `r prettyNum(x["nusers","x"],big.mark=",")` individual Zooniverse users provided `r prettyNum(x["njudgments","x"],big.mark=",")` judgments on `r prettyNum(x["nchunks","x"],big.mark=",")` 500-ms chunks, corresponding to `r prettyNum(dim(data_all)[1],big.mark=",")` LENA&reg; segments.
+For a variety of technical mishaps, `r prettyNum((x["njudgments","x"] - dim(chunks)[1]*5),big.mark=",")` judgments had to be excluded, leaving `r prettyNum(length(levels(factor(chunks$filename))),big.mark=",")` chunks (from `r prettyNum(length(levels(factor(data_all$segmentId_DB))),big.mark=",")` LENA&reg; segments) for which we had five Zooniverse labels (for a total of `r prettyNum(dim(chunks)[1]*5,big.mark=",")` judgments included).
 Among these `r prettyNum(length(levels(factor(chunks$filename))),big.mark=",")` chunks, `r round(sum(chunks$nvotes<2)/dim(chunks)[1]*100,2)`% had no matching labels, `r round(sum(chunks$nvotes==2)/dim(chunks)[1]*100,2)`% had two matching labels (the threshold used for lab-annotated segments), and `r round(sum(chunks$nvotes>=3)/dim(chunks)[1]*100,2)`% had three or more matching labels, or majority agreement within Zooniverse labels.
 Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the `r round(sum(chunks$nvotes>=3)/dim(chunks)[1]*100,2)`%  of chunks that *did* have at least three labels in agreement (i.e., majority agreement).
-This represented `r prettyNum(sum(chunks$nvotes>=3)*5,big.mark=",")` labels for `r prettyNum(sum(chunks$nvotes>=3),big.mark=",")` chunks, corresponding to `r prettyNum(dim(data_all)[1],big.mark=",")` LENA^TM^ segments.
+This represented `r prettyNum(sum(chunks$nvotes>=3)*5,big.mark=",")` labels for `r prettyNum(sum(chunks$nvotes>=3),big.mark=",")` chunks, corresponding to `r prettyNum(dim(data_all)[1],big.mark=",")` LENA&reg; segments.
 Given that each chunk is 500 ms in length, this means about `r (sum(chunks$nvotes>=3)*.5)/60/60` hours of audio data were annotated by 8 different annotators (3 in the laboratory, 5 on Zooniverse).
@@
 
-There was one more layer of pre-processing that was necessary for our analyses: Lab annotators provided judgments at the level of LENA^TM^ segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments.
+There was one more layer of pre-processing that was necessary for our analyses: Lab annotators provided judgments at the level of LENA&reg; segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments.
 Therefore, we needed to combine chunk-level judgments to reconstruct segment-level judgments.
@@
 This is essentially the same thought process expert annotators followed when labeling segment-level audio, as each segment can represent multiple classification subtypes.
-LENA^TM^ segments are not equivalent to child vocalizations: A segment may contain 1 or more vocalizations, which may be of the same type or not.
+LENA&reg; segments are not equivalent to child vocalizations: A segment may contain 1 or more vocalizations, which may be of the same type or not.
 Figure 2 conveys the relationship between chunks and segments, and how judgments are combined into labels.
@@
 
-Above, we concluded that derived metrics integrating information across audio clips (Linguistic and Canonical Proportion, which can be derived from segment- or chunk-level judgments) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA^TM^ segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
+Above, we concluded that derived metrics integrating information across audio clips (Linguistic and Canonical Proportion, which can be derived from segment- or chunk-level judgments) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA&reg; segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
 As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
@@
 
-In the present study we assessed the extent to which child vocalizations from LENA^TM^ daylong recordings can be accurately described based on classifications collected using the citizen science platform Zooniverse.
+In the present study we assessed the extent to which child vocalizations from LENA&reg; daylong recordings can be accurately described based on classifications collected using the citizen science platform Zooniverse.
 Our recordings came from children diagnosed with a genetic syndrome associated with severe language disorder, as well as low-risk controls, and included a range of different ages.
@@
 The first area in which more work would be welcome involves the pre-processing.
-We used LENA^TM^ algorithms to detect vocalizations by the key child, but there are newer, better-performing, open-sourced alternatives that could be used instead [@lavechin2020open].
+We used LENA&reg; algorithms to detect vocalizations by the key child, but there are newer, better-performing, open-sourced alternatives that could be used instead [@lavechin2020open].
 Still, performance is most often established on children whose development is thought to be normative, being raised in small households located in urban environments, whereas it would be important to check for accuracy in more diverse samples, where the children have a diagnosis [e.g., @Dykstra2013], multiple siblings [e.g., @Elo], and/or are growing up in a rural environment [e.g., @cristia2020thorough].
@@
 However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
-Now that this method has been validated, researchers can extract all segments that LENA^TM^ identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
-This approach may be particularly useful in the case of data from children with a diagnosis, whose data tends to be under-represented in LENA^TM^ research, compared to data from normative children.
+Now that this method has been validated, researchers can extract all segments that LENA&reg; identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
+This approach may be particularly useful in the case of data from children with a diagnosis, whose data tends to be under-represented in LENA&reg; research, compared to data from normative children.
 We also hope future work makes use of the scalability of the method by employing larger sample sizes than those used in the present study, where data from only 10 children was used.

### 2021-02-17 202010_jslhr/paper.Rmd
cleaning

@@
 read.csv("../data_analyses/output/chunks_maj_judgments.csv")->chunks
+
+#fix typo
+chunks$segmentId_DB=gsub("(DB.)(.)$","\\1",chunks$segmentId_DB)
+
 read.csv("../data_analyses/output/zoo_lab_maj_judgments.csv")->data_all
@@
 Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r ~ .3 with standardized language scores in a meta-analysis [@wang2020meta].
-Thus despite representing a composite of skills, vocalization count may provide a useful estimate of vocalization development.[^3]
-
-[^3]: LENA&reg;'s child vocalization counts (CVC) aggregates all key child vocalizations detected as being *linguistic*.
+LENA&reg;'s child vocalization counts (CVC) aggregates all key child vocalizations detected as being *linguistic*.
 As a result, an increase in CVC could indicate an actual increase in how many linguistic vocalizations the child produces, or a change in how well they are detected as such by the algorithm.
-Either way, they seem to correlate with vocal development, at least based on their correlation with age (in e.g.) and language scores in [@wang2020meta].
+Thus, despite representing a composite of skills, vocalization count may provide a useful estimate of vocalization development.
 
@@
 One metric that can be derived once children's vocalizations are split into crying, laughing, canonical, and non-canonical is the Linguistic Proportion: the proportion of vocalizations that are linguistic (canonical and non-canonical) out of all vocalizations.
-To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age.
+To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age early on and eventually plateaus.
 
 Yet another metric that could be extracted to estimate relative linguistic complexity is the Canonical Proportion: the proportion of vocalizations that contain a canonical transition or syllable out of all linguistic (canonical and non-canonical) vocalizations.
-One recent study has established that the Canonical Proportion extracted from daylong recordings is significantly correlated with age in a multicultural and multilingual sample [@cychosz2019canonical].
-Importantly, this association was established in a sample of children going up to 3 years of age, and thus well beyond the babbling period, and into first words and word combinations.
-
-There is a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this work focuses on babbling and thus typically on infants under one year of age.
-A critical milestone involves the increasingly common production of canonical syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech [@oller1998late].
+@cychosz2019canonical found that the Canonical Proportion extracted from daylong recordings was significantly correlated with age in a multicultural and multilingual sample of children up to 3 years of age, and thus well beyond the babbling period, and into first words and word combinations.
+That study built on a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this prior work focuses on babbling and thus typically on infants under one year of age.
+In these younger infants, a critical milestone involves the increasingly common production of canonical syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech [@oller1998late].
 Given its adult-like consonant-vowel or vowel-consonant structure, canonical babble is considered to be a starting point on the path to recognizable speech.
-Canonical syllables show a higher complexity given the smooth articulatory transition between a consonant and a vowel (or vice versa), when compared to more primitive sounds such as squeals, or isolated vowels.
-Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development [@oller2000emergence].
+When compared to more primitive sounds such as squeals, or isolated vowels, canonical syllables show a higher complexity given the smooth articulatory transition between a consonant and a vowel (or vice versa).
+Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development [@oller2000emergence;@cychosz2019canonical].
 In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization counts in a sample of children diagnosed with an autism spectrum disorder [@mcdaniel2020predicting].
@@
 The first pertains to how the data are annotated, with human annotation being costly.
-The second relates to how descriptors of vocal development are generated, i.e., how annotations of audio sections are integrated into vocal development metrics.
-
-Crowdsourcing refers to the process whereby a task is solved by a crowd, rather than an individual.
+The second relates to how descriptors of vocal development are generated, i.e., how annotations of individual audio sections are integrated over all data for a given child to derive child-level vocal development metrics.
+
+"Crowdsourcing" refers to the process whereby a task is solved by a crowd, rather than an individual.
 A number of fields, particularly in the data-driven sciences, have already engaged in the collection of data (including annotations) through crowdsourcing, thanks to its low cost and ecological value [@crump2013evaluating;@sescleifer2018systematic].
-For example, a systematic review on crowd-sourced ratings of speech found that "lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology"; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81 [@sescleifer2018systematic].
+For example, a systematic review on crowdsourced ratings of speech found that "lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology"; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81 [@sescleifer2018systematic].
 On the other hand, the systematic review returned only 8 studies (of which only four were published in peer-reviewed journals), suggesting that there is considerable need for further research on this topic.
@@
 Mechanical Turk (MTurk) MTurk is an online labor market created by Amazon to assist "requesters" in hiring and paying "workers" for the completion of computerized tasks.
-Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable [@berinsky2012evaluating], some question marks are raised as some "workers" turn out to be bots, or are poorly motivated (and potentially exploited) humans.
+Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable [@berinsky2012evaluating],  question marks are raised as some "workers" turn out to be bots, or are poorly motivated (and potentially exploited) humans.
 
@@
 
-When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk judgments); and the Canonical Proportion (Canonical)/(all linguistic judgments).
+When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk labels); and the Canonical Proportion (Canonical)/(all linguistic labels).
 
@@
 These stretches of audio assigned to the key child can contain cries, vegetative sounds, and linguistic vocalizations, as well as interstitial pauses.
-We will refer to them as *segments*, to highlight the fact that they are stretches of audio segmented out as belonging to the key child, but they may not be individual vocalizations.
+We refer to them as *segments*, to highlight the fact that they are stretches of audio segmented out as belonging to the key child, but they may not be individual vocalizations.
 
@@
 After this, they completed a brief training on vocal maturity that involved 1) reading relevant literature, 2) reviewing examples of various annotation categories and 3) submitting answers to a brief quiz to test their annotation accuracy.
-Coders then proceeded to the classification, where they made decisions on whether a segment is a cry, laugh, non-canonical syllable(s), canonical syllable(s), or a word.
-Coders were instructed to classify each segment based on the highest level of vocal maturity it contains (e.g., a segment with both canonical and non-canonical syllables would be classified as canonical; a segment with both laughing and non-canonical syllables would be classified as non-canonical).
-For canonical, non-canonical, and word segments, coders indicated how many syllables of each type of vocal maturity they hear (e.g., "ah ma ba" would be 1 non-canonical and 2 canonical syllables).
+Coders then proceeded to the classification, where they made decisions on whether a segment was a cry, laugh, non-canonical syllable(s), canonical syllable(s), or a word.
+Coders were instructed to classify each segment based on the highest level of vocal maturity it contained (e.g., a segment with both canonical and non-canonical syllables would be classified as canonical; a segment with both laughing and non-canonical syllables would be classified as non-canonical).
+For canonical, non-canonical, and word segments, coders indicated how many syllables of each type of vocal maturity they heard (e.g., "ah ma ba" would be 1 non-canonical and 2 canonical syllables).
 For the purposes of the present study, we considered a "word" judgment to be equivalent to a "canonical" judgment.
@@
 
-Coders are also given a "Don’t Mark" option, which they are instructed to use if the segment does not sound like a segment made by the target child, or if there is any overlapping speech or other noise in the background which could affect acoustic analyses.
-Notice, therefore, that this category is broader than the "Junk" category used by Zooniverse annotators.
-
-This process resulted in high levels of inter-rater agreement: Two or more of the three total coders who annotated that segment agreed on 97.22% of the lab-annotated segments.
+Coders were also given a "Don’t Mark" option, which they were instructed to use if the segment did not sound like a segment made by the target child, or if there is any overlapping speech or other noise in the background which could affect acoustic analyses.
+This category is broader than the "Junk" category used by Zooniverse annotators, which is explained in the next section.
+
+This process resulted in high levels of inter-rater agreement: Two or more of the three total coders who annotated that segment agreed on `r round(sum(data_all$Num_Agreement>=2)/dim(data_all)[1]*100,1)`% of the lab-annotated segments.
 
@@
 
-A substantial concern that emerges when considering citizen science annotation for spontaneous speech data is the risk of a privacy breach: even short clips can contain personal information that can expose the identity of the speaker, such as addresses or names, or contain sensitive and private information.
-Fortunately, @Seidl19 determined that when segments are divided into shorter clips (400-600ms), human annotators with little training can code our categories of interest (Canonical, Laughing, Crying, etc.) with a classification quality comparable to the one carried out on full segments.
-We therefore used very short clips (500ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.
-
-Specifically, the segments extracted were automatically cut into clips of exactly 500ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform.
+A substantial concern that emerges when considering citizen science annotation for spontaneous speech data is the risk of a privacy breach: Even short clips can contain personal information that can expose the identity of the speaker, such as addresses or names, or contain sensitive and private information.
+Fortunately, @Seidl19 determined that when segments are divided into shorter clips (400-600 ms), human annotators with little training can code our categories of interest (Canonical, Laughing, Crying, etc.) with a classification quality comparable to the one carried out on full segments.
+We therefore used very short clips (500 ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.
+
+Specifically, the segments extracted were automatically cut into clips of exactly 500 ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform.
 We will call this level *chunk*, to convey that they are part of a bigger unit (the LENA&reg;-identified "segment").
-Note that segments and chunks are collectively referred to as clips because they are parts of the longer audio recording.
-To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms chunks and their corresponding segment.
+Note that segments and chunks are collectively referred to as *clips* because they are parts of the longer audio recording.
+To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500-ms chunks and their corresponding segment.
 Chunks were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling [@bowyer2015panoptes].
@@
 
-Users were asked to assign each 500ms chunk to one out of five possible categories: (1) Canonical, (2) Non-Canonical, (3) Crying, (4) Laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
-Notice that the latter is narrower than the "Don't mark" category of the lab annotation routine, which was used any time there was noise or overlap that could affect acoustic analyses.
+Users were asked to assign each 500-ms chunk to one out of five possible categories: (1) Canonical, (2) Non-Canonical, (3) Crying, (4) Laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
+Notice that the latter is narrower than the "Don't mark" category of the lab annotation routine, which was used any time there was any noise or overlap that could affect acoustic analyses.
 
@@
 
-The following description uses several key concepts: A "judgment" or "label" is a single classification provided by a Zooniverse user in response to a single chunk; a "chunk" is a 500-ms clip extracted from the daylong audio to represent part or all of a segment; a segment is a section of the daylong audio automatically classified as having been uttered by the key child according to the LENA&reg; algorithm.
-Bearing this terminology in mind, we provide an overview of the data and how it was processed before the analyses below.
+The following description uses several key concepts: A "judgment" is a single classification provided by a Zooniverse user in response to a single chunk (or a single classification provided by a lab annotator in response to a single segment); a "chunk" is a 500-ms clip extracted from the daylong audio to represent part or all of a segment; a segment is a section of the daylong audio automatically classified as having been uttered by the key child according to the LENA&reg; algorithm; and a "label" is the final category a chunk or segment was classified as.
+Bearing this terminology in mind, we provide an overview of the data and how it was processed before the main analyses.
 
 An impressive total of `r prettyNum(x["nusers","x"],big.mark=",")` individual Zooniverse users provided `r prettyNum(x["njudgments","x"],big.mark=",")` judgments on `r prettyNum(x["nchunks","x"],big.mark=",")` 500-ms chunks, corresponding to `r prettyNum(dim(data_all)[1],big.mark=",")` LENA&reg; segments.
-For a variety of technical mishaps, `r prettyNum((x["njudgments","x"] - dim(chunks)[1]*5),big.mark=",")` judgments had to be excluded, leaving `r prettyNum(length(levels(factor(chunks$filename))),big.mark=",")` chunks (from `r prettyNum(length(levels(factor(data_all$segmentId_DB))),big.mark=",")` LENA&reg; segments) for which we had five Zooniverse labels (for a total of `r prettyNum(dim(chunks)[1]*5,big.mark=",")` judgments included).
-Among these `r prettyNum(length(levels(factor(chunks$filename))),big.mark=",")` chunks, `r round(sum(chunks$nvotes<2)/dim(chunks)[1]*100,2)`% had no matching labels, `r round(sum(chunks$nvotes==2)/dim(chunks)[1]*100,2)`% had two matching labels (the threshold used for lab-annotated segments), and `r round(sum(chunks$nvotes>=3)/dim(chunks)[1]*100,2)`% had three or more matching labels, or majority agreement within Zooniverse labels.
-Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the `r round(sum(chunks$nvotes>=3)/dim(chunks)[1]*100,2)`%  of chunks that *did* have at least three labels in agreement (i.e., majority agreement).
-This represented `r prettyNum(sum(chunks$nvotes>=3)*5,big.mark=",")` labels for `r prettyNum(sum(chunks$nvotes>=3),big.mark=",")` chunks, corresponding to `r prettyNum(dim(data_all)[1],big.mark=",")` LENA&reg; segments.
-Given that each chunk is 500 ms in length, this means about `r (sum(chunks$nvotes>=3)*.5)/60/60` hours of audio data were annotated by 8 different annotators (3 in the laboratory, 5 on Zooniverse).
+Given that each chunk is 500 ms in length, this means about `r round((x["nchunks","x"]*.5)/60/60,2)` hours of audio data were annotated by 8 different annotators (3 in the laboratory, 5 on Zooniverse).
+For a variety of technical mishaps, `r prettyNum((x["njudgments","x"] - dim(chunks)[1]*5),big.mark=",")` judgments had to be excluded, leaving `r prettyNum(length(levels(factor(chunks$filename))),big.mark=",")` chunks for which we had five Zooniverse judgments (for a total of `r prettyNum(dim(chunks)[1]*5,big.mark=",")` judgments included).
+Among these `r prettyNum(length(levels(factor(chunks$filename))),big.mark=",")` chunks, `r round(sum(chunks$nvotes<2)/dim(chunks)[1]*100,2)`% had no matching judgments, `r round(sum(chunks$nvotes==2)/dim(chunks)[1]*100,2)`% had two matching judgments (the threshold used for lab-annotated segments), and `r round(sum(chunks$nvotes>=3)/dim(chunks)[1]*100,2)`% had three or more matching judgments, or majority agreement within Zooniverse judgments.
+Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the `r round(sum(chunks$nvotes>=3)/dim(chunks)[1]*100,2)`%  of chunks that *did* have at least three judgments in agreement (i.e., majority agreement).
+This represented `r prettyNum(sum(chunks$nvotes>=3)*5,big.mark=",")` judgments for `r prettyNum(sum(chunks$nvotes>=3),big.mark=",")` chunks.
 
@@
 
-There was one more layer of pre-processing that was necessary for our analyses: Lab annotators provided judgments at the level of LENA&reg; segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments.
-Therefore, we needed to combine chunk-level judgments to reconstruct segment-level judgments.
-We did this by considering the majority label extracted from Zooniverse judgments for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were Junk, then the segment was labelled as Junk; then, following a hierarchy (Canonical>Non-canonical>Crying>Laughing), segments with at least one instance of each judgment were labelled as such:
-
-- if one Canonical judgment was present, then the segment was labelled as Canonical, 
+There was one more layer of processing that was necessary for our main analyses: Lab annotators provided judgments at the level of LENA&reg; segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments.
+Therefore, we needed to combine chunk-level labels to reconstruct segment-level labels.
+We did this by considering the majority labels of all chunks associated with each segment as follows:
+
+- If the majority label for all chunks associated with a given segment were Junk, then the segment was labelled as Junk,  
+- else, if one Canonical judgment was present for any of the chunks associated to the segment, then the segment was labelled as Canonical, 
 - else, if a Non-Canonical judgment was present, the segment was labelled as Non-Canonical, 
@@
 
-This is essentially the same thought process expert annotators followed when labeling segment-level audio, as each segment can represent multiple classification subtypes.
-LENA&reg; segments are not equivalent to child vocalizations: A segment may contain 1 or more vocalizations, which may be of the same type or not.
+This is essentially the same thought process Lab annotators followed when jduging segments: a LENA&reg; segment may contain 1 or more vocalizations, which may be of the same type or not.
 Figure 2 conveys the relationship between chunks and segments, and how judgments are combined into labels.
@@
 
-In this section, we report on explorations of under what conditions Zooniverse judgments more closely aligned with laboratory judgments.
+In this section, we report on explorations of under what conditions Zooniverse labels more closely aligned with laboratory labels.
 In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as "Junk" (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora [@cychosz2019canonical].
@@
 For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["cp","lower_Junk"]`) = `r printr(myr["cp","lower_Junk"])`, CI [`r printr(mylb["cp","lower_Junk"])`,`r printr(myhb["cp","lower_Junk"])`], *p* `r printp(myp["cp","lower_Junk"])`]; and for the higher Junk group it was r(`r myn["cp","higher_Junk"]`) = `r printr(myr["cp","higher_Junk"])`, CI [`r printr(mylb["cp","higher_Junk"])`,`r printr(myhb["cp","higher_Junk"])`], *p* `r printp(myp["cp","higher_Junk"])`.
-Thus, it does not seem that a higher proportion of "Junk" judgments is an index of low quality data.
+Thus, it does not seem that a higher proportion of "Junk" labels is an index of low quality data.
 
@@
 
-Above, we concluded that derived metrics integrating information across audio clips (Linguistic and Canonical Proportion, which can be derived from segment- or chunk-level judgments) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA&reg; segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
-As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
+Above, we concluded that derived metrics integrating information across audio clips (Linguistic and Canonical Proportion, which can be derived from segment- or chunk-level labels) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA&reg; segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
+As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed labels across chunks associated to the same segment.
 We therefore repeated our analyses but instead of deriving our proportions for the Zooniverse data from the segment-level composite, we did it based on the individual chunk-level annotations.
@@
 
-Since all of these results are very similar to those obtained when collapsing chunk-level judgments to generate segment-level judgments, we conclude that in the future this step can be skipped.
-Instead, researchers can derive Linguistic and Canonical Proportions directly from citizen scientists' chunk-level judgments [which was in fact what @cychosz2019canonical did].
+Since all of these results are very similar to those obtained when collapsing chunk-level labels to generate segment-level labels, we conclude that in the future this step can be skipped.
+Instead, researchers can derive Linguistic and Canonical Proportions directly from citizen scientists' chunk-level labels [which was in fact what @cychosz2019canonical did].
 
@@
 We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children [following @cychosz2019canonical].
-We randomly sampled 100 segments from each child's data, and recalculated the Linguistic and Canonical Proportions based only on these associated judgments.
+We randomly sampled 100 segments from each child's data, and recalculated the Linguistic and Canonical Proportions based only on these associated labels.
 We repeated this process 50 times, to assess the extent to which the association between laboratory- and Zooniverse-derived metrics varied in each random sample (e.g., the correlation between Linguistic Proportion as estimated using Zooniverse coding on the one hand, and Linguistic Proportion as estimated using laboratory coding on the other).
@@
 For Canonical Proportion, the range of correlations found for all infants was [`r printr(min(resampling$cp_r_all))`,`r printr(max(resampling$cp_r_all))`]; for the Angelman syndrome group [`r printr(min(resampling$cp_r_as))`,`r printr(max(resampling$cp_r_as))`]; for the low-risk group [`r printr(min(resampling$cp_r_lr))`,`r printr(max(resampling$cp_r_lr))`].
-
-<!-- - look at whether kids who have a high "Junk" proportion according to Zooniverse  -->
-<!--      - have lower levels of concordance across Zoo coders -->
-
-<!-- - take all 5 Zoo judgments or 3 or 1 (may need to run the latter two a couple of times to check for potential variance -- if variance is small, we'll just say so, if it's large then we'll have to write a loop and report CIs) -->
 
@@
 Notice that clips are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority.
-To compare laboratory and Zooniverse annotations at the level of segments, chunk-level judgments are combined using the hierarchy Canonical>Non-canonical>Crying>Laughing."}
+To compare laboratory and Zooniverse annotations at the level of segments, chunk-level labels are combined using the hierarchy Canonical>Non-canonical>Crying>Laughing."}
 knitr::include_graphics("fig_levels.key.pdf")

### 2021-02-18 202010_jslhr/postprocess.R
fixed clean_class issue (old file that had only partial data)

@@ -1,0 +1,149 @@
+## PHASE 1 -- clean classifications
+
+read.csv("../data_analyses/output/zoo_anno_info_pu.csv",colClasses="factor",header=T,sep =",")->classifs
+
+dim(classifs)
+#[1] 169803     15
+
+library(rjson)
+getval=function(x) fromJSON(x)[[1]]$value
+getname=function(x) fromJSON(x)[[1]]$Name
+
+
+## Cleaning stuff up ##
+classifs$Answer=unlist(lapply(as.character(classifs[,"annotations"]), getval))
+table(classifs$Answer) #no weird answers
+
+classifs$filename=unlist(lapply(as.character(classifs[,"subject_data"]), getname))
+length(levels(factor(classifs$filename)))
+#33730 chunks
+
+# Remove additional votes when there are more than 5, 
+# and remove clips with fewer than 5
+table(classifs$filename)->sumvotes
+toomany=rownames(sumvotes)[sumvotes>5]
+length(toomany) #271 have too many votes
+justright=rownames(sumvotes)[sumvotes==5]
+length(justright) #33458 have the right amount
+toofew=rownames(sumvotes)[sumvotes<5]
+length(toofew)#1 chunk didn't receive 5 votes
+
+classifs$nclasschunk=NA
+clean=classifs[classifs$filename %in% justright,] #start by adding the normal ones
+for(thischunk in toomany){
+  classifs$nclasschunk[classifs$filename==thischunk]<-1:sum(classifs$filename==thischunk)
+  clean=rbind(clean,classifs[classifs$filename==thischunk & classifs$nclasschunk<=5 & !is.na(classifs$nclasschunk),])
+}
+
+#check that the procedure worked
+clean$filename=as.factor(as.character(clean$filename))
+table(clean$filename)->sumvotes
+sum(sumvotes>5)
+sum(sumvotes==5) #33729 chunks
+sum(sumvotes<5)
+
+rm(classifs) #clean up to make space
+
+zoosj <- read.csv("../data_analyses/output/zoo_subj_info_pu.csv",header=T,sep =",")
+merge(zoosj,clean,by="filename")->chunks
+dim(chunks)
+#[1] 168645     36
+levels(chunks$ChildID) 
+length(levels(chunks$ChildID))#all 20 kids here
+
+rm(clean) #clean up to make space
+
+write.csv(chunks,"../data_analyses/output/clean_classifications.csv")
+
[…]

### 2021-02-18 202010_jslhr/preprocess.R

@@ -1,143 +1,0 @@
-## PHASE 1 -- clean classifications
-zoosj <- read.csv("../data_analyses/output/zoo_subj_info_pu.csv",header=T,sep =",")
-
-read.csv("../data_analyses/output/zoo_anno_info_pu.csv",colClasses="factor",header=T,sep =",")->classifs
-
-dim(classifs)
-#[1] 168852     15
-
-library(rjson)
-getval=function(x) fromJSON(x)[[1]]$value
-getname=function(x) fromJSON(x)[[1]]$Name
-
-
-## Cleaning stuff up ##
-classifs$Answer=unlist(lapply(as.character(classifs[,"annotations"]), getval))
-table(classifs$Answer) #no weird answers
-
-classifs$filename=unlist(lapply(as.character(classifs[,"subject_data"]), getname))
-length(levels(factor(classifs$filename)))
-#33730 chunks
-
-# Remove additional votes when there are more than 5, 
-# and remove clips with fewer than 5
-table(classifs$filename)->sumvotes
-toomany=rownames(sumvotes)[sumvotes>5]
-length(toomany) #271 have too many votes -- now 194
-justright=rownames(sumvotes)[sumvotes==5]
-length(justright) #33458 have the right amount -- now 33535
-toofew=rownames(sumvotes)[sumvotes<5]
-length(toofew)#1 chunk didn't receive 5 votes
-
-classifs$nclasschunk=NA
-clean=classifs[classifs$filename %in% justright,] #start by adding the normal ones
-for(thischunk in toomany){
-  classifs$nclasschunk[classifs$filename==thischunk]<-1:sum(classifs$filename==thischunk)
-  clean=rbind(clean,classifs[classifs$filename==thischunk & classifs$nclasschunk<=5 & !is.na(classifs$nclasschunk),])
-}
-
-#check that the procedure worked
-clean$filename=as.factor(as.character(clean$filename))
-table(clean$filename)->sumvotes
-sum(sumvotes>5)
-sum(sumvotes==5) #33729 chunks
-sum(sumvotes<5)
-
-rm(classifs) #clean up to make space
-
-merge(zoosj,clean,by="filename")->chunks
-dim(chunks)
-#[1] 168645     36
-
-rm(clean) #clean up to make space
-
-write.csv(chunks,"../data_analyses/output/clean_classifications.csv")
-
-## PHASE 2 -- Generate views on the data: majority judgment on the chunk
-
[…]

### 2021-02-18 data_analyses/code/generate_jslhr_data.R

@@ -8,6 +8,7 @@ origmetadata <- read.csv("../files_from_elsewhere/metadata_all_PU.csv",colClasse
 dim(origmetadata)
 origmetadata$filename=paste0(origmetadata$AudioData,".mp3")
 length(levels(factor(origmetadata$filename)))
+levels(origmetadata$ChildID)
 #33730
 
 # #file that has "subject" information -- in Zooniverse terms, that is info about the file that was classified
@@ -31,7 +32,10 @@ for(mych in names(mytab[mytab==2])){
 dim(zoosj_PU)
 #33730
 
-write.csv(zoosj_PU,"../output/zoo_subj_info_pu.csv")
+write.csv(zoosj_PU,"../output/zoo_subj_info_pu.csv") #all 20 kids are in here
+
+
+
 
 read.csv("../output/zoo_subj_info_pu.csv",colClasses="factor",header=T,sep =",")->zoosj
 
@@ -47,17 +51,16 @@ length(levels(factor(classifs$subject_id)))
 
 write.csv(classifs,"../output/zoo_anno_info_pu.csv")
 
+
 #An impressive total of `r length(levels(factor(judgments$user_id)))` 
 #individual Zooniverse users provided `r dim(judgments)[1]` judgments on 
 #`r length(levels(factor(judgments$subject_ids)))` 500-ms chunks
-keyinfo=c(length(levels(factor(judgments$user_id))),
-          dim(judgments)[1],
-          length(levels(factor(judgments$subject_ids))))
+keyinfo=c(length(levels(factor(classifs$user_id))),
+          dim(classifs)[1],
+          length(levels(factor(classifs$subject_ids))))
 names(keyinfo)<-c("nusers","njudgments","nchunks")
 
 write.csv(keyinfo,"../output/key_info.csv")
-
-
 
 
 

### 2021-02-18 202010_jslhr/paper.Rmd
ready for final proofing

@@
 
-#fix typo
-chunks$segmentId_DB=gsub("(DB.)(.)$","\\1",chunks$segmentId_DB)
-
 read.csv("../data_analyses/output/zoo_lab_maj_judgments.csv")->data_all
 
-label_options=c("Canonical"   ,     "Crying"  ,        "Junk"  ,    "Laughing", "Non-Canonical")
+label_options=c("Canonical" , "Non-Canonical"  ,     "Crying"   ,    "Laughing",        "Junk" )
 
@@
 Yet another metric that could be extracted to estimate relative linguistic complexity is the Canonical Proportion: the proportion of vocalizations that contain a canonical transition or syllable out of all linguistic (canonical and non-canonical) vocalizations.
-@cychosz2019canonical found that the Canonical Proportion extracted from daylong recordings was significantly correlated with age in a multicultural and multilingual sample of children up to 3 years of age, and thus well beyond the babbling period, and into first words and word combinations.
+@cychosz2021canonical found that the Canonical Proportion extracted from daylong recordings was significantly correlated with age in a multicultural and multilingual sample of children up to 3 years of age, and thus well beyond the babbling period, and into first words and word combinations.
 That study built on a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this prior work focuses on babbling and thus typically on infants under one year of age.
@@
 When compared to more primitive sounds such as squeals, or isolated vowels, canonical syllables show a higher complexity given the smooth articulatory transition between a consonant and a vowel (or vice versa).
-Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development [@oller2000emergence;@cychosz2019canonical].
+Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development [@oller2000emergence;@cychosz2021canonical].
 In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization counts in a sample of children diagnosed with an autism spectrum disorder [@mcdaniel2020predicting].
@@
 There is one previous study that attempted this approach.
-@cychosz2019canonical drew child vocalization data from a diverse set of corpora centered on children learning one of four languages: English, Tseltal, Tsimane', and Yélî.
+@cychosz2021canonical drew child vocalization data from a diverse set of corpora centered on children learning one of four languages: English, Tseltal, Tsimane', and Yélî.
 For the English and Tsimane' corpora, vocalizations were automatically identified using LENA&reg;, whereas the other two were extracted through manual segmentation.
@@
 We hope that this methodology will open the road to larger scale analyses of children's vocalizations as captured by wearables.
-In addition, we went beyond @cychosz2019canonical's study in two important ways.
+In addition, we went beyond @cychosz2021canonical's study in two important ways.
 First, we relied on the largest and best established citizen science platform: Zooniverse hosts more than 1.6 million users from diverse walks of life, and it offers a completely automatized API system to more easily scale tasks in a transparent and cumulative science fashion.
-In contrast, @cychosz2019canonical actively recruited people to provide annotations, stating that "[a]nnotators included language and speech researchers, undergraduate students, research assistants, and other interested parties, totaling 136 unique annotators".
+In contrast, @cychosz2021canonical actively recruited people to provide annotations, stating that "[a]nnotators included language and speech researchers, undergraduate students, research assistants, and other interested parties, totaling 136 unique annotators".
 A process that draws from research assistants relies on expert resources, whose time is expensive.
 Moreover, the people these researchers recruit tend to be more accustomed to child vocalization data, which may improve their annotation performance, and thus provide an overestimate of the quality of the annotations that can be realistically done within a truly citizen science framework.
-Those authors also had to rely on manual extraction of the resulting data from the iHearUplay team, which added to the administrative load.
 Second, a core goal of the present work was to determine how citizen scientists' annotations fare compared to the current gold standard, laboratory annotations.
-@cychosz2019canonical did not have a gold standard for their data.
+@cychosz2021canonical did not have a gold standard for their data.
 
@@
 
-It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels.
+It was important for generalizability purposes to examine data from children with variable biological age, mental age, and vocal maturity levels.
 We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA&reg; [see meta-analyses in @wang2020meta;@cristia2020accuracy]; as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range.
-Some previous work had shown that vocal development proceeded differently when compared to a normative sample [@hamrick2019capturing;@hamrick2019measurement].
+Some previous work had shown that vocal development proceeded differently among children with an Angelman syndrome diagnosis when compared to a normative sample [@hamrick2019capturing;@hamrick2019measurement].
 Generally, studies on vocal development among children with genetic disorders reveal lower volubility and lower prevalence of canonical syllables [@rankine2017language;@belardi2017retrospective; see also @roche2018early for a review of retrospective video analysis of children diagnosed with Autism Spectrum Disorder, Rett Syndrome, or Fragile X, documenting these and other vocal characteristics].
 Low volubility is one of the contexts in which daylong recordings may be particularly advantageous over alternative recording methods [see @rankine2017language for discussion], since it is challenging to measure vocal development accurately based on shorter recordings (as the child may not vocalize at all in a 1-2h recording).
-Additionally, the ideal method will be sensitive not only to the large changes that occur over normative development, but also to the potentially smaller individual differences found among children.
+Additionally, the ideal method is sensitive not only to the large changes that occur over normative development, but also to the potentially smaller individual differences found among children.
 Such a level of sensitivity is important if these methods are ever to be used to detect potential effects of intervention.
@@
 Specifically, the segments extracted were automatically cut into clips of exactly 500 ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform.
-We will call this level *chunk*, to convey that they are part of a bigger unit (the LENA&reg;-identified "segment").
+We call this level *chunk*, to convey that they are part of a bigger unit (the LENA&reg;-identified "segment").
 Note that segments and chunks are collectively referred to as *clips* because they are parts of the longer audio recording.
@@
 
-Citizen scientists could access our project, "Maturity of Baby Sounds", by logging in, or as anonymous participants.
-Before starting the annotation, participants were given a quick tutorial which walked them through the steps of the annotation workflow (i.e., how to play a sound, how to make a selection; see Figure 1).
+Citizen scientists could access our project, "Maturity of Baby Sounds", by logging in, or as anonymous annotators.
+Before starting the annotation, citizen scientists were given a quick tutorial which walked them through the steps of the annotation workflow (i.e., how to play a sound, how to make a selection; see Figure 1).
 In this tutorial, we included one audio example for each category in order to make the classification task as smooth as possible.
@@
 
-<!-- lh said: I'm also wondering what the breakdown of lab annotations was for the Zooniverse segments without majority agreement -->
+```{r clean up, echo=F}
+#dim(data_all) #11980    16
+# remove non-majority labels from lab
+data_all[data_all$Num_Agreement>0,]->data_all
+#dim(data_all) # 11647    16
+
+# remove non-majority labels from zooniverse
+data_all[data_all$Zoon_classif %in% label_options,]->data_all
+#dim(data_all) # 11268    16
+
+#and reset the factors for cleanliness
+data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=label_options)
+data_all$lab=factor(data_all$lab, levels=label_options)
+sample_data<-cbind(data_all$lab,data_all$Zoon_classif)
+
+table_zoon=round(table(data_all$Zoon)/sum(table(data_all$Zoon))*100,2)
+table_lab=round(table(data_all$lab)/sum(table(data_all$lab))*100,2)
+```
 
@@
 Figure 2 conveys the relationship between chunks and segments, and how judgments are combined into labels.
+After removing segments for which there was no majority label in the lab and/or Zooniverse coding, we were left with data for `r dim(data_all)[1]` segments.
 
 ----- INSERT FIGURE 2 HERE -----
-
-```{r clean up, echo=F}
-# remove non-majority labels from lab
-data_all[data_all$Num_Agreement>2,]->data_all
-
-# remove non-majority labels from zooniverse
-data_all[data_all$Zoon_classif %in% label_options,]->data_all
-
-#and reset the factors for cleanliness
-data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=label_options)
-data_all$lab=factor(data_all$lab, levels=label_options)
-sample_data<-cbind(data_all$lab,data_all$Zoon_classif)
-
-table_zoon=round(table(data_all$Zoon)/sum(table(data_all$Zoon))*100)
-table_lab=round(table(data_all$lab)/sum(table(data_all$lab))*100)
-```
 
@@
   thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"])
-    myr[thisvar,"AS"]=round(thiscor$estimate,3)
+    myr[thisvar,"AS"]=printr(thiscor$estimate)
     mylb[thisvar,"AS"]=round(thiscor$conf.int[1],3)
@@
   thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"])
-    myr[thisvar,"LR"]=round(thiscor$estimate,3)
+    myr[thisvar,"LR"]=printr(thiscor$estimate)
     mylb[thisvar,"LR"]=round(thiscor$conf.int[1],3)
@@
 
-    #add r to figure
-
 }
+
+#matrix that will hold all r results
+corage<-t(myr)[,c("l_lp","z_lp","l_cp","z_cp")]
+
 ```
@@
 Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups.
-There was a near-zero relationship among the children diagnosed with Angelman syndrome r(`r myn["l_lp","AS"]`) = `r printr(myr["l_lp","AS"])`, CI [`r printr(mylb["l_lp","AS"])`,`r printr(myhb["l_lp","AS"])`], *p* = `r printp(myp["l_lp","AS"])`]; and a significant association among low-risk control children r(`r myn["l_lp","LR"]`) = `r printr(myr["l_lp","LR"])`, CI [`r printr(mylb["l_lp","LR"])`,`r printr(myhb["l_lp","LR"])`], *p* = `r printp(myp["l_lp","LR"])`].
-The Canonical Proportion exhibited non-significant developmental decreases among children diagnosed with Angelman syndrome r(`r myn["l_cp","AS"]`) = `r printr(myr["l_cp","AS"])`, CI [`r printr(mylb["l_cp","AS"])`,`r printr(myhb["l_cp","AS"])`], *p* = `r printp(myp["l_cp","AS"])`]; and marginal developmental increases among low-risk control r(`r myn["l_cp","LR"]`) = `r printr(myr["l_cp","LR"])`, CI [`r printr(mylb["l_cp","LR"])`,`r printr(myhb["l_cp","LR"])`], *p* = `r printp(myp["l_cp","LR"])`].
-
-Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(`r myn["z_lp","AS"]`) = `r printr(myr["z_lp","AS"])`, CI [`r printr(mylb["z_lp","AS"])`,`r printr(myhb["z_lp","AS"])`], *p* = `r printp(myp["z_lp","AS"])`]; whereas low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r printr(myr["z_lp","LR"])`, CI [`r printr(mylb["z_lp","LR"])`,`r printr(myhb["z_lp","LR"])`], *p* = `r printp(myp["z_lp","LR"])`].
-Similarly, there were non-significant developmental decreases in the Canonical Proportion among children with Angelman syndrome r(`r myn["z_cp","AS"]`) = `r printr(myr["z_cp","AS"])`, CI [`r printr(mylb["z_cp","AS"])`,`r printr(myhb["z_cp","AS"])`], *p* = `r printp(myp["z_cp","AS"])`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r printr(myr["z_cp","LR"])`, CI [`r printr(mylb["z_cp","LR"])`,`r printr(myhb["z_cp","LR"])`], *p* = `r printp(myp["z_cp","LR"])`].
+There was a near-zero relationship among the children diagnosed with Angelman syndrome r(`r myn["l_lp","AS"]`) = `r (myr["l_lp","AS"])`, CI [`r printr(mylb["l_lp","AS"])`,`r printr(myhb["l_lp","AS"])`], *p* = `r printp(myp["l_lp","AS"])`]; and a significant association among low-risk control children r(`r myn["l_lp","LR"]`) = `r (myr["l_lp","LR"])`, CI [`r printr(mylb["l_lp","LR"])`,`r printr(myhb["l_lp","LR"])`], *p* = `r printp(myp["l_lp","LR"])`].
+The Canonical Proportion exhibited a non-significant developmental decrease among children diagnosed with Angelman syndrome r(`r myn["l_cp","AS"]`) = `r (myr["l_cp","AS"])`, CI [`r printr(mylb["l_cp","AS"])`,`r printr(myhb["l_cp","AS"])`], *p* = `r printp(myp["l_cp","AS"])`]; and a marginal developmental increase among low-risk control r(`r myn["l_cp","LR"]`) = `r (myr["l_cp","LR"])`, CI [`r printr(mylb["l_cp","LR"])`,`r printr(myhb["l_cp","LR"])`], *p* = `r printp(myp["l_cp","LR"])`].
+
+Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(`r myn["z_lp","AS"]`) = `r (myr["z_lp","AS"])`, CI [`r printr(mylb["z_lp","AS"])`,`r printr(myhb["z_lp","AS"])`], *p* = `r printp(myp["z_lp","AS"])`]; whereas low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r (myr["z_lp","LR"])`, CI [`r printr(mylb["z_lp","LR"])`,`r printr(myhb["z_lp","LR"])`], *p* = `r printp(myp["z_lp","LR"])`].
+Similarly, there was a non-significant developmental decrease in the Canonical Proportion among children with Angelman syndrome r(`r myn["z_cp","AS"]`) = `r (myr["z_cp","AS"])`, CI [`r printr(mylb["z_cp","AS"])`,`r printr(myhb["z_cp","AS"])`], *p* = `r printp(myp["z_cp","AS"])`]; and a marginal developmental increase among low-risk control children, r(`r myn["z_cp","LR"]`) = `r (myr["z_cp","LR"])`, CI [`r printr(mylb["z_cp","LR"])`,`r printr(myhb["z_cp","LR"])`], *p* = `r printp(myp["z_cp","LR"])`].
 
@@
 
-Next, we discuss the correspondence between citizen science classifications and the laboratory gold standard, at the level of individual segments.
-Results were visualized with a confusion matrix showing precision and recall (Figure 4): the diagonal elements show the number of correct segment-level classifications for each class while the off-diagonal elements show non-matching classifications.
+Next, we discuss the correspondence between citizen science labels and the laboratory gold standard, at the level of individual segments.
+Results were visualized with a confusion matrix showing precision and recall (Figure 4): the diagonal elements show the number of correct segment-level labels for each class while the off-diagonal elements show non-matching labels.
 
@@
 #mycf
-```
-
------ INSERT FIGURE 4 HERE -----
-
-These visualizations suggest that performance is moderate to good, which was confirmed via statistical analyses.
-The overall (weighted) accuracy is `r round(mycf$overall["Accuracy"]*100)`%, CI = [`r round(mycf$overall["AccuracyLower"]*100)`,`r round(mycf$overall["AccuracyUpper"]*100)`],  kappa is `r printr(mycf$overall["Kappa"])`, and the Gwet's AC1 coefficient is `r printr(mygac$est)`, CI = [`r printr(mygac$lb)`,`r printr(mygac$ub)`].
-
-### Child level descriptors
-
-```{r matr2}
-#create matrix to hold r results
-myr=matrix(NA,nrow=2,ncol=3)
-rownames(myr)<-c("lp","cp")
-colnames(myr)<-c("AS","LR","all")
-mylb=myhb=myn=myp =myr #this is an ugly way of creating 5 matrices
-
-  #get cor properties
-  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_lp"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
-    myr[1,"AS"]=round(thiscor$estimate,3)
-    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
-    myn[1,"AS"]=round(thiscor$parameter,3)
-    myp[1,"AS"]=round(thiscor$p.value,3)
-
-  #repeat for LR
-
-      #get cor properties
-  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"],proportions[proportions$Diagnosis!="AngelmanSyndrome","z_lp"])
-    myr[1,"LR"]=round(thiscor$estimate,3)
-    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
-    myn[1,"LR"]=round(thiscor$parameter,3)
-    myp[1,"LR"]=round(thiscor$p.value,3)
-
-#repeat for all cor
-
-      thiscor=cor.test(proportions$z_lp,proportions$l_lp)
-    myr[1,"all"]=round(thiscor$estimate,3)
-    mylb[1,"all"]=round(thiscor$conf.int[1],3)
-    myhb[1,"all"]=round(thiscor$conf.int[2],3)
-    myn[1,"all"]=round(thiscor$parameter,3)
-    myp[1,"all"]=round(thiscor$p.value,3)
-
-    # repeat whole for cp
-
-  #get cor properties
-  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_cp"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
-    myr[2,"AS"]=round(thiscor$estimate,3)
-    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
-    myn[2,"AS"]=round(thiscor$parameter,3)
-    myp[2,"AS"]=round(thiscor$p.value,3)
-
-  #repeat for LR
-
-      #get cor properties
-  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"],proportions[proportions$Diagnosis!="AngelmanSyndrome","z_cp"])
-    myr[2,"LR"]=round(thiscor$estimate,3)
-    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
-    myn[2,"LR"]=round(thiscor$parameter,3)
-    myp[2,"LR"]=round(thiscor$p.value,3)
-
-#repeat for all cor
-
-      thiscor=cor.test(proportions$z_cp,proportions$l_cp)
-    myr[2,"all"]=round(thiscor$estimate,3)
-    mylb[2,"all"]=round(thiscor$conf.int[1],3)
-    myhb[2,"all"]=round(thiscor$conf.int[2],3)
-    myn[2,"all"]=round(thiscor$parameter,3)
-    myp[2,"all"]=round(thiscor$p.value,3)
-
-mycortab=t(myr)
-rownames(mycortab)<-paste("All seg",rownames(mycortab))
-```
-
-Although the classification at the segment level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of children's individual development.
-We illustrate this association in Figure 5.
-Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r printr(myr["lp","all"])`, CI [`r printr(mylb["lp","all"])`,`r printr(myhb["lp","all"])`], *p* `r printp(myp["lp","all"])`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r printr(myr["lp","AS"])`, CI [`r printr(mylb["lp","AS"])`,`r printr(myhb["lp","AS"])`], *p* = `r printp(myp["lp","AS"])`]; low-risk control r(`r myn["lp","LR"]`) = `r printr(myr["lp","LR"])`, CI [`r printr(mylb["lp","LR"])`,`r printr(myhb["lp","LR"])`], *p* = `r printp(myp["lp","LR"])`].
-
------ INSERT FIGURE 5 HERE -----
-
-Similarly, a strong positive correlation is found in the Canonical Proportion r(`r myn["cp","all"]`) = `r printr(myr["cp","all"])`, CI [`r printr(mylb["cp","all"])`,`r printr(myhb["cp","all"])`], *p* `r printp(myp["cp","all"])`].
-When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman syndrome: r(`r myn["cp","AS"]`) = `r printr(myr["cp","AS"])`, CI [`r printr(mylb["cp","AS"])`,`r printr(myhb["cp","AS"])`], *p* = `r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r printr(myr["cp","LR"])`, CI [`r printr(mylb["cp","LR"])`,`r printr(myhb["cp","LR"])`], *p* `r printp(myp["cp","LR"])`.
-
-## Additional analyses
-
-In this section, we report on explorations of under what conditions Zooniverse labels more closely aligned with laboratory labels.
-In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as "Junk" (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora [@cychosz2019canonical].
-A high proportion of "Junk" may indicate that automated segmentation was errorful for those children, and may be a sign that the rest of the data could be compromised as well.
-
-```{r prelim-Junk, echo=F}
-JunkXdiag=t.test(proportions$z_Junk ~ proportions$Diagnosis)
-
-#create matrix to hold r results
-myr=matrix(NA,nrow=2,ncol=2)
-rownames(myr)<-c("lp","cp")
-colnames(myr)<-c("lower_Junk","higher_Junk")
-mylb=myhb=myn=myp =myr #this is an ugly way of creating 5 matrices
-
-  #get cor properties
-  thiscor=cor.test(proportions[proportions$median_Junk=="lower_Junk","z_lp"],proportions[proportions$median_Junk=="lower_Junk","l_lp"])
-    myr[1,"lower_Junk"]=round(thiscor$estimate,3)
-    mylb[1,"lower_Junk"]=round(thiscor$conf.int[1],3)
-    myhb[1,"lower_Junk"]=round(thiscor$conf.int[2],3)
-    myn[1,"lower_Junk"]=round(thiscor$parameter,3)
-    myp[1,"lower_Junk"]=round(thiscor$p.value,3)
-
-  #repeat for LR
-
-      #get cor properties
-  thiscor=cor.test(proportions[proportions$median_Junk!="lower_Junk","l_lp"],proportions[proportions$median_Junk!="lower_Junk","z_lp"])
-    myr[1,"higher_Junk"]=round(thiscor$estimate,3)
-    mylb[1,"higher_Junk"]=round(thiscor$conf.int[1],3)
-    myhb[1,"higher_Junk"]=round(thiscor$conf.int[2],3)
-    myn[1,"higher_Junk"]=round(thiscor$parameter,3)
-    myp[1,"higher_Junk"]=round(thiscor$p.value,3)
-
-    # repeat whole for cp
-
-  #get cor properties
-  thiscor=cor.test(proportions[proportions$median_Junk=="lower_Junk","z_cp"],proportions[proportions$median_Junk=="lower_Junk","l_cp"])
-    myr[2,"lower_Junk"]=round(thiscor$estimate,3)
-    mylb[2,"lower_Junk"]=round(thiscor$conf.int[1],3)
-    myhb[2,"lower_Junk"]=round(thiscor$conf.int[2],3)
-    myn[2,"lower_Junk"]=round(thiscor$parameter,3)
-    myp[2,"lower_Junk"]=round(thiscor$p.value,3)
-
-  #repeat for higher_Junk
-
-      #get cor properties
-  thiscor=cor.test(proportions[proportions$median_Junk!="lower_Junk","l_cp"],proportions[proportions$median_Junk!="lower_Junk","z_cp"])
-    myr[2,"higher_Junk"]=round(thiscor$estimate,3)
-    mylb[2,"higher_Junk"]=round(thiscor$conf.int[1],3)
-    myhb[2,"higher_Junk"]=round(thiscor$conf.int[2],3)
-    myn[2,"higher_Junk"]=round(thiscor$parameter,3)
-    myp[2,"higher_Junk"]=round(thiscor$p.value,3)
-```
-
-We investigated this hypothesis by calculating the proportion of their data labeled as "Junk" for each individual child.
-There was no significant difference in the proportions of their data labeled as "Junk" for children with Angelman syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the low-risk group (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, *p* = `r printp(JunkXdiag$p.value)`.
-We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk".
-Results were similar across these two post-hoc subgroups (see Figure 6).
-For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["lp","lower_Junk"]`) = `r printr(myr["lp","lower_Junk"])`, CI [`r printr(mylb["lp","lower_Junk"])`,`r printr(myhb["lp","lower_Junk"])`], *p* = `r printp(myp["lp","lower_Junk"])`]; and for the higher Junk group it was r(`r myn["lp","higher_Junk"]`) = `r printr(myr["lp","higher_Junk"])`, CI [`r printr(mylb["lp","higher_Junk"])`,`r printr(myhb["lp","higher_Junk"])`], *p* `r printp(myp["lp","higher_Junk"])`.
-For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["cp","lower_Junk"]`) = `r printr(myr["cp","lower_Junk"])`, CI [`r printr(mylb["cp","lower_Junk"])`,`r printr(myhb["cp","lower_Junk"])`], *p* `r printp(myp["cp","lower_Junk"])`]; and for the higher Junk group it was r(`r myn["cp","higher_Junk"]`) = `r printr(myr["cp","higher_Junk"])`, CI [`r printr(mylb["cp","higher_Junk"])`,`r printr(myhb["cp","higher_Junk"])`], *p* `r printp(myp["cp","higher_Junk"])`.
-Thus, it does not seem that a higher proportion of "Junk" labels is an index of low quality data.
-
------ INSERT FIGURE 6 HERE -----
-
-Above, we concluded that derived metrics integrating information across audio clips (Linguistic and Canonical Proportion, which can be derived from segment- or chunk-level labels) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA&reg; segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
-As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed labels across chunks associated to the same segment.
-We therefore repeated our analyses but instead of deriving our proportions for the Zooniverse data from the segment-level composite, we did it based on the individual chunk-level annotations.
-To facilitate comparison, correlations across laboratory- and Zooniverse-derived proportions are provided in Table 2.
-
------ INSERT TABLE 2 HERE -----
-
-```{r bychunk-age, echo=F}
-read.csv("zoo_chunk_majority_agreement.csv")->chunk
-bychild=table(chunk$Answer,chunk$ChildID)
-cans=bychild["Canonical",]
-noncans=bychild["Non-Canonical",]
-Junk=bychild["Junk",]
-sum_bychild=table(chunk$ChildID)
-
-ling_prop =(cans+noncans)/(sum_bychild-Junk)
-can_prop =cans/(cans+noncans)
-
-#check that they match before merging
-#demo_data$ChildID==names(ling_prop)
-#demo_data$ChildID==names(can_prop)
-
-props_bychild_chunk=cbind(ling_prop,can_prop,demo_data)
-colnames(props_bychild_chunk)[1:2]<-c("chi","ling_prop")
-
-z_lp_as=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
-
-z_lp_lr=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
-
-z_cp_as=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
-
-z_cp_lr=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
-```
-
-Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman syndrome group: r(`r z_lp_as$parameter`) = `r printr(z_lp_as$estimate)`, CI [`r printr(z_lp_as$conf.int[1])`,`r printr(z_lp_as$conf.int[2])`], *p* = `r printp(z_lp_as$p.value)`], but increased with age  in the low-risk control group: r(`r z_lp_lr$parameter`) = `r printr(z_lp_lr$estimate)`, CI [`r printr(z_lp_lr$conf.int[1])`,`r printr(z_lp_lr$conf.int[2])`], *p* = `r printp(z_lp_lr$p.value)`.
-Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman syndrome r(`r z_cp_as$parameter`) = `r printr(z_cp_as$estimate)`, CI [`r printr(z_cp_as$conf.int[1])`,`r printr(z_cp_as$conf.int[2])`], *p* = `r printp(z_cp_as$p.value)`]; and developmental increases among low-risk control r(`r z_cp_lr$parameter`) = `r printr(z_cp_lr$estimate)`, CI [`r printr(z_cp_lr$conf.int[1])`,`r printr(z_cp_lr$conf.int[2])`], *p* = `r printp(z_cp_lr$p.value)`.
-
-```{r zoo-lab-chunk, echo=F}
-
-#create matrix to hold r results
-myr=matrix(NA,nrow=2,ncol=3)
-rownames(myr)<-c("lp","cp")
-colnames(myr)<-c("AS","LR","all")
-mylb=myhb=myn=myp =myr #this is an ugly way of creating 5 matrices
-
-proportions=proportions[order(proportions$ChildID),]
-
-  #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
-    myr[1,"AS"]=round(thiscor$estimate,3)
-    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
-    myn[1,"AS"]=round(thiscor$parameter,3)
-    myp[1,"AS"]=round(thiscor$p.value,3)
-
-  #repeat for LR
-
-      #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"])
-    myr[1,"LR"]=round(thiscor$estimate,3)
-    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
-    myn[1,"LR"]=round(thiscor$parameter,3)
-    myp[1,"LR"]=round(thiscor$p.value,3)
-
-#repeat for all cor
-
-      thiscor=cor.test(props_bychild_chunk$ling_prop,proportions$l_lp)
-    myr[1,"all"]=round(thiscor$estimate,3)
-    mylb[1,"all"]=round(thiscor$conf.int[1],3)
-    myhb[1,"all"]=round(thiscor$conf.int[2],3)
-    myn[1,"all"]=round(thiscor$parameter,3)
-    myp[1,"all"]=round(thiscor$p.value,3)
-
-    # repeat whole for cp
-
-  #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
-    myr[2,"AS"]=round(thiscor$estimate,3)
-    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
-    myn[2,"AS"]=round(thiscor$parameter,3)
-    myp[2,"AS"]=round(thiscor$p.value,3)
-
-  #repeat for LR
-
-      #get cor properties
-  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"])
-    myr[2,"LR"]=round(thiscor$estimate,3)
-    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
-    myn[2,"LR"]=round(thiscor$parameter,3)
-    myp[2,"LR"]=round(thiscor$p.value,3)
-
-#repeat for all cor
-
-      thiscor=cor.test(props_bychild_chunk$can_prop,proportions$l_cp)
-    myr[2,"all"]=round(thiscor$estimate,3)
-    mylb[2,"all"]=round(thiscor$conf.int[1],3)
-    myhb[2,"all"]=round(thiscor$conf.int[2],3)
-    myn[2,"all"]=round(thiscor$parameter,3)
-    myp[2,"all"]=round(thiscor$p.value,3)
-
- mycortab=rbind(mycortab,t(myr))
-rownames(mycortab)[4:6]<-paste("Chunks",rownames(t(myr)))
-
-```
-
-As for correlations between Zooniverse- and laboratory-derived metrics (e.g., linguistic proportion as estimated using Zooniverse coding, versus laboratory coding), we observe very similar levels of correlation: Linguistic Proportion overall r(`r myn["lp","all"]`) = `r printr(myr["lp","all"])`, CI [`r printr(mylb["lp","all"])`,`r printr(myhb["lp","all"])`], *p* `r printp(myp["lp","all"])`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r printr(myr["lp","AS"])`, CI [`r printr(mylb["lp","AS"])`,`r printr(myhb["lp","AS"])`], *p* = `r printp(myp["lp","AS"])`]; low-risk control r(`r myn["lp","LR"]`) = `r printr(myr["lp","LR"])`, CI [`r printr(mylb["lp","LR"])`,`r printr(myhb["lp","LR"])`], *p* = `r printp(myp["lp","LR"])`].
-As for Canonical Proportion overall r(`r myn["cp","all"]`) = `r printr(myr["cp","all"])`, CI [`r printr(mylb["cp","all"])`,`r printr(myhb["cp","all"])`], *p* `r printp(myp["cp","all"])`; Angelman syndrome: r(`r myn["cp","AS"]`) = `r printr(myr["cp","AS"])`, CI [`r printr(mylb["cp","AS"])`,`r printr(myhb["cp","AS"])`], *p* = `r printp(myp["cp","AS"])`];  low-risk control infants r(`r myn["cp","LR"]`) = `r printr(myr["cp","LR"])`, CI [`r printr(mylb["cp","LR"])`,`r printr(myhb["cp","LR"])`], *p* `r printp(myp["cp","LR"])`.
-
-Since all of these results are very similar to those obtained when collapsing chunk-level labels to generate segment-level labels, we conclude that in the future this step can be skipped.
-Instead, researchers can derive Linguistic and Canonical Proportions directly from citizen scientists' chunk-level labels [which was in fact what @cychosz2019canonical did].
-
-```{r downsample}
-
-#we start from the data base that has all the info
-nsegs=data.frame(table(data_all$ChildID))
-colnames(nsegs)<-c("ChildID","nsegs")
-
-merge(nsegs,demo_data)->nsegs
-nseg_t=t.test(nsegs$nsegs~nsegs$Diagnosis)
-
-resampling=matrix(NA,nrow=50,ncol=6)
-colnames(resampling)<-c("lp_r_all","lp_r_as","lp_r_lr","cp_r_all","cp_r_as","cp_r_lr")
-for(i in 1:50){
-  sel_seg=NULL
-  for(thischi in levels(data_all$ChildID)) sel_seg<-c(sel_seg,as.character(data_all$segmentId_DB[sample(data_all$segmentId_DB[data_all$ChildID==thischi],100)]))
-
-  sampling=data_all[as.character(data_all$segmentId_DB) %in% as.character(sel_seg),]
-
-  #get the ns by child, then calculate the linguistic proportion & canonical proportion, separately for zooniverse & lab
-  ztab=table(sampling$ChildID,sampling$Zoon_classif)
-  z_lp =rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
-  z_cp =ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
-
-  ltab=table(sampling$ChildID,sampling$lab)
-  l_lp =rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
-  l_cp =ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
-
-  thistab=cbind(demo_data,z_lp,z_cp,l_lp,l_cp)
-
-  resampling[i,"lp_r_all"]=cor.test(thistab$z_lp,thistab$l_lp)$estimate
-  resampling[i,"lp_r_as"]=cor.test(thistab$z_lp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
-  resampling[i,"lp_r_lr"]=cor.test(thistab$z_lp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate
-
-   resampling[i,"cp_r_all"]=cor.test(thistab$z_cp,thistab$l_cp)$estimate
-  resampling[i,"cp_r_as"]=cor.test(thistab$z_cp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
-  resampling[i,"cp_r_lr"]=cor.test(thistab$z_cp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate
- }
-myr=colMeans(resampling)
-mycortab=rbind(mycortab,cbind(myr[c("lp_r_as" ,"lp_r_lr","lp_r_all")],myr[c("cp_r_as" ,"cp_r_lr","cp_r_all")]))
-rownames(mycortab)[7:9]<-c("100 seg AS","100 seg LR","100 seg all")
-round(mycortab,3)->mycortab
-
-resampling=data.frame(resampling)
-```
-
-Next, we looked at whether having more segments from each child may lead to more reliable metrics.
-We observed a non-significant trend [Welch t(`r nseg_t$parameter`)=`r nseg_t$statistic`] for lower number of segments in the Angelman syndrome group (M = `r nseg_t$estimate[1]`, range `r min(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`) than among low-risk control infants (M = `r nseg_t$estimate[2]`, range `r min(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`), likely because of the lower volubility of the former children.
-Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications [@cychosz2019canonical].
-We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children [following @cychosz2019canonical].
-We randomly sampled 100 segments from each child's data, and recalculated the Linguistic and Canonical Proportions based only on these associated labels.
-We repeated this process 50 times, to assess the extent to which the association between laboratory- and Zooniverse-derived metrics varied in each random sample (e.g., the correlation between Linguistic Proportion as estimated using Zooniverse coding on the one hand, and Linguistic Proportion as estimated using laboratory coding on the other).
-
-The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table 2), particularly for the low-risk younger infants, and for Linguistic Proportion.
-This may indicate that, especially in some groups of infants, 100 segments may not be sufficient to capture a stable estimate of the child's Linguistic and Canonical Proportions.
-We also observed quite a bit of variance across the 50 runs.
-For Linguistic Proportion, the range of correlations found for all infants was [`r printr(min(resampling$lp_r_all))`,`r printr(max(resampling$lp_r_all))`]; for the Angelman syndrome group [`r printr(min(resampling$lp_r_as))`,`r printr(max(resampling$lp_r_as))`]; for the low-risk group [`r printr(min(resampling$lp_r_lr))`,`r printr(max(resampling$lp_r_lr))`].
-For Canonical Proportion, the range of correlations found for all infants was [`r printr(min(resampling$cp_r_all))`,`r printr(max(resampling$cp_r_all))`]; for the Angelman syndrome group [`r printr(min(resampling$cp_r_as))`,`r printr(max(resampling$cp_r_as))`]; for the low-risk group [`r printr(min(resampling$cp_r_lr))`,`r printr(max(resampling$cp_r_lr))`].
-
-## Discussion
-
-In the present study we assessed the extent to which child vocalizations from LENA&reg; daylong recordings can be accurately described based on classifications collected using the citizen science platform Zooniverse.
-Our recordings came from children diagnosed with a genetic syndrome associated with severe language disorder, as well as low-risk controls, and included a range of different ages.
-Excerpts of these recordings had previously been annotated by trained experts in the lab, and served as the "gold standard".
-
-In descriptive terms, we found that category frequency was similar across laboratory and citizen science annotations.
-Correlations between age and our two child-level metrics revealed different developmental patterns between our sample of older children with a diagnosis of Angelman syndrome and younger low-risk controls.
-Given that the two groups of children differ on both of these features (age and diagnosis), we cannot empirically tease apart these two differences.
-Nonetheless, inspection of Figure 3 suggests different interpretations for the two derived metrics.
-Overall, the pattern for Linguistic Proportion is compatible with a non-linear trajectory common to the two groups, whereby the Linguistic Proportion increases rapidly up to 20 months, and hovers around 90-100% thereafter.
-In this case, the lack of correlation with age in the Angelman syndrome group suggests that there is little systematic variance in this metric after 20 months of age.
-In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and relatively stable levels of Canonical Proportion among the children in the Angelman syndrome group.
-
-Next, we looked at agreement between laboratory and Zooniverse annotation at the level of individual segments.
-We found that some categories had high precision and recall -- notably Non-canonical segments were detected quite accurately, and to a lesser extent the same was true for Canonical segments.
-This is interesting because it suggests that Zooniverse annotations may be trusted particularly for detecting these two types of speech-like segments.
-
-In other cases, precision was fairly high but recall was lower.
-This was the case of the "Junk/Don't mark" category.
-Higher precision and low recall indicates that when Zooniverse annotations return a "Junk" judgment, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as "Don't mark" in the laboratory.
-Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the "Junk" category less) than laboratory annotators, who have access to the whole vocalization.
-This is consistent with the fact that our lab definition of "Junk" was more stringent than the Zooniverse definition.
-Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use "Don't mark" if there was *any* noise or overlapping speech that would potentially affect acoustic analyses.
-
-Both Crying and Laughing showed moderate to good recall, but low precision.
-This means that a high proportion of segments detected as these two categories in the lab can be classified in the appropriate category by Zooniverse annotators, but Zooniverse annotators also put into these categories many segments that lab annotators classified in a different manner.
-Most saliently, there is a problematic level of confusion between Crying and Non-Canonical, whereby lab-detected Crying was almost as likely to be classified as such or Non-canonical in Zooniverse (the recall proportion was 49 and 47% respectively), and even worse, 55% of the segments Zooniverse annotators classified as Crying were actually Non-Canonical according to lab annotations.
-This is likely due to the fact that lab annotators had access to the whole segment, and could even listen to the context of that segment, so they could accurately interpret a segment as crying.
-Thus, it is possible that additional analyses, using for instance pitch or duration characteristics at the level of the whole segment rather than the chunk could improve classification performance, but for the time being, we cannot recommend Zooniverse annotations to identify crying bouts.
-
-As for Laughing, the recall is quite good but precision is low, with confusion involving "Junk" and Non-Canonical to similar extents.
-We are tentative in our interpretation of this result because the category was very sparsely represented, with barely 200 segments classified as Laughing in the whole data set.
-We thus recommend additional work, although we expect that perhaps there will be a similar problem (and solution) as indicated for Crying, whereby determination of Laughing versus Non-Canonical will require inspection of segment-level properties.
-
-Turning now to the child-level analyses, we were particularly interested in derived metrics because this is more typically the goal of such analyses, for instance to study potential individual or group differences, or investigate the impact of an intervention.
-Our results show that, despite agreement variation at the level of the segment, derived metrics were highly correlated across laboratory and Zooniverse data, meaning that variability in precision and recall did not substantially impact key outcome variables.
-Errors only seemed systematic for the Linguistic Proportion (and not for Canonical Proportion), and they were due to lower Linguistic Proportions from Zooniverse than laboratory data.
-This is due to the fact that a substantial number of Crying segments were classified as Non-Canonical by Zooniverse annotators.
-Nonetheless, since these errors were systematic across infants, they do not affect the study of group or individual variation.
-
-Our exploratory analyses then dug into some methodological aspects of our data.
-We found that, although individuals vary widely in terms of what proportion of their data is classified as Junk by Zooniverse annotators, this is not a sign that their data as a whole is compromised, with similar levels of correlations between laboratory- and Zooniverse-derived metrics across median groups based on proportion of their data classified as Junk.
-
-Next, we found that more data led to more reliable measurements: Specifically, our correlations based on 300-500 segments per child were markedly higher than those based on 100 segments per child, particularly when looking at our low-risk, younger participants.
-It is important to note that an advantage of Zooniverse, compared to laboratory annotations, is that it is easier to scale up: More data from each individual child can be annotated this way, which will help countering noise.
-
-In sum, we believe these data show that young children's vocalizations can be accurately described by relying on the judgments of Zooniverse's citizen scientists, particularly when the goal is to describe vocalizations at a broader level than individual chunks or segments.
-It is important to note that from a methodological point of view, the use of crowdsourcing is not merely a question of being an ecological technique, but it can also offer benefits in the area of scientific rigor and reproducibility.
-As previously mentioned in the Introduction section, Zooniverse offers both an automated system of data handling as well as an enormous pool of users: access to a large-scale group of blinded participants can in itself help researchers overcome the methodological compromises that sometimes lead to biased results.
-Additionally, crowd-sourcing and open-science frameworks can come together to encourage replication of existing work: open-source software, shared data and citizen science platforms would make it possible, and actually easy, to run identical studies and evaluate research results independently.
-
-### Further research directions
-
-These findings open up a new, exciting avenue for future research.
-Our pipeline to upload data on Zooniverse is scalable and quickly adapted to new datasets, including data on infants of different ages, learning different languages, or living in environments with different acoustic properties (e.g., rural versus urban environments).
-This means large existing datasets of daylong recordings, notably those hosted on Homebank [@vandam2016homebank], can be reliably annotated in an ecological way using citizen science.
-
-Notwithstanding our enthusiasm, we would like to point out some limitations of our results that should be addressed in future research.
-The first area in which more work would be welcome involves the pre-processing.
-We used LENA&reg; algorithms to detect vocalizations by the key child, but there are newer, better-performing, open-sourced alternatives that could be used instead [@lavechin2020open].
-Still, performance is most often established on children whose development is thought to be normative, being raised in small households located in urban environments, whereas it would be important to check for accuracy in more diverse samples, where the children have a diagnosis [e.g., @Dykstra2013], multiple siblings [e.g., @Elo], and/or are growing up in a rural environment [e.g., @cristia2020thorough].
-Speech technology developments would also be welcome to complement citizen scientists' classifications, particularly for categories that may be hard to classify based on short samples.
-There is a great deal of work to be done in this sense.
-The Introduction mentioned a challenge that established a baseline performance with unweighted average recall (UAR) of 54% (when recall is averaged across the 5 classes, Laughing, Crying, Canonical, Non-canonical, and Junk, giving equal weight to each of them rather than weighing them based on their frequency of occurrence).
-As reported in Table 1, our own method leads to a UAR of 57%, so one may think there is not much to gain with the automated method.
-However, notice that the automated method does lead to much better performance than Zooniverse classifications for Crying, and to some improvements in Canonical and Junk.
-For instance, the team who won the challenge in 2019 improved UAR by about 2%, primarily through gains in the laughing class obtained by adding training data [@yeh2019using].
-That state of the art was challenged by @kaya2020combining, who obtained a UAR of 61% on the same data as the challenge, thanks to improvements in all of the classes but for Laughing.
-Thus, there may be some gains to be obtained in specific classes.
-
-Our results warrant additional research on two metrics we used to describe children’s linguistic productions: Linguistic Proportion and Canonical Proportion.
-Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion.
-We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant to emotional disorders [see a review in @halpern2016excessive].
-Additionally, while Canonical Proportion has been studied in much previous work, we believe further research is necessary using larger scales (saliently increasing the sheer number of children studied, to have sufficient power to detect group effects) and with broader coverage [notably including children growing up in a broad range of languages and cultures, as in @cychosz2019canonical], to make sure that assumptions regarding cross-linguistic and maturational universality are amply justified and not merely a result of having low power to detect differences, or simply sampling from similar populations.
-
-Assuming these child-level metrics continue to be supported by such additional data, it would be interesting for future investigations to expand the analysis of Linguistic Proportion and Canonical Proportion to recover differences and similarities in language development between low-risk children and children with disorders other than Angelman syndrome.
-That is, it is at present unclear whether relatively lower Canonical Proportions will also be found in other populations, or whether this is more specific to Angelman syndrome.
-For instance, some previous work has found differences in babbling even when comparing late talkers with toddlers with no specific diagnosis  [@fasolo2008babbling].
-We thus encourage further research inspecting Canonical Proportion in other populations, as well as methodological work that attempts to generate metrics that are similarly easy to gather at scale.
-
-It is also important to note that work on children with non-normative development poses in itself some challenges.
-Our data contained a lower number of vocalizations for children diagnosed with Angelman syndrome, even though we selected much younger low-risk control children as the comparison group.
-However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
-Now that this method has been validated, researchers can extract all segments that LENA&reg; identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
-This approach may be particularly useful in the case of data from children with a diagnosis, whose data tends to be under-represented in LENA&reg; research, compared to data from normative children.
-We also hope future work makes use of the scalability of the method by employing larger sample sizes than those used in the present study, where data from only 10 children was used.
-
-Another open question we hope future work looks into is the assessment of within-participant changes.
-Our dataset only contained one daylong recording per child (with one exception), so other datasets in which a greater number of recordings are collected longitudinally can bring novel insights.
-Is the age-related decline we observed among children in the Angelman syndrome group truly a developmental decline that happens by necessity in this population, or might it reflect partially other aspects of children's experience and behavior?
-Additionally, longitudinal work within the same infant may be needed to track the "natural history" of vocalizations to better contextualize how these metrics should be interpreted.
-It is also an open question how large must effects be, at the individual level, in order to be detected with this method.
-For example, there is interest in using naturalistic recordings as outcome measures for clinical (behavioral or pharmacological) trials.
-The fact that agreement across annotators is high even in the data coming from children diagnosed with Angelman syndrome is encouraging, but it would be important to establish whether absolute levels of Canonical Proportion, and/or rates of change over longer time spans, may be good candidates for tracking the effects of any treatment.
-
-Finally, broader generalizations of our technique could be of interest to readers of this paper.
-Any researcher is able to undergo the procedure to create Zooniverse projects (https://www.zooniverse.org/about/faq).
-For instance, a project could be set up to annotate potentially disordered speech by older adults at risk of, or diagnosed with, neurodegenerative diseases.
-Although it would be important to similarly validate such annotations, we believe there is strong promise for such tasks, because many neurodegenerative disorders affect local aspects of the speech, which are detected at the syllable level.
-In contrast, tasks that require more speech or audio context may not be well suited to citizen science platforms, because they would require playing longer clips, which may reveal identifying or sensitive information.
-Although citizen scientists themselves are well-intentioned, the platform could be used by others who want to exploit the system for other means.
-
-## Conclusion
-
-In this study we validated the quality of annotations obtained through a citizen science platform, Zooniverse, as compared to a gold standard of human expert annotators.
-We analyzed the correspondence between annotations at the individual segment level, and the individual-child level, using Canonical Proportion and Linguistic Proportion as descriptors.
-We found moderate to good accuracy in the first, and strong positive correlations in the latter.
-We can conclude that citizen scientists are a reliable source of fast and ecological annotation of speech data, particularly when results are combined into child-level descriptors.
-The same methodology may be applied to several research questions in the study of language acquisition and language disorders.
-This finding is particularly welcome in an era when wearables open new avenues for studying human behavior and development in an ecological manner.
-
-\newpage
-
-# Acknowledgements
-
-We are grateful to the families who contributed their data; to the Zooniverse volunteers that made this work possible; and the audience and technical committee of SLT 2020 for helpful feedback on preliminary analyses.
-
-# References
-
-<!-- These lines ensure references are set with hanging indents in PDF documents; they are ignored in Word. -->
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-``` {r fig-zoo, comment=F, message=F, hide = T, warning = F, echo = F,  fig.cap ="Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needed to make; as well as the Field Guide, which remained accessible on the right of the screen throughout the session, and which allowed users to revise examples of the different categories."}
-knitr::include_graphics("zooniverse-pufig.pdf")
-```
-
-```{r fig-process, comment=F, message=F, hide = T, warning = F, echo = F, fig.cap = "Correspondence between LENA-identified segments (which can contain one or more vocalizations) and chunks.
-Segments and chunks are collectively referred to as clips because they are parts of the longer audio recording.
-Notice that clips are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority.
-To compare laboratory and Zooniverse annotations at the level of segments, chunk-level labels are combined using the hierarchy Canonical>Non-canonical>Crying>Laughing."}
-knitr::include_graphics("fig_levels.key.pdf")
-
-```
-
-```{r fig-corage, echo=F,fig.cap ="Correlations between child-level descriptors and age as a function of metric (Linguistic Proportion in the top row, Canonical Proportion in the bottom row), annotation method, and child group (red=low-risk, and black= Angelman syndrome)."}
-
-prettynames=c("Linguistic Proportion (Zoon.)","Canonical Proportion (Zoon.)",
-             "Linguistic Proportion (Lab)","Canonical Proportion (Lab)" )
-names(prettynames)<-c("z_lp","z_cp","l_lp","l_cp")
-mycols=c("black","red")
-names(mycols)<-c("Low-RiskControl","AngelmanSyndrome")
-mypch=c(4,20)
-names(mypch)<-c("Low-RiskControl","AngelmanSyndrome")
-
-lp_range=range(proportions[,c("z_lp","l_lp")])
-cp_range=range(proportions[,c("z_cp","l_cp")])
-
-#jpeg("../Results/corage.jpg",width=20,height=20,units="cm",res=300)
-layout(matrix(c(1:4), 2, 2, byrow = F))
-par(mar = c(4, 4, 1, 1))
-
-#create matrix to hold r results
-myr=matrix(NA,nrow=4,ncol=2)
-rownames(myr)<-c("z_lp","z_cp","l_lp","l_cp")
-colnames(myr)<-c("AS","LR")
-mylb=myhb=myn=myp =myr #this is an ugly way of creating 5 matrices
-
-for(thisvar in c("z_lp","z_cp","l_lp","l_cp")) {
-
-  if(thisvar %in% c("z_lp","l_lp")) myrange=lp_range else myrange=cp_range
-
-  plot(proportions[,thisvar]~proportions$Age, pch=mypch[proportions$Diagnosis],xlab="Age (months)",ylab=prettynames[thisvar],
-       col=mycols[proportions$Diagnosis],
-         ylim=myrange)
-  abline(lm(proportions[,thisvar]~proportions$Age,subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black")
-
-  #get cor properties
-  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"])
-    myr[thisvar,"AS"]=round(thiscor$estimate,3)
-    mylb[thisvar,"AS"]=round(thiscor$conf.int[1],3)
-    myhb[thisvar,"AS"]=round(thiscor$conf.int[2],3)
-    myn[thisvar,"AS"]=round(thiscor$parameter,3)
-    myp[thisvar,"AS"]=round(thiscor$p.value,3)
-
-    #add r to figure
-  text(mean(proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"]),mean(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar]),paste0("r=",myr[thisvar,"AS"]),col="black")
-
-  #repeat for LR  
-    abline(lm(proportions[,thisvar]~proportions$Age,subset=c(proportions$Diagnosis!="AngelmanSyndrome")),col="red")
-
-      #get cor properties
-  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"])
-    myr[thisvar,"LR"]=round(thiscor$estimate,3)
-    mylb[thisvar,"LR"]=round(thiscor$conf.int[1],3)
-    myhb[thisvar,"LR"]=round(thiscor$conf.int[2],3)
-    myn[thisvar,"LR"]=round(thiscor$parameter,3)
-    myp[thisvar,"LR"]=round(thiscor$p.value,3)
-
-    #add r to figure
-
-    text(mean(proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"]),mean(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar]),paste0("r=",myr[thisvar,"LR"]),col="red")
-}
-#dev.off()
-```
-
-```{r fig-prec-rec, echo=F,fig.cap ="Number of segments as a function of their majority label from laboratory annotations (rows) versus Zooniverse annotations (columns).
-Percentages (and shading) indicate precision (left) and recall (right).
-For precision, the percentage indicates what percentage of the segments that have a given majority label according to Zooniverse data are attributed a given label in Lab data (i.e., columns add up to 100%).
-For recall, the percentage indicates what percentage of the segments that have a given majority label according to Lab data are attributed a given label in Zooniverse data (i.e., rows add up to 100%)."}
-
+
+#compose precision
 colsums=colSums(conf_tab)
@@
 colnames(prop_cat)[3]<-"pr"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","pr")])
-prec_plot = ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
- geom_tile(aes(fill= rescale(pr)), colour = "white") +
-  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=4) +
-  geom_text(aes(label = Freq), vjust = 1,size=4) +
-  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
-     theme(legend.position = "none") +
-  xlab("Zooniverse") + ylab("Lab") +
-  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+data.frame(conf_tab)->stall_p
+stall_p$id=paste(stall_p$Lab,stall_p$Zooniverse)
+stall_p=merge(stall_p,prop_cat[c("id","pr")])
 
@@
 colnames(prop_cat)[3]<-"rec"
-data.frame(conf_tab)->stall
-stall$id=paste(stall$Lab,stall$Zooniverse)
-stall=merge(stall,prop_cat[c("id","rec")])
-rec_plot = ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
+data.frame(conf_tab)->stall_r
+stall_r$id=paste(stall_r$Lab,stall_r$Zooniverse)
+stall_r=merge(stall_r,prop_cat[c("id","rec")])
+```
+
+----- INSERT FIGURE 4 HERE -----
+
+These visualizations suggest that performance was moderate to good, which was confirmed via statistical analyses.
+The overall (weighted) accuracy was `r round(mycf$overall["Accuracy"]*100)`%, CI = [`r round(mycf$overall["AccuracyLower"]*100)`,`r round(mycf$overall["AccuracyUpper"]*100)`],  kappa is `r printr(mycf$overall["Kappa"])`, and the Gwet's AC1 coefficient is `r printr(mygac$est)`, CI = [`r printr(mygac$lb)`,`r printr(mygac$ub)`].
+
+### Child level descriptors
+
+```{r matr2}
+#create matrix to hold r results
+myr=matrix(NA,nrow=2,ncol=3)
+rownames(myr)<-c("lp","cp")
+colnames(myr)<-c("AS","LR","all")
+mylb=myhb=myn=myp =myr #this is an ugly way of creating 5 matrices
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_lp"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
+    myr[1,"AS"]=round(thiscor$estimate,3)
+    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
+    myn[1,"AS"]=round(thiscor$parameter,3)
+    myp[1,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"],proportions[proportions$Diagnosis!="AngelmanSyndrome","z_lp"])
+    myr[1,"LR"]=round(thiscor$estimate,3)
+    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
+    myn[1,"LR"]=round(thiscor$parameter,3)
+    myp[1,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(proportions$z_lp,proportions$l_lp)
+    myr[1,"all"]=round(thiscor$estimate,3)
+    mylb[1,"all"]=round(thiscor$conf.int[1],3)
+    myhb[1,"all"]=round(thiscor$conf.int[2],3)
+    myn[1,"all"]=round(thiscor$parameter,3)
+    myp[1,"all"]=round(thiscor$p.value,3)
+
+    # repeat whole for cp
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_cp"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
+    myr[2,"AS"]=round(thiscor$estimate,3)
+    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
+    myn[2,"AS"]=round(thiscor$parameter,3)
+    myp[2,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"],proportions[proportions$Diagnosis!="AngelmanSyndrome","z_cp"])
+    myr[2,"LR"]=round(thiscor$estimate,3)
+    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
+    myn[2,"LR"]=round(thiscor$parameter,3)
+    myp[2,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(proportions$z_cp,proportions$l_cp)
+    myr[2,"all"]=round(thiscor$estimate,3)
+    mylb[2,"all"]=round(thiscor$conf.int[1],3)
+    myhb[2,"all"]=round(thiscor$conf.int[2],3)
+    myn[2,"all"]=round(thiscor$parameter,3)
+    myp[2,"all"]=round(thiscor$p.value,3)
+
+mycortab=t(myr)
+rownames(mycortab)<-paste("All seg",rownames(mycortab))
+
+```
+
+Although the classification at the segment level was only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of children's individual development.
+We illustrate this association in Figure 5.
+Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r printr(myr["lp","all"])`, CI [`r printr(mylb["lp","all"])`,`r printr(myhb["lp","all"])`], *p* `r printp(myp["lp","all"])`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remained high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r printr(myr["lp","AS"])`, CI [`r printr(mylb["lp","AS"])`,`r printr(myhb["lp","AS"])`], *p* = `r printp(myp["lp","AS"])`]; low-risk control r(`r myn["lp","LR"]`) = `r printr(myr["lp","LR"])`, CI [`r printr(mylb["lp","LR"])`,`r printr(myhb["lp","LR"])`], *p* = `r printp(myp["lp","LR"])`].
+
+----- INSERT FIGURE 5 HERE -----
+
+Similarly, a strong positive correlation was found in the Canonical Proportion r(`r myn["cp","all"]`) = `r printr(myr["cp","all"])`, CI [`r printr(mylb["cp","all"])`,`r printr(myhb["cp","all"])`], *p* `r printp(myp["cp","all"])`].
+When we split by participant group, correlations remained high although we do note they were somewhat lower for the children with Angelman syndrome: r(`r myn["cp","AS"]`) = `r printr(myr["cp","AS"])`, CI [`r printr(mylb["cp","AS"])`,`r printr(myhb["cp","AS"])`], *p* = `r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r printr(myr["cp","LR"])`, CI [`r printr(mylb["cp","LR"])`,`r printr(myhb["cp","LR"])`], *p* `r printp(myp["cp","LR"])`.
+
+## Additional analyses
+
+In this section, we report on exploratory analyses, aimed at establishing conditions under which Zooniverse labels aligned with laboratory labels more or less closely.
+In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as "Junk" (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora [@cychosz2021canonical].
+A high proportion of "Junk" may indicate that automated segmentation was errorful for those children, and may be a sign that the rest of the data could be compromised as well.
+
+```{r prelim-Junk, echo=F}
+JunkXdiag=t.test(proportions$z_Junk ~ proportions$Diagnosis)
+
+#create matrix to hold r results
+myr=matrix(NA,nrow=2,ncol=2)
+rownames(myr)<-c("lp","cp")
+colnames(myr)<-c("lower_Junk","higher_Junk")
+mylb=myhb=myn=myp =myr #this is an ugly way of creating 5 matrices
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$median_Junk=="lower_Junk","z_lp"],proportions[proportions$median_Junk=="lower_Junk","l_lp"])
+    myr[1,"lower_Junk"]=round(thiscor$estimate,3)
+    mylb[1,"lower_Junk"]=round(thiscor$conf.int[1],3)
+    myhb[1,"lower_Junk"]=round(thiscor$conf.int[2],3)
+    myn[1,"lower_Junk"]=round(thiscor$parameter,3)
+    myp[1,"lower_Junk"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$median_Junk!="lower_Junk","l_lp"],proportions[proportions$median_Junk!="lower_Junk","z_lp"])
+    myr[1,"higher_Junk"]=round(thiscor$estimate,3)
+    mylb[1,"higher_Junk"]=round(thiscor$conf.int[1],3)
+    myhb[1,"higher_Junk"]=round(thiscor$conf.int[2],3)
+    myn[1,"higher_Junk"]=round(thiscor$parameter,3)
+    myp[1,"higher_Junk"]=round(thiscor$p.value,3)
+
+    # repeat whole for cp
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$median_Junk=="lower_Junk","z_cp"],proportions[proportions$median_Junk=="lower_Junk","l_cp"])
+    myr[2,"lower_Junk"]=round(thiscor$estimate,3)
+    mylb[2,"lower_Junk"]=round(thiscor$conf.int[1],3)
+    myhb[2,"lower_Junk"]=round(thiscor$conf.int[2],3)
+    myn[2,"lower_Junk"]=round(thiscor$parameter,3)
+    myp[2,"lower_Junk"]=round(thiscor$p.value,3)
+
+  #repeat for higher_Junk
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$median_Junk!="lower_Junk","l_cp"],proportions[proportions$median_Junk!="lower_Junk","z_cp"])
+    myr[2,"higher_Junk"]=round(thiscor$estimate,3)
+    mylb[2,"higher_Junk"]=round(thiscor$conf.int[1],3)
+    myhb[2,"higher_Junk"]=round(thiscor$conf.int[2],3)
+    myn[2,"higher_Junk"]=round(thiscor$parameter,3)
+    myp[2,"higher_Junk"]=round(thiscor$p.value,3)
+
+```
+
+We investigated this hypothesis by calculating the proportion of data labeled as "Junk" for each individual child.
+There was no significant difference in the proportions of their data labeled as "Junk" for children with Angelman syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"]*100,2)`%, SD = `r round(sd(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_Junk"])*100,2)`%) compared to the low-risk group (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"]*100,2)`%, SD = `r round(sd(proportions[proportions$Diagnosis!="AngelmanSyndrome","z_Junk"])*100,2)`%): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, *p* = `r printp(JunkXdiag$p.value)`.
+We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk".
+Results were similar across these two post-hoc subgroups (see Figure 6).
+For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["lp","lower_Junk"]`) = `r printr(myr["lp","lower_Junk"])`, CI [`r printr(mylb["lp","lower_Junk"])`,`r printr(myhb["lp","lower_Junk"])`], *p* = `r printp(myp["lp","lower_Junk"])`]; and for the higher Junk group it was r(`r myn["lp","higher_Junk"]`) = `r printr(myr["lp","higher_Junk"])`, CI [`r printr(mylb["lp","higher_Junk"])`,`r printr(myhb["lp","higher_Junk"])`], *p* `r printp(myp["lp","higher_Junk"])`.
+For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["cp","lower_Junk"]`) = `r printr(myr["cp","lower_Junk"])`, CI [`r printr(mylb["cp","lower_Junk"])`,`r printr(myhb["cp","lower_Junk"])`], *p* `r printp(myp["cp","lower_Junk"])`]; and for the higher Junk group it was r(`r myn["cp","higher_Junk"]`) = `r printr(myr["cp","higher_Junk"])`, CI [`r printr(mylb["cp","higher_Junk"])`,`r printr(myhb["cp","higher_Junk"])`], *p* `r printp(myp["cp","higher_Junk"])`.
+Thus, it does not seem that a higher proportion of "Junk" labels is an index of low quality data.
+
+----- INSERT FIGURE 6 HERE -----
+
+Above, we concluded that derived metrics integrating information across audio clips (Linguistic and Canonical Proportion, which can be derived from segment- or chunk-level labels) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse annotators) to segments (the original LENA&reg; segments presented to laboratory annotators), because one can calculate the proportions directly from the chunk-level judgment.
+As a result, there was one stage in our post-processing that may not have been necessary, whereby we collapsed labels across chunks associated to the same segment.
+We therefore repeated our analyses but instead of deriving our proportions for the Zooniverse data from the segment-level composite, we did it based on the individual chunk-level annotations.
+To facilitate comparison, correlations across laboratory- and Zooniverse-derived proportions are provided in Table 2.
+
+----- INSERT TABLE 2 HERE -----
+
+```{r bychunk-age, echo=F}
+read.csv("zoo_chunk_majority_agreement.csv")->chunk
+bychild=table(chunk$Answer,chunk$ChildID)
+cans=bychild["Canonical",]
+noncans=bychild["Non-Canonical",]
+Junk=bychild["Junk",]
+sum_bychild=table(chunk$ChildID)
+
+ling_prop =(cans+noncans)/(sum_bychild-Junk)
+can_prop =cans/(cans+noncans)
+
+#check that they match before merging
+#demo_data$ChildID==names(ling_prop)
+#demo_data$ChildID==names(can_prop)
+
+props_bychild_chunk=cbind(ling_prop,can_prop,demo_data)
+colnames(props_bychild_chunk)[1:2]<-c("chi","ling_prop")
+
+z_lp_as=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
+
+z_lp_lr=cor.test(props_bychild_chunk$ling_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
+
+z_cp_as=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis=="AngelmanSyndrome"])
+
+z_cp_lr=cor.test(props_bychild_chunk$can_prop[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"],props_bychild_chunk$Age[props_bychild_chunk$Diagnosis!="AngelmanSyndrome"])
+
+#matrix that will hold all r results
+corage=cbind(corage[,c("l_lp","z_lp")],
+             t(cbind(printr(z_lp_as$estimate),printr(z_lp_lr$estimate))),
+             corage[,c("l_cp","z_cp")],
+             t(cbind(printr(z_cp_as$estimate),printr(z_cp_lr$estimate)))
+             )
+corage=t(corage)
+rownames(corage)<-c("LP (lab)","LP (Zoon. Seg.)","LP (Zoon. Chunks)",
+                    "CP (lab)","CP (Zoon. Seg.)","CP (Zoon. Chunks)")
+```
+
+Regarding correlations with age, we found very similar results to the analyses based on laboratory annotations, as well as those based on Zooniverse annotations collapsed at the segment level (see Table 3).
+That is, the Linguistic Proportion was not strongly associated with age in the Angelman syndrome group: r(`r z_lp_as$parameter`) = `r printr(z_lp_as$estimate)`, CI [`r printr(z_lp_as$conf.int[1])`,`r printr(z_lp_as$conf.int[2])`], *p* = `r printp(z_lp_as$p.value)`], but increased with age  in the low-risk control group: r(`r z_lp_lr$parameter`) = `r printr(z_lp_lr$estimate)`, CI [`r printr(z_lp_lr$conf.int[1])`,`r printr(z_lp_lr$conf.int[2])`], *p* = `r printp(z_lp_lr$p.value)`.
+Likewise, the Canonical Proportion did not exhibit the same pattern across the groups, with a non-significant developmental decrease found among children with Angelman syndrome r(`r z_cp_as$parameter`) = `r printr(z_cp_as$estimate)`, CI [`r printr(z_cp_as$conf.int[1])`,`r printr(z_cp_as$conf.int[2])`], *p* = `r printp(z_cp_as$p.value)`]; and a marginal developmental increase among low-risk control r(`r z_cp_lr$parameter`) = `r printr(z_cp_lr$estimate)`, CI [`r printr(z_cp_lr$conf.int[1])`,`r printr(z_cp_lr$conf.int[2])`], *p* = `r printp(z_cp_lr$p.value)`.
+
+----- INSERT TABLE 3 HERE -----
+
+```{r zoo-lab-chunk, echo=F}
+
+#create matrix to hold r results
+myr=matrix(NA,nrow=2,ncol=3)
+rownames(myr)<-c("lp","cp")
+colnames(myr)<-c("AS","LR","all")
+mylb=myhb=myn=myp =myr #this is an ugly way of creating 5 matrices
+
+proportions=proportions[order(proportions$ChildID),]
+
+  #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_lp"])
+    myr[1,"AS"]=round(thiscor$estimate,3)
+    mylb[1,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[1,"AS"]=round(thiscor$conf.int[2],3)
+    myn[1,"AS"]=round(thiscor$parameter,3)
+    myp[1,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","ling_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_lp"])
+    myr[1,"LR"]=round(thiscor$estimate,3)
+    mylb[1,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[1,"LR"]=round(thiscor$conf.int[2],3)
+    myn[1,"LR"]=round(thiscor$parameter,3)
+    myp[1,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(props_bychild_chunk$ling_prop,proportions$l_lp)
+    myr[1,"all"]=round(thiscor$estimate,3)
+    mylb[1,"all"]=round(thiscor$conf.int[1],3)
+    myhb[1,"all"]=round(thiscor$conf.int[2],3)
+    myn[1,"all"]=round(thiscor$parameter,3)
+    myp[1,"all"]=round(thiscor$p.value,3)
+
+    # repeat whole for cp
+
+  #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis=="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis=="AngelmanSyndrome","l_cp"])
+    myr[2,"AS"]=round(thiscor$estimate,3)
+    mylb[2,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[2,"AS"]=round(thiscor$conf.int[2],3)
+    myn[2,"AS"]=round(thiscor$parameter,3)
+    myp[2,"AS"]=round(thiscor$p.value,3)
+
+  #repeat for LR
+
+      #get cor properties
+  thiscor=cor.test(props_bychild_chunk[props_bychild_chunk$Diagnosis!="AngelmanSyndrome","can_prop"],proportions[proportions$Diagnosis!="AngelmanSyndrome","l_cp"])
+    myr[2,"LR"]=round(thiscor$estimate,3)
+    mylb[2,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[2,"LR"]=round(thiscor$conf.int[2],3)
+    myn[2,"LR"]=round(thiscor$parameter,3)
+    myp[2,"LR"]=round(thiscor$p.value,3)
+
+#repeat for all cor
+
+      thiscor=cor.test(props_bychild_chunk$can_prop,proportions$l_cp)
+    myr[2,"all"]=round(thiscor$estimate,3)
+    mylb[2,"all"]=round(thiscor$conf.int[1],3)
+    myhb[2,"all"]=round(thiscor$conf.int[2],3)
+    myn[2,"all"]=round(thiscor$parameter,3)
+    myp[2,"all"]=round(thiscor$p.value,3)
+
+ mycortab=rbind(mycortab,t(myr))
+rownames(mycortab)[4:6]<-paste("Chunks",rownames(t(myr)))
+
+```
+
+As for correlations between Zooniverse- and laboratory-derived metrics (e.g., the correlation between the Linguistic Proportion as estimated using Zooniverse coding at the chunk level and the Linguistic Proportion via laboratory coding at the segment level), we observe very similar levels of correlation to the segment-based analyses (see Table 2): Linguistic Proportion overall r(`r myn["lp","all"]`) = `r printr(myr["lp","all"])`, CI [`r printr(mylb["lp","all"])`,`r printr(myhb["lp","all"])`], *p* `r printp(myp["lp","all"])`] between Linguistic Proportion by child from the Zooniverse at the chunk level and the lab annotators' data at the segment level.
+When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r printr(myr["lp","AS"])`, CI [`r printr(mylb["lp","AS"])`,`r printr(myhb["lp","AS"])`], *p* = `r printp(myp["lp","AS"])`; low-risk control r(`r myn["lp","LR"]`) = `r printr(myr["lp","LR"])`, CI [`r printr(mylb["lp","LR"])`,`r printr(myhb["lp","LR"])`], *p* = `r printp(myp["lp","LR"])`.
+As for Canonical Proportion overall r(`r myn["cp","all"]`) = `r printr(myr["cp","all"])`, CI [`r printr(mylb["cp","all"])`,`r printr(myhb["cp","all"])`], *p* `r printp(myp["cp","all"])`; Angelman syndrome: r(`r myn["cp","AS"]`) = `r printr(myr["cp","AS"])`, CI [`r printr(mylb["cp","AS"])`,`r printr(myhb["cp","AS"])`], *p* = `r printp(myp["cp","AS"])`;  low-risk control infants r(`r myn["cp","LR"]`) = `r printr(myr["cp","LR"])`, CI [`r printr(mylb["cp","LR"])`,`r printr(myhb["cp","LR"])`], *p* `r printp(myp["cp","LR"])`.
+
+Since all of these results are very similar to those obtained when collapsing across chunk-level labels to generate segment-level labels, we conclude that in the future this step can be skipped.
+Instead, researchers can derive Linguistic and Canonical Proportions directly from citizen scientists' chunk-level labels [which was in fact what @cychosz2021canonical did].
+
+```{r downsample}
+
+#we start from the data base that has all the info
+nsegs=data.frame(table(data_all$ChildID))
+colnames(nsegs)<-c("ChildID","nsegs")
+
+merge(nsegs,demo_data)->nsegs
+nseg_t=t.test(nsegs$nsegs~nsegs$Diagnosis)
+
+resampling=matrix(NA,nrow=50,ncol=6)
+colnames(resampling)<-c("lp_r_all","lp_r_as","lp_r_lr","cp_r_all","cp_r_as","cp_r_lr")
+for(i in 1:50){
+  sel_seg=NULL
+  for(thischi in levels(data_all$ChildID)) sel_seg<-c(sel_seg,as.character(data_all$segmentId_DB[sample(data_all$segmentId_DB[data_all$ChildID==thischi],100)]))
+
+  sampling=data_all[as.character(data_all$segmentId_DB) %in% as.character(sel_seg),]
+
+  #get the ns by child, then calculate the linguistic proportion & canonical proportion, separately for zooniverse & lab
+  ztab=table(sampling$ChildID,sampling$Zoon_classif)
+  z_lp =rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
+  z_cp =ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
+
+  ltab=table(sampling$ChildID,sampling$lab)
+  l_lp =rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
+  l_cp =ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
+
+  thistab=cbind(demo_data,z_lp,z_cp,l_lp,l_cp)
+
+  resampling[i,"lp_r_all"]=cor.test(thistab$z_lp,thistab$l_lp)$estimate
+  resampling[i,"lp_r_as"]=cor.test(thistab$z_lp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
+  resampling[i,"lp_r_lr"]=cor.test(thistab$z_lp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_lp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate
+
+   resampling[i,"cp_r_all"]=cor.test(thistab$z_cp,thistab$l_cp)$estimate
+  resampling[i,"cp_r_as"]=cor.test(thistab$z_cp[thistab$Diagnosis=="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis=="AngelmanSyndrome"])$estimate
+  resampling[i,"cp_r_lr"]=cor.test(thistab$z_cp[thistab$Diagnosis!="AngelmanSyndrome"],thistab$l_cp[thistab$Diagnosis!="AngelmanSyndrome"])$estimate
+ }
+myr=colMeans(resampling)
+mycortab=rbind(mycortab,cbind(myr[c("lp_r_as" ,"lp_r_lr","lp_r_all")],myr[c("cp_r_as" ,"cp_r_lr","cp_r_all")]))
+rownames(mycortab)[7:9]<-c("100 seg AS","100 seg LR","100 seg all")
+for(i in 1:dim(mycortab)[1]) for(j in 1:dim(mycortab)[2]) mycortab[i,j]<-printr(as.numeric(as.character(mycortab[i,j])))
+
+resampling=data.frame(resampling)
+```
+
+Next, we looked at whether having more segments from each child may lead to more reliable metrics.
+We observed a non-significant trend [Welch t(`r nseg_t$parameter`)=`r nseg_t$statistic`] for lower number of segments in the Angelman syndrome group (M = `r nseg_t$estimate[1]`, range `r min(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis=="AngelmanSyndrome"])`) than among low-risk control infants (M = `r nseg_t$estimate[2]`, range `r min(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`-`r max(nsegs$nsegs[nsegs$Diagnosis!="AngelmanSyndrome"])`), likely because of the lower volubility among the former group of children.
+Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications [@cychosz2021canonical].
+We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data were extracted from each child by under-sampling.
+Since @cychosz2021canonical extracted 100 segments from each child in their study, we randomly sampled 100 segments from each of our children's data, and recalculated the Linguistic and Canonical Proportions based only on these 100 labels.
+We repeated this process 50 times, to assess the extent to which the association between laboratory- and Zooniverse-derived metrics varied in each random sample (e.g., the correlation between Linguistic Proportion as estimated using Zooniverse coding on the one hand, and Linguistic Proportion as estimated using laboratory coding on the other).
+
+The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table 2, "100 seg" rows), particularly for the low-risk younger infants, and for Linguistic Proportion.
+This may indicate that, especially in some groups of infants, 100 segments may not be sufficient to capture a stable estimate of the child's Linguistic and Canonical Proportions.
+We also observed quite a bit of variance across the 50 runs.
+For Linguistic Proportion, the range of correlations found for all infants was [`r printr(min(resampling$lp_r_all))`,`r printr(max(resampling$lp_r_all))`]; for the Angelman syndrome group [`r printr(min(resampling$lp_r_as))`,`r printr(max(resampling$lp_r_as))`]; for the low-risk group [`r printr(min(resampling$lp_r_lr))`,`r printr(max(resampling$lp_r_lr))`].
+For Canonical Proportion, the range of correlations found for all infants was [`r printr(min(resampling$cp_r_all))`,`r printr(max(resampling$cp_r_all))`]; for the Angelman syndrome group [`r printr(min(resampling$cp_r_as))`,`r printr(max(resampling$cp_r_as))`]; for the low-risk group [`r printr(min(resampling$cp_r_lr))`,`r printr(max(resampling$cp_r_lr))`].
+
+## Discussion
+
+In the present study we assessed the extent to which child vocalizations from LENA&reg; daylong recordings can be accurately described based on classifications collected using the citizen science platform Zooniverse.
+Our recordings came from children diagnosed with a genetic syndrome associated with severe language disorder, as well as low-risk controls, and included a range of different ages.
+Excerpts of these recordings had previously been annotated by trained experts in the lab, who provided the gold standard.
+
+In descriptive terms, we found that category frequency was similar across laboratory and citizen science annotations.
+Correlations between age and our two child-level metrics revealed different developmental patterns between our sample of older children with a diagnosis of Angelman syndrome and younger low-risk controls.
+Given that the two groups of children differ on both of these features (age and diagnosis), we cannot empirically tease apart these two differences.
+Nonetheless, inspection of Figure 3 suggests different interpretations for the two derived metrics.
+Overall, the pattern for Linguistic Proportion is compatible with a non-linear trajectory common to the two groups, whereby the Linguistic Proportion increases rapidly up to 20 months, and hovers around 90-100% thereafter.
+In this case, the lack of correlation with age in the Angelman syndrome group suggests that there is little systematic variance in this metric after 20 months of age.
+In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and relatively stable levels of Canonical Proportion among the children in the Angelman syndrome group [in line with previous work; @hamrick2019capturing;@hamrick2019measurement].
+
+Next, we looked at agreement between laboratory and Zooniverse annotation at the level of individual segments.
+We found that some categories had high precision and recall -- notably Non-canonical segments were detected quite accurately, and to a lesser extent the same was true for Canonical segments.
+This is interesting because it suggests that Zooniverse annotations may be trusted particularly for detecting these two types of speech-like segments.
+
+In other cases, precision was fairly high but recall was lower.
+This was the case of the "Junk/Don't mark" category.
+The combination of high precision and low recall indicates that when Zooniverse annotations return a "Junk" judgment, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as "Don't mark" in the laboratory.
+Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the "Junk" category less) than laboratory annotators, who have access to the whole segment (and its context).
+This is consistent with the fact that our lab definition of "Junk" was more stringent than the Zooniverse definition.
+Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use "Don't mark" if there was *any* noise or overlapping speech that could affect acoustic analyses.
+
+Both Crying and Laughing showed moderate to good recall, but low precision.
+This means that a high proportion of segments detected as these two categories in the lab can be classified in the appropriate category by Zooniverse annotators, but Zooniverse annotators also put into these categories many segments that lab annotators classified in a different manner.
+Most saliently, there is a problematic level of confusion between Crying and Non-Canonical, whereby lab-detected Crying was almost as likely to be classified as such or Non-canonical in Zooniverse (the recall was `r round(stall_r[stall_r$Lab=="Crying" & stall_r$Zooniverse=="Crying","rec"])` and `r round(stall_r[stall_r$Lab=="Crying" & stall_r$Zooniverse=="Non-Canonical","rec"])`% respectively), and even worse, 55% of the segments Zooniverse annotators classified as Crying were actually Non-Canonical according to lab annotations.
+This is probably due to the fact that lab annotators had access to the whole segment, and could even listen to the context of that segment, so they could accurately interpret a segment as crying.
+Thus, it is possible that additional analyses, using for instance pitch or duration characteristics at the level of the whole segment rather than the chunk could improve classification performance, but for the time being, we cannot recommend Zooniverse annotations to identify crying bouts.
+
+As for Laughing, the recall is quite good but precision is low, with confusion involving "Junk" and Non-Canonical to similar extents.
+We are tentative in our interpretation of this result because the category was very sparsely represented, with barely `r table(data_all$lab)["Laughing"]` segments classified as Laughing in the whole data set.
+We thus recommend additional work, although we expect that perhaps there will be a similar problem (and solution) as indicated for Crying, whereby determination of Laughing versus Non-Canonical will require inspection of segment-level properties.
+
+Turning now to the child-level analyses, we were particularly interested in derived metrics because this is more typically the goal of such analyses, for instance to study potential individual or group differences, or to investigate the impact of an intervention on a child or group of children.
+Our results show that, despite only moderate levels of agreement at the level of the segment, our two derived metrics (Linguistic and Canonical Proportions) were highly correlated across laboratory and Zooniverse data, meaning that variability in precision and recall did not substantially impact key outcome variables.
+Errors only seemed systematic for the Linguistic Proportion (and not for Canonical Proportion), and they were due to lower Linguistic Proportions from Zooniverse than laboratory data.
+This is due to the fact that a substantial number of Crying segments were classified as Non-Canonical by Zooniverse annotators.
+Nonetheless, since these errors were systematic across infants, they do not affect the study of group or individual variation.
+
+Our exploratory analyses then dug into some methodological aspects of our data.
+We found that, although individuals vary widely in terms of what proportion of their data is classified as Junk by Zooniverse annotators, this is not a sign that their data as a whole is compromised, with similar levels of correlations between laboratory- and Zooniverse-derived metrics across median groups based on proportion of their data classified as Junk.
+
+Next, we found that more data led to more reliable measurements: Specifically, our correlations based on `r min(nsegs$nsegs)`-`r max(nsegs$nsegs)` segments per child were markedly higher than those based on 100 segments per child, particularly when looking at our low-risk, younger participants.
+It is important to note that an advantage of Zooniverse, compared to laboratory annotations, is that it is easier to scale up: More data from each individual child can be annotated this way, which helps countering noise.
+
+In sum, we believe these data show that young children's vocalizations can be accurately described by relying on the judgments of Zooniverse's citizen scientists, particularly when the goal is to describe vocalizations at a broader level than individual chunks or segments.
+It is important to note that from a methodological point of view, the use of crowdsourcing is not merely a question of being an ecological technique, but it can also offer benefits in the area of scientific rigor and reproducibility.
+As previously mentioned in the Introduction section, Zooniverse offers both an automated system of data handling as well as an enormous pool of users: access to a large-scale group of blinded annotators can in itself help researchers overcome the methodological compromises that sometimes lead to biased results.
+Additionally, crowd-sourcing and open-science frameworks can come together to encourage replication of existing work: open-source software, shared data and citizen science platforms would make it possible, and actually easy, to run identical studies and evaluate research results independently.
+
+### Further research directions
+
+These findings open up a new, exciting avenue for future research.
+Our pipeline to upload data on Zooniverse is scalable and quickly adapted to new datasets, including data on infants of different ages, learning different languages, or living in environments with different acoustic properties (e.g., rural versus urban environments).
+This means large existing datasets of daylong recordings, notably those hosted on Homebank [@vandam2016homebank], can be reliably annotated in an ecological way using citizen science.
+
+Notwithstanding our enthusiasm, we would like to point out some limitations of our results that should be addressed in future research.
+The first area in which more work would be welcome involves the pre-processing.
+We used LENA&reg; algorithms to detect vocalizations by the key child, but there are newer, better-performing, open-sourced alternatives that could be used instead [@lavechin2020open].
+Still, performance is most often established on children whose development is thought to be normative, being raised in small households located in urban environments, whereas it would be important to check for accuracy in more diverse samples, where the children have a diagnosis [e.g., @Dykstra2013], multiple siblings [e.g., @Elo], and/or are growing up in a rural environment [e.g., @cristia2020thorough].
+Speech technology developments would also be welcome to complement citizen scientists' classifications, particularly for categories that may be hard to classify based on short samples.
+There is a great deal of work to be done in this sense.
+The Introduction mentioned a challenge that established a baseline performance with unweighted average recall (UAR) of 54% (when recall is averaged across the 5 classes, Laughing, Crying, Canonical, Non-canonical, and Junk, giving equal weight to each of them rather than weighing them based on their frequency of occurrence).
+As reported in Table 1, our own method leads to a UAR of `r round(mean(stall_r$rec[stall_r$Lab==stall_r$Zooniverse]))`%, so one may think there is not much to gain with the automated method.
+However, notice that the automated method does lead to much better performance than Zooniverse classifications for Crying, and to some improvements in Canonical and Junk.
+For instance, the team who won the challenge in 2019 improved UAR by about 2%, primarily through gains in the laughing class obtained by adding training data [@yeh2019using].
+That state of the art was challenged by @kaya2020combining, who obtained a UAR of 61% on the same data as the challenge, thanks to improvements in all of the classes but for Laughing.
+Even more recently, @yao2020classification reported on a series of classifiers focused on crying, with astounding improvements in this class.
+Thus, there may be some gains to be obtained in specific vocalization types, particularly those that are rare (like Laughing) and those which require context (Crying).
+
+Our results warrant additional research on two metrics we used to describe children’s linguistic productions: Linguistic Proportion and Canonical Proportion.
+Linguistic Proportion has not been investigated extensively in previous research, and correlations with age were weaker than those for Canonical Proportion.
+We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant to emotional disorders [see a review in @halpern2016excessive].
+Additionally, while metrics conceptually related to our Canonical Proportion have been studied in much previous work, we believe further research is necessary using larger scales (saliently increasing the sheer number of children studied, to have sufficient power to detect group effects) and with broader coverage [notably including children growing up in a broad range of languages and cultures, as in @cychosz2021canonical], to make sure that assumptions regarding cross-linguistic and maturational universality are amply justified and not merely a result of having low power to detect differences, or simply sampling from similar populations.
+
+Assuming these child-level metrics continue to be supported by such additional data, it would be interesting for future investigations to expand the analysis of Linguistic Proportion and Canonical Proportion to recover differences and similarities in language development between low-risk children and children with disorders other than Angelman syndrome.
+That is, it is at present unclear whether relatively lower Canonical Proportions are also be found in other populations, or whether this is more specific to Angelman syndrome.
+For instance, some previous work has found differences in babbling even when comparing late talkers with toddlers with no specific diagnosis  [@fasolo2008babbling].
+We thus encourage further research inspecting Canonical Proportion in other populations, as well as methodological work that attempts to generate metrics that are similarly easy to gather at scale.
+
+We also acknowledge that work on children with non-normative development poses in itself some challenges.
+Our data contained a lower number of vocalizations for children diagnosed with Angelman syndrome, even though we selected much younger low-risk control children as the comparison group.
+However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
+Now that this method has been validated, researchers can extract all segments that LENA&reg; identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
+This approach may be particularly useful in the case of data from children with a diagnosis, whose data may be under-represented in LENA&reg; research, compared to data from normative children.
+We also hope future work makes use of the scalability of the method by employing larger sample sizes than those used in the present study, where data were drawn from only 10 children in each group.
+
+Another open question we hope future work looks into is the assessment of within-participant changes.
+Our dataset only contained one daylong recording per child (with one exception), so other datasets in which a greater number of recordings are collected longitudinally can bring novel insights.
+Is the age-related decline we observed among children in the Angelman syndrome group truly a developmental decline that happens by necessity in this population, or might it reflect partially other aspects of children's experience and behavior?
+Additionally, longitudinal work within the same infant may be needed to track the "natural history" of vocalizations to better contextualize how these metrics should be interpreted.
+It is also an open question how large  effects must be, at the individual level, in order to be detected with this method.
+For example, there is interest in using naturalistic recordings as outcome measures for clinical (behavioral or pharmacological) trials.
+The fact that agreement across annotators is high even in the data coming from children diagnosed with Angelman syndrome is encouraging, but it would be important to establish whether absolute levels of Canonical Proportion, and/or rates of change over longer time spans, may be good candidates for tracking the effects of a given treatment.
+
+Finally, broader generalizations of our technique could be of interest to readers of this paper.
+Any researcher is able to undergo the procedure to create Zooniverse projects (https://www.zooniverse.org/about/faq).
+For instance, a project could be set up to annotate potentially disordered speech by older adults at risk of, or diagnosed with, neurodegenerative diseases.
+Although it would be important to similarly validate such annotations, we believe there is strong promise for such tasks, because many neurodegenerative disorders affect local aspects of the speech, which are detected at the syllable level.
+In contrast, tasks that require more speech or audio context may not be well suited to citizen science platforms, because they would require playing longer clips, which may reveal identifying or sensitive information.
+Although citizen scientists themselves are well-intentioned, the platform could be used by others who want to exploit the system for nefarious goals.
+
+## Conclusion
+
+In this study, we validated the quality of annotations obtained through a citizen science platform, Zooniverse, as compared to a gold standard of human expert annotators.
+We analyzed the correspondence between annotations at the individual segment level, and the individual child level, using Canonical Proportion and Linguistic Proportion as descriptors.
+We found moderate to good accuracy in the former, and strong positive correlations in the latter.
+We can conclude that citizen scientists are a reliable source of fast and ecological annotation of speech data, particularly when results are combined into child-level descriptors.
+The same methodology may be applied to several research questions in the study of language acquisition and language disorders.
+This finding is particularly welcome in an era when wearables open new avenues for studying human behavior and development in an ecological manner.
+
+\newpage
+
+# Acknowledgements
+
+We are grateful to the families who contributed their data; to the Zooniverse volunteers that made this work possible; and the audience and technical committee of SLT 2020 for helpful feedback on preliminary analyses.
+
+``` {r fig-zoo, comment=F, message=F, hide = T, warning = F, echo = F,  fig.cap ="Screen captures of the 6 stages of the tutorial, three of which are shown here: The initial explanation of the trial, examples of canonical vocalizations, and the classification choice the user needed to make; as well as the Field Guide, which remained accessible on the right of the screen throughout the session, and which allowed users to revise examples of the different categories."}
+knitr::include_graphics("zooniverse-pufig.pdf")
+```
+
+```{r fig-process, comment=F, message=F, hide = T, warning = F, echo = F, fig.cap = "Correspondence between LENA-identified segments (which can contain one or more vocalizations) and chunks.
+Segments and chunks are collectively referred to as clips because they are parts of the longer audio recording.
+Notice that clips are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority.
+To compare laboratory and Zooniverse annotations at the level of segments, chunk-level labels are combined using the hierarchy Canonical > Non-canonical > Crying > Laughing > Junk."}
+knitr::include_graphics("fig_levels.key.pdf")
+
+```
+
+```{r fig-corage, echo=F,fig.cap ="Correlations between child-level descriptors and age as a function of metric (Linguistic Proportion in the top row, Canonical Proportion in the bottom row), annotation method, and child group (red circles = low-risk, and black crosses = Angelman syndrome)."}
+
+prettynames=c("Linguistic Proportion (Zoon.)","Canonical Proportion (Zoon.)",
+             "Linguistic Proportion (Lab)","Canonical Proportion (Lab)" )
+names(prettynames)<-c("z_lp","z_cp","l_lp","l_cp")
+mycols=c("black","red")
+names(mycols)<-c("Low-RiskControl","AngelmanSyndrome")
+mypch=c(4,20)
+names(mypch)<-c("Low-RiskControl","AngelmanSyndrome")
+
+lp_range=range(proportions[,c("z_lp","l_lp")])
+cp_range=range(proportions[,c("z_cp","l_cp")])
+
+#jpeg("../Results/corage.jpg",width=20,height=20,units="cm",res=300)
+layout(matrix(c(1:4), 2, 2, byrow = F))
+par(mar = c(4, 4, 1, 1))
+
+#create matrix to hold r results
+myr=matrix(NA,nrow=4,ncol=2)
+rownames(myr)<-c("z_lp","z_cp","l_lp","l_cp")
+colnames(myr)<-c("AS","LR")
+mylb=myhb=myn=myp =myr #this is an ugly way of creating 5 matrices
+
+for(thisvar in c("z_lp","z_cp","l_lp","l_cp")) {
+
+  if(thisvar %in% c("z_lp","l_lp")) myrange=lp_range else myrange=cp_range
+
+  plot(proportions[,thisvar]~proportions$Age, pch=mypch[proportions$Diagnosis],xlab="Age (months)",ylab=prettynames[thisvar],
+       col=mycols[proportions$Diagnosis],
+         ylim=myrange)
+  abline(lm(proportions[,thisvar]~proportions$Age,subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black")
+
+  #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"])
+    myr[thisvar,"AS"]=round(thiscor$estimate,3)
+    mylb[thisvar,"AS"]=round(thiscor$conf.int[1],3)
+    myhb[thisvar,"AS"]=round(thiscor$conf.int[2],3)
+    myn[thisvar,"AS"]=round(thiscor$parameter,3)
+    myp[thisvar,"AS"]=round(thiscor$p.value,3)
+
+    #add r to figure
+  text(mean(proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"]),mean(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar]),paste0("r=",myr[thisvar,"AS"]),col="black")
+
+  #repeat for LR  
+    abline(lm(proportions[,thisvar]~proportions$Age,subset=c(proportions$Diagnosis!="AngelmanSyndrome")),col="red")
+
+      #get cor properties
+  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"])
+    myr[thisvar,"LR"]=round(thiscor$estimate,3)
+    mylb[thisvar,"LR"]=round(thiscor$conf.int[1],3)
+    myhb[thisvar,"LR"]=round(thiscor$conf.int[2],3)
+    myn[thisvar,"LR"]=round(thiscor$parameter,3)
+    myp[thisvar,"LR"]=round(thiscor$p.value,3)
+
+    #add r to figure
+
+    text(mean(proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"]),mean(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar]),paste0("r=",myr[thisvar,"LR"]),col="red")
+}
+#dev.off()
+```
+
+```{r fig-prec-rec, echo=F,fig.cap ="Number of segments as a function of their majority label from laboratory annotations (rows) versus Zooniverse annotations (columns).
+Percentages (and shading) indicate precision (left) and recall (right).
+For precision, the percentage indicates what percentage of the segments that have a given majority label according to Zooniverse data are attributed a given label in Lab data (i.e., columns add up to 100%).
+For recall, the percentage indicates what percentage of the segments that have a given majority label according to Lab data are attributed a given label in Zooniverse data (i.e., rows add up to 100%)."}
+
+stall_p$Lab=factor(stall_p$Lab,levels=label_options)
+stall_p$Zooniverse=factor(stall_p$Zooniverse,levels=label_options)
+
+prec_plot = ggplot(data = stall_p, mapping = aes(y = Lab, x=Zooniverse)) +
+ geom_tile(aes(fill= rescale(pr)), colour = "white") +
+  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=4) +
+  geom_text(aes(label = Freq), vjust = 1,size=4) +
+  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
+     theme(legend.position = "none") +
+  xlab("Zooniverse") + ylab("Lab") +
+  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+rec_plot = ggplot(data = stall_r, mapping = aes(y = Lab, x=Zooniverse)) +
  geom_tile(aes(fill= rescale(rec)), colour = "white") +
@@
 
-```{r fig-corlab-zoo, echo=F,fig.cap ="Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of child group (red=low-risk, and black=Angelman syndrome)."}
+```{r fig-corlab-zoo, echo=F,fig.cap ="Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of child group (red circles = low-risk, and black crosses = Angelman syndrome)."}
 
@@
 
-mycortab=t(myr)
-rownames(mycortab)<-paste("All seg",rownames(mycortab))
-```
-
-```{r fig-cor-Junk, echo=F,fig.cap ="Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of whether children's Junk proportion was higher or lower than the median for all children (blue=lower, and green=higher)."}
-
-mycols=c("blue","darkgreen")
+```
+
+```{r fig-cor-Junk, echo=F,fig.cap ="Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of whether children's Junk proportion was higher or lower than the median for all children (light blue wheels = lower, and green diamonds = higher)."}
+
+mycols=c("lightblue","darkgreen")
 mypch=c(10,18)
@@
 
-  abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$median_Junk=="lower_Junk")),col="blue")
+  abline(lm(proportions[,"z_lp"]~proportions[,"l_lp"],subset=c(proportions$median_Junk=="lower_Junk")),col="lightblue")
 
@@
 
-  abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$median_Junk=="lower_Junk")),col="blue")
+  abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$median_Junk=="lower_Junk")),col="lightblue")
 
@@
 ```{r tab-LENA-challenge, comment = F, message = F, hide = T, warning = F, echo = F, results = "asis"}
-mycaption="Comparison of recall percentages obtained with the baseline algorithm created by the ComParE team, in the context of the ComParE 2019 BabySounds subchallenge (Schuller et al., 2019), LENATM labels (note LENATM does not distinguish between Canonical and Non-canonical), and those obtained in this study through Zooniverse annotations.
-Label frequency indicates the prevalence of the relevant label (Crying, Laughing, etc.) in each dataset (the same Lab dataset was used for both LENATM and Zooniverse).
+mycaption="Comparison of recall percentages obtained with the baseline algorithm created by the ComParE team, in the context of the ComParE 2019 BabySounds subchallenge (Schuller et al., 2019; C2019B), LENA labels (note LENA did not distinguish between Canonical and Non-canonical in this analysis), and those obtained in this study through Zooniverse annotations.
+Label frequency indicates the prevalence of the relevant label (Crying, Laughing, etc.) in each dataset (the same Lab dataset was used for both LENA and Zooniverse).
 UAR stands for unweighted average recall, WAR for weighted average recall (which takes into account label frequency)."
 
-read.table("tab-lena-challenge.txt",header=F,sep ="\t",skip =1)->tab1
+read.table("tab-lena-challenge.txt",header=T,sep ="\t")->tab1
+
+#add zoo & lab data
+recalls=round(stall_r$rec[stall_r$Lab==stall_r$Zooniverse],1)
+names(recalls)<-c("Canonical","Crying","Junk","Laughing","Non-can.")
+recalls=recalls[as.character(tab1[2:6,"X"])]
+freqs=table(data_all$Zoon)
+names(freqs)<-c("Canonical","Crying","Junk","Laughing","Non-can.")
+freqs=freqs[as.character(tab1[2:6,"X"])]
+w=freqs/sum(freqs)
+UAR=round(mean(recalls),1)
+WAR=round(weighted.mean(recalls,w),1)
+tab1$z=c("Zoon.",recalls, UAR,WAR)
+tab1$f=c("Zoon.",freqs, "","")
+tab1=tab1[,c("X","Recall","x","z","Label.frequency","f")]
+write.csv(tab1,"tab1-final.csv",row.names=F)
 #olnames(tab1)<-c("","Recall","Recall","Recall","Label Frequency","Label Frequency")
@@
 
-kable(tab1,cap =mycaption, col.names = NULL) %>%
-  collapse_rows(columns = 1:6, valign = "middle") #%>%
+kable(tab1,cap =mycaption, col.names = NULL) 
+#%>%
+#  collapse_rows(columns = 1:6, valign = "middle") #%>%
 #add_header_above(c(" " = 1, "Recall"= 3, "Label Frequency" = 2))
@@
 ```
+
+```{r tab-corage, results="as.is"}
+
+kable(corage,cap ="Pearson correlation coefficients between derived metrics (Linguistic Proportion = LP; Canonical Proportion = CP) and age, for the Angelman syndrome (AS) group data and low-risk (LR) group data.
+Metrics were derived from laboratory annotations at the segment level (lab); from Zooniverse annotations collapsed at the segment level (Zoon.
+Seg); or directly from the chunk-level labels (Zoon.
+Chunks).
+")
+write.csv(corage,"mycorage.csv")
+```
+
+# References
+
+<!-- These lines ensure references are set with hanging indents in PDF documents; they are ignored in Word. -->
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}

### 2021-02-18 202010_jslhr/paper.tex
ready for final proofing

--- 2021-01-07 202010_jslhr/paper.tex
@@
 The funders had no impact on this study.
-All authors approved the final manuscript as submitted and agree to be accountable for all aspects of the work and have no conflict of interests to disclose.
+All authors approved the final manuscript as submitted, agree to be accountable for all aspects of the work and have no conflict of interests to disclose.
 
@@
 \abstract{
-\textbf{Purpose:} Recording young children's vocalizations through wearables is a promising method.
+\textbf{Purpose:} Recording young children's vocalizations through wearables is a promising method to assess language development.
 However, accurately and rapidly annotating these files remains challenging.
@@
 
-\textbf{Method:} Segments identified by LENA\textsuperscript{TM} as produced by the key child were extracted from one daylong recording for each of 20 participants: 10 low-risk control children and 10 children diagnosed with Angelman syndrome, a neurogenetic syndrome characterized by severe language impairments.
+\textbf{Method:} Segments identified by LENA® as produced by the key child were extracted from one daylong recording for each of 20 participants: 10 low-risk control children and 10 children diagnosed with Angelman syndrome, a neurogenetic syndrome characterized by severe language impairments.
 Speech samples were annotated by trained annotators in the laboratory as well as by citizen scientists on Zooniverse.
@@
 \textbf{Conclusion:} Annotations obtained through a citizen science platform can help us overcome challenges posed by the process of annotating daylong speech recordings.
-Particularly when used in composites or derived metrics, such annotations can be used to investigate early markers of language delays in non-typically developing children.
+Particularly when used in composites or derived metrics, such annotations can be used to investigate early markers of language delays.
 }
@@
 Some research suggests that spontaneous behavior, captured for instance via home videos, could provide important indices to development (Belardi et al., 2017; Overby, Belardi, \& Schreiber, 2020).
-Advancements in the field of wearable technologies, such as LENA\textsuperscript{TM} recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally (Oller et al., 2010; Rankine et al., 2017; VanDam \& Yoshinaga-Itano, 2019).
+Advancements in the field of wearable technologies, such as LENA® recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally (Oller et al., 2010; Rankine et al., 2017; VanDam \& Yoshinaga-Itano, 2019).
 Wearable recorders allow data collection to happen in the child's natural environment, and at a large scale, which may be particularly helpful for children whose speech is not easily elicited.
@@
 
-A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis, and it is beyond the scope of this paper to provide a full summary of this work {[}but see \textbf{XX ADD REFERENCE} for a recent summary{]}.
-Moreover, the way in which vocal development has been studied in the past may not easily translate to long-form recordings.
-As mentioned earlier, long-form recordings have several advantages, including capturing the child's vocal patterns in their natural environment and being able to accumulate a great amount of data easily, which is particularly useful for diagnoses and pathologies characterized by low levels of vocal production.
+A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis.
+It is beyond the scope of this paper to provide a full summary of this extensive body work, mainly because the way in which vocal development has been studied in the past may not easily translate to long-form recordings.
+Long-form recordings have several advantages, including capturing the child's vocal patterns in their natural environment and being able to accumulate a great amount of data easily, which is particularly useful for diagnoses and pathologies characterized by low levels of vocal production.
 However, this also means that the audiorecording is harder to process than an audiorecording gathered in more manicured and stable conditions: In a typical daylong recording, the child will sometimes be in a quiet room with just her primary caregiver, but later in a noisy supermarket or having dinner with family and friends.
@@
 
-The most commonly used software was created by the LENA\textsuperscript{TM} Foundation, and it returns a diarization of the audio signal split into key talkers, including the child wearing the device, as well as a split of these audio attributed to the child into three categories: crying, other fixed and vegetative signals, and speech vocalizations.
+The most commonly used software was created by the LENA® Foundation, and it returns a diarization of the audio signal split into key talkers, including the child wearing the device, as well as a split of sections attributed to the child into three categories: crying, other fixed and vegetative signals, and speech vocalizations.
 These subcategories have not been widely validated (Cristia et al., 2020).
-Moreover, LENA\textsuperscript{TM} lacks an important distinction between more and less advanced vocalizations.
+Moreover, until recently LENA® lacked an important distinction between more and less advanced vocalizations.\footnote{The latest version of the LENA® software does tag canonical utterances separately from non-canonical ones.}
 
 Other automatized algorithms have been developed in the last two years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical (Schuller et al., 2019).
-The split of speech vocalizations into canonical and non-canonical, and the overt inclusion of laughing, allow a somewhat finer-grained picture of children's vocal activity than what the LENA\textsuperscript{TM} algorithm currently offers.
-Another advantage of this new line of research is that there exists a challenge where performance has been benchmarked, so unlike LENA\textsuperscript{TM}'s algorithms, we know which algorithms perform better than others.
-Indeed, the ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55\%.
+The overt inclusion of laughing allows a somewhat finer-grained picture of children's vocal activity than what the LENA® algorithm currently offers.
+Another advantage of this new line of research is that there exists a challenge where performance has been benchmarked, so unlike LENA®'s algorithms, we know which algorithms perform better than others.
+Indeed, the ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with an unweighted average recall of around 55\% on the test set.
 By re-using the same test set, researchers can develop new algorithms and prove that they can outperform this baseline (or any subsequent best-performing algorithm benchmarked in the same test set).
@@
 Table 1 provides an idea of the accuracy across different methods.
-Specifically, we calculated accuracy of the LENA\textsuperscript{TM}'s subtypes\footnote{If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as \enquote{Junk} or not categorized.} based on the laboratory annotations used in the present study; and we also show the results of ComParE 2019 BabySounds sub-challenge, although note it is not based on the same data as the other two columns.
-Table 1 shows that LENA\textsuperscript{TM} has a good recall for speech-like vocalizations and crying but not laughing, and by design it does not distinguish between canonical and non-canonical vocalizations.
-The ComParE 2019 BabySounds sub-challenge's baseline model achieved a more moderate recall on crying than LENA, but clearly outperforms the LENA\textsuperscript{TM} algorithm in the Laughing category, with the additional advantage of distinguishing canonical and non-canonical -- although note that performance in non-canonical is barely above a chance level of 20\% (based on 5 categories, assuming equal probabilities).
-Notice that the baseline model also included a category called \enquote{Junk} for sections that turned out not to be the child's vocalizations after all, but LENA\textsuperscript{TM} here has a low recall because only sections that LENA\textsuperscript{TM} classified as being child vocalizations are considered in the present study, and thus this class's category cannot but be low-performing for LENA\textsuperscript{TM}.
-That is, since we selected only segments attributed to the key child, any segment that turns out not to be a child vocalization is a LENA\textsuperscript{TM} error.
-We return to the comparison between LENA\textsuperscript{TM}, the ComParE 2019 BabySounds sub-challenge baseline, and our Zooniverse-based method in the Discussion, but for now, we simply point out that there are new algorithms being developed, but their performance --although competitive when compared to LENA\textsuperscript{TM}-- is underwhelming.
+Specifically, we calculated accuracy of the LENA®'s subtypes\footnote{If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as \enquote{Junk} or not categorized.
+When these data were analyzed, the LENA® software did not distinguish canonical and non-canonical.} based on the laboratory annotations used in the present study; and we also show the results of ComParE 2019 BabySounds sub-challenge, although note it is not based on the same data as the other two columns.
+Table 1 shows that LENA® has a good recall for speech-like vocalizations and crying but not laughing, and at the time it did not distinguish between canonical and non-canonical vocalizations.
+The ComParE 2019 BabySounds sub-challenge's baseline model achieved a more moderate recall on crying than LENA, but clearly outperforms the LENA® algorithm in the Laughing category, with the additional advantage of distinguishing canonical and non-canonical -- although note that performance in non-canonical is barely above a chance level of 20\% (based on 5 categories, assuming equal probabilities).
+Notice that the baseline model also included a category called \enquote{Junk} for sections that turned out not to be the child's vocalizations after all.
+LENA® has a low recall for the Junk category because only sections that LENA® classified as being child vocalizations are considered in the present analysis.
+As a result, any segment that turns out not to be a child vocalization is a LENA® error.
+We return to the comparison between LENA®, the ComParE 2019 BabySounds sub-challenge baseline, and our Zooniverse-based method in the Discussion; for now, we simply point out that there are new algorithms being developed, but their performance --although competitive when compared to LENA®-- is underwhelming.
 
@@
 
-One possibility that has only recently begun to be explored is the use of derived metrics.
-So instead of evaluating algorithms and other annotation procedures on their accuracy at individual labels, one can derive a metric that more closely relates to vocalization development.
+One possibility that has only recently begun to be explored is the use of \emph{derived metrics}.
+So instead of evaluating algorithms and other annotation procedures on accuracy of individual sections of the audio, one can integrate over time (e.g., over the whole day-long audio) to derive a metric that more closely relates to vocalization development.
 
-For example, most work using LENA\textsuperscript{TM} software reports on children's vocalization counts -- this is a derived metric because it is not simply a description of sections of the audio, but instead it integrates over a certain period of time.
-Although this has been sometimes criticized as being more about quantity than quality (McDaniel, Yoder, Estes, \& Rogers, 2020), this is a promising metric of individuals' vocalization development because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA\textsuperscript{TM} (Cristia et al., 2020).
+For example, most work using LENA® software reports on children's vocalization counts -- a derived metric because it is not simply a description of sections of the audio, but instead it integrates over the whole recording length.
+Although linguistic vocalization count has been criticized as being more about quantity than quality (McDaniel, Yoder, Estes, \& Rogers, 2020), it is a promising metric of individuals' vocalization development because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA® (Cristia et al., 2020).
 Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r \textasciitilde{} .3 with standardized language scores in a meta-analysis (Wang, Williams, Dilley, \& Houston, 2020).
-Thus despite representing a composite of skills, vocalization count may provide a useful estimate of vocalization development.
+LENA®'s child vocalization counts (CVC) aggregates all key child vocalizations detected as being \emph{linguistic}.
+As a result, an increase in CVC could indicate an actual increase in how many linguistic vocalizations the child produces, or a change in how well they are detected as such by the algorithm.
+Thus, despite representing a composite of skills, vocalization count may provide a useful estimate of vocalization development.
 
@@
 One metric that can be derived once children's vocalizations are split into crying, laughing, canonical, and non-canonical is the Linguistic Proportion: the proportion of vocalizations that are linguistic (canonical and non-canonical) out of all vocalizations.
-To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age.
+To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age early on and eventually plateaus.
 
 Yet another metric that could be extracted to estimate relative linguistic complexity is the Canonical Proportion: the proportion of vocalizations that contain a canonical transition or syllable out of all linguistic (canonical and non-canonical) vocalizations.
-One recent study has established that the Canonical Proportion extracted from daylong recordings is significantly correlated with age in a multicultural and multilingual sample (Cychosz et al., 2019).
-Importantly, this association was established in a sample of children going up to 3 years of age, and thus well beyond the babbling period, and into first words and word combinations.
-
-There is a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this work focuses on babbling and thus typically on infants under one year of age.
-A critical milestone involves the increasingly common production of canonical syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech (Oller, Eilers, Neal, \& Cobo-Lewis, 1998).
+Cychosz et al.
+(2021) found that the Canonical Proportion extracted from daylong recordings was significantly correlated with age in a multicultural and multilingual sample of children up to 3 years of age, and thus well beyond the babbling period, and into first words and word combinations.
+That study built on a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this prior work focuses on babbling and thus typically on infants under one year of age.
+In these younger infants, a critical milestone involves the increasingly common production of canonical syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech (Oller, Eilers, Neal, \& Cobo-Lewis, 1998).
 Given its adult-like consonant-vowel or vowel-consonant structure, canonical babble is considered to be a starting point on the path to recognizable speech.
-Canonical syllables show a higher complexity given the smooth articulatory transition between a consonant and a vowel (or vice versa), when compared to more primitive sounds such as squeals, or isolated vowels.
-Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development (Oller, 2000).
+When compared to more primitive sounds such as squeals, or isolated vowels, canonical syllables show a higher complexity given the smooth articulatory transition between a consonant and a vowel (or vice versa).
+Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development (Cychosz et al., 2021; Oller, 2000).
 In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization counts in a sample of children diagnosed with an autism spectrum disorder (McDaniel et al., 2020).
@@
 The first pertains to how the data are annotated, with human annotation being costly.
-The second relates to how descriptors of vocal development are generated, i.e., how annotations of audio sections are integrated into vocal development metrics.
+The second relates to how descriptors of vocal development are generated, i.e., how annotations of individual audio sections are integrated over all data for a given child to derive child-level vocal development metrics.
 
-Crowdsourcing refers to the process whereby a task is solved by a crowd, rather than an individual.
+\enquote{Crowdsourcing} refers to the process whereby a task is solved by a crowd, rather than an individual.
 A number of fields, particularly in the data-driven sciences, have already engaged in the collection of data (including annotations) through crowdsourcing, thanks to its low cost and ecological value (Crump, McDonnell, \& Gureckis, 2013; Sescleifer, Francoisse, \& Lin, 2018).
-For example, a systematic review on crowd-sourced ratings of speech found that \enquote{lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology}; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81 (Sescleifer et al., 2018).
+For example, a systematic review on crowdsourced ratings of speech found that \enquote{lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology}; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81 (Sescleifer et al., 2018).
 On the other hand, the systematic review returned only 8 studies (of which only four were published in peer-reviewed journals), suggesting that there is considerable need for further research on this topic.
@@
 Mechanical Turk (MTurk) MTurk is an online labor market created by Amazon to assist \enquote{requesters} in hiring and paying \enquote{workers} for the completion of computerized tasks.
-Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable (Berinsky, Huber, \& Lenz, 2012), some question marks are raised as some \enquote{workers} turn out to be bots, or are poorly motivated (and potentially exploited) humans.
+Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable (Berinsky, Huber, \& Lenz, 2012), question marks are raised as some \enquote{workers} turn out to be bots, or are poorly motivated (and potentially exploited) humans.
 
@@
 Cychosz et al.
-(2019) drew child vocalization data from a diverse set of corpora centered on children learning one of four languages: English, Tseltal, Tsimane', and Yélî.
-For the English and Tsimane' corpora, vocalizations were automatically identified using LENA\textsuperscript{TM}, whereas the other two were extracted through manual segmentation.
+(2021) drew child vocalization data from a diverse set of corpora centered on children learning one of four languages: English, Tseltal, Tsimane', and Yélî.
+For the English and Tsimane' corpora, vocalizations were automatically identified using LENA®, whereas the other two were extracted through manual segmentation.
 Segments were then split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received three classifications into Crying, Laughing, Canonical, Non-Canonical, or Junk (with the latter tag used for clips that did not contain a child voice).
@@
 They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the \enquote{Junk} category was very high.
-Additionally, they found that the correlation between Canonical Proportion and age was much weaker in two corpora where vocalizations were segmented with LENA\textsuperscript{TM} than through manual segmentation.
+Additionally, they found that the correlation between Canonical Proportion and age was much weaker in two corpora where vocalizations were segmented with LENA® than through manual segmentation.
 In sum, these analyses within corpora reveal that data may not be uniformly useful, with potential variation across corpora, children, and methods.
@@
 In addition, we went beyond Cychosz et al.
-(2019)\enquote{s study in two important ways.
+(2021)\enquote{s study in two important ways.
 First, we relied on the largest and best established citizen science platform: Zooniverse hosts more than 1.6 million users from diverse walks of life, and it offers a completely automatized API system to more easily scale tasks in a transparent and cumulative science fashion.
 In contrast, Cychosz et al.
-(2019) actively recruited people to provide annotations, stating that \enquote{{[}a{]}nnotators included language and speech researchers, undergraduate students, research assistants, and other interested parties, totaling 136 unique annotators}.
+(2021) actively recruited people to provide annotations, stating that \enquote{{[}a{]}nnotators included language and speech researchers, undergraduate students, research assistants, and other interested parties, totaling 136 unique annotators}.
 A process that draws from research assistants relies on expert resources, whose time is expensive.
 Moreover, the people these researchers recruit tend to be more accustomed to child vocalization data, which may improve their annotation performance, and thus provide an overestimate of the quality of the annotations that can be realistically done within a truly citizen science framework.
-Those authors also had to rely on manual extraction of the resulting data from the iHearUplay team, which added to the administrative load.
 Second, a core goal of the present work was to determine how citizen scientists} annotations fare compared to the current gold standard, laboratory annotations.
 Cychosz et al.
-(2019) did not have a gold standard for their data.
+(2021) did not have a gold standard for their data.
 
@@
 
-When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk judgments); and the Canonical Proportion (Canonical)/(all linguistic judgments).
+When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk labels); and the Canonical Proportion (Canonical)/(all linguistic labels).
 
-It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels.
-We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA\textsuperscript{TM} (see meta-analyses in Wang et al., 2020; Cristia et al., 2020); as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range.
-Some previous work had shown that vocal development proceeded differently when compared to a normative sample (Hamrick et al., 2019; Hamrick \& Tonnsen, 2019).
+It was important for generalizability purposes to examine data from children with variable biological age, mental age, and vocal maturity levels.
+We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA® (see meta-analyses in Wang et al., 2020; Cristia et al., 2020); as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range.
+Some previous work had shown that vocal development proceeded differently among children with an Angelman syndrome diagnosis when compared to a normative sample (Hamrick et al., 2019; Hamrick \& Tonnsen, 2019).
 Generally, studies on vocal development among children with genetic disorders reveal lower volubility and lower prevalence of canonical syllables (Belardi et al., 2017; Rankine et al., 2017; see also Roche et al., 2018 for a review of retrospective video analysis of children diagnosed with Autism Spectrum Disorder, Rett Syndrome, or Fragile X, documenting these and other vocal characteristics).
 Low volubility is one of the contexts in which daylong recordings may be particularly advantageous over alternative recording methods (see Rankine et al., 2017 for discussion), since it is challenging to measure vocal development accurately based on shorter recordings (as the child may not vocalize at all in a 1-2h recording).
-Additionally, the ideal method will be sensitive not only to the large changes that occur over normative development, but also to the potentially smaller individual differences found among children.
+Additionally, the ideal method is sensitive not only to the large changes that occur over normative development, but also to the potentially smaller individual differences found among children.
 Such a level of sensitivity is important if these methods are ever to be used to detect potential effects of intervention.
@@
 Computer code used to collect the Zooniverse judgments is available from GitHub (Semenzin \& Cristia, 2020).
-This manuscript is reproducible thanks to the use of RMd (Baumer \& Udwin, 2015) and Papaja (Aust \& Barth, 2017) on R (Team \& others, 2013).
+This manuscript is reproducible thanks to the use of RMd (Baumer \& Udwin, 2015) and Papaja (Aust \& Barth, 2017) on R (R Core Team \& others, 2013).
 
@@
 
-Recordings were obtained with the Language ENvironment Analysis (LENA\textsuperscript{TM}) Digital Language Processor, a lightweight recorder (\textless{} 60 g, 5.5 cm x 8.5 cm x 1.5 cm) designed to be worn inside a breast-pocket of purpose-made clothing.
+Recordings were obtained with the Language ENvironment Analysis (LENA®) Digital Language Processor, a lightweight recorder (\textless{} 60 g, 5.5 cm x 8.5 cm x 1.5 cm) designed to be worn inside a breast-pocket of purpose-made clothing.
 In our data, every child contributed at least one daylong recording, with one child contributing two recordings (duration 11.58-16.00 hours, mean 15.05 hours).
 
-\hypertarget{data-preprocessing-lenatm}{%
-\subsubsection{\texorpdfstring{Data preprocessing (LENA\textsuperscript{TM})}{Data preprocessing (LENATM)}}\label{data-preprocessing-lenatm}}
+\hypertarget{data-preprocessing-lena}{%
+\subsubsection{Data preprocessing (LENA®)}\label{data-preprocessing-lena}}
 
 The first step we took before beginning the annotation process was the identification of children's vocalizations within day-long recordings.
-We did this automatically, using the LENA\textsuperscript{TM} proprietary software (Xu et al., 2008).
-Specifically, the LENA\textsuperscript{TM} software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child's environment (e.g., Female-Adult, TV).
+We did this automatically, using the LENA® proprietary software (Xu et al., 2008).
+Specifically, the LENA® software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child's environment (e.g., Female-Adult, TV).
 Key child labels have a precision of about 60\% and a recall of about 50\% (Cristia, Lavechin, et al., 2020).
 These stretches of audio assigned to the key child can contain cries, vegetative sounds, and linguistic vocalizations, as well as interstitial pauses.
-We will refer to them as \emph{segments}, to highlight the fact that they are stretches of audio segmented out as belonging to the key child, but they may not be individual vocalizations.
+We refer to them as \emph{segments}, to highlight the fact that they are stretches of audio segmented out as belonging to the key child, but they may not be individual vocalizations.
 
@@
 Annotation was carried out at the segment level during lab coding.
-A subset of segments LENA\textsuperscript{TM} identified as the key child were selected for human annotation.
+A subset of segments LENA® identified as the key child were selected for human annotation.
 Segments were selected from 30 five-minute sections of the child's recording: 10 sections during which the child was vocalizing at their highest rate, and 20 sections drawn randomly from the remainder of the recording.
@@
 After this, they completed a brief training on vocal maturity that involved 1) reading relevant literature, 2) reviewing examples of various annotation categories and 3) submitting answers to a brief quiz to test their annotation accuracy.
-Coders then proceeded to the classification, where they made decisions on whether a segment is a cry, laugh, non-canonical syllable(s), canonical syllable(s), or a word.
-Coders were instructed to classify each segment based on the highest level of vocal maturity it contains (e.g., a segment with both canonical and non-canonical syllables would be classified as canonical; a segment with both laughing and non-canonical syllables would be classified as non-canonical).
-For canonical, non-canonical, and word segments, coders indicated how many syllables of each type of vocal maturity they hear (e.g., \enquote{ah ma ba} would be 1 non-canonical and 2 canonical syllables).
+Coders then proceeded to the classification, where they made decisions on whether a segment was a cry, laugh, non-canonical syllable(s), canonical syllable(s), or a word.
+Coders were instructed to classify each segment based on the highest level of vocal maturity it contained (e.g., a segment with both canonical and non-canonical syllables would be classified as canonical; a segment with both laughing and non-canonical syllables would be classified as non-canonical).
+For canonical, non-canonical, and word segments, coders indicated how many syllables of each type of vocal maturity they heard (e.g., \enquote{ah ma ba} would be 1 non-canonical and 2 canonical syllables).
 For the purposes of the present study, we considered a \enquote{word} judgment to be equivalent to a \enquote{canonical} judgment.
@@
 
-Coders are also given a \enquote{Don't Mark} option, which they are instructed to use if the segment does not sound like a segment made by the target child, or if there is any overlapping speech or other noise in the background which could affect acoustic analyses.
-Notice, therefore, that this category is broader than the \enquote{Junk} category used by Zooniverse annotators.
+Coders were also given a \enquote{Don't Mark} option, which they were instructed to use if the segment did not sound like a segment made by the target child, or if there is any overlapping speech or other noise in the background which could affect acoustic analyses.
+This category is broader than the \enquote{Junk} category used by Zooniverse annotators, which is explained in the next section.
 
-This process resulted in high levels of inter-rater agreement: Two or more of the three total coders who annotated that segment agreed on 97.22\% of the lab-annotated segments.
+This process resulted in high levels of inter-rater agreement: Two or more of the three total coders who annotated that segment agreed on 97.20\% of the lab-annotated segments.
 
@@
 
-A substantial concern that emerges when considering citizen science annotation for spontaneous speech data is the risk of a privacy breach: even short clips can contain personal information that can expose the identity of the speaker, such as addresses or names, or contain sensitive and private information.
+A substantial concern that emerges when considering citizen science annotation for spontaneous speech data is the risk of a privacy breach: Even short clips can contain personal information that can expose the identity of the speaker, such as addresses or names, or contain sensitive and private information.
 Fortunately, Seidl et al.
-(2019) determined that when segments are divided into shorter clips (400-600ms), human annotators with little training can code our categories of interest (Canonical, Laughing, Crying, etc.) with a classification quality comparable to the one carried out on full segments.
-We therefore used very short clips (500ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.
+(2019) determined that when segments are divided into shorter clips (400-600 ms), human annotators with little training can code our categories of interest (Canonical, Laughing, Crying, etc.) with a classification quality comparable to the one carried out on full segments.
+We therefore used very short clips (500 ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.
 
-Specifically, the segments extracted were automatically cut into clips of exactly 500ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform.
-We will call this level \emph{chunk}, to convey that they are part of a bigger unit (the LENA\textsuperscript{TM}-identified \enquote{segment}).
-Note that segments and chunks are collectively referred to as clips because they are parts of the longer audio recording.
-To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms chunks and their corresponding segment.
+Specifically, the segments extracted were automatically cut into clips of exactly 500 ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform.
+We call this level \emph{chunk}, to convey that they are part of a bigger unit (the LENA®-identified \enquote{segment}).
+Note that segments and chunks are collectively referred to as \emph{clips} because they are parts of the longer audio recording.
+To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500-ms chunks and their corresponding segment.
 Chunks were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling (Bowyer, Lintott, Hines, Allan, \& Paget, 2015).
 
-Citizen scientists could access our project, \enquote{Maturity of Baby Sounds}, by logging in, or as anonymous participants.
-Before starting the annotation, participants were given a quick tutorial which walked them through the steps of the annotation workflow (i.e., how to play a sound, how to make a selection; see Figure 1).
+Citizen scientists could access our project, \enquote{Maturity of Baby Sounds}, by logging in, or as anonymous annotators.
+Before starting the annotation, citizen scientists were given a quick tutorial which walked them through the steps of the annotation workflow (i.e., how to play a sound, how to make a selection; see Figure 1).
 In this tutorial, we included one audio example for each category in order to make the classification task as smooth as possible.
@@
 
-Users were asked to assign each 500ms chunk to one out of five possible categories: (1) Canonical, (2) Non-Canonical, (3) Crying, (4) Laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
-Notice that the latter is narrower than the \enquote{Don't mark} category of the lab annotation routine, which was used any time there was noise or overlap that could affect acoustic analyses.
+Users were asked to assign each 500-ms chunk to one out of five possible categories: (1) Canonical, (2) Non-Canonical, (3) Crying, (4) Laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
+Notice that the latter is narrower than the \enquote{Don't mark} category of the lab annotation routine, which was used any time there was any noise or overlap that could affect acoustic analyses.
 
@@
 
-An impressive total of 4,825 individual Zooniverse users provided labels for the Maturity of Baby Sounds project, of which the present data set is one part.
-For this project, we collected a total of 169,767 judgments provided for 33,731 500-ms chunks, corresponding to 11,980 LENA\textsuperscript{TM} segments.
-Nearly a fifth of chunks did not have at least 3 labels in agreement out of the 5 Zooniverse labels (N = 6,585, 19\% of all chunks).
-Of the chunks without a majority agreement, 4341 (66\%) contained one or two Junk judgments (out of 5), 6523 (99,9\%) had at least two matching judgments (the threshold used for lab-annotated segments), and only 61 (0,01\%) had 5 different judgments.
-Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the 81\% of chunks that did have at least 3 labels in agreement; this represented 135,725 labels for 27,145 chunks, corresponding to 11,593 LENA\textsuperscript{TM} segments.
-As the segments average 1.12 seconds in length, this means about 3.8 hours of audio data were annotated by 8 different annotators (3 in the laboratory, 5 on Zooniverse).
+The following description uses several key concepts: A \enquote{judgment} is a single classification provided by a Zooniverse user in response to a single chunk (or a single classification provided by a lab annotator in response to a single segment); a \enquote{chunk} is a 500-ms clip extracted from the daylong audio to represent part or all of a segment; a segment is a section of the daylong audio automatically classified as having been uttered by the key child according to the LENA® algorithm; and a \enquote{label} is the final category a chunk or segment was classified as.
+Bearing this terminology in mind, we provide an overview of the data and how it was processed before the main analyses.
 
-Lab annotators provided judgments at the level of LENA\textsuperscript{TM} segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments.
-Therefore, we needed to combine chunk-level judgments to reconstruct segment-level judgments.
+An impressive total of 1,532 individual Zooniverse users provided 169,803 judgments on 33,730 500-ms chunks, corresponding to 11,980 LENA® segments.
+Given that each chunk is 500 ms in length, this means about 4.68 hours of audio data were annotated by 8 different annotators (3 in the laboratory, 5 on Zooniverse).
+For a variety of technical mishaps, 1,158 judgments had to be excluded, leaving 33,729 chunks for which we had five Zooniverse judgments (for a total of 168,645 judgments included).
+Among these 33,729 chunks, 0.18\% had no matching judgments, 19.52\% had two matching judgments (the threshold used for lab-annotated segments), and 80.30\% had three or more matching judgments, or majority agreement within Zooniverse judgments.
+Future work may explore different ways of setting the minimal requirement for convergence, but for further analyses here, we focused on the 80.30\% of chunks that \emph{did} have at least three judgments in agreement (i.e., majority agreement).
+This represented 135,420 judgments for 27,084 chunks.
 
-We did this by considering the majority label extracted from Zooniverse judgments for all chunks associated with each segment, and then following these rules: If the majority label for all chunks associated with a given segment were Junk, then the segment was labelled as Junk; then, following a hierarchy (Canonical\textgreater{}Non-canonical\textgreater{}Crying\textgreater{}Laughing), segments with at least one instance of each judgment were labelled as such:
+There was one more layer of processing that was necessary for our main analyses: Lab annotators provided judgments at the level of LENA® segments, whereas Zooniverse annotators provided judgments on 500 ms chunks, which are typically smaller than the segments.
+Therefore, we needed to combine chunk-level labels to reconstruct segment-level labels.
+We did this by considering the majority labels of all chunks associated with each segment as follows:
 
@@
 \item
-  if one Canonical judgment was present, then the segment was labelled as Canonical,
+  If the majority label for all chunks associated with a given segment were Junk, then the segment was labelled as Junk,\\
+\item
+  else, if one Canonical judgment was present for any of the chunks associated to the segment, then the segment was labelled as Canonical,
 \item
@@
 
-This is essentially the same thought process expert annotators followed when labeling segment-level audio, as each segment can represent multiple classification subtypes.
-LENA\textsuperscript{TM} segments are not equivalent to child vocalizations: A segment may contain 1 or more vocalizations, which may be of the same type or not.
+This is essentially the same thought process Lab annotators followed when jduging segments: a LENA® segment may contain 1 or more vocalizations, which may be of the same type or not.
 Figure 2 conveys the relationship between chunks and segments, and how judgments are combined into labels.
+After removing segments for which there was no majority label in the lab and/or Zooniverse coding, we were left with data for 11268 segments.
 
@@
 In this section, we provide descriptive analyses of our dataset.
-According to lab annotators, 14\% of segments were Canonical, 60\% Non-Canonical, 1\% Laughing, and 4\% Crying, with the remaining 21\% being categorized as \enquote{Don't mark}.
-Zooniverse data revealed a similar distribution: 15\% Canonical, 60\% Non-Canonical, 3\% Laughing, 7\% Crying, 15\% Junk.
+According to lab annotators, 15.82\% of segments were Canonical, 57.07\% Non-Canonical, 1.65\% Laughing, and 5.23\% Crying, with the remaining 20.23\% being categorized as \enquote{Don't mark}.
+Zooniverse data revealed a similar distribution: 14.71\% Canonical, 60.01\% Non-Canonical, 3.63\% Laughing, 8.87\% Crying, 12.79\% Junk.
 Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical + Non-Canonical) / All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical / (Canonical + Non-Canonical) (i.e., we remove Junk, Crying, and Laughing).
@@
 Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups.
-There was a near-zero relationship among the children diagnosed with Angelman syndrome r(8) = -.091, CI {[}-.682,.571{]}, \emph{p} = .802{]}; and a significant association among low-risk control children r(8) = .7, CI {[}.126,.923{]}, \emph{p} = .024{]}.
-The Canonical Proportion exhibited non-significant developmental decreases among children diagnosed with Angelman syndrome r(8) = -.393, CI {[}-.82,.314{]}, \emph{p} = .261{]}; and marginal developmental increases among low-risk control r(8) = .516, CI {[}-.168,.865{]}, \emph{p} = .126{]}.
+There was a near-zero relationship among the children diagnosed with Angelman syndrome r(8) = .015, CI {[}-.621,.638{]}, \emph{p} = .968{]}; and a significant association among low-risk control children r(8) = .741, CI {[}.21,.935{]}, \emph{p} = .014{]}.
+The Canonical Proportion exhibited a non-significant developmental decrease among children diagnosed with Angelman syndrome r(8) = -.557, CI {[}-.878,.112{]}, \emph{p} = .095{]}; and a marginal developmental increase among low-risk control r(8) = .607, CI {[}-.036,.895{]}, \emph{p} = .063{]}.
 
-Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(8) = -.174, CI {[}-.724,.512{]}, \emph{p} = .630{]}; whereas low-risk control children showed a significant increase with age r(8) = .629, CI {[}-.001,.902{]}, \emph{p} = .051{]}.
-Similarly, there were non-significant developmental decreases in the Canonical Proportion among children with Angelman syndrome r(8) = -.407, CI {[}-.825,.299{]}, \emph{p} = .243{]}; and marginal developmental increases among low-risk control children, r(8) = .514, CI {[}-.171,.864{]}, \emph{p} = .129{]}.
+Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(8) = -.131, CI {[}-.703,.543{]}, \emph{p} = .717{]}; whereas low-risk control children showed a significant increase with age r(8) = .672, CI {[}.073,.915{]}, \emph{p} = .033{]}.
+Similarly, there was a non-significant developmental decrease in the Canonical Proportion among children with Angelman syndrome r(8) = -.404, CI {[}-.824,.302{]}, \emph{p} = .247{]}; and a marginal developmental increase among low-risk control children, r(8) = .604, CI {[}-.042,.894{]}, \emph{p} = .064{]}.
 
@@
 
-These visualizations suggest that performance is moderate to good, which was confirmed via statistical analyses.
-The overall (weighted) accuracy is 73\%, CI = {[}72,74{]}, kappa is .531, and the Gwet's AC1 coefficient is .682, CI = {[}.669,.694{]}.
+These visualizations suggest that performance was moderate to good, which was confirmed via statistical analyses.
+The overall (weighted) accuracy was 67\%, CI = {[}66,68{]}, kappa is .448, and the Gwet's AC1 coefficient is .609, CI = {[}.598,.619{]}.
 
@@
 
-Although the classification at the segment level is only moderately accurate, what we are ultimately interested in is whether citizen scientists' classifications are able to provide a reliable snapshot of children's individual development.
+Although the classification at the segment level was only moderately accurate, what we are ultimately interested in is whether citizen scientists' classifications are able to provide a reliable snapshot of children's individual development.
 We illustrate this association in Figure 5.
-Looking at all 20 children together, we found a strong positive correlation r(18) = .794, CI {[}.543,.915{]}, \emph{p} \textless{} .001{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman syndrome r(8) = .878, CI {[}.555,.971{]}, \emph{p} = .001{]}; low-risk control r(8) = .828, CI {[}.415,.958{]}, \emph{p} = .003{]}.
+Looking at all 20 children together, we found a strong positive correlation r(18) = .821, CI {[}.594,.927{]}, \emph{p} \textless{} .001{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
+When we split by participant group, correlations remained high: for Angelman syndrome r(8) = .89, CI {[}.592,.974{]}, \emph{p} = .001{]}; low-risk control r(8) = .83, CI {[}.419,.959{]}, \emph{p} = .003{]}.
 
@@
 
-Similarly, a strong positive correlation is found in the Canonical Proportion r(18) = .92, CI {[}.806,.968{]}, \emph{p} \textless{} .001{]}.
-When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman syndrome: r(8) = .681, CI {[}.091,.917{]}, \emph{p} = 0.03{]}; than the low-risk control children r(8) = .984, CI {[}.932,.996{]}, \emph{p} \textless{} .001.
+Similarly, a strong positive correlation was found in the Canonical Proportion r(18) = .938, CI {[}.848,.976{]}, \emph{p} \textless{} .001{]}.
+When we split by participant group, correlations remained high although we do note they were somewhat lower for the children with Angelman syndrome: r(8) = .855, CI {[}.487,.965{]}, \emph{p} = 0.00{]}; than the low-risk control children r(8) = .966, CI {[}.858,.992{]}, \emph{p} \textless{} .001.
 
@@
 
-In this section, we report on explorations of under what conditions Zooniverse judgments more closely aligned with laboratory judgments.
-In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as \enquote{Junk} (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora (Cychosz et al., 2019).
+In this section, we report on exploratory analyses, aimed at establishing conditions under which Zooniverse labels aligned with laboratory labels more or less closely.
+In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as \enquote{Junk} (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora (Cychosz et al., 2021).
 A high proportion of \enquote{Junk} may indicate that automated segmentation was errorful for those children, and may be a sign that the rest of the data could be compromised as well.
 
-We investigated this hypothesis by calculating the proportion of their data labeled as \enquote{Junk} for each individual child.
-There was no significant difference in the proportions of their data labeled as \enquote{Junk} for children with Angelman syndrome (M = 0.14) compared to the low-risk group (M = 0.16): Welch's t(13.35) = -0.58, \emph{p} = .569.
+We investigated this hypothesis by calculating the proportion of data labeled as \enquote{Junk} for each individual child.
+There was no significant difference in the proportions of their data labeled as \enquote{Junk} for children with Angelman syndrome (M = 12.89\%, SD = 7.95\%) compared to the low-risk group (M = 14.25\%, SD = 3.97\%): Welch's t(13.23) = -0.48, \emph{p} = .636.
 We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as \enquote{Junk}.
 Results were similar across these two post-hoc subgroups (see Figure 6).
-For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = .863, CI {[}.511,.967{]}, \emph{p} = .001{]}; and for the higher Junk group it was r(8) = .895, CI {[}.608,.975{]}, \emph{p} \textless{} .001.
-For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = .957, CI {[}.824,.99{]}, \emph{p} \textless{} .001{]}; and for the higher Junk group it was r(8) = .895, CI {[}.608,.975{]}, \emph{p} \textless{} .001.
-Thus, it does not seem that a higher proportion of \enquote{Junk} judgments is an index of low quality data.
+For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = .885, CI {[}.576,.973{]}, \emph{p} = .001{]}; and for the higher Junk group it was r(8) = .864, CI {[}.513,.967{]}, \emph{p} .001.
+For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = .958, CI {[}.828,.99{]}, \emph{p} \textless{} .001{]}; and for the higher Junk group it was r(8) = .93, CI {[}.726,.984{]}, \emph{p} \textless{} .001.
+Thus, it does not seem that a higher proportion of \enquote{Junk} labels is an index of low quality data.
 
@@
 
-Above, we concluded that derived metrics integrating information across audio clips (Linguistic and Canonical Proportion, which can be derived from segment- or chunk-level judgments) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA\textsuperscript{TM} segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment.
-As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed judgments across chunks associated to the same segment.
+Above, we concluded that derived metrics integrating information across audio clips (Linguistic and Canonical Proportion, which can be derived from segment- or chunk-level labels) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse annotators) to segments (the original LENA® segments presented to laboratory annotators), because one can calculate the proportions directly from the chunk-level judgment.
+As a result, there was one stage in our post-processing that may not have been necessary, whereby we collapsed labels across chunks associated to the same segment.
 We therefore repeated our analyses but instead of deriving our proportions for the Zooniverse data from the segment-level composite, we did it based on the individual chunk-level annotations.
@@
 
-Regarding correlations with age, we found that the Linguistic Proportion was not strongly associated with age in the Angelman syndrome group: r(8) = -.038, CI {[}-.652,.606{]}, \emph{p} = .918{]}, but increased with age in the low-risk control group: r(8) = .716, CI {[}.157,.927{]}, \emph{p} = .020.
-Similarly, the Canonical Proportion did not exhibit the same pattern across the groups, with developmental decreases found among children with Angelman syndrome r(8) = -.38, CI {[}-.815,.328{]}, \emph{p} = .278{]}; and developmental increases among low-risk control r(8) = .573, CI {[}-.088,.884{]}, \emph{p} = .083.
+Regarding correlations with age, we found very similar results to the analyses based on laboratory annotations, as well as those based on Zooniverse annotations collapsed at the segment level (see Table 3).
+That is, the Linguistic Proportion was not strongly associated with age in the Angelman syndrome group: r(8) = -.038, CI {[}-.652,.606{]}, \emph{p} = .918{]}, but increased with age in the low-risk control group: r(8) = .716, CI {[}.157,.927{]}, \emph{p} = .020.
+Likewise, the Canonical Proportion did not exhibit the same pattern across the groups, with a non-significant developmental decrease found among children with Angelman syndrome r(8) = -.38, CI {[}-.815,.328{]}, \emph{p} = .278{]}; and a marginal developmental increase among low-risk control r(8) = .573, CI {[}-.088,.884{]}, \emph{p} = .083.
 
-As for correlations between Zooniverse- and laboratory-derived metrics (e.g., linguistic proportion as estimated using Zooniverse coding, versus laboratory coding), we observe very similar levels of correlation: Linguistic Proportion overall r(18) = .798, CI {[}.549,.917{]}, \emph{p} \textless{} .001{]} between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.
-When we split by participant group, correlations remain high: for Angelman syndrome r(8) = .845, CI {[}.46,.963{]}, \emph{p} = .002{]}; low-risk control r(8) = .794, CI {[}.328,.949{]}, \emph{p} = .006{]}.
-As for Canonical Proportion overall r(18) = .949, CI {[}.873,.98{]}, \emph{p} \textless{} .001; Angelman syndrome: r(8) = .724, CI {[}.173,.93{]}, \emph{p} = .018{]}; low-risk control infants r(8) = .972, CI {[}.881,.993{]}, \emph{p} \textless{} .001.
+----- INSERT TABLE 3 HERE -----
 
-Since all of these results are very similar to those obtained when collapsing chunk-level judgments to generate segment-level judgments, we conclude that in the future this step can be skipped.
-Instead, researchers can derive Linguistic and Canonical Proportions directly from citizen scientists' chunk-level judgments (which was in fact what Cychosz et al., 2019 did).
+As for correlations between Zooniverse- and laboratory-derived metrics (e.g., the correlation between the Linguistic Proportion as estimated using Zooniverse coding at the chunk level and the Linguistic Proportion via laboratory coding at the segment level), we observe very similar levels of correlation to the segment-based analyses (see Table 2): Linguistic Proportion overall r(18) = .845, CI {[}.644,.937{]}, \emph{p} \textless{} .001{]} between Linguistic Proportion by child from the Zooniverse at the chunk level and the lab annotators' data at the segment level.
+When we split by participant group, correlations remain high: for Angelman syndrome r(8) = .845, CI {[}.459,.962{]}, \emph{p} = .002; low-risk control r(8) = .841, CI {[}.45,.962{]}, \emph{p} = .002.
+As for Canonical Proportion overall r(18) = .963, CI {[}.906,.985{]}, \emph{p} \textless{} .001; Angelman syndrome: r(8) = .866, CI {[}.519,.968{]}, \emph{p} = .001; low-risk control infants r(8) = .974, CI {[}.892,.994{]}, \emph{p} \textless{} .001.
+
+Since all of these results are very similar to those obtained when collapsing across chunk-level labels to generate segment-level labels, we conclude that in the future this step can be skipped.
+Instead, researchers can derive Linguistic and Canonical Proportions directly from citizen scientists' chunk-level labels (which was in fact what Cychosz et al., 2021 did).
 
 Next, we looked at whether having more segments from each child may lead to more reliable metrics.
-We observed a non-significant trend {[}Welch t(17.88)=-0.53{]} for lower number of segments in the Angelman syndrome group (M = 348.70, range 163-599) than among low-risk control infants (M = 375.50, range 224-627), likely because of the lower volubility of the former children.
-Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications (Cychosz et al., 2019).
-We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data is extracted from each child by sampling 100 segments from all children (following Cychosz et al., 2019).
-We randomly sampled 100 segments from each child's data, and recalculated the Linguistic and Canonical Proportions based only on these associated judgments.
+We observed a non-significant trend {[}Welch t(17.94)=-1.20{]} for lower number of segments in the Angelman syndrome group (M = 512.20, range 302-901) than among low-risk control infants (M = 614.60, range 343-1038), likely because of the lower volubility among the former group of children.
+Moreover, these numbers of segments are much larger than those used in previous work relying on citizen science classifications (Cychosz et al., 2021).
+We therefore asked whether the correlations between laboratory and Zooniverse child-level estimates are affected by how much data were extracted from each child by under-sampling.
+Since Cychosz et al.
+(2021) extracted 100 segments from each child in their study, we randomly sampled 100 segments from each of our children's data, and recalculated the Linguistic and Canonical Proportions based only on these 100 labels.
 We repeated this process 50 times, to assess the extent to which the association between laboratory- and Zooniverse-derived metrics varied in each random sample (e.g., the correlation between Linguistic Proportion as estimated using Zooniverse coding on the one hand, and Linguistic Proportion as estimated using laboratory coding on the other).
 
-The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table 2), particularly for the low-risk younger infants, and for Linguistic Proportion.
+The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table 2, \enquote{100 seg} rows), particularly for the low-risk younger infants, and for Linguistic Proportion.
 This may indicate that, especially in some groups of infants, 100 segments may not be sufficient to capture a stable estimate of the child's Linguistic and Canonical Proportions.
 We also observed quite a bit of variance across the 50 runs.
-For Linguistic Proportion, the range of correlations found for all infants was {[}.442,.893{]}; for the Angelman syndrome group {[}.490,.969{]}; for the low-risk group {[}.289,.989{]}.
-For Canonical Proportion, the range of correlations found for all infants was {[}.750,.943{]}; for the Angelman syndrome group {[}.790,.994{]}; for the low-risk group {[}.547,.939{]}.
+For Linguistic Proportion, the range of correlations found for all infants was {[}.569,.878{]}; for the Angelman syndrome group {[}.493,.95{]}; for the low-risk group {[}.481,.911{]}.
+For Canonical Proportion, the range of correlations found for all infants was {[}.815,.964{]}; for the Angelman syndrome group {[}-.001,.94{]}; for the low-risk group {[}.859,.992{]}.
 
@@
 
-In the present study we assessed the extent to which child vocalizations from LENA\textsuperscript{TM} daylong recordings can be accurately described based on classifications collected using the citizen science platform Zooniverse.
+In the present study we assessed the extent to which child vocalizations from LENA® daylong recordings can be accurately described based on classifications collected using the citizen science platform Zooniverse.
 Our recordings came from children diagnosed with a genetic syndrome associated with severe language disorder, as well as low-risk controls, and included a range of different ages.
-Excerpts of these recordings had previously been annotated by trained experts in the lab, and served as the \enquote{gold standard}.
+Excerpts of these recordings had previously been annotated by trained experts in the lab, who provided the gold standard.
 
@@
 In this case, the lack of correlation with age in the Angelman syndrome group suggests that there is little systematic variance in this metric after 20 months of age.
-In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and relatively stable levels of Canonical Proportion among the children in the Angelman syndrome group.
+In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and relatively stable levels of Canonical Proportion among the children in the Angelman syndrome group (in line with previous work; Hamrick et al., 2019; Hamrick \& Tonnsen, 2019).
 
@@
 This was the case of the \enquote{Junk/Don't mark} category.
-Higher precision and low recall indicates that when Zooniverse annotations return a \enquote{Junk} judgment, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as \enquote{Don't mark} in the laboratory.
-Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the \enquote{Junk} category less) than laboratory annotators, who have access to the whole vocalization.
+The combination of high precision and low recall indicates that when Zooniverse annotations return a \enquote{Junk} judgment, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as \enquote{Don't mark} in the laboratory.
+Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the \enquote{Junk} category less) than laboratory annotators, who have access to the whole segment (and its context).
 This is consistent with the fact that our lab definition of \enquote{Junk} was more stringent than the Zooniverse definition.
-Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use \enquote{Don't mark} if there was \emph{any} noise or overlapping speech that would potentially affect acoustic analyses.
+Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use \enquote{Don't mark} if there was \emph{any} noise or overlapping speech that could affect acoustic analyses.
 
@@
 This means that a high proportion of segments detected as these two categories in the lab can be classified in the appropriate category by Zooniverse annotators, but Zooniverse annotators also put into these categories many segments that lab annotators classified in a different manner.
-Most saliently, there is a problematic level of confusion between Crying and Non-Canonical, whereby lab-detected Crying was almost as likely to be classified as such or Non-canonical in Zooniverse (the recall proportion was 49 and 47\% respectively), and even worse, 55\% of the segments Zooniverse annotators classified as Crying were actually Non-Canonical according to lab annotations.
-This is likely due to the fact that lab annotators had access to the whole segment, and could even listen to the context of that segment, so they could accurately interpret a segment as crying.
+Most saliently, there is a problematic level of confusion between Crying and Non-Canonical, whereby lab-detected Crying was almost as likely to be classified as such or Non-canonical in Zooniverse (the recall was 51 and 46\% respectively), and even worse, 55\% of the segments Zooniverse annotators classified as Crying were actually Non-Canonical according to lab annotations.
+This is probably due to the fact that lab annotators had access to the whole segment, and could even listen to the context of that segment, so they could accurately interpret a segment as crying.
 Thus, it is possible that additional analyses, using for instance pitch or duration characteristics at the level of the whole segment rather than the chunk could improve classification performance, but for the time being, we cannot recommend Zooniverse annotations to identify crying bouts.
@@
 As for Laughing, the recall is quite good but precision is low, with confusion involving \enquote{Junk} and Non-Canonical to similar extents.
-We are tentative in our interpretation of this result because the category was very sparsely represented, with barely 200 segments classified as Laughing in the whole data set.
+We are tentative in our interpretation of this result because the category was very sparsely represented, with barely 186 segments classified as Laughing in the whole data set.
 We thus recommend additional work, although we expect that perhaps there will be a similar problem (and solution) as indicated for Crying, whereby determination of Laughing versus Non-Canonical will require inspection of segment-level properties.
 
-Turning now to the child-level analyses, we were particularly interested in derived metrics because this is more typically the goal of such analyses, for instance to study potential individual or group differences, or investigate the impact of an intervention.
-Our results show that, despite agreement variation at the level of the segment, derived metrics were highly correlated across laboratory and Zooniverse data, meaning that variability in precision and recall did not substantially impact key outcome variables.
+Turning now to the child-level analyses, we were particularly interested in derived metrics because this is more typically the goal of such analyses, for instance to study potential individual or group differences, or to investigate the impact of an intervention on a child or group of children.
+Our results show that, despite only moderate levels of agreement at the level of the segment, our two derived metrics (Linguistic and Canonical Proportions) were highly correlated across laboratory and Zooniverse data, meaning that variability in precision and recall did not substantially impact key outcome variables.
 Errors only seemed systematic for the Linguistic Proportion (and not for Canonical Proportion), and they were due to lower Linguistic Proportions from Zooniverse than laboratory data.
@@
 
-Next, we found that more data led to more reliable measurements: Specifically, our correlations based on 300-500 segments per child were markedly higher than those based on 100 segments per child, particularly when looking at our low-risk, younger participants.
-It is important to note that an advantage of Zooniverse, compared to laboratory annotations, is that it is easier to scale up: More data from each individual child can be annotated this way, which will help countering noise.
+Next, we found that more data led to more reliable measurements: Specifically, our correlations based on 302-1038 segments per child were markedly higher than those based on 100 segments per child, particularly when looking at our low-risk, younger participants.
+It is important to note that an advantage of Zooniverse, compared to laboratory annotations, is that it is easier to scale up: More data from each individual child can be annotated this way, which helps countering noise.
 
@@
 It is important to note that from a methodological point of view, the use of crowdsourcing is not merely a question of being an ecological technique, but it can also offer benefits in the area of scientific rigor and reproducibility.
-As previously mentioned in the Introduction section, Zooniverse offers both an automated system of data handling as well as an enormous pool of users: access to a large-scale group of blinded participants can in itself help researchers overcome the methodological compromises that sometimes lead to biased results.
+As previously mentioned in the Introduction section, Zooniverse offers both an automated system of data handling as well as an enormous pool of users: access to a large-scale group of blinded annotators can in itself help researchers overcome the methodological compromises that sometimes lead to biased results.
 Additionally, crowd-sourcing and open-science frameworks can come together to encourage replication of existing work: open-source software, shared data and citizen science platforms would make it possible, and actually easy, to run identical studies and evaluate research results independently.
@@
 The first area in which more work would be welcome involves the pre-processing.
-We used LENA\textsuperscript{TM} algorithms to detect vocalizations by the key child, but there are newer, better-performing, open-sourced alternatives that could be used instead (Lavechin, Bousbib, Bredin, Dupoux, \& Cristia, 2020).
+We used LENA® algorithms to detect vocalizations by the key child, but there are newer, better-performing, open-sourced alternatives that could be used instead (Lavechin, Bousbib, Bredin, Dupoux, \& Cristia, 2020).
 Still, performance is most often established on children whose development is thought to be normative, being raised in small households located in urban environments, whereas it would be important to check for accuracy in more diverse samples, where the children have a diagnosis (e.g., Dykstra et al., 2013), multiple siblings (e.g., Elo, 2016), and/or are growing up in a rural environment (e.g., Cristia, Lavechin, et al., 2020).
@@
 The Introduction mentioned a challenge that established a baseline performance with unweighted average recall (UAR) of 54\% (when recall is averaged across the 5 classes, Laughing, Crying, Canonical, Non-canonical, and Junk, giving equal weight to each of them rather than weighing them based on their frequency of occurrence).
-As reported in Table 1, our own method leads to a UAR of 57\%, so one may think there is not much to gain with the automated method.
+As reported in Table 1, our own method leads to a UAR of 59\%, so one may think there is not much to gain with the automated method.
 However, notice that the automated method does lead to much better performance than Zooniverse classifications for Crying, and to some improvements in Canonical and Junk.
@@
 That state of the art was challenged by Kaya, Verkholyak, Markitantov, and Karpov (2020), who obtained a UAR of 61\% on the same data as the challenge, thanks to improvements in all of the classes but for Laughing.
-Thus, there may be some gains to be obtained in specific classes.
+Even more recently, Yao, Micheletti, Johnson, and Barbaro (2020) reported on a series of classifiers focused on crying, with astounding improvements in this class.
+Thus, there may be some gains to be obtained in specific vocalization types, particularly those that are rare (like Laughing) and those which require context (Crying).
 
@@
 We hope additional research is devoted to further investigating this measure as a child-level descriptor, since it may be more relevant to emotional disorders (see a review in Halpern \& Coelho, 2016).
-Additionally, while Canonical Proportion has been studied in much previous work, we believe further research is necessary using larger scales (saliently increasing the sheer number of children studied, to have sufficient power to detect group effects) and with broader coverage (notably including children growing up in a broad range of languages and cultures, as in Cychosz et al., 2019), to make sure that assumptions regarding cross-linguistic and maturational universality are amply justified and not merely a result of having low power to detect differences, or simply sampling from similar populations.
+Additionally, while metrics conceptually related to our Canonical Proportion have been studied in much previous work, we believe further research is necessary using larger scales (saliently increasing the sheer number of children studied, to have sufficient power to detect group effects) and with broader coverage (notably including children growing up in a broad range of languages and cultures, as in Cychosz et al., 2021), to make sure that assumptions regarding cross-linguistic and maturational universality are amply justified and not merely a result of having low power to detect differences, or simply sampling from similar populations.
 
 Assuming these child-level metrics continue to be supported by such additional data, it would be interesting for future investigations to expand the analysis of Linguistic Proportion and Canonical Proportion to recover differences and similarities in language development between low-risk children and children with disorders other than Angelman syndrome.
-That is, it is at present unclear whether relatively lower Canonical Proportions will also be found in other populations, or whether this is more specific to Angelman syndrome.
+That is, it is at present unclear whether relatively lower Canonical Proportions are also be found in other populations, or whether this is more specific to Angelman syndrome.
 For instance, some previous work has found differences in babbling even when comparing late talkers with toddlers with no specific diagnosis (Fasolo, Majorano, \& D'Odorico, 2008).
@@
 
-It is also important to note that work on children with non-normative development poses in itself some challenges.
+We also acknowledge that work on children with non-normative development poses in itself some challenges.
 Our data contained a lower number of vocalizations for children diagnosed with Angelman syndrome, even though we selected much younger low-risk control children as the comparison group.
 However, this limitation could be easily overcome in the future: We were limited here because we wanted to use the exact same segments that laboratory annotators had inspected.
-Now that this method has been validated, researchers can extract all segments that LENA\textsuperscript{TM} identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
-This approach may be particularly useful in the case of data from children with a diagnosis, whose data tends to be under-represented in LENA\textsuperscript{TM} research, compared to data from normative children.
-We also hope future work makes use of the scalability of the method by employing larger sample sizes than those used in the present study, where data from only 10 children was used.
+Now that this method has been validated, researchers can extract all segments that LENA® identifies as being the key child, thus maximally using available data to get more reliable estimates, as suggested by one of our exploratory analyses.
+This approach may be particularly useful in the case of data from children with a diagnosis, whose data may be under-represented in LENA® research, compared to data from normative children.
+We also hope future work makes use of the scalability of the method by employing larger sample sizes than those used in the present study, where data were drawn from only 10 children in each group.
 
@@
 Additionally, longitudinal work within the same infant may be needed to track the \enquote{natural history} of vocalizations to better contextualize how these metrics should be interpreted.
-It is also an open question how large must effects be, at the individual level, in order to be detected with this method.
+It is also an open question how large effects must be, at the individual level, in order to be detected with this method.
 For example, there is interest in using naturalistic recordings as outcome measures for clinical (behavioral or pharmacological) trials.
-The fact that agreement across annotators is high even in the data coming from children diagnosed with Angelman syndrome is encouraging, but it would be important to establish whether absolute levels of Canonical Proportion, and/or rates of change over longer time spans, may be good candidates for tracking the effects of any treatment.
+The fact that agreement across annotators is high even in the data coming from children diagnosed with Angelman syndrome is encouraging, but it would be important to establish whether absolute levels of Canonical Proportion, and/or rates of change over longer time spans, may be good candidates for tracking the effects of a given treatment.
 
@@
 In contrast, tasks that require more speech or audio context may not be well suited to citizen science platforms, because they would require playing longer clips, which may reveal identifying or sensitive information.
-Although citizen scientists themselves are well-intentioned, the platform could be used by others who want to exploit the system for other means.
+Although citizen scientists themselves are well-intentioned, the platform could be used by others who want to exploit the system for nefarious goals.
 
@@
 
-In this study we validated the quality of annotations obtained through a citizen science platform, Zooniverse, as compared to a gold standard of human expert annotators.
-We analyzed the correspondence between annotations at the individual segment level, and the individual-child level, using Canonical Proportion and Linguistic Proportion as descriptors.
-We found moderate to good accuracy in the first, and strong positive correlations in the latter.
+In this study, we validated the quality of annotations obtained through a citizen science platform, Zooniverse, as compared to a gold standard of human expert annotators.
+We analyzed the correspondence between annotations at the individual segment level, and the individual child level, using Canonical Proportion and Linguistic Proportion as descriptors.
+We found moderate to good accuracy in the former, and strong positive correlations in the latter.
 We can conclude that citizen scientists are a reliable source of fast and ecological annotation of speech data, particularly when results are combined into child-level descriptors.
@@
 We are grateful to the families who contributed their data; to the Zooniverse volunteers that made this work possible; and the audience and technical committee of SLT 2020 for helpful feedback on preliminary analyses.
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
 
@@
 Notice that clips are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority.
-To compare laboratory and Zooniverse annotations at the level of segments, chunk-level judgments are combined using the hierarchy Canonical\textgreater{}Non-canonical\textgreater{}Crying\textgreater{}Laughing.}
+To compare laboratory and Zooniverse annotations at the level of segments, chunk-level labels are combined using the hierarchy Canonical \textgreater{} Non-canonical \textgreater{} Crying \textgreater{} Laughing \textgreater{} Junk.}
 \end{figure}
@@
 \includegraphics{paper_files/figure-latex/fig-corage-1.pdf}
-\caption{\label{fig:fig-corage}Correlations between child-level descriptors and age as a function of metric (Linguistic Proportion in the top row, Canonical Proportion in the bottom row), annotation method, and child group (red=low-risk, and black= Angelman syndrome).}
+\caption{\label{fig:fig-corage}Correlations between child-level descriptors and age as a function of metric (Linguistic Proportion in the top row, Canonical Proportion in the bottom row), annotation method, and child group (red circles = low-risk, and black crosses = Angelman syndrome).}
 \end{figure}
@@
 \includegraphics{paper_files/figure-latex/fig-corlab-zoo-1.pdf}
-\caption{\label{fig:fig-corlab-zoo}Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of child group (red=low-risk, and black=Angelman syndrome).}
+\caption{\label{fig:fig-corlab-zoo}Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of child group (red circles = low-risk, and black crosses = Angelman syndrome).}
 \end{figure}
@@
 \includegraphics{paper_files/figure-latex/fig-cor-Junk-1.pdf}
-\caption{\label{fig:fig-cor-Junk}Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of whether children's Junk proportion was higher or lower than the median for all children (blue=lower, and green=higher).}
+\caption{\label{fig:fig-cor-Junk}Correlations between child-level descriptors derived from Zooniverse data (y axis) versus lab data (x axis) as a function of whether children's Junk proportion was higher or lower than the median for all children (light blue wheels = lower, and green diamonds = higher).}
 \end{figure}
@@
 
-\caption{\label{tab:tab-LENA-challenge}Comparison of recall percentages obtained with the baseline algorithm created by the ComParE team, in the context of the ComParE 2019 BabySounds subchallenge (Schuller et al., 2019), LENATM labels (note LENATM does not distinguish between Canonical and Non-canonical), and those obtained in this study through Zooniverse annotations.
-Label frequency indicates the prevalence of the relevant label (Crying, Laughing, etc.) in each dataset (the same Lab dataset was used for both LENATM and Zooniverse).
+\caption{\label{tab:tab-LENA-challenge}Comparison of recall percentages obtained with the baseline algorithm created by the ComParE team, in the context of the ComParE 2019 BabySounds subchallenge (Schuller et al., 2019; C2019B), LENA labels (note LENA did not distinguish between Canonical and Non-canonical in this analysis), and those obtained in this study through Zooniverse annotations.
+Label frequency indicates the prevalence of the relevant label (Crying, Laughing, etc.) in each dataset (the same Lab dataset was used for both LENA and Zooniverse).
 UAR stands for unweighted average recall, WAR for weighted average recall (which takes into account label frequency).}
@@
 \hline
- & C2019B & LENA & Zoon. & C2019B & Lab\\
+ & C2019B & LENA & Zoon. & C2019B & Zoon.\\
 \hline
-Crying & 70.6 & 78.0 & 49.0 & 669 & 581\\
-\cline{1-6}
-Laughing & 41.5 & 12.0 & 56.0 & 149 & 137\\
-\cline{1-6}
-Non-can. & 24.1 &  & 77.0 & 4485 & 5512\\
-\cline{1-2}
-\cline{4-6}
-Canonical & 66.4 & \multirow{-2}{*}{\raggedright\arraybackslash 73.0} & 59.0 & 1426 & 1715\\
-\cline{1-6}
-Junk & 67.3 & 1.0 & 44.0 & 4575 & 2169\\
-\cline{1-6}
-UAR & 54.0 & 41.0 & 57.0 &  & \\
-\cline{1-4}
-WAR & 49.9 & 57.0 & 65.0 & \multirow{-2}{*}{\raggedright\arraybackslash } & \multirow{-2}{*}{\raggedright\arraybackslash }\\
+Crying & 70.6 & 78.0 & 50.9 & 669 & 6762\\
+\hline
+Laughing & 41.5 & 56.0 & 62.9 & 149 & 409\\
+\hline
+Non-can. & 24.1 & 77.0 & 78.9 & 4485 & 1441\\
+\hline
+Canonical & 66.4 & 59.0 & 58.3 & 1426 & 1657\\
+\hline
+Junk & 67.3 & 1.0 & 43.5 & 4575 & 999\\
+\hline
+UAR & 54.0 & 41.0 & 58.9 &  & \\
+\hline
+WAR & 49.9 & 57.0 & 55.3 &  & \\
 \hline
@@
 \centering
-\begin{tabular}[t]{r|r}
+\begin{tabular}[t]{l|l|l}
 \hline
-Linguistic Proportion & Canonical Proportion\\
+  & Linguistic Proportion & Canonical Proportion\\
 \hline
-0.878 & 0.681\\
+All seg AS & .89 & .855\\
 \hline
-0.828 & 0.984\\
+Chunks AS & .845 & .866\\
 \hline
-0.794 & 0.920\\
+100 seg AS & .791 & .632\\
+\hline
+All seg LR & .83 & .966\\
+\hline
+Chunks LR & .841 & .974\\
+\hline
+100 seg LR & .787 & .943\\
+\hline
+All seg all & .821 & .938\\
+\hline
+Chunks all & .845 & .963\\
+\hline
+100 seg all & .771 & .905\\
 \hline
@@
 \end{table}
+
+\begin{table}
+
+\caption{\label{tab:tab-corage}Pearson correlation coefficients between derived metrics (Linguistic Proportion = LP; Canonical Proportion = CP) and age, for the Angelman syndrome (AS) group data and low-risk (LR) group data.
+Metrics were derived from laboratory annotations at the segment level (lab); from Zooniverse annotations collapsed at the segment level (Zoon.
+Seg); or directly from the chunk-level labels (Zoon.
+Chunks).
+}
+\centering
+\begin{tabular}[t]{l|l|l}
+\hline
+  & AS & LR\\
+\hline
+LP (lab) & .015 & .741\\
+\hline
+LP (Zoon. Seg.) & -.131 & .672\\
+\hline
+LP (Zoon. Chunks) & -.038 & .716\\
+\hline
+CP (lab) & -.557 & .607\\
+\hline
+CP (Zoon. Seg.) & -.404 & .604\\
+\hline
+CP (Zoon. Chunks) & -.38 & .573\\
+\hline
+\end{tabular}
+\end{table}
+
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
 
@@
 
-\leavevmode\hypertarget{ref-cychosz2019canonical}{}%
-Cychosz, M., Cristia, A., Bergelson, E., Casillas, M., Baudet, G., Warlaumont, A. S., \ldots{} Seidl, A. (2019). Canonical babble development in a large-scale crosslinguistic corpus.
+\leavevmode\hypertarget{ref-cychosz2021canonical}{}%
+Cychosz, M., Cristia, A., Bergelson, E., Casillas, M., Baudet, G., Warlaumont, A. S., \ldots{} Seidl, A. (2021). Canonical babble development in a large-scale crosslinguistic corpus.
 
@@
 
+\leavevmode\hypertarget{ref-team2013r}{}%
+R Core Team, \& others. (2013). R: A language and environment for statistical computing.
+
 \leavevmode\hypertarget{ref-roche2018early}{}%
@@
 (2018).
-Early vocal development in autism spectrum disorder, rett syndrome, and fragile x syndrome: Insights from studies using retrospective video analysis.
+Early vocal development in autism spectrum disorder, Rett syndrome, and Fragile X syndrome: Insights from studies using retrospective video analysis.
 \emph{Advances in Neurodevelopmental Disorders}, \emph{2}(1), 49--61.
@@
 
-\leavevmode\hypertarget{ref-team2013r}{}%
-Team, R. C., \& others. (2013). R: A language and environment for statistical computing.
-
 \leavevmode\hypertarget{ref-vandam2016homebank}{}%
@@
 (2016).
+HomeBank: An online repository of daylong child-centered audio recordings.
 \emph{Seminars in Speech and Language}, \emph{37}(2), 128--142.
@@
 Signal processing for young child speech language development.
-In \emph{First workshop on Child, Computer and Interaction}.
+In \emph{First Workshop on Child, Computer and Interaction}.
+
+\leavevmode\hypertarget{ref-yao2020classification}{}%
+Yao, X., Micheletti, M., Johnson, M., \& Barbaro, K. de. (2020). Classification of infant crying in real-world home environments using deep learning. \emph{arXiv Preprint arXiv:2005.07036}.
 
@@
 (2019).
-Using Attention Networks and Adversarial Augmentation for Styrian Dialect, Continuous Sleepiness and Baby Sound Recognition.
+Using attention networks and adversarial augmentation for Styrian dialect, continuous sleepiness and baby sound recognition.
 In \emph{INTERSPEECH} (pp.

### 2021-02-19 202010_jslhr/paper.Rmd
fixed figs in doc

@@
     corresponding : no
-  - name          : "Bridgette Kelleher"
+  - name          : "Bridgette L. Kelleher"
     affiliation   : "2"
@@
 author_note: >
-       AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017); and the J.
+       AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017), and the J.
 S.
-McDonnell Foundation (Understanding Human Cognition Scholar Award).
-LH acknowledges the National Institute of Deafness and Other Communication Disorders (F31DC018219).
-BK acknowledges the National Institute of Mental Health (K23MH111955) and the Kinley Trust.
+McDonnell Foundation (Understanding Human Cognition Scholar Award); LH  the National Institute of Deafness and Other Communication Disorders (F31DC018219); BK the National Institute of Mental Health (K23MH111955) and the Kinley Trust.
 This publication uses data generated via the Zooniverse.org platform, development of which is funded by generous support, including a Global Impact Award from Google, and by a grant from the Alfred P.

### 2021-02-19 202010_jslhr/paper.Rmd
added

--- 2021-02-18 202010_jslhr/paper.Rmd
@@
 author_note: >
-       AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017); and the J.
+       AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017) and the J.
 S.
-McDonnell Foundation (Understanding Human Cognition Scholar Award).
-LH acknowledges the National Institute of Deafness and Other Communication Disorders (F31DC018219).
-BK acknowledges the National Institute of Mental Health (K23MH111955) and the Kinley Trust.
+McDonnell Foundation (Understanding Human Cognition Scholar Award); LH the National Institute of Deafness and Other Communication Disorders (F31DC018219); BK  the National Institute of Mental Health (K23MH111955) and the Kinley Trust.
 This publication uses data generated via the Zooniverse.org platform, development of which is funded by generous support, including a Global Impact Award from Google, and by a grant from the Alfred P.

### 2021-02-19 202010_jslhr/paper.Rmd

--- 2021-02-19 202010_jslhr/paper.Rmd
@@
 author_note: >
-       AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017), and the J.
+       AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017) and the J.
 S.

### 2021-02-19 202010_jslhr/paper.tex

@@
 \shorttitle{Children's vocalizations and citizen science}
-\author{Chiara Semenzin\textsuperscript{1}, Lisa Hamrick\textsuperscript{2}, Amanda Seidl\textsuperscript{2}, Bridgette Kelleher\textsuperscript{2}, \& Alejandrina Cristia\textsuperscript{1}}
+\author{Chiara Semenzin\textsuperscript{1}, Lisa Hamrick\textsuperscript{2}, Amanda Seidl\textsuperscript{2}, Bridgette L. Kelleher\textsuperscript{2}, \& Alejandrina Cristia\textsuperscript{1}}
 \affiliation{
@@
 \textsuperscript{1} Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University\\\textsuperscript{2} Purdue University}
-\authornote{AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017); and the J.
+\authornote{AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017), and the J.
 S.
-McDonnell Foundation (Understanding Human Cognition Scholar Award).
-LH acknowledges the National Institute of Deafness and Other Communication Disorders (F31DC018219).
-BK acknowledges the National Institute of Mental Health (K23MH111955) and the Kinley Trust.
+McDonnell Foundation (Understanding Human Cognition Scholar Award); LH  the National Institute of Deafness and Other Communication Disorders (F31DC018219); BK the National Institute of Mental Health (K23MH111955) and the Kinley Trust.
 This publication uses data generated via the Zooniverse.org platform, development of which is funded by generous support, including a Global Impact Award from Google, and by a grant from the Alfred P.

### 2021-03-02 202010_jslhr/paper.Rmd
minor

@@
 ```{r setup, include = FALSE}
-library("papaja")
-library(tidyverse)
-library(dplyr)
-library(ggplot2)
-library(caret)
-library(scales)
-library(kableExtra)
-library(huxtable) #merging cells in tab1
-library(rel)
-library(ggpubr)
-library(lme4)
+library(groundhog)
+groundhog.day="2021-02-23"
+pkgs=c('papaja','tidyverse','dplyr','ggplot2','caret','scales','kableExtra','rel','ggpubr','lme4','rmarkdown')
+groundhog.library(pkgs, groundhog.day)
+
 #knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = FALSE, fig.pos = "T")
@@
 #document session info
-capture.output(sessionInfo(),file="lastknit_session_info.txt")
+capture.output(sessionInfo(),file="paper-lastknit_session_info.txt")
 

### 2021-03-02 202010_jslhr/sup_mat.Rmd
minor

@@
 ```{r setup, include=FALSE}
+library(groundhog)
+groundhog.day="2021-02-23"
+pkgs=c("knitr", 'tidyverse','dplyr','ggplot2','caret','scales','kableExtra','rel','rmarkdown')
+groundhog.library(pkgs, groundhog.day)
+
 knitr::opts_chunk$set(echo = TRUE)
 knitr::opts_chunk$set(fig.width=12, fig.height=8)
-library(tidyverse)
-library(dplyr)
-library(ggplot2)
-library(caret)
-library(scales)
-library(kableExtra)
-library(rel)
+
 dodiv=function(x) x/sum(x, na.rm=T)

### 2021-03-08 202010_jslhr/paper.Rmd
attempts at increasing reproducibility

@@
 ```{r setup, include = FALSE}
+
 library(groundhog)
 groundhog.day="2021-02-23"
-pkgs=c('papaja','tidyverse','dplyr','ggplot2','caret','scales','kableExtra','rel','ggpubr','lme4','rmarkdown')
-groundhog.library(pkgs, groundhog.day)
+#groundhog.library('papaja', groundhog.day) #NOTE! THIS IS A POINT OF POTENTIAL NON-REPRODUCIBILITY
+library(papaja) #NOTE!
+groundhog.library('tidyverse', groundhog.day)
+groundhog.library('dplyr', groundhog.day)
+groundhog.library('ggplot2', groundhog.day)
+groundhog.library('rmarkdown', groundhog.day)
+groundhog.library('caret', groundhog.day)
+groundhog.library('scales', groundhog.day)
+groundhog.library('kableExtra', groundhog.day)
+groundhog.library('rel', groundhog.day)
+groundhog.library('ggpubr', groundhog.day)
+groundhog.library('lme4', groundhog.day)
+groundhog.library('e1071', groundhog.day)
 
@@
 #document session info
-capture.output(sessionInfo(),file="paper-lastknit_session_info.txt")
+capture.output(sessionInfo(),file="lastknit_session_info.txt")
 
@@
 The first step we took before beginning the annotation process was the identification of children's vocalizations within day-long recordings.
-We did this automatically, using the LENA&reg; proprietary software [@xu2008signal].
-Specifically, the LENA&reg; software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child’s environment (e.g., Female-Adult, TV).
+We did this automatically, using the LENA&reg; proprietary software [@xu2008signal], with versions ranging from LENA Pro 3.5 to LENA HUB 3.3 to 4.0.
+In broad terms, the LENA&reg; software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child’s environment (e.g., Female-Adult, TV).
 Key child labels have a precision of about 60% and a recall of about 50% [@cristia2020thorough].
@@
 This category is broader than the "Junk" category used by Zooniverse annotators, which is explained in the next section.
+Training materials for laboratory coders are available from @hamrick2021si.
 
@@
 colnames(resampling)<-c("lp_r_all","lp_r_as","lp_r_lr","cp_r_all","cp_r_as","cp_r_lr")
+
 for(i in 1:50){
   sel_seg=NULL
-  for(thischi in levels(data_all$ChildID)) sel_seg<-c(sel_seg,as.character(data_all$segmentId_DB[sample(data_all$segmentId_DB[data_all$ChildID==thischi],100)]))
-
+
+  #the next line samples 100 voc IDs from each child
+  for(thischi in levels(factor(data_all$ChildID))) sel_seg<-c(sel_seg,as.character(sample(data_all$segmentId_DB[data_all$ChildID==thischi],100)))
+
+  #generate a subset using those voc IDs
   sampling=data_all[as.character(data_all$segmentId_DB) %in% as.character(sel_seg),]
@@
 }
+
 myr=colMeans(resampling)

### 2021-03-08 202010_jslhr/paper.tex

@@
             pdftitle={Describing vocalizations in young children: A big data approach through citizen science annotation},
+            pdfauthor={Chiara Semenzin1, Lisa Hamrick2, Amanda Seidl2, Bridgette L. Kelleher2, \& Alejandrina Cristia1},
             pdfborder={0 0 0},
@@
   {}{\typeout{Failed to patch abstract.}}
+\patchcmd{\HyOrg@maketitle}
+  {\section{\protect\normalfont{\@title}}}
+  {\section*{\protect\normalfont{\@title}}}
+  {}{\typeout{Failed to patch title.}}
 \makeatother
 \shorttitle{Children's vocalizations and citizen science}
-\author{Chiara Semenzin\textsuperscript{1}, Lisa Hamrick\textsuperscript{2}, Amanda Seidl\textsuperscript{2}, Bridgette L. Kelleher\textsuperscript{2}, \& Alejandrina Cristia\textsuperscript{1}}
-\affiliation{
-\vspace{0.5cm}
-\textsuperscript{1} Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University\\\textsuperscript{2} Purdue University}
-\authornote{AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017), and the J.
-S.
-McDonnell Foundation (Understanding Human Cognition Scholar Award); LH  the National Institute of Deafness and Other Communication Disorders (F31DC018219); BK the National Institute of Mental Health (K23MH111955) and the Kinley Trust.
-This publication uses data generated via the Zooniverse.org platform, development of which is funded by generous support, including a Global Impact Award from Google, and by a grant from the Alfred P.
-Sloan Foundation.
-The funders had no impact on this study.
-All authors approved the final manuscript as submitted, agree to be accountable for all aspects of the work and have no conflict of interests to disclose.
-
-Correspondence concerning this article should be addressed to Alejandrina Cristia, 29 rue d'Ulm, 75005, Paris, France. E-mail: alecristia@gmail.com}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 \title{Describing vocalizations in young children: A big data approach through citizen science annotation}
-
+\author{Chiara Semenzin\textsuperscript{1}, Lisa Hamrick\textsuperscript{2}, Amanda Seidl\textsuperscript{2}, Bridgette L. Kelleher\textsuperscript{2}, \& Alejandrina Cristia\textsuperscript{1}}
 \date{}
+
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University\\\textsuperscript{2} Purdue University}
 
@@
 The first step we took before beginning the annotation process was the identification of children's vocalizations within day-long recordings.
-We did this automatically, using the LENA® proprietary software (Xu et al., 2008).
-Specifically, the LENA® software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child's environment (e.g., Female-Adult, TV).
+We did this automatically, using the LENA® proprietary software (Xu et al., 2008), with versions ranging from LENA Pro 3.5 to LENA HUB 3.3 to 4.0.
+In broad terms, the LENA® software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child's environment (e.g., Female-Adult, TV).
 Key child labels have a precision of about 60\% and a recall of about 50\% (Cristia, Lavechin, et al., 2020).
@@
 This category is broader than the \enquote{Junk} category used by Zooniverse annotators, which is explained in the next section.
+Training materials for laboratory coders are available from Hamrick, Seidl, and Kelleher (2021).
 
@@
 We also observed quite a bit of variance across the 50 runs.
-For Linguistic Proportion, the range of correlations found for all infants was {[}.569,.878{]}; for the Angelman syndrome group {[}.493,.95{]}; for the low-risk group {[}.481,.911{]}.
-For Canonical Proportion, the range of correlations found for all infants was {[}.815,.964{]}; for the Angelman syndrome group {[}-.001,.94{]}; for the low-risk group {[}.859,.992{]}.
+For Linguistic Proportion, the range of correlations found for all infants was {[}.647,.888{]}; for the Angelman syndrome group {[}.517,.95{]}; for the low-risk group {[}.605,.903{]}.
+For Canonical Proportion, the range of correlations found for all infants was {[}.849,.962{]}; for the Angelman syndrome group {[}.222,.942{]}; for the low-risk group {[}.818,.977{]}.
 
@@
 \hline
-100 seg AS & .791 & .632\\
+100 seg AS & .806 & .721\\
 \hline
@@
 \hline
-100 seg LR & .787 & .943\\
+100 seg LR & .795 & .939\\
 \hline
@@
 \hline
-100 seg all & .771 & .905\\
+100 seg all & .783 & .909\\
 \hline
@@
 
+\leavevmode\hypertarget{ref-hamrick2021si}{}%
+Hamrick, L., Seidl, A., \& Kelleher, B. (2021). Annotation manual. Retrieved from \url{https://osf.io/x3nqe/?view_only=4d42fdb50108467bb37ec807c53672f3}
+
 \leavevmode\hypertarget{ref-hamrick2019measurement}{}%

### 2021-03-08 202010_jslhr/sup_mat.Rmd

@@
 groundhog.day="2021-02-23"
-pkgs=c("knitr", 'tidyverse','dplyr','ggplot2','caret','scales','kableExtra','rel','rmarkdown')
-groundhog.library(pkgs, groundhog.day)
+groundhog.library('tidyverse', groundhog.day)
+groundhog.library('dplyr', groundhog.day)
+groundhog.library('ggplot2', groundhog.day)
+groundhog.library('rmarkdown', groundhog.day)
+groundhog.library('caret', groundhog.day)
+groundhog.library('scales', groundhog.day)
+groundhog.library('kableExtra', groundhog.day)
+groundhog.library('rel', groundhog.day)
+groundhog.library('ggpubr', groundhog.day)
+groundhog.library('lme4', groundhog.day)
+groundhog.library('e1071', groundhog.day)
+groundhog.library('rjson', groundhog.day)
+
+# Seed for random number generation
+set.seed(42)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+
+#document session info
+capture.output(sessionInfo(),file="supmat-lastknit_session_info.txt")
 
@@
 - 2020-11-03 first version
-
-## Read data in
-
-```{r}
-# read datasets
-
-demo_data=read.csv("../Derived_Data/demo-data.tsv",sep="\t")
-data_ang <- read.csv("../Derived_Data/classifications_PU_zoon_final17.csv",header=T,sep=",")
-data_td <- read.csv("../Derived_Data/classifications_PU_zoon_final.csv")
-data_all<-rbind(data_ang, data_td)
-
-#add filenames to demo data
-demo_data_fn <- demo_data %>% 
-     left_join(select(data_all, filename, ChildID), by = c("ChildID"))
-demo_data_fn<-unique(demo_data_fn)
-
-#remove the word mixed that takes up space and is unnecessary
-data_all$Zoon_classif=factor(gsub("Mixed_","",as.character(data_all$Zoon_classif),fixed=T))
-#relevel the factor so that it's easier to read
-data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=c("Canonical","Non-Canonical",                   "Crying","Laughing","Junk",levels(data_all$Zoon_classif)[grep("_",levels(data_all$Zoon_classif))]))
-# create lab column with easier to read correspondance
-data_all$lab<-as.character(data_all$Major_Choice)
-data_all$lab[data_all$lab=="Non-canonical syllables"]<-"Non-Canonical"
-data_all$lab[data_all$lab=="Canonical syllables"]<-"Canonical"
-data_all$lab[data_all$lab %in% c("Don't mark","None")]<-"Junk"
-data_all$lab=factor(data_all$lab,levels=levels(data_all$Zoon_classif))
-#apply same factor levels as zooniverse so that we can do symmetrical confusion matrices
-```
+- 2021-03-08 increased reproducibility
 
@@
 
-```{r}
+```{r gen-seg-allinfo}
+## PHASE 3 -- Generate views on the data: majority judgment on the segment
+
+#dictionary that relates chunk to segment created by Chiara
+dict <- fromJSON(file="../data_analyses/files_from_elsewhere/dict_4.json") 
+dict.simple=unlist(dict)
+length(dict.simple) 
+dict=data.frame(cbind(dict.simple,gsub(".$","",as.character(names(dict.simple)))))
+colnames(dict)<-c("chunk","segmentId_DB")
+length(levels(factor(dict$chunk)))
+#33728 chunks
+length(levels(factor(dict$segmentId_DB)))
+#12170 segments
+#NOTE!!! THERE ARE TWO CHUNKS MISSING FROM THIS DICTIONARY!!!
+
+#laboratory annotations
+read.csv("../data_analyses/files_from_elsewhere/result_final_lisa.csv")->lab_jud
+dim(lab_jud)
+#11982 segments
+
+sum(lab_jud$segmentId_DB %in% levels(factor(dict$segmentId_DB))) #11980 in common
+sum(!(lab_jud$segmentId_DB %in% levels(factor(dict$segmentId_DB)))) #2 fund in lab but not dict
+sum(!(levels(factor(dict$segmentId_DB)) %in% lab_jud$segmentId_DB )) #190 found in dict but not in lab
+
+#assuming that we should follow lab data, I'll kick out any segment not in there
+
+#chunk info
+read.csv("../data_analyses/output/chunks_maj_judgments.csv",colClasses="factor")->maj_jud
+maj_jud$AudioData=gsub(".mp3","",maj_jud$filename)
+
+# next I need to match up the segments with the chunks
+sum(dict.simple %in% maj_jud$AudioData) # 33727 found
+sum(!(dict.simple %in% maj_jud$AudioData)) # 1 not found
+
+rownames(dict)<-dict$chunk
+maj_jud$segmentId_DB <- dict[maj_jud$AudioData,"segmentId_DB"]
+length(levels(factor(maj_jud$segmentId_DB))) #12170 segments
+
+#in postprocess, the following code is executed, but not here, because we want to keep info about chunk-majority-judgment within each segment
+  # #generate majority at the segment level using our rule:
+  # # canonical > non-canonical > crying > laughing > junk
+  # table(maj_jud$segmentId_DB,maj_jud$Answer)->mytab
+  # mytype<-ifelse(mytab[,"Canonical"]>0,"Canonical",
+  #                ifelse(mytab[,"Non-Canonical"]>0,"Non-Canonical",
+  #                       ifelse(mytab[,"Crying"]>0,"Crying",
+  #                              ifelse(mytab[,"Laughing"]>0,"Laughing",
+  #                                     ifelse(mytab[,"Junk"]>0,"Junk",""
+  #                                     )))))
+
+table(maj_jud$segmentId_DB,maj_jud$Answer)->mytab #we still create the table with the number of judgments
+
+#but the mytype vector will include n's of each type
+mytype <- rep("",n=length(levels(factor(maj_jud$segmentId_DB))))
+mytype<-ifelse(mytab[,"Canonical"]>0,paste(mytype,"Ca",mytab[,"Canonical"],sep=""),mytype)
+mytype<-ifelse(mytab[,"Non-Canonical"]>0,paste(mytype,"N",mytab[,"Non-Canonical"],sep=""),mytype)
+mytype<-ifelse(mytab[,"Crying"]>0,paste(mytype,"Cr",mytab[,"Crying"],sep=""),mytype)
+mytype<-ifelse(mytab[,"Laughing"]>0,paste(mytype,"L",mytab[,"Laughing"],sep=""),mytype)
+mytype<-ifelse(mytab[,"Junk"]>0,paste(mytype,"J",mytab[,"Junk"],sep=""),mytype)
+
+levels(factor(mytype))
+#there are over 350 combinations!! So we won't do that
+
+mytype <- rep("",n=length(levels(factor(maj_jud$segmentId_DB))))
+mytype<-ifelse(mytab[,"Canonical"]>0,paste(mytype,"Ca",sep=""),mytype)
+mytype<-ifelse(mytab[,"Non-Canonical"]>0,paste(mytype,"N",sep=""),mytype)
+mytype<-ifelse(mytab[,"Crying"]>0,paste(mytype,"Cr",sep=""),mytype)
+mytype<-ifelse(mytab[,"Laughing"]>0,paste(mytype,"L",sep=""),mytype)
+mytype<-ifelse(mytab[,"Junk"]>0,paste(mytype,"J",sep=""),mytype)
+levels(factor(mytype))
+# about 30 unique combinations, let's go for it
+
+zoo_jud=cbind(row.names(mytab),mytype)
+colnames(zoo_jud)<-c("segmentId_DB","Answer")
+
+merge(zoo_jud,lab_jud,by="segmentId_DB")->all_jud
+dim(all_jud)
+#11980 segments
+levels(factor(all_jud$ChildID)) 
+length(levels(factor(all_jud$ChildID)))#all 20 kids here
+
+```
+
+```{r cm}
+
 table(data_all$lab)
@@
 
-## Collapse across "mixed"
+```{r read-in}
 
-```{r}
-#given results above, we map the mixed
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical"]<-"Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Non-Canonical_Crying"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Crying"]<-"Crying"
-data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Crying"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing_Crying"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Canonical"]<-"Canonical"
-# +
-data_all$Zoon_classif[data_all$Zoon_classif=="Canonical_Crying"]<-"Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Canonical_Laughing"]<-"Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Laughing_Canonical_Crying"]<-"Non-Canonical"
+read.csv("../data_analyses/output/key_info.csv")->x
+rownames(x)<-x$X
 
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Laughing"]<-"Crying"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Canonical_Laughing"]<-"Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Laughing_Non-Canonical"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Non-Canonical"]<-"Non-Canonical"
-data_all$Zoon_classif[data_all$Zoon_classif=="Crying_Non-Canonical_Laughing"]<-"Non-Canonical" 
-data_all$Zoon_classif[data_all$Zoon_classif=="Non-Canonical_Laughing"]<-"Non-Canonical"
+read.csv("../data_analyses/output/chunks_maj_judgments.csv")->chunks
 
-```
+read.csv("../data_analyses/output/zoo_lab_maj_judgments.csv")->data_all
 
-```{r}
-#and reset the factors for cleanliness
-data_all$Zoon_classif=factor(data_all$Zoon_classif)
-data_all$lab=factor(data_all$lab)
-sample_data<-cbind(data_all$lab,data_all$Zoon_classif)
+label_options=c("Canonical" , "Non-Canonical"  ,     "Crying"   ,    "Laughing",        "Junk" )
+
+#use better names
+data_all$Zoon_classif=data_all$Answer
+
+# create lab column with easier to read correspondance
+data_all$lab<-as.character(data_all$Major_Choice)
+data_all$lab[data_all$lab=="Non-canonical syllables"]<-"Non-Canonical"
+data_all$lab[data_all$lab=="Canonical syllables"]<-"Canonical"
+data_all$lab[data_all$lab %in% c("Don't mark","None")]<-"Junk"
+data_all$lab=factor(data_all$lab,levels=label_options)
+#apply same factor levels as zooniverse so that we can do symmetrical confusion matrices
+
+#add binomials for Linguistic Proportion
+data_all$lab_ling=ifelse(data_all$lab %in% c("Canonical","Non-Canonical"),1,0)
+data_all$zoo_ling=ifelse(data_all$Zoon_classif %in% c("Canonical","Non-Canonical"),1,0)
+data_all$lab_ling[data_all$lab=="Junk"]<-NA
+data_all$zoo_ling[data_all$lab=="Junk"]<-NA
+
+#add binomials for Canonical Proportion
+data_all$lab_can=data_all$zoo_can=NA
+data_all$lab_can[data_all$lab=="Canonical"]<-1
+data_all$lab_can[data_all$lab=="Non-Canonical"]<-0
+data_all$zoo_can[data_all$Zoon_classif=="Canonical"]<-1
+data_all$zoo_can[data_all$Zoon_classif=="Non-Canonical"]<-0
+
+demo_data=read.csv("demo-data.tsv",sep ="\t")
+#add filenames to demo data, to be used later
+demo_data_fn <- demo_data %>% 
+     left_join(select(data_all, filename, ChildID), by = c("ChildID"))
+demo_data_fn<-unique(demo_data_fn)
 

### 2021-03-16 202010_jslhr/paper.Rmd
increased reproducibility

@@
 
-read.csv("../data_analyses/output/key_info.csv")->x
+read.csv("key_info.csv")->x
 rownames(x)<-x$X
 
-read.csv("../data_analyses/output/chunks_maj_judgments.csv")->chunks
-
-read.csv("../data_analyses/output/zoo_lab_maj_judgments.csv")->data_all
+read.csv(".chunks_maj_judgments.csv")->chunks
+
+read.csv("zoo_lab_maj_judgments.csv")->data_all
 

### 2021-03-16 202010_jslhr/sup_mat.Rmd
increased reproducibility

@@
 #dictionary that relates chunk to segment created by Chiara
-dict <- fromJSON(file="../data_analyses/files_from_elsewhere/dict_4.json") 
+dict <- fromJSON(file="dict_4.json") 
 dict.simple=unlist(dict)
@@
 #laboratory annotations
-read.csv("../data_analyses/files_from_elsewhere/result_final_lisa.csv")->lab_jud
+read.csv("lab_annotations.csv")->lab_jud
 dim(lab_jud)
@@
 #chunk info
-read.csv("../data_analyses/output/chunks_maj_judgments.csv",colClasses="factor")->maj_jud
+read.csv("chunks_maj_judgments.csv",colClasses="factor")->maj_jud
 maj_jud$AudioData=gsub(".mp3","",maj_jud$filename)
@@
 
-#but the mytype vector will include n's of each type
-mytype <- rep("",n=length(levels(factor(maj_jud$segmentId_DB))))
-mytype<-ifelse(mytab[,"Canonical"]>0,paste(mytype,"Ca",mytab[,"Canonical"],sep=""),mytype)
-mytype<-ifelse(mytab[,"Non-Canonical"]>0,paste(mytype,"N",mytab[,"Non-Canonical"],sep=""),mytype)
-mytype<-ifelse(mytab[,"Crying"]>0,paste(mytype,"Cr",mytab[,"Crying"],sep=""),mytype)
-mytype<-ifelse(mytab[,"Laughing"]>0,paste(mytype,"L",mytab[,"Laughing"],sep=""),mytype)
-mytype<-ifelse(mytab[,"Junk"]>0,paste(mytype,"J",mytab[,"Junk"],sep=""),mytype)
-
-levels(factor(mytype))
-#there are over 350 combinations!! So we won't do that
+# #but the mytype vector will include n's of each type
+# mytype <- rep("",n=length(levels(factor(maj_jud$segmentId_DB))))
+# mytype<-ifelse(mytab[,"Canonical"]>0,paste(mytype,"Ca",mytab[,"Canonical"],sep=""),mytype)
+# mytype<-ifelse(mytab[,"Non-Canonical"]>0,paste(mytype,"N",mytab[,"Non-Canonical"],sep=""),mytype)
+# mytype<-ifelse(mytab[,"Crying"]>0,paste(mytype,"Cr",mytab[,"Crying"],sep=""),mytype)
+# mytype<-ifelse(mytab[,"Laughing"]>0,paste(mytype,"L",mytab[,"Laughing"],sep=""),mytype)
+# mytype<-ifelse(mytab[,"Junk"]>0,paste(mytype,"J",mytab[,"Junk"],sep=""),mytype)
+# 
+# levels(factor(mytype))
+# #there are over 350 combinations!! So we won't do that
 
@@
 
+# create columns with names that match the following chunks
+all_jud$Zoon_classif = all_jud$Answer
+all_jud$lab = all_jud$Major_Choice
+
+# for the lab case, simplify
+all_jud$lab[all_jud$lab=="Canonical syllables"]<-"Ca"
+all_jud$lab[all_jud$lab=="Words"]<-"Ca"
+all_jud$lab[all_jud$lab=="Crying"]<-"Cr"
+all_jud$lab[all_jud$lab=="Laughing"]<-"L"
+all_jud$lab[all_jud$lab=="Don't mark"]<-"J"
+all_jud$lab[all_jud$lab=="None"]<-""
+all_jud$lab[all_jud$lab=="Non-canonical syllables"]<-"N"
+
+table(all_jud$lab)
+table(all_jud$Zoon_classif)
+
+#remove classes with fewer than 10 instances
+table(all_jud$Zoon_classif)[table(all_jud$Zoon_classif)<10]
+all_jud=all_jud[!(all_jud$Zoon_classif %in% names(table(all_jud$Zoon_classif)[table(all_jud$Zoon_classif)<10])),]
+
+#remove classes with no majority judgment
+all_jud=all_jud[all_jud$Zoon_classif !="",]
+all_jud=all_jud[all_jud$lab !="",]
+dim(all_jud)
+
+all_jud$Zoon_classif=factor(all_jud$Zoon_classif)
+all_jud$lab=factor(all_jud$lab, levels=levels(all_jud$Zoon_classif))
 ```
@@
 
-table(data_all$lab)
-table(data_all$Zoon_classif)
-mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
+mycf=confusionMatrix(all_jud$lab, all_jud$Zoon_classif, dnn = c("Lab","Zooniverse"))
 conf_tab=mycf$table
@@
 my_conf_tab=conf_tab
-for(i in 1:18) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
+for(i in 1:dim(my_conf_tab)[2]) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
 colSums(my_conf_tab)
@@
   geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=2) +
-  geom_text(aes(label = Freq), vjust = 1,size=1) +
+#  geom_text(aes(label = Freq), vjust = 1,size=1) +
   scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
@@
 my_conf_tab=conf_tab
-for(i in 1:18) my_conf_tab[,i]=my_conf_tab[,i]/rowsums[i]
+for(i in 1:dim(my_conf_tab)[1]) my_conf_tab[,i]=my_conf_tab[,i]/rowsums[i]
 rowSums(my_conf_tab)
@@
 
+## Read in & clean up final data
+
 ```{r read-in}
 
-read.csv("../data_analyses/output/key_info.csv")->x
+read.csv("key_info.csv")->x
 rownames(x)<-x$X
 
-read.csv("../data_analyses/output/chunks_maj_judgments.csv")->chunks
+read.csv(".chunks_maj_judgments.csv")->chunks
 
-read.csv("../data_analyses/output/zoo_lab_maj_judgments.csv")->data_all
+read.csv("zoo_lab_maj_judgments.csv")->data_all
 
@@
 
+```{r clean up, echo=F}
+#dim(data_all) #11980    16
+# remove non-majority labels from lab
+data_all[data_all$Num_Agreement>0,]->data_all
+#dim(data_all) # 11647    16
+
+# remove non-majority labels from zooniverse
+data_all[data_all$Zoon_classif %in% label_options,]->data_all
+#dim(data_all) # 11268    16
+
+#and reset the factors for cleanliness
+data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=label_options)
+data_all$lab=factor(data_all$lab, levels=label_options)
+sample_data<-cbind(data_all$lab,data_all$Zoon_classif)
+
+table_zoon=round(table(data_all$Zoon)/sum(table(data_all$Zoon))*100,2)
+table_lab=round(table(data_all$lab)/sum(table(data_all$lab))*100,2)
+```
+
+```{r prepare data by child, echo=F}
+
+#get the ns by child, then calculate the linguistic proportion & canonical proportion, separately for zooniverse & lab
+ztab=table(data_all$filename,data_all$Zoon_classif)
+z_lp =rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
+z_cp =ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])
+
+ltab=table(data_all$filename,data_all$lab)
+l_lp =rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
+l_cp =ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])
+
+#and also get Junk from zooniverse
+z_Junk=ztab[,c("Junk")]/rowSums(ztab)
+
+#put all the proportions together
+if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) proportions=cbind(rownames(ztab),z_lp,z_cp,l_lp,l_cp,z_Junk) else print("oops this code needs to be more complex because we don't have the same kids for the two proportions")
+colnames(proportions)[1]<-"filename"
+
+#ages=aggregate(data_all$Age,by=list(data_all$ChildID),mean) #this is a weird way of adding ages, since all of the ages for a given child should be the same if there is only one recording, and if there are multiple recordings, then we should not get the mean
+#improvement: now we merge with a demo data tab, but note this is merged with child id, so the problem of multiple recs per child is still there
+
+# Created demo_data with filenames. Use filenames instead of childIDs to merge proportions and demo data.
+
+merge(proportions,demo_data_fn,by="filename")->proportions
+colnames(proportions)[dim(proportions)[2]]<-"Age"
+
+#cbinding results in text, so we numerize the proportions
+for(thisvar in c("z_lp","z_cp","l_lp","l_cp","z_Junk")) proportions[,thisvar]=as.numeric(as.character(proportions[,thisvar]))
+
+#add median Junk
+proportions$median_Junk<-ifelse(proportions$z_Junk<median(proportions$z_Junk),"lower_Junk","higher_Junk")
+
+#summary(proportions)
+
+merge(data_all,proportions,all=T)->data_all
+```
+
 ## Separate confusion matrices for Angelman syndrome children
@@
 ```{r}
-data_as_td<-left_join(data_all,demo_data,on="ChildID")
 # CM with just AS kids
-data_AS<-subset(data_as_td, Diagnosis=="AngelmanSyndrome")
+data_AS<-subset(data_all, Diagnosis=="AngelmanSyndrome")
 mycf=confusionMatrix(data_AS$lab, data_AS$Zoon_classif, dnn = c("Lab","Zooniverse"))
@@
 # CM with just TD kids
-data_TD<-subset(data_as_td, Diagnosis=="Low-RiskControl")
+data_TD<-subset(data_all, Diagnosis=="Low-RiskControl")
 mycf=confusionMatrix(data_TD$lab, data_TD$Zoon_classif, dnn = c("Lab","Zooniverse"))

### 2021-03-23 202010_jslhr/paper.Rmd
switched to renv

@@
 ```{r setup, include = FALSE}
-
-library(groundhog)
-groundhog.day="2021-02-23"
-#groundhog.library('papaja', groundhog.day) #NOTE! THIS IS A POINT OF POTENTIAL NON-REPRODUCIBILITY
-library(papaja) #NOTE!
-groundhog.library('tidyverse', groundhog.day)
-groundhog.library('dplyr', groundhog.day)
-groundhog.library('ggplot2', groundhog.day)
-groundhog.library('rmarkdown', groundhog.day)
-groundhog.library('caret', groundhog.day)
-groundhog.library('scales', groundhog.day)
-groundhog.library('kableExtra', groundhog.day)
-groundhog.library('rel', groundhog.day)
-groundhog.library('ggpubr', groundhog.day)
-groundhog.library('lme4', groundhog.day)
-groundhog.library('e1071', groundhog.day)
+library(rel)
+library(dplyr)
+library(ggplot2)
+library(ggpubr)
+library(tidyverse)
+library(lme4)
+library(e1071)
+library(papaja)
+library(caret)
+library(scales)
+library(kableExtra)
 
@@
 
-read.csv(".chunks_maj_judgments.csv")->chunks
+read.csv("chunks_maj_judgments.csv")->chunks
 

### 2021-03-23 202010_jslhr/sup_mat.Rmd
switched to renv

@@
 ```{r setup, include=FALSE}
-library(groundhog)
-groundhog.day="2021-02-23"
-groundhog.library('tidyverse', groundhog.day)
-groundhog.library('dplyr', groundhog.day)
-groundhog.library('ggplot2', groundhog.day)
-groundhog.library('rmarkdown', groundhog.day)
-groundhog.library('caret', groundhog.day)
-groundhog.library('scales', groundhog.day)
-groundhog.library('kableExtra', groundhog.day)
-groundhog.library('rel', groundhog.day)
-groundhog.library('ggpubr', groundhog.day)
-groundhog.library('lme4', groundhog.day)
-groundhog.library('e1071', groundhog.day)
-groundhog.library('rjson', groundhog.day)
+library('tidyverse')
+library('dplyr')
+library('ggplot2')
+library('rmarkdown')
+library('caret')
+library('scales')
+library('kableExtra')
+library('rel')
+library('ggpubr')
+library('lme4')
+library('e1071')
+library('rjson')
 

### 2021-05-26 202010_jslhr/paper.Rmd
correcting typos

@@
 To compare laboratory and Zooniverse annotations at the level of segments, chunk-level labels are combined using the hierarchy Canonical > Non-canonical > Crying > Laughing > Junk."}
-knitr::include_graphics("fig_levels.key.pdf")
+knitr::include_graphics("fig_levels.pdf")
 
@@
 names(prettynames)<-c("z_lp","z_cp","l_lp","l_cp")
-mycols=c("black","red")
+mycols=c("red","black")
 names(mycols)<-c("Low-RiskControl","AngelmanSyndrome")
@@
 colnames(myr)<-c("AS","LR")
-mylb=myhb=myn=myp =myr #this is an ugly way of creating 5 matrices
+mylb=myhb=myn=myp =myr #this is an ugly way of creating 6 matrices
+
+#to improve legibility of figures, providing the coordinates to offset the text
+mytext<-rbind(
+  cbind(-.1,-.1,0,-.2), #"z_lp"
+  cbind(+.1,+.1,-.1,+.3), #"z_cp"
+  cbind(-.1,-.1,0,-.2), #"l_lp"
+  cbind(+.1,+.1,-.1,+.25) #"l_cp"
+)
+rownames(mytext)<-c("z_lp","z_cp","l_lp","l_cp")
+colnames(mytext)<-c("AS.x","AS.y","LR.x","LR.y")
 
@@
          ylim=myrange)
-  abline(lm(proportions[,thisvar]~proportions$Age,subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black")
+  abline(lm(proportions[,thisvar]~proportions$Age,subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black",lty=2)
 
@@
     #add r to figure
-  text(mean(proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"]),mean(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar]),paste0("r=",myr[thisvar,"AS"]),col="black")
+  text(
+    mean(proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"])+mytext[thisvar,"AS.x"],mean(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar])+mytext[thisvar,"AS.y"],
+    paste0("r=",myr[thisvar,"AS"]),col="black")
 
@@
 
-    text(mean(proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"]),mean(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar]),paste0("r=",myr[thisvar,"LR"]),col="red")
+    text(mean(proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"])+mytext[thisvar,"LR.x"],
+         mean(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar])+mytext[thisvar,"LR.y"],
+         paste0("r=",myr[thisvar,"LR"]),col="red")
 }
@@
 
-  abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black")
+  abline(lm(proportions[,"z_cp"]~proportions[,"l_cp"],subset=c(proportions$Diagnosis=="AngelmanSyndrome")),col="black",lty=2)
 
@@
 recalls=round(stall_r$rec[stall_r$Lab==stall_r$Zooniverse],1)
-names(recalls)<-c("Canonical","Crying","Junk","Laughing","Non-can.")
+names(recalls)<-stall_r$Lab[stall_r$Lab==stall_r$Zooniverse]
+names(recalls)[names(recalls)=="Non-Canonical"]<-"Non-can."
+  #c("Canonical","Crying","Junk","Laughing","Non-can.")
 recalls=recalls[as.character(tab1[2:6,"X"])]
 freqs=table(data_all$Zoon)
-names(freqs)<-c("Canonical","Crying","Junk","Laughing","Non-can.")
+names(freqs)[names(freqs)=="Non-Canonical"]<-"Non-can."
 freqs=freqs[as.character(tab1[2:6,"X"])]

### 2021-05-26 202010_jslhr/paper.tex
correcting typos

@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
 \PassOptionsToPackage{hyphens}{url}
 %
-\documentclass[english,,man]{apa6}
+\documentclass[
+  english,
+  ,man]{apa6}
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amssymb,amsmath}
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
   \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
+\IfFileExists{microtype.sty}{% use microtype if available
   \usepackage[]{microtype}
@@
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
   \IfFileExists{parskip.sty}{%
@@
     \setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\usepackage{hyperref}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \hypersetup{
@@
   pdfauthor={Chiara Semenzin1, Lisa Hamrick2, Amanda Seidl2, Bridgette L. Kelleher2, \& Alejandrina Cristia1},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{graphicx,grffile}
+  pdflang={en-EN},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage{graphicx}
 \makeatletter
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 % Manuscript styling
@@
 \captionsetup[figure]{font={stretch=1.5}}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[shorthands=off,main=english]{babel}
-\else
-  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
+\ifxetex
+  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
   \usepackage{polyglossia}
   \setmainlanguage[]{english}
+\else
+  \usepackage[main=english]{babel}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
 \fi
+\ifluatex
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 Some research suggests that spontaneous behavior, captured for instance via home videos, could provide important indices to development (Belardi et al., 2017; Overby, Belardi, \& Schreiber, 2020).
-Advancements in the field of wearable technologies, such as LENA® recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally (Oller et al., 2010; Rankine et al., 2017; VanDam \& Yoshinaga-Itano, 2019).
+Advancements in the field of wearable technologies, such as LENA® recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally (D.
+K.
+Oller et al., 2010; Rankine et al., 2017; VanDam \& Yoshinaga-Itano, 2019).
 Wearable recorders allow data collection to happen in the child's natural environment, and at a large scale, which may be particularly helpful for children whose speech is not easily elicited.
@@
 The most commonly used software was created by the LENA® Foundation, and it returns a diarization of the audio signal split into key talkers, including the child wearing the device, as well as a split of sections attributed to the child into three categories: crying, other fixed and vegetative signals, and speech vocalizations.
-These subcategories have not been widely validated (Cristia et al., 2020).
+These subcategories have not been widely validated (Cristia, Bulgarelli, \& Bergelson, 2020).
 Moreover, until recently LENA® lacked an important distinction between more and less advanced vocalizations.\footnote{The latest version of the LENA® software does tag canonical utterances separately from non-canonical ones.}
@@
 Table 1 provides an idea of the accuracy across different methods.
-Specifically, we calculated accuracy of the LENA®'s subtypes\footnote{If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as \enquote{Junk} or not categorized.
+Specifically, we calculated accuracy of the LENA®'s subtypes\footnote{If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as ``Junk'' or not categorized.
 When these data were analyzed, the LENA® software did not distinguish canonical and non-canonical.} based on the laboratory annotations used in the present study; and we also show the results of ComParE 2019 BabySounds sub-challenge, although note it is not based on the same data as the other two columns.
@@
 The ComParE 2019 BabySounds sub-challenge's baseline model achieved a more moderate recall on crying than LENA, but clearly outperforms the LENA® algorithm in the Laughing category, with the additional advantage of distinguishing canonical and non-canonical -- although note that performance in non-canonical is barely above a chance level of 20\% (based on 5 categories, assuming equal probabilities).
-Notice that the baseline model also included a category called \enquote{Junk} for sections that turned out not to be the child's vocalizations after all.
+Notice that the baseline model also included a category called ``Junk'' for sections that turned out not to be the child's vocalizations after all.
 LENA® has a low recall for the Junk category because only sections that LENA® classified as being child vocalizations are considered in the present analysis.
@@
 For example, most work using LENA® software reports on children's vocalization counts -- a derived metric because it is not simply a description of sections of the audio, but instead it integrates over the whole recording length.
-Although linguistic vocalization count has been criticized as being more about quantity than quality (McDaniel, Yoder, Estes, \& Rogers, 2020), it is a promising metric of individuals' vocalization development because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA® (Cristia et al., 2020).
+Although linguistic vocalization count has been criticized as being more about quantity than quality (McDaniel, Yoder, Estes, \& Rogers, 2020), it is a promising metric of individuals' vocalization development because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA® (Cristia, Bulgarelli, \& Bergelson, 2020).
 Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r \textasciitilde{} .3 with standardized language scores in a meta-analysis (Wang, Williams, Dilley, \& Houston, 2020).
@@
 That study built on a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this prior work focuses on babbling and thus typically on infants under one year of age.
-In these younger infants, a critical milestone involves the increasingly common production of canonical syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech (Oller, Eilers, Neal, \& Cobo-Lewis, 1998).
+In these younger infants, a critical milestone involves the increasingly common production of canonical syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech (D.
+K.
+Oller, Eilers, Neal, \& Cobo-Lewis, 1998).
 Given its adult-like consonant-vowel or vowel-consonant structure, canonical babble is considered to be a starting point on the path to recognizable speech.
 When compared to more primitive sounds such as squeals, or isolated vowels, canonical syllables show a higher complexity given the smooth articulatory transition between a consonant and a vowel (or vice versa).
-Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development (Cychosz et al., 2021; Oller, 2000).
-In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization counts in a sample of children diagnosed with an autism spectrum disorder (McDaniel et al., 2020).
+Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development (Cychosz et al., 2021; D.
+Oller, 2000).
+In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization counts in a sample of children diagnosed with an autism spectrum disorder (McDaniel, Yoder, Estes, \& Rogers, 2020).
 
@@
 
-\enquote{Crowdsourcing} refers to the process whereby a task is solved by a crowd, rather than an individual.
+``Crowdsourcing'' refers to the process whereby a task is solved by a crowd, rather than an individual.
 A number of fields, particularly in the data-driven sciences, have already engaged in the collection of data (including annotations) through crowdsourcing, thanks to its low cost and ecological value (Crump, McDonnell, \& Gureckis, 2013; Sescleifer, Francoisse, \& Lin, 2018).
-For example, a systematic review on crowdsourced ratings of speech found that \enquote{lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology}; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81 (Sescleifer et al., 2018).
+For example, a systematic review on crowdsourced ratings of speech found that ``lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology''; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81 (Sescleifer, Francoisse, \& Lin, 2018).
 On the other hand, the systematic review returned only 8 studies (of which only four were published in peer-reviewed journals), suggesting that there is considerable need for further research on this topic.
 
-Mechanical Turk (MTurk) MTurk is an online labor market created by Amazon to assist \enquote{requesters} in hiring and paying \enquote{workers} for the completion of computerized tasks.
-Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable (Berinsky, Huber, \& Lenz, 2012), question marks are raised as some \enquote{workers} turn out to be bots, or are poorly motivated (and potentially exploited) humans.
+Mechanical Turk (MTurk) MTurk is an online labor market created by Amazon to assist ``requesters'' in hiring and paying ``workers'' for the completion of computerized tasks.
+Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable ({``Evaluating online labor markets for experimental research,''} n.d.), question marks are raised as some ``workers'' turn out to be bots, or are poorly motivated (and potentially exploited) humans.
 
@@
 
-One of the most successful platforms hosting citizen science projects is Zooniverse (zooniverse.org; Borne \& Zooniverse Team, 2011).
+One of the most successful platforms hosting citizen science projects is Zooniverse {[}zooniverse.org; Borne and Zooniverse Team (2011){]}.
 The website hosts a multitude of interdisciplinary projects that have allowed the public to take part in cutting-edge scientific research, from marine biology to papyrology.
@@
 The authors also studied how their Canonical Proportion related to age within each corpus that had a sufficient number of children varying in age.
-They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the \enquote{Junk} category was very high.
+They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the ``Junk'' category was very high.
 Additionally, they found that the correlation between Canonical Proportion and age was much weaker in two corpora where vocalizations were segmented with LENA® than through manual segmentation.
@@
 In addition, we went beyond Cychosz et al.
-(2021)\enquote{s study in two important ways.
+(2021)`s study in two important ways.
 First, we relied on the largest and best established citizen science platform: Zooniverse hosts more than 1.6 million users from diverse walks of life, and it offers a completely automatized API system to more easily scale tasks in a transparent and cumulative science fashion.
 In contrast, Cychosz et al.
-(2021) actively recruited people to provide annotations, stating that \enquote{{[}a{]}nnotators included language and speech researchers, undergraduate students, research assistants, and other interested parties, totaling 136 unique annotators}.
-A process that draws from research assistants relies on expert resources, whose time is expensive.
+(2021) actively recruited people to provide annotations, stating that ``{[}a{]}nnotators included language and speech researchers, undergraduate students, research assistants, and other interested parties, totaling 136 unique annotators.'' A process that draws from research assistants relies on expert resources, whose time is expensive.
 Moreover, the people these researchers recruit tend to be more accustomed to child vocalization data, which may improve their annotation performance, and thus provide an overestimate of the quality of the annotations that can be realistically done within a truly citizen science framework.
-Second, a core goal of the present work was to determine how citizen scientists} annotations fare compared to the current gold standard, laboratory annotations.
+Second, a core goal of the present work was to determine how citizen scientists' annotations fare compared to the current gold standard, laboratory annotations.
 Cychosz et al.
@@
 It was important for generalizability purposes to examine data from children with variable biological age, mental age, and vocal maturity levels.
-We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA® (see meta-analyses in Wang et al., 2020; Cristia et al., 2020); as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range.
-Some previous work had shown that vocal development proceeded differently among children with an Angelman syndrome diagnosis when compared to a normative sample (Hamrick et al., 2019; Hamrick \& Tonnsen, 2019).
+We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA® (Cristia, Bulgarelli, \& Bergelson, 2020; see meta-analyses in Wang, Williams, Dilley, \& Houston, 2020); as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range.
+Some previous work had shown that vocal development proceeded differently among children with an Angelman syndrome diagnosis when compared to a normative sample (Hamrick, Seidl, \& Kelleher, 2019; Hamrick \& Tonnsen, 2019).
 Generally, studies on vocal development among children with genetic disorders reveal lower volubility and lower prevalence of canonical syllables (Belardi et al., 2017; Rankine et al., 2017; see also Roche et al., 2018 for a review of retrospective video analysis of children diagnosed with Autism Spectrum Disorder, Rett Syndrome, or Fragile X, documenting these and other vocal characteristics).
@@
 Returning to our overall predictions with these goals in mind, we were particularly interested in describing potential differences in accuracy of derived metrics as a function of our two participant groups.
-Previous work suggests that children with Angelman syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the Canonical Proportion was expected to increase with age (Hamrick et al., 2019; Hamrick \& Tonnsen, 2019).
+Previous work suggests that children with Angelman syndrome show decreases in Canonical Proportion with age, whereas in typically-developing infants, conversely, the Canonical Proportion was expected to increase with age (Hamrick, Seidl, \& Kelleher, 2019; Hamrick \& Tonnsen, 2019).
 Although we could not rely on previous work to make predictions regarding the Linguistic Proportion, we reasoned that we should observe an increase in the Linguistic Proportion for both populations.
@@
 In broad terms, the LENA® software attempts to assign segments of speech either to the key child (the one wearing the recorder), or to one of 15 other categories in the child's environment (e.g., Female-Adult, TV).
-Key child labels have a precision of about 60\% and a recall of about 50\% (Cristia, Lavechin, et al., 2020).
+Key child labels have a precision of about 60\% and a recall of about 50\% (Cristia et al., 2020).
 These stretches of audio assigned to the key child can contain cries, vegetative sounds, and linguistic vocalizations, as well as interstitial pauses.
@@
 Coders were instructed to classify each segment based on the highest level of vocal maturity it contained (e.g., a segment with both canonical and non-canonical syllables would be classified as canonical; a segment with both laughing and non-canonical syllables would be classified as non-canonical).
-For canonical, non-canonical, and word segments, coders indicated how many syllables of each type of vocal maturity they heard (e.g., \enquote{ah ma ba} would be 1 non-canonical and 2 canonical syllables).
-For the purposes of the present study, we considered a \enquote{word} judgment to be equivalent to a \enquote{canonical} judgment.
-This may be false sometimes (e.g., \enquote{oh} may be coded as a real word by laboratory experts, although it is a non-canonical syllable), but we could not have a \enquote{word} judgment in our Zooniverse annotations because that requires context (which compromises privacy -- see next section).
-
-Coders were also given a \enquote{Don't Mark} option, which they were instructed to use if the segment did not sound like a segment made by the target child, or if there is any overlapping speech or other noise in the background which could affect acoustic analyses.
-This category is broader than the \enquote{Junk} category used by Zooniverse annotators, which is explained in the next section.
+For canonical, non-canonical, and word segments, coders indicated how many syllables of each type of vocal maturity they heard (e.g., ``ah ma ba'' would be 1 non-canonical and 2 canonical syllables).
+For the purposes of the present study, we considered a ``word'' judgment to be equivalent to a ``canonical'' judgment.
+This may be false sometimes (e.g., ``oh'' may be coded as a real word by laboratory experts, although it is a non-canonical syllable), but we could not have a ``word'' judgment in our Zooniverse annotations because that requires context (which compromises privacy -- see next section).
+
+Coders were also given a ``Don't Mark'' option, which they were instructed to use if the segment did not sound like a segment made by the target child, or if there is any overlapping speech or other noise in the background which could affect acoustic analyses.
+This category is broader than the ``Junk'' category used by Zooniverse annotators, which is explained in the next section.
 Training materials for laboratory coders are available from Hamrick, Seidl, and Kelleher (2021).
@@
 A substantial concern that emerges when considering citizen science annotation for spontaneous speech data is the risk of a privacy breach: Even short clips can contain personal information that can expose the identity of the speaker, such as addresses or names, or contain sensitive and private information.
-Fortunately, Seidl et al.
-(2019) determined that when segments are divided into shorter clips (400-600 ms), human annotators with little training can code our categories of interest (Canonical, Laughing, Crying, etc.) with a classification quality comparable to the one carried out on full segments.
+Fortunately, Seidl, Warlaumont, and Cristia (2019) determined that when segments are divided into shorter clips (400-600 ms), human annotators with little training can code our categories of interest (Canonical, Laughing, Crying, etc.) with a classification quality comparable to the one carried out on full segments.
 We therefore used very short clips (500 ms) as these are unlikely to contain more than two syllables, and thus prevent the identification of any personal information.
@@
 Specifically, the segments extracted were automatically cut into clips of exactly 500 ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform.
-We call this level \emph{chunk}, to convey that they are part of a bigger unit (the LENA®-identified \enquote{segment}).
+We call this level \emph{chunk}, to convey that they are part of a bigger unit (the LENA®-identified ``segment'').
 Note that segments and chunks are collectively referred to as \emph{clips} because they are parts of the longer audio recording.
@@
 
-Citizen scientists could access our project, \enquote{Maturity of Baby Sounds}, by logging in, or as anonymous annotators.
+Citizen scientists could access our project, ``Maturity of Baby Sounds,'' by logging in, or as anonymous annotators.
 Before starting the annotation, citizen scientists were given a quick tutorial which walked them through the steps of the annotation workflow (i.e., how to play a sound, how to make a selection; see Figure 1).
@@
 Users were asked to assign each 500-ms chunk to one out of five possible categories: (1) Canonical, (2) Non-Canonical, (3) Crying, (4) Laughing, and (5) Junk (overlapping speech, non-infant speech, silence, external sounds).
-Notice that the latter is narrower than the \enquote{Don't mark} category of the lab annotation routine, which was used any time there was any noise or overlap that could affect acoustic analyses.
+Notice that the latter is narrower than the ``Don't mark'' category of the lab annotation routine, which was used any time there was any noise or overlap that could affect acoustic analyses.
 
@@
 
-The following description uses several key concepts: A \enquote{judgment} is a single classification provided by a Zooniverse user in response to a single chunk (or a single classification provided by a lab annotator in response to a single segment); a \enquote{chunk} is a 500-ms clip extracted from the daylong audio to represent part or all of a segment; a segment is a section of the daylong audio automatically classified as having been uttered by the key child according to the LENA® algorithm; and a \enquote{label} is the final category a chunk or segment was classified as.
+The following description uses several key concepts: A ``judgment'' is a single classification provided by a Zooniverse user in response to a single chunk (or a single classification provided by a lab annotator in response to a single segment); a ``chunk'' is a 500-ms clip extracted from the daylong audio to represent part or all of a segment; a segment is a section of the daylong audio automatically classified as having been uttered by the key child according to the LENA® algorithm; and a ``label'' is the final category a chunk or segment was classified as.
 Bearing this terminology in mind, we provide an overview of the data and how it was processed before the main analyses.
@@
 In this section, we provide descriptive analyses of our dataset.
-According to lab annotators, 15.82\% of segments were Canonical, 57.07\% Non-Canonical, 1.65\% Laughing, and 5.23\% Crying, with the remaining 20.23\% being categorized as \enquote{Don't mark}.
-Zooniverse data revealed a similar distribution: 14.71\% Canonical, 60.01\% Non-Canonical, 3.63\% Laughing, 8.87\% Crying, 12.79\% Junk.
+According to lab annotators, 15.82\% of segments were Canonical, 57.07\% Non-Canonical, 1.65\% Laughing, and 5.23\% Crying, with the remaining 20.23\% being categorized as ``Don't mark.'' Zooniverse data revealed a similar distribution: 14.71\% Canonical, 60.01\% Non-Canonical, 3.63\% Laughing, 8.87\% Crying, 12.79\% Junk.
 Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical + Non-Canonical) / All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical / (Canonical + Non-Canonical) (i.e., we remove Junk, Crying, and Laughing).
@@
 In this section, we report on exploratory analyses, aimed at establishing conditions under which Zooniverse labels aligned with laboratory labels more or less closely.
-In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as \enquote{Junk} (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora (Cychosz et al., 2021).
-A high proportion of \enquote{Junk} may indicate that automated segmentation was errorful for those children, and may be a sign that the rest of the data could be compromised as well.
-
-We investigated this hypothesis by calculating the proportion of data labeled as \enquote{Junk} for each individual child.
-There was no significant difference in the proportions of their data labeled as \enquote{Junk} for children with Angelman syndrome (M = 12.89\%, SD = 7.95\%) compared to the low-risk group (M = 14.25\%, SD = 3.97\%): Welch's t(13.23) = -0.48, \emph{p} = .636.
-We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as \enquote{Junk}.
-Results were similar across these two post-hoc subgroups (see Figure 6).
+In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as ``Junk'' (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora (Cychosz et al., 2021).
+A high proportion of ``Junk'' may indicate that automated segmentation was errorful for those children, and may be a sign that the rest of the data could be compromised as well.
+
+We investigated this hypothesis by calculating the proportion of data labeled as ``Junk'' for each individual child.
+There was no significant difference in the proportions of their data labeled as ``Junk'' for children with Angelman syndrome (M = 12.89\%, SD = 7.95\%) compared to the low-risk group (M = 14.25\%, SD = 3.97\%): Welch's t(13.23) = -0.48, \emph{p} = .636.
+We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as ``Junk.'' Results were similar across these two post-hoc subgroups (see Figure 6).
 For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = .885, CI {[}.576,.973{]}, \emph{p} = .001{]}; and for the higher Junk group it was r(8) = .864, CI {[}.513,.967{]}, \emph{p} .001.
 For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(8) = .958, CI {[}.828,.99{]}, \emph{p} \textless{} .001{]}; and for the higher Junk group it was r(8) = .93, CI {[}.726,.984{]}, \emph{p} \textless{} .001.
-Thus, it does not seem that a higher proportion of \enquote{Junk} labels is an index of low quality data.
+Thus, it does not seem that a higher proportion of ``Junk'' labels is an index of low quality data.
 
@@
 
-The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table 2, \enquote{100 seg} rows), particularly for the low-risk younger infants, and for Linguistic Proportion.
+The mean correlations across 50 runs were lower than those recovered using all segments from each child (see Table 2, ``100 seg'' rows), particularly for the low-risk younger infants, and for Linguistic Proportion.
 This may indicate that, especially in some groups of infants, 100 segments may not be sufficient to capture a stable estimate of the child's Linguistic and Canonical Proportions.
@@
 In this case, the lack of correlation with age in the Angelman syndrome group suggests that there is little systematic variance in this metric after 20 months of age.
-In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and relatively stable levels of Canonical Proportion among the children in the Angelman syndrome group (in line with previous work; Hamrick et al., 2019; Hamrick \& Tonnsen, 2019).
+In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and relatively stable levels of Canonical Proportion among the children in the Angelman syndrome group {[}in line with previous work; Hamrick, Seidl, and Kelleher (2019);Hamrick and Tonnsen (2019){]}.
 
@@
 In other cases, precision was fairly high but recall was lower.
-This was the case of the \enquote{Junk/Don't mark} category.
-The combination of high precision and low recall indicates that when Zooniverse annotations return a \enquote{Junk} judgment, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as \enquote{Don't mark} in the laboratory.
-Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the \enquote{Junk} category less) than laboratory annotators, who have access to the whole segment (and its context).
-This is consistent with the fact that our lab definition of \enquote{Junk} was more stringent than the Zooniverse definition.
-Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use \enquote{Don't mark} if there was \emph{any} noise or overlapping speech that could affect acoustic analyses.
+This was the case of the ``Junk/Don't mark'' category.
+The combination of high precision and low recall indicates that when Zooniverse annotations return a ``Junk'' judgment, this tends to be correct, but Zooniverse annotations may not detect all elements labeled as ``Don't mark'' in the laboratory.
+Inspection of confusion matrices and overall frequencies suggest that such a pattern of errors could be due to citizen scientists accepting more clips (i.e., using the ``Junk'' category less) than laboratory annotators, who have access to the whole segment (and its context).
+This is consistent with the fact that our lab definition of ``Junk'' was more stringent than the Zooniverse definition.
+Indeed, laboratory coders were instructed to be sensitive since the aim was also to study acoustic characteristics of the segments (which is unrelated to the current project), and therefore they would use ``Don't mark'' if there was \emph{any} noise or overlapping speech that could affect acoustic analyses.
 
@@
 
-As for Laughing, the recall is quite good but precision is low, with confusion involving \enquote{Junk} and Non-Canonical to similar extents.
+As for Laughing, the recall is quite good but precision is low, with confusion involving ``Junk'' and Non-Canonical to similar extents.
 We are tentative in our interpretation of this result because the category was very sparsely represented, with barely 186 segments classified as Laughing in the whole data set.
@@
 We used LENA® algorithms to detect vocalizations by the key child, but there are newer, better-performing, open-sourced alternatives that could be used instead (Lavechin, Bousbib, Bredin, Dupoux, \& Cristia, 2020).
-Still, performance is most often established on children whose development is thought to be normative, being raised in small households located in urban environments, whereas it would be important to check for accuracy in more diverse samples, where the children have a diagnosis (e.g., Dykstra et al., 2013), multiple siblings (e.g., Elo, 2016), and/or are growing up in a rural environment (e.g., Cristia, Lavechin, et al., 2020).
+Still, performance is most often established on children whose development is thought to be normative, being raised in small households located in urban environments, whereas it would be important to check for accuracy in more diverse samples, where the children have a diagnosis (e.g., Dykstra et al., 2013), multiple siblings (e.g., Elo, 2016), and/or are growing up in a rural environment (e.g., Cristia et al., 2020).
 Speech technology developments would also be welcome to complement citizen scientists' classifications, particularly for categories that may be hard to classify based on short samples.
@@
 Is the age-related decline we observed among children in the Angelman syndrome group truly a developmental decline that happens by necessity in this population, or might it reflect partially other aspects of children's experience and behavior?
-Additionally, longitudinal work within the same infant may be needed to track the \enquote{natural history} of vocalizations to better contextualize how these metrics should be interpreted.
+Additionally, longitudinal work within the same infant may be needed to track the ``natural history'' of vocalizations to better contextualize how these metrics should be interpreted.
 It is also an open question how large effects must be, at the individual level, in order to be detected with this method.
@@
 \centering
-\includegraphics{fig_levels.key.pdf}
+\includegraphics{fig_levels.pdf}
 \caption{\label{fig:fig-process}Correspondence between LENA-identified segments (which can contain one or more vocalizations) and chunks.
@@
 \hline
-Crying & 70.6 & 78.0 & 50.9 & 669 & 6762\\
-\hline
-Laughing & 41.5 & 56.0 & 62.9 & 149 & 409\\
-\hline
-Non-can. & 24.1 & 77.0 & 78.9 & 4485 & 1441\\
-\hline
-Canonical & 66.4 & 59.0 & 58.3 & 1426 & 1657\\
-\hline
-Junk & 67.3 & 1.0 & 43.5 & 4575 & 999\\
+Crying & 70.6 & 78 & 50.9 & 669 & 999\\
+\hline
+Laughing & 41.5 & 12 & 62.9 & 149 & 409\\
+\hline
+Non-can. & 24.1 & 73 & 78.9 & 4485 & 6762\\
+\hline
+Canonical & 66.4 &  & 58.3 & 1426 & 1657\\
+\hline
+Junk & 67.3 & 1.0 & 43.5 & 4575 & 1441\\
 \hline
@@
 \hline
-WAR & 49.9 & 57.0 & 55.3 &  & \\
+WAR & 49.9 & 57.0 & 68.3 &  & \\
 \hline
@@
 \hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
 \leavevmode\hypertarget{ref-aust2017papaja}{}%
-Aust, F., \& Barth, M. (2017). Papaja: Create APA manuscripts with R Markdown.
+Aust, F., \& Barth, M. (2017). Papaja: Create {APA manuscripts with R Markdown}.
 
@@
 (2017).
-A retrospective video analysis of canonical babbling and volubility in infants with Fragile X Syndrome at 9--12 months of age.
+A retrospective video analysis of canonical babbling and volubility in infants with {Fragile X Syndrome} at 9--12 months of age.
 \emph{Journal of Autism and Developmental Disorders}, \emph{47}(4), 1193--1206.
-
-\leavevmode\hypertarget{ref-berinsky2012evaluating}{}%
-Berinsky, A. J., Huber, G. A., \& Lenz, G. S. (2012). Evaluating online labor markets for experimental research: Amazon.com's Mechanical Turk. \emph{Political Analysis}, \emph{20}(3), 351--368.
 
@@
 Panoptes, a project building tool for citizen science.
-In \emph{Proceedings of the AAAI Conference on Human Computation and Crowdsourcing (HCOMP'15)}.
+In \emph{{Proceedings of the AAAI Conference on Human Computation and Crowdsourcing (HCOMP'15)}}.
 San Diego, CA.
@@
 \leavevmode\hypertarget{ref-casillas2019step}{}%
-Casillas, M., \& Cristia, A. (2019). A step-by-step guide to collecting and analyzing long-format speech environment (LFSE) recordings. \emph{Collabra}, \emph{5}(1).
+Casillas, M., \& Cristia, A. (2019). A step-by-step guide to collecting and analyzing long-format speech environment ({LFSE}) recordings. \emph{Collabra}, \emph{5}(1).
 
@@
 (2020).
-Accuracy of the Language Environment Analysis system segmentation and metrics: A systematic review.
+Accuracy of the {Language Environment Analysis} system segmentation and metrics: A systematic review.
 \emph{Journal of Speech, Language, and Hearing Research}, \emph{63}(4), 1093--1105.
@@
 (2020).
-A thorough evaluation of the Language Environment Analysis (LENA) system.
+A thorough evaluation of the {Language Environment Analysis (LENA)} system.
 \emph{Behavior Research Methods}, 1--20.
@@
 \leavevmode\hypertarget{ref-crump2013evaluating}{}%
-Crump, M. J., McDonnell, J. V., \& Gureckis, T. M. (2013). Evaluating Amazon's Mechanical Turk as a tool for experimental behavioral research. \emph{PloS One}, \emph{8}(3), e57410.
+Crump, M. J., McDonnell, J. V., \& Gureckis, T. M. (2013). Evaluating{ Amazon's Mechanical Turk} as a tool for experimental behavioral research. \emph{PloS One}, \emph{8}(3), e57410.
 
@@
 (2013).
-Using the Language Environment Analysis (LENA) system in preschool classrooms with children with autism spectrum disorders.
+{Using the Language Environment Analysis (LENA) system in preschool classrooms with children with autism spectrum disorders}.
 \emph{Autism}, \emph{17}(5), 582--594.
-\url{https://doi.org/https://doi.org/10.1177/1362361312446206}
+https://doi.org/\url{https://doi.org/10.1177/1362361312446206}
 
 \leavevmode\hypertarget{ref-Elo}{}%
-Elo, H. (2016). \emph{Acquiring language as a twin}. Tampere, Finland: Tampere University Press.
+Elo, H. (2016). \emph{{Acquiring language as a twin}}. Tampere, Finland: Tampere University Press.
+
+\leavevmode\hypertarget{ref-berinsky2012evaluating}{}%
+Evaluating online labor markets for experimental research: Amazon.com's mechanical turk. (n.d.).
 
@@
 (2018).
-Language ENvironment analysis (LENA) system investigation of day long recordings in children: A literature review.
+{Language ENvironment analysis (LENA)} system investigation of day long recordings in children: A literature review.
 \emph{Journal of Communication Disorders}, \emph{72}, 77--85.
@@
 \leavevmode\hypertarget{ref-hamrick2019capturing}{}%
-Hamrick, L., Seidl, A., \& Kelleher, B.
+Hamrick, L.
+R., Seidl, A., \& Kelleher, B.
+L.
 (2019).
-Capturing variability in early language skills among children with Angelman syndrome: A novel semi-automated approach.
+Capturing variability in early language skills among children with {A}ngelman syndrome: A novel semi-automated approach.
 \emph{Presented at the Angelman Biomarkers and Outcome Measures Alliance Scientific Meeting}.
@@
 \leavevmode\hypertarget{ref-hamrick2021si}{}%
-Hamrick, L., Seidl, A., \& Kelleher, B. (2021). Annotation manual. Retrieved from \url{https://osf.io/x3nqe/?view_only=4d42fdb50108467bb37ec807c53672f3}
+Hamrick, L. R., Seidl, A., \& Kelleher, B. L. (2021). Annotation manual. Retrieved from \url{https://osf.io/x3nqe/?view_only=4d42fdb50108467bb37ec807c53672f3}
 
 \leavevmode\hypertarget{ref-hamrick2019measurement}{}%
-Hamrick, L., \& Tonnsen, B.
+Hamrick, L.
+R., \& Tonnsen, B.
+L.
 (2019).
-Measurement and characterization of early social communication in children with Angelman syndrome.
+{Measurement and characterization of early social communication in children with Angelman syndrome}.
 In A.
 Sadhwani \& K.
-Okoniewski (Eds.), \emph{Advances in characterizing developmental trajectories in Angelman syndrome across the lifespan, Gatlinburg conference}.
+C.
+Okoniewski (Eds.), \emph{Advances in characterizing developmental trajectories in {A}ngelman syndrome across the lifespan, {G}atlinburg conference}.
 San Antonio, TX.
@@
 \emph{Proceedings of Interspeech}.
-\url{https://doi.org/https://arxiv.org/abs/2005.12656v2}
+https://doi.org/\url{https://arxiv.org/abs/2005.12656v2}
 
@@
 
+\leavevmode\hypertarget{ref-team2013r}{}%
+R Core Team, \& others. (2013). R: A language and environment for statistical computing.
+
 \leavevmode\hypertarget{ref-rankine2017language}{}%
@@
 (2017).
-Language ENvironment Analysis (LENA) in Phelan-McDermid syndrome: Validity and suggestions for use in minimally verbal children with Autism Spectrum Disorder.
+{Language ENvironment Analysis (LENA) in Phelan-McDermid syndrome: Validity and suggestions for use in minimally verbal children with Autism Spectrum Disorder}.
 \emph{Journal of Autism and Developmental Disorders}, \emph{47}(6), 1605--1617.
-
-\leavevmode\hypertarget{ref-team2013r}{}%
-R Core Team, \& others. (2013). R: A language and environment for statistical computing.
 
@@
 (2018).
-Early vocal development in autism spectrum disorder, Rett syndrome, and Fragile X syndrome: Insights from studies using retrospective video analysis.
+{Early vocal development in autism spectrum disorder, Rett syndrome, and Fragile X syndrome: Insights from studies using retrospective video analysis}.
 \emph{Advances in Neurodevelopmental Disorders}, \emph{2}(1), 49--61.
@@
 (2019).
-The INTERSPEECH 2019 Computational Paralinguistics Challenge: Styrian Dialects, Continuous Sleepiness, Baby Sounds \& Orca Activity.
+{The INTERSPEECH 2019 Computational Paralinguistics Challenge: Styrian Dialects, Continuous Sleepiness, Baby Sounds \& Orca Activity}.
 In \emph{Interspeech} (pp.
@@
 \leavevmode\hypertarget{ref-vandam2019use}{}%
-VanDam, M., \& Yoshinaga-Itano, C. (2019). Use of the LENA Autism Screen with children who are deaf or hard of hearing. \emph{Medicina}, \emph{55}(8), 495.
+VanDam, M., \& Yoshinaga-Itano, C. (2019). Use of the {LENA Autism Screen} with children who are deaf or hard of hearing. \emph{Medicina}, \emph{55}(8), 495.
 
@@
 (2020).
-A meta-analysis of the predictability of LENA™ automated measures for child language development.
+A meta-analysis of the predictability of {LENA}™ automated measures for child language development.
 \emph{Developmental Review}, \emph{57}, 100921.
@@
 Signal processing for young child speech language development.
-In \emph{First Workshop on Child, Computer and Interaction}.
+In \emph{First {Workshop on Child, Computer and Interaction}}.
 
@@
 (2019).
-Using attention networks and adversarial augmentation for Styrian dialect, continuous sleepiness and baby sound recognition.
+Using attention networks and adversarial augmentation for {S}tyrian dialect, continuous sleepiness and baby sound recognition.
 In \emph{INTERSPEECH} (pp.
@@
 
+\end{CSLReferences}
+
 \clearpage
